### 2020-11-12 draft.tex
first commit

\documentclass{article}
\usepackage[utf8]{inputenc}

\usepackage[dutch]{babel}     	
\usepackage{graphicx}                   % om afbeeldingen in te kunnen laden   
\usepackage{tikz,tikz-3dplot}                       % om afbeeldingen te kunnen maken
\usetikzlibrary{calc}                   % meer opties in tikz (oa coordinaten bepalen)

\usepackage{wrapfig}              % meer controle over figuren
\usepackage{floatrow}
\usepackage[format=hang]{caption}
\usepackage{listings}                   % voor mooie code
\usepackage{amsmath, amssymb, amsfonts} % om wiskunde te typen
\usepackage{scalerel}                   % om symbolen te schalen
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{bm}                         % Wiskundesymbolen vet met commando \bm{}
\usepackage{libertine}                  % lettertype
\usepackage{inconsolata}                % lettertype code
\usepackage[T1]{fontenc}                % betere lettertype encodering voor talen met accenten
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{units}
\usepackage{multicol}
\usepackage{appendix}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{physics}
\usepackage[noend]{algpseudocode}
\usepackage{tikz-feynman}
\usepackage{subcaption}
\usepackage{xfrac}
\tikzfeynmanset{compat=1.0.0}
\floatname{algorithm}{Algoritme}

\title{Thesis: draft}
\author{Nina Botte }
\date{November 2020}

\begin{document}

\maketitle

\section{Introduction}
The past decade social media has become ever more important in our daily lives. Nowadays most of us have, beside a real world life, a virtual life on social media. However this increased engagement on social media may have a huge impact, not only on the individual scale, but also on a broader, societal, scale. It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings. The broad range of interactions people undergo on social media with people from all over the world can thus not be underestimated in the process of opinion formation. It is also known that some people change their opinions more easily than others, that is once in a while you encounter somebody that is really stubborn and persistent towards its own opinion. The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance. The aim of this thesis is thus to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
These opinion dynamics models have two important layers. On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamic models to reproduce the opinion dynamics and opinion formation in real life.

\subsection{Complex networks}

Complex networks have become more and more popular for analyzing complex, dynamical systems. They are used in fields such as physics, economics, social sciences, biology, etc. These different fields are very diverge, but they have at least one common ground: they often deal with a large number of variables/components that interact with each other. Or in other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the individual components. %(maybe check cursus compl en crit?)
One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks. A network, also often called a graph, is a structure composed of nodes or vertices and a set of links (edges) that indicate the interactions between the nodes. Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems. One can find the language to describe networks in mathematical graph theory. However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable. The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way. However we do want our models to display properties similar to the ones seen in real systems. Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks. Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic. Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc. These concepts will be explained in depth in section \ref{sec2}, but let us already anticipate a bit on the case of social networks. It is found that many real life social systems have a power law degree distribution. This heterogeneous or scale free degree distribution represents one of the three general properties of social networks. The other two are short distances, also referred to as small-world phenomenon, and high clustering. %https://www.nature.com/articles/srep01783
Ideally, our theoretical network should exhibit these three properties. Some theoretical networks, that are thoroughly studied, may not have all of these three properties. However, they might still be used, because of their simplicity and ability to reproduce analytic results. When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems.
% to do: read links 19, 21, 22, 23, 24 of https://arxiv.org/pdf/0711.3199.pdf

% maybe also talk about degree distributions (poisson, power law), power law behavior of many real systems, clustering + maybe talk about networks that will be used in this thesis?

\subsection{Opinion dynamics}

% biblio: https://arxiv.org/pdf/0711.3199.pdf
% biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9

\section{'section2'}
\label{sec2}
\end{document}

### 2020-11-18 draft.tex
updated text + made bibtex file for references

@@
 \section{Introduction}
+
 The past decade social media has become ever more important in our daily lives.
 Nowadays most of us have, beside a real world life, a virtual life on social media.
-However this increased engagement on social media may have a huge impact, not only on the individual scale, but also on a broader, societal, scale.
+However, this increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal, scale.
 It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings.
 The broad range of interactions people undergo on social media with people from all over the world can thus not be underestimated in the process of opinion formation.
+The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
  It is also known that some people change their opinions more easily than others, that is once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
-The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.
-The aim of this thesis is thus to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
-These opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamic models to reproduce the opinion dynamics and opinion formation in real life.
+The aim of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+Opinion dynamics models have two important layers.
+On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation in real life.
 
@@
 These different fields are very diverge, but they have at least one common ground: they often deal with a large number of variables/components that interact with each other.
-Or in other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the individual components.
-%(maybe check cursus compl en crit?)
+Or in other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of individual components alone.
+Complex systems often display phenomena such as non-linearity, emergence, spontanuous order,...%(maybe check cursus compl en crit?)
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
@@
 Ideally, our theoretical network should exhibit these three properties.
-Some theoretical networks, that are thoroughly studied, may not have all of these three properties.
+Some theoretical networks, that are thoroughly studied, do not possess all of these three properties.
 However, they might still be used, because of their simplicity and ability to reproduce analytic results.
 When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems.
+\\
+One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system.
+Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, a clustered random network, the Watts-Strogatz model, etc.
+These will be explained thoroughly in section \ref{sec2}.
 % to do: read links 19, 21, 22, 23, 24 of https://arxiv.org/pdf/0711.3199.pdf
@@
 
+As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg.
+stubbornness), the information someone is exposed to, etc.
+Opinion dynamics models should try to capture these complex processes in a simplified way.
+Several models have been developed ranging from simple binary models to more complex, continuous approaches.
+The basic idea of all opinion dynamic models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
+These models are obviously a simplification of real world opinion dynamics.
+It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) en disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc.
+\\
+The opinion of the agents in the model can be either discrete or continuous.
+This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions.
+This might come across as a huge simplification of the real world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
+republicans or democrats, being left-minded or right-minded, cat or dog, renting or buying a house,...).
+%The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,...
+% not finished, describe different models
+%The structure of social networks plays an important role in aspects as peer pressure, whereas the algorithmic personalization used on social media platforms may interfere with the information someone is exposed to.
+Algorithmic personalization is the individualized way in which social media platforms show posts/information to each of its users.
+
+% talk about updating rules used in thesis + concepts as resistance etc. 
+%https://arxiv.org/pdf/1605.06326.pdf opinie dynamica
+
 % biblio: https://arxiv.org/pdf/0711.3199.pdf
@@
 
+\subsection{Complexity and social physics}
+
+Another important tool in the study of opinion dynamics/formations in large groups of people is the use of statistical physics.
+This branch of physics offers a framework that relates microscopic properties of atoms, molecules,...
+to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
+The observation that large number of people display collective, 'macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics.
+The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics.
+The 'microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, 'macroscopic' groups such as human societies.
+These 'macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc.
+The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among individuals.
+%https://arxiv.org/pdf/1605.06326.pdf complexity
+% talk about history
+
+% talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...)
 \section{'section2'}
 \label{sec2}
+
+% bibliography
+% \cite{Sirbu2016}
+\bibliographystyle{plain} % can change this style
+\bibliography{references}
+% to run this go to correct location in terminal --> type: pdflatex name, (file name.tex) --> type: bibtex name --> type: pdflatex name --> type: pdflatex name 
 \end{document}

### 2020-11-20 draft.tex
worked on the references + changed documentclass to book

@@
-\documentclass{article}
+\documentclass[11 pt , letterpaper , twoside , openright]{book}
 \usepackage[utf8]{inputenc}
 
-\usepackage[dutch]{babel}     	
+\usepackage[english]{babel}     	
 \usepackage{graphicx}                   % om afbeeldingen in te kunnen laden   
@@
 
-\title{Thesis: draft}
-\author{Nina Botte }
-\date{November 2020}
-
 \begin{document}
 
-\maketitle
-
-\section{Introduction}
+\chapter{Introduction}
 
@@
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation in real life.
+On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.
 
-\subsection{Complex networks}
+\section{Complex networks}
 
 Complex networks have become more and more popular for analyzing complex, dynamical systems.
-They are used in fields such as physics, economics, social sciences, biology, etc.
+They are used in fields such as physics, economics, social sciences, biology, etc \cite{Costa2008}.
 These different fields are very diverge, but they have at least one common ground: they often deal with a large number of variables/components that interact with each other.
-Or in other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of individual components alone.
-Complex systems often display phenomena such as non-linearity, emergence, spontanuous order,...%(maybe check cursus compl en crit?)
+In other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of the individual components alone \cite{Mata2020}.
+Complex systems often display phenomena such as non-linearity, emergence, spontanuous order,...
+%(maybe check cursus compl en crit?)
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links (edges) that indicate the interactions between the nodes.
+A network, also often called a graph, is a structure composed of nodes or vertices and a set of links (edges) that indicate the interactions between the nodes \cite{Mata2020}.
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
 One can find the language to describe networks in mathematical graph theory.
-However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable.
+However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.\\
 The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way.
-However we do want our models to display properties similar to the ones seen in real systems.
+However we do want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
 Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
-Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic.
+Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.
 Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc.
@@
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
-The other two are short distances, also referred to as small-world phenomenon, and high clustering.
-%https://www.nature.com/articles/srep01783
+The other two are short distances, also referred to as small-world phenomenon, and high clustering \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 These will be explained thoroughly in section \ref{sec2}.
+%references to chapter 2 need to be adapted
 % to do: read links 19, 21, 22, 23, 24 of https://arxiv.org/pdf/0711.3199.pdf
@@
 
-\subsection{Opinion dynamics}
+\section{Opinion dynamics}
 
@@
 Opinion dynamics models should try to capture these complex processes in a simplified way.
-Several models have been developed ranging from simple binary models to more complex, continuous approaches.
+Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.
 The basic idea of all opinion dynamic models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
 These models are obviously a simplification of real world opinion dynamics.
-It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) en disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc.
+It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) en disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
 \\
@@
 republicans or democrats, being left-minded or right-minded, cat or dog, renting or buying a house,...).
+\\
+
 %The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,...
-% not finished, describe different models
-%The structure of social networks plays an important role in aspects as peer pressure, whereas the algorithmic personalization used on social media platforms may interfere with the information someone is exposed to.
+The structure of social networks plays an important role in aspects as peer pressure, whereas the algorithmic personalization used on social media platforms may interfere with the information someone is exposed to.
 Algorithmic personalization is the individualized way in which social media platforms show posts/information to each of its users.
 
+% not finished, describe different models
 % talk about updating rules used in thesis + concepts as resistance etc. 
-%https://arxiv.org/pdf/1605.06326.pdf opinie dynamica
-
-% biblio: https://arxiv.org/pdf/0711.3199.pdf
 % biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9
 
-\subsection{Complexity and social physics}
+\section{Complexity and social physics}
 
@@
 to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
-The observation that large number of people display collective, 'macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics.
+The observation that large number of people display collective, 'macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
 The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics.
@@
 These 'macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc.
-The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among individuals.
-%https://arxiv.org/pdf/1605.06326.pdf complexity
+The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.
 % talk about history
@@
 % talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...)
-\section{'section2'}
+\chapter{'section2'}
 \label{sec2}
 
+\newpage
 % bibliography
-% \cite{Sirbu2016}
-\bibliographystyle{plain} % can change this style
+\bibliographystyle{abbrv} % can change this style
 \bibliography{references}

### 2020-11-20 draft.tex
changed references.bib + did some more writing

@@
 
+% talk about difference between simple networks and complex network??
+
 \section{Opinion dynamics}
@@
 \\
-
+The rules that determine how an agent updates his or her opinion can include many different aspect.
+They can, for example, be as simple as a majority rule (i.e.
+the majority model) or can include a more probabilistic way of updating.
+One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of a node on others.
 %The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,...
@@
 % talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...)
-\chapter{'section2'}
+\chapter{Networks}
 \label{sec2}
+
+%https://www.researchgate.net/publication/324312765_What_is_a_Complex_Network_CDT-2
+
+%https://www.nature.com/articles/s42005-020-0359-6
+
+\section{Definition and measurements}
+
+Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$ where the element of $E$ determine connections between elements of $V$
+\begin{equation}
+	E \subseteq \{\{x, y\}| x, y \in V \}.
+\end{equation}
+If self-loops are not allowed, we need the following condition on the elements of $E$
+\begin{equation}
+	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \}.
+\end{equation}
+
+% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections)
+
+\section{Network models}
+% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
+
+\chapter{Opinion dynamics}
+% talk about opinion dynamics, different models, resistance, etc
 

### 2020-11-25 draft.tex
changed the '' so that it looks better

@@
 In other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of the individual components alone \cite{Mata2020}.
-Complex systems often display phenomena such as non-linearity, emergence, spontanuous order,...
+Complex systems often display phenomena such as non-linearity, emergence, spontaneous order,...
 %(maybe check cursus compl en crit?)
@@
 \\
-The rules that determine how an agent updates his or her opinion can include many different aspect.
+The rules that determine how an agent updates his or her opinion can include many different aspects.
 They can, for example, be as simple as a majority rule (i.e.
@@
 to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
-The observation that large number of people display collective, 'macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
+The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
 The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics.
-The 'microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, 'macroscopic' groups such as human societies.
-These 'macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc.
+The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies.
+These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc.
 The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.

### 2020-11-27 draft.tex
continued writing (chapter 2)

@@
 \usepackage{xfrac}
-\tikzfeynmanset{compat=1.0.0}
+\tikzfeynmanset{compat=1.1.0}
 \floatname{algorithm}{Algoritme}
@@
 
-\section{Definition and measurements}
+\section{Definition and network measurements}
 
-Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$ where the element of $E$ determine connections between elements of $V$
+Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$ where the element of $E$ determine connections between elements of $V$ \cite{Costa2018}
 \begin{equation}
@@
 \end{equation}
+A graph with no self-loops is a simple graph.
+The nodes $x$ and $y$ of an edge $\{x, y\}$ are called the endpoints of the edge.
+It is possible that nodes are not joined by any edge, such nodes are called isolated.
+If two or more edges have the same pair of endpoints, the graph is called a multi-graph.
+This thesis is not concerned with multi-graphs, nor are self-loops allowed.
+Graphs in which the edges have an orientation are called directed graphs.
+If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
+A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.\\
+
+\subsection{Adjacency matrix}
+
+The adjacency matrix \textbf{A} is a $N \times N$ matrix (for a graph with $N$ nodes) where the elements indicate whether two nodes are connected by an edge or not \cite{Mata2020}
+\begin{equation}
+	A_{ij} = 
+	\begin{cases}
+		1 & \text{if nodes $i$ and $j$ are connected}\\
+		0 & \text{otherwise}
+	\end{cases}
+\end{equation}
+For a simple, undirected graph, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$).
+Directed graphs have asymmetric adjacency matrices.
+If we are dealing with weighted graphs, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
+\begin{equation}
+	A_{ij} = 
+	\begin{cases}
+		w_{ij} & \text{if nodes $i$ and $j$ are connected}\\
+		0 & \text{otherwise}
+	\end{cases}
+\end{equation}
+with, generally, $0 \leqslant w_{ij} \leqslant 1$ \cite{Mata2020}. \\
+\subsection{Degree and degree distribution}
+
+The adjacency matrix contains a lot of information such as the degree of a node.
+The degree $k_i$ of a node $i$ is the number of edges attached to the node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of the node $i$.
+The degree can be obtained from the adjacency matrix in the following way \cite{Mata2020}
+\begin{equation}
+	k_i = \sum_{j=1}^N A_{ij}.
+\end{equation}
+In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\text{in}}$ and the number of outgoing edges $k_i^{\text{out}}$ of the node $i$.
+These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
+\begin{equation}
+	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}.
+\end{equation}
+The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$.
+For weighted graphs the degree is easily generalized to the weighted degree $s_i$, also often called strength \cite{Ioannis2007}
+\begin{equation}
+	s_i = \sum_{j=1}^N w_{ij}.
+\end{equation}
+%start talking about degree distribution (global, degree = local)
 

### 2020-11-28 draft.tex
changed some really small things in the text

@@
 Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
+%maybe give definition of temporal network
 Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.
@@
 These concepts will be explained in depth in section \ref{sec2}, but let us already anticipate a bit on the case of social networks.
-It is found that many real life social systems have a power law degree distribution.
+It is found that many real life social systems have a power law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.

### 2020-12-11 draft.tex
continued writing in chapter 2

@@
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of a node on others.
+% add reference to chapter 2
 %The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,...
@@
 The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$.
-For weighted graphs the degree is easily generalized to the weighted degree $s_i$, also often called strength \cite{Ioannis2007}
+For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
 \begin{equation}
@@
 \end{equation}
-%start talking about degree distribution (global, degree = local)
+From the degree of the nodes in the network one can construct the degree distribution.
+The degree distribution $P(k)$ represent the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$.
+The average degree can be obtained from the degree distribution \cite{Mata2020}
+\begin{equation}
+	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k)
+\end{equation}
+where $N$ is the total number of nodes in the network.
+The degree distribution also allows us to classify networks.
+The two most important classes are homogeneous and heterogeneous networks.
+Homogeneous networks have a bell curved degree distribution, eg.
+a Poisson distribution.
+In this case, most of the nodes have a degree closed to the average degree $k$.
+Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
+% explain differences  more, consequences of power-law + need to include figure?
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network.
+% correct statement?
+%maybe explain why we care about degree, centrality measure etc
 
-% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections)
+\subsection{Clustering coefficient and transitivity}
+
+The clustering coefficient is another important measure of network topology.
+It is a measure for the number of triangles in a network and determines the connectivity in the neighborhood of a node $i$: if a node $i$ has a high clustering coefficient, its neighbors are likely to be directly connected to each other \cite{Li2017}.
+Here a triangle is defined as a loop of length three, this is a sequence of nodes $x, y, z, x$ such that $\{x, y\}, \{y, z\}$ and $\{z, x\}$ are edges of the network.
+The clustering coefficient is thus a way to measure the degree to which nodes in a network tend to cluster \cite{Li2017}.
+The local clustering coefficient of a node $i$ is defined as
+\begin{equation}
+	C_i = \frac{2n_i}{k_i(k_i-1)}
+\end{equation}
+where $n_i$ is the number of edges that actually exist between the nodes in the neighborhood of $i$ and $k_i(k_i-1)/2$ is the maximum number of edges that could exist between them (note that this expression is only valid for an undirected graph, since for a directed graph $e_{ij} \neq e_{ji}$ and we have $k_i(k_i-1)$ possible edges between the neighbors of node i).
+% should give more mathimatical expression for n_i?
+The average clustering coefficient is then given as the average of the local clustering coefficients
+\begin{equation}
+	\bar{C} = \frac{1}{N}\sum_{i = 1}^N C_i
+\end{equation}
+where $N$ denotes, once again, the number of nodes in the network.
+The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Li2017} \cite{Iyer2013}.
+\\
+A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as \cite{F.Costa2007}
+\begin{equation}
+	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}.
+\end{equation}
+A connected triple is defined as a set of three nodes with at least two or three edges between them.
+The factor three arises from the fact that each triangle contributes to three different connected triples in the network: one centered at each node in the triangle \cite{F.Costa2007}.
+% talk about difference transitivity and clustering + read https://arxiv.org/pdf/cond-mat/0505185.pdf
+
+\cite{Albert2002} % not used yet, explains statistical mechanics in complex networks
+
+% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
+% explain small-world (milgram), high clustering, power-law
 

### 2020-12-12 draft.tex
continued writing chapter 2

@@
 
-Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$ where the element of $E$ determine connections between elements of $V$ \cite{Costa2018}
+Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
 \begin{equation}
@@
 % explain differences  more, consequences of power-law + need to include figure?
-They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network.
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
 % correct statement?
@@
 The local clustering coefficient of a node $i$ is defined as
-\begin{equation}
+\begin{equation}\label{clus}
 	C_i = \frac{2n_i}{k_i(k_i-1)}
@@
 The average clustering coefficient is then given as the average of the local clustering coefficients
-\begin{equation}
+\begin{equation}\label{avClus}
 	\bar{C} = \frac{1}{N}\sum_{i = 1}^N C_i
@@
 where $N$ denotes, once again, the number of nodes in the network.
-The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Li2017} \cite{Iyer2013}.
+The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Iyer2013} \cite{Li2017}.
 \\
-A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as \cite{F.Costa2007}
-\begin{equation}
+A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as (valid for undirected, unweighted networks) \cite{F.Costa2007}
+\begin{equation}\label{globalTrans}
 	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}.
 \end{equation}
-A connected triple is defined as a set of three nodes with at least two or three edges between them.
+A connected triple is defined as a set of three nodes with at least two edges between them, so that each node can be reached from the other two (either directly or indirectly).
 The factor three arises from the fact that each triangle contributes to three different connected triples in the network: one centered at each node in the triangle \cite{F.Costa2007}.
-% talk about difference transitivity and clustering + read https://arxiv.org/pdf/cond-mat/0505185.pdf
-
-\cite{Albert2002} % not used yet, explains statistical mechanics in complex networks
+Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$.
+These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
+\begin{align}
+	N_{\Delta} &= \sum_{k > j > i} a_{ij}a_{ik}a_{jk} \\
+	N_3 &= \sum_{k > j > i} (a_{ij}a_{ik} + a_{ji}a_{jk} + a_{ki}a_{kj})
+\end{align}
+where $a_{ij}$ are the elements of the adjacency matrix. It is possible to define the transitivity of one node as \cite{F.Costa2007}
+\begin{equation}\label{trans}
+	T_i = \frac{N_{\Delta}(i)}{N_3(i)}
+\end{equation}
+where $N_{\Delta}(i)$ represents the number of triangles that involve node $i$ and $N_3(i)$ is the number of connected triples with $i$ as the central node \cite{F.Costa2007}
+\begin{align}
+	N_{\Delta}(i) &= \sum_{k > j} a_{ij}a_{ik}a_{jk} \\
+	N_3(i) &= \sum_{k > j} a_{ij}a_{ik}
+\end{align}
+It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$ and thus represents the number of neighbors of $i$.
+It is thus obvious that the local clustering coefficient $C_i$ (Eq.
+(\ref{clus})) of a node and the transitivity $T_i$ (Eq.
+(\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
+In the case of the average clustering coefficient $\bar{C}$ and the global transitivity $T$ this is however not the case.
+The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node.
+This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.
 
@@
 \section{Network models}
+
+In order to study the topological structures of real-world networks, several theoretical network models have been developed.
+These theoretical models usually have a simpler representation and well known properties that can be derived mathematically.
+They are widely studied and used extensively.
+The study of complex networks relies heavily on the knowledge and understanding of these models.
+In this section we will review some of the most important ones.
+
+\subsection{The Erd\H{o}s-R\'{e}nyi network}
+
+In 1959 the two mathematicians Paul Erd\H{o}s and Alfr\'{e}d R\'{e}nyi proposed a model to generate random networks \cite{F.Costa2007}.
+Idependently of them, Solomonoff and Rapoport already proposed the model in 1951.
+The model constructs random networks with $N$ nodes in the following way \cite{Albert2002}
+\begin{enumerate}
+	\item Start with $N$ disconnected nodes,
+	\item For each pair of nodes, connect them with a predefined probability $p$.
+\end{enumerate} 
+If $p=1$, we obtain a network with the maximum number of edges $N(N-1)/2$. This is called the fully connected network.\\
+The probability that a node is connected to $k$ other nodes and not to $N-k$ others is given by
+\begin{equation}
+	p^k (1-p)^{N-1-k}.
+\end{equation}
+The number of ways to choose the $k$ nodes among the $N-1$ possible candidates is given by the binomial coefficient
+\begin{equation}
+	\binom{N-1}{k} = \frac{(N-1)!}{k!(N-1-k)!}
+\end{equation}
+and thus the probability of being connected to $k$ other nodes becomes
+\begin{equation}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}.
+\end{equation}
+% continue expression for large N
+
 % different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...

### 2020-12-14 draft.tex
continued writing on ER network

@@
 The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
- It is also known that some people change their opinions more easily than others, that is once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
+ It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
 The aim of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
@@
 One can find the language to describe networks in mathematical graph theory.
-However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.\\
+However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
+One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\
 The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way.
@@
 Some theoretical networks, that are thoroughly studied, do not possess all of these three properties.
-However, they might still be used, because of their simplicity and ability to reproduce analytic results.
+However, they might still be used, because of their simplicity and ability to produce analytic results.
 When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems.
@@
 In 1959 the two mathematicians Paul Erd\H{o}s and Alfr\'{e}d R\'{e}nyi proposed a model to generate random networks \cite{F.Costa2007}.
-Idependently of them, Solomonoff and Rapoport already proposed the model in 1951.
+Independently of them, Solomonoff and Rapoport already proposed the model in 1951.
 The model constructs random networks with $N$ nodes in the following way \cite{Albert2002}
@@
 \end{equation}
-and thus the probability of being connected to $k$ other nodes becomes
+and the probability of being connected to exactly $k$ other nodes becomes
 \begin{equation}
@@
 \end{equation}
+We thus derived the degree distribution of a random network. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
+\begin{equation}\label{avDeg}
+\begin{split}
+	\left<k\right> &= \sum_{k=0}^{N-1} k\  p_k  \\
+&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}.
+\end{split}
+\end{equation}
+In order to simplify this equation we can make use of the following formula \cite{Hopcroft2006}
+\begin{equation} 
+	(p+q)^n = \sum_{k=0}^n \binom{n}{k} p^k q^{n-k}.
+\end{equation}
+Differentiating both sides of this equation gives us \cite{Hopcroft2006}
+\begin{equation}
+\begin{split}
+	n(p+q)^{n-1} &= \sum_{k=0}^n \binom{n}{k} k p^{k-1} q^{n-k} \\
+	&= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} q^{n-k}.
+\end{split}
+\end{equation}
+Finally, by substituting $q = 1-p$, we get \cite{Hopcroft2006}
+\begin{equation}
+\begin{split}
+	n (p+(1-p))^{n-1} &= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} (1-p)^{n-k} \\
+	np &= \sum_{k=0}^n \binom{n}{k} k p^{k} (1-p)^{n-k}
+\end{split}
+\end{equation}
+where the right hand side is equal to Equation (\ref{avDeg}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
+% derivation for large k https://barabasi.com/f/624.pdf
+\cite{Albert2014} large N
+
 % continue expression for large N

### 2020-12-17 draft.tex
added comment in introduction
continued writing the introduction + section about network communities

@@
 
+% add research question and hypotheses!!
+% also talk about algorithmic personalisation!!
+
 The past decade social media has become ever more important in our daily lives.
 Nowadays most of us have, beside a real world life, a virtual life on social media.
-However, this increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal, scale.
+This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale.
 It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings.
@@
 The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
+One important feature of opinion dynamics on social media is the use of algorithmic personalization.
+Social media allow for simultaneous or asynchronous interactions between people without geographical constraints and thus allow for the spread of information at a faster pace and at an unprecedented scale \cite{Perra2019}.
+People are, however, still bound by temporal and cognitive constraints.
+In order to assure a more pleasant/convenient experience on their platforms, social media use algorithms to order/filter the posts on an individuals time line according to what might be relevant to that individual \cite{Perra2019}.
+It is thus important to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
 It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
@@
 Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc.
-These concepts will be explained in depth in section \ref{sec2}, but let us already anticipate a bit on the case of social networks.
+These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks.
 It is found that many real life social systems have a power law degree distribution \cite{Muchnik2013}.
@@
 One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system.
-Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, a clustered random network, the Watts-Strogatz model, etc.
-These will be explained thoroughly in section \ref{sec2}.
+Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model, the Watts-Strogatz model, etc.
+These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
 %references to chapter 2 need to be adapted
@@
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of a node on others.
+\\
+Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms.
+The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
+Here popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
+In Chapter \ref{chap3} %also give section
+a detailed description of the filtering algorithms used in  this thesis will be given.
 % add reference to chapter 2
@@
 
-\section{Complexity and social physics}
+\section{Statistical and social physics}
 
@@
 % talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...)
+
+\section{This thesis: hypotheses and goals}
+
+As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
+So on one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.\\
+In one of their previous papers (\cite{Perra2019}) prof.Nicola Perra and prof.
+Luis E.C.
+Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
+They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).
+Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers.
+In the case of nudging, they found that the population opinion moved to the nudged opinion relatively fast, even in the case of small nudging.
+They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
+This thesis will build on their model and incorporate these new features.
+We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient.
+We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can observe convergence to one opinion inside the communities.
+On the other hand we want to see the influence of stubborn agents on the formation and evolution of opinions in the network.
+
+% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that
+% give reference to sections were community structure etc are explained!
+
 \chapter{Networks}
-\label{sec2}
+\label{chap2}
 
@@
 \section{Definition and network measurements}
+\label{netDef}
 
@@
 
+\subsection{Network communities}
+
+% continue working on this based on http://networksciencebook.com/chapter/9#basics and https://arxiv.org/pdf/1503.06277.pdf
+
+Network communities is another feature that (complex) networks may posses.
+A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}.
+More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}.
+In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}.
+The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}.
+The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it.
+Several community definitions are consistent with the density hypothesis.
+Let us consider three possible definitions: maximum cliques, strong community and weak community \cite{Albert2016}.\\
+\newline
+\textbf{Maximum cliques}\\
+\newline
+A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}.
+A community based on this definition would then be the largest clique in the network.
+This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
+\newline
+\textbf{Strong communities}\\
+\newline
+A strong community is defined such that each node inside the community has more links to other nodes inside the same community than to nodes outside the community \cite{Albert2016}.
+Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
+If we define the internal degree $k_i^{int}(C)$ as the number of links of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of links of node $i$ with nodes outside $C$, then we have the following condition in order that $C$ is a strong community \cite{Albert2016}
+\begin{equation}
+	k_i^{int}(C) > k_i^{ext}(C).
+\end{equation}
+\\
+\newline
+\textbf{Weak communities}\\
+\newline
+A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}.
+We thus have the following condition for a weak community $C$ \cite{Albert2016}
+\begin{equation}
+	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C).
+\end{equation}
 % talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
 % explain small-world (milgram), high clustering, power-law
+% explain community structure + difference with clusetering!!
+% explain connectivity before network communities
 
 \section{Network models}
+\label{netModel}
 
@@
 \chapter{Opinion dynamics}
+\label{chap3}
 % talk about opinion dynamics, different models, resistance, etc

### 2020-12-18 draft.tex
corrected some mistakes + added reference

@@
 It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
-The aim of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
 Opinion dynamics models have two important layers.
@@
 Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
-%maybe give definition of temporal network
+Temporal networks are networks where the edges are not always active, but instead become active for some periods of time \cite{Holme2012}.
 Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.
@@
 These models are obviously a simplification of real world opinion dynamics.
-It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) en disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
+It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
 \\
@@
 They can, for example, be as simple as a majority rule (i.e.
-the majority model) or can include a more probabilistic way of updating.
+the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating.
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of a node on others.
@@
 As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
-So on one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.\\
-In one of their previous papers (\cite{Perra2019}) prof.Nicola Perra and prof.
+So on one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.
+Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\
+In one of their previous papers (\cite{Perra2019}) prof.
+Nicola Perra and prof.
 Luis E.C.
@@
 This thesis will build on their model and incorporate these new features.
-We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient.
+We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
 We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can observe convergence to one opinion inside the communities.
-On the other hand we want to see the influence of stubborn agents on the formation and evolution of opinions in the network.
+On the other hand we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network.
 
@@
 From the degree of the nodes in the network one can construct the degree distribution.
-The degree distribution $P(k)$ represent the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$.
+The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$.
 The average degree can be obtained from the degree distribution \cite{Mata2020}
@@
 % explain differences  more, consequences of power-law + need to include figure?
-They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
+They are also referred to as scale free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
 % correct statement?
 %maybe explain why we care about degree, centrality measure etc
+
+\subsection{Connectivity}
+% explain connectivity before network communities and also before clustering!
 
@@
 \end{align}
-It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$ and thus represents the number of neighbors of $i$.
+It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
 It is thus obvious that the local clustering coefficient $C_i$ (Eq.
@@
 Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
-If we define the internal degree $k_i^{int}(C)$ as the number of links of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of links of node $i$ with nodes outside $C$, then we have the following condition in order that $C$ is a strong community \cite{Albert2016}
+If we define the internal degree $k_i^{int}(C)$ as the number of links of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of links of node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
 \begin{equation}
-	k_i^{int}(C) > k_i^{ext}(C).
+	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C.
 \end{equation}
@@
 \end{equation}
+
 % talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
 % explain small-world (milgram), high clustering, power-law
-% explain community structure + difference with clusetering!!
-% explain connectivity before network communities
+% explain community structure + difference with clustering!!
 

### 2020-12-18 draft.tex
test

@@
 % also talk about algorithmic personalisation!!
+% test ssh key github
 
@@
 Temporal networks are networks where the edges are not always active, but instead become active for some periods of time \cite{Holme2012}.
+In temporal networks time becomes an explicit element in the network representation \cite{Holme2012}.
 Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.

### 2021-01-18 draft.tex
test2
corrected two typos

@@
 % also talk about algorithmic personalisation!!
-% test ssh key github
 
@@
 People are, however, still bound by temporal and cognitive constraints.
-In order to assure a more pleasant/convenient experience on their platforms, social media use algorithms to order/filter the posts on an individuals time line according to what might be relevant to that individual \cite{Perra2019}.
+In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individuals time line according to what might be relevant to that individual \cite{Perra2019}.
 It is thus important to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
@@
 \end{align}
-It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
+It is not to hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
 It is thus obvious that the local clustering coefficient $C_i$ (Eq.
@@
 A community based on this definition would then be the largest clique in the network.
-This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
+This definition of community might, however, be to restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
 \newline

### 2021-02-01 draft.tex
continued writing on ER network + connectivity

--- 2020-12-18 draft.tex
@@
 This might come across as a huge simplification of the real world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
-republicans or democrats, being left-minded or right-minded, cat or dog, renting or buying a house,...).
+republicans or democrats, cat or dog, renting or buying a house,...).
 \\
@@
 the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating.
-One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of a node on others.
+One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others.
 \\
@@
 a Poisson distribution.
-In this case, most of the nodes have a degree closed to the average degree $k$.
+In this case, most of the nodes have a degree closed to the average degree $\left<k\right>$.
 Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
@@
 \subsection{Connectivity}
-% explain connectivity before network communities and also before clustering!
+In physics literature, the degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
+This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.
+Two nodes $u$ and $v$ of a graph are connected if there exist a path between them.
+If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
+The average degree $\left<k\right>$ is a global measurement of the connectivity of the network \cite{Costa2008}.
+If the average connectivity $\left<k\right>$ is too small (this is, if there are too few edges), there will be many isolated edges and only a few clusters with a small number of nodes.
+If more and more edges are added to the network, the small clusters will grow and will be connected to each other to form larger clusters.
+At some critical value of the connectivity, most nodes will be connected into a giant cluster, which characterizes the percolation of the network \cite{F.Costa2007}.
+% NOT finished; add this as a subsection in the degree section??
 
@@
 % explain small-world (milgram), high clustering, power-law
-% explain community structure + difference with clustering!!
+% explain modularity!! + diameter, shortest path, etc...!
 
@@
 and the probability of being connected to exactly $k$ other nodes becomes
-\begin{equation}
+\begin{equation}\label{degDistRan}
 	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}.
@@
 We thus derived the degree distribution of a random network. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
+% maybe put calculations in appendix?
 \begin{equation}\label{avDeg}
@@
 \end{equation}
-where the right hand side is equal to Equation (\ref{avDeg}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
-% derivation for large k https://barabasi.com/f/624.pdf
-\cite{Albert2014} large N
-
-% continue expression for large N
+where the right hand side is equal to Equation (\ref{avDeg}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.\\
+Most real networks are however sparse, which means that $\left<k\right> \ll N$. In this case the degree distribution is well approximated by a Poisson distribution \cite{Albert2014} 
+\begin{equation}
+	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!}
+\end{equation}
+% In appendix??
+The derivation goes as follows.
+We begin with the exact binomial distribution (Eq.(\ref{degDistRan})) which characterizes a random graph.
+In the case $k \ll N$, we can simplify the binomial in the following way \cite{Albert2014}
+\begin{equation}\label{bin}
+	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!}.
+\end{equation}
+We also note the following equality \cite{Albert2014}
+\begin{equation}
+	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg).
+\end{equation}
+Using the series expansion \cite{Albert2014}
+\begin{equation}
+	\ln(1+x) = \sum_{n=1}^{\inf}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ...  \qquad \forall \ |x| \leq 1
+\end{equation}
+we get \cite{Albert2014}
+\begin{equation}
+	\ln[(1-p)^{N-1-k}] \approx (N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right>
+\end{equation}
+which is valid if $k \ll N$. We thus find \cite{Albert2014}
+\begin{equation}\label{e}
+	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>}.
+\end{equation}
+Substituting Eq. (\ref{e}) and Eq. (\ref{bin}) in Eq. (\ref{degDistRan}) and using $\left<k\right> = (N-1)p$, we finally obtain the Poisson form of the degree distribution \cite{Albert2014}
+\begin{equation}
+\begin{split}
+	p_k &= \binom{N-1}{k}p^k (1-p)^{N-1-k} = \frac{(N-1)^k}{k!} p^k \mathrm{e}^{-\left<k\right>} \\
+&= \frac{(N-1)^k}{k!} \bigg(\frac{\left<k\right>}{N-1} \bigg)^k \mathrm{e}^{-\left<k\right>}\\
+&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!}.
+\end{split}
+\end{equation}
 

### 2021-02-06 draft.tex
continued writing + structured things better (included table of content + appendix)

--- 2020-12-18 draft.tex
@@
 \begin{document}
+\frontmatter
+\tableofcontents
+%\listoffigures
 
+\mainmatter
 \chapter{Introduction}
@@
 \subsection{Connectivity}
-% explain connectivity before network communities and also before clustering!
+In physics literature, the degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
+This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.
+Two nodes $u$ and $v$ of a graph are connected if there exist a path between them.
+If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
+The average degree $\left<k\right>$ is a global measurement of the connectivity of the network \cite{Costa2008}.
+If the average connectivity $\left<k\right>$ is too small (this is, if there are too few edges), there will be many isolated nodes and only a few clusters with a small number of nodes.
+If more and more edges are added to the network, the small clusters will grow and will be connected to each other to form larger clusters.
+At some critical value of the connectivity, most nodes will be connected into a giant cluster, which characterizes the percolation of the network \cite{F.Costa2007}.
+% NOT finished; add this as a subsection in the degree section??
+
+\subsection{Network distances}
+
+% explain length, diameter, average length,...
+Several very useful measurements quantify the notion of length and distance in a network/graph.
+We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequent are adjacent nodes in the network \cite{Goddard2010}.
+The length of a path is then defined as the number of edges in the path \cite{Goddard2010}.
+A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}.
+The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}.
+Geodesic distance is often just called distance \cite{Goddard2010}.
+The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\
+\newline
+\textbf{Average distance}\\
+\newline
+The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
+\begin{equation}
+	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}.
+\end{equation}
+This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have by definition a distance equal to infinity) \cite{F.Costa2007}.
+This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for network with a lot of unconnected pairs of nodes.
+These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}.
+Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
+\begin{equation}
+	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}.
+\end{equation}
+This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}.
+This assumes that the efficiency of sending information between two nodes is reciprocal to their distance \cite{F.Costa2007}.
+The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency
+\begin{equation}
+ 	h = \frac{1}{E}
+\end{equation}
+\newline
+\textbf{Diameter and radius}\\
+\newline
+The diameter $diam(G)$ of a network $G$ is the maximum distance between two nodes \cite{Goddard2010}.
+The eccentricity of a node is the maximum distance from that node to any other node.
+The radius $rad(G)$ is the minimum eccentricity among all nodes of $G$.
+Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
+% finished?
 
@@
 We thus derived the degree distribution of a random network. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
+\begin{equation}\label{avDeg}
+\begin{split}
+	\left<k\right> &= \sum_{k=0}^{N-1} k\  p_k  \\
+&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}.
+\end{split}
+\end{equation}
+This equation can be simplified to $\left<k\right> = (N-1)p$, for $N \gg 1$ this approximately becomes $\left<k\right> = Np$.
+The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
+Most real networks are however sparse, which means that $\left<k\right> \ll N$.
+In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014}
+\begin{equation}
+	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!}
+\end{equation}
+The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. \\
+The average clustering coefficient (Eq. (\ref{globalTrans})) of the Erd\H{o}s-R\'{e}nyi network is given by \cite{Clauset2011}
+\begin{equation}
+	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}.
+\end{equation}
+This can be easily understood since the choosing three nodes out of $N$ is given by $\binom{N}{3}$.
+The chance that they are connected by three edges is given by $p^3$, this than gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples.
+If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
+
+% also discuss average clustering and distance of ER
+
+% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
+
+\chapter{Opinion dynamics}
+\label{chap3}
+% talk about opinion dynamics, different models, resistance, etc
+
+\newpage
+
+\appendix
+\chapter{Network models: derivations}
+\section{Derivations of the Erd\H{o}s-R\'{e}nyi network}\label{ER}
+\subsection{Average degree}\label{degree}
+The expected mean degree is given by \cite{Hopcroft2006}
 \begin{equation}\label{avDeg}
@@
 where the right hand side is equal to Equation (\ref{avDeg}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
-% derivation for large k https://barabasi.com/f/624.pdf
-\cite{Albert2014} large N
 
-% continue expression for large N
-
-% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
-
-\chapter{Opinion dynamics}
-\label{chap3}
-% talk about opinion dynamics, different models, resistance, etc
+\subsection{Poisson form of the degree distribution}\label{poiss}
+We begin with the exact binomial distribution (Eq.(\ref{degDistRan}))
+\begin{equation}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
+\end{equation}
+which characterizes a random graph. In the case $k \ll N$, we can simplify the binomial in the following way \cite{Albert2014}
+\begin{equation}\label{bin}
+	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!}.
+\end{equation}
+We also note the following equality \cite{Albert2014}
+\begin{equation}
+	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg).
+\end{equation}
+Using the series expansion \cite{Albert2014}
+\begin{equation}
+	\ln(1+x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ...  \qquad \forall \ |x| \leq 1
+\end{equation}
+we get \cite{Albert2014}
+\begin{equation}
+	\ln[(1-p)^{N-1-k}] \approx (N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right>
+\end{equation}
+which is valid if $k \ll N$. We thus find \cite{Albert2014}
+\begin{equation}\label{e}
+	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>}.
+\end{equation}
+Substituting Eq. (\ref{e}) and Eq. (\ref{bin}) in Eq. (\ref{degDistRan}) and using $\left<k\right> = (N-1)p$, we finally obtain the Poisson form of the degree distribution \cite{Albert2014}
+\begin{equation}
+\begin{split}
+	p_k &= \binom{N-1}{k}p^k (1-p)^{N-1-k} = \frac{(N-1)^k}{k!} p^k \mathrm{e}^{-\left<k\right>} \\
+&= \frac{(N-1)^k}{k!} \bigg(\frac{\left<k\right>}{N-1} \bigg)^k \mathrm{e}^{-\left<k\right>}\\
+&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!}.
+\end{split}
+\end{equation}
 
 \newpage
+\backmatter
 % bibliography

### 2021-02-08 draft.tex
continued writing: WS model + added references

@@
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
-The other two are short distances, also referred to as small-world phenomenon, and high clustering \cite{Muchnik2013}.
+The other two are short distances, also referred to as "small world phenomenon", and high clustering \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 \begin{equation}
-	E \subseteq \{\{x, y\}| x, y \in V \}.
+	E \subseteq \{\{x, y\}| x, y \in V \}
 \end{equation}
@@
 \begin{equation}
-	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \}.
+	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \}
 \end{equation}
@@
 \begin{equation}
-	k_i = \sum_{j=1}^N A_{ij}.
+	k_i = \sum_{j=1}^N A_{ij}
 \end{equation}
@@
 \begin{equation}
-	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}.
+	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}
 \end{equation}
@@
 \begin{equation}
-	s_i = \sum_{j=1}^N w_{ij}.
+	s_i = \sum_{j=1}^N w_{ij}
 \end{equation}
@@
 \begin{equation}
-	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}.
+	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
 \end{equation}
@@
 \begin{equation}
-	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}.
+	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}
 \end{equation}
@@
 \newline
-The diameter $diam(G)$ of a network $G$ is the maximum distance between two nodes \cite{Goddard2010}.
+The diameter $diam(G)$ of a network $G$ is the maximum distance between any pair of nodes \cite{Goddard2010}.
 The eccentricity of a node is the maximum distance from that node to any other node.
@@
 \begin{equation}\label{globalTrans}
-	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}.
+	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}
 \end{equation}
@@
 \begin{equation}
-	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C.
+	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C
 \end{equation}
-\\
 \newline
@@
 \begin{equation}
-	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C).
+	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C)
 \end{equation}
@@
 
-\subsection{The Erd\H{o}s-R\'{e}nyi network}
+\subsection{The Erd\H{o}s-R\'{e}nyi model}
 
@@
 If $p=1$, we obtain a network with the maximum number of edges $N(N-1)/2$. This is called the fully connected network.\\
+\newline
+\textbf{Degree distribution and average degree}\\
+\newline
 The probability that a node is connected to $k$ other nodes and not to $N-k$ others is given by
 \begin{equation}
-	p^k (1-p)^{N-1-k}.
+	p^k (1-p)^{N-1-k}
 \end{equation}
@@
 \begin{equation}\label{degDistRan}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}.
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
 \end{equation}
@@
 	\left<k\right> &= \sum_{k=0}^{N-1} k\  p_k  \\
-&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}.
+&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}
 \end{split}
@@
 The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. \\
+\newline
+\textbf{Average clustering coefficient}\\
+\newline
 The average clustering coefficient (Eq. (\ref{globalTrans})) of the Erd\H{o}s-R\'{e}nyi network is given by \cite{Clauset2011}
 \begin{equation}
-	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}.
+	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}
 \end{equation}
@@
 If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
+\\
+\newline
+\textbf{Diameter and average distance}\\
+\newline
+A heuristic derivation of the diameter of a random network goes as follows. Consider a random network with average degree $\left<k\right>$, then a node in this network has on average \cite{Albert2014}
+\begin{itemize}
+	\item $\left<k\right>$ nodes at distance one $(d=1)$
+	\item $\left<k\right>^2$ nodes at distance two $(d=2)$
+	\item $\left<k\right>^3$ nodes at distance three $(d=3)$
+	\item ...
+	\item $\left<k\right>^d$ nodes at distance $d$
+\end{itemize}
+The number of nodes at a distance $d$ cannot exceed the number of nodes $N$ in the network. We thus obtain the following equation from which the maximum distance or diameter $d_{max}$ can be obtained
+\begin{equation}
+	N \approx \left<k\right>^{d_{max}}
+\end{equation}
+or 
+\begin{equation}\label{diameter}
+	d_{max} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+\end{equation}
+A more precise calculation can be found in Appendix \ref{ER}, Section \ref{ERdist}.
+Equation (\ref{diameter}) is, however, a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
+This is because $d_{max}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is average over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
+Thus the average distance of a random network is defined by \cite{Albert2014}
+\begin{equation}
+	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+\end{equation}
+Since for large $N$ $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
+Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.
+\\
+\\
+To summarize: the random Erd\H{o}s-R\'{e}nyi network has a bell shaped degree distribution which approximates a Poisson distribution for $N \gg 1$; the average clustering coefficient becomes small for large networks and the average distance depends logarithmic on the system sizes.
+The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the "small world phenomenon" (namely short average distance).\\
+Real social networks, however, usually posses the "small world phenomenon" and thus have both a short average distance and a high clustering coefficient.
+The question then rises: can we construct a model that has both of these properties.
+The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
 
-% also discuss average clustering and distance of ER
+\subsection{The Watts-Strogatz model}
+
+In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering as short average distances.
+Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}.
+The construction goes as follows \cite{Watts1998}
+\begin{enumerate}
+	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
+	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done is such a way that self-loops and edge duplication are avoided
+\end{enumerate}
+The probability $p$ allows us to tune the graph between a regular lattice ($p = 0$) and a random network ($p = 1$) \cite{Watts1998}.
+Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$.
+The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected \cite{Watts1998}.\\
+The regular ring lattice ($p=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
+\begin{align}\label{C0}
+	\bar{C}(0) = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
+\end{align}
+% derivations?? http://www.pitt.edu/~kpele/Materials15/module8.pdf put in appendix!! --> to do
+The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
+The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ \cite{Watts1998}
+\begin{align}
+	\bar{C}(1) \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
+\end{align}
+Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph, even for small $p$ \cite{Newman2000}.
+The clustering coefficient is approximately given by   \cite{Barrat1999}
+\begin{equation}
+	\bar{C} \approx \bar{C}(0)(1-p)^3
+\end{equation} 
+where $\bar{C}(0)$ is given by Equation (\ref{C0}). This can be seen in the following way
+\begin{itemize}
+	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-p)^3$
+	\item The probability that edges are rewired back to each other is negligible
+\end{itemize}
+For rather small values of $p$, the clustering coefficient remains high.
+The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
+
+% also discuss case where edges are added instead of rewired??
 
@@
 \begin{equation} 
-	(p+q)^n = \sum_{k=0}^n \binom{n}{k} p^k q^{n-k}.
+	(p+q)^n = \sum_{k=0}^n \binom{n}{k} p^k q^{n-k}
 \end{equation}
@@
 	n(p+q)^{n-1} &= \sum_{k=0}^n \binom{n}{k} k p^{k-1} q^{n-k} \\
-	&= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} q^{n-k}.
+	&= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} q^{n-k}
 \end{split}
@@
 \begin{equation}\label{bin}
-	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!}.
+	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!}
 \end{equation}
@@
 \begin{equation}
-	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg).
+	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg)
 \end{equation}
@@
 \begin{equation}\label{e}
-	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>}.
+	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>}
 \end{equation}
@@
 &= \frac{(N-1)^k}{k!} \bigg(\frac{\left<k\right>}{N-1} \bigg)^k \mathrm{e}^{-\left<k\right>}\\
-&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!}.
+&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!}
 \end{split}
@@
 
+\subsection{Diameter}\label{ERdist}
+The expected number of nodes at a distance $d$ from a starting node is \cite{Albert2014}
+\begin{equation}\label{dist}
+	N(d) = 1 + \left<k\right> + \left<k\right>^2 + ... + \left<k\right>^d = \frac{\left<k\right>^{d+1} - 1}{\left<k\right> - 1}
+\end{equation}
+$N(d)$ cannot exceed the number of nodes $N$ in the network. The maximum distance or diameter $d_{max}$ of the network can thus be found by setting \cite{Albert2014}
+\begin{equation}
+	N(d_{max}) \approx N
+\end{equation}
+If $\left<k\right> \gg 1$, we can neglect the $-1$ terms in Eq. (\ref{dist}), which gives \cite{Albert2014}
+\begin{equation}
+	N \approx \left<k\right>^{d_{max}}
+\end{equation}
+and thus the diameter of a random network becomes \cite{Albert2014}
+\begin{equation}
+	d_{max} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+\end{equation}
 \newpage

### 2021-02-09 draft.tex
continued writing: assortativity, modularity, etc

@@
 
+\subsection{Assortative and disassortative networks}
+
+Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks.
+More precisely, the notion of assortativity and disassortativity has to do with degree correlations.
+The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}.
+The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
+\begin{equation}
+	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)}
+\end{equation}
+The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
+\begin{equation}
+	k_{nn}(k) = \sum_{k'} k'P(k'|k)
+\end{equation}
+If there are no correlations $k_{nn}$ is independent of $k$ \cite{F.Costa2007}.
+If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative.
+If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
+However, the terms assortativity and disassortativity are also often used without taking degree correlations into account.
+Instead, one may look at a particular node property to define assortative and disassortative network \cite{Thed2014}.
+In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.
+
+% talk about assortativity of social networks?? + talk about node assortativity instead of global assortativity, see Thed2014?
+
 \subsection{Network distances}
@@
 The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
-\begin{equation}
+\begin{equation}\label{avdist}
 	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
@@
 
+\subsection{Modularity}
+
+The modularity $Q$ is a measurement that represents the quality of a particular division of a network into different communities \cite{F.Costa2007}.
+It was proposed after the fundamental problem arised for real networks concerning how to best divide the network into its constituent communities \cite{F.Costa2007}.
+Generally, no a priori information is available about the number of existing communities in real networks \cite{F.Costa2007}.\\
+The modularity for a network with $N$ nodes and $L$ edges and a partition into $n_c$ communities is calculated as follows.
+Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
+The modularity of a community $Q_c$ is then calculated by measuring the difference between the network real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired \cite{Albert2016}.
+This last quantity is given by $p_{ij}$ which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
+Hence, the modularity of a community $Q_c$ becomes \cite{Albert2016}
+\begin{equation}
+	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij})
+\end{equation}
+Using \cite{Albert2016}
+\begin{equation}
+	p_{ij} = \frac{k_ik_j}{2L}
+\end{equation}
+we can derive a simpler form for $Q_c$ \cite{Albert2016}
+% derivation in appendix, see albert2016
+\begin{equation}\label{Qc}
+	Q_c = \frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2
+\end{equation}
+where $L_c$ is the total number of links in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.\\
+To generalize this idea to the whole network, we sum Eq. (\ref{Qc}) over all $n_c$ communities \cite{Albert2016}
+\begin{equation}
+	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]
+\end{equation}
+% properties of modularity albert2016
+% NOT FINISHED
+
 % talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
 % explain small-world (milgram), high clustering, power-law
-% explain modularity!! + diameter, shortest path, etc...!
+% explain modularity!!
 
@@
 \end{equation}
-This can be easily understood since the choosing three nodes out of $N$ is given by $\binom{N}{3}$.
-The chance that they are connected by three edges is given by $p^3$, this than gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples.
+This can be easily understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$.
+The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples.
 If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
@@
 Equation (\ref{diameter}) is, however, a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
-This is because $d_{max}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is average over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
+This is because $d_{max}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
 Thus the average distance of a random network is defined by \cite{Albert2014}
@@
 \end{equation}
-Since for large $N$ $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
+Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
 Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.
@@
 The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the "small world phenomenon" (namely short average distance).\\
-Real social networks, however, usually posses the "small world phenomenon" and thus have both a short average distance and a high clustering coefficient.
+Real social networks, however, usually display the "small world phenomenon" and thus have both a short average distance and a high clustering coefficient.
 The question then rises: can we construct a model that has both of these properties.
 The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
+% also take about the giant cluster, critical probability??
 
@@
 
-In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering as short average distances.
+In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances.
 Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}.
@@
 \end{align}
-% derivations?? http://www.pitt.edu/~kpele/Materials15/module8.pdf put in appendix!! --> to do
-The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
+The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}.
+The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$.
+\\
 The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ \cite{Watts1998}
@@
 
+\subsection{The stochastic block model}
+
+Another random graph model that produces networks containing communities is the stochastic block model (SBM).
+The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
+The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected than with the rest of the network \cite{F.Costa2007}.
+Hence, these networks posses a modular/community structure and this exactly the structure that is easily modeled by the stochastic block model.
+% NOT FINISHED
+
 % different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
@@
 \appendix
-\chapter{Network models: derivations}
-\section{Derivations of the Erd\H{o}s-R\'{e}nyi network}\label{ER}
+\chapter{Network models: Derivations}
+\section{The Erd\H{o}s-R\'{e}nyi model}\label{ER}
 \subsection{Average degree}\label{degree}
@@
 \end{equation}
+
+\section{The Watts-Strogatz model}\label{WS}
+\subsection{The regular ring lattice}\label{ringclus}
+
+\textbf{Clustering coefficient}\\
+\newline
+A triangle in the regular ring lattice requires two edge traversals in the same direction and one in the opposite direction \cite{Pele2015}.
+In the regular ring lattice every node has a fixed degree $K$, this means that each node is connected to $K/2$ neighbors on its left and $K/2$ neighbors on its right.
+It is then clear that the final, backwards step can span at most $K/2$ nodes \cite{Pele2015}; otherwise we would end up at a different node from which we started, which would definitely not form a triangle.\\
+Hence, the number of triangles for a given node is given by choosing the two forward target nodes from the $K/2$ possibilities \cite{Pele2015}
+\begin{equation}\label{triangle}
+	\binom{\frac{K}{2}}{2} = \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)
+\end{equation}
+The number of connected triples per node is given by choosing two nodes out of the $K$ neighbors \cite{Pele2015}
+\begin{equation}\label{triple}
+	\binom{K}{2} = \frac{1}{2} K (K-1)
+\end{equation}
+Since every node in the regular ring lattice we can easily find the total number of triangles and connected triples in the network by multiplying Eq.
+(\ref{triangle}) and Eq.
+(\ref{triple}) with the number of nodes $N$.
+We then obtain the final result for the clustering coefficient in the regular ring lattice \cite{Pele2015}
+\begin{equation}
+	\bar{C} = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)}
+\end{equation}
+\newline
+\textbf{Average distance}\\
+\newline
+The general formula for the average distance is given by Eq. (\ref{avdist})
+\begin{equation}
+	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
+\end{equation}
+where the sum runs over the shortest path between every pair of nodes.
+Assume we have a regular ring lattice with $N$ nodes and each node has $K$ neighbors.
+Now we can calculate the distance between node $i$ and every other node in the ring lattice
+\begin{itemize}
+	\item There are $K$ nodes adjacent to node $i$, so there are $K$ nodes with a distance $d_{ij} = 1$
+	\item There are $K$ nodes that are neighbors of neighbors of node $i$, so there are $K$ nodes with a distance $d_{ij} = 2$
+	\item ...
+	\item There are $K$ nodes that have a distance $d_{ij} = \frac{N}{K}$ between them and node $i$; its easy to see that these are the nodes that are the furthest away from node $i$ 
+\end{itemize}  
+So the average distance for node $i$ becomes \cite{chap9}
+\begin{equation}
+	l_i = \frac{1}{N-1} K \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg)
+\end{equation}
+where the last approximation holds for $N \gg 1$. Since every node in the regular ring lattice is identical we can easily find the average distance of the network \cite{chap9}
+\begin{equation}
+	l = \frac{1}{N} \sum_i l_i = \frac{1}{N} N l_i \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg)
+\end{equation}
+Using \cite{chap9}
+\begin{equation}
+	1 + 2+ 3 +...+ x = \sum_{i = 1}^x i = \frac{x(x+1)}{2} \approx \frac{x^2}{2}
+\end{equation}
+where the last approximation again holds for $x \gg 1$; we finally obtain \cite{chap9}
+\begin{equation}
+	l \approx \frac{K}{N} \frac{\frac{N^2}{K^2}}{2} = \frac{N}{2K}
+\end{equation}
+
 \newpage

### 2021-02-10 draft.tex
continued writing + corrected some small mistakes

@@
 \usepackage{physics}
+\usepackage{interval}
 \usepackage[noend]{algpseudocode}
@@
 This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have by definition a distance equal to infinity) \cite{F.Costa2007}.
-This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for network with a lot of unconnected pairs of nodes.
+This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes.
 These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}.
@@
 \newline
-A strong community is defined such that each node inside the community has more links to other nodes inside the same community than to nodes outside the community \cite{Albert2016}.
+A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}.
 Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
-If we define the internal degree $k_i^{int}(C)$ as the number of links of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of links of node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
+If we define the internal degree $k_i^{int}(C)$ as the number of edges of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges of node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
 \begin{equation}
@@
 Generally, no a priori information is available about the number of existing communities in real networks \cite{F.Costa2007}.\\
-The modularity for a network with $N$ nodes and $L$ edges and a partition into $n_c$ communities is calculated as follows.
+The modularity for a network with $N$ nodes and $L$ edges and partitioned into $n_c$ communities is calculated as follows.
 Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
-The modularity of a community $Q_c$ is then calculated by measuring the difference between the network real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired \cite{Albert2016}.
+The modularity of a community $Q_c$ is then calculated by measuring the difference between the network real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
 This last quantity is given by $p_{ij}$ which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
@@
 \end{equation}
-Using \cite{Albert2016}
+where the factor $\frac{1}{2L}$ is just a normalization factor. Using \cite{Albert2016}
 \begin{equation}
@@
 we can derive a simpler form for $Q_c$ \cite{Albert2016}
-% derivation in appendix, see albert2016
 \begin{equation}\label{Qc}
@@
 \end{equation}
-where $L_c$ is the total number of links in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.\\
+where $L_c$ is the total number of edges in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.\\
 To generalize this idea to the whole network, we sum Eq. (\ref{Qc}) over all $n_c$ communities \cite{Albert2016}
-\begin{equation}
+\begin{equation}\label{mod}
 	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]
 \end{equation}
-% properties of modularity albert2016
-% NOT FINISHED
+The derivation of this formula and thus of Eq.
+(\ref{Qc}) can be found in Appendix \ref{modul}, Section \ref{simplemod}.
+The modularity $Q$ can have values in the interval $\interval{-1}{1}$ and has the following properties \cite{Albert2016}
+\begin{itemize}
+	\item The higher the value of $Q$ is for a partition, the better is the corresponding community structure.
+This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure.
+	\item A modularity $Q$ equal to zero corresponds to the situation where all nodes are put in the same community, this is if the whole network is put into a single community.
+Then the two terms in Eq.
+(\ref{mod}) become equal and $Q$ vanishes.
+	\item If every node is put into a separate community, the first term in Eq. (\ref{mod}) vanishes and we are left with $n_c$ negative terms. Hence, this situation corresponds to a negative modularity.
+\end{itemize}
+As said before, modularity is often used in community detection algorithms.
+It is thus important to be aware of the limitations of modularity.
+One of the biggest shortcomings of modularity maximization as community detection is its resolution limit.
+Modularity maximization forces small communities into larger ones \cite{Albert2016}.
+If two communities $A$ and $B$ are merged the modularity changes by \cite{Albert2016}
+\begin{equation}
+	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2}
+\end{equation}
+where $l_{AB}$ is the number of direct edges that connect nodes in community $A$ with nodes in community $B$; $k_A$ and $k_B$ are the total degree of nodes in communities $A$ and $B$ respectively.
+The derivation of this formula is found in Appendix \ref{modul}, Section \ref{modchange}.
+If $A$ and $B$ are distinct communities, they should remain distinct when $Q$ is maximized \cite{Albert2016}.
+This is however not always the case.
+Consider the case where $\frac{k_Ak_B}{2L} < 1$ and there is at least one edge between $A$ and $B$ ($l_{AB} \geqslant 1$), then $\Delta Q_{AB} \geqslant 0$ and the modularity increases by merging the two communities \cite{Albert2016}.
+Assume for simplicity that $k_A = k_B = k$, then modularity is increased by merging $A$ and $B$ if
+\begin{equation}\label{reslim}
+	k \leqslant \sqrt{2L}
+\end{equation}
+and if there is at least one edge between the two communities; even if $A$ and $B$ are distinct communities \cite{Albert2016}!
+Equation (\ref{reslim}) is called the resolution limit: communities that are smaller than this limit will not be detected by modularity maximization \cite{Albert2016}.
 
@@
 % explain small-world (milgram), high clustering, power-law
-% explain modularity!!
 
@@
 \end{equation}
-This equation can be simplified to $\left<k\right> = (N-1)p$, for $N \gg 1$ this approximately becomes $\left<k\right> = Np$.
+This equation can be simplified to $\left<k\right> = (N-1)p$, and for $N \gg 1$ this approximately becomes $\left<k\right> = Np$.
 The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
@@
 Hence, these networks posses a modular/community structure and this exactly the structure that is easily modeled by the stochastic block model.
+The stochastic block model is often used as a benchmark for community detection algorithms.\\
+The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}
+\begin{itemize}
+	\item The number of nodes $N$ 
+	\item A partition of the node set $\{1, 2, ..., N\}$ into disjoint subsets $C_1, ..., C_k$, these are the communities
+	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$ represent the edge probabilities inside the community $C_i$ and the other elements $M_{ij}$, $i, j = 1, ..., k$ and $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
+\end{itemize}
+% discuss case where probailities on diag are bigger than off-diag --> assortative SBM
 % NOT FINISHED
+% clauset2017 http://tuvalu.santafe.edu/~aaronc/courses/5352/csci5352_2017_L6.pdf
 
@@
 \appendix
+\chapter{Network measurements: Derivations}
+\section{Modularity}\label{modul}
+\subsection{Simple form of network modularity}\label{simplemod}
+
+Using \cite{Albert2016}
+\begin{equation}
+	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij})
+\end{equation}
+and \cite{Albert2016}
+\begin{equation}
+	p_{ij} = \frac{k_ik_j}{2L}
+\end{equation}
+we can write the modularity of the whole network as \cite{Albert2016}
+\begin{equation}\label{modnetw}
+	Q = \frac{1}{2L}\sum_{i, j = 1}^N \bigg(A_{ij} - \frac{k_ik_j}{2L}\bigg) \delta_{c_i, c_j}
+\end{equation}
+where $c_i$ represents the community to which node $i$ belongs and where the Dirac-delta function $\delta_{c_i, c_j}$ expresses that only pairs of nodes that belong to the same community are taken into account.
+We can then rewrite the first term as a sum over communities \cite{Albert2016}
+\begin{equation}\label{1}
+	\frac{1}{2L}\sum_{i, j = 1}^N A_{ij}\delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{2L}\sum_{i, j \in C_c} A_{ij} = \sum_{c=1}^{n_c}\frac{L_c}{L}
+\end{equation}
+where $L_c$ is the number of edges in community $C_c$ and the factor 2 disappears because each edge is counted twice in $A_{ij}$.
+Similarly, the second term in Eq.
+(\ref{modnetw}) can be rewritten as \cite{Albert2016}
+\begin{equation}\label{2}
+	\frac{1}{2L}\sum_{i, j = 1}^N \frac{k_ik_j}{2L} \delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{(2L)^2}\sum_{i, j \in C_c} k_ik_j = \sum_{c=1}^{n_c}\frac{k_c^2}{4L^2}
+\end{equation}
+where $k_c$ is the total degree of the nodes in community $C_c$. Combining Eq. (\ref{1}) and Eq. (\ref{2}) gives the final result \cite{Albert2016}
+\begin{equation}\label{Qfin}
+	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]
+\end{equation}
+
+\subsection{Modularity change after merging two communities}\label{modchange}
+
+Consider two communities $A$ and $B$ and denote with $k_A$ and $k_B$ the total degree of communities $A$ and $B$ respectively \cite{Albert2016}.
+The change in modularity after merging the two communities is calculated as follows (using Eq.
+(\ref{Qfin})) \cite{Albert2016}
+\begin{equation}\label{Q_AB}
+	\Delta Q_{AB} = \bigg[\frac{L_{AB}}{L} - \bigg(\frac{k_{AB}}{2L} \bigg)^2 \bigg] - \bigg[\frac{L_A}{L} - \bigg(\frac{k_A}{2L} \bigg)^2 + \frac{L_B}{2L} - \bigg(\frac{k_B}{2L} \bigg)^2 \bigg]
+\end{equation}
+where $L_{AB} = L_A + L_B + l_{AB}$ is the total number of edges in the merged community; $L_A$ and $L_B$ are the number of edges in communities $A$ and $B$ respectively and $l_{AB}$ is the number of direct edges between communities $A$ and $B$ \cite{Albert2016}.
+$k_{AB} = k_A + k_B$ is the total degree of nodes in the merged community \cite{Albert2016}.
+Inserting the formulas for $L_{AB}$ and $k_{AB}$ in Eq.
+(\ref{Q_AB}) gives
+\begin{equation}
+	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2}
+\end{equation}
+
 \chapter{Network models: Derivations}
@@
 The expected mean degree is given by \cite{Hopcroft2006}
-\begin{equation}\label{avDeg}
+\begin{equation}\label{avDegr}
 \begin{split}
@@
 \end{equation}
-where the right hand side is equal to Equation (\ref{avDeg}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
+where the right hand side is equal to Equation (\ref{avDegr}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
 
@@
 We begin with the exact binomial distribution (Eq.(\ref{degDistRan}))
-\begin{equation}
+\begin{equation}\label{pk}
 	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
@@
 \end{equation}
-Using the series expansion \cite{Albert2014}
+where we made use of the fact that $\left<k\right> = (N-1)p$. Using the series expansion \cite{Albert2014}
 \begin{equation}
-	\ln(1+x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ...  \qquad \forall \ |x| \leq 1
+	\ln(1+x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ...  \qquad \forall \ |x| \leqslant 1
 \end{equation}
@@
 \end{equation}
-Substituting Eq. (\ref{e}) and Eq. (\ref{bin}) in Eq. (\ref{degDistRan}) and using $\left<k\right> = (N-1)p$, we finally obtain the Poisson form of the degree distribution \cite{Albert2014}
+Substituting Eq. (\ref{e}) and Eq. (\ref{bin}) in Eq. (\ref{pk}) and using $\left<k\right> = (N-1)p$, we finally obtain the Poisson form of the degree distribution \cite{Albert2014}
 \begin{equation}
@@
 \end{equation}
-Since every node in the regular ring lattice we can easily find the total number of triangles and connected triples in the network by multiplying Eq.
+Since every node in the regular ring lattice is identical, we can easily find the total number of triangles and connected triples in the network by multiplying Eq.
 (\ref{triangle}) and Eq.
@@
 	\item There are $K$ nodes adjacent to node $i$, so there are $K$ nodes with a distance $d_{ij} = 1$
-	\item There are $K$ nodes that are neighbors of neighbors of node $i$, so there are $K$ nodes with a distance $d_{ij} = 2$
+	\item There are $K$ nodes that are neighbors of the neighbors of node $i$, so there are $K$ nodes with a distance $d_{ij} = 2$
 	\item ...

### 2021-02-11 draft.tex
continued writing on the SBM

@@
 In order to study the topological structures of real-world networks, several theoretical network models have been developed.
-These theoretical models usually have a simpler representation and well known properties that can be derived mathematically.
+These theoretical models usually have a simpler representation and well known properties that can be derived analytically.
 They are widely studied and used extensively.
@@
 
-In 1959 the two mathematicians Paul Erd\H{o}s and Alfr\'{e}d R\'{e}nyi proposed a model to generate random networks \cite{F.Costa2007}.
+In 1959 the two mathematicians Paul Erd\H{o}s and Alfr\'{e}d R\'{e}nyi proposed a model to generate simple random networks \cite{F.Costa2007}.
 Independently of them, Solomonoff and Rapoport already proposed the model in 1951.
@@
 The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
-Most real networks are however sparse, which means that $\left<k\right> \ll N$.
-In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014}
+Most real networks are sparse, which means that $\left<k\right> \ll N$. In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014} 
 \begin{equation}
@@
 The average clustering coefficient (Eq. (\ref{globalTrans})) of the Erd\H{o}s-R\'{e}nyi network is given by \cite{Clauset2011}
-\begin{equation}
+\begin{equation}\label{C}
 	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}
@@
 A more precise calculation can be found in Appendix \ref{ER}, Section \ref{ERdist}.
-Equation (\ref{diameter}) is, however, a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
+Equation (\ref{diameter}) might, however, be a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
 This is because $d_{max}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
-Thus the average distance of a random network is defined by \cite{Albert2014}
-\begin{equation}
+Hence, the average distance of a random network is defined by \cite{Albert2014}
+\begin{equation}\label{d}
 	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}}
 \end{equation}
-Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
+Since for large $N$, $\ln{N} \ll N$, and we find that the average distance is a lot smaller than the size of the network.
 Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.
@@
 Real social networks, however, usually display the "small world phenomenon" and thus have both a short average distance and a high clustering coefficient.
-The question then rises: can we construct a model that has both of these properties.
+The question then rises: can we construct a model that has both of these properties?
 The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
@@
 	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
-	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done is such a way that self-loops and edge duplication are avoided
+	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done in such a way that self-loops and edge duplication are avoided
 \end{enumerate}
@@
 \\
-The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ \cite{Watts1998}
+The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
 \begin{align}
@@
 
-Another random graph model that produces networks containing communities is the stochastic block model (SBM).
+Another random graph model is one that produces networks containing a community structure, namely the stochastic block model (SBM).
 The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
-The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected than with the rest of the network \cite{F.Costa2007}.
-Hence, these networks posses a modular/community structure and this exactly the structure that is easily modeled by the stochastic block model.
-The stochastic block model is often used as a benchmark for community detection algorithms.\\
+The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected with each other than with the rest of the network \cite{F.Costa2007}.
+Hence, these networks posses a modular/community structure and this is exactly the structure that is easily modeled by the stochastic block model.
+The stochastic block model is also often used as a benchmark for community detection algorithms.\\
 The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}
@@
 	\item The number of nodes $N$ 
-	\item A partition of the node set $\{1, 2, ..., N\}$ into disjoint subsets $C_1, ..., C_k$, these are the communities
-	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$ represent the edge probabilities inside the community $C_i$ and the other elements $M_{ij}$, $i, j = 1, ..., k$ and $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
+	\item A partition of the node set $\{1, 2, ..., N\}$ into disjoint subsets $C_1, ..., C_k$; these groups are the communities
+	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$ represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
 \end{itemize}
-% discuss case where probailities on diag are bigger than off-diag --> assortative SBM
+There are no real restrictions to the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures.
+However, the edge probabilities inside a community are often taken to be bigger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
+In this way, the stochastic block model generates a network with a clear community structure.
+A stochastic block model that fulfills this condition is called an assortative SBM \cite{Lee2019}.
+A disassortative stochastic block model refers to the case where the edge probabilities between the communities are higher than those inside the communities ($M_{ii} < M_{ij}$, $i \neq j$) \cite{Gribel2020}.\\
+The stochastic block model includes the Erd\H{o}s-R\'{e}nyi model \cite{Clauset2017}.
+This can be seen in the two different ways.
+First, consider the case where $k = 1$, thus we only have one group.
+The matrix $M$ then reduces to one number $p$ and we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model \cite{Clauset2017}.
+Second, assume we have more than one group or community, $k > 1$, and assume that every element of the matrix $M$ is the same, $M_{ij} = p$ \cite{Clauset2017}.
+Again, we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model.
+In both cases, the mathematical properties of the Erd\H{o}s-R\'{e}nyi model are obtained: a Poisson degree distribution, an average distance that is logarithmic with the system size $N$, vanishing clustering coefficient for large $N$, etc \cite{Clauset2017}.
+The structure within the groups or communities of the stochastic block model is that of a random graph, while the structure between the communities is that of a random bipartite graph \cite{Clauset2017}.\\
+In this thesis, the stochastic block model is generated in the following way.
+First, the parameters of the model are specified.
+Then, for each community $C_i$ an Erd\H{o}s-R\'{e}nyi network is constructed, where the number of nodes equals the size of the community $C_i$ and the edge probability $p$ is given by $M_{ii}$.
+Finally, edges are added between communities $C_i$ and $C_j$ with probability $p = M_{ij}$.\\
+The stochastic block model can be fitted to data \cite{Clauset2017}.
+Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.
+One of the shortcomings of fitting the SBM to infer the community structure, is that real-world networks often have a heterogeneous or power law degree distribution.
+Since the SBM is made of random networks inside each group and random bipartite networks between each pair of groups, the degree distribution of the full network is always a mixture of Poisson distributions \cite{Clauset2017}.
+The consequence of this is that if a network possesses a more skewed degree distribution, the SBM will put the nodes with a high degree in a small group, together with a high probability of connecting to other, larger groups \cite{Clauset2017}.
+The reason behind this is one of the main assumptions of the stochastic block model, namely the fact that the edge probabilities only depend  on which group the node belongs to \cite{Clauset2017}.
+In other words, nodes that belong to the same group are stochastically  equivalent, meaning that they have equivalent connectivity patterns to other nodes \cite{Clauset2017}.
+However, a network with a skewed degree distribution violates this assumption of edge independence and the SBM seeks to explain this over other kinds of latent group structures that we might care about \cite{Clauset2017}.
+Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation \cite{Karrer2011}.
+% talk about community sizes from power law??
 % NOT FINISHED
-% clauset2017 http://tuvalu.santafe.edu/~aaronc/courses/5352/csci5352_2017_L6.pdf
 

### 2021-02-12 draft.tex
continued writing on real world social networks + introduction on opinion dynamics

@@
 % NOT FINISHED
+% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
 
-% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
+\section{Real-world social networks}
+The models discussed in the sections above, describe real-world networks in an idealized, simplified way.
+They may posses some of the characteristics of real-world networks, but they certainly do not always describe real-world networks appropriately.
+The question then arises: what are the main properties/characteristics of real-world networks, and more specifically of real-world social networks?
+Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
+As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties.
+These are the two properties of the "small world phenomenon", namely a short average distance and a high clustering coefficient, and a heterogeneous or power law degree distribution \cite{RealWorld}.
+These properties are valid for both off-line and on-line social networks \cite{Zhang2014}.
+Lets discuss them in a bit more detail.\\
+\newline
+\textbf{Short average distance}\\
+\newline
+In the late 1960's Stanley Milgram performed a, what is now famous, experiment.
+He asked randomly selected people in Boston and Omaha to forward a letter to a distant target person.
+The letter could, however, only be sent to acquaintances, thought to be closer to the target person \cite{RealWorld}.
+The remarkable result of this experiment was that the average number of steps between the sender and the target person was only around six.
+This phenomenon is now often referred to as "six degrees of separation" \cite{RealWorld}.\\
+This phenomenon is also observed in many other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
+Such networks usually have an average distance that depends logarithmically on the system size instead of linearly.
+A short average distance facilitates a fast transmission of information \cite{Zhang2014}.\\
+\newline
+\textbf{High clustering coefficient}\\
+\newline
+In real-world social networks, one often observes the pattern that two acquaintances/friends of an individual are also acquaintances/friends of each other \cite{RealWorld}.
+In graph terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
+This property is not only found in social networks, but also in a wide variety of different real-world networks \cite{RealWorld}.\\
+\newline
+\textbf{Heterogeneous or power law degree distribution}\\
+\newline
+Many real-world networks are found to have a power law degree distribution \cite{RealWorld}
+\begin{equation}
+	P(k) \sim k^{-\gamma}
+\end{equation}
+where the exponent $\gamma$ often lies between 2 and 3.
+Networks with a power law degree distribution are also called heterogeneous or scale free networks.
+The term scale free refers to the fact that there is no characteristic scale in the network.
+Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$, the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
+If the exponent $\gamma$ is smaller or equal to 2 ($\gamma \leqslant 2$), the average degree diverges; if $2 < \gamma < 3$, there is a finite average degree, but a diverging variance \cite{Newman2005}.\\
+\newline
+As said earlier, these three main characteristics are the same for on- and off-line social networks.
+Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}.
+They also often posses a modular/community structure \cite{Ferrara2012}\cite{McGlohon2011}.
+There are, however, some properties that are different between on-line and off-line networks as well.
+One of them might be the degree correlations/mixing pattern.
+It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern \cite{Hu2009}\cite{Zhang2014}.
+For a long time, it was thought that this is also valid for on-line social networks.
+However, recent research has shown that a lot of on-line social networks display disassortative mixture \cite{Hu2009}\cite{Zhang2014}.
+This discrepancy between real-life social networks and on-line social networks may be explained as follows.
+In real-life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}.
+The `ordinary people' don't have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
+In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance/influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
+This leads to a disassortative mixing pattern.
+% also talk about size of groups that people can maintain?
 
@@
 \label{chap3}
+The study of opinion dynamics in a group of people, as already mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
+The network layer that describes the underlying structure and correlations between the agents/people is intensively described in the previous section.
+In this section we will focus on the second important layer, namely that of the opinion dynamics models.
+These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
+Hence, these opinion dynamics are the dynamics that run on the underlying network structure.
+Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.
+Some major attention will also be given to the content curation/filtering algorithms used by social media companies.
+This is because this thesis deals with opinion dynamics on social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.
 % talk about opinion dynamics, different models, resistance, etc

### 2021-02-13 draft.tex
continued writing opinion dynamics chapter

@@
 \usepackage{interval}
+\usepackage[colorlinks,citecolor=red,urlcolor=blue,bookmarks=false,hypertexnames=true]{hyperref} 
 \usepackage[noend]{algpseudocode}
@@
 This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale.
-It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings.
+It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings \cite{Perra2019}.
 The broad range of interactions people undergo on social media with people from all over the world can thus not be underestimated in the process of opinion formation.
@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links (edges) that indicate the interactions between the nodes \cite{Mata2020}.
+A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes \cite{Mata2020}.
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
@@
 In temporal networks time becomes an explicit element in the network representation \cite{Holme2012}.
-Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.
+Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.\\
 Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc.
 These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks.
-It is found that many real life social systems have a power law degree distribution \cite{Muchnik2013}.
+It is found that many real-life social systems have a power law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
-The other two are short distances, also referred to as "small world phenomenon", and high clustering \cite{Muchnik2013}.
+The other two are short distances and high clustering, these two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 \\
+\newline
 One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system.
@@
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others.
+Furthermore, the agents can have a parameter that determines whether they are active or not.
 \\
@@
 This branch of physics offers a framework that relates microscopic properties of atoms, molecules,...
-to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
+to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.\\
 The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
@@
 Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\
-In one of their previous papers (\cite{Perra2019}) prof.
+\newline
+In one of their previous papers (see reference \cite{Perra2019}) prof.
 Nicola Perra and prof.
@@
 Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
-They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).
-Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers.
-In the case of nudging, they found that the population opinion moved to the nudged opinion relatively fast, even in the case of small nudging.
+They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).\\ Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}.
+On the other hand, a more heterogeneous contact pattern could hamper this formation of echo chambers \cite{Perra2019}.
+If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
+In the case of nudging, they found that the population opinion moved to the nudged opinion relatively fast, even in the case of small nudging.\\
 They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
@@
 We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can observe convergence to one opinion inside the communities.
-On the other hand we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network.
+On the other hand, we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network.
 
@@
 The nodes $x$ and $y$ of an edge $\{x, y\}$ are called the endpoints of the edge.
-It is possible that nodes are not joined by any edge, such nodes are called isolated.
+It is possible that nodes are not joined by any edge, such nodes are called isolated or disconnected.
 If two or more edges have the same pair of endpoints, the graph is called a multi-graph.
@@
 In physics literature, the degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
-This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.
-Two nodes $u$ and $v$ of a graph are connected if there exist a path between them.
-If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
+This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.\\
+Two nodes $u$ and $v$ of a graph are connected if there exist a path between them. If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
 The average degree $\left<k\right>$ is a global measurement of the connectivity of the network \cite{Costa2008}.
 If the average connectivity $\left<k\right>$ is too small (this is, if there are too few edges), there will be many isolated nodes and only a few clusters with a small number of nodes.
-If more and more edges are added to the network, the small clusters will grow and will be connected to each other to form larger clusters.
-At some critical value of the connectivity, most nodes will be connected into a giant cluster, which characterizes the percolation of the network \cite{F.Costa2007}.
+If more and more edges are added to the network, the small clusters will grow and will tend to connect to each other to form larger clusters.
+At some critical value of the connectivity, most nodes will be connected into a giant cluster (the giant component), which characterizes the percolation of the network \cite{F.Costa2007}.
 % NOT finished; add this as a subsection in the degree section??
@@
 More precisely, the notion of assortativity and disassortativity has to do with degree correlations.
+\\
 The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}.
@@
 \end{equation}
-If there are no correlations $k_{nn}$ is independent of $k$ \cite{F.Costa2007}.
+If there are no correlations $k_{nn}$ is independent of $k$, and such a network is called a neutral network \cite{F.Costa2007}.
 If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative.
@@
 Several very useful measurements quantify the notion of length and distance in a network/graph.
-We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequent are adjacent nodes in the network \cite{Goddard2010}.
+We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
 The length of a path is then defined as the number of edges in the path \cite{Goddard2010}.
@@
 The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\
+\newline
 \newline
@@
 \end{equation}
-This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have by definition a distance equal to infinity) \cite{F.Costa2007}.
+This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}.
 This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes.
@@
 It is a measure for the number of triangles in a network and determines the connectivity in the neighborhood of a node $i$: if a node $i$ has a high clustering coefficient, its neighbors are likely to be directly connected to each other \cite{Li2017}.
-Here a triangle is defined as a loop of length three, this is a sequence of nodes $x, y, z, x$ such that $\{x, y\}, \{y, z\}$ and $\{z, x\}$ are edges of the network.
+\\Here a triangle is defined as a loop of length three, this is a sequence of nodes $x, y, z, x$ such that $\{x, y\}, \{y, z\}$ and $\{z, x\}$ are edges of the network.
 The clustering coefficient is thus a way to measure the degree to which nodes in a network tend to cluster \cite{Li2017}.
@@
 \\
+\newline
 A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as (valid for undirected, unweighted networks) \cite{F.Costa2007}
@@
 \begin{align}
-	N_{\Delta} &= \sum_{k > j > i} a_{ij}a_{ik}a_{jk} \\
-	N_3 &= \sum_{k > j > i} (a_{ij}a_{ik} + a_{ji}a_{jk} + a_{ki}a_{kj})
+	N_{\Delta} &= \sum_{k > j > i} A_{ij}A_{ik}A_{jk} \\
+	N_3 &= \sum_{k > j > i} (A_{ij}A_{ik} + A_{ji}A_{jk} + A_{ki}A_{kj})
 \end{align}
-where $a_{ij}$ are the elements of the adjacency matrix. It is possible to define the transitivity of one node as \cite{F.Costa2007}
+where $A_{ij}$ are the elements of the adjacency matrix. It is possible to define the transitivity of one node as \cite{F.Costa2007}
 \begin{equation}\label{trans}
@@
 \begin{align}
-	N_{\Delta}(i) &= \sum_{k > j} a_{ij}a_{ik}a_{jk} \\
-	N_3(i) &= \sum_{k > j} a_{ij}a_{ik}
+	N_{\Delta}(i) &= \sum_{k > j} A_{ij}A_{ik}A_{jk} \\
+	N_3(i) &= \sum_{k > j} A_{ij}A_{ik}
 \end{align}
@@
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
-In the case of the average clustering coefficient $\bar{C}$ and the global transitivity $T$ this is however not the case.
+In the case of the average clustering coefficient $\bar{C}$ and the global transitivity $T$ this is, however, not the case.
 The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node.
@@
 Modularity maximization forces small communities into larger ones \cite{Albert2016}.
-If two communities $A$ and $B$ are merged the modularity changes by \cite{Albert2016}
+If two communities $A$ and $B$ are merged, the modularity changes by \cite{Albert2016}
 \begin{equation}
@@
 The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. \\
+\newline
+\newline
 \newline
@@
 The stochastic block model is also often used as a benchmark for community detection algorithms.\\
+\newline
 The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}
@@
 A disassortative stochastic block model refers to the case where the edge probabilities between the communities are higher than those inside the communities ($M_{ii} < M_{ij}$, $i \neq j$) \cite{Gribel2020}.\\
+\newline
 The stochastic block model includes the Erd\H{o}s-R\'{e}nyi model \cite{Clauset2017}.
@@
 The stochastic block model can be fitted to data \cite{Clauset2017}.
-Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.
+Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.\\
 One of the shortcomings of fitting the SBM to infer the community structure, is that real-world networks often have a heterogeneous or power law degree distribution.
@@
 Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
+\newline
 As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties.
@@
 Lets discuss them in a bit more detail.\\
+\newline
 \newline
@@
 Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$, the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
-If the exponent $\gamma$ is smaller or equal to 2 ($\gamma \leqslant 2$), the average degree diverges; if $2 < \gamma < 3$, there is a finite average degree, but a diverging variance \cite{Newman2005}.\\
+If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges; if $2 < \gamma < 3$, there is a finite average degree, but a diverging variance \cite{Newman2005}.\\
 \newline
@@
 They also often posses a modular/community structure \cite{Ferrara2012}\cite{McGlohon2011}.
+\\
 There are, however, some properties that are different between on-line and off-line networks as well.
@@
 The study of opinion dynamics in a group of people, as already mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
-The network layer that describes the underlying structure and correlations between the agents/people is intensively described in the previous section.
-In this section we will focus on the second important layer, namely that of the opinion dynamics models.
+The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
+In this chapter we will focus on the second important layer, namely that of the opinion dynamics models.
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
@@
 Some major attention will also be given to the content curation/filtering algorithms used by social media companies.
-This is because this thesis deals with opinion dynamics on social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.
+This is because this thesis deals with opinion dynamics on social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
+There are two major groups of opinion dynamics models, namely binary models and continuous approaches.
+These will be discussed in the following two sections.
+This thesis only deals with binary opinion dynamics models, so these will be thoroughly described; the continuous models will be quickly reviewed in order to give a more complete overview.
 % talk about opinion dynamics, different models, resistance, etc
+\section{Binary models}
+
+Binary models are models where each agent can have one of the two opinions A or B (or, equivalently, 0 or 1).
+A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates \cite{Nguyen2020}.
+
+\subsection{The majority model}
+
+One of the simplest opinion dynamics models is the binary majority model.
+In this model each agent has one of the two opinions A or B (or 0 or 1).
+Initially, these opinions can be equally distributed or not.
+If the distribution is not equal, but instead eg.
+80/20, there is a predominant opinion and a minority opinion.\\
+The following updating rule is then imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
+Thus, initially each agent has an opinion, then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
+Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity and does an opinion, that is, initially, a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
+This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016}.
+It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
+
+\section{Continuous models}
+
+\section{Algorithmic personalization}
+
+On-line social platforms use algorithms to order and filter the posts that appear on the time line of an individual \cite{Perra2019}.
+They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
+If social media companies didn't use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
+This in turn would make the on-line social experience a lot more unpleasant and inconvenient.
+The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that one can process \cite{Bozdag2013}.
+% search for literature to back this up! (Bozdag2013, read this!!)
+% talk about the algorithms proposed by perra and rocha
 

### 2021-02-15 draft.tex
continued writing opinion dynamics chapter + changed/added some small things here and there

@@
 The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies.
-These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc.
-The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.
+These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc \cite{Buchanan2007}.
+The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.\\
+\newline
+The statistical approach of social phenomena has one big difficulty compared to the description of gases and other physical phenomena.
+In regular physics, the microscopic constituents are `simple' atoms or molecules, whose dynamics and interactions with each other are often well known and can be put in relatively simple laws with only a few parameters \cite{Castellano2009}.
+Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}.
+Even if the dynamics/behavior of individuals was well know, than the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
+Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}.
+\\
+However, the use of statistical physics may also offer a way out.
+In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}.
+Instead, the relevant aspects for the global behavior are often higher level features such as symmetries, conservation laws, dimensionality, etc \cite{Castellano2009}.
+This introduces a concept of universality.
+With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
+An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
+Note that this is not only an important factor in social physics, but also in regular physical research and science in general.
 % talk about history
@@
 The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
-In this chapter we will focus on the second important layer, namely that of the opinion dynamics models.
+This chapter focuses on the second important layer, namely that of the opinion dynamics models.
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
 Hence, these opinion dynamics are the dynamics that run on the underlying network structure.
-Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.
-Some major attention will also be given to the content curation/filtering algorithms used by social media companies.
-This is because this thesis deals with opinion dynamics on social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
+Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.\\
+Some major attention will also be given to the content curation/filtering algorithms used by social media companies (see Section \ref{filter}).
+This is needed because this thesis deals with opinion dynamics on on-line social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
+\newline
+There will also be some attention on the activation mechanism (see Section \ref{actMech}).
+This is the mechanism that describes the timings in which content/information is exchanged between users \cite{Perra2019}.
+Such a mechanism may change the network from a static one to a more temporal one, where nodes become active and inactive over time.
+\\
+\newline
 There are two major groups of opinion dynamics models, namely binary models and continuous approaches.
-These will be discussed in the following two sections.
-This thesis only deals with binary opinion dynamics models, so these will be thoroughly described; the continuous models will be quickly reviewed in order to give a more complete overview.
-% talk about opinion dynamics, different models, resistance, etc
-\section{Binary models}
+These will be discussed in the following two sections, Section \ref{Bin} and Section \ref{Con}, respectively.
+This thesis only deals with binary opinion dynamics models, so these will be thoroughly described; the continuous models will be quickly reviewed in order to give a more complete overview.\\
+\newline
+Finally, in Section \ref{stubb}, the introduction of stubborn people is discussed and, in Section \ref{thesisOp}, a summary of the opinion dynamics model used in this thesis is given.
+
+\section{Binary models}\label{Bin}
+% Talk about Ising model???
 
@@
 Thus, initially each agent has an opinion, then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
-Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity and does an opinion, that is, initially, a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity and does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
-This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016}.
+This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan (also note that even when the opinions are equally distributed, there may still be a slight imbalance due to the random assignement of an opinion to each node; instead of a perfect 50/50 distribution, one may end up with a eg.
+49.8/50.2 distribution; in this sense there is still some kind of imbalance and a notion of `minority' and `majority' opinion)\cite{Benjamini2016}.
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
+% https://arxiv.org/pdf/0710.3256.pdf Contains a mean field mathimatical description, interesting?
 
-\section{Continuous models}
+\subsection{The probabilistic majority model}
 
-\section{Algorithmic personalization}
+The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature.
+In this model the fractions of neighbors of user $i$ with opinion A and B are determined.
+Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
+For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors has opinion B), user $i$ will updates his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
+Thus, user $i$ no longer updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
+This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}. 
+% https://arxiv.org/pdf/0710.3256.pdf read this (overview, discusses several models)!!
+
+\subsection{The Voter model}
+% https://arxiv.org/pdf/0710.3256.pdf
+
+\section{Continuous models}\label{Con}
+% https://arxiv.org/pdf/0710.3256.pdf, bounded confidence methods
+% https://www.researchgate.net/publication/235737446_Mixing_Beliefs_Among_Interacting_Agents defuant model
+% http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf
+% http://jasss.soc.surrey.ac.uk/9/3/8.html
+
+\section{Algorithmic personalization}\label{filter}
 
@@
 They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
-If social media companies didn't use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
-This in turn would make the on-line social experience a lot more unpleasant and inconvenient.
-The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that one can process \cite{Bozdag2013}.
-% search for literature to back this up! (Bozdag2013, read this!!)
-% talk about the algorithms proposed by perra and rocha
+If social media companies woudn't use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
+This in turn would make the on-line social experience a lot more unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
+The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that one can process \cite{Bozdag2013}.\\
+\newline
+The backside of such algorithmic filtering is that it fine-tunes the information it shows based on the user.
+Even two users with the same friends may then be exposed to different content, based on their previous activities and history \cite{Bozdag2013}.
+This may lead to the formation of echo chambers in which users get `trapped' in content that confirms their own mindset \cite{Bozdag2013}.
+\\
+It is thus important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
+Prof.
+Nicola Perra and prof.
+Luis E.C.
+Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{Modelling opinion dynamics in the age of algorithmic personalization} (see reference \cite{Perra2019}).
+Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}
+\begin{itemize}
+	\item Popularity filtering: promoting content that is popular across the platform
+	\item Semantic filtering: promoting posts that are similar to previous consumed posts
+	\item Collaborative filtering: promoting content that people similar or close to us consume
+\end{itemize} 
+The following paragraph briefly reviews the filtering algorithms that prof. Nicola Perra and prof. Luis E.C. Rocha proposed. \\
+\newline
+At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from its friends \cite{Perra2019}.
+The filtering algorithm then selects $Q$ items from this list and shows them on the user's time line; the following selection methods are used \cite{Perra2019}
+\begin{itemize}
+	\item The reference method (REF): this method randomly selects $Q$ items from the list $L_i(t)$
+	\item The old method (OLD): the $Q$ oldest post in $L_i(t)$ are displayed on the time line of the user
+	\item The recent method (REC): shows the $Q$ most recent items from $L_i(t)$
+	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the users and then selects the first $Q$ elements
+\end{itemize}
+Prof.
+Nicola Perra and prof.
+Luis E.C.
+Rocha showed that the REF, OLD and REC methods behave similar in many cases and that the PR method often displays different behavior (such as a larger formation of echo chambers compared to the other three methods) \cite{Perra2019}.
+In this thesis the REC method is mainly used as a reference/benchmark method and the PR method is used as the main filtering algorithm.
+
+\section{The activation mechanism}\label{actMech}
+
+\section{The presence of stubborn people}\label{stubb}
+
+\section{The opinion dynamics model used in this thesis}\label{thesisOp}
+
+% https://arxiv.org/pdf/0710.3256.pdf Talk about agent-based models??
 

### 2021-02-16 draft.tex
made some small adaptations

@@
 The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
+\newline
 One important feature of opinion dynamics on social media is the use of algorithmic personalization.
@@
 It is thus important to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
+\newline
 It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
 One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+\newline
 Opinion dynamics models have two important layers.
 On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.
+% refer to relevant sections in introduction?
 
@@
 %(maybe check cursus compl en crit?)
+\\
+\newline
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
@@
 The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way.
-However we do want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
+However, we do want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
 Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
@@
 Opinion dynamics models should try to capture these complex processes in a simplified way.
-Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.
+Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
 The basic idea of all opinion dynamic models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
@@
 \\
+\newline
 The opinion of the agents in the model can be either discrete or continuous.
@@
 \\
+\newline
 The rules that determine how an agent updates his or her opinion can include many different aspects.
@@
 the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating.
+A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}.
+\\
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others.
 Furthermore, the agents can have a parameter that determines whether they are active or not.
-\\
+The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
+\newline
 Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms.
@@
 Here popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
-In Chapter \ref{chap3} %also give section
-a detailed description of the filtering algorithms used in  this thesis will be given.
+In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
 % add reference to chapter 2
@@
 Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}.
-Even if the dynamics/behavior of individuals was well know, than the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
+Even if the dynamics/behavior of individuals would be well know, then the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
 Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}.
@@
 \newline
-As said earlier, these three main characteristics are the same for on- and off-line social networks.
+As said earlier, these three main characteristics are valid for on- and off-line social networks.
 Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}.
@@
 One of them might be the degree correlations/mixing pattern.
-It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern \cite{Hu2009}\cite{Zhang2014}.
+It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern (this distinguishes them from other real-world networks, eg.
+biological networks \cite{F.Costa2007}, which often have a disassortative pattern) \cite{Hu2009}\cite{Zhang2014}.
 For a long time, it was thought that this is also valid for on-line social networks.
@@
 \newline
-There are two major groups of opinion dynamics models, namely binary models and continuous approaches.
+There are two major groups of opinion dynamics models, namely binary/discrete models and continuous approaches.
 These will be discussed in the following two sections, Section \ref{Bin} and Section \ref{Con}, respectively.
@@
 Binary models are models where each agent can have one of the two opinions A or B (or, equivalently, 0 or 1).
-A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates \cite{Nguyen2020}.
+A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.
 
@@
 Thus, initially each agent has an opinion, then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
-Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity and does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+\newline
+Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity; and, does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+\\
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
 This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan (also note that even when the opinions are equally distributed, there may still be a slight imbalance due to the random assignement of an opinion to each node; instead of a perfect 50/50 distribution, one may end up with a eg.
-49.8/50.2 distribution; in this sense there is still some kind of imbalance and a notion of `minority' and `majority' opinion)\cite{Benjamini2016}.
+49.8/50.2 distribution; in this sense there is still some kind of notion of `minority' and `majority' opinion) \cite{Benjamini2016}.
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
@@
 For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors has opinion B), user $i$ will updates his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
-Thus, user $i$ no longer updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
+Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
 This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}. 
@@
 This in turn would make the on-line social experience a lot more unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
-The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that one can process \cite{Bozdag2013}.\\
+The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that human beings can process \cite{Bozdag2013}.\\
 \newline
@@
 \newline
-At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from its friends \cite{Perra2019}.
+At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
 The filtering algorithm then selects $Q$ items from this list and shows them on the user's time line; the following selection methods are used \cite{Perra2019}
@@
 	\item The recent method (REC): shows the $Q$ most recent items from $L_i(t)$
-	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the users and then selects the first $Q$ elements
+	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $Q$ elements
 \end{itemize}

### 2021-02-17 draft.tex
continued writing in chapter on opinion dynamics

@@
 \usepackage{titlesec}
+\usepackage{bbold}
 \usepackage{algorithm}
@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes \cite{Mata2020}.
+A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
@@
 The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
-This chapter focuses on the second important layer, namely that of the opinion dynamics models.
+The current chapter focuses on the second important layer, namely that of the opinion dynamics models.
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
@@
 Binary models are models where each agent can have one of the two opinions A or B (or, equivalently, 0 or 1).
-A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.
+A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
+In this section, some binary models will be discussed; note, however, that the models described here are far from all the models that currently exist.
 
@@
 The following updating rule is then imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
-Thus, initially each agent has an opinion, then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
-\newline
-Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive such a system to converge to unanimity; and, does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+Thus, initially each agent is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
+\newline
+Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive the system to converge to unanimity; and, does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
 \\
@@
 
-\subsection{The Voter model}
+\subsection{The voter model}
+
+Another extremely simple binary model, maybe even simpler than the majority model, is the voter model.
+The model works as follows: each agent in the network has a binary opinion (A or B); at each time step, an agent is randomly selected along with one of its neighbors and, subsequently, the agent adopts the opinion of this neighbor \cite{Castellano2009}.
+The agent thus mimics its neighbor \cite{Castellano2009}.
+\\
+The pressure of the majority in only felt in an averaged way, in the sense a neighbor that carries the majority opinion has more chance to be randomly selected \cite{Castellano2009}.\\
+Lots of studies have been performed on this model. In this thesis, however, we will not make use of the voter model. Hence, we will not go into further detail. 
 % https://arxiv.org/pdf/0710.3256.pdf
@@
 \section{Continuous models}\label{Con}
+
+In continuous opinion dynamics models, the opinion variable of the agents in the network is no longer discrete, in contrast to binary or discrete models, but takes real values (often in the interval $\interval{0}{1}$ or $\interval{-1}{1}$).\\
+In continuous opinion dynamics models, some concepts that are valid for the discrete models no longer apply (eg.
+the concept of a majority/minority opinion or the concept of equality of opinions); hence, they require a bit of a different framework \cite{Castellano2009}.
+\\
+The basis of most continuous models is a set of $N$ nodes or agents, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
+In contrast to discrete/binary models, all agents usually start with  different opinions \cite{Castellano2009}.
+The possible outcomes are more complex than in the discrete models and often has opinion clusters emerging in the final states (where opinion clusters refer to clusters of nodes with the same opinion) \cite{Castellano2009}.
+The number of opinion clusters can be one (consensus), two (polarization) or more (fragmentation) \cite{Castellano2009}.
+\\
+Several models have been proposed, we will only discuss a fraction of them.
+
+\subsection{The classical models}
+
+In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$) \cite{Krause2002}.
+The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
+\begin{equation}\label{refMod}
+	x(t+1) = W x(t)
+\end{equation}
+where $W$ is the fixed stochastic matrix, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
+Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}.
+The model they proposed assumes that an agent $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}.
+Thus this variation has the following updating rule (in matrix notation) \cite{Krause2002}
+\begin{equation}
+	x(t+1) = Gx(0) + (\mathbb{1}-G)Wx(t)
+\end{equation}
+where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix. If all the $g_i$ are equal to zero, we obtain the model of Eq. (\ref{refMod}) \cite{Krause2002}.\\
+A possible time variant extension is \cite{Krause2002}
+\begin{equation}
+	x(t+1) = W(t)x(t)
+\end{equation}
+where now the weights are time dependent.
+This model can be used for situations in which so called "hardening of positions" occur \cite{Krause2002}.
+This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}.
+% talk about extension http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf
+
+\subsection{Bounded confidence models}
+
+In bounded confidence models, each node has, beside an opinion, a threshold $\epsilon$, also called an uncertainty or tolerance (note that $\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
+This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are then ignored \cite{Deffuant2006}.
+In other words, an agent with opinion $x_i$ is only affected/influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
+The two most popular bounded confidence models are the Deffuant model and the Hegselmann-Krause model \cite{Castellano2009}.
+\\
+\newline
+\textbf{The Deffuant model}\\
+\newline
+Assume we have a population of $N$ agents, each with an opinion $x_i$ that takes real values in the interval $\interval{0}{1}$ \cite{Castellano2009}.
+Furthermore, the threshold $\epsilon$ is the same for all the agents.
+At each time step, an agent $i$ is randomly selected together with one of its neighbors $j$ (which is also randomly selected) \cite{Castellano2009}.
+If the opinions of the two nodes $i$ and $j$ are too far apart nothing happens; this is the case if $\abs{x_i(t) - x_j(t)} \geqslant \epsilon$.
+If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, then the opinions are updated to \cite{Castellano2009}
+\begin{equation}
+\begin{split}
+	x_i(t+1) &= x_i(t) + \mu [x_j(t) - x_i(t)]\\
+	x_j(t+1) &= x_j(t) + \mu [x_i(t) - x_j(t)]
+\end{split}
+\end{equation}
+The parameter $\mu$ is called the convergence parameter; its value lies in the interval $\interval{0}{1/2}$ \cite{Castellano2009}. \\
+The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two agents move closer to each other \cite{Castellano2009}. \\
+The final state is a situation where several opinion clusters coexist.
+This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (so the difference in opinions between all agents in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
+Only agents inside the same cluster keep on interacting with each other, which makes the opinion inside a cluster converge to a single value \cite{Castellano2009}.
+Usually, the number and sizes of the clusters depends in the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
+\newline
+\textbf{The Hegselmann-Krause model}\\
+\newline
+The Hegselmann-Krause model is very similar to the Deffuant model.
+Again, there are $N$ agents, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
+Each agent interacts with those neighbors with an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
+The difference with the Deffuant model lies in the updating rule \cite{Castellano2009}.
+In the Deffuant model, only the opinion of one randomly selected neighbor is taken into account, whereas the Hegselmann-Krause model considers all compatible neighbors \cite{Castellano2009}.
+The updating rule becomes \cite{Castellano2009}
+\begin{equation}
+	x_i(t+1) = \frac{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}x_j(t)}{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}}
+\end{equation}
+where $A_{ij}$ are the elements of the adjacency matrix. This model thus takes the average opinion of the compatible neighbors into account \cite{Castellano2009}.\\
+The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary states \cite{Castellano2009}.
+The number of clusters decreases if $\epsilon$ increased and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
+\newline
+There are, of course, many other variations of the bounded confidence model.
+As mentioned earlier, the threshold $\epsilon$ can be different for each node, so that each node has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
+\epsilon_r$ \cite{Krause2002}.
+% talk about relative agreement model?? http://jasss.soc.surrey.ac.uk/9/3/8.html and http://jasss.soc.surrey.ac.uk/5/4/1.html
+
 % https://arxiv.org/pdf/0710.3256.pdf, bounded confidence methods

### 2021-03-02 draft.tex
made some very small changes + comments

@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.
+A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
@@
 Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
-The basic idea of all opinion dynamic models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
+The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
 These models are obviously a simplification of real world opinion dynamics.
@@
 The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
-Here popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
+Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
 In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
@@
 An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
-Note that this is not only an important factor in social physics, but also in regular physical research and science in general.
-% talk about history
-
-% talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...)
+Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
+
+% talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...) --> partly done?
 
@@
 
-% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that
-% give reference to sections were community structure etc are explained!
+% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that NOT FINAL YET
+% give reference to sections were community structure etc are explained! (done?)
 
@@
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
-Hence, these opinion dynamics are the dynamics that run on the underlying network structure.
+Hence, these opinion dynamics are the dynamics that run on the underlying network structure.\\
 Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.\\
+\newline
 Some major attention will also be given to the content curation/filtering algorithms used by social media companies (see Section \ref{filter}).
@@
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
-This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan (also note that even when the opinions are equally distributed, there may still be a slight imbalance due to the random assignement of an opinion to each node; instead of a perfect 50/50 distribution, one may end up with a eg.
-49.8/50.2 distribution; in this sense there is still some kind of notion of `minority' and `majority' opinion) \cite{Benjamini2016}.
+This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016} (also note that even when the opinions are equally distributed, there may still be a slight imbalance due to the random assignement of an opinion to each node; instead of a perfect 50/50 distribution, one may end up with a eg.
+49.8/50.2 distribution; in this sense there is still some kind of notion of `minority' and `majority' opinion).
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
@@
 Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
-For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors has opinion B), user $i$ will updates his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
+For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will updates his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
 Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
@@
 \\
-The pressure of the majority in only felt in an averaged way, in the sense a neighbor that carries the majority opinion has more chance to be randomly selected \cite{Castellano2009}.\\
-Lots of studies have been performed on this model. In this thesis, however, we will not make use of the voter model. Hence, we will not go into further detail. 
+The pressure of the majority in only felt in an averaged way, in the sense that a neighbor that carries the majority opinion has more chance to be randomly selected \cite{Castellano2009}.\\
+Lots of studies have been performed on this model.
+In this thesis, however, we will not make use of the voter model.
+Hence, we will not go into further detail.
+% cite papers that have done studies on this model??
 % https://arxiv.org/pdf/0710.3256.pdf
@@
 
-In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$) \cite{Krause2002}.
+In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance/influence of the neighbors) \cite{Krause2002}.
 The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
@@
 \end{equation}
-where $W$ is the fixed stochastic matrix, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
+where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
 Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}.
@@
 where $A_{ij}$ are the elements of the adjacency matrix. This model thus takes the average opinion of the compatible neighbors into account \cite{Castellano2009}.\\
-The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary states \cite{Castellano2009}.
+The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary opinion clusters \cite{Castellano2009}.
 The number of clusters decreases if $\epsilon$ increased and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
@@
 Rocha showed that the REF, OLD and REC methods behave similar in many cases and that the PR method often displays different behavior (such as a larger formation of echo chambers compared to the other three methods) \cite{Perra2019}.
-In this thesis the REC method is mainly used as a reference/benchmark method and the PR method is used as the main filtering algorithm.
+In this thesis the REC and REF methods are mainly used as a reference/benchmark method and the PR method is used as the main filtering algorithm.
 
@@
 \section{The presence of stubborn people}\label{stubb}
+% maybe also mention here the model with the threshold!!
 

### 2021-03-02 draft.tex
wrote activation mechanism section, not finished

@@
 At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $Q$ items from this list and shows them on the user's time line; the following selection methods are used \cite{Perra2019}
+The filtering algorithm then selects $Q$ items from this list and shows them on the user's time line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
 \begin{itemize}
@@
 
+In the most simple opinion dynamics models, which makes use of a static network, users are always active and thus always exchanging opinions/posts with each other.
+In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented.
+\\
+It might, however, be a more realistic situation if nodes/users become active/inactive over time.
+This would be a better representation of real social media platforms, where user are active for some periods of time and then become inactive again.
+This kind of activation mechanism of the user gives the network a more temporal character.
+The underlying network structure doesn't change, but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't.\\
+\newline
+In this thesis we follow the same activation mechanism as was proposed in the paper of prof.
+Nicola Perra and prof.
+Luis E.C.
+Rocha (see reference \cite{Perra2019}).
+Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$).
+In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes.
+Prof.
+Nicola Perra and prof.
+Luis E.C.
+Rocha choose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
+They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2019}.
+However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}.
+\\
+Hence, for simplicity, this thesis implements the constant activation probability variant.
+It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
+
+% not finished, read this https://www.nature.com/articles/srep00469 and https://www.nature.com/articles/srep03006, might contain useful info
+
 \section{The presence of stubborn people}\label{stubb}

### 2021-03-06 draft.tex
wrote section activation mechanisms + added some commentary notes + added one reference

@@
 
-\section{The activation mechanism}\label{actMech}
-
-In the most simple opinion dynamics models, which makes use of a static network, users are always active and thus always exchanging opinions/posts with each other.
+\section{The activation mechanism: Static to temporal}\label{actMech}
+
+In the most simple opinion dynamics models, which make use of a static network, users are always active and thus always exchanging opinions/posts with each other.
 In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented.
+Another possibility is updating all users at the same time during each time step.
 \\
 It might, however, be a more realistic situation if nodes/users become active/inactive over time.
-This would be a better representation of real social media platforms, where user are active for some periods of time and then become inactive again.
-This kind of activation mechanism of the user gives the network a more temporal character.
-The underlying network structure doesn't change, but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't.\\
+This would be a better representation of real social media platforms, where users are active for some periods of time and then become inactive again.
+This kind of activation mechanism of the users gives the network a more temporal character.
+The underlying network structure doesn't change (in the sense that each user/node has a fixed number of neighbors that doesn't change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't.
+Hence, in this sense, this does affect the network structure in some way, because if a user, for example, doesn't become active for a long period of time, he/she doesn't participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
+This may affect the dynamics, since it may seem as if the network has a different average degree and different things like connectivity etc.
+\\
 \newline
@@
 Rocha choose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
-They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2019}.
+They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
 However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}.
@@
 It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
-
-% not finished, read this https://www.nature.com/articles/srep00469 and https://www.nature.com/articles/srep03006, might contain useful info
+\\
+If a user is active, he/she first spreads his current opinion to all his/her neighbors (for example by posting something).
+Then, the active user updates his own opinion according to the opinions on his/her time line \cite{Perra2019}.
+This updating goes according to the implemented opinion dynamics model (eg.
+majority model, probabilistic majority model,...).
+In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.\\
+It is important to note that what somebody sees on his/her time line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
+Every user has a "hidden time line $L_i(t)$" which contains the actual time line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
+At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this "hidden time line $L_i(t)$" and shows them on the actual time line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
+
+% NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where nodes can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, eg.
+if a friend has an opinion far from your, you might be more likely to defriend that one and, vice versa, you might be more likely to become new friends with somebody who has an opinion close to yours)
 
@@
 % https://arxiv.org/pdf/0710.3256.pdf Talk about agent-based models??
+
+% NOTE: maybe make somewhere in thesis link to real-world examples; for example the election system in USA: formation of echo chambers, different states voting for a different party
 

### 2021-03-12 draft.tex
wrote part about stubborness, not finished

@@
 
-\subsection{Bounded confidence models}
+\subsection{Bounded confidence models}\label{bounded}
 
@@
 \section{The presence of stubborn people}\label{stubb}
+
+It is well known that in real-life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others.
+They are convinced that they carry the truth with them and that the other people, with different world-views, are not capable of giving them any new and useful insights.
+They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs.
+Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is a tendency that people with extreme opinions are more persistent towards these opinions than people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\
+Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of such extreme and persistent groups a lot bigger.
+Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people.
+So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument).
+Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no constraints.
+Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge impacts.
+\\
+What is the impact of having stubborn/extreme people in your society?
+Does it matter if the stubborn people are grouped together or if they are distributed evenly throughout your system?
+Are there ways to reduce the possible impacts of stubbornness and/or extremism?
+Can these stubborn groups lead to polarization or even overtake the opinions in the system?
+To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.
+
+\subsection{Stubbornness in bounded confidence models}
+
+The bounded confidence models, discussed in Section \ref{bounded}, can be easily extended to included some notion of stubbornness.
+Guillaume Deffuant proposed such an extension where stubbornness is correlated to extremism \cite{Deffuant2006}.
+The model he proposed goes as follows.
+Remember that in the bounded confidence models, the agents have an opinion $x_i$ (which is a real number, often between $\interval{0}{1}$ or between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
+Guillaume Deffuant then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or the extreme opinion $1$ \cite{Deffuant2006}.
+These extremists are given a threshold equal to $\epsilon_e$ \cite{Deffuant2006}.
+The other agents are considered to be moderate and can have opinions between $-1$ and $1$.
+The moderates are given a threshold $\epsilon$, with $\epsilon > \epsilon_e$ \cite{Deffuant2006}.
+This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.
+% talk about results here??
+
+\subsection{Stubbornness in binary models}
+% not finished!!
+Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models.
+There are, however, many other ways to include some kind of stubborn people in these models.
+In the following, two different possibilities will be discussed.
+One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
+In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions.
+However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
+\newline
+\textbf{Including a stubbornness parameter}\\
+\newline
+One way of creating those stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
+This resistance parameters can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
+\begin{itemize}
+	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...)
+	\item If $r_i = 1$, the node is completely stubborn and there is no change of convincing him/here of the other idea
+	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion
+\end{itemize}
+The resistance parameter $r_i$ can be different for all nodes or can be the same.
+One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
+Many other possibilities can be found according to how the stubborn people are distributed and according to how stubborn they are made.
+This makes the model very flexible.\\
+\newline
+\textbf{The threshold model}
+\newline
+% NOT FINISHED!!
+
+Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study.
+
+% talk about epsilon parameter in deffuant model to implement stubbornness + talk about model thesis, binary so no real extreme/moderate distinction
+
 % maybe also mention here the model with the threshold!!

### 2021-03-13 draft.tex
continued writing stubbornness section

@@
 
-\subsection{The majority model}
+\subsection{The majority model}\label{majModel}
 
@@
 Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people.
-So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument).
+So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc).
 Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no constraints.
@@
 One way of creating those stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
-This resistance parameters can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
+This resistance parameter can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
 \begin{itemize}
@@
 	\item If $r_i = 1$, the node is completely stubborn and there is no change of convincing him/here of the other idea
-	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion
+	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so 
 \end{itemize}
-The resistance parameter $r_i$ can be different for all nodes or can be the same.
+The resistance parameter $r_i$ can be different or the same for all nodes.
 One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
@@
 \newline
-% NOT FINISHED!!
-
-Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study.
+In the threshold model one introduces a notion of stubbornness by introducing a threshold parameter $T$.
+This parameter can be either the same for all the nodes ($T$) or can be made individual ($T_i$).
+The main idea is then that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
+The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with another opinion they might be convinced of that opinion as well.
+The more persistent you are, the more friends you need in order to be convinced of another idea.
+\\
+There are different possibilities to implement this kind of model.
+One way, which makes use of the majority model, is as follows.
+First the fraction of friends with opinion A and opinion B, $p_A$ and $p_B$ respectively, is determined; then the following rules are used to update the opinion
+\begin{itemize}
+	\item If $p_A > p_B$ and $p_A > T$, opinion A is adopted
+	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted
+	\item In all other cases, the current opinion is not changed
+\end{itemize}
+If $T = 0$, this model reduces to the pure majority model described in Section \ref{majModel}.
+If $T = 1$, the nodes don't change their opinion.
+For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
+% Talk about probabilistic implementation?
+\newline
+Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study, many more exist or can be invented.
 

### 2021-03-13 draft.tex
made some small adjustements

@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
+A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
 One can find the language to describe networks in mathematical graph theory.
-However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
+However, complex systems in real-life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
 One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\
@@
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
-The other two are short distances and high clustering, these two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
+The other two are short distances and high clustering.
+These two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions.
-This might come across as a huge simplification of the real world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
+This might come across as a huge simplification of the real world complexity of opinion formation, but also in real-life people often have to choose between two competing opinions (eg.
 republicans or democrats, cat or dog, renting or buying a house,...).
@@
 % add reference to chapter 2
-%The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,...
-The structure of social networks plays an important role in aspects as peer pressure, whereas the algorithmic personalization used on social media platforms may interfere with the information someone is exposed to.
-Algorithmic personalization is the individualized way in which social media platforms show posts/information to each of its users.
 
@@
 With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
-An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
+An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real-life are compatible with the outputs generated by the models \cite{Castellano2009}.
 Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
@@
 \end{equation}
-If there are no correlations $k_{nn}$ is independent of $k$, and such a network is called a neutral network \cite{F.Costa2007}.
+If there are no degree correlations, $k_{nn}$ is independent of $k$.
+Such a network is called a neutral network \cite{F.Costa2007}.
 If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative.
@@
 
-Network communities is another feature that (complex) networks may posses.
+Network communities are another feature that (complex) networks may posses.
 A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}.
@@
 Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
-If we define the internal degree $k_i^{int}(C)$ as the number of edges of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges of node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
+If we define the internal degree $k_i^{int}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
 \begin{equation}
@@
 \end{equation}
-Since for large $N$, $\ln{N} \ll N$, and we find that the average distance is a lot smaller than the size of the network.
+Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
 Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.
@@
 Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$.
-The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected \cite{Watts1998}.\\
+The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
 The regular ring lattice ($p=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}

### 2021-03-16 draft.tex
made some more small adjustements

@@
 The past decade social media has become ever more important in our daily lives.
-Nowadays most of us have, beside a real world life, a virtual life on social media.
+Nowadays most of us have, beside a real-world life, a virtual life on social media.
 This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale.
+\\
 It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings \cite{Perra2019}.
@@
 \newline
+It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
+One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+\newline
 One important feature of opinion dynamics on social media is the use of algorithmic personalization.
@@
 People are, however, still bound by temporal and cognitive constraints.
-In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individuals time line according to what might be relevant to that individual \cite{Perra2019}.
-It is thus important to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
-\newline
-It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
-One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time line according to what might be relevant to that individual \cite{Perra2019}.
+It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
 \newline
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions, while on the other hand we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.
+On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.
 % refer to relevant sections in introduction?
@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
+A network, also often called a graph (both terms will be used in this thesis), is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
@@
 The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way.
-However, we do want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
+However, we still want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
 Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
@@
 These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks.
-It is found that many real-life social systems have a power law degree distribution \cite{Muchnik2013}.
+It is found that many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
@@
 The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
-These models are obviously a simplification of real world opinion dynamics.
+These models are obviously a simplification of real-world opinion dynamics.
 It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
@@
 This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions.
-This might come across as a huge simplification of the real world complexity of opinion formation, but also in real-life people often have to choose between two competing opinions (eg.
+This might come across as a huge simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
 republicans or democrats, cat or dog, renting or buying a house,...).
@@
 With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
-An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real-life are compatible with the outputs generated by the models \cite{Castellano2009}.
+An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
 Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
@@
 As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
-So on one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.
+On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.
 Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\
@@
 If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
-In the case of nudging, they found that the population opinion moved to the nudged opinion relatively fast, even in the case of small nudging.\\
+In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging.\\
 They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
@@
 
-% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that NOT FINAL YET
+% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that NOT FINAL YET --> maybe talk about fact that you play around with distributing the opinions over the different communities?
++ talk about investigating real networks??
 % give reference to sections were community structure etc are explained! (done?)
@@
 For a simple, undirected graph, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$).
-Directed graphs have asymmetric adjacency matrices.
+Directed graphs can have asymmetric adjacency matrices.
 If we are dealing with weighted graphs, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
@@
 The adjacency matrix contains a lot of information such as the degree of a node.
-The degree $k_i$ of a node $i$ is the number of edges attached to the node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of the node $i$.
-The degree can be obtained from the adjacency matrix in the following way \cite{Mata2020}
+The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$.
+The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
 \begin{equation}
@@
 \end{equation}
-In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\text{in}}$ and the number of outgoing edges $k_i^{\text{out}}$ of the node $i$.
+In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\text{in}}$ and the number of outgoing edges $k_i^{\text{out}}$ of node $i$.
 These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
@@
 a Poisson distribution.
-In this case, most of the nodes have a degree closed to the average degree $\left<k\right>$.
+In this case, most of the nodes have a degree close to the average degree $\left<k\right>$.
 Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
@@
 % correct statement?
++ talk about fact that av degree can diverge; explain scale free concept a bit more detailed
 %maybe explain why we care about degree, centrality measure etc
@@
 However, the terms assortativity and disassortativity are also often used without taking degree correlations into account.
-Instead, one may look at a particular node property to define assortative and disassortative network \cite{Thed2014}.
+Instead, one may look at a particular node property to classify networks as either assortative or disassortative \cite{Thed2014}.
 In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.
@@
 This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}.
-This assumes that the efficiency of sending information between two nodes is reciprocal to their distance \cite{F.Costa2007}.
-The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency
+This assumes that the efficiency of sending information between two nodes is reciprocal to their distance.
+The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency \cite{F.Costa2007}
 \begin{equation}
@@
 where $N$ denotes, once again, the number of nodes in the network.
-The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Iyer2013} \cite{Li2017}.
+The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Heer2020}\cite{Iyer2013} \cite{Li2017}.
 \\
@@
 It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
-It is thus obvious that the local clustering coefficient $C_i$ (Eq.
+It is then obvious that the local clustering coefficient $C_i$ (Eq.
 (\ref{clus})) of a node and the transitivity $T_i$ (Eq.
@@
 Several community definitions are consistent with the density hypothesis.
-Let us consider three possible definitions: maximum cliques, strong community and weak community \cite{Albert2016}.\\
+Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
 \newline
@@
 The modularity of a community $Q_c$ is then calculated by measuring the difference between the network real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
-This last quantity is given by $p_{ij}$ which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
+This last quantity is given by $p_{ij}$, which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
 Hence, the modularity of a community $Q_c$ becomes \cite{Albert2016}
@@
 \begin{itemize}
-	\item The higher the value of $Q$ is for a partition, the better is the corresponding community structure.
+	\item The higher the value of $Q$ is for a certain partition, the better is the corresponding community structure.
 This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure.
@@
 \newline
-The probability that a node is connected to $k$ other nodes and not to $N-k$ others is given by
-\begin{equation}
-	p^k (1-p)^{N-1-k}
-\end{equation}
-The number of ways to choose the $k$ nodes among the $N-1$ possible candidates is given by the binomial coefficient
-\begin{equation}
-	\binom{N-1}{k} = \frac{(N-1)!}{k!(N-1-k)!}
-\end{equation}
-and the probability of being connected to exactly $k$ other nodes becomes
+The degree distribution of the random network with $N$ nodes is given by
 \begin{equation}\label{degDistRan}
@@
 \end{equation}
-We thus derived the degree distribution of a random network. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
+The derivation can be found in Appendix \ref{ER}, Section \ref{degdis}. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
 \begin{equation}\label{avDeg}
@@
 This can be easily understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$.
-The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples.
+The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates).
 If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
@@
 \end{align}
-Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph, even for small $p$ \cite{Newman2000}.
+Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph (which is small), even for small $p$ \cite{Newman2000}.
 The clustering coefficient is approximately given by   \cite{Barrat1999}
@@
 	\item A partition of the node set $\{1, 2, ..., N\}$ into disjoint subsets $C_1, ..., C_k$; these groups are the communities
-	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$ represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
+	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
 \end{itemize}
@@
 Finally, edges are added between communities $C_i$ and $C_j$ with probability $p = M_{ij}$.\\
+\newline
 The stochastic block model can be fitted to data \cite{Clauset2017}.
 Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.\\
-One of the shortcomings of fitting the SBM to infer the community structure, is that real-world networks often have a heterogeneous or power law degree distribution.
+One of the shortcomings of fitting the SBM to infer the community structure, is that real-world networks often have a heterogeneous or power-law degree distribution.
 Since the SBM is made of random networks inside each group and random bipartite networks between each pair of groups, the degree distribution of the full network is always a mixture of Poisson distributions \cite{Clauset2017}.
@@
 However, a network with a skewed degree distribution violates this assumption of edge independence and the SBM seeks to explain this over other kinds of latent group structures that we might care about \cite{Clauset2017}.
-Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation \cite{Karrer2011}.
+Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
+We will not discuss their variation, since that would lead us to far astray, but interested readers can find more detailed information in reference \cite{Karrer2011}.
 % talk about community sizes from power law??
@@
 As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties.
-These are the two properties of the "small world phenomenon", namely a short average distance and a high clustering coefficient, and a heterogeneous or power law degree distribution \cite{RealWorld}.
+These are the two properties of the "small world phenomenon", namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
 These properties are valid for both off-line and on-line social networks \cite{Zhang2014}.
@@
 \newline
-\textbf{Heterogeneous or power law degree distribution}\\
-\newline
-Many real-world networks are found to have a power law degree distribution \cite{RealWorld}
+\textbf{Heterogeneous or power-law degree distribution}\\
+\newline
+Many real-world networks are found to have a power-law degree distribution \cite{RealWorld}
 \begin{equation}
@@
 where the exponent $\gamma$ often lies between 2 and 3.
-Networks with a power law degree distribution are also called heterogeneous or scale free networks.
+Networks with a power-law degree distribution are also called heterogeneous or scale free networks.
 The term scale free refers to the fact that there is no characteristic scale in the network.
@@
 This discrepancy between real-life social networks and on-line social networks may be explained as follows.
-In real-life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}.
+In real life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}.
 The `ordinary people' don't have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
@@
 
-It is well known that in real-life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others.
+It is well known that in real life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others.
 They are convinced that they carry the truth with them and that the other people, with different world-views, are not capable of giving them any new and useful insights.
@@
 \section{The Erd\H{o}s-R\'{e}nyi model}\label{ER}
+\subsection{Degree distribution}\label{degdis}
+The probability that a node is connected to $k$ other nodes and not to $N-1-k$ others is given by
+\begin{equation}
+	p^k (1-p)^{N-1-k}
+\end{equation}
+The number of ways to choose the $k$ nodes among the $N-1$ possible candidates is given by the binomial coefficient
+\begin{equation}
+	\binom{N-1}{k} = \frac{(N-1)!}{k!(N-1-k)!}
+\end{equation}
+and the probability of being connected to exactly $k$ other nodes becomes
+\begin{equation}\label{degDistRan}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
+\end{equation}
+We thus derived the degree distribution of a random network.
+
 \subsection{Average degree}\label{degree}

### 2021-03-17 draft.tex
made some small changes

@@
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
-We will not discuss their variation, since that would lead us to far astray, but interested readers can find more detailed information in reference \cite{Karrer2011}.
+We will not discuss their variation, since that would lead us too far astray, but interested readers can find more detailed information in reference \cite{Karrer2011}.
 % talk about community sizes from power law??
@@
 \\
-There are, however, some properties that are different between on-line and off-line networks as well.
+There are, however, some properties that are different between on-line and off-line social networks as well.
 One of them might be the degree correlations/mixing pattern.
@@
 There are two major groups of opinion dynamics models, namely binary/discrete models and continuous approaches.
-These will be discussed in the following two sections, Section \ref{Bin} and Section \ref{Con}, respectively.
+These will be discussed in Section \ref{Bin} and Section \ref{Con}, respectively.
 This thesis only deals with binary opinion dynamics models, so these will be thoroughly described; the continuous models will be quickly reviewed in order to give a more complete overview.\\
@@
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
+This can be explained by the fact that if the network is sparse enough, we may have disconnected parts; each of these disconnected groups may then evolve to a different consensus opinion, so that over the entire network different opinions coexist \cite{Nguyen2020}.
 % https://arxiv.org/pdf/0710.3256.pdf Contains a mean field mathimatical description, interesting?
@@
 Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
-For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will updates his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
+For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
 Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
@@
 
-\subsection{The classical models}
+\subsection{The classical models}\label{classMod}
 
@@
 In bounded confidence models, each node has, beside an opinion, a threshold $\epsilon$, also called an uncertainty or tolerance (note that $\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
-This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are then ignored \cite{Deffuant2006}.
+This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
 In other words, an agent with opinion $x_i$ is only affected/influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
@@
 This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (so the difference in opinions between all agents in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
-Only agents inside the same cluster keep on interacting with each other, which makes the opinion inside a cluster converge to a single value \cite{Castellano2009}.
+Only agents inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
 Usually, the number and sizes of the clusters depends in the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
@@
 This in turn would make the on-line social experience a lot more unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
-The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that human beings can process \cite{Bozdag2013}.\\
+The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that humans can process \cite{Bozdag2013}.\\
 \newline
@@
 	\item The recent method (REC): shows the $Q$ most recent items from $L_i(t)$
-	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $Q$ elements
+	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $Q$ elements.
+This filtering method is based on the semantic filtering explained above
 \end{itemize}
@@
 Hence, in this sense, this does affect the network structure in some way, because if a user, for example, doesn't become active for a long period of time, he/she doesn't participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
-This may affect the dynamics, since it may seem as if the network has a different average degree and different things like connectivity etc.
+This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity, clustering coefficient, etc may behave differently.
 \\
@@
 Luis E.C.
-Rocha choose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
+Rocha chose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
 They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is a tendency that people with extreme opinions are more persistent towards these opinions than people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\
+\newline
 Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of such extreme and persistent groups a lot bigger.
@@
 So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc).
-Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no constraints.
+Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
 Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge impacts.
 \\
+\newline
 What is the impact of having stubborn/extreme people in your society?
@@
 This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.
+\\
+\newline
+Another example of introducing some kind of persistence towards the own opinion of the nodes is the "hardening of positions" that was briefly discussed at the end of Section \ref{classMod}.
 % talk about results here??
@@
 	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...)
-	\item If $r_i = 1$, the node is completely stubborn and there is no change of convincing him/here of the other idea
+	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea
 	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so 
@@
 This makes the model very flexible.\\
-\newline
-\textbf{The threshold model}
+\newpage
+\noindent
+\textbf{The threshold model}\\
 \newline
@@
 This parameter can be either the same for all the nodes ($T$) or can be made individual ($T_i$).
-The main idea is then that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
-The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with another opinion they might be convinced of that opinion as well.
+The main idea is that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
+The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might be convinced of that opinion as well.
 The more persistent you are, the more friends you need in order to be convinced of another idea.
 \\
+\newline
 There are different possibilities to implement this kind of model.
 One way, which makes use of the majority model, is as follows.
-First the fraction of friends with opinion A and opinion B, $p_A$ and $p_B$ respectively, is determined; then the following rules are used to update the opinion
+First the fraction of friends with opinion A and opinion B, $p_A$ and $p_B$ respectively, are determined; then the following rules are used to update the opinion
 \begin{itemize}
@@
 For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
-% Talk about probabilistic implementation?
+% Talk about probabilistic implementation? + talk about fact that nothing happens for T < 0.5???
 \newline
@@
 \end{equation}
-where $c_i$ represents the community to which node $i$ belongs and where the Dirac-delta function $\delta_{c_i, c_j}$ expresses that only pairs of nodes that belong to the same community are taken into account.
+where $c_i$ represents the community to which node $i$ belongs and where the Dirac-delta function $\delta_{c_i, c_j}$ expresses the fact that only pairs of nodes that belong to the same community are taken into account.
 We can then rewrite the first term as a sum over communities \cite{Albert2016}
@@
 \subsection{Degree distribution}\label{degdis}
-The probability that a node is connected to $k$ other nodes and not to $N-1-k$ others is given by
+The probability that a node is connected to $k$ other nodes and not to the $N-1-k$ others is given by
 \begin{equation}
@@
 \end{equation}
-Differentiating both sides of this equation gives us \cite{Hopcroft2006}
+Differentiating both sides of this equation (with respect to $p$) gives us \cite{Hopcroft2006}
 \begin{equation}

### 2021-03-19 draft.tex
made some small changes
added a small section about the SBM-WS in the stochastic block model section

@@
 	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
-	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done in such a way that self-loops and edge duplication are avoided
+	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
+Rewiring is done in such a way that self-loops and edge duplication are avoided
 \end{enumerate}
-The probability $p$ allows us to tune the graph between a regular lattice ($p = 0$) and a random network ($p = 1$) \cite{Watts1998}.
+The probability $\beta$ allows us to tune the graph between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}.
 Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$.
 The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
-The regular ring lattice ($p=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
+The regular ring lattice ($\beta=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
 \begin{align}\label{C0}
@@
 \\
-The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
+The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq.
+(\ref{C}) and Eq.
+(\ref{d})) \cite{Watts1998}
 \begin{align}
@@
 \end{align}
-Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph (which is small), even for small $p$ \cite{Newman2000}.
+Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random graph (which is small), even for small $\beta$ \cite{Newman2000}.
 The clustering coefficient is approximately given by   \cite{Barrat1999}
 \begin{equation}
-	\bar{C} \approx \bar{C}(0)(1-p)^3
+	\bar{C} \approx \bar{C}(0)(1-\beta)^3
 \end{equation} 
@@
 \begin{itemize}
-	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-p)^3$
+	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$
 	\item The probability that edges are rewired back to each other is negligible
 \end{itemize}
-For rather small values of $p$, the clustering coefficient remains high.
+For rather small values of $\beta$, the clustering coefficient remains high.
 The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
@@
 We will not discuss their variation, since that would lead us too far astray, but interested readers can find more detailed information in reference \cite{Karrer2011}.
+\\
+\newline
+An extension that we are going to explore a bit deeper is the stochastic Watts-Strogatz block model (SBM-WS).\\
+\newline
+\textbf{The stochastic Watts-Strogatz block model (SBM-WS)}\\
+\newline
+As said before, the stochastic block model is a very flexible model and there are many different extensions.
+One extension that is used in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
+In this way, we obtain a version of the stochastic block model that has both a community structure and a high clustering coefficient.\\
+The construction of the model is very similar as for the regular stochastic block model.
+For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $K$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
+The result is a network that can be tuned to have both a community structure and a high clustering coefficient.
+This is different from the regular stochastic block model that can only account for a community structure and lacks a high clustering.
 % talk about community sizes from power law??
+% talk about extention with WS as constituents instead of ER
 % NOT FINISHED
-% different subsections where you talk about different graph models: erdos-renyi, clustered random, wats strogatz, barabasi albert, etc...
 
@@
 and the probability of being connected to exactly $k$ other nodes becomes
-\begin{equation}\label{degDistRan}
+\begin{equation}
 	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}

### 2021-03-22 draft.tex
worked on lay-out + added some more chapters

@@
 \documentclass[11 pt , letterpaper , twoside , openright]{book}
+\renewcommand{\baselinestretch}{1.15} 
 \usepackage[utf8]{inputenc}
 
-\usepackage[english]{babel}     	
+\usepackage[english, dutch]{babel} 
+\usepackage{fancyhdr}
+%\pagestyle{empty}
+\newenvironment{abstract}%
+{\cleardoublepage\null \vfill\begin{center}\bfseries \abstractname \end{center}}{\vfill\null}
+
 \usepackage{graphicx}                   % om afbeeldingen in te kunnen laden   
@@
 \usepackage[noend]{algpseudocode}
-\usepackage{tikz-feynman}
 \usepackage{subcaption}
 \usepackage{xfrac}
-\tikzfeynmanset{compat=1.1.0}
 \floatname{algorithm}{Algoritme}
+
+\addto\captionsenglish{\renewcommand{\contentsname}{Table of Contents}
+}
 
@@
 \frontmatter
+
+\selectlanguage{english}
+\thispagestyle{empty}
+\begin{abstract}
+The past decade, the rise of social media has become undeniable.
+More and more people are getting engaged on some type of social media platform and hence, it is probably not wrong to say that nowadays most of us have, beside our real life, a virtual life on social media.
+This rapidly increased engagement on social media undoubtedly has many advantages.
+However, every upside has a downside and social media is no different.
+This thesis deals with the impact of social media on the formation and evolution of opinions in society.
+Social media allows for the spread of information at a faster pace and at an unprecedented scale.
+However, people are still bounded by time- and cognitive constraints.
+In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to.
+% definetely not finished!! --> this will be put in summary! (think about how you want lay-out: chapter of abstract style?)
+\end{abstract}
+
+\chapter*{Acknowledgments} 
+\addcontentsline{toc}{chapter}{Acknowledgments}
+
+\chapter*{Summary}
+\addcontentsline{toc}{chapter}{Summary}
+
+\selectlanguage{dutch}
+\chapter*{Nederlandstalige samenvatting}
+\addcontentsline{toc}{chapter}{Nederlandstalige samenvatting}
+
+\selectlanguage{english}
 \tableofcontents
+\addcontentsline{toc}{chapter}{Table of Contents}
 %\listoffigures
@@
 % also talk about algorithmic personalisation!!
+% ADD MORE REFERENCES (previous research, contextualization etc)
 
@@
 \newline
-In one of their previous papers (see reference \cite{Perra2019}) prof.
+In one of their previous papers (\textit{"Modelling opinion dynamics in the age of algorithmic personalization"}, see reference \cite{Perra2019}) prof.
 Nicola Perra and prof.
@@
 We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
-We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can observe convergence to one opinion inside the communities.
+We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can, for example, observe convergence to one opinion inside the communities.
+We will also pay some attention to networks that have both a high clustering coefficient and a community structure.
+Can we observe echo chambers in those kinds of networks?
+If yes, can we determine which property might be the main driver for the formation of echo chambers: high clustering or community structure?
+\\
+% ALSO mention real-world network??
+\newline
 On the other hand, we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network.
+Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness?
+What about the possible interplay between the filtering algorithms and the introduction of stubborn people?
+Does the distribution of stubborn people play a role (eg.
+randomly distributed versus clustered together inside communities)?
+\\
 
@@
 The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
+\newpage
+\noindent
 Most real networks are sparse, which means that $\left<k\right> \ll N$. In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014} 
@@
 The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. \\
-\newline
-\newline
 \newline
@@
 \newline
-\newline
 \textbf{Short average distance}\\
@@
 \newline
-Finally, in Section \ref{stubb}, the introduction of stubborn people is discussed and, in Section \ref{thesisOp}, a summary of the opinion dynamics model used in this thesis is given.
+Finally, in Section \ref{stubb}, the introduction of stubborn people is discussed.
 
@@
 Luis E.C.
-Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{Modelling opinion dynamics in the age of algorithmic personalization} (see reference \cite{Perra2019}).
+Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
 Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}
@@
 However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
-\newline
+\newpage
+\noindent
 \textbf{Including a stubbornness parameter}\\
@@
 This makes the model very flexible.\\
-\newpage
-\noindent
+\newline
 \textbf{The threshold model}\\
@@
 
-\section{The opinion dynamics model used in this thesis}\label{thesisOp}
+%\section{The opinion dynamics model used in this thesis}\label{thesisOp}
 
@@
 % NOTE: maybe make somewhere in thesis link to real-world examples; for example the election system in USA: formation of echo chambers, different states voting for a different party
+\chapter{Materials and methods in this thesis}
+
+The previous two chapters, Chapter \ref{chap2} and \ref{chap3}, gave a broad introduction/overview of networks and opinion dynamics models respectively.
+In this chapter we will summarize the specific networks and opinion dynamics models that are used in this thesis.
+\\
+Section \ref{networkmodel} gives an overview of the network models that are used in this thesis.
+In chapter \ref{chap2}, Section \ref{netModel} we already discussed their main properties; here we will mainly discuss why we use these networks.
+Which of their properties are of interest in this thesis?
+What are there benefits and possible shortcomings regarding to the goals of this thesis?\\
+In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.\\ % + explain where they come from etc?
+Finally, Section \ref{opinionmodel} discusses the opinion dynamics model used in this thesis. Again we will give some focus on possible advantages and disadvantages of the model.
+
+\section{Network models}\label{networkmodel}
+\section{Real-world networks}\label{realnetwork}
+\section{Opinion dynamics models}\label{opinionmodel}
+% use pseudo code here??
+
+\chapter{Results}
+
+\chapter{Discussion}
+
+\chapter{Conclusions and outlook}
 

### 2021-03-23 draft.tex
small change in stubborn section

@@
 
+% DON'T FORGET: popular text somewhere!!
+
 \mainmatter
@@
 	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea
-	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so 
+	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
+A uniform random number $f$ between 0 and 1 is drawn.
+If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion
 \end{itemize}

### 2021-03-24 draft.tex
started writing abstract

@@
 
+\chapter*{Acknowledgments} 
+\addcontentsline{toc}{chapter}{Acknowledgments}
+
 \selectlanguage{english}
-\thispagestyle{empty}
 \begin{abstract}
+\addcontentsline{toc}{chapter}{Abstract}
+The past decade, the rise of social media has become undeniable.
+More and more people are getting engaged on some type of social media platform and hence, it is probably not wrong to say that nowadays most of us have, beside our real life, a virtual life on social media.
+This rapidly increased engagement on social media undoubtedly has many advantages.
+However, every upside has a downside and social media is no different.
+This thesis deals with the impact of social media on the formation and evolution of opinions in society.\\
+Social media allows for the spread of information at a faster pace and at an unprecedented scale.
+However, people are still bounded by time- and cognitive constraints.
+In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to.
+This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in "echo chambers" that confirm their own beliefs.
+It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
+People are, however, more than sheep that thoughtlessly follow their shepherd (contrary to what some books might lead us to believe).
+They are capable of critical thinking and may be hesitant to change their current opinion.
+This aspect of potential stubbornness should not be ignored.\\
+In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
+We will investigate which network properties (eg.
+community structure, clustering coefficient, etc) may enhance or hamper the formation of echo chambers.
+We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness.
+Beside that, we will also pay some attention to the possible ways of distributing the opinions; we will compare randomly distributed cases to cases where entire communities are given one single opinion.
+% definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
+\end{abstract}
+
+\selectlanguage{dutch}
+\begin{abstract}
+\addcontentsline{toc}{chapter}{Samenvatting}
 The past decade, the rise of social media has become undeniable.
@@
 \end{abstract}
-
-\chapter*{Acknowledgments} 
-\addcontentsline{toc}{chapter}{Acknowledgments}
-
-\chapter*{Summary}
-\addcontentsline{toc}{chapter}{Summary}
-
-\selectlanguage{dutch}
-\chapter*{Nederlandstalige samenvatting}
-\addcontentsline{toc}{chapter}{Nederlandstalige samenvatting}
 

### 2021-03-24 draft.tex
worked on lay-out

@@
-\documentclass[11 pt , letterpaper , twoside , openright]{book}
+\documentclass[11 pt , letterpaper , oneside , openright]{book}
 \renewcommand{\baselinestretch}{1.15} 
@@
 \usepackage{fancyhdr}
-%\pagestyle{empty}
+
+\setlength{\headheight}{13.6pt}
+
 \newenvironment{abstract}%
@@
 \usepackage{graphicx}                   % om afbeeldingen in te kunnen laden
-\usepackage{tikz,tikz-3dplot}                       % om afbeeldingen te kunnen maken
-\usetikzlibrary{calc}                   % meer opties in tikz (oa coordinaten bepalen)
 
@@
 
+\pagestyle{plain}
+
 \chapter*{Acknowledgments} 
@@
 \addcontentsline{toc}{chapter}{Samenvatting}
-The past decade, the rise of social media has become undeniable.
-More and more people are getting engaged on some type of social media platform and hence, it is probably not wrong to say that nowadays most of us have, beside our real life, a virtual life on social media.
-This rapidly increased engagement on social media undoubtedly has many advantages.
-However, every upside has a downside and social media is no different.
-This thesis deals with the impact of social media on the formation and evolution of opinions in society.
-Social media allows for the spread of information at a faster pace and at an unprecedented scale.
-However, people are still bounded by time- and cognitive constraints.
-In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to.
 % definetely not finished!! --> this will be put in summary! (think about how you want lay-out: chapter of abstract style?)
@@
 \mainmatter
+
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\chaptername \ \thechapter}
+\rhead{\rightmark}
+\cfoot{\thepage}
+
 \chapter{Introduction}
@@
 
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\appendixname \ \thechapter}
+\rhead{\rightmark}
+\cfoot{\thepage}
+
 \appendix
+\addappheadtotoc
+
 \chapter{Network measurements: Derivations}
@@
 \backmatter
+
+\pagestyle{plain}
 % bibliography

### 2021-03-25 draft.tex
continued working on lay-out

@@
 \usepackage[utf8]{inputenc}
+
+\usepackage[dvipsnames]{xcolor}
 
@@
 \usepackage{interval}
-\usepackage[colorlinks,citecolor=red,urlcolor=blue,bookmarks=false,hypertexnames=true]{hyperref} 
+\usepackage[colorlinks,citecolor=NavyBlue,linkcolor=NavyBlue% 
+,urlcolor=NavyBlue,linktocpage=true,bookmarks=false,hypertexnames=true]{hyperref} 
 \usepackage[noend]{algpseudocode}
@@
 
+\usepackage{sectsty}
+
+\chapternumberfont{\color{NavyBlue}}
+\chaptertitlefont{\color{black}\sectionrule{3ex}{3pt}{-1ex}{1pt}}
+
 \begin{document}
+
 \frontmatter
+\selectlanguage{english}
+
+\begin{titlepage}
+\begin{center}
+
+\vspace*{.06\textheight}
+{\scshape\LARGE \textcolor{NavyBlue}{Ghent University}\par}\vspace{1.5cm} % University name
+\textsc{\Large Master Thesis}\\[0.5cm] % Thesis type
+
+\hrule % Horizontal line
+\vspace{0.4cm} 
+{\huge \bfseries Opinion dynamics on social media with stubborn actors\par}\vspace{0.4cm} % Thesis title
+\hrule % Horizontal line
+\vspace{1.5cm}
+
+\begin{minipage}[t]{0.4\textwidth}
+\begin{flushleft} \large
+\emph{Author:}\\
+\textcolor{NavyBlue}{Nina \textsc{Botte}} % Author name - remove the \href bracket to remove the link
+\end{flushleft}
+\end{minipage}
+\begin{minipage}[t]{0.4\textwidth}
+\begin{flushright} \large
+\emph{Supervisor:} \\
+\textcolor{NavyBlue}{Prof. Dr. Luis E.C. \textsc{Rocha}} % Supervisor name - remove the \href bracket to remove the link  
+\end{flushright}
+\end{minipage}\\[3cm]
+
+\vfill
+
+\large \textit{A thesis submitted in partial fulfillment of the requirements\\ for the degree of Master of Science in Physics and Astronomy}\\[0.3cm] % University requirement text
+\textit{in the}\\[0.4cm]
+\textcolor{NavyBlue}{Faculty of \textsc{Science}}\\\textcolor{NavyBlue}{Department of \textsc{Physics and Astronomy}}\\[2cm] % Research group name and department name
+
+\vfill
+
+{\large \today}\\[4cm] % Date
+%\includegraphics[scale=0.7]{ugent_logo.png} % University/department logo - uncomment to place it
+
+\vfill
+\end{center}
+\end{titlepage}
 
@@
 \fancyhf{}
-\lhead{\chaptername \ \thechapter}
+\lhead{\textcolor{NavyBlue}{\chaptername} \ \textcolor{NavyBlue}{\thechapter}}
 \rhead{\rightmark}
@@
 A network, also often called a graph (both terms will be used in this thesis), is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
+\newline
 Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
@@
 (\ref{mod}) become equal and $Q$ vanishes.
+	\newline
 	\item If every node is put into a separate community, the first term in Eq. (\ref{mod}) vanishes and we are left with $n_c$ negative terms. Hence, this situation corresponds to a negative modularity.
@@
 \fancyhf{}
-\lhead{\appendixname \ \thechapter}
+\lhead{\textcolor{NavyBlue}{\appendixname} \ \textcolor{NavyBlue}{\thechapter}}
 \rhead{\rightmark}

### 2021-03-25 draft.tex
wrote part on network models in chapter 5, NOT FINISHED, still need to include shortcommings of the models

@@
 In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
-We will investigate which network properties (eg.
-community structure, clustering coefficient, etc) may enhance or hamper the formation of echo chambers.
+We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers.
 We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness.
@@
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.
-% refer to relevant sections in introduction?
-
-\section{Complex networks}
+On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\ Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks.
+The second layer, the opinion dynamics layer, is introduced in Section \ref{Opinion}.
+Section \ref{statPhys} introduces concepts such as statistical and social physics.
+Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are discussed.
+
+\section{Complex networks}\label{complNet}
 
@@
 
-\section{Opinion dynamics}
+\section{Opinion dynamics}\label{Opinion}
 
@@
 
-\section{Statistical and social physics}
+\section{Statistical and social physics}\label{statPhys}
 
@@
 
-\section{This thesis: hypotheses and goals}
+\section{This thesis: hypotheses and goals}\label{goal}
 
@@
 If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
-In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging.\\
+In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.\\
 They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
@@
 In this way, we obtain a version of the stochastic block model that has both a community structure and a high clustering coefficient.\\
-The construction of the model is very similar as for the regular stochastic block model.
+The construction of the model is very similar to that of the regular stochastic block model.
 For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $K$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
@@
 In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.\\ % + explain where they come from etc?
-Finally, Section \ref{opinionmodel} discusses the opinion dynamics model used in this thesis. Again we will give some focus on possible advantages and disadvantages of the model.
+Finally, Section \ref{opinionmodel} discusses the opinion dynamics model that is implemented in this thesis. Again we will give some focus on possible advantages and disadvantages of the model.
 
 \section{Network models}\label{networkmodel}
+
+This thesis uses three main network models, namely
+\begin{itemize}
+	\item The Erd\H{o}s-R\'{e}nyi model
+	\item The Watts-Strogatz model
+	\item The stochastic block model, of which two types are implemented
+		\begin{itemize}
+			\item[\tiny$\bullet$] The regular stochastic block model (SBM)
+			\item[\tiny$\bullet$] The stochastic Watts-Strogatz block model (SBM-WS)
+		\end{itemize}
+\end{itemize}
+Each of these networks serve specific goals which will shortly be discussed in the following paragraphs.\\
+\newline
+\textbf{The Erd\H{o}s-R\'{e}nyi model}\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model is mainly used as a benchmark model.
+In this thesis we focus intensively on two network features that may impact the formation of echo chambers, namely a high clustering coefficient and a community structure.
+It is therefore necessary to have a model that lacks both of these features to compare the results with.
+The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that doesn't display both of these features and is hence the perfect comparison model.
+\\
+It is expected that the random nature of the model and the lack of a high clustering and a community structure will result in the absence of echo chambers.\\
+The Erd\H{o}s-R\'{e}nyi model is also used in the regular stochastic block model to construct the constituent communities.\\
+% talk about expected behaviour?? 
+\newline
+\textbf{The Watts-Strogatz model}\\
+\newline
+The Watts-Strogatz model is used in this thesis as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
+In the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} Prof.
+Perra and Prof.
+Rocha already showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
+Here, we will use the model to investigate the interplay between high clustering and the presence of stubborn people and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+We will investigate whether a community structure behaves similar to a high clustering regarding the appearance of echo chambers.\\
+Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
+\newline
+\textbf{The stochastic block model}\\
+\newline
+One of the main features of the stochastic block model is the ability to generate networks with a community structure.
+We will therefore use this model to investigate the impact of community structure/modularity on the evolution of opinions and the formation of echo chambers.\\
+It is expected that a community structure will enhance the formation of echo chambers compared to the random network case.
+The question remains, however, whether the influence of a community structure on the formation of echo chambers is of the same caliber as that of a high clustering.\\
+The stochastic Watts-Strogatz block model is used as a model that has both a high clustering and a community structure.
+We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
+\newline
+All the above models have a homogeneous degree distribution.
+In this thesis, we didn't focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
+It might, however, be an interesting extension to investigate such types of models as well.
+% NOT FINISHED: talk about shortcommings --> models, neither of them describes real networks completely accurate, etc
+
 \section{Real-world networks}\label{realnetwork}

### 2021-03-26 draft.tex
made some small adjustements in SBM section

@@
 %\includegraphics[scale=0.7]{ugent_logo.png} % University/department logo - uncomment to place it
-
+% https://www.overleaf.com/project/605b4c033a439c5f0f15c7e5 --> copywright?
 \vfill
@@
 
-Another random graph model is one that produces networks containing a community structure, namely the stochastic block model (SBM).
+The stochastic block model (SBM) is another random graph model that is able to produce networks containing a community structure.
 The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
@@
 Hence, these networks posses a modular/community structure and this is exactly the structure that is easily modeled by the stochastic block model.
-The stochastic block model is also often used as a benchmark for community detection algorithms.\\
+The stochastic block model is also often used as a benchmark for community detection algorithms \cite{Abbe2018}.\\
 \newline
@@
 	\item The number of nodes $N$ 
-	\item A partition of the node set $\{1, 2, ..., N\}$ into disjoint subsets $C_1, ..., C_k$; these groups are the communities
+	\item A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities
 	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
 \end{itemize}
-There are no real restrictions to the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures.
+There are no real restrictions to the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures \cite{Karrer2011}.
 However, the edge probabilities inside a community are often taken to be bigger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
@@
 This can be seen in the two different ways.
-First, consider the case where $k = 1$, thus we only have one group.
+First, consider the case where $k = 1$, thus we only have one community and all nodes belong to this community.
 The matrix $M$ then reduces to one number $p$ and we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model \cite{Clauset2017}.
-Second, assume we have more than one group or community, $k > 1$, and assume that every element of the matrix $M$ is the same, $M_{ij} = p$ \cite{Clauset2017}.
+Second, assume we have more than one group or community, $k > 1$, and assume that every element of the matrix $M$ is the same, $M_{ij} = p$ and $M_{ii}=p$ \cite{Clauset2017}.
 Again, we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model.
@@
 Since the SBM is made of random networks inside each group and random bipartite networks between each pair of groups, the degree distribution of the full network is always a mixture of Poisson distributions \cite{Clauset2017}.
-The consequence of this is that if a network possesses a more skewed degree distribution, the SBM will put the nodes with a high degree in a small group, together with a high probability of connecting to other, larger groups \cite{Clauset2017}.
+The consequence of this is that if a network possesses a more skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
 The reason behind this is one of the main assumptions of the stochastic block model, namely the fact that the edge probabilities only depend  on which group the node belongs to \cite{Clauset2017}.
 In other words, nodes that belong to the same group are stochastically  equivalent, meaning that they have equivalent connectivity patterns to other nodes \cite{Clauset2017}.
-However, a network with a skewed degree distribution violates this assumption of edge independence and the SBM seeks to explain this over other kinds of latent group structures that we might care about \cite{Clauset2017}.
+However, a network with a skewed degree distribution violates this assumption of edge independence \cite{Clauset2017}.
+To say it in the words of Karrer and Newman: "Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...]." (Karrer and Newman, p.1, reference \cite{Karrer2011}).
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.

### 2021-03-27 draft.tex
made some small adjustments and continued writing in chapter 4

@@
 People are, however, still bound by temporal and cognitive constraints.
-In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time line according to what might be relevant to that individual \cite{Perra2019}.
+In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}.
 It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
@@
 It is found that many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
-This heterogeneous or scale free degree distribution represents one of the three general properties of social networks.
+This heterogeneous or scale-free degree distribution represents one of the three general properties of social networks.
 The other two are short distances and high clustering.
@@
 % explain differences  more, consequences of power-law + need to include figure?
-They are also referred to as scale free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
 % correct statement?
@@
 \end{itemize}
-There are no real restrictions to the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures \cite{Karrer2011}.
+There are no real restrictions on the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures \cite{Karrer2011}.
 However, the edge probabilities inside a community are often taken to be bigger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
@@
 The matrix $M$ then reduces to one number $p$ and we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model \cite{Clauset2017}.
-Second, assume we have more than one group or community, $k > 1$, and assume that every element of the matrix $M$ is the same, $M_{ij} = p$ and $M_{ii}=p$ \cite{Clauset2017}.
+Second, assume we have more than one group or community, $k > 1$, and assume that every element of the matrix $M$ is the same, $M_{ij} = p$, $\forall \ i, j$ \cite{Clauset2017}.
 Again, we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model.
@@
 In this way, we obtain a version of the stochastic block model that has both a community structure and a high clustering coefficient.\\
+\newline
 The construction of the model is very similar to that of the regular stochastic block model.
@@
 % talk about community sizes from power law??
-% talk about extention with WS as constituents instead of ER
+
 % NOT FINISHED
@@
 where the exponent $\gamma$ often lies between 2 and 3.
-Networks with a power-law degree distribution are also called heterogeneous or scale free networks.
-The term scale free refers to the fact that there is no characteristic scale in the network.
+Networks with a power-law degree distribution are also called heterogeneous or scale-free networks.
+The term scale-free refers to the fact that there is no characteristic scale in the network.
 Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$, the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
@@
 The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature.
-In this model the fractions of neighbors of user $i$ with opinion A and B are determined.
+In this model the fraction of neighbors of user $i$ with opinion A and the fraction of neighbors with opinion B are determined.
 Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
@@
 Hence, we will not go into further detail.
-% cite papers that have done studies on this model??
-% https://arxiv.org/pdf/0710.3256.pdf
+Interested readers are referred to references \cite{Castellano2009}, \cite{Fernandez2014} and related references therein.% cite papers that have done studies on this model??
 
@@
 The model they proposed assumes that an agent $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}.
-Thus this variation has the following updating rule (in matrix notation) \cite{Krause2002}
+This variation thus has the following updating rule (in matrix notation) \cite{Krause2002}
 \begin{equation}
@@
 
-In bounded confidence models, each node has, beside an opinion, a threshold $\epsilon$, also called an uncertainty or tolerance (note that $\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
+In bounded confidence models, each node has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance (note that $\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
 This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
 In other words, an agent with opinion $x_i$ is only affected/influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
-The two most popular bounded confidence models are the Deffuant model and the Hegselmann-Krause model \cite{Castellano2009}.
-\\
+\\
+The two most popular bounded confidence models are the Deffuant model and the Hegselmann-Krause model \cite{Castellano2009}. \\
 \newline
@@
 Only agents inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
-Usually, the number and sizes of the clusters depends in the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
+Usually, the number and sizes of the clusters depends on the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
 \newline
@@
 
-On-line social platforms use algorithms to order and filter the posts that appear on the time line of an individual \cite{Perra2019}.
+On-line social platforms use algorithms to order and filter the posts that appear on the time-line of an individual \cite{Perra2019}.
 They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
@@
 At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $Q$ items from this list and shows them on the user's time line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
+The filtering algorithm then selects $Q$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
 \begin{itemize}
 	\item The reference method (REF): this method randomly selects $Q$ items from the list $L_i(t)$
-	\item The old method (OLD): the $Q$ oldest post in $L_i(t)$ are displayed on the time line of the user
+	\item The old method (OLD): the $Q$ oldest post in $L_i(t)$ are displayed on the time-line of the user
 	\item The recent method (REC): shows the $Q$ most recent items from $L_i(t)$
@@
 \\
+\newline
 If a user is active, he/she first spreads his current opinion to all his/her neighbors (for example by posting something).
-Then, the active user updates his own opinion according to the opinions on his/her time line \cite{Perra2019}.
+Then, the active user updates his own opinion according to the opinions on his/her time-line \cite{Perra2019}.
 This updating goes according to the implemented opinion dynamics model (eg.
@@
 In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.\\
-It is important to note that what somebody sees on his/her time line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
-Every user has a "hidden time line $L_i(t)$" which contains the actual time line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
-At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this "hidden time line $L_i(t)$" and shows them on the actual time line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
+It is important to note that what somebody sees on his/her time-line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
+Every user has a "hidden time-line $L_i(t)$" which contains the actual time-line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
+At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this "hidden time-line $L_i(t)$" and shows them on the actual time-line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
 
@@
 
+\newpage
 \section{The presence of stubborn people}\label{stubb}
@@
 It is well known that in real life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others.
-They are convinced that they carry the truth with them and that the other people, with different world-views, are not capable of giving them any new and useful insights.
+They are convinced that they carry the truth with them and that other people, with different world-views, are not capable of giving them any new and useful insights.
 They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs.
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
 \newline
@@
 Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people.
-So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc).
+So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a wide reach and a significant impact on society).
 Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
-Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge impacts.
+Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge, and largely unknown, impacts.
 \\
@@
 Can these stubborn groups lead to polarization or even overtake the opinions in the system?
+Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism?
 To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.
@@
 The moderates are given a threshold $\epsilon$, with $\epsilon > \epsilon_e$ \cite{Deffuant2006}.
-This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.
-\\
-\newline
+This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
+Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in reference \cite{Deffuant2006}. \\
+\newpage
+\noindent
 Another example of introducing some kind of persistence towards the own opinion of the nodes is the "hardening of positions" that was briefly discussed at the end of Section \ref{classMod}.
-% talk about results here??
 
@@
 In the following, two different possibilities will be discussed.
+\\
 One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
 In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions.
-However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
-\newpage
-\noindent
+However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.
+Whether this resistance is a sign of critical thinking or ignorance/being convinced that you know it best, is up to the reader to decide.
+The models simply introduce some type of resistance and make no distinction between possible underlying reasons for this resistance.\\
+\newline
 \textbf{Including a stubbornness parameter}\\
 \newline
-One way of creating those stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
+One way of creating stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
 This resistance parameter can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
@@
 	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea
+	\newline
 	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
@@
 The main idea is that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
-The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might be convinced of that opinion as well.
+The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
 The more persistent you are, the more friends you need in order to be convinced of another idea.
@@
 \newline
-There are different possibilities to implement this kind of model.
+There are different possibilities of implementing this kind of model.
 One way, which makes use of the majority model, is as follows.
@@
 	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted
-	\item In all other cases, the current opinion is not changed
+	\item In all other cases, the current opinion of the node is not changed
 \end{itemize}
@@
 \begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model
-	\item The Watts-Strogatz model
+	\item The Erd\H{o}s-R\'{e}nyi model (ER)
+	\item The Watts-Strogatz model (WS)
 	\item The stochastic block model, of which two types are implemented
@@
 The Erd\H{o}s-R\'{e}nyi model is mainly used as a benchmark model.
-In this thesis we focus intensively on two network features that may impact the formation of echo chambers, namely a high clustering coefficient and a community structure.
+In this thesis we focus intensively on two network features/topologies that may impact the formation of echo chambers, namely a high clustering coefficient and a community structure.
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
@@
 The Watts-Strogatz model is used in this thesis as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-In the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} Prof.
+In the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}) Prof.
 Perra and Prof.
 Rocha already showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
-Here, we will use the model to investigate the interplay between high clustering and the presence of stubborn people and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn people and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
 We will investigate whether a community structure behaves similar to a high clustering regarding the appearance of echo chambers.\\
@@
 It is expected that a community structure will enhance the formation of echo chambers compared to the random network case.
-The question remains, however, whether the influence of a community structure on the formation of echo chambers is of the same caliber as that of a high clustering.\\
+The question remains, however, whether the influence of a community structure on the formation of echo chambers is of the same caliber as that of a high clustering.
+It is also of interest to investigate the possible relation between the appearance of echo chambers and the modularity of the network.\\
 The stochastic Watts-Strogatz block model is used as a model that has both a high clustering and a community structure.
@@
 In this thesis, we didn't focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
-It might, however, be an interesting extension to investigate such types of models as well.
+It might, however, be an interesting extension to investigate such types of models as well.\\
+\newline
+The above models obviously have their added value and benefits.
+One of their most important plus points is their simplicity.
+They are the most simple network models that one can create and each of them have characteristics and topologies that can be easily tuned (clustering for WS, community structure for SBM).
+This makes it really easy to investigate the effect of different network topologies (clustering, community structure) without the need of taking the interplay with other possible complex network features and topologies into account.\\
+This plus side is, however, also one of the possible downsides.
+Real networks aren't simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
+In other words, network models may lack many features of real-world networks.
+It is thus always a bit tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world networks.\\
+The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.\\
+The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis.
+But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world.
+Results that are obtained from them should not be thoughtlessly extrapolated to real-world networks.
 % NOT FINISHED: talk about shortcommings --> models, neither of them describes real networks completely accurate, etc

### 2021-03-31 draft.tex
changed some comments
implemented lay-out feedback from Yolan

@@
-\documentclass[11 pt , letterpaper , oneside , openright]{book}
+\documentclass[11 pt , letterpaper , twoside , openright]{book}
 \renewcommand{\baselinestretch}{1.15} 
@@
 
+\usepackage{geometry}
+\geometry{
+ a4paper,
+ left=30mm,
+ bottom=25mm,
+ right = 20mm
+}
+
+\usepackage{emptypage}
+
 \begin{document}
@@
 
-{\large \today}\\[4cm] % Date
+{\large Academic year 2020-2021}\\[4cm] % Date
 %\includegraphics[scale=0.7]{ugent_logo.png} % University/department logo - uncomment to place it
@@
 \pagestyle{plain}
-
-\chapter*{Acknowledgments} 
+\renewcommand{\abstractname}{Acknowledgments}
+\begin{abstract}
 \addcontentsline{toc}{chapter}{Acknowledgments}
 
+\end{abstract}
+
+\renewcommand{\abstractname}{Abstract}
 \selectlanguage{english}
@@
 \addcontentsline{toc}{chapter}{Samenvatting}
-% definetely not finished!! --> this will be put in summary! (think about how you want lay-out: chapter of abstract style?)
+% definetely not finished!!
 \end{abstract}
@@
 % add research question and hypotheses!!
-% also talk about algorithmic personalisation!!
 % ADD MORE REFERENCES (previous research, contextualization etc)
@@
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\ Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks.
+On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\
+\newline
+Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks.
 The second layer, the opinion dynamics layer, is introduced in Section \ref{Opinion}.
@@
 These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
-%references to chapter 2 need to be adapted
 % to do: read links 19, 21, 22, 23, 24 of https://arxiv.org/pdf/0711.3199.pdf
-
-% maybe also talk about degree distributions (poisson, power law), power law behavior of many real systems, clustering + maybe talk about networks that will be used in this thesis?
 
@@
 In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
-% add reference to chapter 2
-
-% not finished, describe different models
-% talk about updating rules used in thesis + concepts as resistance etc. 
+
 % biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9
@@
 \end{equation}
-In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\text{in}}$ and the number of outgoing edges $k_i^{\text{out}}$ of node $i$.
+In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$.
 These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
@@
 
+\newpage
 \subsection{Network distances}
@@
 % finished?
-
+\newpage
 \subsection{Clustering coefficient and transitivity}
@@
 \begin{equation}\label{avClus}
-	\bar{C} = \frac{1}{N}\sum_{i = 1}^N C_i
+	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i
 \end{equation}
@@
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
-In the case of the average clustering coefficient $\bar{C}$ and the global transitivity $T$ this is, however, not the case.
+In the case of the average clustering coefficient $\left<C\right>$ and the global transitivity $T$ this is, however, not the case.
 The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node.
@@
 Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
-If we define the internal degree $k_i^{int}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
-\begin{equation}
-	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C
+If we define the internal degree $k_i^{\textrm{int}}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{\textrm{ext}}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
+\begin{equation}
+	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \qquad \forall i \in C
 \end{equation}
@@
 \begin{equation}
-	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C)
+	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C)
 \end{equation}
@@
 (\ref{mod}) become equal and $Q$ vanishes.
-	\newline
 	\item If every node is put into a separate community, the first term in Eq. (\ref{mod}) vanishes and we are left with $n_c$ negative terms. Hence, this situation corresponds to a negative modularity.
@@
 % explain small-world (milgram), high clustering, power-law
-
+\newpage
 \section{Network models}
@@
 The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
-\newpage
-\noindent
+\newline
 Most real networks are sparse, which means that $\left<k\right> \ll N$. In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014} 
@@
 \begin{equation}\label{C}
-	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}
+	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}
 \end{equation}
@@
 The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates).
-If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
+If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<C\right> \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
 \\
@@
 The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
-The regular ring lattice ($\beta=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
+The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<C(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
 \begin{align}\label{C0}
-	\bar{C}(0) = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
+	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
 \end{align}
@@
 \\
-The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq.
+The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<C(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
 (\ref{C}) and Eq.
@@
 \begin{align}
-	\bar{C}(1) \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
+	\left<C(1)\right> \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
 \end{align}
@@
 \begin{equation}
-	\bar{C} \approx \bar{C}(0)(1-\beta)^3
-\end{equation} 
-where $\bar{C}(0)$ is given by Equation (\ref{C0}). This can be seen in the following way
+	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3
+\end{equation} 
+where $\left<C(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
 \begin{itemize}
@@
 An extension that we are going to explore a bit deeper is the stochastic Watts-Strogatz block model (SBM-WS).\\
-\newline
+\newpage
+\noindent
 \textbf{The stochastic Watts-Strogatz block model (SBM-WS)}\\
@@
 Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
+\newpage
+\noindent
 For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
@@
 Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in reference \cite{Deffuant2006}. \\
-\newpage
-\noindent
+\newline
 Another example of introducing some kind of persistence towards the own opinion of the nodes is the "hardening of positions" that was briefly discussed at the end of Section \ref{classMod}.
@@
 	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea
-	\newline
 	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
@@
 
-% talk about epsilon parameter in deffuant model to implement stubbornness + talk about model thesis, binary so no real extreme/moderate distinction
-
-% maybe also mention here the model with the threshold!!
-
 %\section{The opinion dynamics model used in this thesis}\label{thesisOp}
@@
 Results that are obtained from them should not be thoughtlessly extrapolated to real-world networks.
-% NOT FINISHED: talk about shortcommings --> models, neither of them describes real networks completely accurate, etc
 

### 2021-03-31 draft.tex
implemented lay-out feedback from Yolan, part 2

@@
 \usepackage{listings}                   % voor mooie code
-\usepackage{amsmath, amssymb, amsfonts} % om wiskunde te typen
+\usepackage[fleqn]{amsmath} 
+\usepackage{amssymb, amsfonts} % om wiskunde te typen
 \usepackage{scalerel}                   % om symbolen te schalen
@@
 It is a measure for the number of triangles in a network and determines the connectivity in the neighborhood of a node $i$: if a node $i$ has a high clustering coefficient, its neighbors are likely to be directly connected to each other \cite{Li2017}.
-\\Here a triangle is defined as a loop of length three, this is a sequence of nodes $x, y, z, x$ such that $\{x, y\}, \{y, z\}$ and $\{z, x\}$ are edges of the network.
+\\
+\newline
+A triangle is defined as a loop of length three, this is a sequence of nodes $x, y, z, x$ such that $\{x, y\}, \{y, z\}$ and $\{z, x\}$ are edges of the network.
 The clustering coefficient is thus a way to measure the degree to which nodes in a network tend to cluster \cite{Li2017}.
@@
 \end{equation}
+\newline
 A connected triple is defined as a set of three nodes with at least two edges between them, so that each node can be reached from the other two (either directly or indirectly).
 The factor three arises from the fact that each triangle contributes to three different connected triples in the network: one centered at each node in the triangle \cite{F.Costa2007}.
-Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$.
-These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
+\\
+\newline
+Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
 \begin{align}
@@
 This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
-\newline
+\newpage
+\noindent
 \textbf{Strong communities}\\
@@
 \end{itemize}
-The number of nodes at a distance $d$ cannot exceed the number of nodes $N$ in the network. We thus obtain the following equation from which the maximum distance or diameter $d_{max}$ can be obtained
-\begin{equation}
-	N \approx \left<k\right>^{d_{max}}
+The number of nodes at a distance $d$ cannot exceed the number of nodes $N$ in the network.
+We thus obtain the following equation from which the maximum distance or diameter $d_{\textrm{max}}$ can be obtained
+\begin{equation}
+	N \approx \left<k\right>^{d_{\textrm{max}}}
 \end{equation}
@@
 \begin{equation}\label{diameter}
-	d_{max} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
 \end{equation}
@@
 Equation (\ref{diameter}) might, however, be a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
-This is because $d_{max}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
+This is because $d_{\textrm{max}}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
 Hence, the average distance of a random network is defined by \cite{Albert2014}
@@
 \begin{equation}
-	N(d_{max}) \approx N
+	N(d_{\textrm{max}}) \approx N
 \end{equation}
@@
 \begin{equation}
-	N \approx \left<k\right>^{d_{max}}
+	N \approx \left<k\right>^{d_{\textrm{max}}}
 \end{equation}
@@
 \begin{equation}
-	d_{max} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
 \end{equation}
@@
 \begin{equation}
-	\bar{C} = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)}
+	\left<C\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)}
 \end{equation}

### 2021-03-31 draft.tex
implemented lay-out feedback from Yolan, part 3

@@
 \begin{equation}
-	E \subseteq \{\{x, y\}| x, y \in V \}
+	E \subseteq \{\{x, y\}| x, y \in V \}\ .
 \end{equation}
@@
 \begin{equation}
-	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \}
+	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \} \ .
 \end{equation}
@@
 	\begin{cases}
-		1 & \text{if nodes $i$ and $j$ are connected}\\
-		0 & \text{otherwise}
+		1 & \text{if nodes $i$ and $j$ are connected},\\
+		0 & \text{otherwise}.
 	\end{cases}
@@
 	\begin{cases}
-		w_{ij} & \text{if nodes $i$ and $j$ are connected}\\
-		0 & \text{otherwise}
+		w_{ij} & \text{if nodes $i$ and $j$ are connected},\\
+		0 & \text{otherwise},
 	\end{cases}
@@
 \begin{equation}
-	k_i = \sum_{j=1}^N A_{ij}
+	k_i = \sum_{j=1}^N A_{ij}\ .
 \end{equation}
@@
 These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
-\begin{equation}
-	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}
-\end{equation}
+\begin{align}
+	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \ , && k_i^{\text{out}} = \sum_{j=1}^N A_{ij} \ .
+\end{align}
 The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$.
@@
 \begin{equation}
-	s_i = \sum_{j=1}^N w_{ij}
+	s_i = \sum_{j=1}^N w_{ij} \ .
 \end{equation}
@@
 \begin{equation}
-	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k)
+	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k) \ ,
 \end{equation}
@@
 \begin{equation}
-	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)}
+	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)} \ .
 \end{equation}
@@
 \begin{equation}
-	k_{nn}(k) = \sum_{k'} k'P(k'|k)
+	k_{nn}(k) = \sum_{k'} k'P(k'|k) \ .
 \end{equation}
@@
 \newline
-\newline
 \textbf{Average distance}\\
@@
 \begin{equation}\label{avdist}
-	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
+	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ .
 \end{equation}
@@
 \begin{equation}
-	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}
+	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}} \ .
 \end{equation}
@@
 \begin{equation}
- 	h = \frac{1}{E}
+ 	h = \frac{1}{E} \ .
 \end{equation}
@@
 \begin{equation}\label{clus}
-	C_i = \frac{2n_i}{k_i(k_i-1)}
+	C_i = \frac{2n_i}{k_i(k_i-1)} \ ,
 \end{equation}
@@
 \begin{equation}\label{avClus}
-	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i
+	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i \ ,
 \end{equation}
@@
 \begin{equation}\label{globalTrans}
-	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}
+	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}} \ .
 \end{equation}
@@
 \begin{align}
-	N_{\Delta} &= \sum_{k > j > i} A_{ij}A_{ik}A_{jk} \\
-	N_3 &= \sum_{k > j > i} (A_{ij}A_{ik} + A_{ji}A_{jk} + A_{ki}A_{kj})
+	N_{\Delta} &= \sum_{k > j > i} A_{ij}A_{ik}A_{jk} \ , \\
+	N_3 &= \sum_{k > j > i} (A_{ij}A_{ik} + A_{ji}A_{jk} + A_{ki}A_{kj}) \ ,
 \end{align}
@@
 \begin{equation}\label{trans}
-	T_i = \frac{N_{\Delta}(i)}{N_3(i)}
+	T_i = \frac{N_{\Delta}(i)}{N_3(i)} \ ,
 \end{equation}
@@
 \begin{align}
-	N_{\Delta}(i) &= \sum_{k > j} A_{ij}A_{ik}A_{jk} \\
-	N_3(i) &= \sum_{k > j} A_{ij}A_{ik}
+	N_{\Delta}(i) &= \sum_{k > j} A_{ij}A_{ik}A_{jk} \ , \\
+	N_3(i) &= \sum_{k > j} A_{ij}A_{ik} \ .
 \end{align}
@@
 \begin{equation}
-	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \qquad \forall i \in C
+	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \ , \qquad \forall i \in C \ .
 \end{equation}
@@
 \begin{equation}
-	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C)
+	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C) \ .
 \end{equation}
@@
 \begin{equation}
-	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij})
+	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij}) \ ,
 \end{equation}
@@
 \begin{equation}
-	p_{ij} = \frac{k_ik_j}{2L}
+	p_{ij} = \frac{k_ik_j}{2L} \ ,
 \end{equation}
@@
 \begin{equation}\label{Qc}
-	Q_c = \frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2
+	Q_c = \frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \ ,
 \end{equation}
@@
 \begin{equation}\label{mod}
-	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]
+	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg] \ .
 \end{equation}
@@
 \begin{equation}
-	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2}
+	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2} \ ,
 \end{equation}
@@
 \begin{equation}\label{reslim}
-	k \leqslant \sqrt{2L}
+	k \leqslant \sqrt{2L} \ ,
 \end{equation}
@@
 \begin{equation}\label{degDistRan}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
 \end{equation}
@@
 	\left<k\right> &= \sum_{k=0}^{N-1} k\  p_k  \\
-&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}
+&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
 \end{split}
@@
 \begin{equation}
-	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!}
+	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!} \ .
 \end{equation}
@@
 \begin{equation}\label{C}
-	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}
+	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1} \ .
 \end{equation}
@@
 \begin{equation}
-	N \approx \left<k\right>^{d_{\textrm{max}}}
+	N \approx \left<k\right>^{d_{\textrm{max}}} \ ,
 \end{equation}
@@
 \begin{equation}\label{diameter}
-	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}} \ .
 \end{equation}
@@
 \begin{equation}\label{d}
-	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}} \ .
 \end{equation}
@@
 \begin{align}\label{C0}
-	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
+	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
 \end{align}
@@
 \begin{align}
-	\left<C(1)\right> \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
+	\left<C(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
 \end{align}
@@
 \begin{equation}
-	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3
+	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3 \ ,
 \end{equation} 
@@
 \begin{equation}
-	P(k) \sim k^{-\gamma}
+	P(k) \sim k^{-\gamma} \ ,
 \end{equation}
@@
 \begin{equation}\label{refMod}
-	x(t+1) = W x(t)
+	x(t+1) = W x(t) \ ,
 \end{equation}
@@
 \begin{equation}
-	x(t+1) = Gx(0) + (\mathbb{1}-G)Wx(t)
+	x(t+1) = Gx(0) + (\mathbb{1}-G)Wx(t) \ ,
 \end{equation}
@@
 \begin{equation}
-	x(t+1) = W(t)x(t)
+	x(t+1) = W(t)x(t) \ ,
 \end{equation}
@@
 \begin{split}
-	x_i(t+1) &= x_i(t) + \mu [x_j(t) - x_i(t)]\\
-	x_j(t+1) &= x_j(t) + \mu [x_i(t) - x_j(t)]
+	x_i(t+1) &= x_i(t) + \mu [x_j(t) - x_i(t)] \ ,\\
+	x_j(t+1) &= x_j(t) + \mu [x_i(t) - x_j(t)] \ .
 \end{split}
@@
 \begin{equation}
-	x_i(t+1) = \frac{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}x_j(t)}{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}}
+	x_i(t+1) = \frac{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}x_j(t)}{\sum_{j:\abs{x_i(t) - x_j(t)}<\epsilon}A_{ij}} \ ,
 \end{equation}
@@
 \begin{equation}\label{modnetw}
-	Q = \frac{1}{2L}\sum_{i, j = 1}^N \bigg(A_{ij} - \frac{k_ik_j}{2L}\bigg) \delta_{c_i, c_j}
+	Q = \frac{1}{2L}\sum_{i, j = 1}^N \bigg(A_{ij} - \frac{k_ik_j}{2L}\bigg) \delta_{c_i, c_j} \ ,
 \end{equation}
@@
 \begin{equation}\label{1}
-	\frac{1}{2L}\sum_{i, j = 1}^N A_{ij}\delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{2L}\sum_{i, j \in C_c} A_{ij} = \sum_{c=1}^{n_c}\frac{L_c}{L}
+	\frac{1}{2L}\sum_{i, j = 1}^N A_{ij}\delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{2L}\sum_{i, j \in C_c} A_{ij} = \sum_{c=1}^{n_c}\frac{L_c}{L} \ ,
 \end{equation}
@@
 \begin{equation}\label{2}
-	\frac{1}{2L}\sum_{i, j = 1}^N \frac{k_ik_j}{2L} \delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{(2L)^2}\sum_{i, j \in C_c} k_ik_j = \sum_{c=1}^{n_c}\frac{k_c^2}{4L^2}
+	\frac{1}{2L}\sum_{i, j = 1}^N \frac{k_ik_j}{2L} \delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{(2L)^2}\sum_{i, j \in C_c} k_ik_j = \sum_{c=1}^{n_c}\frac{k_c^2}{4L^2} \ ,
 \end{equation}
@@
 \begin{equation}\label{Qfin}
-	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]
+	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg] \ .
 \end{equation}
@@
 \begin{equation}\label{Q_AB}
-	\Delta Q_{AB} = \bigg[\frac{L_{AB}}{L} - \bigg(\frac{k_{AB}}{2L} \bigg)^2 \bigg] - \bigg[\frac{L_A}{L} - \bigg(\frac{k_A}{2L} \bigg)^2 + \frac{L_B}{2L} - \bigg(\frac{k_B}{2L} \bigg)^2 \bigg]
+	\Delta Q_{AB} = \bigg[\frac{L_{AB}}{L} - \bigg(\frac{k_{AB}}{2L} \bigg)^2 \bigg] - \bigg[\frac{L_A}{L} - \bigg(\frac{k_A}{2L} \bigg)^2 + \frac{L_B}{2L} - \bigg(\frac{k_B}{2L} \bigg)^2 \bigg] \ ,
 \end{equation}
@@
 \begin{equation}
-	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2}
+	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2} \ .
 \end{equation}
@@
 \begin{equation}
-	p^k (1-p)^{N-1-k}
+	p^k (1-p)^{N-1-k} \ .
 \end{equation}
@@
 \begin{equation}
-	\binom{N-1}{k} = \frac{(N-1)!}{k!(N-1-k)!}
+	\binom{N-1}{k} = \frac{(N-1)!}{k!(N-1-k)!} \ ,
 \end{equation}
@@
 \begin{equation}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
 \end{equation}
@@
 \begin{equation} 
-	(p+q)^n = \sum_{k=0}^n \binom{n}{k} p^k q^{n-k}
+	(p+q)^n = \sum_{k=0}^n \binom{n}{k} p^k q^{n-k} \ .
 \end{equation}
@@
 	n(p+q)^{n-1} &= \sum_{k=0}^n \binom{n}{k} k p^{k-1} q^{n-k} \\
-	&= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} q^{n-k}
+	&= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} q^{n-k} \ .
 \end{split}
@@
 	n (p+(1-p))^{n-1} &= \frac{1}{p} \sum_{k=0}^n \binom{n}{k} k p^{k} (1-p)^{n-k} \\
-	np &= \sum_{k=0}^n \binom{n}{k} k p^{k} (1-p)^{n-k}
+	np &= \sum_{k=0}^n \binom{n}{k} k p^{k} (1-p)^{n-k} \ ,
 \end{split}
@@
 \begin{equation}\label{pk}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}
+	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ ,
 \end{equation}
@@
 \begin{equation}\label{bin}
-	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!}
+	\binom{N-1}{k} = \frac{(N-1)(N-1-1)...(N-1-k+1)}{k!} \approx \frac{(N-1)^k}{k!} \ .
 \end{equation}
@@
 \begin{equation}
-	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg)
+	\ln[(1-p)^{N-1-k}] = (N-1-k)\ln\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \ ,
 \end{equation}
@@
 \begin{equation}
-	\ln(1+x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ...  \qquad \forall \ |x| \leqslant 1
+	\ln(1+x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}x^n = x - \frac{x^2}{2} + \frac{x^3}{3} - ... \ ,  \qquad \forall \ |x| \leqslant 1 \ ,
 \end{equation}
@@
 \begin{equation}
-	\ln[(1-p)^{N-1-k}] \approx (N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right>
+	\ln[(1-p)^{N-1-k}] \approx (N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right> \ ,
 \end{equation}
@@
 \begin{equation}\label{e}
-	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>}
+	(1-p)^{N-1-k} = \mathrm{e}^{-\left<k\right>} \ .
 \end{equation}
@@
 &= \frac{(N-1)^k}{k!} \bigg(\frac{\left<k\right>}{N-1} \bigg)^k \mathrm{e}^{-\left<k\right>}\\
-&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!}
+&= \mathrm{e}^{-\left<k\right>}\frac{\left<k\right>^k}{k!} \ .
 \end{split}
@@
 \begin{equation}\label{dist}
-	N(d) = 1 + \left<k\right> + \left<k\right>^2 + ... + \left<k\right>^d = \frac{\left<k\right>^{d+1} - 1}{\left<k\right> - 1}
+	N(d) = 1 + \left<k\right> + \left<k\right>^2 + ... + \left<k\right>^d = \frac{\left<k\right>^{d+1} - 1}{\left<k\right> - 1} \ .
 \end{equation}
@@
 \begin{equation}
-	N(d_{\textrm{max}}) \approx N
+	N(d_{\textrm{max}}) \approx N \ .
 \end{equation}
@@
 \begin{equation}
-	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}}
+	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}} \ .
 \end{equation}
@@
 \begin{equation}\label{triangle}
-	\binom{\frac{K}{2}}{2} = \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)
+	\binom{\frac{K}{2}}{2} = \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg) \ .
 \end{equation}
@@
 \begin{equation}\label{triple}
-	\binom{K}{2} = \frac{1}{2} K (K-1)
+	\binom{K}{2} = \frac{1}{2} K (K-1) \ .
 \end{equation}
@@
 \begin{equation}
-	\left<C\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)}
+	\left<C\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)} \ .
 \end{equation}
@@
 \begin{equation}
-	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
+	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ ,
 \end{equation}
@@
 \begin{equation}
-	l_i = \frac{1}{N-1} K \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg)
+	l_i = \frac{1}{N-1} K \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ ,
 \end{equation}
@@
 \begin{equation}
-	l = \frac{1}{N} \sum_i l_i = \frac{1}{N} N l_i \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg)
+	l = \frac{1}{N} \sum_i l_i = \frac{1}{N} N l_i \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ .
 \end{equation}
@@
 \begin{equation}
-	1 + 2+ 3 +...+ x = \sum_{i = 1}^x i = \frac{x(x+1)}{2} \approx \frac{x^2}{2}
+	1 + 2+ 3 +...+ x = \sum_{i = 1}^x i = \frac{x(x+1)}{2} \approx \frac{x^2}{2} \ ,
 \end{equation}
@@
 \begin{equation}
-	l \approx \frac{K}{N} \frac{\frac{N^2}{K^2}}{2} = \frac{N}{2K}
+	l \approx \frac{K}{N} \frac{\frac{N^2}{K^2}}{2} = \frac{N}{2K} \ .
 \end{equation}

### 2021-04-02 draft.tex
inserted test figure

@@
 \usepackage[noend]{algpseudocode}
+\usepackage[labelfont=bf]{caption}
 \usepackage{subcaption}
@@
 This aspect of potential stubbornness should not be ignored.\\
-In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
+In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.\\
 We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers.
 We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness.
-Beside that, we will also pay some attention to the possible ways of distributing the opinions; we will compare randomly distributed cases to cases where entire communities are given one single opinion.
+\\
+Beside that, we will also pay some attention to the possible ways of distributing the opinions and the stubborn people and how this may effect the opinion dynamics in the system. \\
+Finally, we will test the model on real-world networks.
 % definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
@@
 Which of their properties are of interest in this thesis?
-What are there benefits and possible shortcomings regarding to the goals of this thesis?\\
+What are there benefits and possible shortcomings regarding to our goals?\\
 In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.\\ % + explain where they come from etc?
@@
 In other words, network models may lack many features of real-world networks.
-It is thus always a bit tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world networks.\\
+It is thus always a bit tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
 The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.\\
+\newline
 The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis.
@@
 \chapter{Results}
+
+TEST FIGURE!!
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=14cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_WS_fracStubb_PR.png}}
+\caption{This is an image from a text that uses color to teach music. Make really really long caption, as long as textwidth}
+\captionsetup{width=\textwidth, labelfont=bf}
+\label{fig}
+\end{figure}
+\noindent
+The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis.
+But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world.
+Results that are obtained from them should not be thoughtlessly extrapolated to real-world networks.
 

### 2021-04-03 draft.tex
small update

@@
 Finally, we will test the model on real-world networks.
+Some question that may arise are: Do the results agree with those obtained from the network models?
+What are the similarities or differences in the results?
+What are the possible underlying reason/structures for these similarities/differences?\\
+
 % definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
@@
 
+% talk about using real-world networks to compare against network models??
+
 \section{Opinion dynamics}\label{Opinion}
@@
 What are there benefits and possible shortcomings regarding to our goals?\\
-In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.\\ % + explain where they come from etc?
+In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.
+The reasons of choosing these specific networks will be explained, together with there possible strengths and weaknesses.\\ % + explain where they come from etc?
 Finally, Section \ref{opinionmodel} discusses the opinion dynamics model that is implemented in this thesis. Again we will give some focus on possible advantages and disadvantages of the model.
@@
 \section{Real-world networks}\label{realnetwork}
+% talk about reasons why you chose the networks you chose: eg. size (computation time),...
+
+Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies.
+However, most of these companies are not to keen on sharing any form of data.
+This makes it very hard to find accurate and complete social media network data.
+Hence, we chose to test the model on other types of social networks, either on-line or off-line.
+Some possibilities are any form of friendship networks, e-mail networks, collaboration networks, etc.
+\\
+\newline
+One of the decisive properties of the real-world networks that were eventually used, was there size.
+The main, and actually only, reason for this was the computation time.
+If the networks would be to large, it would take weeks, sometimes even months before there would be any kind of results.\\
+The size of the network models that were used, was around a thousand nodes.
+So for the real-world networks, we also used networks with a size of magnitude of thousand.
+This obviously benefits the computation time; it is, however, unknown how the network size affects the results.
+It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic.
 \section{Opinion dynamics models}\label{opinionmodel}
@@
 
+% put here: be carefull with extrapolating data/results to real world cases, need for more real data to compare with (eg data from role of facebook in elections in US), this lack of comparison between real data and network model data/analytical work = one of the downfalls/backsides of social physics??
+% relevance to society of knowing the impact of social media on opinion dynamics 
+% important to know the limits of the model, may be intuitive, but intuition maybe misleading (learned in physics) --> why we need experimental data to compare with!
+
 \chapter{Conclusions and outlook}

### 2021-04-05 draft.tex
added figure of first result

@@
 
-TEST FIGURE!!
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=14cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_WS_fracStubb_PR.png}}
-\caption{This is an image from a text that uses color to teach music. Make really really long caption, as long as textwidth}
-\captionsetup{width=\textwidth, labelfont=bf}
-\label{fig}
+\section{Impact of network structure}
+
+\subsection{Evolution of opinions}
+% write explanation here! Also mention all model parameters (eg Q=10, etc)!!
+\begin{figure}[h!]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_50-50_PR_REC.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10x100_10-001-0001_50-50_PR_REC.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+Four different network models are used (each with an average degree around 10 and a total number of nodes equal to 1000): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{cl} = 0.1,\ p_{add} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{cl} = 0.03,\ p_{add} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+Each figure is the result of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
 \end{figure}
-\noindent
-The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis.
-But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world.
-Results that are obtained from them should not be thoughtlessly extrapolated to real-world networks.
+\newpage
+\subsection{Echo chambers}
+
+\section{Impact of stubborn people}
+
+\section{Random versus community distributed opinions/stubborn agents}
 

### 2021-04-06 draft.tex
added more figures

@@
 \subsection{Evolution of opinions}
-% write explanation here! Also mention all model parameters (eg Q=10, etc)!!
-\begin{figure}[h!]
+\textbf{Starting conditions $\bm{P_0(0) = 0.5}$ and $\bm{P_1(0)=0.5}$}\\
+\newline
+ write explanation here! Also mention all model parameters (eg Q=10, etc)!!
+\begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
@@
   \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Four different network models are used (each with an average degree around 10 and a total number of nodes equal to 1000): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{cl} = 0.1,\ p_{add} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{cl} = 0.03,\ p_{add} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each figure is the result of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
 \end{figure}
 \newpage
+\noindent
+\textbf{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}\\
+\newline
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+\end{figure}
+
 \subsection{Echo chambers}
+\textbf{Starting conditions $\bm{P_0(0) = 0.5}$ and $\bm{P_1(0)=0.5}$}\\
+\newline
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC.png}
+    \caption{REC}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR.png}
+    \caption{PR}
+  \end{subfigure}
+  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+\textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100 ($10\times 100$), a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$).
+The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$.
+Each plot is the average of $10^2$ independent simulations.}%more explanation?
+\end{figure}
+\newpage
+\noindent
+\textbf{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}\\
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
+Each plot is the average of $10^2$ independent simulations.}
+\end{figure}
 

### 2021-04-06 draft.tex
latest update

--- 2021-04-05 draft.tex
@@
 At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $Q$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
+The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
 \begin{itemize}
-	\item The reference method (REF): this method randomly selects $Q$ items from the list $L_i(t)$
-	\item The old method (OLD): the $Q$ oldest post in $L_i(t)$ are displayed on the time-line of the user
-	\item The recent method (REC): shows the $Q$ most recent items from $L_i(t)$
-	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $Q$ elements.
+	\item The reference method (REF): this method randomly selects $S$ items from the list $L_i(t)$
+	\item The old method (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user
+	\item The recent method (REC): shows the $S$ most recent items from $L_i(t)$
+	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
 This filtering method is based on the semantic filtering explained above
@@
 % NOTE: maybe make somewhere in thesis link to real-world examples; for example the election system in USA: formation of echo chambers, different states voting for a different party
-\chapter{Materials and methods in this thesis}
+\chapter{Materials and methods in this thesis}\label{chap4}
 
@@
 
-\subsection{Evolution of opinions}
-% write explanation here! Also mention all model parameters (eg Q=10, etc)!!
-\begin{figure}[h!]
+In this section we will discuss some results related to the network structure and network topologies of the different network models that we used.
+We will mainly focus on the effect of a high clustering coefficient and a community structure.
+Two different opinion distributions are investigated
+\begin{itemize}
+	\item An initial $50/50$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.5$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.5$
+	\item An initial $20/80$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.2$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$
+\end{itemize}
+For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.
+
+\subsection{Starting conditions $\bm{P_0(0) = 0.5}$ and $\bm{P_1(0)=0.5}$}
+We start with the following opinion distribution: a fraction of nodes with opinion 0 equal to 0.5 ($P_0(0) = 0.5$) and a fraction of nodes with opinion 1 equal to 0.5 ($P_1(0) = 0.5$).
+The opinions are randomly distributed over the network.
+The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} % correct section??
+is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
+Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around 10
+\begin{itemize}
+	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$
+	\item The Watts-Strogatz model (WS) with mean degree $K=10$; two different values of the rewire probability $\beta$ are used, each of them has a slightly different value of the average clustering coefficient, namely $\beta = 0.01$ with $\left<C\right> \sim 0.60$ and $\beta = 0.06$ with $\left<C\right> \sim 0.55$.
+This model is used to investigate the effect of clustering
+	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+This model is used to investigate the effect of a community structure.
+	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
+This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$)
+\end{itemize}
+\textbf{Evolution of the opinion fractions}\\
+\newline
+Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of opinion 0 over time.
+Two different filtering algorithms are compared to each other, namely the PR method and the REC method (see Chapter \ref{chap3}, Section \ref{filter} for a detailed explanation).
+None of the network topologies, nor the filtering algorithms, is able to break the initial $50/50$ opinion distribution.
+Having a high clustering or a clear community structure has no impact on the prevalence of the opinions.
+The status quo is maintained and the system is not pushed towards a different equilibrium.
+\newpage
+\begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
@@
   \end{subfigure}
+  \captionsetup{format=plain}
   \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Four different network models are used (each with an average degree around 10 and a total number of nodes equal to 1000): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100 ($10\times100$), a high modularity case ($p_{cl} = 0.1,\ p_{add} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{cl} = 0.03,\ p_{add} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100 ($10\times100$), each community has the following WS parameters: $K = 10,\ \beta = 0.01$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each figure is the result of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50}
 \end{figure}
 \newpage
-\subsection{Echo chambers}
+\noindent
+\textbf{Echo chambers}\\
+\newline
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC.png}
+    \caption{REC}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR.png}
+    \caption{PR}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+\textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$).
+The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$.
+Each plot is the average of $10^2$ independent simulations.}%more explanation?
+\end{figure}
+
+\subsection{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}
+\textbf{Evolution of the opinion fractions}\\
+\newline
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+\end{figure}
+\textbf{Echo chambers}\\
+\newline
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
+Each plot is the average of $10^2$ independent simulations.}
+\end{figure}
 

### 2021-04-07 draft.tex
worked on results section

@@
 \addcontentsline{toc}{chapter}{Table of Contents}
-%\listoffigures
+\listoffigures
+\addcontentsline{toc}{chapter}{List of Figures}
 
@@
 \begin{itemize}
-	\item $\left<k\right>$ nodes at distance one $(d=1)$
-	\item $\left<k\right>^2$ nodes at distance two $(d=2)$
-	\item $\left<k\right>^3$ nodes at distance three $(d=3)$
-	\item ...
-	\item $\left<k\right>^d$ nodes at distance $d$
+	\item $\left<k\right>$ nodes at distance one $(d=1)$,
+	\item $\left<k\right>^2$ nodes at distance two $(d=2)$,
+	\item $\left<k\right>^3$ nodes at distance three $(d=3)$,
+	\item ...,
+	\item $\left<k\right>^d$ nodes at distance $d$.
 \end{itemize}
@@
 \begin{enumerate}
-	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
+	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side).
 	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
-Rewiring is done in such a way that self-loops and edge duplication are avoided
+Rewiring is done in such a way that self-loops and edge duplication are avoided.
 \end{enumerate}
@@
 \begin{itemize}
-	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$
-	\item The probability that edges are rewired back to each other is negligible
+	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$.
+	\item The probability that edges are rewired back to each other is negligible.
 \end{itemize}
@@
 \begin{itemize}
-	\item The number of nodes $N$ 
-	\item A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities
-	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$
+	\item The number of nodes $N$, 
+	\item A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities,
+	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.
 \end{itemize}
@@
 \begin{itemize}
-	\item Popularity filtering: promoting content that is popular across the platform
-	\item Semantic filtering: promoting posts that are similar to previous consumed posts
-	\item Collaborative filtering: promoting content that people similar or close to us consume
+	\item Popularity filtering: promoting content that is popular across the platform,
+	\item Semantic filtering: promoting posts that are similar to previous consumed posts,
+	\item Collaborative filtering: promoting content that people similar or close to us consume.
 \end{itemize} 
@@
 \begin{itemize}
-	\item The reference method (REF): this method randomly selects $S$ items from the list $L_i(t)$
-	\item The old method (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user
-	\item The recent method (REC): shows the $S$ most recent items from $L_i(t)$
+	\item The reference method (REF): this method randomly selects $S$ items from the list $L_i(t)$.
+	\item The old method (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user.
+	\item The recent method (REC): shows the $S$ most recent items from $L_i(t)$.
 	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
-This filtering method is based on the semantic filtering explained above
+This filtering method is based on the semantic filtering explained above.
 \end{itemize}
@@
 \begin{itemize}
-	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...)
-	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea
+	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...).
+	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea.
 	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
 A uniform random number $f$ between 0 and 1 is drawn.
-If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion
+If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.
 \end{itemize}
@@
 \begin{itemize}
-	\item If $p_A > p_B$ and $p_A > T$, opinion A is adopted
-	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted
-	\item In all other cases, the current opinion of the node is not changed
+	\item If $p_A > p_B$ and $p_A > T$, opinion A is adopted,
+	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted,
+	\item In all other cases, the current opinion of the node is not changed.
 \end{itemize}
@@
 \begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER)
-	\item The Watts-Strogatz model (WS)
+	\item The Erd\H{o}s-R\'{e}nyi model (ER),
+	\item The Watts-Strogatz model (WS),
 	\item The stochastic block model, of which two types are implemented
 		\begin{itemize}
-			\item[\tiny$\bullet$] The regular stochastic block model (SBM)
-			\item[\tiny$\bullet$] The stochastic Watts-Strogatz block model (SBM-WS)
+			\item[\tiny$\bullet$] The regular stochastic block model (SBM),
+			\item[\tiny$\bullet$] The stochastic Watts-Strogatz block model (SBM-WS).
 		\end{itemize}
@@
 \begin{itemize}
-	\item An initial $50/50$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.5$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.5$
-	\item An initial $20/80$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.2$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$
+	\item An initial $50/50$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.5$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.5$.
+	\item An initial $20/80$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.2$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$.
+Hence, we are dealing with a minority and a majority opinion.
 \end{itemize}
-For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.
+For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.\\
+\newline
+Before sharing the results, a quick disclaimer: Some of the results that are shared in this section, were already obtained by prof.
+Nicola Perra and prof.
+Luis E.C.
+Rocha in their paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+These results include the effect of a high clustering coefficient (Watts-Strogatz model) and a random network model.
+Similar results (albeit with a different network size and average degree) are included in this section to give a complete overview and to be able to compare a community structure to a high clustering coefficient.
 
@@
 is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
+\newpage
+\noindent
 Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around 10
 \begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$
+	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
 	\item The Watts-Strogatz model (WS) with mean degree $K=10$; two different values of the rewire probability $\beta$ are used, each of them has a slightly different value of the average clustering coefficient, namely $\beta = 0.01$ with $\left<C\right> \sim 0.60$ and $\beta = 0.06$ with $\left<C\right> \sim 0.55$.
-This model is used to investigate the effect of clustering
+This model is used to investigate the effect of clustering.
 	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
@@
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$)
+This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
 \end{itemize}
@@
 \newline
-Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of opinion 0 over time.
+Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of opinion 0 over time, together with the standard deviation $\sigma$.
 Two different filtering algorithms are compared to each other, namely the PR method and the REC method (see Chapter \ref{chap3}, Section \ref{filter} for a detailed explanation).
@@
   \captionsetup{format=plain}
-  \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
 Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR.png}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR_8x7.png}
     \caption{PR}
@@
   \captionsetup{format=plain}
-  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$.
-Each plot is the average of $10^2$ independent simulations.}%more explanation?
+Each plot is the average of $10^2$ independent simulations.}
+\label{echo_50_50}
 \end{figure}
+\noindent
+Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$.
+The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
+The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0).
+Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1.
+The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
+Error bars are omitted for the sake of an improved visualization.\\
+These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\
+In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR method.
+The results for both filtering algorithms look qualitatively the same.
+The main difference is that the formation of echo chambers, using the PR method, is about ten times as big compared to the formation of echo chambers using the REC method.
+\\
+The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a very limited, sign of echo chambers, for both the REC method and the PR method.
+Networks with a high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
+Going to slightly higher values of the clustering coefficient ($\left<C\right> \sim 0.55$ to $\left<C\right> \sim 0.60$) in the Watts-Strogatz model results in significantly larger echo chambers.
+This suggests that the formation of echo chambers is really sensitive to changes in the clustering coefficient.
+The stochastic block model (SBM) with a high modularity, hence, a clear community structure, does show an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller than compared to networks with a high clustering.
+The stochastic block model with a low modularity has no, or very limited, echo chambers, similar as for the Erd\H{o}s-R\'{e}nyi
+model.
+This is to be expected, since a stochastic block model with a low modularity, and hence a very blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network.
+The stochastic Watts-Strogatz block model (SBM-WS) has both a community structure and a high average clustering.
+The formation of echo chambers when using this model is approximately of the same size compared to the formation of echo chambers in a Watts-Strogatz model with a similar average clustering.
+This suggests that it is mainly the high clustering that drives the system to form echo chambers, while the community structure plays a less important role.
+However, as said before, the stochastic block model with a high modularity also has a (smaller) formation of echo chambers, so having a clear and strong community structure definitely plays some role in phenomena such as polarization and the formation of opinion clusters.
+\\
+Note that for an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (see Figure \ref{ev_op_50_50}).
+Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially lead to polarization and formation of echo chambers.
 
 \subsection{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}
+
+The initial fraction of nodes with opinion 0 is now equal to $P_0(0) = 0.2$, opinion 0 thus becomes the minority opinion.
+The initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$, hence this is the majority opinion.
+The opinions are, initially, randomly distributed over the system.
+Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
+\newline
+Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $10$
+\begin{itemize}
+	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
+	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$.
+The value of the average clustering coefficient is equal to $\left<C\right> \sim 0.55$.
+This model is used to investigate the effect of clustering.
+	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+This model is used to investigate the effect of a community structure.
+	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
+This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
+\end{itemize}
 \textbf{Evolution of the opinion fractions}\\
 \newline
+Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion 0, which is the minority opinion in this case, over time, together with the standard deviation $\sigma$.
+For each network model two filtering algorithms are compared, namely the REC filtering algorithm and the PR filtering algorithm.
+\\
+The REC method is not able to break the prevalence of opinion 0 over time for neither of the four different network models.
+When using the PR method, the results look very different.
+The prevalence of the minority opinion, $P_0(t)$, decreases over time.
+This decrease is a direct consequence of the PR filtering algorithm.
+Since the PR method orders the posts according to the users' preferences, nodes with the majority opinion 1 will mainly get to see posts with the similar opinion 1 on their time-line, which limits the chance of changing their opinion to the minority opinion.
+Nodes with the minority opinion 0, on the other hand, will obviously see posts with opinion 0 on their time-line when using the PR method.
+However, since there are more nodes with opinion 1 than nodes with opinion 0, the nodes with the minority opinion 0 will also get to see a significant number of posts with the majority opinion 1 on their time-line.
+The conclusion is that, when using the PR method, there is an imbalance in which opinion can switch to which: nodes with the minority opinion 0 have a chance of switching to the majority opinion 1, whereas nodes with the majority opinion 1 have (almost) no chance of switching to the minority opinion 0.
+This imbalance leads to the decrease of the minority opinion 0.
+
+% Talk about effect network topologies??
+See paper + simulation with N = 10.000 and N = 100.000 --> mention possible discrepancies and effect of network size??.
+Having a high clustering coefficient and/or a clear community structure plays no role in the e
+
 \begin{figure}[H]
@@
   \captionsetup{format=plain}
-  \caption{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
 Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+\label{ev_op_20_80}
 \end{figure}
@@
   \captionsetup{format=plain}
-  \caption{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 Each plot is the average of $10^2$ independent simulations.}
+\label{echo_20_80}
 \end{figure}
+
+\subsection{Relation between formation of echo chambers and community structure}
+
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_12x10}
+	\label{echo_vs_mod}
+	\captionsetup{format=plain}
+	\caption[Relation between the formation of echo chambers and community structure (modularity)]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+The initial opinion distribution is $50/50$.
+The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10.
+The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around 10.
+Each data point is the result of $10^2$ independent simulations.}
+\end{figure}
+% mention that error bars represent standard error + reason + mention values for p_cl and p_add together with modularity, maybe clustering
 

### 2021-04-08 draft.tex
continued working on results section

@@
 \subsection{Stubbornness in binary models}
+\label{stubbBin}
 % not finished!!
@@
 
+In this chapter the results are shared and discussed. The chapter in subdivided into three main sections
+\begin{itemize}
+	\item The impact of network structure: the effect of different network topologies on the evolution of the opinions and the formation of echo chambers is reviewed.
+Main focus lies on networks with a high clustering coefficient and/or a community structure.
+	\item The impact of stubborn people: stubborn people are introduced in the system; in this section the effects of those stubborn people on the evolution of opinions and the formation of echo chambers will be discussed.
+	\item Random versus community distributed opinions/stubborn agents: this section investigates different ways of distributing the opinions and/or the stubborn agents in a network model with a community structure (SBM).
+\end{itemize}
+
 \section{Impact of network structure}
@@
 \subsection{Starting conditions $\bm{P_0(0) = 0.5}$ and $\bm{P_1(0)=0.5}$}
+\label{50-50}
 We start with the following opinion distribution: a fraction of nodes with opinion 0 equal to 0.5 ($P_0(0) = 0.5$) and a fraction of nodes with opinion 1 equal to 0.5 ($P_1(0) = 0.5$).
@@
 is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
-\newpage
-\noindent
+\newline
 Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around 10
@@
 \textbf{Echo chambers}\\
-\newline
 \begin{figure}[H]
@@
 This model is used to investigate the effect of clustering.
+	\newpage
 	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
@@
 Having a high clustering coefficient and/or a clear community structure plays no role in the e
-
+\newpage
 \begin{figure}[H]
@@
 \end{figure}
+\newpage
+\noindent
 \textbf{Echo chambers}\\
 \newline
+Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models.
+For each network, the REC and PR method are compared to each other.
+The interpretation of the figure is the same as was described in Subsection \ref{50-50}, the echo chamber paragraph.\\
+The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity) look qualitatively the same.
+Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz model agree with each other as well.
+This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
+There is no visible effect for having a clear community structure.
+This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared.
+There is (almost) no difference in the results for a high and a low modularity.\\
+\newline
+Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
+When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered.
+This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
+There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates an increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
+From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions.
+The opinion distribution is however shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
+\\
+The PR method shows a completely different tendency.
+There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1.
+This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
+\\
+\newline
+Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
+For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+
 \begin{figure}[H]
@@
     \caption{WS}
+    \label{ws}
   \end{subfigure}
@@
     \caption{SBM}
+  	\label{sbm}    
   \end{subfigure}
@@
     \caption{SBM-WS}
+    \label{sbm-ws}
   \end{subfigure}
@@
 	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_12x10}
-	\label{echo_vs_mod}
 	\captionsetup{format=plain}
@@
 Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_mod}
 \end{figure}
-% mention that error bars represent standard error + reason + mention values for p_cl and p_add together with modularity, maybe clustering
-
+
+Now, lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$.
+In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
+The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$).
+This is a measure for the formation of echo chambers (opinion clusters) in the system.
+The data points represent the value of the echo chamber size in networks with a different modularity.
+The initial opinion distribution was always $50/50$.
+\newpage
+\noindent
+In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around 10)
+\begin{itemize}
+	\item $Q=0 \rightarrow$ Erd\H{o}s-R\'{e}nyi model with edge probability $p=0.01$,
+	\item $Q \sim 0.2 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.03$ and the edge probability between communities is $p_{\text{add}} = 0.008$,
+	\item $Q \sim 0.47 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.05$ and the edge probability between communities is $p_{\text{add}} = 0.005$,
+	\item $Q \sim 0.6 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.07$ and the edge probability between communities is $p_{\text{add}} = 0.004$,
+	\item $Q \sim 0.9 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.1$ and the edge probability between communities is $p_{\text{add}} = 0.001$.
+\end{itemize}
+The data points are shown together with the standard error $SE = \sigma /\sqrt{N}$.
+The standard error $SE$ is chosen here instead of the standard deviation $\sigma$.
+This is because the standard deviation is quite large.
+Each data point is the average of $10^2$ independent simulations.
+The dispersion/variation, which is measured by the standard deviation, in the set of the $10^2$ obtained values is quite large.
+However, if we run the $10^2$ simulations over and over again, each time calculating the average, the obtained value of the average is always more or less the same.
+This suggests that the value of the calculated average lies close to the real average.
+Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.\\
+\newline
+The PR filtering algorithm is compared to the REC filtering algorithm.
+For most values of the modularity, the REC method produces almost no echo chambers.
+If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than two times as much echo chambers than in the beginning at $t=0$.\\
+For the PR method, the increase in the size of the echo chambers with increasing modularity is much more significant.
+From a modularity approximately equal to $Q \sim 0.5$, we can see a sharp increase in the formation of echo chambers.
+For modularity values smaller than $Q \sim 0.5$, the formation of echo chambers remains very small, even when using the PR method.\\
+These results strongly suggest that a community structure enhances the formation of echo chambers in the system, mostly when using the PR filtering algorithm.
+A quite clear community structure, represented by a high modularity, is however needed to really see the effect of this community structure on the formation of opinion clusters.
+
+\newpage
 \section{Impact of stubborn people}
+In this section, the impact of introducing stubborn people will be reviewed. We focus on two different models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
+\begin{itemize}
+	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
+	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
+\end{itemize}
+
+\subsection{Including a stubbornness parameter}
+\label{stubbpar}
+Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
+\begin{itemize}
+	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
+	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
+\end{itemize}
+For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
+We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around 10)
+\begin{itemize}
+	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
+	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
+	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+This model is used to investigate the effect of a community structure.
+	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
+This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
+\end{itemize}
+For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
+\newline
+\textbf{Making a fraction of the nodes completely stubborn}\\
+\newline
+nsdjk\\
+\newline
+\textbf{Making all nodes partly stubborn}\\
+\newline
+jsqnbd\\
+
+\subsection{The majority threshold model}
+\label{majThres}
 
@@
 \chapter{Conclusions and outlook}
-
+% also discuss need to investigate effect of changing N, S and p_act
 \newpage

### 2021-04-09 draft.tex
latest update

@@
 The conclusion is that, when using the PR method, there is an imbalance in which opinion can switch to which: nodes with the minority opinion 0 have a chance of switching to the majority opinion 1, whereas nodes with the majority opinion 1 have (almost) no chance of switching to the minority opinion 0.
-This imbalance leads to the decrease of the minority opinion 0.
+This imbalance leads to the decrease of the minority opinion 0.\\
+\newline
+From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist for all the network models, $P_0(500) \sim 0$.
+This result differs from the conclusion that prof.
+N.
+Perra and prof.
+L.E.C.
+Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+They found that the minority opinion was able to survive over time, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
+% explain further, but first test WS, K=6, beta=0.01, N=1000, 20-80, PR
 
@@
 For most values of the modularity, the REC method produces almost no echo chambers.
-If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than two times as much echo chambers than in the beginning at $t=0$.\\
+If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
 For the PR method, the increase in the size of the echo chambers with increasing modularity is much more significant.
@@
 \newline
-nsdjk\\
+We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.\\
 \newline

### 2021-04-10 draft.tex
new figure results section

@@
 
-\subsection{Relation between formation of echo chambers and community structure}
+\subsection{Relation between formation of echo chambers and community structure}\label{echoVSmod}
 
@@
 For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
-\newline
+\newpage
+\noindent
 \textbf{Making a fraction of the nodes completely stubborn}\\
 \newline
-We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.\\
-\newline
+We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$).
+We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.
+In all cases, the stubborn nodes are randomly distributed over the network.\\
+\newline
+Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm.
+The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.
+Data points are again shown together with their standard error, $SE$, for the same reasons as explained in Section \ref{echoVSmod}.\\
+The Erd\H{o}s-R\'{e}nyi model (ER) shows no appearance of echo chambers.
+Introducing a fraction of completely stubborn nodes doesn't change this behavior.
+The stochastic block model with a low modularity behaves similar as the Erd\H{o}s-R\'{e}nyi model (no formation of echo chambers).
+Note that the data points for the SBM with a low modularity are not visible in Figure \ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
+The other three networks (WS, SBM-WS and SBM with a high modularity) all show qualitatively the same behavior.
+If an increasing fraction of completely stubborn nodes is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn people.
+This can be explained as follows.
+Having a fraction of completely stubborn nodes, reduces the effective number of neighbors that can change their opinion.
+This in turn leads to the hampering of the formation of echo chambers.\\
+The WS and SBM-WS networks are the high clustering networks and for a small fraction of stubborn people they show the largest echo chambers.
+The stochastic block model with a high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for a fraction of stubborn nodes smaller than fracRes $\leqslant 0.1$.
+This leads us to make the same conclusion as before: a community structure enhances the formation of echo chambers compared to a random network, however not as significant as networks with a high clustering coefficient.
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus fraction of completely stubborn nodes, PR]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around 10: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.}
+\label{echo_vs_frac_complStubb}
+\end{figure}
+\noindent
 \textbf{Making all nodes partly stubborn}\\

### 2021-04-10 draft.tex
new figure results section

@@
 For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+% explain black dotted line??
 
@@
 \end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
+    \caption{REF}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
+    \caption{REC}
+    \label{ws}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes, REF and REC]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
+\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around 10: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.}
+\label{echo_vs_frac_complStubb_REF-REC}
+\end{figure}
+% eplain REC/REF figure!!
 \noindent

### 2021-04-11 draft.tex
added extra result

@@
 
-In this chapter the results are shared and discussed. The chapter in subdivided into three main sections
+In this chapter the results are shared and discussed. The chapter in subdivided into four main sections
 \begin{itemize}
@@
 	\item Random versus community distributed opinions/stubborn agents: this section investigates different ways of distributing the opinions and/or the stubborn agents in a network model with a community structure (SBM).
+	\item Real-world networks: the opinion dynamics model of this thesis will be ran on some real-world networks.
+This section will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
 \end{itemize}
@@
 
-In this section we will discuss some results related to the network structure and network topologies of the different network models that we used.
+In this section we will discuss some results related to the network structure and network topologies of the different network models that are used in this thesis.
 We will mainly focus on the effect of a high clustering coefficient and a community structure.
@@
 \newline
-Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around 10
+Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
 \begin{itemize}
@@
 This model is used to investigate the effect of clustering.
-	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
 This model is used to investigate the effect of a community structure.
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
@@
 \textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$).
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$).
 The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$.
@@
 Note that for an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (see Figure \ref{ev_op_50_50}).
-Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially lead to polarization and formation of echo chambers.
+Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a more clustered opinion distribution and thus lead to polarization and formation of echo chambers.
 
@@
 \newline
-Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $10$
+Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
 \begin{itemize}
@@
 They found that the minority opinion was able to survive over time, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
+However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
+In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
+In both cases the maximum length of the actual time-line is set equal to $S=20$.
+The difference in the results is most likely an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
+This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
+In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
+In this figure we can see that the minority opinion 0 doesn't vanish from the system over time, but instead reaches a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.
 % explain further, but first test WS, K=6, beta=0.01, N=1000, 20-80, PR
@@
 Having a high clustering coefficient and/or a clear community structure plays no role in the e
-\newpage
 \begin{figure}[H]
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used (each with an average degree around 10 and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
@@
 \newpage
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
+    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
+    \label{ws1000}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
+    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
+    \label{ws10000}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$.
+In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+\label{ev_op_WS_k=6}
+\end{figure}
 \noindent
@@
 \newline
-Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models.
+Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
 For each network, the REC and PR method are compared to each other.
 The interpretation of the figure is the same as was described in Subsection \ref{50-50}, the echo chamber paragraph.\\
+\newline
 The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity) look qualitatively the same.
@@
 There is (almost) no difference in the results for a high and a low modularity.\\
-\newline
+\newpage
+\noindent
 Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
@@
 For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+% remake plots?? because for PR minority opinion vanishes does why more echo chambers??? 
 % explain black dotted line??
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around 10: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
@@
 The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10.
-The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around 10.
+The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around $\left<k\right> \sim 10$.
 Each data point is the result of $10^2$ independent simulations.}
@@
 \noindent
-In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around 10)
+In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
 \begin{itemize}
@@
 For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
-We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around 10)
+We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
 \begin{itemize}
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around 10: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.}
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
 \label{echo_vs_frac_complStubb}
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around 10: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.}
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
 \label{echo_vs_frac_complStubb_REF-REC}
@@
 \section{Random versus community distributed opinions/stubborn agents}
+
+\section{Real-world networks}
 

### 2021-04-11 draft.tex
added extra result

@@
 This imbalance leads to the decrease of the minority opinion 0.\\
-\newline
-From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist for all the network models, $P_0(500) \sim 0$.
-This result differs from the conclusion that prof.
-N.
-Perra and prof.
-L.E.C.
-Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
-They found that the minority opinion was able to survive over time, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
-However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
-In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
-In both cases the maximum length of the actual time-line is set equal to $S=20$.
-The difference in the results is most likely an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
-This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
-In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion 0 doesn't vanish from the system over time, but instead reaches a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.
-% explain further, but first test WS, K=6, beta=0.01, N=1000, 20-80, PR
-
-% Talk about effect network topologies??
-See paper + simulation with N = 10.000 and N = 100.000 --> mention possible discrepancies and effect of network size??.
-Having a high clustering coefficient and/or a clear community structure plays no role in the e
+\newpage
+\noindent
 \begin{figure}[H]
@@
     \caption{WS}
+    \label{ws_20-80_op}
   \end{subfigure}
@@
     \caption{SBM-WS}
+    \label{sbm-ws_20-80_op}
   \end{subfigure}
@@
 \end{figure}
-\newpage
+\noindent
+From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist, $P_0(500) \sim 0$.
+This is certainly the case for the low clustering networks (ER and SBM).
+The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_0(500) \sim 0.007$ for the SBM-WS network (Figure \ref{sbm-ws_20-80_op}) and an average value of $P_0(500) \sim 0.009$ for the WS network (Figure \ref{ws_20-80_op}).\\
+This result differs from the conclusion that prof.
+N.
+Perra and prof.
+L.E.C.
+Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
+However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
+In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
+In both cases the maximum length of the actual time-line is set equal to $S=20$.
+The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
+This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
+In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
+In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500 \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.
+
 \begin{figure}[H]
@@
 \newline
-The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity) look qualitatively the same.
-Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz model agree with each other as well.
+The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
+Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz block model agree with each other as well.
 This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
-There is no visible effect for having a clear community structure.
+There are no visible effects of having a clear community structure.
 This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared.
 There is (almost) no difference in the results for a high and a low modularity.\\
-\newpage
-\noindent
+\newline
 Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
@@
 For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
-% remake plots?? because for PR minority opinion vanishes does why more echo chambers??? 
+An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
+This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity.
+In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180.\\
+This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
+Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population.
+Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
+The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
+This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution.
+In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+% talk about small spike around 1 for PR method??
 % explain black dotted line??

### 2021-04-11 draft.tex
latest update

@@
 The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
-This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
+This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results (together with the standard deviation $\sigma$) for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
 In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500 \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.
+In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.
 
@@
 This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
-There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates an increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
+There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
 From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions.
-The opinion distribution is however shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
+The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
 \\
@@
 \end{itemize}
-
-\subsection{Including a stubbornness parameter}
-\label{stubbpar}
-Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
-\begin{itemize}
-	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
-	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
-\end{itemize}
-For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
 We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
@@
 For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
+Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with a activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.
+
+\subsection{Including a stubbornness parameter}
+\label{stubbpar}
+Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
+\begin{itemize}
+	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
+	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
+\end{itemize}
+For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
+
 \newpage
@@
 \end{figure}
+\noindent
+Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
+The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
+Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease.
+This is especially clear for the high clustering networks (WS and SBM-WS).
+The SBM with a high modularity also displays a tiny bump.
+The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\
+These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn agents enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
+% explain potential reasons (eg echo chambers around stubborn nodes --> keep on pushing their opinion to the neighbors, not the case for PR) + check those + need for futher research.
+If the fraction of completely stubborn nodes becomes bigger than fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
+This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions.
+This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.
 
@@
 \end{figure}
-% eplain REC/REF figure!!
+% don't forget the evolution of opinions!!
 \noindent

### 2021-04-13 draft.tex
latest update

@@
 % use pseudo code here??
+% explain the main model and state that this is used in thesis unless specified otherwise
 
@@
 For each network, the REC and PR method are compared to each other.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50}, the echo chamber paragraph.\\
+The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.\\
 \newline
@@
 \section{Impact of stubborn people}
-In this section, the impact of introducing stubborn people will be reviewed. We focus on two different models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
+In this section, the impact of introducing stubborn people will be reviewed. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
 \begin{itemize}
@@
 	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
+	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<C\right> \sim 0.55$.
+This model is used to investigate the effect of clustering.
 	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
@@
 \end{itemize}
-For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
-Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with a activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.
+For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.
 
@@
 For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
+Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.
 
@@
 Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
+Data points are again shown together with their standard error $SE$.
 The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
@@
     \caption{REC}
-    \label{ws}
   \end{subfigure}
@@
 \end{figure}
-% don't forget the evolution of opinions!!
+% don't forget the evolution of opinions + also for 20/80, do each network, PR, REC, REF for eg.
+stubb 0.4 (so two fig (50-50 and 20-80) with 4 subplots (one for each network), each plot has PR, REF, REC) + if all stubb frac behaves same, explain in text, if not make extra plot!!
 \noindent
@@
 \newline
-jsqnbd\\
+All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
+\newline
+Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod} and the data points are shown together with their standard error $SE$.
+The results look very similar to the case where a fraction of the nodes is completely stubborn and when using the PR method, which is depicted in Figure \ref{echo_vs_frac_complStubb}.\\
+The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers, as before, and this is not changed by making the nodes more and more stubborn.
+The same result is obtained for the stochastic block model with a low modularity (the data points lie again behind those of the ER network).\\
+The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
+If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.\\
+This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
+If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
+If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
+In Figure \ref{echo_vs_all_frac_stubb_PR} this is represented by an echo chamber size equal to one for $r=1$.
+
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, PR]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_all_frac_stubb_PR}
+\end{figure}
+\newpage
+\noindent
+In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
+Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
+Data points are again displayed with their standard error $SE$.\\
+We can make similar observations as we did in the "Making a fraction of the nodes completely stubborn" paragraph.
+The results for the REF and REC method look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.\\
+The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases.
+For the REF method, this increases reaches a maximum around $r \sim 0.7$ and then starts to slightly decrease between $r\sim 0.7$ and $r\sim 0.9$.
+Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers (echo chamber values around 17 for $r\sim 0.9$ to echo chamber values equal to one for $r=1$).
+The REC method also seems to have a maximum around $r \sim 0.7$ for the WS model, the SBM-WS model on the other hand displays a continuous increase until $r \sim 0.9$.
+In both case we have again a sharp drop from $r \sim 0.9$ to $r=1$.
+If $r=1$, the nodes are completely stubborn and thus keep their initial opinion.
+Hence, the number of like minded neighborhoods doesn't change over time, which gives a normalized number of echo chambers equal to one.
+It is however remarkable that for very high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is very large and even larger than if the nodes are not stubborn.
+We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithm in the case of networks with a high clustering coefficient.
+This is the complete opposite of the case where the PR method is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
+% REASON??
+\\
+If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.
+The increase is however not as strong as for the high clustering networks.\\
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) don't seem to have any dependency on the stubbornness level of the nodes.
+This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithm that drives this increase in the formation of echo chambers with increasing $r$.
+% talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REF_8x7.png}
+    \caption{REF}
+    \label{REF_all_frac_stubb}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REC_8x7.png}
+    \caption{REC}
+    \label{REC_all_frac_stubb}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, REF and REC]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
+\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_all_frac_stubb_REF-REC}
+\end{figure}
 
@@
 \label{majThres}
+
+In this subsection, the results of the majority threshold model are shared and discussed.
+This model is explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
+Note, that we don't make use of a probabilistic majority model here, but instead a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
+The rest of the model (filtering algorithms, activation mechanism,...) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
+The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
+The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.
+% explain opinion dynamics or refer to relevant sections!!!
+
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_threshold_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, PR]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_threshold_PR}
+\end{figure}
+% standard deviation in this figure --> explain reason: emphasis on spread in data --> shows that at beginning remains more or less constant
 

### 2021-04-14 draft.tex
latest update

@@
 \end{figure}
+\newpage
 \noindent
 \textbf{Echo chambers}\\
-\newline
-Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
-For each network, the REC and PR method are compared to each other.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.\\
-\newline
-The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
-Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz block model agree with each other as well.
-This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
-There are no visible effects of having a clear community structure.
-This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared.
-There is (almost) no difference in the results for a high and a low modularity.\\
-\newline
-Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
-When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered.
-This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
-There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
-From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions.
-The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
-\\
-The PR method shows a completely different tendency.
-There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1.
-This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
-\\
-\newline
-Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
-For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
-An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
-This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity.
-In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180.\\
-This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
-Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population.
-Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
-The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
-This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution.
-In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
-% talk about small spike around 1 for PR method??
-% explain black dotted line??
 
@@
 \end{figure}
+\newpage
+\noindent
+Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
+For each network, the REC and PR method are compared to each other.
+The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.\\
+\newline
+The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
+Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz block model agree with each other as well.
+This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
+There are no visible effects of having a clear community structure.
+This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared.
+There is (almost) no difference in the results for a high and a low modularity.\\
+\newline
+Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
+When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered.
+This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
+There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
+From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions.
+The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
+\\
+The PR method shows a completely different tendency.
+There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1.
+This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
+\\
+\newline
+Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
+For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
+This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity.
+In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180.
+These large spikes, indicating a large increase of echo chambers around the minority opinion, don't necessary mean that we have a huge absolute number of echo chambers around the minority opinion 0 at $t=500$.
+They merely indicate that compared to the beginning at $t=0$ the increase is large.
+If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
+This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
+Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population.
+Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
+The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
+This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution.
+In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+% talk about small spike around 1 for PR method??
+% explain black dotted line??
 
@@
 The initial opinion distribution was always $50/50$.
-\newpage
-\noindent
+\\
+\newline
 In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
@@
 
-\newpage
 \section{Impact of stubborn people}
@@
 This leads us to make the same conclusion as before: a community structure enhances the formation of echo chambers compared to a random network, however not as significant as networks with a high clustering coefficient.
-\begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
-	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn nodes, PR]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_frac_complStubb}
-\end{figure}
-\noindent
+\\
+\newline
 Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
@@
 \begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_frac_complStubb}
+\end{figure}
+
+\begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes, REF and REC]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
 \end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{PR}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{REC}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
+\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_frac_complStubb_PR-REC_20-80}
+\end{figure}
+% explain figure + difference in y-ax
+
 % don't forget the evolution of opinions + also for 20/80, do each network, PR, REC, REF for eg.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, PR]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, REF and REC]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
 The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
-The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.
+The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.\\
 % explain opinion dynamics or refer to relevant sections!!!
+\newline
+Figure \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
+In this figure, the PR filtering algorithm is used.
+Data points are now shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.\\
+The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat, for $T \leqslant 0.5$.
+This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
+The majority opinion will always be larger than this threshold (the sum of the two opinion fractions is always equal to one).
+If $T \geqslant 0.5$, the threshold start to have an effect.
+The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
+Similar observations can be made for the stochastic block model with a high modularity.
+The only difference is of a quantitative nature (the formation of echo chambers is not as large as for the high clustering networks).\\
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points lie behind those of the Erd\H{o}s-R\'{e}nyi network) do not enhance the formation of echo chambers in any way.
+Including a threshold $T$ does not change this tendency.
+% make coupling to ev of op??
 
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, PR]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
 \end{figure}
-% standard deviation in this figure --> explain reason: emphasis on spread in data --> shows that at beginning remains more or less constant
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REF_8x7.png}
+    \caption{REF}
+    \label{REF_thres}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REC_8x7.png}
+    \caption{REC}
+    \label{REC_thres}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_threshold_REF-REC}
+\end{figure}
+\noindent
+Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
+The initial opinion distribution was again $50/50$.
+The results for the REF and REC method strongly resemble each other.
+Data points are again shown together with their standard deviation $\sigma$ to really emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 100$).
+These large errors are a direct result of the majority model.
+If we run the majority model once, the model forces the system to converge to a single opinion.
+To which of the two opinions the model converges is a matter of random fluctuations.
+So, on average, the opinion distribution will remain $50/50$.
+However, each run the system converges to one of the two opinions, with the result that the number of echo chambers around this opinion strongly increases.
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion 1 (and thus resulting in large echo chambers around opinion 1) and there will have been simulations that converge to opinion 0 (resulting in almost no echo chambers around opinion 1).\\
+Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
+This is not due to any network features, but a direct consequence of the usage of the majority model.
+As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes kind of a huge, single echo chamber).
+The PR method seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
+This can be clearly seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than when using the REC/REF methods.\\
+\newline
+Figure \ref{ev_op_majT_one_sim} illustrates this particular behavior of the majority model more clearly.
+The figure shows the evolution of opinion 0 over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion 0; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion 0 as to opinion 1, the opinion to which the system eventually converges is a consequence of random fluctuations).
+% explain figure!!
+
+% make link to figures op frac+ hampering PR + network structures!!
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_majT=0_one_sim_8x7.png}
+    \caption{ER}
+    \label{er_majT}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_majT=0_one_sim_8x7.png}
+    \caption{WS}
+    \label{ws_majT}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_majT=0_one_sim_8x7.png}
+    \caption{SBM}
+    \label{sbm_majT}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_majT=0_one_sim_8x7.png}
+    \caption{SBM-WS}
+    \label{sbm-ws_majT}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.
+Results for one simulation of the majority threshold model with $T=0$ ]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.}
+\label{ev_op_majT_one_sim}
+\end{figure}
 

### 2021-04-15 draft.tex
latest update

@@
 
-In this chapter the results are shared and discussed. The chapter in subdivided into four main sections
+In this chapter the results are shared and discussed. The chapter is subdivided into four main sections
 \begin{itemize}
@@
 This model is used to investigate the effect of clustering.
-	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
 This model is used to investigate the effect of a community structure.
@@
 The stochastic Watts-Strogatz block model (SBM-WS) has both a community structure and a high average clustering.
-The formation of echo chambers when using this model is approximately of the same size compared to the formation of echo chambers in a Watts-Strogatz model with a similar average clustering.
+The formation of echo chambers when using this model is approximately of the same size as the formation of echo chambers in a Watts-Strogatz model with a similar average clustering.
 This suggests that it is mainly the high clustering that drives the system to form echo chambers, while the community structure plays a less important role.
@@
 	\newpage
-	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
 This model is used to investigate the effect of a community structure.
@@
 They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
-However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
+The results that they obtained are, however, for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
 In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
@@
 In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.
+In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.
 
@@
     \caption{ER}
+    \label{er_echo_20-80}
   \end{subfigure}
@@
 For each network, the REC and PR method are compared to each other.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.\\
+The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.
+Error bars are again omitted for a better visualization.\\
 \newline
@@
 This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
+This reduced visibility of the minority opinion 0 in these neighborhoods drives them eventually to adopt opinion 1, with an increase in the number of echo chambers around the majority opinion 1 as a consequence.
+Remember from Figure \ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion.
+Hence, the increase in the formation of echo chambers around the majority opinion 1 is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
+This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the nodes carrying the minority opinion 0 becomes equal to zero.
 \\
@@
 These large spikes, indicating a large increase of echo chambers around the minority opinion, don't necessary mean that we have a huge absolute number of echo chambers around the minority opinion 0 at $t=500$.
-They merely indicate that compared to the beginning at $t=0$ the increase is large.
+They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion 0 at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion 0 at $t=500$).
 If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
 This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
-Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population.
+Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks.
 Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
@@
 
-Now, lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$.
+Lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$.
 In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
@@
 \\
-\newline
+\newpage
+\noindent
 In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
@@
 A quite clear community structure, represented by a high modularity, is however needed to really see the effect of this community structure on the formation of opinion clusters.
-
+\newpage
 \section{Impact of stubborn people}
@@
 This model is used to investigate the effect of clustering.
-	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
+	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
 This model is used to investigate the effect of a community structure.
@@
 	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
-	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
+	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$ to completely stubborn, $r = 1$).
 \end{itemize}
@@
 In all cases, the stubborn nodes are randomly distributed over the network.\\
-\newline
-Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm.
+
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_frac_complStubb}
+\end{figure}
+\noindent
+Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm and for an initial $50/50$ opinion distribution.
 The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.
@@
 Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
-Data points are again shown together with their standard error $SE$.
+The results are again obtained for an initial $50/50$ opinion distribution.
+Data points are shown together with their standard error $SE$.
 The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
@@
 \begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
-	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_frac_complStubb}
-\end{figure}
-
-\begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
@@
     \caption{REC}
+    \label{REC_frac_compl_stubb}
   \end{subfigure}
@@
 \end{figure}
+\noindent
+Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion).
+Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm.
+The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure.
+The interpretation of the $y$-axis is slightly different than before.
+In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+Thus we first normalized the echo chambers around opinion 0 and those around opinion 1 and than took the average of both.
+In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity.
+We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
+Hence, we chose for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
+Note that by taking the average, we neglect the skewed echo chamber distribution.
+This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn people and are less interested in the distribution of those echo chambers.
+\\
+\newline
+When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes.
+This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm as depicted in Figure \ref{echo_vs_frac_complStubb}.
+Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
+This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}.
+The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
+It is however not recommended to compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution.
+The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
+When comparing the different network models, these skewed distributions should not be overlooked.
+The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to compare the different models against each other.
+\\
+\newline
+% explain reason behavior REC!!
+The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, look again different than those for the PR filtering algorithm.
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn nodes, followed by a decrease.
+This result is very similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}.
+The stochastic block model with a high modularity doesn't really show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn nodes smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn agents increases further, we can start to see a decrease in the formation of echo chambers.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn agents (echo chamber size around $1.2$).
+This very limited, almost non-significant formation of echo chambers for the ER model and the SBM with a low modularity, in combination with the REC filtering algorithm, is also visible in Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
+For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
+We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
 
@@
     \caption{PR}
+    \label{PR_frac_compl_stubb_20-80}
   \end{subfigure}
@@
     \caption{REC}
+    \label{REC_frac_compl_stubb_20-80}
   \end{subfigure}
@@
 \end{figure}
-% explain figure + difference in y-ax
 
@@
 \end{figure}
+\noindent
+Figure \ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
+Data points are shown together with their standard error $SE$, which is very small and not really visible in the figure.
+Figure \ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Figure \ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
+The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+In case of the PR filtering algorithm (Figure \ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models for a stubbornness parameter less than $r \leqslant 0.4$.
+If the stubbornness parameter becomes bigger than $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
+This behavior is slightly different than in the case of an initial $50/50$ distribution, combined with the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
+In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if the stubbornness parameter became bigger than $r>0.1$).
+This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
+As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
+It should be noted that these types of figures are not really suited to compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+In order to perform such a comparison, one should look at the complete distribution (see eg.
+Figure \ref{echo_20_80}), as was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+\newline
+The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC method (see Figure \ref{REC_all_frac_stubb}).
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to a stubbornness parameter $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
+The increase seems, however, less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
+This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
+The stochastic block model with a high modularity also shows a slight increase in the formation of echo chambers up to $r = 0.8$.
+This increase can even be observed for the low modularity case, albeit less strong.
+The Erd\H{o}s-R\'{e}nyi network seems to wiggle a bit around a formation of echo chambers equal to one.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{PR}
+    \label{PR_all_frac_stubb_20-80}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_REC_20-80_8x7.png}
+    \caption{REC}
+    \label{REC_all_frac_stubb_20-80}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution.
+\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_all_frac_stubb_PR-REC_20-80}
+\end{figure}
 
@@
 The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
-The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.\\
+The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.
 % explain opinion dynamics or refer to relevant sections!!!
-\newline
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_threshold_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_threshold_PR}
+\end{figure}
+\noindent
 Figure \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
 In this figure, the PR filtering algorithm is used.
-Data points are now shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.\\
+Data points are shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.\\
 The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat, for $T \leqslant 0.5$.
@@
 \begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_threshold_PR_12x10.png}
-	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_threshold_PR}
-\end{figure}
-
-\begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
@@
 \noindent
-Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
+Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
 The initial opinion distribution was again $50/50$.
@@
 After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion 1 (and thus resulting in large echo chambers around opinion 1) and there will have been simulations that converge to opinion 0 (resulting in almost no echo chambers around opinion 1).\\
-Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
+Another consequence of this converging to one of the two opinions every single simulation, that can be observed in Figure \ref{echo_vs_threshold_REF-REC}, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
 This is not due to any network features, but a direct consequence of the usage of the majority model.
@@
 This can be clearly seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than when using the REC/REF methods.\\
-\newline
-Figure \ref{ev_op_majT_one_sim} illustrates this particular behavior of the majority model more clearly.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than when using the REC/REF methods, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
+If the threshold $T \geqslant 0.5$, it starts to have an impact on the system.
+However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold doesn't become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
+At this point, none of the nodes can be convinced of another opinion and hence keep their initial opinion.
+The number of echo chambers doesn't change over time, which results in an echo chamber size equal to one in the figure.\\
+\newline
+Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
 The figure shows the evolution of opinion 0 over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion 0; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion 0 as to opinion 1, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the results for each of the four network model: ER, WS, SBM (high and low modularity) and SBM-WS.
+For each model the REC and PR filtering algorithm are compared to each other.\\
+From this figure it becomes clear that the PR method strongly hampers the convergence to a single opinion each independent simulation.
+It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC method.
+This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
+Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Figure \ref{sbm_majT}.
+For the REC method, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
+For the PR method, this is not really the case, but this is probably due to the fact that the PR method already strongly hampers the convergence to a single opinion; having a high or low modularity does not really give an additional effect on top of this PR filtering effect.
 % explain figure!!

### 2021-04-16 draft.tex
latest update

@@
 In all cases, the stubborn nodes are randomly distributed over the network.\\
+We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.
 
@@
 \noindent
+So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn agents.
+But what about the evolution of the opinions over time?
+Are these affected by introducing stubborn people in the system?
+\\
+\newline
+In Figure \ref{ev_op_50_50_stubb_REC} the evolution of the fraction of nodes with opinion 0 for an initial $50/50$ opinion distribution is shown.
+Data points are shown together with the standard deviation $\sigma$.
+This figure shows the results for the REC filtering algorithm.
+The four network models (ER, WS, SBM (high and low modularity) and SBM-WS) are compared to each other.
+For each model we look at three different fractions of completely stubborn people: no stubborn people, fracRes $=0$; a fraction of completely stubborn people equal to fracRes $=0.4$ and a fraction of completely stubborn people equal to fracRes $=0.8$.
+Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\
+\newline
+Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm.
+Data points are again represented together with the standard deviation $\sigma$.
+Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models.
+The results don't differ from those obtained for the REC filtering algorithm.
+The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_stubb_REC}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_stubb_PR}
+\end{figure}
+\noindent
+Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution.
+Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg.
+an initial $20/80$ distribution.\\
+\newline
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{PR}
+    \label{PR_frac_compl_stubb_20-80}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{REC}
+    \label{REC_frac_compl_stubb_20-80}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
+\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_frac_complStubb_PR-REC_20-80}
+\end{figure}
+\noindent
 Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion).
@@
 For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
+\newpage
+\noindent
 We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
+\\
+\newline
+So much for the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
+Lets now discuss the evolution of the opinions in the system.
+In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC method does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion.
+Are these results affected by introducing stubborn agents?\\
+\newline
+Figure \ref{ev_op_20_80_stubb_REC} shows the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ distribution and when using the REC filtering algorithm.
+Data point are shown together with their standard deviation $\sigma$.
+Four network models are compared (ER, WS, SBM (high and low modularity) and SBM-WS).
+For each network model we compare three different fractions of completely stubborn nodes (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
+From this figure it becomes clear that introducing stubborn people in combination with the REC filtering algorithm is not able to break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+None of the network topologies, nor the stubborn agents can break the status-quo.
+This result is similar to the case of an initial $50/50$ opinion distribution in combination with a fraction of stubborn people, as discussed for Figure \ref{ev_op_50_50_stubb_REC}.\\
+\newline
+Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
+Data points are presented along with their standard deviation $\sigma$.
+Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
+If there are no stubborn people, the PR leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
+Introducing a fraction of stubborn agents changes this scenario.
+For an increasing fraction of stubborn people, the decrease of the minority opinion becomes less and less strong.
+This result is the same for all the different network topologies.
+If there are no stubborn people, the PR method forces nodes with the minority opinion to switch to the majority opinion.
+If we however introduce a fraction of stubborn people in the system, then we will also end up with some stubborn nodes that carry the minority opinion.
+These stubborn nodes with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
+The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{PR}
-    \label{PR_frac_compl_stubb_20-80}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{REC}
-    \label{REC_frac_compl_stubb_20-80}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_20-80_8x7.png}
+    \caption{SBM-WS}
   \end{subfigure}
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
-\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_stubb_REC}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_stubb_PR}
+\end{figure}
+% don't forget the evolution of opinions + also for 20/80, do each network, PR, REC, REF for eg.
+stubb 0.4 (so two fig (50-50 and 20-80) with 4 subplots (one for each network), each plot has PR, REF, REC) + if all stubb frac behaves same, explain in text, if not make extra plot!!
+\newpage
+\noindent
+\textbf{Making all nodes partly stubborn}\\
+\newline
+Instead of introducing a fraction of completely stubborn people in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so).
+All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
+\begin{figure}[H]
+	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
+	\captionsetup{format=plain}
+	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_frac_complStubb_PR-REC_20-80}
+\label{echo_vs_all_frac_stubb_PR}
 \end{figure}
-
-% don't forget the evolution of opinions + also for 20/80, do each network, PR, REC, REF for eg.
-stubb 0.4 (so two fig (50-50 and 20-80) with 4 subplots (one for each network), each plot has PR, REF, REC) + if all stubb frac behaves same, explain in text, if not make extra plot!!
 \noindent
-\textbf{Making all nodes partly stubborn}\\
-\newline
-All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
-\newline
 Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
@@
 If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
-In Figure \ref{echo_vs_all_frac_stubb_PR} this is represented by an echo chamber size equal to one for $r=1$.
-
-\begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
-	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_all_frac_stubb_PR}
-\end{figure}
-\newpage
-\noindent
+In Figure \ref{echo_vs_all_frac_stubb_PR} this is represented by an echo chamber size equal to one for $r=1$.\\
+\newline
 In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
@@
 \end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_all_frac_stubb_REC}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac__stubb_PR_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_all_frac_stubb_PR}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_all_frac_stubb_REC}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_all_frac_stubb_PR}
+\end{figure}
+% explain ev of op figures + make link to those with frac compl stubb
 

### 2021-04-17 draft.tex
plus minus finished stubborn people section in results

@@
 \end{itemize}
-For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
+For both cases we will look at both the evolution of the opinions and the formation of echo chambers and we will do this for two different initial conditions: a balanced initial opinion distribution ($50/50$) and an unbalanced initial opinion distribution ($20/80$).
+\\
 Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.
@@
 We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.
+Then we will discuss the formation of echo chambers and the evolution of the opinions for an initial $20/80$ opinion distribution.
 
@@
 \\
-\newline
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
+    \caption{REF}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
+    \caption{REC}
+    \label{REC_frac_compl_stubb}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
+\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_frac_complStubb_REF-REC}
+\end{figure}
+\noindent
 Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
@@
 This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions.
-This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
-    \caption{REF}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
-    \caption{REC}
-    \label{REC_frac_compl_stubb}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
-\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_frac_complStubb_REF-REC}
-\end{figure}
-\noindent
+This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.\\
+\newline
+This discrepancy between the results for the PR method and those for the REF/REC method in combination with a small fraction of stubborn nodes can be potentially explained with the help of the following simple example.
+Assume we have three nodes that form a triangle.
+Two nodes initially carry opinion 0 and one node carries opinion 1.\\
+Lets first discuss the PR method and assume that none of the nodes are stubborn.
+In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the nodes adopt the same opinion).
+This can be easily seen: the one node with opinion 1 is likely to switch to opinion 0, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR method wants to promote posts with opinion 1 does not matter here, since there are no neighbors with opinion 1).
+The two nodes with opinion 0 will have more chance of keeping that opinion than switching to opinion 1 in case of the PR method.
+This is because the PR method will promote the posts of the friend with the same opinion 0 over the posts of the friend with opinion 1.
+The end result will be that there is a high chance that all the three nodes will adopt opinion 0 and hence, form an echo chamber.
+Now assume that we make the node with opinion 1 stubborn.
+This node is then no longer capable of switching to opinion 0.
+However, the two nodes with opinion 0 still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion 0, and hence reduces the visibility of the posts of the stubborn friend with opinion 1.
+The system will now not form an echo chamber, since the stubborn node with opinion 1 cannot switch to opinion 0 and the PR method reduces the chance of changing the opinion of the nodes with opinion 0 to opinion 1.
+The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn people.\\
+What happens if we use the REF/REC filtering algorithm?
+Again we have one node with opinion 1 and two nodes with opinion 0 in the triangle.
+Assume first that none of the nodes are stubborn.
+For the PR method, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber.
+For the REF/REC method this chance is however smaller, which can be seen as follows.
+The node with opinion 1 still has a high chance of switching to opinion 0, because all his friends (which are the two other nodes in the triangle) carry this opinion.
+However, the two nodes with opinion 0 will now not have a higher chance of keeping this opinion than to switch to opinion 1, as was the case for the PR method.
+Instead, they have an equal chance of keeping their opinion as to change to the other opinion.
+This is because both the REF and REC filtering algorithms don't discriminate between the two different opinions of the friends of the nodes with opinion 0.
+Both opinion 0 and opinion 1 have an equal chance of being promoted on the time-line of the nodes with opinion 0.
+This reduced tendency of the nodes with opinion 0 to not change their opinion (compared to using the PR method) hampers the ability of the triangle to form an echo chamber.
+This would explain why, in case of no stubborn nodes, the REC/REF method shows a lower formation of echo chambers compared to the PR method.
+Now, lets again assume that we make the node with opinion 1 stubborn.
+In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle.
+For the REC/REF method, things are different.
+The stubborn node with opinion 1 will obviously not change opinion.
+The two nodes with opinion 0, on the other hand, will now have a higher chance of switching to opinion 1 (which was not the case when using the PR filtering algorithm).
+This can be seen as follows.
+The REF/REC filtering algorithms do not discriminate between the two different opinions of the two friends of the nodes with opinion 0.
+If opinion 1 is, however, stubborn it will keep on pushing that opinion on the time-lines of the two nodes with opinion 0.
+This will eventually lead to an increase of posts with opinion 1, compared to posts with opinion 0, on the time-lines of the two nodes with opinion 0.
+The result will then be that the nodes with opinion 0 will have a higher chance of switching to opinion 1, which enhances the formation of echo chambers in the triangle.
+This would explain the increase in the formation of echo chambers for a small fraction of stubborn people when using the REF or REC filtering algorithms.
+\\
+If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease again which can also be easily seen with the help of the triangle example.
+If the fraction of stubborn people becomes too high, we might end up with the case where both the node with opinion 1 and one of the nodes with opinion 0 are stubborn.
+This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three nodes in the triangle carry the same opinion.\\
+% make figure 
+\newline
 So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn agents.
@@
 Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\
-\newline
-Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm.
-Data points are again represented together with the standard deviation $\sigma$.
-Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models.
-The results don't differ from those obtained for the REC filtering algorithm.
-The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.
 
@@
 \noindent
+Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm.
+Data points are again represented together with the standard deviation $\sigma$.
+Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models.
+The results don't differ from those obtained for the REC filtering algorithm.
+The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
+\newline
 Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution.
@@
 We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
-Hence, we chose for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
+Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
 Note that by taking the average, we neglect the skewed echo chamber distribution.
@@
 When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes.
-This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm as depicted in Figure \ref{echo_vs_frac_complStubb}.
+This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}.
 Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
 This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}.
+\\
 The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
@@
 For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
-\newpage
-\noindent
+\newline
 We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
@@
 \newline
-So much for the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
+So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
 Lets now discuss the evolution of the opinions in the system.
@@
 For each network model we compare three different fractions of completely stubborn nodes (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
+\\
 From this figure it becomes clear that introducing stubborn people in combination with the REC filtering algorithm is not able to break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
@@
 Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
-If there are no stubborn people, the PR leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
+\\
+If there are no stubborn people, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
 Introducing a fraction of stubborn agents changes this scenario.
@@
 \end{figure}
-% don't forget the evolution of opinions + also for 20/80, do each network, PR, REC, REF for eg.
-stubb 0.4 (so two fig (50-50 and 20-80) with 4 subplots (one for each network), each plot has PR, REF, REC) + if all stubb frac behaves same, explain in text, if not make extra plot!!
+
 \newpage
@@
 All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
+We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
+
 \begin{figure}[H]
@@
 This is the complete opposite of the case where the PR method is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
-% REASON??
-\\
+The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the "Making a fraction of the nodes completely stubborn" paragraph.\\
 If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.
 The increase is however not as strong as for the high clustering networks.\\
+\newline
 The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) don't seem to have any dependency on the stubbornness level of the nodes.
@@
 \noindent
+What about the results for the evolution of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
+Again, we look at the four different network models (ER, WS, SBM (high and low modularity) and SBM-WS).
+For each network model, the results for three different values of the stubbornness parameter $r$ of all the nodes are compared: the nodes are not stubborn, $r=0$; the nodes are partly stubborn with $r = 0.4$ and the nodes are partly stubborn with $r=0.8$.
+Two different filtering algorithms are investigated, namely the REC filtering algorithm and the PR one.\\
+\newline
+Figure \ref{ev_op_50_50_all_frac_stubb_REC} shows the results for the evolution of the fraction of nodes with opinion 0 over time in case of using the REC filtering algorithm.
+From this figure, it becomes clear that making the nodes partly stubborn is not able to break the status-quo.
+The prevalence of the initial opinions is maintained for all the network models and for all the values of the stubbornness parameter $r$.
+This result is exactly the same as the result we obtained for the case in which a fraction of completely stubborn nodes was introduced in the system (in case of an initial $50/50$ opinion distribution and in combination with the REC filtering algorithm), see Figure \ref{ev_op_50_50_stubb_REC}.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_all_frac_stubb_REC}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac__stubb_PR_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_50_50_all_frac_stubb_PR}
+\end{figure}
+\noindent
+Figure \ref{ev_op_50_50_all_frac_stubb_PR} shows the results for the evolution of the fraction of nodes with opinion 0 over time in case of an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+Also for the PR filtering algorithm, the status-quo is maintained for all the network models and for all the values of the stubbornness parameter $r$.
+This result is again the same as the case where we introduced a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution and using the PR filtering algorithm, see Figure \ref{ev_op_50_50_stubb_PR}.\\
+\newline
+Lets now look at the results for an unbalanced initial opinion distribution, eg.
+$20/80$.
+Opinion 0 is again the minority opinion, which makes opinion 1, obviously, the majority opinion.
+First, we will discuss the formation of echo chambers in the system under unbalanced initial opinion conditions and when making all the nodes partly stubborn.
+Then, we will look into the results for the evolution of the opinions with time.\\
+\newline
 Figure \ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
@@
 The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+\newline
 In case of the PR filtering algorithm (Figure \ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models for a stubbornness parameter less than $r \leqslant 0.4$.
@@
 \end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_all_frac_stubb_REC}
-\end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac__stubb_PR_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_all_frac_stubb_PR}
-\end{figure}
-
+\noindent
+We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn.
+Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter r ($r=0$, $r=0.4$ and $r=0.8$).
+Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
+\newline
+Figure \ref{ev_op_20_80_all_frac_stubb_REC} gives the results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm.
+For none of the network models, nor the values of the stubbornness parameter $r$, the initial prevalence of the opinions can be broken.
+This result is analogue to the result for an initial $20/80$ opinion distribution, combined with the REC filtering algorithm, in the case of introducing a fraction of completely stubborn people (see Figure \ref{ev_op_20_80_stubb_REC}).
+
+\newpage
 \begin{figure}[H]
@@
 \end{figure}
-% explain ev of op figures + make link to those with frac compl stubb
+\noindent
+The results for the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
+If the nodes are not stubborn, the PR method strongly reduces the minority opinion over time.
+Making the nodes more and more stubborn reduces this tendency.
+This result is more or less similar to the corresponding case in which a fraction of completely stubborn nodes was introduced in the system, see Figure \ref{ev_op_20_80_stubb_PR}.
+The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the minority opinion compared to making a fraction of the nodes completely stubborn.
+For a stubbornness parameter equal to $r=0.4$, the decrease of the minority opinion is almost not hampered, it is merely slowed down.
+All the nodes need to be really stubborn ($r=0.8$) to really see a hampering in the decrease of the minority opinion.
+Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn people equal to fracRes $=0.4$ was introduced (a final fraction  of the minority opinion equal to 0.07 versus a final fraction of the minority opinion equal to 0.1).
 
@@
 The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
-The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.
+The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.\\
+For this model we will only investigate the case of balanced initial conditions, so an initial $50/50$ opinion distribution.
+We will mainly focus on the formation of echo chambers versus the threshold parameter $T$ and we will compare three different filtering algorithms: the PR, REF and REC filtering algorithm.
+\\
 % explain opinion dynamics or refer to relevant sections!!!
+\newline
+Figure \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
+In this figure, the PR filtering algorithm is used.
+Data points are shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.
+The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.\\
+The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat for a threshold parameter smaller than $T \leqslant 0.5$.
+This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
+This is because the majority opinion of the neighbors of a node will always be larger than the threshold if the threshold is smaller than $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
+If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an effect.
+The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
+Similar observations can be made for the stochastic block model with a high modularity.
+The only difference is of a quantitative nature (the formation of echo chambers is not as large as for the high clustering networks).\\
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points lie behind those of the Erd\H{o}s-R\'{e}nyi network) do not enhance the formation of echo chambers in any way.
+Including a threshold $T$ does not change this tendency.
+% make coupling to ev of op??
+\newpage
 \begin{figure}[H]
@@
 \label{echo_vs_threshold_PR}
-\end{figure}
-\noindent
-Figure \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
-In this figure, the PR filtering algorithm is used.
-Data points are shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.\\
-The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat, for $T \leqslant 0.5$.
-This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
-The majority opinion will always be larger than this threshold (the sum of the two opinion fractions is always equal to one).
-If $T \geqslant 0.5$, the threshold start to have an effect.
-The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
-Similar observations can be made for the stochastic block model with a high modularity.
-The only difference is of a quantitative nature (the formation of echo chambers is not as large as for the high clustering networks).\\
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points lie behind those of the Erd\H{o}s-R\'{e}nyi network) do not enhance the formation of echo chambers in any way.
-Including a threshold $T$ does not change this tendency.
-% make coupling to ev of op??
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REF_8x7.png}
-    \caption{REF}
-    \label{REF_thres}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REC_8x7.png}
-    \caption{REC}
-    \label{REC_thres}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
-\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_threshold_REF-REC}
 \end{figure}
@@
 This can be clearly seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than when using the REC/REF methods, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
-If the threshold $T \geqslant 0.5$, it starts to have an impact on the system.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF methods, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
+If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an impact on the system.
 However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold doesn't become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
@@
 The figure shows the results for each of the four network model: ER, WS, SBM (high and low modularity) and SBM-WS.
-For each model the REC and PR filtering algorithm are compared to each other.\\
+For each model the REC and PR filtering algorithm are compared to each other.
+The threshold parameter is set equal to $T=0$, so there is no effect of any type of stubbornness.\\
 From this figure it becomes clear that the PR method strongly hampers the convergence to a single opinion each independent simulation.
@@
 For the PR method, this is not really the case, but this is probably due to the fact that the PR method already strongly hampers the convergence to a single opinion; having a high or low modularity does not really give an additional effect on top of this PR filtering effect.
-% explain figure!!
-
-% make link to figures op frac+ hampering PR + network structures!!
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REF_8x7.png}
+    \caption{REF}
+    \label{REF_thres}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REC_8x7.png}
+    \caption{REC}
+    \label{REC_thres}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Each data point is the result of $10^2$ independent simulations.}
+\label{echo_vs_threshold_REF-REC}
+\end{figure}
 

### 2021-04-17 draft.tex
plus minus finished stubborn people section in results

@@
 \end{itemize}
-For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.\\
+For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.
+Two filtering algorithms will be compared to each other, namely the REC and the PR filtering algorithm, in order to investigate the possible interplay between those filtering algorithms and the network structures.\\
 \newline
@@
 \newline
-Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of opinion 0 over time, together with the standard deviation $\sigma$.
+Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of nodes with opinion 0, $P_0(t)$, over time, together with the standard deviation $\sigma$.
 Two different filtering algorithms are compared to each other, namely the PR method and the REC method (see Chapter \ref{chap3}, Section \ref{filter} for a detailed explanation).
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
@@
 These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\
+\newline
 In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR method.
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
@@
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
-Four different network models are used (each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$): \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.}

### 2021-04-18 draft.tex
latest update

@@
 
-\section{Impact of network structure}
+\section{Impact of network structure}\label{netw_struc}
 
@@
 
-\subsection{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}
+\subsection{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}\label{20-80}
 
@@
 \end{figure}
-
+\newpage
 \section{Random versus community distributed opinions/stubborn agents}
+
+This section is devoted to results regarding the way the opinions and/or the stubborn people are distributed over the network.
+In all the previous results, the two opinions and the stubborn people (in case there were stubborn people introduced in the system) were, initially, always randomly distributed across the system.
+It might however be more realistic if people with the same opinion are more or less grouped together initially.
+We did introduce a network model with a community structure (the SBM) to represent the group structures found in real-life systems.
+But we haven't yet distributed the opinions according to those groups.
+In subsection \ref{groupVSrandom} we will discuss some results for distributing the opinions in communities versus randomly distributing them.
+
+\subsection{Distributing opinions in communities versus a random opinion distribution}\label{groupVSrandom}
+
+In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion 0.
+The remaining communities are given a $50/50$ opinion distribution.
+This situation may, in an idealized way, represent the situation in which you have groups of people following the same idea together with groups where the two ideas coexist.
+The total opinion distribution over the entire network is no longer $50/50$, but instead opinion 0 becomes the majority opinion (hence, opinion 1 becomes the minority opinion).
+The higher the fraction of single opinion 0 communities, the bigger the difference between this majority and minority opinion becomes.
+The results will always be compared to the corresponding case in which the total opinion distribution is randomly given to each community.
+Thus, eg.
+if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion 0 communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$.
+This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
+We will look at a network with a high modularity, representing the case of more or less isolated communities, and at a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
+More specifically, the two stochastic block models that are implemented are (both models have a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
+\begin{itemize}
+	\item A stochastic block model with 10 communities of size 100.
+The edge probability inside the communities is $p_{\text{cl}} = 0.1$ and the edge probability between communities is $p_{\text{add}} = 0.001$.
+This is the high modularity network.
+	\item A stochastic block model with 10 communities of size 100.
+The edge probability inside the communities is $p_{\text{cl}} = 0.03$ and the edge probability between communities is $p_{\text{add}} = 0.008$.
+This is the low modularity network.
+\end{itemize} 
+The opinion dynamics model that is implemented is described in Chapter \ref{chap4}, Section \ref{opinionmodel}.
+The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line is $S=20$.
+% look both at PR and REC??
+\\
+Both the results for the evolution of the opinions in the system and the formation of echo chambers will be discussed.\\
+\newpage
+\noindent
+\textbf{Evolution of the opinion fractions}\\
+\newline
+Lets first discuss how the evolution of the opinions are (potentially) impacted by distributing the opinions in communities instead of randomly distributing them.\\
+\newline
+Figure \ref{commOp0_01_other_50-50} shows the results for the fraction of nodes with opinion 0 versus time for the case in which a fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the other communities get an initial $50/50$ opinion distribution.
+The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in case of a fraction of $0.1$ single opinion 0 communities).
+The figure shows the results for both the high and the low modularity case.
+For each modularity the REC and PR filtering algorithms are compared to each other.\\
+In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
+The prevalence of the majority opinion 0 remains unchanged over time.
+This result is the same for both the high and low modularity case, see Figures \ref{high_mod_rec} and \ref{low_mod_rec}.\\
+The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced initial opinions, if the opinions are randomly distributed.
+This is a result that was already found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
+Figure \ref{ev_op_20_80}.
+% continue: talk about how comm distr affect this result + compare to low mod + add figures/data to correct Figure file
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_high_mod_8x7.png}
+    \caption{High modularity, REC}
+    \label{high_mod_rec}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_high_mod_8x7.png}
+    \caption{High modularity, PR}
+    \label{high_mod_pr}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_low_mod_8x7.png}
+   \caption{Low modularity, REC}
+    \label{low_mod_rec}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_low_mod_8x7.png}
+    \caption{Low modularity, PR}
+    \label{low_mod_pr}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.
+The stochastic block model is implemented with two different modularities.
+The average degree is always around $\left<k\right> \sim 10$ and the total number of nodes is equal to $N = 10^3$.
+\textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
+Each plot is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
+\label{commOp0_01_other_50-50}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_REC_high_mod_8x7.png}
+    \caption{High modularity, REC}
+    \label{high_mod_rec}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_PR_high_mod_8x7.png}
+    \caption{High modularity, PR}
+    \label{high_mod_pr}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_REC_low_mod_8x7.png}
+   \caption{Low modularity, REC}
+    \label{low_mod_rec}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_PR_low_mod_8x7.png}
+    \caption{Low modularity, PR}
+    \label{low_mod_pr}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.5$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($75/25$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.5$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($75/25$ in this case) is randomly distributed over each community.
+The stochastic block model is implemented with two different modularities.
+The average degree is always around $\left<k\right> \sim 10$ and the total number of nodes is equal to $N = 10^3$.
+\textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
+Each plot is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
+\label{commOp0_05_other_50-50}
+\end{figure}
+\noindent
+\textbf{Echo chambers}\\
+\newline
+njd
 

### 2021-04-20 draft.tex
latest update

@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
 The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$.
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 	\captionsetup{format=plain}
-	\caption[Relation between the formation of echo chambers and community structure (modularity)]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
+	\caption[Relation between the formation of echo chambers and community structure (modularity).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
 \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution.
 \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.
-Results for one simulation of the majority threshold model with $T=0$ ]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
@@
 Figure \ref{ev_op_20_80}.
-% continue: talk about how comm distr affect this result + compare to low mod + add figures/data to correct Figure file
+If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion 0, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
+This makes sense, since the single opinion communities are more or less isolated from the other communities.
+This has as consequence that the other (mixed opinion) communities live under the assumption that the opinion distribution is $50/50$.
+They don't really feel that opinion 0 is a majority opinion, since they are isolated from these single opinion 0 communities.
+In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer stand.
+This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present.
+The community distributed case behaves equivalent to the random distributed case.
+In the low modularity network, the community structure is not clearly present, but becomes blended.
+The different communities talk a lot to each other and there is a lot of communication between them.
+As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion 0 in the same way as if the opinions would be randomly distributed over all the communities.
 
@@
 \end{figure}
+\noindent
+The results found in Figure \ref{commOp0_01_other_50-50} for a fraction of $0.1$ of single opinion 0 communities remain valid if we increase the fraction of single opinion 0 communities.
+This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average fraction of opinion 0 before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion 0 communities.
+The error bars in this figure are omitted for the sake of an improved visualization.
+The community distributed cases are always compared to the corresponding randomly distributed cases.
+Results are shown for both a high modularity network and a low modularity network.
+Figure \ref{rec_comm} shows the results for the REC filtering algorithm.
+Figure \ref{pr_comm} shows those for the PR filtering algorithm.\\
+The higher the fraction of single opinion 0 communities, the higher the initial average fraction of opinion 0.
+This can be seen in both figures by the increase of the solid black line, which represents the initial fraction of nodes with opinion 0.
+In case of the REC filtering algorithm, Figure \ref{rec_comm}, the fraction of opinion 0 remains the same after the time evolution for a fixed fraction of single opinion 0 communities.
+Giving a fraction of the communities a single opinion, opinion 0, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
+This result is valid for both the high and low modularity case.\\
+The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the fraction of opinion 0, which is the majority opinion, over time for a fixed fraction of single opinion 0 communities.
+Making a fraction of the communities single opinion 0 communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
+The low modularity network does not show a sign of this hampering.
+The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was already explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion 0 communities can influence the $50/50$ communities.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_REC_high_mod_8x7.png}
-    \caption{High modularity, REC}
-    \label{high_mod_rec}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_PR_high_mod_8x7.png}
-    \caption{High modularity, PR}
-    \label{high_mod_pr}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_REC_low_mod_8x7.png}
-   \caption{Low modularity, REC}
-    \label{low_mod_rec}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=05_other=50-50_PR_low_mod_8x7.png}
-    \caption{Low modularity, PR}
-    \label{low_mod_pr}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_REC_8x7.png}
+    \caption{REC}
+    \label{rec_comm}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_PR_8x7.png}
+    \caption{PR}
+    \label{pr_comm}
   \end{subfigure}
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.5$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($75/25$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.5$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($75/25$ in this case) is randomly distributed over each community.
-The stochastic block model is implemented with two different modularities.
-The average degree is always around $\left<k\right> \sim 10$ and the total number of nodes is equal to $N = 10^3$.
-\textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
-Each plot is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
-\label{commOp0_05_other_50-50}
+  \caption[Average fraction of opinion 0 before and after the time evolution versus the fraction of single opinion 0 communities.
+Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity network.
+Results for the REC and PR filtering algorithm.]{Average fraction of opinion 0 before and after the time evolution versus the fraction of single opinion 0 communities.
+Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
+\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+Each plot is the result of $10^2$ independent simulations.}
+\label{fracOp0_vs_commOp0_other_50-50}
 \end{figure}
+\newpage
 \noindent
@@
 \newline
-njd
+Figure \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion 0 community while the other communities get an initial $50/50$ opinion distribution.
+As in similar plots before, the value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
+The community distributed case is compared to the corresponding randomly distributed case (which is, in this case, an initial $55/45$ opinion distribution that is randomly distributed in each community).
+The formation of echo chambers in both the high and low modularity network are investigated.
+Figure \ref{rec_comm_echo} shows the results for the REC filtering algorithm, while Figure \ref{pr_comm_echo} shows those for the PR filtering algorithm.
+Error bars are again omitted for an improved visualization.\\
+\newline
+Lets first discuss the results for the REC method.
+The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0, with consequently an increase in more mixed neighborhoods.
+The number of neighborhoods where all the nodes carry opinion 0 thus decreases in favor of neighborhoods where we find both nodes with opinion 0 and nodes with opinion 1.
+Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion 0 communities, reduces and that the opinions become more mixed.
+The fraction of opinion 0 remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion 0 see a decrease in the nodes with opinion 0 and become more mixed.
+We can also observe a very small increase in the number of echo chambers around the minority opinion 1, which indicates the fact that this minority opinion starts to group more together.
+The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion 0 and opinion 1.
+Again, the opinion fractions remain constant over time, but are shifted from a random distribution to a more clustered situation.
+The increase remains however smaller than two, which indicates that there are less than twice as much echo chambers at $t=500$ than at $t=0$.\\
+The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion 0, which could be observed for the community distributed, high modularity case.
+This is due to the fact that the number of echo chambers around opinion 0 at $t=0$, originating from the fraction of single opinion 0 communities, is a lot smaller for the low modularity network than for the high modularity network.
+The reason for this is the much more blended community structure for the low modularity network.
+We can still observe a very small increase in the number of echo chambers around opinion 1.
+This increase in the formation of echo chambers around the minority opinion 1 seems however hampered compared to the randomly distributed case.
+Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80} (see Figure \ref{sbm}).
+\\
+\newline    % make link to opinion fraction 
+The results for the PR filtering algorithm are depicted in Figure \ref{pr_comm_echo}.
+They are quite different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Figure \ref{rec_comm_echo}.\\
+Lets first focus on the high modularity network.
+The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (note that these results were obtained for an initial $20/80$ opinion distribution).
+We can observe a strong increase in the formation of echo chambers around the majority opinion 0, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion 1.
+Remember from Figure \ref{high_mod_pr} that the PR method leads to an increase in the majority opinion 0.
+If we, however, make a fraction equal to $0.1$ of the communities single opinion 0 communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion 0 is strongly hampered.
+This hampering is partly an artifact of the way the opinions are distributed.
+If we give entire communities only opinion 0 while others get a $50/50$ distribution, we have a large number of echo chambers around opinion 0 at $t=0$ due to these single opinion communities.
+Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
+If we, however, would look at the absolute number of echo chambers around opinion 0 at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
+This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0.
+% share data??
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_REC_8x7.png}
+    \caption{REC}
+    \label{rec_comm_echo}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_PR_8x7.png}
+    \caption{PR}
+    \label{pr_comm_echo}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
+Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
+\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+Each plot is the result of $10^2$ independent simulations.}
+\label{echo_01_commOp0_other_50-50}
+\end{figure}
 

### 2021-04-21 draft.tex
continued writing results section

@@
 \\
+% partly rewrite this according to results? Convergence to one opinion in communities --> discussed? Clustered stubborn people --> discussed? Include clustered opinions? Include real-world networks?
 
@@
 The data points represent the value of the echo chamber size in networks with a different modularity.
-The initial opinion distribution was always $50/50$.
+The initial opinion distribution was always $50/50$ an was, initially, randomly distributed over the network.
 \\
@@
 We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn.
-Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter r ($r=0$, $r=0.4$ and $r=0.8$).
+Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$).
 Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
@@
 The higher the fraction of single opinion 0 communities, the bigger the difference between this majority and minority opinion becomes.
-The results will always be compared to the corresponding case in which the total opinion distribution is randomly given to each community.
+The results will always be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
 Thus, eg.
@@
 They don't really feel that opinion 0 is a majority opinion, since they are isolated from these single opinion 0 communities.
-In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer stand.
+In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
 This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present.
@@
 Lets first discuss the results for the REC method.
-The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0, with consequently an increase in more mixed neighborhoods.
+The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0 (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line).
 The number of neighborhoods where all the nodes carry opinion 0 thus decreases in favor of neighborhoods where we find both nodes with opinion 0 and nodes with opinion 1.
@@
 This increase in the formation of echo chambers around the minority opinion 1 seems however hampered compared to the randomly distributed case.
-Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80} (see Figure \ref{sbm}).
-\\
-\newline    % make link to opinion fraction 
+Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80}, see Figure \ref{sbm}.
+\\
+\newpage
+\noindent    % make link to opinion fraction 
 The results for the PR filtering algorithm are depicted in Figure \ref{pr_comm_echo}.
@@
 Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
-If we, however, would look at the absolute number of echo chambers around opinion 0 at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
+If we, however, look at the absolute number of echo chambers around opinion 0 at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case.
 This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0.
-% share data??
+Beside this hampering of echo chambers around the majority opinion 0, the community distributed case also shows an increase of echo chambers around the minority opinion 1.
+Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion 0 (and thus the decrease of the minority opinion 1).
+Hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
+The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network.
+Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion 0 and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion 1.
+From Figure \ref{low_mod_pr}, it was already made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of this minority opinion (when using the PR filtering algorithm).
+This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry this minority opinion.
+The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case.
+However, as for the high modularity case, if we look at the absolute number of echo chambers around the majority opinion at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case.
+% make conclusion about this ?
++ make table with data??
 \begin{figure}[H]
@@
 \end{figure}
+\noindent
+Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, but now a fraction equal to $0.5$ of the communities is only given opinion 0, while the others get a $50/50$ distribution.
+The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion 0 (Figure \ref{echo_01_commOp0_other_50-50}).\\
+\newline
+The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion 0 with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
+A difference with the case where a fraction equal to $0.1$ of the communities are single opinion 0 communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion 1, but instead we observe a decrease in the number of echo chambers around the minority opinion.
+For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion 0 and a minority of the nodes carrying the minority opinion 1.
+\\
+The low modularity, community distributed case seems to wiggle around $1$.
+If we, however, look closely, we can see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller).
+We can observe a small decrease in the formation of echo chambers around the majority opinion 0, with a small increase in more mixed opinion neighborhoods as result.
+We can also observe a small decrease in the number of echo chambers around the minority opinion 0.
+The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion 0 communities, while the others get a $50/50$ opinion distribution.\\
+The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Figure \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
+They show an increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
+\newline
+Results for the PR filtering algorithm also strongly resemble those depicted in Figure \ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion 0 communities.
+The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Figure \ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}).
+They show a strong increase in the formation of echo chambers around the majority opinion 0.\\
+The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion 0, which was also observed in Figure \ref{pr_comm_echo}.\\
+The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion 0.
+This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion 0, hampers the increase in this majority opinion compared to the randomly distributed case (see Figure \ref{high_mod_pr} and Figure \ref{pr_comm}).
+The increase in the number of echo chambers around the minority opinion 0 that was visible in Figure \ref{pr_comm_echo} is no longer present.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
+    \caption{REC}
+    \label{rec_comm05_echo}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
+    \caption{PR}
+    \label{pr_comm05_echo}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
+Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
+\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+Each plot is the result of $10^2$ independent simulations.}
+\label{echo_05_commOp0_other_50-50}
+\end{figure}
 

### 2021-04-22 draft.tex
tweaked some small things

@@
 \usepackage{sectsty}
+\allsectionsfont{\boldmath}
 
@@
 Social media allows for the spread of information at a faster pace and at an unprecedented scale.
-However, people are still bounded by time- and cognitive constraints.
+However, people are still bound by time- and cognitive constraints.
 In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to.
-This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in "echo chambers" that confirm their own beliefs.
+This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
 It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
@@
 
-% definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
+% definetely not finished!! --> add some type of conclusions or not? Change part about distributing stubborn people??
 \end{abstract}
@@
 \selectlanguage{dutch}
+\newgeometry{top=0mm, bottom=15mm}
 \begin{abstract}
 \addcontentsline{toc}{chapter}{Samenvatting}
-% definetely not finished!!
+Sociale media zijn bijna niet meer weg te denken uit ons dagelijks leven.
+De laatste tientallen jaren is het gebruik van sociale media, zowel op persoonlijk vlak als op professioneel vlak, in snel tempo toegenomen.
+Tegenwoordig kan men zeggen dat (bijna) iedereen, naast zijn werkelijke leven, een virtueel leven op sociale media leidt.
+Het gebruik van sociale media heeft ongetwijfeld veel voordelen.
+Goed en kwaad gaan echter vaak hand in hand en men kan niet ontkennen dat het gebruik van sociale media ook veel potentiële, soms ongekende, nadelen kent.
+Eén van die, tot nog toe, ongekende kanten van sociale media is de impact die het heeft op de dynamica van opinies in de maatschappij.
+\\
+Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloedt worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,...
+Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen.
+Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\
+Door het toenemende gebruik van sociale media worden deze echter overspoeld door informatie.
+Om hun gebruikers niet het gevoel te geven dat ze verdrinken in wat ze te zien krijgen op hun sociale media platform, gebruiken deze sociale media platformen  filtering algoritmen.
+Deze filtering algoritmen filteren, zoals de naam al impliceert, en individualiseren de informatie die iemand te zien krijgt op zijn/haar tijdslijn.
+Een gebruiker op sociale media krijgt dus niet alle berichten die de vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn/haar sociale media geschiedenis.
+Dit filteren/individualiseren kan, mogelijks, grote gevolgen hebben.
+Zo zou het kunnen leiden tot de vorming van  ``echo kamers'' waar gebruikers gevangen raken tussen ideeën die hun eigen visie versterken.
+Dit zou grote gevolgen kunnen hebben op bijvoorbeeld polarisatie in de maatschappij, opkomst van extremisme, enzovoort.
+Het is daarom uiterst belangrijk om de impact van deze filtering algoritmen grondig in kaart te brengen.\\
+Mensen zijn echter ook in staat om kritisch te denken, soms ook over-kritisch of selectief kritisch, maar dat terzijde.
+Dit kritisch denken kan ertoe leiden dat mensen niet zomaar hun mening veranderen omdat hun vrienden en/of omgeving dat suggereren.
+Ze zijn meer overtuigd van hun eigen gelijk of menen dat hun eigen visie een correcter standpunt inneemt.
+Deze individuele koppigheid zou men niet mogen verwaarlozen in het ontwerpen van modellen die de dynamica van opinies op sociale media willen onderzoeken.\\
+In deze thesis wordt gebruik gemaakt van ``toy models'' die de interactie tussen verschillende netwerk modellen, filtering algoritmen en individuele terughoudenheid om van mening te veranderen onderzoeken in de evolutie van twee opinies.
+Er worden verschillende netwerk modellen geïntroduceerd die de impact van onder andere clustering en/of groepsstructuur op de evolutie van opinies en de vorming van echo kamers moeten onderzoeken.
+Daarnaast wordt er gebruik gemaakt van enkel simpele filtering algoritmen om hun impact op de dynamica van de opinies te onderzoeken.
+Tot slot wordt er een notie van koppigheid in het systeem geïntroduceerd om mogelijke effecten hiervan in kaart te kunnen brengen.
+\\
+Naast het bovenstaande zal er ook kort aandacht gespendeerd worden aan hoe de twee verschillende opinies verdeeld worden over het systeem.
+Is er bijvoorbeeld een verschil tussen de twee meningen willekeurig over het systeem te verspreiden versus deze in een meer groepsstructuur te verdelen?
+Om de thesis te besluiten zal het opinie-dynamica model, dat in de thesis gebruikt werd, getest worden op netwerken verkregen van data uit de werkelijke wereld (bv.
+van email data of telefoonnetwerken).
+Er zal nagegaan worden of de resultaten van deze werkelijke netwerken gelijkenissen, dan wel verschillen, vertonen met de data verkregen van de netwerk modellen.
 \end{abstract}
 
+\restoregeometry
 \selectlanguage{english}
+\newgeometry{top=25mm, bottom=25mm}
 \tableofcontents
@@
 \addcontentsline{toc}{chapter}{List of Figures}
-
-% DON'T FORGET: popular text somewhere!!
+\restoregeometry
 
@@
 
-% add research question and hypotheses!!
 % ADD MORE REFERENCES (previous research, contextualization etc)
@@
 Complex systems often display phenomena such as non-linearity, emergence, spontaneous order,...
-%(maybe check cursus compl en crit?)
 \\
@@
 The other two are short distances and high clustering.
-These two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
+These two are also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
-% to do: read links 19, 21, 22, 23, 24 of https://arxiv.org/pdf/0711.3199.pdf
 
@@
 In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
-
-% biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9
 
@@
 \newline
-In one of their previous papers (\textit{"Modelling opinion dynamics in the age of algorithmic personalization"}, see reference \cite{Perra2019}) prof.
+In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}) prof.
 Nicola Perra and prof.
@@
 \\
-% ALSO mention real-world network??
 \newline
@@
 % partly rewrite this according to results? Convergence to one opinion in communities --> discussed? Clustered stubborn people --> discussed? Include clustered opinions? Include real-world networks?
-
-% talk about goals and hypotheses, what does this thesis wants to investigate, reference paper Rocha --> what did they do, their findings, how does this thesis build on that NOT FINAL YET --> maybe talk about fact that you play around with distributing the opinions over the different communities?
-+ talk about investigating real networks??
-% give reference to sections were community structure etc are explained! (done?)
 
@@
 From the degree of the nodes in the network one can construct the degree distribution.
-The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$.
+The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$ \cite{Newman2003}.
 The average degree can be obtained from the degree distribution \cite{Mata2020}
@@
 % finished?
-\newpage
+
 \subsection{Clustering coefficient and transitivity}
@@
 The clustering coefficient is thus a way to measure the degree to which nodes in a network tend to cluster \cite{Li2017}.
-The local clustering coefficient of a node $i$ is defined as
+The local clustering coefficient of a node $i$ is defined as \cite{Mata2020}
 \begin{equation}\label{clus}
@@
 % should give more mathimatical expression for n_i?
-The average clustering coefficient is then given as the average of the local clustering coefficients
+The average clustering coefficient is then given as the average of the local clustering coefficients \cite{Newman2003}
 \begin{equation}\label{avClus}
@@
 \newline
-A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as (valid for undirected, unweighted networks) \cite{F.Costa2007}
+A measure that is closely related to the clustering coefficient is the transitivity $T$, defined as (valid for undirected, unweighted networks) \cite{F.Costa2007}\cite{Newman2003}
 \begin{equation}\label{globalTrans}
@@
 \newline
-A connected triple is defined as a set of three nodes with at least two edges between them, so that each node can be reached from the other two (either directly or indirectly).
+A connected triple is defined as a set of three nodes with at least two edges between them, so that each node can be reached from the other two (either directly or indirectly) \cite{Newman2003}.
 The factor three arises from the fact that each triangle contributes to three different connected triples in the network: one centered at each node in the triangle \cite{F.Costa2007}.
 \\
-\newline
 Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
@@
 In the case of the average clustering coefficient $\left<C\right>$ and the global transitivity $T$ this is, however, not the case.
-The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node.
+The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node \cite{Newman2003}.
 This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.
@@
 Several community definitions are consistent with the density hypothesis.
-Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
+Let us consider three possible definitions: maximum cliques, strong communities and weak communities \cite{Albert2016}.\\
 \newline
@@
 \newline
-A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}.
+A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}\cite{F.Costa2007}.
 Let us denote a community as $C$ and let $i$ be a node inside the community $C$.
@@
 \newline
-A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}.
+A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}\cite{F.Costa2007}.
 We thus have the following condition for a weak community $C$ \cite{Albert2016}
@@
 % explain small-world (milgram), high clustering, power-law
-\newpage
+
 \section{Network models}
@@
 Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
-Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.
+Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the ``small world phenomenon''; the other property is a high clustering coefficient \cite{Easley2010}.
 \\
@@
 To summarize: the random Erd\H{o}s-R\'{e}nyi network has a bell shaped degree distribution which approximates a Poisson distribution for $N \gg 1$; the average clustering coefficient becomes small for large networks and the average distance depends logarithmic on the system sizes.
-The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the "small world phenomenon" (namely short average distance).\\
-Real social networks, however, usually display the "small world phenomenon" and thus have both a short average distance and a high clustering coefficient.
+The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the ``small world phenomenon'' (namely short average distance).\\
+Real social networks, however, usually display the ``small world phenomenon'' and thus have both a short average distance and a high clustering coefficient.
 The question then rises: can we construct a model that has both of these properties?
@@
 However, a network with a skewed degree distribution violates this assumption of edge independence \cite{Clauset2017}.
-To say it in the words of Karrer and Newman: "Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...]." (Karrer and Newman, p.1, reference \cite{Karrer2011}).
+To say it in the words of Karrer and Newman: ``Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...].'' (Karrer and Newman, p.1, reference \cite{Karrer2011}).
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
@@
 Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
-\newline
 As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties.
-These are the two properties of the "small world phenomenon", namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
+These are the two properties of the ``small world phenomenon'', namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
 These properties are valid for both off-line and on-line social networks \cite{Zhang2014}.
@@
 \newline
-In the late 1960's Stanley Milgram performed a, what is now famous, experiment.
+In the late 1960's Stanley Milgram performed a, what is now famous, experiment \cite{Newman2003}.
 He asked randomly selected people in Boston and Omaha to forward a letter to a distant target person.
@@
 The remarkable result of this experiment was that the average number of steps between the sender and the target person was only around six.
-This phenomenon is now often referred to as "six degrees of separation" \cite{RealWorld}.\\
+This phenomenon is now often referred to as ``six degrees of separation'' \cite{RealWorld}.\\
 This phenomenon is also observed in many other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
-Such networks usually have an average distance that depends logarithmically on the system size instead of linearly.
-A short average distance facilitates a fast transmission of information \cite{Zhang2014}.\\
+Such networks usually have an average distance that depends logarithmically on the system size instead of linearly \cite{Newman2003}.
+A short average distance facilitates a fast transmission of information \cite{Newman2003}\cite{Zhang2014}.\\
 \newline
@@
 where now the weights are time dependent.
-This model can be used for situations in which so called "hardening of positions" occur \cite{Krause2002}.
+This model can be used for situations in which so called ``hardening of positions'' occur \cite{Krause2002}.
 This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}.
@@
 Luis E.C.
-Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
 Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}
@@
 It is important to note that what somebody sees on his/her time-line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
-Every user has a "hidden time-line $L_i(t)$" which contains the actual time-line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
-At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this "hidden time-line $L_i(t)$" and shows them on the actual time-line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
+Every user has a ``hidden time-line $L_i(t)$'' which contains the actual time-line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
+At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this ``hidden time-line $L_i(t)$'' and shows them on the actual time-line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
 
@@
 \newline
-Another example of introducing some kind of persistence towards the own opinion of the nodes is the "hardening of positions" that was briefly discussed at the end of Section \ref{classMod}.
+Another example of introducing some kind of persistence towards the own opinion of the nodes is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
 
@@
 The Watts-Strogatz model is used in this thesis as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-In the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}) Prof.
+In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}) Prof.
 Perra and Prof.
@@
 Luis E.C.
-Rocha in their paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
 These results include the effect of a high clustering coefficient (Watts-Strogatz model) and a random network model.
@@
 
-\subsection{Starting conditions $\bm{P_0(0) = 0.5}$ and $\bm{P_1(0)=0.5}$}
+\subsection{Starting conditions $P_0(0) = 0.5$ and $P_1(0)=0.5$}
 \label{50-50}
@@
 
-\subsection{Starting conditions $\bm{P_0(0) = 0.2}$ and $\bm{P_1(0)=0.8}$}\label{20-80}
+\subsection{Starting conditions $P_0(0) = 0.2$ and $P_1(0)=0.8$}\label{20-80}
 
@@
 L.E.C.
-Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}).
+Rocha made in the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
 They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
@@
 For each network, the REC and PR method are compared to each other.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.
+The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.
 Error bars are again omitted for a better visualization.\\
@@
 Data points are again displayed with their standard error $SE$.\\
-We can make similar observations as we did in the "Making a fraction of the nodes completely stubborn" paragraph.
+We can make similar observations as we did in the ``Making a fraction of the nodes completely stubborn'' paragraph.
 The results for the REF and REC method look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.\\
@@
 This is the complete opposite of the case where the PR method is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
-The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the "Making a fraction of the nodes completely stubborn" paragraph.\\
+The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the ``Making a fraction of the nodes completely stubborn'' paragraph.\\
 If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.

### 2021-04-23 draft.tex
latest update

@@
 
-%https://www.researchgate.net/publication/324312765_What_is_a_Complex_Network_CDT-2
-
-%https://www.nature.com/articles/s42005-020-0359-6
-
 \section{Definition and network measurements}
@@
 This thesis is not concerned with multi-graphs, nor are self-loops allowed.
-Graphs in which the edges have an orientation are called directed graphs.
+Graphs in which the edges have an orientation are called directed graphs \cite{Costa2018}.
 If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
@@
 a Poisson distribution.
-In this case, most of the nodes have a degree close to the average degree $\left<k\right>$.
+In this case, most of the nodes have a degree close to the average degree $\left<k\right>$ \cite{Barabasi2016}.
 Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
-% explain differences  more, consequences of power-law + need to include figure?
-They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}.
-% correct statement?
-+ talk about fact that av degree can diverge; explain scale free concept a bit more detailed
-%maybe explain why we care about degree, centrality measure etc
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Barabasi2016}\cite{Mata2020}.
 
@@
 At some critical value of the connectivity, most nodes will be connected into a giant cluster (the giant component), which characterizes the percolation of the network \cite{F.Costa2007}.
-% NOT finished; add this as a subsection in the degree section??
+% NOT finished --> need for more explanation?
 
@@
 
-% talk about assortativity of social networks?? + talk about node assortativity instead of global assortativity, see Thed2014?
-
 \newpage
@@
 
-% explain length, diameter, average length,...
 Several very useful measurements quantify the notion of length and distance in a network/graph.
@@
 Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
-% finished?
-
+\newpage
+\noindent
 \subsection{Clustering coefficient and transitivity}
@@
 
-% continue working on this based on http://networksciencebook.com/chapter/9#basics and https://arxiv.org/pdf/1503.06277.pdf
-
 Network communities are another feature that (complex) networks may posses.
@@
 Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
-The modularity of a community $Q_c$ is then calculated by measuring the difference between the network real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
+The modularity of a community $Q_c$ is then calculated by measuring the difference between the network's real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
 This last quantity is given by $p_{ij}$, which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
@@
 
-% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
-% explain small-world (milgram), high clustering, power-law
-
 \section{Network models}
@@
 The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
-% also take about the giant cluster, critical probability??
 
@@
 The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
-
-% also discuss case where edges are added instead of rewired??
+The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$.
 
@@
 
-% NOT FINISHED
-
 \section{Real-world social networks}
@@
 The term scale-free refers to the fact that there is no characteristic scale in the network.
-Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$, the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
-If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges; if $2 < \gamma < 3$, there is a finite average degree, but a diverging variance \cite{Newman2005}.\\
+Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
+If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges, however simple scale-free networks with an exponent smaller than two cannot exist, since the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}; if $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance \cite{Newman2005}.
+A diverging variance means that the fluctuations around the average can be arbitrary large, which makes this average degree no longer a characteristic scale for the system \cite{Barabasi2016}.\\
+It should be noted that not all real-world networks have a heterogeneous degree distribution.
+And even if the degree distribution is heterogeneous or fat-tailed, it doesn't always need to follow a power-law.
+The degree distribution can eg.
+also be a power-law with an exponential cut-off or a log-normal distribution.\\
 \newline
@@
 \noindent
-For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}.
+For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
 Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\

### 2021-04-23 draft.tex
latest update

@@
 \newline
-Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of nodes with opinion 0, $P_0(t)$, over time, together with the standard deviation $\sigma$.
-Two different filtering algorithms are compared to each other, namely the PR method and the REC method (see Chapter \ref{chap3}, Section \ref{filter} for a detailed explanation).
-None of the network topologies, nor the filtering algorithms, is able to break the initial $50/50$ opinion distribution.
-Having a high clustering or a clear community structure has no impact on the prevalence of the opinions.
-The status quo is maintained and the system is not pushed towards a different equilibrium.
-\newpage
+We start with the results for the evolution of the opinion fractions in case of an initial $50/50$ opinion distribution.
+Prof.
+N.
+Perra and prof.
+Luis E.C.
+Rocha already showed that, for the case of balanced initial opinions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the different filtering algorithms \cite{Perra2019}.
+In Figure \ref{ev_op_50_50} the results for the network model with a community structure (SBM, Figure \ref{sbm_50-50}) and the network model with both a community structure and a high clustering (SBM-WS, Figure \ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithm.
+Results are shown together with their standard deviation $\sigma$.
+From this figure it becomes clear that having a community structure is also not able to break the status-quo for both of the filtering algorithms.
+
 \begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_50-50_PR_REC.png}
-    \caption{WS}
-  \end{subfigure}
   \begin{subfigure}[b]{0.49\textwidth}
@@
     \caption{SBM}
+    \label{sbm_50-50}
   \end{subfigure}
@@
     \caption{SBM-WS}
+    \label{sbm-ws_50-50}
   \end{subfigure}
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Two different network models with a community structure are represented, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
 \label{ev_op_50_50}
 \end{figure}
-\newpage
 \noindent
 \textbf{Echo chambers}\\
+\newline
+Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$.
+The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
+The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0).
+Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1.
+The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
+Error bars are omitted for the sake of an improved visualization.\\
+These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\
+
 \begin{figure}[H]
@@
 \noindent
-Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$.
-The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0).
-Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1.
-The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
-Error bars are omitted for the sake of an improved visualization.\\
-These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\
-\newline
 In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR method.
@@
 This model is used to investigate the effect of clustering.
-	\newpage
 	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
@@
 This imbalance leads to the decrease of the minority opinion 0.\\
-\newpage
-\noindent
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
-    \caption{WS}
-    \label{ws_20-80_op}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
-    \caption{SBM-WS}
-    \label{sbm-ws_20-80_op}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
-\label{ev_op_20_80}
-\end{figure}
-\noindent
+\newline
 From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist, $P_0(500) \sim 0$.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
+    \caption{WS}
+    \label{ws_20-80_op}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
+    \caption{SBM-WS}
+    \label{sbm-ws_20-80_op}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+\label{ev_op_20_80}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
@@
 \end{figure}
-\newpage
 \noindent
 \textbf{Echo chambers}\\
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
-    \caption{ER}
-    \label{er_echo_20-80}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
-    \caption{WS}
-    \label{ws}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
-    \caption{SBM}
-  	\label{sbm}    
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
-    \caption{SBM-WS}
-    \label{sbm-ws}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
-Each plot is the average of $10^2$ independent simulations.}
-\label{echo_20_80}
-\end{figure}
-\newpage
-\noindent
+\newline
 Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
@@
 Error bars are again omitted for a better visualization.\\
-\newline
 The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
@@
 \newline
-Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model.
+Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
 When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered.
@@
 In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
+    \caption{ER}
+    \label{er_echo_20-80}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
+    \caption{WS}
+    \label{ws}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
+    \caption{SBM}
+  	\label{sbm}    
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
+    \caption{SBM-WS}
+    \label{sbm-ws}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
+The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
+Each plot is the average of $10^2$ independent simulations.}
+\label{echo_20_80}
+\end{figure}
+
 % talk about small spike around 1 for PR method??
@@
 \\
-\newline
-In Figure \ref{ev_op_50_50_stubb_REC} the evolution of the fraction of nodes with opinion 0 for an initial $50/50$ opinion distribution is shown.
-Data points are shown together with the standard deviation $\sigma$.
-This figure shows the results for the REC filtering algorithm.
-The four network models (ER, WS, SBM (high and low modularity) and SBM-WS) are compared to each other.
-For each model we look at three different fractions of completely stubborn people: no stubborn people, fracRes $=0$; a fraction of completely stubborn people equal to fracRes $=0.4$ and a fraction of completely stubborn people equal to fracRes $=0.8$.
-Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_stubb_REC}
-\end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_stubb_PR}
-\end{figure}
-\noindent
-Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm.
-Data points are again represented together with the standard deviation $\sigma$.
-Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models.
-The results don't differ from those obtained for the REC filtering algorithm.
-The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
+The short answer is no.
+Introducing stubborn people in the system does not affect the evolution of the opinions over time for neither the REC, nor the PR filtering algorithm.
+The balanced opinion distribution is maintained over time.
+This result is valid for all the four implemented network models (ER, WS, SBM and SBM-WS).
+Network topologies such as a high clustering coefficient and/or a community structure do not change this result.
+\\% Include figure??
 \newline
@@
 \\
-\newline
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_stubb_PR}
+\end{figure}
+\noindent
 So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
@@
 Are these results affected by introducing stubborn agents?\\
-\newline
-Figure \ref{ev_op_20_80_stubb_REC} shows the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ distribution and when using the REC filtering algorithm.
-Data point are shown together with their standard deviation $\sigma$.
-Four network models are compared (ER, WS, SBM (high and low modularity) and SBM-WS).
-For each network model we compare three different fractions of completely stubborn nodes (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
-\\
-From this figure it becomes clear that introducing stubborn people in combination with the REC filtering algorithm is not able to break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
-None of the network topologies, nor the stubborn agents can break the status-quo.
-This result is similar to the case of an initial $50/50$ opinion distribution in combination with a fraction of stubborn people, as discussed for Figure \ref{ev_op_50_50_stubb_REC}.\\
-\newline
+As for the balanced initial conditions, introducing stubborn people does not break the prevalence of the initial opinions for the REC filtering algorithm.
+Introducing stubborn people in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
 Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
@@
 The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_20_80_stubb_REC}
-\end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_20_80_stubb_PR}
-\end{figure}
-
-\newpage
-\noindent
+\\
+\newline
 \textbf{Making all nodes partly stubborn}\\
@@
 What about the results for the evolution of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
-Again, we look at the four different network models (ER, WS, SBM (high and low modularity) and SBM-WS).
-For each network model, the results for three different values of the stubbornness parameter $r$ of all the nodes are compared: the nodes are not stubborn, $r=0$; the nodes are partly stubborn with $r = 0.4$ and the nodes are partly stubborn with $r=0.8$.
-Two different filtering algorithms are investigated, namely the REC filtering algorithm and the PR one.\\
-\newline
-Figure \ref{ev_op_50_50_all_frac_stubb_REC} shows the results for the evolution of the fraction of nodes with opinion 0 over time in case of using the REC filtering algorithm.
-From this figure, it becomes clear that making the nodes partly stubborn is not able to break the status-quo.
-The prevalence of the initial opinions is maintained for all the network models and for all the values of the stubbornness parameter $r$.
-This result is exactly the same as the result we obtained for the case in which a fraction of completely stubborn nodes was introduced in the system (in case of an initial $50/50$ opinion distribution and in combination with the REC filtering algorithm), see Figure \ref{ev_op_50_50_stubb_REC}.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_all_frac_stubb_REC}
-\end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac__stubb_PR_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_50_50_all_frac_stubb_PR}
-\end{figure}
-\noindent
-Figure \ref{ev_op_50_50_all_frac_stubb_PR} shows the results for the evolution of the fraction of nodes with opinion 0 over time in case of an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-Also for the PR filtering algorithm, the status-quo is maintained for all the network models and for all the values of the stubbornness parameter $r$.
-This result is again the same as the case where we introduced a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution and using the PR filtering algorithm, see Figure \ref{ev_op_50_50_stubb_PR}.\\
+Well, making all the nodes partly stubborn does not disturb/change the evolution of the opinions.
+The balanced initial conditions remain balanced over time, as was the case for no stubbornness in the system.
+This result is valid for all the network models (ER, WS, SBM and SBM-WS) and for both the REC and PR filtering algorithm.\\
 \newline
@@
 The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
-\newline
-In case of the PR filtering algorithm (Figure \ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models for a stubbornness parameter less than $r \leqslant 0.4$.
-If the stubbornness parameter becomes bigger than $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
-This behavior is slightly different than in the case of an initial $50/50$ distribution, combined with the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
-In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if the stubbornness parameter became bigger than $r>0.1$).
-This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
-As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
-It should be noted that these types of figures are not really suited to compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
-In order to perform such a comparison, one should look at the complete distribution (see eg.
-Figure \ref{echo_20_80}), as was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
-\newline
-The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC method (see Figure \ref{REC_all_frac_stubb}).
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to a stubbornness parameter $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
-The increase seems, however, less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
-This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
-The stochastic block model with a high modularity also shows a slight increase in the formation of echo chambers up to $r = 0.8$.
-This increase can even be observed for the low modularity case, albeit less strong.
-The Erd\H{o}s-R\'{e}nyi network seems to wiggle a bit around a formation of echo chambers equal to one.
 
@@
 \noindent
+In case of the PR filtering algorithm (Figure \ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models for a stubbornness parameter less than $r \leqslant 0.4$.
+If the stubbornness parameter becomes bigger than $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
+This behavior is slightly different than in the case of an initial $50/50$ distribution, combined with the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
+In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if the stubbornness parameter became bigger than $r>0.1$).
+This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
+As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
+It should be noted that these types of figures are not really suited to compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+In order to perform such a comparison, one should look at the complete distribution (see eg.
+Figure \ref{echo_20_80}), as was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+\newline
+The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC method (see Figure \ref{REC_all_frac_stubb}).
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to a stubbornness parameter $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
+The increase seems, however, less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
+This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
+The stochastic block model with a high modularity also shows a slight increase in the formation of echo chambers up to $r = 0.8$.
+This increase can even be observed for the low modularity case, albeit less strong.
+The Erd\H{o}s-R\'{e}nyi network seems to wiggle a bit around a formation of echo chambers equal to one.\\
+\newline
+
 We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn.
@@
 Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
-\newline
-Figure \ref{ev_op_20_80_all_frac_stubb_REC} gives the results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm.
-For none of the network models, nor the values of the stubbornness parameter $r$, the initial prevalence of the opinions can be broken.
-This result is analogue to the result for an initial $20/80$ opinion distribution, combined with the REC filtering algorithm, in the case of introducing a fraction of completely stubborn people (see Figure \ref{ev_op_20_80_stubb_REC}).
-
-\newpage
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_REC_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_REC_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_REC_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_REC_20-80_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the REC filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_20_80_all_frac_stubb_REC}
-\end{figure}
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_20_80_all_frac_stubb_PR}
-\end{figure}
-\noindent
+The results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm are not affected by making all the nodes more and more stubborn.
+The prevalence of the opinions is maintained an the status quo is not broken.\\
+The PR filtering algorithm gives us another story.
 The results for the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
@@
 
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_all_frac_stubb_PR}
+\end{figure}
+\noindent
 \subsection{The majority threshold model}

### 2021-04-24 draft.tex
latest update
continued writing chapter 4, opinion dynamics in this thesis

@@
 \usepackage{xfrac}
-\floatname{algorithm}{Algoritme}
+\floatname{algorithm}{Algorithm}
 
@@
 \emph{Supervisor:} \\
-\textcolor{NavyBlue}{Prof. Dr. Luis E.C. \textsc{Rocha}} % Supervisor name - remove the \href bracket to remove the link  
+\textcolor{NavyBlue}{Prof. Dr. Luis E.C. \textsc{Rocha}}\\
+\emph{Co-Supervisor:}\\
+\textcolor{NavyBlue}{Prof. Dr. Jan \textsc{Ryckebusch}}
+ % Supervisor name - remove the \href bracket to remove the link  
 \end{flushright}
@@
 This can be explained by the fact that if the network is sparse enough, we may have disconnected parts; each of these disconnected groups may then evolve to a different consensus opinion, so that over the entire network different opinions coexist \cite{Nguyen2020}.
-% https://arxiv.org/pdf/0710.3256.pdf Contains a mean field mathimatical description, interesting?
-
-\subsection{The probabilistic majority model}
+
+\subsection{The probabilistic majority model}\label{probMaj}
 
@@
 This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}.
-% https://arxiv.org/pdf/0710.3256.pdf read this (overview, discusses several models)!!
 
@@
 Hence, we will not go into further detail.
-Interested readers are referred to references \cite{Castellano2009}, \cite{Fernandez2014} and related references therein.% cite papers that have done studies on this model??
+Interested readers are referred to references \cite{Castellano2009}, \cite{Fernandez2014} and related references therein.
 
@@
 This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}.
-% talk about extension http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf
 
@@
 If the opinions of the two nodes $i$ and $j$ are too far apart nothing happens; this is the case if $\abs{x_i(t) - x_j(t)} \geqslant \epsilon$.
-If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, then the opinions are updated to \cite{Castellano2009}
+If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, then the opinions are updated to \cite{Castellano2009}\cite{Deffuant2000}
 \begin{equation}
@@
 As mentioned earlier, the threshold $\epsilon$ can be different for each node, so that each node has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
-\epsilon_r$ \cite{Krause2002}.
-% talk about relative agreement model?? http://jasss.soc.surrey.ac.uk/9/3/8.html and http://jasss.soc.surrey.ac.uk/5/4/1.html
-
-% https://arxiv.org/pdf/0710.3256.pdf, bounded confidence methods
-% https://www.researchgate.net/publication/235737446_Mixing_Beliefs_Among_Interacting_Agents defuant model
-% http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf
-% http://jasss.soc.surrey.ac.uk/9/3/8.html
+\epsilon_r$ \cite{Krause2002}.\\
+Another extension is the relative agreement interaction model. Interested readers are referred to references \cite{Deffuant2006} and \cite{Deffuant2002}.
 
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
 \newline
@@
 \label{stubbBin}
-% not finished!!
+
 Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models.
@@
 For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
-% Talk about probabilistic implementation? + talk about fact that nothing happens for T < 0.5???
 \newline
 Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study, many more exist or can be invented.
-
-%\section{The opinion dynamics model used in this thesis}\label{thesisOp}
-
-% https://arxiv.org/pdf/0710.3256.pdf Talk about agent-based models??
 
@@
 The Erd\H{o}s-R\'{e}nyi model is also used in the regular stochastic block model to construct the constituent communities.\\
-% talk about expected behaviour?? 
 \newline
@@
 It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic.
-\section{Opinion dynamics models}\label{opinionmodel}
-% use pseudo code here??
-% explain the main model and state that this is used in thesis unless specified otherwise
+
+% Discuss actual networks we used!! + reference them??
+
+\section{Opinion dynamics model}\label{opinionmodel}
+
+In Chapter \ref{chap3} we already discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
+The goal of this section is to put every thing together and summarize the opinion dynamics model of this thesis.\\
+\newline
+This thesis follows the opinion dynamics model that was designed by prof.
+Perra and prof.
+Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}.
+\\
+We start with a network with $N$ users.
+Initially, each user is given one of the two opinions 0 or 1.
+Each time step a random fraction equal to $p= 0.1$ of these users becomes active, where $p$ is the activation probability.
+As discussed in Chapter \ref{chap3}, Section \ref{actMech}, this thesis makes use of a constant activation probability.
+Every time step, all active users update their opinion synchronously.
+This updating is done as follows.
+First, the active user sends his current opinion to all his neighbors (active or not).
+Then, he updates his own opinion according to the opinions on his time-line.
+This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}.
+If another model instead of the probabilistic majority model is used, eg.
+the majority model, this will be explicitly stated.
+If not, one can assume that the probabilistic majority model is implemented.
+It is important to realize that what an active user sees on his/her time-line is only a selection of all the posts that his/her friends have posted.
+Remember from Chapter \ref{chap3}, Section \ref{actMech} that each user has a hidden time-line $L_i(t)$.
+This hidden time-line contains the time-line of the user at the last time he/she was active ($t_{\text{last}}$) plus all the post that his/her friends posted from $t_{\text{last}}$ until the user becomes active again at time $t$.
+If a user becomes active, the filtering algorithm (REF, REC or PR), discussed in Chapter \ref{chap3}, Section \ref{filter}, selects $S$ items from this hidden time-line and shows them on the actual time-line of the user.
+We will usually set $S = 20$, which quantifies the maximum length of the actual time-line.\\
+
+\begin{algorithm}
+\caption{Updating the opinions of the active nodes at time $t$}
+\label{alg:update}
+\begin{algorithmic}[1]
+	\State \(L \leftarrow \text{hidden time-line}\)
+	\State \(T \leftarrow \text{actual time-line}\)
+	\State \(S \leftarrow 20\) \Comment{Maximum length of time-line $T$}
+	\State \(r\) \Comment{Stubbornness parameter of each node, takes values between $0$ and $1$}
+	\ForAll{active user}
+		\State \(\text{op} \leftarrow \text{current opinion}\)
+		\ForAll{neighbor} \Comment{Add opinion of active user to hidden time-line of each neighbor}
+			\State \(L.\textsc{add}\text{(op)}\)
+		\EndFor
+		\State \(f \leftarrow \text{rand(0, 1)}\) \Comment{Draw random number between $0$ and $1$}
+		\If{$f > r$}
+			\State \(\text{op} \leftarrow \Call{Update opinion}{L}\) \Comment{Update opinion of active user}
+		\Else
+			\State \(\text{op} \leftarrow \text{current opinion}\) \Comment{Don't update opinion of active user}
+		\EndIf
+	\EndFor
+
+	\Function{Update opinion}{$L$}
+		\State \(\textsc{Order}(L)\) \Comment{Order hidden time-line according to REF, REC or PR filtering}
+		\If{$\text{len}(L) > S$}
+			\State \(T \leftarrow L.\textsc{resize}(S)\) \Comment{Time-line $T$ = $S$ first elements of ordered hidden time-line}
+		\EndIf
+		\State \(\text{op} \leftarrow \textsc{probMaj}(T)\) \Comment{Update opinion according to probabilistic majority model}
+	\State \Return op \Comment{Return updated opinion}
+	\EndFunction
+\end{algorithmic}
+\end{algorithm}
+\noindent
+If we want to introduce some kind of stubbornness in the model we will mainly use the stubbornness parameter, as was explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
+Lets briefly summarize this concept of stubbornness.
+Each node is given a stubbornness parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$ the node can update his opinion, if $f \leqslant r_i$ the node will keep his current opinion.\\
+Some results for the majority threshold model, also explained in Chapter \ref{chap3}, Section \ref{stubbBin}, will be shared as well.
+This is, however, not the main stubbornness model that we will focus on.
+If this model is used, this will be explicitly stated.
+Otherwise one can assume that we are making use of the stubbornness parameter.\\
+In Algorithm \ref{alg:update}, the pseudo-code for updating the opinions of the active nodes in case of using a stubbornness parameter $r$ at time step $t$ is displayed. \\
+\newline
+This model is a very simple one, since it only allows for two different opinions in the system.
+The designed filtering algorithms are very simple and intuitive as well.
+It also introduces some randomness by using the probabilistic majority model instead of the classic majority model.
+This might make the model a bit more realistic.
+This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.
+A simple model is more tractable and allows for a better understanding of what is going on in the dynamics.
+However, this might be a downside as well.
+Real-world opinion dynamics are, probably, far from simple.
+People don't update their opinion according to a simple probabilistic opinion dynamics model.
+Beside that, the cases where people have to choose between only two opinions are rather limited.
+Filtering algorithms, used by real social media companies, are unfortunately not publicly shared.
+We can however assume that they are not as simple as the ones we used in this thesis.
+Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
+This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results.
+\\
+This doesn't mean that simple models should be thrown away.
+Definitely not!
+They might still be capable of capturing a lot of the dynamics that are seen in the real world.
+Not because a system appears very complex, that it cannot be modeled by something far more simple.
+This has been proven by physics research over and over again.
+Beside that, if you don't know the dynamics/underlying processes of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
+There is no need to make things too complex if simple works too.
+There is however a need to compare data from (computational) models to experimental, real-world data.
+A need for critical thinking and careful analyzing of results can also not be denied.
+% put this section here?? Reference??
 

### 2021-04-27 draft.tex
latest update

@@
 In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to.
-This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
+This filtering and individualizing of information may have some potentially dangerous downsides in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
 It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
@@
 Section \ref{statPhys} introduces concepts such as statistical and social physics.
-Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are discussed.
+Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are summarized and discussed.
 
@@
 The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
-\newline
+\newpage
+\noindent
 Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms.
@@
 The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
-The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$.
+The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.
 
@@
 In contrast to discrete/binary models, all agents usually start with  different opinions \cite{Castellano2009}.
-The possible outcomes are more complex than in the discrete models and often has opinion clusters emerging in the final states (where opinion clusters refer to clusters of nodes with the same opinion) \cite{Castellano2009}.
+The possible outcomes are more complex than in the discrete models and often have opinion clusters emerging in the final states (where opinion clusters refer to clusters of nodes with the same opinion) \cite{Castellano2009}.
 The number of opinion clusters can be one (consensus), two (polarization) or more (fragmentation) \cite{Castellano2009}.
@@
 Every user has a ``hidden time-line $L_i(t)$'' which contains the actual time-line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
-At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this ``hidden time-line $L_i(t)$'' and shows them on the actual time-line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
+At the time $t$ that the user becomes active again, the filtering algorithm selects $S$ post from this ``hidden time-line $L_i(t)$'' and shows them on the actual time-line of the user; these $S$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
 
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
-\newline
-Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of such extreme and persistent groups a lot bigger.
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+\newline
+Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups a lot bigger.
 Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people.
-So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a wide reach and a significant impact on society).
+So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
 Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
@@
 Can these stubborn groups lead to polarization or even overtake the opinions in the system?
-Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism?
+Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism/polarization?
 To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.
@@
 	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
-A uniform random number $f$ between 0 and 1 is drawn.
+An uniform random number $f$ between 0 and 1 is drawn.
 If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.
@@
 The reasons of choosing these specific networks will be explained, together with there possible strengths and weaknesses.\\ % + explain where they come from etc?
-Finally, Section \ref{opinionmodel} discusses the opinion dynamics model that is implemented in this thesis. Again we will give some focus on possible advantages and disadvantages of the model.
+Finally, Section \ref{opinionmodel} summarizes and discusses the opinion dynamics model that is implemented in this thesis.
+Again we will give some focus on possible advantages and disadvantages of the model.
 
@@
 The designed filtering algorithms are very simple and intuitive as well.
-It also introduces some randomness by using the probabilistic majority model instead of the classic majority model.
+Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
 This might make the model a bit more realistic.
@@
 \\
-Note that for an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (see Figure \ref{ev_op_50_50}).
+Note that for an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (see Figure \ref{ev_op_50_50} and related results in reference \cite{Perra2019}).
 Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a more clustered opinion distribution and thus lead to polarization and formation of echo chambers.
@@
 If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
-This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
+This enhancing of the formation of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
 Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks.
@@
 	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
-	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$ to completely stubborn, $r = 1$).
+	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$, to completely stubborn, $r = 1$).
 \end{itemize}
@@
 The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
-It is however not recommended to compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution.
+It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution.
 The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
 When comparing the different network models, these skewed distributions should not be overlooked.
-The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to compare the different models against each other.
+The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to (quantitatively) compare the different models against each other.
 \\
@@
 For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
-\newline
+\newpage
+\noindent
 We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
 \\
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM-WS}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
-\label{ev_op_20_80_stubb_PR}
-\end{figure}
-\noindent
+\newline
 So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
@@
 The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.
-\\
-\newline
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{ER}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{WS}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
+    \caption{SBM-WS}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+\label{ev_op_20_80_stubb_PR}
+\end{figure}
+\newpage
+\noindent
 \textbf{Making all nodes partly stubborn}\\
@@
 All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
-We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
+We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where-after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
 
@@
 As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
-It should be noted that these types of figures are not really suited to compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
-In order to perform such a comparison, one should look at the complete distribution (see eg.
+It should be noted that these types of figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
 Figure \ref{echo_20_80}), as was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
@@
 \newline
-
 We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.
+The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The majority threshold model is implemented.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
@@
 However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold doesn't become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
-At this point, none of the nodes can be convinced of another opinion and hence keep their initial opinion.
+At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
 The number of echo chambers doesn't change over time, which results in an echo chamber size equal to one in the figure.\\
@@
 The figure shows the evolution of opinion 0 over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion 0; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion 0 as to opinion 1, the opinion to which the system eventually converges is a consequence of random fluctuations).
-The figure shows the results for each of the four network model: ER, WS, SBM (high and low modularity) and SBM-WS.
+The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
 For each model the REC and PR filtering algorithm are compared to each other.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.
+The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+The majority threshold model is implemented.
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
@@
 Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
-If we, however, look at the absolute number of echo chambers around opinion 0 at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case.
+If we, however, would look at the absolute number of echo chambers around opinion 0 at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
 This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0.
-Beside this hampering of echo chambers around the majority opinion 0, the community distributed case also shows an increase of echo chambers around the minority opinion 1.
-Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion 0 (and thus the decrease of the minority opinion 1).
-Hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
+Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion 0 (and thus the decrease of the minority opinion 1); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed cases versus the random distributed case.
+Beside this hampering of echo chambers around the majority opinion 0, the community distributed case also shows an increase of echo chambers around the minority opinion 1 (which is not visible for the randomly distributed case).
+This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
 The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network.
@@
 The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case.
-However, as for the high modularity case, if we look at the absolute number of echo chambers around the majority opinion at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case.
+However, as for the high modularity case, if we would look at the absolute number of echo chambers around the majority opinion at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
 % make conclusion about this ?
@@
 The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion 0 (Figure \ref{echo_01_commOp0_other_50-50}).\\
-\newline
 The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion 0 with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
@@
 The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Figure \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
-They show an increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
+They show a small increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
 \newline
@@
 
+We made use of simple toy models of real social media platforms to generate results regarding the evolution of opinions and (possible) formation of echo chambers on these platforms.
+These toy models consist of several layers, including a network layer and an opinion dynamics layer.
+Simple filtering algorithms were implemented to replicate and simulate the content curation used by these real social media companies.
+Furthermore, a notion of stubbornness was introduced to investigate the interplay between this individual resistance to change, network structure and filtering algorithm.
+For simplicity we only allowed two competing opinions (0 and 1) in the system.\\
+\newline
+We investigated the effect of the network structure on the dynamics of the opinions in the system.
+The main focus was put on the difference between networks with a high clustering coefficient and no community structure (WS) compared to networks with a clear community structure and a low clustering coefficient (SBM).
+It was found that a clear community structure does enhance the formation of echo chambers (mainly when using the PR filtering algorithm).
+This enhancement is however not as strong as for networks with a high clustering coefficient.
+Networks that have both a high clustering coefficient and a high modularity (clear community structure) (SBM-WS) behave similar as the high clustering networks (WS).
+This indicates that it is mainly the high clustering that drives the formation of echo chambers.
+There is however a clear correlation between modularity and formation of echo chambers: a higher modularity results in a larger formation of echo chambers (in case of the PR filtering algorithm).
+The enhancement is just not as strong as the enhancement in the formation of echo chambers introduced by a high clustering.
+\\
+% also talk about evolution of opinions??
+\newline
+When introducing a notion of stubbornness in the system (with the help of a stubbornness parameter $r$), we found that the interplay between the PR filtering algorithm and this individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
+The REF and REC filtering algorithms, on the other hand, showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if the stubbornness was further increased.
+This is mainly seen in the high clustering networks, however a tiny bump can also be observed in the high modularity network.
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen in both the case of balanced initial conditions ($50/50$) and unbalanced initial conditions ($20/80$).\\
+% talk about evolution of opinions??
+\newline
+It was also found that how the opinions are initially distributed over the communities in a network with a community structure (SBM) does makes a difference.
+We investigate a scenario in which a fraction of the communities was given a single opinion, opinion 0, while the remaining communities were given an initial $50/50$ distribution.
+This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.
+This was done for both a high modularity and a low modularity network and for both the REC and PR filtering algorithms.\\
+The REC method did not show any difference in the evolution of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.\\
+For the PR filtering algorithm it was found that the high modularity network hampers the increase in the majority opinion in the community distributed scenario compared to the randomly distributed scenario.
+For the low modularity network, such a discrepancy was not found.\\
+% talk about echo chambers??
+
 % put here: be carefull with extrapolating data/results to real world cases, need for more real data to compare with (eg data from role of facebook in elections in US), this lack of comparison between real data and network model data/analytical work = one of the downfalls/backsides of social physics??

### 2021-04-28 draft.tex
latest update

@@
 \vspace{0.4cm} 
-{\huge \bfseries Opinion dynamics on social media with stubborn actors\par}\vspace{0.4cm} % Thesis title
+{\huge \bfseries Opinion dynamics on social networks with stubborn actors\par}\vspace{0.4cm} % Thesis title
 \hrule % Horizontal line
@@
 \\
+\newline
 It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings \cite{Perra2019}.
@@
 One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\
+\newline
 The advantage of working with network models is that they reduce the level of complexity encountered in the real world, so that one can treat these systems in a more practical way.
@@
 Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.\\
+\newline
 Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc.
@@
 Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
+\newline
 The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
@@
 \\
+\newline
 One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others.
@@
 The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
-\newpage
-\noindent
+\newline
 Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms.
@@
 This branch of physics offers a framework that relates microscopic properties of atoms, molecules,...
-to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.\\
+to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
 The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
@@
 Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}.
-\\
-However, the use of statistical physics may also offer a way out.
+The use of statistical physics may, however, also offer a way out.
 In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}.
@@
 \newline
-In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}) prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
-They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).\\ Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}.
+In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}) Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
+They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).
+Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}.
 On the other hand, a more heterogeneous contact pattern could hamper this formation of echo chambers \cite{Perra2019}.
 If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
-In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.\\
+In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.
 They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
@@
 If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
-A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.\\
+\newline
+A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.
 
@@
 \end{equation}
-with, generally, $0 \leqslant w_{ij} \leqslant 1$ \cite{Mata2020}. \\
+with, generally, $0 \leqslant w_{ij} \leqslant 1$ \cite{Mata2020}.
+
 \subsection{Degree and degree distribution}
@@
 \subsection{Connectivity}
+
 In physics literature, the degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
 This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.\\
+\newline
 Two nodes $u$ and $v$ of a graph are connected if there exist a path between them. If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
+\newline
 The average degree $\left<k\right>$ is a global measurement of the connectivity of the network \cite{Costa2008}.
@@
 \\
+\newline
 The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}.
@@
 If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
-However, the terms assortativity and disassortativity are also often used without taking degree correlations into account.
+\newline
+The terms assortativity and disassortativity are also often used without taking degree correlations into account.
 Instead, one may look at a particular node property to classify networks as either assortative or disassortative \cite{Thed2014}.
@@
 
-\newpage
 \subsection{Network distances}
@@
 \begin{equation}\label{clus}
-	C_i = \frac{2n_i}{k_i(k_i-1)} \ ,
+	cc_i = \frac{2n_i}{k_i(k_i-1)} \ ,
 \end{equation}
@@
 \begin{equation}\label{avClus}
-	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i \ ,
+	\left<cc\right> = \frac{1}{N}\sum_{i = 1}^N cc_i \ ,
 \end{equation}
@@
 \\
+\newpage
+\noindent
 Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
@@
 It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
-It is then obvious that the local clustering coefficient $C_i$ (Eq.
+It is then obvious that the local clustering coefficient $cc_i$ (Eq.
 (\ref{clus})) of a node and the transitivity $T_i$ (Eq.
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
-In the case of the average clustering coefficient $\left<C\right>$ and the global transitivity $T$ this is, however, not the case.
+In the case of the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$ this is, however, not the case.
 The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node \cite{Newman2003}.
@@
 Let us consider three possible definitions: maximum cliques, strong communities and weak communities \cite{Albert2016}.\\
-\newline
+\newpage
+\noindent
 \textbf{Maximum cliques}\\
@@
 This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
-\newpage
-\noindent
+\newline
 \textbf{Strong communities}\\
@@
 Generally, no a priori information is available about the number of existing communities in real networks \cite{F.Costa2007}.\\
+\newline
 The modularity for a network with $N$ nodes and $L$ edges and partitioned into $n_c$ communities is calculated as follows.
@@
 \begin{equation}\label{C}
-	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1} \ .
+	\left<cc\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1} \ .
 \end{equation}
@@
 The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates).
-If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<C\right> \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
+If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<cc\right> \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
 \\
@@
 \\
-\\
+\newline
 To summarize: the random Erd\H{o}s-R\'{e}nyi network has a bell shaped degree distribution which approximates a Poisson distribution for $N \gg 1$; the average clustering coefficient becomes small for large networks and the average distance depends logarithmic on the system sizes.
 The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the ``small world phenomenon'' (namely short average distance).\\
+\newline
 Real social networks, however, usually display the ``small world phenomenon'' and thus have both a short average distance and a high clustering coefficient.
@@
 The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
-The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<C(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
+\newline
+The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<cc(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
 \begin{align}\label{C0}
-	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
+	\left<cc(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
 \end{align}
@@
 \\
-The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<C(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
+\newline
+The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
 (\ref{C}) and Eq.
@@
 \begin{align}
-	\left<C(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
+	\left<cc(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
 \end{align}
@@
 \begin{equation}
-	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3 \ ,
-\end{equation} 
-where $\left<C(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
+	\left<cc\right> \approx \left<cc(0)\right>(1-\beta)^3 \ ,
+\end{equation} 
+where $\left<cc(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
 \begin{itemize}
@@
 The structure within the groups or communities of the stochastic block model is that of a random graph, while the structure between the communities is that of a random bipartite graph \cite{Clauset2017}.\\
+\newline
 In this thesis, the stochastic block model is generated in the following way.
@@
 The stochastic block model can be fitted to data \cite{Clauset2017}.
-Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.\\
+Given a choice of the number of communities $k$ and an observed network $G$, the stochastic block model can be used to infer the latent community structure and the edge probability matrix $M$ \cite{Clauset2017}.
 One of the shortcomings of fitting the SBM to infer the community structure, is that real-world networks often have a heterogeneous or power-law degree distribution.
@@
 An extension that we are going to explore a bit deeper is the stochastic Watts-Strogatz block model (SBM-WS).\\
-\newpage
-\noindent
+\newline
 \textbf{The stochastic Watts-Strogatz block model (SBM-WS)}\\
@@
 This phenomenon is now often referred to as ``six degrees of separation'' \cite{RealWorld}.\\
+\newline
 This phenomenon is also observed in many other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
@@
 A diverging variance means that the fluctuations around the average can be arbitrary large, which makes this average degree no longer a characteristic scale for the system \cite{Barabasi2016}.\\
+\newline
 It should be noted that not all real-world networks have a heterogeneous degree distribution.
@@
 \\
+\newline
 There are, however, some properties that are different between on-line and off-line social networks as well.
@@
 Hence, these opinion dynamics are the dynamics that run on the underlying network structure.\\
+\newline
 Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.\\
@@
 
-Binary models are models where each agent can have one of the two opinions A or B (or, equivalently, 0 or 1).
+Binary models are models where each agent can have one of the two opinions A or B.
 A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
+\newline
 In this section, some binary models will be discussed; note, however, that the models described here are far from all the models that currently exist.
@@
 One of the simplest opinion dynamics models is the binary majority model.
-In this model each agent has one of the two opinions A or B (or 0 or 1).
+In this model each agent has one of the two opinions A or B.
 Initially, these opinions can be equally distributed or not.
 If the distribution is not equal, but instead eg.
-80/20, there is a predominant opinion and a minority opinion.\\
+80/20, there is a predominant opinion and a minority opinion.
 The following updating rule is then imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
@@
 \\
+\newline
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
@@
 Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
-\newpage
-\noindent
+\newline
 For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
 Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
+\newline
 This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}.
@@
 
-Another extremely simple binary model, maybe even simpler than the majority model, is the voter model.
+Another simple binary model, maybe even simpler than the majority model, is the voter model.
 The model works as follows: each agent in the network has a binary opinion (A or B); at each time step, an agent is randomly selected along with one of its neighbors and, subsequently, the agent adopts the opinion of this neighbor \cite{Castellano2009}.
@@
 \\
+\newline
 The pressure of the majority in only felt in an averaged way, in the sense that a neighbor that carries the majority opinion has more chance to be randomly selected \cite{Castellano2009}.\\
+\newline
 Lots of studies have been performed on this model.
@@
 In continuous opinion dynamics models, the opinion variable of the agents in the network is no longer discrete, in contrast to binary or discrete models, but takes real values (often in the interval $\interval{0}{1}$ or $\interval{-1}{1}$).\\
+\newline
 In continuous opinion dynamics models, some concepts that are valid for the discrete models no longer apply (eg.
@@
 \\
+\newline
 The basis of most continuous models is a set of $N$ nodes or agents, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
@@
 \\
+\newline
 Several models have been proposed, we will only discuss a fraction of them.
@@
 where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
+\newline
 Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}.
@@
 where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix. If all the $g_i$ are equal to zero, we obtain the model of Eq. (\ref{refMod}) \cite{Krause2002}.\\
+\newline
 A possible time variant extension is \cite{Krause2002}
@@
 \\
+\newline
 The two most popular bounded confidence models are the Deffuant model and the Hegselmann-Krause model \cite{Castellano2009}. \\
@@
 The parameter $\mu$ is called the convergence parameter; its value lies in the interval $\interval{0}{1/2}$ \cite{Castellano2009}. \\
-The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two agents move closer to each other \cite{Castellano2009}. \\
+\newline
+The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two agents move closer to each other \cite{Castellano2009}.
 The final state is a situation where several opinion clusters coexist.
@@
 where $A_{ij}$ are the elements of the adjacency matrix. This model thus takes the average opinion of the compatible neighbors into account \cite{Castellano2009}.\\
+\newline
 The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary opinion clusters \cite{Castellano2009}.
@@
 As mentioned earlier, the threshold $\epsilon$ can be different for each node, so that each node has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
-\epsilon_r$ \cite{Krause2002}.\\
-Another extension is the relative agreement interaction model. Interested readers are referred to references \cite{Deffuant2006} and \cite{Deffuant2002}.
+\epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to references \cite{Deffuant2006} and \cite{Deffuant2002}.
 
@@
 \\
-It is thus important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
-Prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
+\newline
+It is important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
+Perra and Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
 Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}
@@
 \end{itemize} 
-The following paragraph briefly reviews the filtering algorithms that prof. Nicola Perra and prof. Luis E.C. Rocha proposed. \\
+The following paragraph briefly reviews the filtering algorithms that Perra and Rocha proposed. \\
 \newline
@@
 \begin{itemize}
-	\item The reference method (REF): this method randomly selects $S$ items from the list $L_i(t)$.
-	\item The old method (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user.
-	\item The recent method (REC): shows the $S$ most recent items from $L_i(t)$.
-	\item The preference method (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
-This filtering method is based on the semantic filtering explained above.
+	\item The reference filtering algorithm (REF): this filtering algorithm randomly selects $S$ items from the list $L_i(t)$.
+	\item The old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user.
+	\item The recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$.
+	\item The preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
+This filtering algorithm is based on the semantic filtering explained above.
 \end{itemize}
-Prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha showed that the REF, OLD and REC methods behave similar in many cases and that the PR method often displays different behavior (such as a larger formation of echo chambers compared to the other three methods) \cite{Perra2019}.
-In this thesis the REC and REF methods are mainly used as a reference/benchmark method and the PR method is used as the main filtering algorithm.
+Perra and Rocha showed that the REF, OLD and REC filtering algorithms behave similar in many cases and that the PR filtering algorithm often displays different behavior (such as a larger formation of echo chambers compared to the other three filtering algorithms) \cite{Perra2019}.
+In this thesis the REC and REF filtering algorithms are mainly used as a reference/benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
 
@@
 \\
+\newline
 It might, however, be a more realistic situation if nodes/users become active/inactive over time.
@@
 \newline
-In this thesis we follow the same activation mechanism as was proposed in the paper of prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha (see reference \cite{Perra2019}).
+In this thesis we follow the same activation mechanism as was proposed in the paper of Perra and Rocha (see reference \cite{Perra2019}).
 Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$).
 In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes.
-Prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha chose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
+Perra and Rocha chose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
 They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
@@
 \\
-Hence, for simplicity, this thesis implements the constant activation probability variant.
+\newline
+For simplicity, this thesis implements the constant activation probability variant.
 It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
@@
 In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.\\
+\newline
 It is important to note that what somebody sees on his/her time-line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
@@
 
-\newpage
 \section{The presence of stubborn people}\label{stubb}
@@
 This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
+\newline
 Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in reference \cite{Deffuant2006}. \\
@@
 \\
+\newline
 One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
@@
 \\
+\newline
 Section \ref{networkmodel} gives an overview of the network models that are used in this thesis.
@@
 What are there benefits and possible shortcomings regarding to our goals?\\
+\newline
 In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.
 The reasons of choosing these specific networks will be explained, together with there possible strengths and weaknesses.\\ % + explain where they come from etc?
+\newline
 Finally, Section \ref{opinionmodel} summarizes and discusses the opinion dynamics model that is implemented in this thesis.
@@
 \\
+\newline
 It is expected that the random nature of the model and the lack of a high clustering and a community structure will result in the absence of echo chambers.\\
+\newline
 The Erd\H{o}s-R\'{e}nyi model is also used in the regular stochastic block model to construct the constituent communities.\\
@@
 The Watts-Strogatz model is used in this thesis as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}) Prof.
-Perra and Prof.
-Rocha already showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
+\newline
+In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}) Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
 Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn people and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
 We will investigate whether a community structure behaves similar to a high clustering regarding the appearance of echo chambers.\\
+\newline
 Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
@@
 We will therefore use this model to investigate the impact of community structure/modularity on the evolution of opinions and the formation of echo chambers.\\
+\newline
 It is expected that a community structure will enhance the formation of echo chambers compared to the random network case.
@@
 It is also of interest to investigate the possible relation between the appearance of echo chambers and the modularity of the network.\\
+\newline
 The stochastic Watts-Strogatz block model is used as a model that has both a high clustering and a community structure.
@@
 They are the most simple network models that one can create and each of them have characteristics and topologies that can be easily tuned (clustering for WS, community structure for SBM).
-This makes it really easy to investigate the effect of different network topologies (clustering, community structure) without the need of taking the interplay with other possible complex network features and topologies into account.\\
+This makes it really easy to investigate the effect of different network topologies (clustering, community structure) without the need of taking the interplay with other possible complex network features and topologies into account.
 This plus side is, however, also one of the possible downsides.
@@
 It is thus always a bit tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
+\newline
 The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.\\
@@
 If the networks would be to large, it would take weeks, sometimes even months before there would be any kind of results.\\
+\newline
 The size of the network models that were used, was around a thousand nodes.
@@
 \newline
-This thesis follows the opinion dynamics model that was designed by prof.
-Perra and prof.
-Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}.
-\\
+This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}.
+\\
+\newline
 We start with a network with $N$ users.
-Initially, each user is given one of the two opinions 0 or 1.
+Initially, each user is given one of the two opinions A or B.
 Each time step a random fraction equal to $p= 0.1$ of these users becomes active, where $p$ is the activation probability.
@@
 For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$ the node can update his opinion, if $f \leqslant r_i$ the node will keep his current opinion.\\
+\newline
 Some results for the majority threshold model, also explained in Chapter \ref{chap3}, Section \ref{stubbBin}, will be shared as well.
@@
 Otherwise one can assume that we are making use of the stubbornness parameter.\\
+\newline
 In Algorithm \ref{alg:update}, the pseudo-code for updating the opinions of the active nodes in case of using a stubbornness parameter $r$ at time step $t$ is displayed. \\
 \newline
-This model is a very simple one, since it only allows for two different opinions in the system.
-The designed filtering algorithms are very simple and intuitive as well.
+This model is a simple one, since it only allows for two different opinions in the system.
+The designed filtering algorithms are simple and intuitive as well.
 Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
@@
 \\
+\newline
 This doesn't mean that simple models should be thrown away.
-Definitely not!
 They might still be capable of capturing a lot of the dynamics that are seen in the real world.
-Not because a system appears very complex, that it cannot be modeled by something far more simple.
+Not because a system appears complex, that it cannot be modeled by something far more simple.
 This has been proven by physics research over and over again.
@@
 
-In this chapter the results are shared and discussed. The chapter is subdivided into four main sections
-\begin{itemize}
-	\item The impact of network structure: the effect of different network topologies on the evolution of the opinions and the formation of echo chambers is reviewed.
-Main focus lies on networks with a high clustering coefficient and/or a community structure.
-	\item The impact of stubborn people: stubborn people are introduced in the system; in this section the effects of those stubborn people on the evolution of opinions and the formation of echo chambers will be discussed.
-	\item Random versus community distributed opinions/stubborn agents: this section investigates different ways of distributing the opinions and/or the stubborn agents in a network model with a community structure (SBM).
-	\item Real-world networks: the opinion dynamics model of this thesis will be ran on some real-world networks.
-This section will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
-\end{itemize}
-
-\section{Impact of network structure}\label{netw_struc}
-
-In this section we will discuss some results related to the network structure and network topologies of the different network models that are used in this thesis.
-We will mainly focus on the effect of a high clustering coefficient and a community structure.
-Two different opinion distributions are investigated
-\begin{itemize}
-	\item An initial $50/50$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.5$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.5$.
-	\item An initial $20/80$ opinion distribution: the initial fraction of nodes with opinion 0 is equal to $P_0(0) = 0.2$ and the initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$.
-Hence, we are dealing with a minority and a majority opinion.
-\end{itemize}
-For both opinion distributions we investigate the evolution of the opinion fractions with time and the formation of echo chambers.
+This chapter is subdivided into four main sections.
+We first discuss the impact of the network structure on the evolution of the opinions and the formation of echo chambers.
+Then, we introduce some kind of stubbornness in the system.
+The impact of how the opinions/stubborn people are initially distributed in a network with a community structure (SBM) will be discussed afterwards.
+Finally, we will run the opinion dynamics model of this thesis on some real-world networks.
+We will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
+
+\section{Network structure}\label{netw_struc}
+
+In this section we will discuss the results related to the network topologies of the different network models that are used in this thesis.
+We will focus on the effect of a high clustering coefficient and a community structure.
+Two different opinion distributions are investigated, namely balanced starting conditions and unbalanced starting conditions.
+For both opinion distributions we investigate the evolution of the prevalence of the opinions with time and the formation of echo chambers.
 Two filtering algorithms will be compared to each other, namely the REC and the PR filtering algorithm, in order to investigate the possible interplay between those filtering algorithms and the network structures.\\
-\newline
-Before sharing the results, a quick disclaimer: Some of the results that are shared in this section, were already obtained by prof.
+%\newline
+%Before sharing the results, a quick disclaimer: Some of the results that are shared in this section, were already obtained by prof.
 Nicola Perra and prof.
@@
 
-\subsection{Starting conditions $P_0(0) = 0.5$ and $P_1(0)=0.5$}
+\subsection{Balanced starting conditions}
 \label{50-50}
-We start with the following opinion distribution: a fraction of nodes with opinion 0 equal to 0.5 ($P_0(0) = 0.5$) and a fraction of nodes with opinion 1 equal to 0.5 ($P_1(0) = 0.5$).
+In this configuration $50 \%$ of the nodes start with opinion A ($P_{\text{A}}(0) = 0.5$) and $50 \%$ of the nodes start with opinion B ($P_{\text{B}}(0) = 0.5$).
 The opinions are randomly distributed over the network.
-The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} % correct section??
-is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
-\newline
-Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
-\begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-	\item The Watts-Strogatz model (WS) with mean degree $K=10$; two different values of the rewire probability $\beta$ are used, each of them has a slightly different value of the average clustering coefficient, namely $\beta = 0.01$ with $\left<C\right> \sim 0.60$ and $\beta = 0.06$ with $\left<C\right> \sim 0.55$.
-This model is used to investigate the effect of clustering.
-	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
-This model is used to investigate the effect of a community structure.
-	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} is used.
+We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
+\newline
+Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
+The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
+To investigate the effect of clustering on the dynamics, we use the Watts-Strogatz model (WS).
+Two different values of the rewire probability $\beta$ are used, each with a slightly different value of the clustering coefficient $\left<cc\right>$: $\beta = 0.01$ with $\left<cc\right> = 0.648 \pm 0.003$ and $\beta = 0.06$ with $\left<cc\right> = 0.557 \pm 0.007$.
+The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
+Two different modularities are investigated: a high modularity $Q = 0.908 \pm 0.005$ with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
-\end{itemize}
-\textbf{Evolution of the opinion fractions}\\
+This is a network with both a high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and a high modularity ($Q = 0.909 \pm 0.004$).
+\\
+\newline
+\textbf{Evolution of the prevalence of the opinions}\\
 \newline
 We start with the results for the evolution of the opinion fractions in case of an initial $50/50$ opinion distribution.
-Prof.
-N.
-Perra and prof.
-Luis E.C.
-Rocha already showed that, for the case of balanced initial opinions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the different filtering algorithms \cite{Perra2019}.
-In Figure \ref{ev_op_50_50} the results for the network model with a community structure (SBM, Figure \ref{sbm_50-50}) and the network model with both a community structure and a high clustering (SBM-WS, Figure \ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithm.
-Results are shown together with their standard deviation $\sigma$.
-From this figure it becomes clear that having a community structure is also not able to break the status-quo for both of the filtering algorithms.
+Perra and Rocha showed that, for the case of balanced initial opinions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the different filtering algorithms \cite{Perra2019}.
+The results for the network model with community structure (SBM, Figure \ref{sbm_50-50}) and the network model with both community structure and high clustering (SBM-WS, Figure \ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithms.
+A community structure is also not able to break the status quo for both of the filtering algorithms.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
-    \caption{SBM}
     \label{sbm_50-50}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10x100_10-001-0001_50-50_PR_REC.png}
-    \caption{SBM-WS}
     \label{sbm-ws_50-50}
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
-Two different network models with a community structure are represented, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+  \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ distribution.
+Two different network models with a community structure are represented, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
+Each curve is the average of $10^2$ independent simulations and data points are shown, together with their standard deviation $\sigma$, for every 50 time steps to improve visualization.}
 \label{ev_op_50_50}
@@
 \newline
-Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$.
-The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0).
-Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1.
+Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
+The value shown on the $y$-axis, $F_N(\left<P_{\text{B}}^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
+The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion B (these are thus neighborhoods in which most users adopt opinion A).
+Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion B.
 The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
-Error bars are omitted for the sake of an improved visualization.\\
-These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\
+Error bars are omitted for the sake of an improved visualization.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC.png}
-    \caption{REC}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+      \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR_8x7.png}
-    \caption{PR}
   \end{subfigure}
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-\textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$).
-The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$.
-Each plot is the average of $10^2$ independent simulations.}
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; The Watts-Strogatz model (WS) with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
+The starting conditions are $P_{\text{A}}(0) = 0.5$ and $P_{\text{B}}(0) = 0.5$.
+Each curve is the average of $10^2$ independent simulations.}
 \label{echo_50_50}
@@
 \noindent
-In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR method.
-The results for both filtering algorithms look qualitatively the same.
-The main difference is that the formation of echo chambers, using the PR method, is about ten times as big compared to the formation of echo chambers using the REC method.
-\\
-The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a very limited, sign of echo chambers, for both the REC method and the PR method.
+In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR filtering algorithms.
+The results for both filtering algorithms are qualitatively the same.
+The main difference is that the formation of echo chambers, using the PR filtering algorithm, is about ten times as big compared to the formation of echo chambers using the REC filtering algorithm.
+\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a very limited, sign of echo chambers, for both the REC filtering algorithm and the PR filtering algorithm.
 Networks with a high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
-Going to slightly higher values of the clustering coefficient ($\left<C\right> \sim 0.55$ to $\left<C\right> \sim 0.60$) in the Watts-Strogatz model results in significantly larger echo chambers.
+Going to slightly higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the Watts-Strogatz model results in significantly larger echo chambers.
 This suggests that the formation of echo chambers is really sensitive to changes in the clustering coefficient.
@@
 \\
-Note that for an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (see Figure \ref{ev_op_50_50} and related results in reference \cite{Perra2019}).
+\newline
+For an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (Fig.
+\ref{ev_op_50_50} and related results in reference \cite{Perra2019}).
 Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a more clustered opinion distribution and thus lead to polarization and formation of echo chambers.
 
-\subsection{Starting conditions $P_0(0) = 0.2$ and $P_1(0)=0.8$}\label{20-80}
-
-The initial fraction of nodes with opinion 0 is now equal to $P_0(0) = 0.2$, opinion 0 thus becomes the minority opinion.
-The initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$, hence this is the majority opinion.
-The opinions are, initially, randomly distributed over the system.
-Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
-\newline
-Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
-\begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$.
-The value of the average clustering coefficient is equal to $\left<C\right> \sim 0.55$.
-This model is used to investigate the effect of clustering.
-	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
-This model is used to investigate the effect of a community structure.
-	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+\subsection{Unbalanced starting conditions}\label{20-80}
+
+We will now investigate the effect of unbalanced starting conditions, that is one opinion is initially more prevalent than the other.
+The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$, opinion A thus becomes the minority opinion.
+The initial prevalence of opinion B is $P_\text{B}(0) = 0.8$, hence this is the majority opinion.
+The opinions are, initially, randomly distributed over the system.\\ %Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
+\newline
+We look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
+The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
+To investigate the effect of clustering on the dynamics, we use the Watts-Strogatz model (WS).
+The rewire probability is $\beta = 0.06$, which gives a clustering coefficient equal to $\left<cc\right> = 0.557 \pm 0.007$.
+The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
+Two different modularities are investigated: a high modularity $Q = 0.908 \pm 0.005$ with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
-\end{itemize}
-\textbf{Evolution of the opinion fractions}\\
-\newline
-Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion 0, which is the minority opinion in this case, over time, together with the standard deviation $\sigma$.
-For each network model two filtering algorithms are compared, namely the REC filtering algorithm and the PR filtering algorithm.
-\\
-The REC method is not able to break the prevalence of opinion 0 over time for neither of the four different network models.
-When using the PR method, the results look very different.
-The prevalence of the minority opinion, $P_0(t)$, decreases over time.
-This decrease is a direct consequence of the PR filtering algorithm.
-Since the PR method orders the posts according to the users' preferences, nodes with the majority opinion 1 will mainly get to see posts with the similar opinion 1 on their time-line, which limits the chance of changing their opinion to the minority opinion.
-Nodes with the minority opinion 0, on the other hand, will obviously see posts with opinion 0 on their time-line when using the PR method.
-However, since there are more nodes with opinion 1 than nodes with opinion 0, the nodes with the minority opinion 0 will also get to see a significant number of posts with the majority opinion 1 on their time-line.
-The conclusion is that, when using the PR method, there is an imbalance in which opinion can switch to which: nodes with the minority opinion 0 have a chance of switching to the majority opinion 1, whereas nodes with the majority opinion 1 have (almost) no chance of switching to the minority opinion 0.
-This imbalance leads to the decrease of the minority opinion 0.\\
-\newline
-From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist, $P_0(500) \sim 0$.
-This is certainly the case for the low clustering networks (ER and SBM).
-The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_0(500) \sim 0.007$ for the SBM-WS network (Figure \ref{sbm-ws_20-80_op}) and an average value of $P_0(500) \sim 0.009$ for the WS network (Figure \ref{ws_20-80_op}).\\
-This result differs from the conclusion that prof.
-N.
-Perra and prof.
-L.E.C.
-Rocha made in the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
-They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}.
-The results that they obtained are, however, for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
+This is a network with both a high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and a high modularity ($Q = 0.909 \pm 0.004$).\\
+\newline
+\textbf{Evolution of the prevalence of the opinions}\\
+\newline
+Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion A, which is the minority opinion, over time.
+For each network model two filtering algorithms are compared, the REC filtering algorithm and the PR filtering algorithm.
+\\
+\newline
+The REC filtering algorithm maintains the status quo; the prevalence of opinion A remains stable over time for the four different network models.
+When using the PR filtering algorithm, the results are very different.
+The prevalence of the minority opinion, $P_\text{A}(t)$, decreases over time.
+This decrease is a consequence of the PR filtering algorithm.
+Since the PR filtering algorithm orders the posts according to the users' preferences, nodes with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
+Nodes with the minority opinion A, on the other hand, will obviously see posts with opinion A on their time-line when using the PR filtering algorithm.
+However, since there are more nodes with opinion B than nodes with opinion A, the nodes with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
+When using the PR filtering algorithm, there is an imbalance in the sense that nodes with the minority opinion A have a chance of switching to the majority opinion B, whereas nodes with the majority opinion B have (almost) no chance of switching to the minority opinion A.
+This imbalance leads to the decrease of the minority opinion A.\\
+\newline
+Figure \ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
+This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
+The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Figure \ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Figure \ref{ws_20-80_op}).\\
+\newline
+This result differs from reference \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
+The results that where obtained in reference \cite{Perra2019} are, however, for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
 In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
@@
 The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
-This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results (together with the standard deviation $\sigma$) for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
-In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.
+This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the prevalence of opinion A for an initial $20/80$ opinion distribution and using the PR filtering algorithm; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
+In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the maximum length of the actual time-line is $S=20$.
+In this figure we can see that the minority opinion A reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with a value of $P_\text{A}(500) = 0.027 \pm 0.023$ in Figure \ref{ws1000} and a value of $P_\text{A}(500) = 0.030 \pm 0.018$ in Figure \ref{ws10000}.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+      \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
-    \caption{WS}
     \label{ws_20-80_op}
@@
   \begin{subfigure}[b]{0.49\textwidth}    
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
-    \caption{SBM-WS}
     \label{sbm-ws_20-80_op}
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+The bars represent the standard deviation $\sigma$.}
 \label{ev_op_20_80}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{$N = 10^3$}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
-    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
     \label{ws1000}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{$N = 10^4$}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
-    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
     \label{ws10000}
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution.
-The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$.
-In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$.
+The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
+The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$ and \textbf{(b)} $N=10^4$.
+In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the PR filtering filtering algorithm is used.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}
 \label{ev_op_WS_k=6}
@@
 Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
-For each network, the REC and PR method are compared to each other.
+For each network, the REC and PR filtering algorithm are compared to each other.
 The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.
@@
 This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
+\newline
 There are no visible effects of having a clear community structure.
@@
 Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
-When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered.
+When using the REC filtering algorithm, the formation of echo chambers around the majority opinion 1 is hampered.
 This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
 There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
-From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions.
+From figure \ref{ev_op_20_80} it was already concluded that the REC filtering algorithm doesn't break the prevalence of the opinions.
 The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
 \\
-The PR method shows a completely different tendency.
+\newline
+The PR filtering algorithm shows a completely different tendency.
 There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1.
-This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
+This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
 This reduced visibility of the minority opinion 0 in these neighborhoods drives them eventually to adopt opinion 1, with an increase in the number of echo chambers around the majority opinion 1 as a consequence.
@@
 Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
-For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
 An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
@@
 If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
+\newline
 This enhancing of the formation of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
-Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks.
+Remember from Figure \ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks.
 Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
@@
 This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution.
-In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+In case of the PR filtering algorithm this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
-    \caption{ER}
     \label{er_echo_20-80}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
-    \caption{WS}
     \label{ws}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
-    \caption{SBM}
   	\label{sbm}    
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
-    \caption{SBM-WS}
     \label{sbm-ws}
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
 The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
-Each plot is the average of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.}
 \label{echo_20_80}
@@
 \\
-\newpage
-\noindent
+\newline
 In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
@@
 The PR filtering algorithm is compared to the REC filtering algorithm.
-For most values of the modularity, the REC method produces almost no echo chambers.
+For most values of the modularity, the REC filtering algorithm produces almost no echo chambers.
 If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
-For the PR method, the increase in the size of the echo chambers with increasing modularity is much more significant.
+\newline
+For the PR filtering algorithm, the increase in the size of the echo chambers with increasing modularity is much more significant.
 From a modularity approximately equal to $Q \sim 0.5$, we can see a sharp increase in the formation of echo chambers.
-For modularity values smaller than $Q \sim 0.5$, the formation of echo chambers remains very small, even when using the PR method.\\
+For modularity values smaller than $Q \sim 0.5$, the formation of echo chambers remains very small, even when using the PR filtering algorithm.\\
+\newline
 These results strongly suggest that a community structure enhances the formation of echo chambers in the system, mostly when using the PR filtering algorithm.
 A quite clear community structure, represented by a high modularity, is however needed to really see the effect of this community structure on the formation of opinion clusters.
-\newpage
+
 \section{Impact of stubborn people}
+
 In this section, the impact of introducing stubborn people will be reviewed. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
@@
 	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<C\right> \sim 0.55$.
+	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<cc\right> \sim 0.55$.
 This model is used to investigate the effect of clustering.
@@
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
+This is a network with both a high clustering ($\left<cc\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
 \end{itemize}
-For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.
+For each network model, we compare the REF, REC and PR filtering algorithms to each other to analyze the possible interplay between stubbornness and filtering algorithm.
 
@@
 \\
-Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.
-
-\newpage
-\noindent
+\newline
+Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
+\newline
 \textbf{Making a fraction of the nodes completely stubborn}\\
@@
 In all cases, the stubborn nodes are randomly distributed over the network.\\
+\newline
 We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
+Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
 Each data point is the result of $10^2$ independent simulations.}
@@
 Data points are again shown together with their standard error, $SE$, for the same reasons as explained in Section \ref{echoVSmod}.\\
+\newline
 The Erd\H{o}s-R\'{e}nyi model (ER) shows no appearance of echo chambers.
@@
 Note that the data points for the SBM with a low modularity are not visible in Figure \ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
+\newline
 The other three networks (WS, SBM-WS and SBM with a high modularity) all show qualitatively the same behavior.
@@
 This in turn leads to the hampering of the formation of echo chambers.\\
+\newline
 The WS and SBM-WS networks are the high clustering networks and for a small fraction of stubborn people they show the largest echo chambers.
@@
 This leads us to make the same conclusion as before: a community structure enhances the formation of echo chambers compared to a random network, however not as significant as networks with a high clustering coefficient.
-\\
 \begin{figure}[H]
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REF}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
-    \caption{REF}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
-    \caption{REC}
     \label{REC_frac_compl_stubb}
@@
 Data points are shown together with their standard error $SE$.
-The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
+The results for the REF and REC filtering algorithms behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
 Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease.
@@
 The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\
+\newline
 These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn agents enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
@@
 \newline
-This discrepancy between the results for the PR method and those for the REF/REC method in combination with a small fraction of stubborn nodes can be potentially explained with the help of the following simple example.
+This discrepancy between the results for the PR filtering algorithm and those for the REF/REC filtering algorithms in combination with a small fraction of stubborn nodes can be potentially explained with the help of the following simple example.
 Assume we have three nodes that form a triangle.
-Two nodes initially carry opinion 0 and one node carries opinion 1.\\
-Lets first discuss the PR method and assume that none of the nodes are stubborn.
+Two nodes initially carry opinion 0 and one node carries opinion 1.
+Lets first discuss the PR filtering algorithm and assume that none of the nodes are stubborn.
 In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the nodes adopt the same opinion).
-This can be easily seen: the one node with opinion 1 is likely to switch to opinion 0, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR method wants to promote posts with opinion 1 does not matter here, since there are no neighbors with opinion 1).
-The two nodes with opinion 0 will have more chance of keeping that opinion than switching to opinion 1 in case of the PR method.
-This is because the PR method will promote the posts of the friend with the same opinion 0 over the posts of the friend with opinion 1.
+This can be easily seen: the one node with opinion 1 is likely to switch to opinion 0, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion 1 does not matter here, since there are no neighbors with opinion 1).
+The two nodes with opinion 0 will have more chance of keeping that opinion than switching to opinion 1 in case of the PR filtering algorithm.
+This is because the PR filtering algorithm will promote the posts of the friend with the same opinion 0 over the posts of the friend with opinion 1.
 The end result will be that there is a high chance that all the three nodes will adopt opinion 0 and hence, form an echo chamber.
@@
 However, the two nodes with opinion 0 still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion 0, and hence reduces the visibility of the posts of the stubborn friend with opinion 1.
-The system will now not form an echo chamber, since the stubborn node with opinion 1 cannot switch to opinion 0 and the PR method reduces the chance of changing the opinion of the nodes with opinion 0 to opinion 1.
-The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn people.\\
+The system will now not form an echo chamber, since the stubborn node with opinion 1 cannot switch to opinion 0 and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion 0 to opinion 1.
+The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn people.
 What happens if we use the REF/REC filtering algorithm?
@@
 Assume first that none of the nodes are stubborn.
-For the PR method, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber.
-For the REF/REC method this chance is however smaller, which can be seen as follows.
+For the PR filtering algorithm, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber.
+For the REF/REC filtering algorithms this chance is however smaller, which can be seen as follows.
 The node with opinion 1 still has a high chance of switching to opinion 0, because all his friends (which are the two other nodes in the triangle) carry this opinion.
-However, the two nodes with opinion 0 will now not have a higher chance of keeping this opinion than to switch to opinion 1, as was the case for the PR method.
+However, the two nodes with opinion 0 will now not have a higher chance of keeping this opinion than to switch to opinion 1, as was the case for the PR filtering algorithm.
 Instead, they have an equal chance of keeping their opinion as to change to the other opinion.
@@
 Both opinion 0 and opinion 1 have an equal chance of being promoted on the time-line of the nodes with opinion 0.
-This reduced tendency of the nodes with opinion 0 to not change their opinion (compared to using the PR method) hampers the ability of the triangle to form an echo chamber.
-This would explain why, in case of no stubborn nodes, the REC/REF method shows a lower formation of echo chambers compared to the PR method.
+This reduced tendency of the nodes with opinion 0 to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
+This would explain why, in case of no stubborn nodes, the REC/REF filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
 Now, lets again assume that we make the node with opinion 1 stubborn.
 In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle.
-For the REC/REF method, things are different.
+For the REC/REF filtering algorithms, things are different.
 The stubborn node with opinion 1 will obviously not change opinion.
@@
 This would explain the increase in the formation of echo chambers for a small fraction of stubborn people when using the REF or REC filtering algorithms.
-\\
 If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease again which can also be easily seen with the help of the triangle example.
@@
 Are these affected by introducing stubborn people in the system?
-\\
 The short answer is no.
@@
 Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg.
-an initial $20/80$ distribution.\\
-\newline
+an initial $20/80$ distribution.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{PR}
     \label{PR_frac_compl_stubb_20-80}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
-    \caption{REC}
     \label{REC_frac_compl_stubb_20-80}
@@
 \\
+\newline
 The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
@@
 For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
-\newpage
-\noindent
-We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
+\newline
+We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC filtering algorithm).
 \\
@@
 Lets now discuss the evolution of the opinions in the system.
-In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC method does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
 The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion.
-Are these results affected by introducing stubborn agents?\\
+Are these results affected by introducing stubborn agents?
 As for the balanced initial conditions, introducing stubborn people does not break the prevalence of the initial opinions for the REC filtering algorithm.
 Introducing stubborn people in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
+\newline
 Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
@@
 \\
+\newline
 If there are no stubborn people, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
@@
 This result is the same for all the different network topologies.
-If there are no stubborn people, the PR method forces nodes with the minority opinion to switch to the majority opinion.
+If there are no stubborn people, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
 If we however introduce a fraction of stubborn people in the system, then we will also end up with some stubborn nodes that carry the minority opinion.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
-    \caption{SBM-WS}
   \end{subfigure}
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
 \label{ev_op_20_80_stubb_PR}
 \end{figure}
-\newpage
 \noindent
@@
 All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
+\newline
 We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where-after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
@@
 The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod} and the data points are shown together with their standard error $SE$.
-The results look very similar to the case where a fraction of the nodes is completely stubborn and when using the PR method, which is depicted in Figure \ref{echo_vs_frac_complStubb}.\\
+The results look very similar to the case where a fraction of the nodes is completely stubborn and when using the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_frac_complStubb}.\\
+\newline
 The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers, as before, and this is not changed by making the nodes more and more stubborn.
 The same result is obtained for the stochastic block model with a low modularity (the data points lie again behind those of the ER network).\\
+\newline
 The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
-If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.\\
+If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
 This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
@@
 Data points are again displayed with their standard error $SE$.\\
+\newline
 We can make similar observations as we did in the ``Making a fraction of the nodes completely stubborn'' paragraph.
-The results for the REF and REC method look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.\\
+The results for the REF and REC filtering algorithms look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
 The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases.
-For the REF method, this increases reaches a maximum around $r \sim 0.7$ and then starts to slightly decrease between $r\sim 0.7$ and $r\sim 0.9$.
+For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slightly decrease between $r\sim 0.7$ and $r\sim 0.9$.
 Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers (echo chamber values around 17 for $r\sim 0.9$ to echo chamber values equal to one for $r=1$).
-The REC method also seems to have a maximum around $r \sim 0.7$ for the WS model, the SBM-WS model on the other hand displays a continuous increase until $r \sim 0.9$.
+The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model, the SBM-WS model on the other hand displays a continuous increase until $r \sim 0.9$.
 In both case we have again a sharp drop from $r \sim 0.9$ to $r=1$.
@@
 We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithm in the case of networks with a high clustering coefficient.
-This is the complete opposite of the case where the PR method is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
-The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the ``Making a fraction of the nodes completely stubborn'' paragraph.\\
+This is the complete opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
+The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the ``Making a fraction of the nodes completely stubborn'' paragraph.
 If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REF}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REF_8x7.png}
-    \caption{REF}
     \label{REF_all_frac_stubb}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REC_8x7.png}
-    \caption{REC}
     \label{REC_all_frac_stubb}
@@
 Figure \ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Figure \ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
-The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{PR}
     \label{PR_all_frac_stubb_20-80}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_REC_20-80_8x7.png}
-    \caption{REC}
     \label{REC_all_frac_stubb_20-80}
@@
 This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
+\newline
 As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
@@
 \newline
-The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC method (see Figure \ref{REC_all_frac_stubb}).
+The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC filtering algorithm (see Figure \ref{REC_all_frac_stubb}).
 The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to a stubbornness parameter $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
@@
 This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
+\newline
 The stochastic block model with a high modularity also shows a slight increase in the formation of echo chambers up to $r = 0.8$.
@@
 Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
+\newline
 The results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm are not affected by making all the nodes more and more stubborn.
-The prevalence of the opinions is maintained an the status quo is not broken.\\
+The prevalence of the opinions is maintained an the status quo is not broken.
 The PR filtering algorithm gives us another story.
 The results for the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
-If the nodes are not stubborn, the PR method strongly reduces the minority opinion over time.
+If the nodes are not stubborn, the PR filtering algorithm strongly reduces the minority opinion over time.
 Making the nodes more and more stubborn reduces this tendency.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{ER}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{WS}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{SBM}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
-    \caption{SBM-WS}
   \end{subfigure}
@@
   \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
-Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
 \label{ev_op_20_80_all_frac_stubb_PR}
@@
 The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.\\
+\newline
 The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat for a threshold parameter smaller than $T \leqslant 0.5$.
@@
 The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
+\newline
 Similar observations can be made for the stochastic block model with a high modularity.
@@
 % make coupling to ev of op??
-\newpage
 \begin{figure}[H]
@@
 The initial opinion distribution was again $50/50$.
-The results for the REF and REC method strongly resemble each other.
+The results for the REF and REC filtering algorithms strongly resemble each other.
 Data points are again shown together with their standard deviation $\sigma$ to really emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 100$).
@@
 After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion 1 (and thus resulting in large echo chambers around opinion 1) and there will have been simulations that converge to opinion 0 (resulting in almost no echo chambers around opinion 1).\\
+\newline
 Another consequence of this converging to one of the two opinions every single simulation, that can be observed in Figure \ref{echo_vs_threshold_REF-REC}, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
@@
 As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes kind of a huge, single echo chamber).
-The PR method seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
+The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
 This can be clearly seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF methods, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF filtering algorithms, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
 If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an impact on the system.
@@
 The threshold parameter is set equal to $T=0$, so there is no effect of any type of stubbornness.\\
-From this figure it becomes clear that the PR method strongly hampers the convergence to a single opinion each independent simulation.
-It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC method.
+\newline
+From this figure it becomes clear that the PR filtering algorithm strongly hampers the convergence to a single opinion each independent simulation.
+It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
 This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
 Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Figure \ref{sbm_majT}.
-For the REC method, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
-For the PR method, this is not really the case, but this is probably due to the fact that the PR method already strongly hampers the convergence to a single opinion; having a high or low modularity does not really give an additional effect on top of this PR filtering effect.
+For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
+For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not really give an additional effect on top of this PR filtering effect.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REF}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REF_8x7.png}
-    \caption{REF}
     \label{REF_thres}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REC_8x7.png}
-    \caption{REC}
     \label{REC_thres}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_majT=0_one_sim_8x7.png}
-    \caption{ER}
     \label{er_majT}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_majT=0_one_sim_8x7.png}
-    \caption{WS}
     \label{ws_majT}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_majT=0_one_sim_8x7.png}
-    \caption{SBM}
     \label{sbm_majT}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_majT=0_one_sim_8x7.png}
-    \caption{SBM-WS}
     \label{sbm-ws_majT}
@@
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$.
+Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.}
@@
 \end{figure}
-\newpage
+
 \section{Random versus community distributed opinions/stubborn agents}
@@
 This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
+\newline
 We will look at a network with a high modularity, representing the case of more or less isolated communities, and at a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
@@
 \\
+\newline
 Both the results for the evolution of the opinions in the system and the formation of echo chambers will be discussed.\\
-\newpage
-\noindent
+\newline
 \textbf{Evolution of the opinion fractions}\\
@@
 For each modularity the REC and PR filtering algorithms are compared to each other.\\
+\newline
 In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
@@
 This result is the same for both the high and low modularity case, see Figures \ref{high_mod_rec} and \ref{low_mod_rec}.\\
+\newline
 The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced initial opinions, if the opinions are randomly distributed.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{High modularity, REC}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_high_mod_8x7.png}
-    \caption{High modularity, REC}
     \label{high_mod_rec}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{High modularity, PR}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_high_mod_8x7.png}
-    \caption{High modularity, PR}
     \label{high_mod_pr}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+   \caption{Low modularity, REC}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_low_mod_8x7.png}
-   \caption{Low modularity, REC}
     \label{low_mod_rec}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{Low modularity, PR}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_low_mod_8x7.png}
-    \caption{Low modularity, PR}
     \label{low_mod_pr}
@@
 \textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
-Each plot is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
+Each curve is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
 \label{commOp0_01_other_50-50}
@@
 Figure \ref{pr_comm} shows those for the PR filtering algorithm.\\
+\newline
 The higher the fraction of single opinion 0 communities, the higher the initial average fraction of opinion 0.
@@
 This result is valid for both the high and low modularity case.\\
+\newline
 The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the fraction of opinion 0, which is the majority opinion, over time for a fixed fraction of single opinion 0 communities.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_REC_8x7.png}
-    \caption{REC}
     \label{rec_comm}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_PR_8x7.png}
-    \caption{PR}
     \label{pr_comm}
@@
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each plot is the result of $10^2$ independent simulations.}
+Each curve is the result of $10^2$ independent simulations.}
 \label{fracOp0_vs_commOp0_other_50-50}
 \end{figure}
-\newpage
 \noindent
@@
 \newline
-Lets first discuss the results for the REC method.
+Lets first discuss the results for the REC filtering algorithm.
 The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0 (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line).
@@
 The increase remains however smaller than two, which indicates that there are less than twice as much echo chambers at $t=500$ than at $t=0$.\\
+\newline
 The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion 0, which could be observed for the community distributed, high modularity case.
@@
 \\
-\newpage
-\noindent    % make link to opinion fraction 
+\newline    % make link to opinion fraction 
 The results for the PR filtering algorithm are depicted in Figure \ref{pr_comm_echo}.
 They are quite different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Figure \ref{rec_comm_echo}.\\
+\newline
 Lets first focus on the high modularity network.
@@
 We can observe a strong increase in the formation of echo chambers around the majority opinion 0, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion 1.
-Remember from Figure \ref{high_mod_pr} that the PR method leads to an increase in the majority opinion 0.
+Remember from Figure \ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion 0.
 If we, however, make a fraction equal to $0.1$ of the communities single opinion 0 communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion 0 is strongly hampered.
@@
 This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
+\newline
 The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_REC_8x7.png}
-    \caption{REC}
     \label{rec_comm_echo}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_PR_8x7.png}
-    \caption{PR}
     \label{pr_comm_echo}
@@
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each plot is the result of $10^2$ independent simulations.}
+Each curve is the result of $10^2$ independent simulations.}
 \label{echo_01_commOp0_other_50-50}
@@
 The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion 0 (Figure \ref{echo_01_commOp0_other_50-50}).\\
+\newline
 The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion 0 with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
@@
 \\
+\newline
 The low modularity, community distributed case seems to wiggle around $1$.
@@
 The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion 0 communities, while the others get a $50/50$ opinion distribution.\\
+\newline
 The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Figure \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
@@
 They show a strong increase in the formation of echo chambers around the majority opinion 0.\\
+\newline
 The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion 0, which was also observed in Figure \ref{pr_comm_echo}.\\
+\newline
 The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion 0.
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
-    \caption{REC}
     \label{rec_comm05_echo}
@@
   \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
-    \caption{PR}
     \label{pr_comm05_echo}
@@
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each plot is the result of $10^2$ independent simulations.}
+Each curve is the result of $10^2$ independent simulations.}
 \label{echo_05_commOp0_other_50-50}
@@
 Furthermore, a notion of stubbornness was introduced to investigate the interplay between this individual resistance to change, network structure and filtering algorithm.
-For simplicity we only allowed two competing opinions (0 and 1) in the system.\\
+For simplicity we only allowed two competing opinions (A and B) in the system.\\
 \newline
@@
 This is mainly seen in the high clustering networks, however a tiny bump can also be observed in the high modularity network.
-This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen in both the case of balanced initial conditions ($50/50$) and unbalanced initial conditions ($20/80$).\\
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen in both the case of balanced initial conditions and unbalanced initial conditions.\\
 % talk about evolution of opinions??
@@
 This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.
-This was done for both a high modularity and a low modularity network and for both the REC and PR filtering algorithms.\\
-The REC method did not show any difference in the evolution of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.\\
+This was done for both a high modularity and a low modularity network and for both the REC and PR filtering algorithms.
+\\
+\newline
+The REC filtering algorithm did not show any difference in the evolution of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.
 For the PR filtering algorithm it was found that the high modularity network hampers the increase in the majority opinion in the community distributed scenario compared to the randomly distributed scenario.
@@
 \begin{equation}
-	\left<C\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)} \ .
+	\left<cc\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)} \ .
 \end{equation}

### 2021-04-29 draft.tex
latest update

@@
 \\
-Beside that, we will also pay some attention to the possible ways of distributing the opinions and the stubborn people and how this may effect the opinion dynamics in the system. \\
+Beside that, we will also pay some attention to the possible ways of distributing the opinions and the stubborn actors and how this may effect the opinion dynamics in the system. \\
 Finally, we will test the model on real-world networks.
@@
 \selectlanguage{english}
-\newgeometry{top=25mm, bottom=25mm}
+\newgeometry{top=20mm, bottom=25mm}
 \tableofcontents
@@
 \addcontentsline{toc}{chapter}{List of Figures}
+\listoftables
+\addcontentsline{toc}{chapter}{List of Tables}
 \restoregeometry
@@
 However, we still want our models to display properties similar to the ones seen in real systems \cite{Mata2020}.
-Since many real systems are not static but evolve in time, this means that we don't only need to deal with static networks but also with temporal networks.
+Since many real systems are not static but evolve in time, this means that we do not only need to deal with static networks but also with temporal networks.
 Temporal networks are networks where the edges are not always active, but instead become active for some periods of time \cite{Holme2012}.
@@
 Furthermore, the agents can have a parameter that determines whether they are active or not.
-The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
+The introduction of stubborn actors will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
 \newline
@@
 As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
-On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network.
+On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn actors in the network.
 Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\
@@
 In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.
-They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure.
+They, however, did not investigate the effect of stubborn actors in the network, nor did they investigate networks with a community structure.
 This thesis will build on their model and incorporate these new features.
@@
 \newline
-On the other hand, we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network.
+On the other hand, we want to determine the impact of stubborn actors on the formation and evolution of opinions in the network.
 Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness?
-What about the possible interplay between the filtering algorithms and the introduction of stubborn people?
-Does the distribution of stubborn people play a role (eg.
+What about the possible interplay between the filtering algorithms and the introduction of stubborn actors?
+Does the distribution of stubborn actors play a role (eg.
 randomly distributed versus clustered together inside communities)?
@@
 Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
-They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Barabasi2016}\cite{Mata2020}.
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree is not a characteristic scale for the network \cite{Barabasi2016}\cite{Mata2020}.
 
@@
 The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}.
-The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it.
+The density hypothesis narrows what could be considered a community, but it does not uniquely define it.
 Several community definitions are consistent with the density hypothesis.
@@
 A community based on this definition would then be the largest clique in the network.
-This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\
+This definition of community might, however, be too restrictive and, beside that, large cliques do not appear very frequent in networks \cite{Albert2016}.\\
 \newline
@@
 It should be noted that not all real-world networks have a heterogeneous degree distribution.
-And even if the degree distribution is heterogeneous or fat-tailed, it doesn't always need to follow a power-law.
+And even if the degree distribution is heterogeneous or fat-tailed, it does not always need to follow a power-law.
 The degree distribution can eg.
@@
 In real life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}.
-The `ordinary people' don't have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
+The `ordinary people' do not have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
 In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance/influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
@@
 \newline
-Finally, in Section \ref{stubb}, the introduction of stubborn people is discussed.
+Finally, in Section \ref{stubb}, the introduction of stubborn actors is discussed.
 
@@
 They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
-If social media companies woudn't use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
+If social media companies would not use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
 This in turn would make the on-line social experience a lot more unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
@@
 This kind of activation mechanism of the users gives the network a more temporal character.
-The underlying network structure doesn't change (in the sense that each user/node has a fixed number of neighbors that doesn't change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't.
-Hence, in this sense, this does affect the network structure in some way, because if a user, for example, doesn't become active for a long period of time, he/she doesn't participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
+The underlying network structure does not change (in the sense that each user/node has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
+Hence, in this sense, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
 This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity, clustering coefficient, etc may behave differently.
@@
 
-\section{The presence of stubborn people}\label{stubb}
+\section{The presence of stubborn actors}\label{stubb}
 
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion doesn't necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme does not need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
 \newline
 Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups a lot bigger.
-Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people.
+Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore would not reach too many people.
 So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
@@
 Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models.
-There are, however, many other ways to include some kind of stubborn people in these models.
+There are, however, many other ways to include some kind of stubborn actors in these models.
 In the following, two different possibilities will be discussed.
@@
 \newline
-One way of creating stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
+One way of creating stubborn actors in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
 This resistance parameter can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
@@
 One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
-Many other possibilities can be found according to how the stubborn people are distributed and according to how stubborn they are made.
+Many other possibilities can be found according to how the stubborn actors are distributed and according to how stubborn they are made.
 This makes the model very flexible.\\
@@
 If $T = 0$, this model reduces to the pure majority model described in Section \ref{majModel}.
-If $T = 1$, the nodes don't change their opinion.
+If $T = 1$, the nodes do not change their opinion.
 For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
@@
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
-The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that doesn't display both of these features and is hence the perfect comparison model.
+The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the perfect comparison model.
 \\
@@
 In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}) Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
-Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn people and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
 We will investigate whether a community structure behaves similar to a high clustering regarding the appearance of echo chambers.\\
@@
 All the above models have a homogeneous degree distribution.
-In this thesis, we didn't focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
+In this thesis, we did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
 It might, however, be an interesting extension to investigate such types of models as well.\\
@@
 This plus side is, however, also one of the possible downsides.
-Real networks aren't simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
+Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
 In other words, network models may lack many features of real-world networks.
@@
 		\Else
-			\State \(\text{op} \leftarrow \text{current opinion}\) \Comment{Don't update opinion of active user}
+			\State \(\text{op} \leftarrow \text{current opinion}\) \Comment{Do not update opinion of active user}
 		\EndIf
@@
 Real-world opinion dynamics are, probably, far from simple.
-People don't update their opinion according to a simple probabilistic opinion dynamics model.
+People do not update their opinion according to a simple probabilistic opinion dynamics model.
 Beside that, the cases where people have to choose between only two opinions are rather limited.
@@
 \newline
-This doesn't mean that simple models should be thrown away.
+This does not mean that simple models should be thrown away.
 They might still be capable of capturing a lot of the dynamics that are seen in the real world.
@@
 This has been proven by physics research over and over again.
-Beside that, if you don't know the dynamics/underlying processes of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
+Beside that, if you do not know the dynamics/underlying processes of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
 There is no need to make things too complex if simple works too.
@@
 Then, we introduce some kind of stubbornness in the system.
-The impact of how the opinions/stubborn people are initially distributed in a network with a community structure (SBM) will be discussed afterwards.
+The impact of how the opinions/stubborn actors are initially distributed in a network with a community structure (SBM) will be discussed afterwards.
 Finally, we will run the opinion dynamics model of this thesis on some real-world networks.
@@
   \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ distribution.
-Two different network models with a community structure are represented, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Two different network models with a community structure are represented, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
@@
 \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; The Watts-Strogatz model (WS) with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model with $p = 0.01$; The stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; The stochastic Watts-Strogatz model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; The Watts-Strogatz model with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
 The starting conditions are $P_{\text{A}}(0) = 0.5$ and $P_{\text{B}}(0) = 0.5$.
@@
 \begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
+  \begin{subfigure}[t]{0.49\textwidth}
+    \subcaption{ER}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
       \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
+  \begin{subfigure}[t]{0.49\textwidth}
+    \subcaption{WS}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}    
-    \caption{SBM}
+  \begin{subfigure}[t]{0.49\textwidth}    
+    \subcaption{SBM}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+  \begin{subfigure}[t]{0.49\textwidth}
+    \subcaption{SBM-WS}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
@@
   \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
@@
 \newline
-Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
+Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
 For each network, the REC and PR filtering algorithm are compared to each other.
@@
 There are no visible effects of having a clear community structure.
-This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared.
+Figure \ref{sbm} shows the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure).
 There is (almost) no difference in the results for a high and a low modularity.\\
@@
 Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
-When using the REC filtering algorithm, the formation of echo chambers around the majority opinion 1 is hampered.
-This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
-There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0.
-From figure \ref{ev_op_20_80} it was already concluded that the REC filtering algorithm doesn't break the prevalence of the opinions.
-The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters.
+When using the REC filtering algorithm, the formation of echo chambers around the majority opinion B is hampered.
+This is probably a consequence of the absence of clustering and the presence of short average path lengths in these models, since $80\%$ of the nodes carry this opinion.
+There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion A.
+From figure \ref{ev_op_20_80} it was already concluded that the REC filtering algorithm does not break the prevalence of the opinions.
+The opinion distribution is however slightly shifted from the case where the minority opinion A is more or less randomly distributed over the network to the case where the minority opinion A groups together in opinion clusters.
 \\
@@
 The PR filtering algorithm shows a completely different tendency.
-There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1.
-This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1.
-This reduced visibility of the minority opinion 0 in these neighborhoods drives them eventually to adopt opinion 1, with an increase in the number of echo chambers around the majority opinion 1 as a consequence.
+There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion B.
+This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion A in the neighborhood of nodes with the majority opinion B.
+This reduced visibility of the minority opinion A in these neighborhoods drives them eventually to adopt opinion B, with an increase in the number of echo chambers around the majority opinion B as a consequence.
 Remember from Figure \ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion.
-Hence, the increase in the formation of echo chambers around the majority opinion 1 is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
-This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the nodes carrying the minority opinion 0 becomes equal to zero.
+Hence, the increase in the formation of echo chambers around the majority opinion B is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
+This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the nodes carrying the minority opinion A becomes equal to zero.
 \\
@@
 Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
-For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
-An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
+For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
+An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
 This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity.
-In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180.
-These large spikes, indicating a large increase of echo chambers around the minority opinion, don't necessary mean that we have a huge absolute number of echo chambers around the minority opinion 0 at $t=500$.
-They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion 0 at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion 0 at $t=500$).
-If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
+In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero are not actually reaching infinity, but are cut off around values equal to 100/180.
+These large spikes, indicating a large increase of echo chambers around the minority opinion, do not necessary mean that we have a huge absolute number of echo chambers around the minority opinion A at $t=500$.
+They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion A at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion A at $t=500$).
+If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion B and that only a minor fraction carries the minority opinion A.\\
 \newline
 This enhancing of the formation of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
-Remember from Figure \ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks.
+Remember from Figure \ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for the high clustering networks.
 Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
 The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
-This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution.
+This means that the average fraction of nodes with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
 In case of the PR filtering algorithm this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.
-The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$.
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
+The starting conditions are $P_\text{A}(0) = 0.2$ and $P_\text{B}(0) = 0.8$.
 Each curve is the average of $10^2$ independent simulations.}
@@
 
-\begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_12x10}
-	\captionsetup{format=plain}
-	\caption[Relation between the formation of echo chambers and community structure (modularity).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-The initial opinion distribution is $50/50$.
-The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10.
-The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around $\left<k\right> \sim 10$.
-Each data point is the result of $10^2$ independent simulations.}
-\label{echo_vs_mod}
-\end{figure}
-
-Lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$.
 In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
-The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$).
+The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion A and the normalized fraction of nodes with all neighbors carrying opinion B (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
 This is a measure for the formation of echo chambers (opinion clusters) in the system.
@@
 \newline
-In increasing order of modularity, the used network models are (all with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
-\begin{itemize}
-	\item $Q=0 \rightarrow$ Erd\H{o}s-R\'{e}nyi model with edge probability $p=0.01$,
-	\item $Q \sim 0.2 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.03$ and the edge probability between communities is $p_{\text{add}} = 0.008$,
-	\item $Q \sim 0.47 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.05$ and the edge probability between communities is $p_{\text{add}} = 0.005$,
-	\item $Q \sim 0.6 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.07$ and the edge probability between communities is $p_{\text{add}} = 0.004$,
-	\item $Q \sim 0.9 \rightarrow$ SBM with 10 communities of size 100; the edge probability inside the communities is $p_{\text{cl}} = 0.1$ and the edge probability between communities is $p_{\text{add}} = 0.001$.
-\end{itemize}
 The data points are shown together with the standard error $SE = \sigma /\sqrt{N}$.
@@
 This suggests that the value of the calculated average lies close to the real average.
-Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.\\
-\newline
-The PR filtering algorithm is compared to the REC filtering algorithm.
+Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.
+
+\begin{figure}[H]
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_10x8}
+	\captionsetup{format=plain}
+	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+The initial opinion distribution is $50/50$.
+The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10.
+The other four data points are the results for SBM's with 10 communities of size 100 ($N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and average degree $\left<k\right> \sim 10$.
+Each data point is the average of $10^2$ independent simulations.
+The bars represent the standard error $SE$.}
+\label{echo_vs_mod}
+\end{figure}
+\noindent
+Table \ref{tab1} displays the network models that are used in Figure \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
+
+\begin{table}[H]
+\centering
+\setlength{\tabcolsep}{10pt} % Default value: 6pt
+\renewcommand{\arraystretch}{1.5} % Default value: 1
+\begin{tabular}{c | c}
+$\bm{Q}$ & \textbf{Network Model} \\
+\hline
+$0$ & Erd\H{o}s-R\'{e}nyi model; $p=0.01$ \\
+$0.220 \pm 0.007$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ \\
+$0.476 \pm 0.009$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.05$ and $p_{\text{add}} = 0.005$ \\
+$0.624 \pm 0.008$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.07$ and $p_{\text{add}} = 0.004$ \\
+$0.908 \pm 0.005$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ \\
+\end{tabular}
+\captionsetup{format=plain}
+\caption{Network models used in Figure \ref{echo_vs_mod}.}
+\label{tab1}
+\end{table}
+\noindent
+The PR and REC filtering algorithms are compared.
 For most values of the modularity, the REC filtering algorithm produces almost no echo chambers.
-If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
-\newline
-For the PR filtering algorithm, the increase in the size of the echo chambers with increasing modularity is much more significant.
-From a modularity approximately equal to $Q \sim 0.5$, we can see a sharp increase in the formation of echo chambers.
-For modularity values smaller than $Q \sim 0.5$, the formation of echo chambers remains very small, even when using the PR filtering algorithm.\\
-\newline
-These results strongly suggest that a community structure enhances the formation of echo chambers in the system, mostly when using the PR filtering algorithm.
-A quite clear community structure, represented by a high modularity, is however needed to really see the effect of this community structure on the formation of opinion clusters.
-
-\section{Impact of stubborn people}
-
-In this section, the impact of introducing stubborn people will be reviewed. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
-\begin{itemize}
-	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
-	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
-\end{itemize}
-We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
-\begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<cc\right> \sim 0.55$.
-This model is used to investigate the effect of clustering.
-	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$).
-This model is used to investigate the effect of a community structure.
-	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$.
+If the modularity becomes high enough we see a small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than two.
+This means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
+\newline
+For the PR filtering algorithm, the increase in the size of the echo chambers with increasing modularity is more significant.
+From a modularity equal to $Q = 0.476 \pm 0.009$, we see a sharp increase in the formation of echo chambers.
+For $Q < 0.476$, the formation of echo chambers remains small, even when using the PR filtering algorithm.\\
+\newline
+These results suggest that the community structure enhances the formation of echo chambers, mostly when using the PR filtering algorithm.
+A strong community structure, high modularity, is however required to see the effect of community structure on the formation of opinion clusters.
+
+\section{Stubborn actors}
+
+In this section, we will study the effect of stubborn actors on the opinion dynamics.
+We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}.
+The results of the first model, including a stubbornness parameter, are given in Section \ref{stubbpar}.
+The results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
+\newline
+We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness.
+Each of the network models have a total number of nodes $N=10^3$ and an average degree $\left<k\right> \sim 10$.
+The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
+The interplay between stubbornness and clustering is studies with the Watts-Strogatz model (WS) with mean degree $K=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
+The effect of community structure is investigated with the stochastic block model (SBM) with 10 communities of size 100.
+Two different modularities are implemented: a high modularity, $Q = 0.908 \pm 0.005$, with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity, $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Finally, the interplay between stubbornness and both clustering and community structure is studied with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
+Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> = 10$ and rewire probability $\beta = 0.01$.
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<cc\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
-\end{itemize}
+The clustering coefficient of this network is $\left<cc\right> = 0.553 \pm 0.004$ and the modularity is $Q = 0.909 \pm 0.004$.\\
+\newline
 For each network model, we compare the REF, REC and PR filtering algorithms to each other to analyze the possible interplay between stubbornness and filtering algorithm.
@@
 \label{stubbpar}
-Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
-\begin{itemize}
-	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
-	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$, to completely stubborn, $r = 1$).
-\end{itemize}
-For both cases we will look at both the evolution of the opinions and the formation of echo chambers and we will do this for two different initial conditions: a balanced initial opinion distribution ($50/50$) and an unbalanced initial opinion distribution ($20/80$).
-\\
-\newline
-Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
-\newline
-\textbf{Making a fraction of the nodes completely stubborn}\\
-\newline
-We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$).
+
+In this section, we study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of nodes completely stubborn, (ii) all nodes partly stubborn.\\
+\newline
+For both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers.
+This is done this for two different starting conditions: a balanced starting conditions and unbalanced starting conditions.
+\\
+\newline
+The opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
+\newline
+\textbf{Completely stubborn actors}\\
+\newline
+We make a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$).
 We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.
@@
 \begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_10x8.png}
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_frac_complStubb}
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm and for an initial $50/50$ opinion distribution.
+Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the PR filtering algorithm and for an initial $50/50$ opinion distribution.
 The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.
-Data points are again shown together with their standard error, $SE$, for the same reasons as explained in Section \ref{echoVSmod}.\\
+The dotted black line represents an echo chamber size equal to one.
+This means that the number of echo chambers remains constant over time.\\
 \newline
 The Erd\H{o}s-R\'{e}nyi model (ER) shows no appearance of echo chambers.
-Introducing a fraction of completely stubborn nodes doesn't change this behavior.
+Introducing a fraction of completely stubborn actors does not change this behavior.
 The stochastic block model with a low modularity behaves similar as the Erd\H{o}s-R\'{e}nyi model (no formation of echo chambers).
-Note that the data points for the SBM with a low modularity are not visible in Figure \ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
+The data points for the SBM with a low modularity are not visible in Figure \ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
 \newline
 The other three networks (WS, SBM-WS and SBM with a high modularity) all show qualitatively the same behavior.
-If an increasing fraction of completely stubborn nodes is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn people.
+If an increasing fraction of completely stubborn actors is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn actors.
 This can be explained as follows.
-Having a fraction of completely stubborn nodes, reduces the effective number of neighbors that can change their opinion.
+Having a fraction of completely stubborn actors, reduces the effective number of neighbors that can change their opinion.
 This in turn leads to the hampering of the formation of echo chambers.\\
 \newline
-The WS and SBM-WS networks are the high clustering networks and for a small fraction of stubborn people they show the largest echo chambers.
-The stochastic block model with a high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for a fraction of stubborn nodes smaller than fracRes $\leqslant 0.1$.
+The WS and SBM-WS networks are the high clustering networks.
+For a small fraction of stubborn actors they show the largest echo chambers.
+The stochastic block model with a high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for fracRes $\leqslant 0.1$.
 This leads us to make the same conclusion as before: a community structure enhances the formation of echo chambers compared to a random network, however not as significant as networks with a high clustering coefficient.
 \begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \caption{REF}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \caption{REC}
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution.
-\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+\textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc \right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_frac_complStubb_REF-REC}
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms.
-The results are again obtained for an initial $50/50$ opinion distribution.
-Data points are shown together with their standard error $SE$.
-The results for the REF and REC filtering algorithms behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
-Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease.
+Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the REF and REC filtering algorithms.
+The results are obtained for an initial $50/50$ opinion distribution.
+The results for the REF and REC filtering algorithms behave qualitatively the same, but are different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
+Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors, we first observe an increase in the formation of echo chambers, followed by a decrease.
 This is especially clear for the high clustering networks (WS and SBM-WS).
-The SBM with a high modularity also displays a tiny bump.
-The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\
-\newline
-These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn agents enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
+The SBM with a high modularity displays a tiny bump.
+The ER and the SBM with a low modularity (data points for this model lie behind the ER data points) do not seem to have a significant bump.\\
+\newline
+These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn actors enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
 % explain potential reasons (eg echo chambers around stubborn nodes --> keep on pushing their opinion to the neighbors, not the case for PR) + check those + need for futher research.
-If the fraction of completely stubborn nodes becomes bigger than fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
-This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions.
+If fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
+This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that will not change opinions.
 This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.\\
 \newline
-This discrepancy between the results for the PR filtering algorithm and those for the REF/REC filtering algorithms in combination with a small fraction of stubborn nodes can be potentially explained with the help of the following simple example.
+This discrepancy between the results for the PR filtering algorithm and those for the REF/REC filtering algorithms in combination with a small fraction of stubborn actors can be potentially explained with the help of the following simple example.
 Assume we have three nodes that form a triangle.
-Two nodes initially carry opinion 0 and one node carries opinion 1.
+Two nodes initially carry opinion A and one node carries opinion B.
 Lets first discuss the PR filtering algorithm and assume that none of the nodes are stubborn.
 In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the nodes adopt the same opinion).
-This can be easily seen: the one node with opinion 1 is likely to switch to opinion 0, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion 1 does not matter here, since there are no neighbors with opinion 1).
-The two nodes with opinion 0 will have more chance of keeping that opinion than switching to opinion 1 in case of the PR filtering algorithm.
-This is because the PR filtering algorithm will promote the posts of the friend with the same opinion 0 over the posts of the friend with opinion 1.
-The end result will be that there is a high chance that all the three nodes will adopt opinion 0 and hence, form an echo chamber.
-Now assume that we make the node with opinion 1 stubborn.
-This node is then no longer capable of switching to opinion 0.
-However, the two nodes with opinion 0 still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion 0, and hence reduces the visibility of the posts of the stubborn friend with opinion 1.
-The system will now not form an echo chamber, since the stubborn node with opinion 1 cannot switch to opinion 0 and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion 0 to opinion 1.
-The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn people.
+This can be easily seen: the one node with opinion B is likely to switch to opinion A, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
+The two nodes with opinion A will have more chance of keeping that opinion than switching to opinion B in case of the PR filtering algorithm.
+This is because the PR filtering algorithm will promote the posts of the friend with the same opinion A over the posts of the friend with opinion B.
+The end result will be that there is a high chance that all the three nodes will adopt opinion A and hence, form an echo chamber.
+Now assume that we make the node with opinion B stubborn.
+This node is no longer capable of switching to opinion A.
+However, the two nodes with opinion A still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion A, and hence reduces the visibility of the posts of the stubborn friend with opinion B.
+The system will now not form an echo chamber, since the stubborn node with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion A to opinion B.
+The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
 What happens if we use the REF/REC filtering algorithm?
-Again we have one node with opinion 1 and two nodes with opinion 0 in the triangle.
+Again we have one node with opinion B and two nodes with opinion A in the triangle.
 Assume first that none of the nodes are stubborn.
@@
 For the REF/REC filtering algorithms this chance is however smaller, which can be seen as follows.
-The node with opinion 1 still has a high chance of switching to opinion 0, because all his friends (which are the two other nodes in the triangle) carry this opinion.
-However, the two nodes with opinion 0 will now not have a higher chance of keeping this opinion than to switch to opinion 1, as was the case for the PR filtering algorithm.
+The node with opinion B still has a high chance of switching to opinion A, because all his friends (which are the two other nodes in the triangle) carry this opinion.
+However, the two nodes with opinion A will now not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
 Instead, they have an equal chance of keeping their opinion as to change to the other opinion.
-This is because both the REF and REC filtering algorithms don't discriminate between the two different opinions of the friends of the nodes with opinion 0.
-Both opinion 0 and opinion 1 have an equal chance of being promoted on the time-line of the nodes with opinion 0.
-This reduced tendency of the nodes with opinion 0 to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
-This would explain why, in case of no stubborn nodes, the REC/REF filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
-Now, lets again assume that we make the node with opinion 1 stubborn.
+This is because both the REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the nodes with opinion A.
+Both opinion A and opinion B have an equal chance of being promoted on the time-line of the nodes with opinion A.
+This reduced tendency of the nodes with opinion A to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
+This would explain why, in case of no stubborn actors, the REC/REF filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
+Now, lets again assume that we make the node with opinion B stubborn.
 In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle.
 For the REC/REF filtering algorithms, things are different.
-The stubborn node with opinion 1 will obviously not change opinion.
-The two nodes with opinion 0, on the other hand, will now have a higher chance of switching to opinion 1 (which was not the case when using the PR filtering algorithm).
+The stubborn node with opinion B will not change opinion.
+The two nodes with opinion A, on the other hand, will now have a higher chance of switching to opinion B (which was not the case when using the PR filtering algorithm).
 This can be seen as follows.
-The REF/REC filtering algorithms do not discriminate between the two different opinions of the two friends of the nodes with opinion 0.
-If opinion 1 is, however, stubborn it will keep on pushing that opinion on the time-lines of the two nodes with opinion 0.
-This will eventually lead to an increase of posts with opinion 1, compared to posts with opinion 0, on the time-lines of the two nodes with opinion 0.
-The result will then be that the nodes with opinion 0 will have a higher chance of switching to opinion 1, which enhances the formation of echo chambers in the triangle.
-This would explain the increase in the formation of echo chambers for a small fraction of stubborn people when using the REF or REC filtering algorithms.
+The REF/REC filtering algorithms do not discriminate between the two different opinions of the two friends of the nodes with opinion A.
+If the node with opinion B is, however, stubborn it will keep on pushing that opinion on the time-lines of the two nodes with opinion A.
+This will eventually lead to an increase of posts with opinion B, compared to posts with opinion A, on the time-lines of the two nodes with opinion A.
+The result will then be that the nodes with opinion A will have a higher chance of switching to opinion B, which enhances the formation of echo chambers in the triangle.
+This would explain the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
 If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease again which can also be easily seen with the help of the triangle example.
-If the fraction of stubborn people becomes too high, we might end up with the case where both the node with opinion 1 and one of the nodes with opinion 0 are stubborn.
+If the fraction of stubborn actors becomes too high, we might end up with the case where both the node with opinion B and one of the nodes with opinion A are stubborn.
 This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three nodes in the triangle carry the same opinion.\\
@@
 \newline
-So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn agents.
-But what about the evolution of the opinions over time?
-Are these affected by introducing stubborn people in the system?
+So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn actors.
+But what about the evolution of the prevalence of the opinions over time?
+Are these affected by introducing stubborn actors in the system?
 The short answer is no.
-Introducing stubborn people in the system does not affect the evolution of the opinions over time for neither the REC, nor the PR filtering algorithm.
+Introducing stubborn actors in the system does not affect the evolution of the prevalence of the opinions over time for neither the REC, nor the PR filtering algorithm.
 The balanced opinion distribution is maintained over time.
@@
 \newline
-Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution.
-Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg.
-an initial $20/80$ distribution.
+Lets now study the situation of unbalanced starting conditions, combined with a fraction of stubborn actors.
+The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and the initial prevalence of opinion B is $P_\text{B}(0) = 0.8$.
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution.
 \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_frac_complStubb_PR-REC_20-80}
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion).
+Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion).
 Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm.
@@
 The interpretation of the $y$-axis is slightly different than before.
-In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
-Thus we first normalized the echo chambers around opinion 0 and those around opinion 1 and than took the average of both.
+In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both.
 In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity.
-We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
-Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
+We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
+Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
 Note that by taking the average, we neglect the skewed echo chamber distribution.
-This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn people and are less interested in the distribution of those echo chambers.
-\\
-\newline
-When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes.
+This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers.
+\\
+\newline
+When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
 This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}.
-Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
+Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn actors equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
 This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}.
@@
 The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
-It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution.
+It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of an initial $20/80$ opinion distribution.
 The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
 When comparing the different network models, these skewed distributions should not be overlooked.
-The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to (quantitatively) compare the different models against each other.
+The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other.
 \\
@@
 The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, look again different than those for the PR filtering algorithm.
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn nodes, followed by a decrease.
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease.
 This result is very similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}.
-The stochastic block model with a high modularity doesn't really show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn nodes smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn agents increases further, we can start to see a decrease in the formation of echo chambers.
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn agents (echo chamber size around $1.2$).
+The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn actors smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we can start to see a decrease in the formation of echo chambers.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn actors (echo chamber size around $1.2$).
 This very limited, almost non-significant formation of echo chambers for the ER model and the SBM with a low modularity, in combination with the REC filtering algorithm, is also visible in Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
-For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
-\newline
-We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC filtering algorithm).
-\\
-\newline
-So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn people in the system.
+For an increasing fraction of the completely stubborn actors, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
+\newline
+We can conclude that the formation of echo chambers versus the fraction of completely stubborn actors, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn actors for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm).
+\\
+\newline
+So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn actors in the system.
 Lets now discuss the evolution of the opinions in the system.
-In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn agents the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
 The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion.
-Are these results affected by introducing stubborn agents?
-As for the balanced initial conditions, introducing stubborn people does not break the prevalence of the initial opinions for the REC filtering algorithm.
-Introducing stubborn people in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
-\newline
-Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion 0 versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
+Are these results affected by introducing stubborn actors?
+As for the balanced initial conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
+Introducing stubborn actors in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
+\newline
+Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion A versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
 Data points are presented along with their standard deviation $\sigma$.
-Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
-\\
-\newline
-If there are no stubborn people, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
-Introducing a fraction of stubborn agents changes this scenario.
-For an increasing fraction of stubborn people, the decrease of the minority opinion becomes less and less strong.
+Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
+\\
+\newline
+If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
+Introducing a fraction of stubborn actors changes this scenario.
+For an increasing fraction of stubborn actors, the decrease of the minority opinion becomes less and less strong.
 This result is the same for all the different network topologies.
-If there are no stubborn people, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
-If we however introduce a fraction of stubborn people in the system, then we will also end up with some stubborn nodes that carry the minority opinion.
-These stubborn nodes with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
-The consequence is that these stubborn people with the minority opinion hamper the decrease of that minority opinion.
+If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
+If we however introduce a fraction of stubborn actors in the system, then we will also end up with some stubborn actors that carry the minority opinion.
+These stubborn actors with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
+The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
@@
 \newline
-Instead of introducing a fraction of completely stubborn people in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so).
+Instead of introducing a fraction of completely stubborn actors in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so).
 All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
@@
 \begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
 	\captionsetup{format=plain}
 	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_all_frac_stubb_PR}
@@
 If $r=1$, the nodes are completely stubborn and thus keep their initial opinion.
-Hence, the number of like minded neighborhoods doesn't change over time, which gives a normalized number of echo chambers equal to one.
+Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
 It is however remarkable that for very high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is very large and even larger than if the nodes are not stubborn.
@@
 This is the complete opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
-The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn nodes, see the ``Making a fraction of the nodes completely stubborn'' paragraph.
+The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Making a fraction of the nodes completely stubborn'' paragraph.
 If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.
@@
 \newline
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) don't seem to have any dependency on the stubbornness level of the nodes.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) do not seem to have any dependency on the stubbornness level of the nodes.
 This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithm that drives this increase in the formation of echo chambers with increasing $r$.
@@
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_all_frac_stubb_REF-REC}
@@
 $20/80$.
-Opinion 0 is again the minority opinion, which makes opinion 1, obviously, the majority opinion.
+Opinion A is again the minority opinion, which makes opinion B, obviously, the majority opinion.
 First, we will discuss the formation of echo chambers in the system under unbalanced initial opinion conditions and when making all the nodes partly stubborn.
@@
 \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_all_frac_stubb_PR-REC_20-80}
@@
 \newline
-The results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm are not affected by making all the nodes more and more stubborn.
+The results for the fraction of nodes with opinion A versus time in case of the REC filtering algorithm are not affected by making all the nodes more and more stubborn.
 The prevalence of the opinions is maintained an the status quo is not broken.
 The PR filtering algorithm gives us another story.
-The results for the evolution of the fraction of nodes with opinion 0 over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
+The results for the evolution of the fraction of nodes with opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
 If the nodes are not stubborn, the PR filtering algorithm strongly reduces the minority opinion over time.
 Making the nodes more and more stubborn reduces this tendency.
-This result is more or less similar to the corresponding case in which a fraction of completely stubborn nodes was introduced in the system, see Figure \ref{ev_op_20_80_stubb_PR}.
+This result is more or less similar to the corresponding case in which a fraction of completely stubborn actors was introduced in the system, see Figure \ref{ev_op_20_80_stubb_PR}.
 The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the minority opinion compared to making a fraction of the nodes completely stubborn.
@@
 All the nodes need to be really stubborn ($r=0.8$) to really see a hampering in the decrease of the minority opinion.
-Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn people equal to fracRes $=0.4$ was introduced (a final fraction  of the minority opinion equal to 0.07 versus a final fraction of the minority opinion equal to 0.1).
+Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced (a final fraction  of the minority opinion equal to 0.07 versus a final fraction of the minority opinion equal to 0.1).
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
@@
 This model is explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
-Note, that we don't make use of a probabilistic majority model here, but instead a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
+Note, that we do not make use of a probabilistic majority model here, but instead a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
 The rest of the model (filtering algorithms, activation mechanism,...) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
@@
 \begin{figure}[H]
-	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_threshold_PR_12x10.png}
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_threshold_PR_12x10.png}
 	\captionsetup{format=plain}
@@
 The majority threshold model is implemented.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_threshold_PR}
@@
 However, each run the system converges to one of the two opinions, with the result that the number of echo chambers around this opinion strongly increases.
-After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion 1 (and thus resulting in large echo chambers around opinion 1) and there will have been simulations that converge to opinion 0 (resulting in almost no echo chambers around opinion 1).\\
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).\\
 \newline
@@
 If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an impact on the system.
-However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold doesn't become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
+However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
 At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
-The number of echo chambers doesn't change over time, which results in an echo chamber size equal to one in the figure.\\
+The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.\\
 \newline
 Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
-The figure shows the evolution of opinion 0 over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion 0; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion 0 as to opinion 1, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the evolution of opinion A over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
 The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
@@
 \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the result of $10^2$ independent simulations.}
+Each data point is the average of $10^2$ independent simulations.}
 \label{echo_vs_threshold_REF-REC}
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.
-Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution.
+  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ opinion distribution.
+Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ distribution.
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
@@
 
-\section{Random versus community distributed opinions/stubborn agents}
-
-This section is devoted to results regarding the way the opinions and/or the stubborn people are distributed over the network.
-In all the previous results, the two opinions and the stubborn people (in case there were stubborn people introduced in the system) were, initially, always randomly distributed across the system.
+\section{Random versus community distributed opinions/stubborn actors}
+
+This section is devoted to results regarding the way the opinions and/or the stubborn actors are distributed over the network.
+In all the previous results, the two opinions and the stubborn actors (in case there were stubborn actors introduced in the system) were, initially, always randomly distributed across the system.
 It might however be more realistic if people with the same opinion are more or less grouped together initially.
 We did introduce a network model with a community structure (the SBM) to represent the group structures found in real-life systems.
-But we haven't yet distributed the opinions according to those groups.
+But we have not yet distributed the opinions according to those groups.
 In subsection \ref{groupVSrandom} we will discuss some results for distributing the opinions in communities versus randomly distributing them.
@@
 
-In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion 0.
+In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion A.
 The remaining communities are given a $50/50$ opinion distribution.
 This situation may, in an idealized way, represent the situation in which you have groups of people following the same idea together with groups where the two ideas coexist.
-The total opinion distribution over the entire network is no longer $50/50$, but instead opinion 0 becomes the majority opinion (hence, opinion 1 becomes the minority opinion).
-The higher the fraction of single opinion 0 communities, the bigger the difference between this majority and minority opinion becomes.
+The total opinion distribution over the entire network is no longer $50/50$, but instead opinion A becomes the majority opinion (hence, opinion B becomes the minority opinion).
+The higher the fraction of single opinion A communities, the bigger the difference between this majority and minority opinion becomes.
 The results will always be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
 Thus, eg.
-if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion 0 communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$.
+if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion A communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$.
 This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
@@
 \newline
-Figure \ref{commOp0_01_other_50-50} shows the results for the fraction of nodes with opinion 0 versus time for the case in which a fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the other communities get an initial $50/50$ opinion distribution.
-The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in case of a fraction of $0.1$ single opinion 0 communities).
+Figure \ref{commOp0_01_other_50-50} shows the results for the fraction of nodes with opinion A versus time for the case in which a fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the other communities get an initial $50/50$ opinion distribution.
+The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in case of a fraction of $0.1$ single opinion A communities).
 The figure shows the results for both the high and the low modularity case.
@@
 In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
-The prevalence of the majority opinion 0 remains unchanged over time.
+The prevalence of the majority opinion A remains unchanged over time.
 This result is the same for both the high and low modularity case, see Figures \ref{high_mod_rec} and \ref{low_mod_rec}.\\
@@
 Figure \ref{ev_op_20_80}.
-If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion 0, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
+If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion A, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
 This makes sense, since the single opinion communities are more or less isolated from the other communities.
 This has as consequence that the other (mixed opinion) communities live under the assumption that the opinion distribution is $50/50$.
-They don't really feel that opinion 0 is a majority opinion, since they are isolated from these single opinion 0 communities.
+They do not really feel that opinion A is a majority opinion, since they are isolated from these single opinion A communities.
 In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
@@
 The different communities talk a lot to each other and there is a lot of communication between them.
-As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion 0 in the same way as if the opinions would be randomly distributed over all the communities.
+As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over all the communities.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion 0 communities, while the others get an initial $50/50$ opinion distribution.
+  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
 Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.
@@
 \textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
-Each curve is the result of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
+Each curve is the average of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
 \label{commOp0_01_other_50-50}
@@
 \noindent
-The results found in Figure \ref{commOp0_01_other_50-50} for a fraction of $0.1$ of single opinion 0 communities remain valid if we increase the fraction of single opinion 0 communities.
-This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average fraction of opinion 0 before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion 0 communities.
+The results found in Figure \ref{commOp0_01_other_50-50} for a fraction of $0.1$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
+This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average fraction of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
 The error bars in this figure are omitted for the sake of an improved visualization.
@@
 \newline
-The higher the fraction of single opinion 0 communities, the higher the initial average fraction of opinion 0.
-This can be seen in both figures by the increase of the solid black line, which represents the initial fraction of nodes with opinion 0.
-In case of the REC filtering algorithm, Figure \ref{rec_comm}, the fraction of opinion 0 remains the same after the time evolution for a fixed fraction of single opinion 0 communities.
-Giving a fraction of the communities a single opinion, opinion 0, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
+The higher the fraction of single opinion A communities, the higher the initial average fraction of opinion A.
+This can be seen in both figures by the increase of the solid black line, which represents the initial fraction of nodes with opinion A.
+In case of the REC filtering algorithm, Figure \ref{rec_comm}, the fraction of opinion A remains the same after the time evolution for a fixed fraction of single opinion A communities.
+Giving a fraction of the communities a single opinion, opinion A, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
 This result is valid for both the high and low modularity case.\\
 \newline
-The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the fraction of opinion 0, which is the majority opinion, over time for a fixed fraction of single opinion 0 communities.
-Making a fraction of the communities single opinion 0 communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
+The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the fraction of opinion A, which is the majority opinion, over time for a fixed fraction of single opinion A communities.
+Making a fraction of the communities single opinion A communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
 The low modularity network does not show a sign of this hampering.
-The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was already explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion 0 communities can influence the $50/50$ communities.
+The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was already explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
 
@@
   \captionsetup{format=plain}
-  \caption[Average fraction of opinion 0 before and after the time evolution versus the fraction of single opinion 0 communities.
+  \caption[Average fraction of opinion A before and after the time evolution versus the fraction of single opinion A communities.
 Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Average fraction of opinion 0 before and after the time evolution versus the fraction of single opinion 0 communities.
+Results for the REC and PR filtering algorithm.]{Average fraction of opinion A before and after the time evolution versus the fraction of single opinion A communities.
 Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each curve is the result of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.}
 \label{fracOp0_vs_commOp0_other_50-50}
@@
 \newline
-Figure \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion 0 community while the other communities get an initial $50/50$ opinion distribution.
+Figure \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
 As in similar plots before, the value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
@@
 Lets first discuss the results for the REC filtering algorithm.
-The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0 (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line).
-The number of neighborhoods where all the nodes carry opinion 0 thus decreases in favor of neighborhoods where we find both nodes with opinion 0 and nodes with opinion 1.
-Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion 0 communities, reduces and that the opinions become more mixed.
-The fraction of opinion 0 remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion 0 see a decrease in the nodes with opinion 0 and become more mixed.
-We can also observe a very small increase in the number of echo chambers around the minority opinion 1, which indicates the fact that this minority opinion starts to group more together.
-The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion 0 and opinion 1.
+The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line).
+The number of neighborhoods where all the nodes carry opinion A thus decreases in favor of neighborhoods where we find both nodes with opinion A and nodes with opinion B.
+Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion A communities, reduces and that the opinions become more mixed.
+The fraction of opinion A remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion A see a decrease in the nodes with opinion A and become more mixed.
+We can also observe a very small increase in the number of echo chambers around the minority opinion B, which indicates the fact that this minority opinion starts to group more together.
+The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion A and opinion B.
 Again, the opinion fractions remain constant over time, but are shifted from a random distribution to a more clustered situation.
@@
 \newline
-The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion 0, which could be observed for the community distributed, high modularity case.
-This is due to the fact that the number of echo chambers around opinion 0 at $t=0$, originating from the fraction of single opinion 0 communities, is a lot smaller for the low modularity network than for the high modularity network.
+The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion A, which could be observed for the community distributed, high modularity case.
+This is due to the fact that the number of echo chambers around opinion A at $t=0$, originating from the fraction of single opinion A communities, is a lot smaller for the low modularity network than for the high modularity network.
 The reason for this is the much more blended community structure for the low modularity network.
-We can still observe a very small increase in the number of echo chambers around opinion 1.
-This increase in the formation of echo chambers around the minority opinion 1 seems however hampered compared to the randomly distributed case.
+We can still observe a very small increase in the number of echo chambers around opinion B.
+This increase in the formation of echo chambers around the minority opinion B seems however hampered compared to the randomly distributed case.
 Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80}, see Figure \ref{sbm}.
@@
 The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (note that these results were obtained for an initial $20/80$ opinion distribution).
-We can observe a strong increase in the formation of echo chambers around the majority opinion 0, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion 1.
-Remember from Figure \ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion 0.
-If we, however, make a fraction equal to $0.1$ of the communities single opinion 0 communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion 0 is strongly hampered.
+We can observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion B.
+Remember from Figure \ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion A.
+If we, however, make a fraction equal to $0.1$ of the communities single opinion A communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion A is strongly hampered.
 This hampering is partly an artifact of the way the opinions are distributed.
-If we give entire communities only opinion 0 while others get a $50/50$ distribution, we have a large number of echo chambers around opinion 0 at $t=0$ due to these single opinion communities.
-Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
-If we, however, would look at the absolute number of echo chambers around opinion 0 at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
-This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0.
-Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion 0 (and thus the decrease of the minority opinion 1); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed cases versus the random distributed case.
-Beside this hampering of echo chambers around the majority opinion 0, the community distributed case also shows an increase of echo chambers around the minority opinion 1 (which is not visible for the randomly distributed case).
+If we give entire communities only opinion A while others get a $50/50$ distribution, we have a large number of echo chambers around opinion A at $t=0$ due to these single opinion communities.
+Moreover, the number of opinion A echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
+If we, however, would look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
+This shows that the observed hampering of the formation of echo chambers around the majority opinion A in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion A.
+Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed cases versus the random distributed case.
+Beside this hampering of echo chambers around the majority opinion A, the community distributed case also shows an increase of echo chambers around the minority opinion B (which is not visible for the randomly distributed case).
 This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
@@
 The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network.
-Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion 0 and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion 1.
+Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion B.
 From Figure \ref{low_mod_pr}, it was already made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of this minority opinion (when using the PR filtering algorithm).
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
-A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each curve is the result of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.}
 \label{echo_01_commOp0_other_50-50}
@@
 \noindent
-Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, but now a fraction equal to $0.5$ of the communities is only given opinion 0, while the others get a $50/50$ distribution.
-The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion 0 (Figure \ref{echo_01_commOp0_other_50-50}).\\
-\newline
-The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion 0 with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
-A difference with the case where a fraction equal to $0.1$ of the communities are single opinion 0 communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion 1, but instead we observe a decrease in the number of echo chambers around the minority opinion.
-For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion 0 and a minority of the nodes carrying the minority opinion 1.
+Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, but now a fraction equal to $0.5$ of the communities is only given opinion A, while the others get a $50/50$ distribution.
+The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion A (Figure \ref{echo_01_commOp0_other_50-50}).\\
+\newline
+The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
+A difference with the case where a fraction equal to $0.1$ of the communities are single opinion A communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion B, but instead we observe a decrease in the number of echo chambers around the minority opinion.
+For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion A and a minority of the nodes carrying the minority opinion B.
 \\
@@
 If we, however, look closely, we can see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller).
-We can observe a small decrease in the formation of echo chambers around the majority opinion 0, with a small increase in more mixed opinion neighborhoods as result.
-We can also observe a small decrease in the number of echo chambers around the minority opinion 0.
-The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion 0 communities, while the others get a $50/50$ opinion distribution.\\
+We can observe a small decrease in the formation of echo chambers around the majority opinion A, with a small increase in more mixed opinion neighborhoods as result.
+We can also observe a small decrease in the number of echo chambers around the minority opinion A.
+The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion A communities, while the others get a $50/50$ opinion distribution.\\
 \newline
@@
 \newline
-Results for the PR filtering algorithm also strongly resemble those depicted in Figure \ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion 0 communities.
+Results for the PR filtering algorithm also strongly resemble those depicted in Figure \ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion A communities.
 The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Figure \ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}).
-They show a strong increase in the formation of echo chambers around the majority opinion 0.\\
-\newline
-The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion 0, which was also observed in Figure \ref{pr_comm_echo}.\\
-\newline
-The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion 0.
-This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion 0, hampers the increase in this majority opinion compared to the randomly distributed case (see Figure \ref{high_mod_pr} and Figure \ref{pr_comm}).
-The increase in the number of echo chambers around the minority opinion 0 that was visible in Figure \ref{pr_comm_echo} is no longer present.
+They show a strong increase in the formation of echo chambers around the majority opinion A.\\
+\newline
+The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion A, which was also observed in Figure \ref{pr_comm_echo}.\\
+\newline
+The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion A.
+This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (see Figure \ref{high_mod_pr} and Figure \ref{pr_comm}).
+The increase in the number of echo chambers around the minority opinion A that was visible in Figure \ref{pr_comm_echo} is no longer present.
 
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
-A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$.
+Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution.
+A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
 \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
-Each curve is the result of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.}
 \label{echo_05_commOp0_other_50-50}
@@
 It was also found that how the opinions are initially distributed over the communities in a network with a community structure (SBM) does makes a difference.
-We investigate a scenario in which a fraction of the communities was given a single opinion, opinion 0, while the remaining communities were given an initial $50/50$ distribution.
+We investigate a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ distribution.
 This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.

### 2021-04-30 draft.tex
latest update

@@
   \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution.
-\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_frac_complStubb_PR-REC_20-80}
@@
 Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion).
-Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm.
-The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure.
+The PR and the REC filtering algorithms are compared.
+The standard errors $SE$ are small and not visible in the figure.
+The dotted black line represent the situation in which the number of echo chambers remains constant over time.
 The interpretation of the $y$-axis is slightly different than before.
@@
 We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
-Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
-Note that by taking the average, we neglect the skewed echo chamber distribution.
-This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers.
-\\
-\newline
-When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
+Hence, we opted for a different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B), then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
+By taking the average, we neglect the skewed echo chamber distribution.
+This is, however, not a problem.
+We want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers.
+\\
+\newline
+When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
 This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}.
-Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn actors equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
-This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}.
+For an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
+This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was explained for Figure \ref{echo_20_80}.
 \\
@@
 The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
-It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of an initial $20/80$ opinion distribution.
-The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
+It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of unbalanced starting conditions.
+The reason is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
 When comparing the different network models, these skewed distributions should not be overlooked.
@@
 % explain reason behavior REC!!
-The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, look again different than those for the PR filtering algorithm.
+The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, are different than those for the PR filtering algorithm.
 The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease.
-This result is very similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}.
-The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn actors smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we can start to see a decrease in the formation of echo chambers.
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn actors (echo chamber size around $1.2$).
-This very limited, almost non-significant formation of echo chambers for the ER model and the SBM with a low modularity, in combination with the REC filtering algorithm, is also visible in Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
-For an increasing fraction of the completely stubborn actors, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
+This result is similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}.
+The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we see a decrease in the formation of echo chambers.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show no significant formation of echo chambers.
+For an increasing fraction of the completely stubborn actors, the echo chambers eventually converge to one (for all the network models), indicating a constant amount of echo chambers over time.\\
 \newline
@@
 \newline
-So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn actors in the system.
-Lets now discuss the evolution of the opinions in the system.
-In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
-The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion.
+Lets now discuss the evolution of the prevalence of the opinions.
+In Figure \ref{ev_op_20_80}, we saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+The PR filtering algorithm, on the other hand, resulted in a decrease of the prevalence of the minority opinion.
 Are these results affected by introducing stubborn actors?
 As for the balanced initial conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
-Introducing stubborn actors in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
-\newline
-Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion A versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm.
-Data points are presented along with their standard deviation $\sigma$.
-Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
-\\
-\newline
-If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution.
+The status quo is maintained over time.
+Introducing stubborn actors combined with the PR filtering algorithm, on the other hand, does have an impact in case of unbalanced initial conditions.\\
+\newline
+Figure \ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time in case of unbalanced starting conditions and using the PR filtering algorithm.
+We compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS).
+For each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
+\\
+\newline
+If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of unbalanced starting conditions.
 Introducing a fraction of stubborn actors changes this scenario.
 For an increasing fraction of stubborn actors, the decrease of the minority opinion becomes less and less strong.
-This result is the same for all the different network topologies.
+This result is valid for all the different network topologies.
 If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
-If we however introduce a fraction of stubborn actors in the system, then we will also end up with some stubborn actors that carry the minority opinion.
-These stubborn actors with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
+If we introduce a fraction of stubborn actors in the system, then we will end up with stubborn actors that carry the minority opinion.
+These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
 The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
-Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}%more explanation?
 \label{ev_op_20_80_stubb_PR}
@@
 \noindent
-\textbf{Making all nodes partly stubborn}\\
-\newline
-Instead of introducing a fraction of completely stubborn actors in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so).
+\textbf{All actors partly stubborn}\\
+\newline
+In the second configuration, all the nodes are partly stubborn.
+They still have some chance of changing their opinion, they are however less eager to do so.
 All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
 \newline
-We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where-after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).
+We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate the results for unbalanced starting conditions.
 
 \begin{figure}[H]
-	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_10x8.png}
 	\captionsetup{format=plain}
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_all_frac_stubb_PR}
@@
 Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod} and the data points are shown together with their standard error $SE$.
-The results look very similar to the case where a fraction of the nodes is completely stubborn and when using the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_frac_complStubb}.\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers, as before, and this is not changed by making the nodes more and more stubborn.
-The same result is obtained for the stochastic block model with a low modularity (the data points lie again behind those of the ER network).\\
+The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod}.
+The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig.
+\ref{echo_vs_frac_complStubb}).\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers.
+This is not changed by making the nodes more and more stubborn.
+The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\
 \newline
@@
 If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
-In Figure \ref{echo_vs_all_frac_stubb_PR} this is represented by an echo chamber size equal to one for $r=1$.\\
+This situation is represented by the black dotted line.\\
 \newline
 In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
-Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
-Data points are again displayed with their standard error $SE$.\\
-\newline
-We can make similar observations as we did in the ``Making a fraction of the nodes completely stubborn'' paragraph.
-The results for the REF and REC filtering algorithms look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
+Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.\\
+\newline
+We can make similar observations as we did in the ``Completely stubborn actors'' paragraph.
+The results for the REF and REC filtering algorithms are qualitatively similar to each other and differ from the results obtained with the PR filtering algorithm (Fig.
+\ref{echo_vs_all_frac_stubb_PR}).
 The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases.
-For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slightly decrease between $r\sim 0.7$ and $r\sim 0.9$.
-Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers (echo chamber values around 17 for $r\sim 0.9$ to echo chamber values equal to one for $r=1$).
-The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model, the SBM-WS model on the other hand displays a continuous increase until $r \sim 0.9$.
-In both case we have again a sharp drop from $r \sim 0.9$ to $r=1$.
+For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slowly decrease between $r\sim 0.7$ and $r\sim 0.9$.
+Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers.
+The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model.
+The SBM-WS model, on the other hand, displays a continuous increase until $r \sim 0.9$.
+In both case we have a sharp drop from $r \sim 0.9$ to $r=1$.
 If $r=1$, the nodes are completely stubborn and thus keep their initial opinion.
 Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
-It is however remarkable that for very high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is very large and even larger than if the nodes are not stubborn.
-We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithm in the case of networks with a high clustering coefficient.
-This is the complete opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
-The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Making a fraction of the nodes completely stubborn'' paragraph.
-If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers.
+This is represented by the black dotted line.
+It is remarkable that for high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is large and even larger than if the nodes are not stubborn.
+We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithms in case of networks with a high clustering coefficient.
+This is the opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
+The reason for this discrepancy is probably of a similar nature as for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Completely stubborn actors'' paragraph.
+If we zoom in on the stochastic block model with a high modularity, we see a slight increase in the formation of echo chambers.
 The increase is however not as strong as for the high clustering networks.\\
 \newline
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) do not seem to have any dependency on the stubbornness level of the nodes.
-This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithm that drives this increase in the formation of echo chambers with increasing $r$.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points are hidden behind those of the ER network) do not have any dependency on the stubbornness level of the nodes.
+This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithms that drives this increase in the formation of echo chambers with increasing $r$.
 % talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
@@
   \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
-\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+\textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_all_frac_stubb_REF-REC}
@@
 \noindent
-What about the results for the evolution of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
-Well, making all the nodes partly stubborn does not disturb/change the evolution of the opinions.
+What about the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
+Making all the nodes partly stubborn does not disturb/change the evolution of the opinions.
 The balanced initial conditions remain balanced over time, as was the case for no stubbornness in the system.
@@
 \newline
-Lets now look at the results for an unbalanced initial opinion distribution, eg.
-$20/80$.
-Opinion A is again the minority opinion, which makes opinion B, obviously, the majority opinion.
-First, we will discuss the formation of echo chambers in the system under unbalanced initial opinion conditions and when making all the nodes partly stubborn.
-Then, we will look into the results for the evolution of the opinions with time.\\
+Lets now study the results for unbalanced starting conditions ($20/80$).
+Opinion A is the minority opinion, which makes opinion B the majority opinion.
+First, we discuss the formation of echo chambers.
+Then, we investigate the results for the evolution of the prevalence of the opinions.\\
 \newline
 Figure \ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
-Data points are shown together with their standard error $SE$, which is very small and not really visible in the figure.
+The standard error $SE$ is small and not visible in the figure.
 Figure \ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Figure \ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
@@
   \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution.
-\textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
 \label{echo_vs_all_frac_stubb_PR-REC_20-80}
@@
 \noindent
-In case of the PR filtering algorithm (Figure \ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models for a stubbornness parameter less than $r \leqslant 0.4$.
-If the stubbornness parameter becomes bigger than $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
-This behavior is slightly different than in the case of an initial $50/50$ distribution, combined with the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_all_frac_stubb_PR}.
-In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if the stubbornness parameter became bigger than $r>0.1$).
+In case of the PR filtering algorithm (Fig.
+\ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models if $r \leqslant 0.4$.
+If $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
+This behavior is slightly different than in case of an initial $50/50$ distribution, combined with the PR filtering algorithm (Fig.
+\ref{echo_vs_all_frac_stubb_PR}).
+In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if $r>0.1$).
 This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
 \newline
-As was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
-It should be noted that these types of figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+As was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
+These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
 In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
-Figure \ref{echo_20_80}), as was already explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
-\newline
-The results for the REC filtering algorithm (Figure \ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution, combined with the REC filtering algorithm (see Figure \ref{REC_all_frac_stubb}).
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to a stubbornness parameter $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
-The increase seems, however, less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
+Figure \ref{echo_20_80}), as was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+\newline
+The results for the REC filtering algorithm (Fig.
+\ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution (Fig.
+\ref{REC_all_frac_stubb}).
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
+The increase seems less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
 This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
 \newline
-The stochastic block model with a high modularity also shows a slight increase in the formation of echo chambers up to $r = 0.8$.
+The stochastic block model with a high modularity shows a slight increase in the formation of echo chambers up to $r = 0.8$.
 This increase can even be observed for the low modularity case, albeit less strong.
-The Erd\H{o}s-R\'{e}nyi network seems to wiggle a bit around a formation of echo chambers equal to one.\\
-\newline
-We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn.
-Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$).
+The Erd\H{o}s-R\'{e}nyi network wiggles a bit around a formation of echo chambers equal to one.\\
+\newline
+We will now discuss the results for the evolution of the prevalence of the opinions in case of unbalanced starting conditions and making all the nodes partly stubborn.
+We investigate four different network models.
+For each model the results for three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$) are compared.
 Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
 \newline
-The results for the fraction of nodes with opinion A versus time in case of the REC filtering algorithm are not affected by making all the nodes more and more stubborn.
+The prevalence of opinion A versus time in case of the REC filtering algorithm is not affected by making all the nodes more and more stubborn.
 The prevalence of the opinions is maintained an the status quo is not broken.
 The PR filtering algorithm gives us another story.
-The results for the evolution of the fraction of nodes with opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
-If the nodes are not stubborn, the PR filtering algorithm strongly reduces the minority opinion over time.
+The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
+If the nodes are not stubborn, the PR filtering algorithm reduces the minority opinion over time.
 Making the nodes more and more stubborn reduces this tendency.
-This result is more or less similar to the corresponding case in which a fraction of completely stubborn actors was introduced in the system, see Figure \ref{ev_op_20_80_stubb_PR}.
+This result is more or less similar to the corresponding case in which a fraction of completely stubborn actors was introduced in the system (Fig.
+\ref{ev_op_20_80_stubb_PR}).
 The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the minority opinion compared to making a fraction of the nodes completely stubborn.
-For a stubbornness parameter equal to $r=0.4$, the decrease of the minority opinion is almost not hampered, it is merely slowed down.
-All the nodes need to be really stubborn ($r=0.8$) to really see a hampering in the decrease of the minority opinion.
-Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced (a final fraction  of the minority opinion equal to 0.07 versus a final fraction of the minority opinion equal to 0.1).
+For $r=0.4$, the decrease of the minority opinion is almost not hampered, it is merely slowed down.
+All the nodes need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the minority opinion.
+Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$).
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
-Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}%more explanation?
 \label{ev_op_20_80_all_frac_stubb_PR}
@@
 This model is explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
-Note, that we do not make use of a probabilistic majority model here, but instead a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
+We do not make use of a probabilistic majority model, but a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
 The rest of the model (filtering algorithms, activation mechanism,...) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
@@
 The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.\\
-For this model we will only investigate the case of balanced initial conditions, so an initial $50/50$ opinion distribution.
-We will mainly focus on the formation of echo chambers versus the threshold parameter $T$ and we will compare three different filtering algorithms: the PR, REF and REC filtering algorithm.
+\newline
+For this model we only investigate the case of balanced starting conditions.
+We focus on the formation of echo chambers versus the threshold parameter $T$ and we compare three different filtering algorithms: the PR, REF and REC filtering algorithms.
 \\
@@
 \newline
-The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat for a threshold parameter smaller than $T \leqslant 0.5$.
+The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat for $T \leqslant 0.5$.
 This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
-This is because the majority opinion of the neighbors of a node will always be larger than the threshold if the threshold is smaller than $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
-If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an effect.
+This is because the majority opinion of the neighbors of a node will always be larger than the threshold if $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
+If $T \geqslant 0.5$, it starts to have an effect.
 The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
@@
 The only difference is of a quantitative nature (the formation of echo chambers is not as large as for the high clustering networks).\\
+\newline
 The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points lie behind those of the Erd\H{o}s-R\'{e}nyi network) do not enhance the formation of echo chambers in any way.
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard deviation $\sigma$.}
 \label{echo_vs_threshold_PR}
@@
 Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
-The initial opinion distribution was again $50/50$.
-The results for the REF and REC filtering algorithms strongly resemble each other.
+The initial opinion distribution was $50/50$.
+The results for the REF and REC filtering algorithms resemble each other.
 Data points are again shown together with their standard deviation $\sigma$ to really emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 100$).
@@
 To which of the two opinions the model converges is a matter of random fluctuations.
-So, on average, the opinion distribution will remain $50/50$.
+On average, the opinion distribution will remain $50/50$.
 However, each run the system converges to one of the two opinions, with the result that the number of echo chambers around this opinion strongly increases.
-After the $10^2$ independent simulations, the spread in the formation of echo chambers will be very large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).\\
-\newline
-Another consequence of this converging to one of the two opinions every single simulation, that can be observed in Figure \ref{echo_vs_threshold_REF-REC}, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).\\
+\newline
+Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
 This is not due to any network features, but a direct consequence of the usage of the majority model.
@@
 The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
-This can be clearly seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF filtering algorithms, again an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
-If the threshold becomes bigger than $T \geqslant 0.5$, it starts to have an impact on the system.
-However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.6$, after which we start to see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
+This can be seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF filtering algorithms, an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
+If $T \geqslant 0.5$, it starts to have an impact on the system.
+However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.6$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
 At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
-The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.\\
+The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.
+A constant amount of echo chambers over time is represented by the black dotted line.\\
 \newline
 Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
-The figure shows the evolution of opinion A over time for a single simulation (note that we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
 The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
-For each model the REC and PR filtering algorithm are compared to each other.
-The threshold parameter is set equal to $T=0$, so there is no effect of any type of stubbornness.\\
-\newline
-From this figure it becomes clear that the PR filtering algorithm strongly hampers the convergence to a single opinion each independent simulation.
+For each model the REC and PR filtering algorithms are compared.
+The threshold parameter is set equal to $T=0$, so there is no effect of stubbornness.\\
+\newline
+It becomes clear that the PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
 It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
@@
 For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
-For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not really give an additional effect on top of this PR filtering effect.
+For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
 
@@
 The majority threshold model is implemented.
-\textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm.
+\textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
 The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.}
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard deviation $\sigma$.}
 \label{echo_vs_threshold_REF-REC}
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ opinion distribution.
-Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ distribution.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ opinion distribution.
+Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ distribution.
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
-Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm.}
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.}
 \label{ev_op_majT_one_sim}
@@
 
-\section{Random versus community distributed opinions/stubborn actors}
+\section{Random versus community distributed}
 

### 2021-05-01 draft.tex
latest update

@@
 Then, we introduce some kind of stubbornness in the system.
-The impact of how the opinions/stubborn actors are initially distributed in a network with a community structure (SBM) will be discussed afterwards.
+The impact of how the opinions are initially distributed in a network with community structure (SBM) will be discussed afterwards.
 Finally, we will run the opinion dynamics model of this thesis on some real-world networks.
@@
 Two different network models with a community structure are represented, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
+For each model the REC and the PR filtering algorithms are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown, together with their standard deviation $\sigma$, for every 50 time steps to improve visualization.}
@@
 \newline
-We look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
+We study four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
@@
 Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
+For each model the REC and the PR filtering algorithms are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.
-Error bars are again omitted for a better visualization.\\
+Error bars are omitted for a better visualization.\\
+\newline
 The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
@@
 Figure \ref{sbm} shows the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure).
-There is (almost) no difference in the results for a high and a low modularity.\\
-\newline
+There is (almost) no difference in the results for a high and a low modularity.
+\newpage
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
+    \label{er_echo_20-80}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
+    \label{ws}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
+  	\label{sbm}    
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
+    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
+    \label{sbm-ws}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+For each model the REC and the PR filtering algorithms are compared.
+The starting conditions are $P_\text{A}(0) = 0.2$ and $P_\text{B}(0) = 0.8$.
+Each curve is the average of $10^2$ independent simulations.}
+\label{echo_20_80}
+\end{figure}
+\noindent
 Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
@@
 
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
-    \label{er_echo_20-80}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
-    \label{ws}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
-    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
-  	\label{sbm}    
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
-    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
-    \label{sbm-ws}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.
-The starting conditions are $P_\text{A}(0) = 0.2$ and $P_\text{B}(0) = 0.8$.
-Each curve is the average of $10^2$ independent simulations.}
-\label{echo_20_80}
-\end{figure}
-
 % talk about small spike around 1 for PR method??
@@
 This suggests that the value of the calculated average lies close to the real average.
-Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.
+Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.\\
+\newline
+Table \ref{tab1} displays the network models that are used in Figure \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
+
+\begin{table}[H]
+\centering
+\setlength{\tabcolsep}{10pt} % Default value: 6pt
+\renewcommand{\arraystretch}{1.5} % Default value: 1
+\begin{tabular}{c | c}
+$\bm{Q}$ & \textbf{Network Model} \\
+\hline
+$0$ & Erd\H{o}s-R\'{e}nyi model; $p=0.01$ \\
+$0.220 \pm 0.007$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ \\
+$0.476 \pm 0.009$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.05$ and $p_{\text{add}} = 0.005$ \\
+$0.624 \pm 0.008$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.07$ and $p_{\text{add}} = 0.004$ \\
+$0.908 \pm 0.005$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ \\
+\end{tabular}
+\caption{Network models used in Figure \ref{echo_vs_mod}.}
+\label{tab1}
+\end{table}
 
@@
 \noindent
-Table \ref{tab1} displays the network models that are used in Figure \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
-
-\begin{table}[H]
-\centering
-\setlength{\tabcolsep}{10pt} % Default value: 6pt
-\renewcommand{\arraystretch}{1.5} % Default value: 1
-\begin{tabular}{c | c}
-$\bm{Q}$ & \textbf{Network Model} \\
-\hline
-$0$ & Erd\H{o}s-R\'{e}nyi model; $p=0.01$ \\
-$0.220 \pm 0.007$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ \\
-$0.476 \pm 0.009$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.05$ and $p_{\text{add}} = 0.005$ \\
-$0.624 \pm 0.008$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.07$ and $p_{\text{add}} = 0.004$ \\
-$0.908 \pm 0.005$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ \\
-\end{tabular}
-\captionsetup{format=plain}
-\caption{Network models used in Figure \ref{echo_vs_mod}.}
-\label{tab1}
-\end{table}
-\noindent
 The PR and REC filtering algorithms are compared.
@@
 A strong community structure, high modularity, is however required to see the effect of community structure on the formation of opinion clusters.
-
+\newpage
 \section{Stubborn actors}
@@
 We can conclude that the formation of echo chambers versus the fraction of completely stubborn actors, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn actors for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm).
-\\
-\newline
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}%more explanation?
+\label{ev_op_20_80_stubb_PR}
+\end{figure}
+\noindent
 Lets now discuss the evolution of the prevalence of the opinions.
@@
 The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-Bars represent the standard deviation $\sigma$.}%more explanation?
-\label{ev_op_20_80_stubb_PR}
-\end{figure}
-\noindent
+\\
+\newline
 \textbf{All actors partly stubborn}\\
@@
 \newline
-We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate the results for unbalanced starting conditions.
+We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate the results for unbalanced starting conditions.\\
+\newline
+Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod}.
+The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig.
+\ref{echo_vs_frac_complStubb}).\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers.
+This is not changed by making the nodes more and more stubborn.
+The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\
+\newline
+The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
+If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
+This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
+If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
+If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
+This situation is represented by the black dotted line.
 
@@
 \noindent
-Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod}.
-The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig.
-\ref{echo_vs_frac_complStubb}).\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers.
-This is not changed by making the nodes more and more stubborn.
-The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\
-\newline
-The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
-If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
-This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
-If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
-If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
-This situation is represented by the black dotted line.\\
-\newline
 In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
-Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.\\
-\newline
+Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
+\newpage
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REF}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REF_8x7.png}
+    \label{REF_all_frac_stubb}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REC_8x7.png}
+    \label{REC_all_frac_stubb}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
+\textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$.}
+\label{echo_vs_all_frac_stubb_REF-REC}
+\end{figure}
+\noindent
 We can make similar observations as we did in the ``Completely stubborn actors'' paragraph.
@@
 % talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REF}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REF_8x7.png}
-    \label{REF_all_frac_stubb}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REC_8x7.png}
-    \label{REC_all_frac_stubb}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
-\textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
-Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard error $SE$.}
-\label{echo_vs_all_frac_stubb_REF-REC}
-\end{figure}
-\noindent
+\\
+\newline
 What about the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
@@
 Figure \ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Figure \ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
-The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.
+The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+\newline
+In case of the PR filtering algorithm (Fig.
+\ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models if $r \leqslant 0.4$.
+If $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
+This behavior is slightly different than in case of an initial $50/50$ distribution, combined with the PR filtering algorithm (Fig.
+\ref{echo_vs_all_frac_stubb_PR}).
+In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if $r>0.1$).
+This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
+\newline
+As was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
+These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
+Figure \ref{echo_20_80}), as was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.
 
@@
 \noindent
-In case of the PR filtering algorithm (Fig.
-\ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models if $r \leqslant 0.4$.
-If $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
-This behavior is slightly different than in case of an initial $50/50$ distribution, combined with the PR filtering algorithm (Fig.
-\ref{echo_vs_all_frac_stubb_PR}).
-In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if $r>0.1$).
-This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
-\newline
-As was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
-These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
-In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
-Figure \ref{echo_20_80}), as was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
-\newline
 The results for the REC filtering algorithm (Fig.
-\ref{REC_all_frac_stubb_20-80}) resemble those for an initial $50/50$ opinion distribution (Fig.
+\ref{REC_all_frac_stubb_20-80}) resemble those for balanced starting conditions (Fig.
 \ref{REC_all_frac_stubb}).
@@
 This increase can even be observed for the low modularity case, albeit less strong.
-The Erd\H{o}s-R\'{e}nyi network wiggles a bit around a formation of echo chambers equal to one.\\
-\newline
+The Erd\H{o}s-R\'{e}nyi network wiggles a bit around a formation of echo chambers equal to one.
+\newpage
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{ER}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{WS}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{SBM-WS}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}%more explanation?
+\label{ev_op_20_80_all_frac_stubb_PR}
+\end{figure}
+\noindent
 We will now discuss the results for the evolution of the prevalence of the opinions in case of unbalanced starting conditions and making all the nodes partly stubborn.
@@
 
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-Bars represent the standard deviation $\sigma$.}%more explanation?
-\label{ev_op_20_80_all_frac_stubb_PR}
-\end{figure}
-\noindent
 \subsection{The majority threshold model}
@@
 \\
-% explain opinion dynamics or refer to relevant sections!!!
 \newline
@@
 % make coupling to ev of op??
+% explain opinion dynamics or refer to relevant sections!!!
 \begin{figure}[H]
-	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_threshold_PR_12x10.png}
+	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_threshold_PR_10x8.png}
 	\captionsetup{format=plain}
@@
 Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
-The initial opinion distribution was $50/50$.
+The initial opinion distribution is $50/50$.
 The results for the REF and REC filtering algorithms resemble each other.
-Data points are again shown together with their standard deviation $\sigma$ to really emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 100$).
-These large errors are a direct result of the majority model.
+Data points are again shown together with their standard deviation $\sigma$ to emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 150$).
+These large errors are a result of the majority model.
 If we run the majority model once, the model forces the system to converge to a single opinion.
@@
 However, each run the system converges to one of the two opinions, with the result that the number of echo chambers around this opinion strongly increases.
-After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).\\
-\newline
-Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
-This is not due to any network features, but a direct consequence of the usage of the majority model.
-As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes kind of a huge, single echo chamber).
-The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
-This can be seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 10$) than those when using the REC/REF filtering algorithms, an indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.\\
-If $T \geqslant 0.5$, it starts to have an impact on the system.
-However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.6$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
-At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
-The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.
-A constant amount of echo chambers over time is represented by the black dotted line.\\
-\newline
-Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
-The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
-The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
-For each model the REC and PR filtering algorithms are compared.
-The threshold parameter is set equal to $T=0$, so there is no effect of stubbornness.\\
-\newline
-It becomes clear that the PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
-It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
-This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
-Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Figure \ref{sbm_majT}.
-For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
-For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).
 
@@
 \end{figure}
+\noindent
+Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
+This is not due to any network features, but a direct consequence of the usage of the majority model.
+As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes kind of a huge, single echo chamber).
+The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
+This can be seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REC/REF filtering algorithms.
+This is another indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.
+If $T \geqslant 0.5$, it starts to have an impact.
+However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
+At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
+The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.
+A constant amount of echo chambers over time is represented by the black dotted line.\\
+\newline
+Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
+The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
+For each model the REC and PR filtering algorithms are compared.
+The threshold parameter is set equal to $T=0$, so there is no effect of stubbornness.\\
+\newline
+It becomes clear that the PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
+It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
+This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
+Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Figure \ref{sbm_majT}.
+For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
+For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
 
@@
 Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithms.}
+For each model the REC and the PR filtering algorithms are compared.}
 \label{ev_op_majT_one_sim}
@@
 
-This section is devoted to results regarding the way the opinions and/or the stubborn actors are distributed over the network.
-In all the previous results, the two opinions and the stubborn actors (in case there were stubborn actors introduced in the system) were, initially, always randomly distributed across the system.
+This section is devoted to results regarding the way the opinions are distributed over the network.
+In all the previous results, the two opinions were, initially, always randomly distributed across the network.
 It might however be more realistic if people with the same opinion are more or less grouped together initially.
@@
 But we have not yet distributed the opinions according to those groups.
-In subsection \ref{groupVSrandom} we will discuss some results for distributing the opinions in communities versus randomly distributing them.
-
-\subsection{Distributing opinions in communities versus a random opinion distribution}\label{groupVSrandom}
-
-In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion A.
+\\
+\newline
+We study the stochastic block model (SBM) and we give a fraction of the communities a single opinion, namely opinion A.
 The remaining communities are given a $50/50$ opinion distribution.
@@
 Thus, eg.
-if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion A communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$.
-This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
-\newline
-We will look at a network with a high modularity, representing the case of more or less isolated communities, and at a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
-More specifically, the two stochastic block models that are implemented are (both models have a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
-\begin{itemize}
-	\item A stochastic block model with 10 communities of size 100.
-The edge probability inside the communities is $p_{\text{cl}} = 0.1$ and the edge probability between communities is $p_{\text{add}} = 0.001$.
-This is the high modularity network.
-	\item A stochastic block model with 10 communities of size 100.
-The edge probability inside the communities is $p_{\text{cl}} = 0.03$ and the edge probability between communities is $p_{\text{add}} = 0.008$.
-This is the low modularity network.
-\end{itemize} 
-The opinion dynamics model that is implemented is described in Chapter \ref{chap4}, Section \ref{opinionmodel}.
+if we have 10 communities of size 100 and we make $10 \%$ of those communities single opinion A communities while the others get a $50/50$ distribution, the total initial opinion distribution is $55/45$.
+This situation is then compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
+\newline
+We will a network with a high modularity, representing the case of more or less isolated communities, and a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
+More specifically, the two stochastic block models that are implemented are: (i) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ ($Q = 0.908 \pm 0.005$); (ii) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ ($Q = 0.220 \pm 0.007$).
+Both networks have a total number of nodes equal to $N = 10^3$ and average degree $\left<k\right> = 10.79 \pm 0.14$ and $\left<k\right> = 10.17 \pm 0.14$, respectively.\\
+\newline
+The opinion dynamics model is described in Chapter \ref{chap4}, Section \ref{opinionmodel}.
 The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line is $S=20$.
-% look both at PR and REC??
-\\
-\newline
-Both the results for the evolution of the opinions in the system and the formation of echo chambers will be discussed.\\
-\newline
-\textbf{Evolution of the opinion fractions}\\
-\newline
-Lets first discuss how the evolution of the opinions are (potentially) impacted by distributing the opinions in communities instead of randomly distributing them.\\
-\newline
-Figure \ref{commOp0_01_other_50-50} shows the results for the fraction of nodes with opinion A versus time for the case in which a fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the other communities get an initial $50/50$ opinion distribution.
+\\
+\newline
+In subsection \ref{groupVSrandomOP} we will discuss the results for the evolution of the prevalence of the opinions. Subsection \ref{groupVSrandomECHO} will discuss the formation of echo chambers.
+
+\subsection{Evolution of the prevalence of the opinions}\label{groupVSrandomOP}
+
+First, we discuss how the evolution of the opinions are impacted by distributing the opinions in communities instead of randomly distributing them.\\
+\newline
+Figure \ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
+A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the other communities get an initial $50/50$ opinion distribution.
 The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in case of a fraction of $0.1$ single opinion A communities).
@@
 \newline
-The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced initial opinions, if the opinions are randomly distributed.
-This is a result that was already found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
+The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced starting conditions and if the opinions are randomly distributed.
+This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
 Figure \ref{ev_op_20_80}.
-If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion A, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
+If a fraction equal to $0.1$ of the communities is given one opinion, namely opinion A, while the remaining communities get a $50/50$ opinion distribution, we see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
 This makes sense, since the single opinion communities are more or less isolated from the other communities.
@@
 The different communities talk a lot to each other and there is a lot of communication between them.
-As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over all the communities.
+As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over the communities.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
 Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.
-The stochastic block model is implemented with two different modularities.
-The average degree is always around $\left<k\right> \sim 10$ and the total number of nodes is equal to $N = 10^3$.
-\textbf{(a)} High modularity ($p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$), REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity ($p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$), REC; \textbf{(d)} Low modularity, PR.
-Each curve is the average of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.}
+The stochastic block model with 10 communities of size 100 is implemented with two different modularities: $Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$, $\left<k\right> = 10.79 \pm 0.14$ and $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$, $\left<k\right> = 10.17 \pm 0.14$.
+$N = 10^3$ for both modularities.
+\textbf{(a)} High modularity, REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity, REC; \textbf{(d)} Low modularity, PR.
+Each curve is the average of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}
 \label{commOp0_01_other_50-50}
@@
 \noindent
-The results found in Figure \ref{commOp0_01_other_50-50} for a fraction of $0.1$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
-This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average fraction of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
+The results found in Figure \ref{commOp0_01_other_50-50} for $10 \%$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
+This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
 The error bars in this figure are omitted for the sake of an improved visualization.
@@
 \newline
-The higher the fraction of single opinion A communities, the higher the initial average fraction of opinion A.
-This can be seen in both figures by the increase of the solid black line, which represents the initial fraction of nodes with opinion A.
-In case of the REC filtering algorithm, Figure \ref{rec_comm}, the fraction of opinion A remains the same after the time evolution for a fixed fraction of single opinion A communities.
+The higher the fraction of single opinion A communities, the higher the initial average prevalence of opinion A.
+This is seen in both figures by the increase of the solid black line, which represents the initial prevalence of opinion A.
+In case of the REC filtering algorithm (Fig.
+\ref{rec_comm}) the prevalence of opinion A remains the same after the time evolution for a fixed fraction of single opinion A communities.
 Giving a fraction of the communities a single opinion, opinion A, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
@@
 \newline
-The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the fraction of opinion A, which is the majority opinion, over time for a fixed fraction of single opinion A communities.
+The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the prevalence of opinion A over time for a fixed fraction of single opinion A communities.
 Making a fraction of the communities single opinion A communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
-The low modularity network does not show a sign of this hampering.
-The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was already explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
+The low modularity network does not exhibit this hampering.
+The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
 
@@
   \captionsetup{format=plain}
-  \caption[Average fraction of opinion A before and after the time evolution versus the fraction of single opinion A communities.
+  \caption[Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
 Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Average fraction of opinion A before and after the time evolution versus the fraction of single opinion A communities.
-Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
-\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+Results for the REC and PR filtering algorithms.]{Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
+Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
+\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
 Each curve is the average of $10^2$ independent simulations.}
@@
 \end{figure}
-\noindent
-\textbf{Echo chambers}\\
+
+\subsection{Echo chambers}\label{groupVSrandomECHO}
+
+We will investigate the effect of community distributing the opinions versus randomly distributing them on the formation of echo chambers. \\
 \newline
 Figure \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
-As in similar plots before, the value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The community distributed case is compared to the corresponding randomly distributed case (which is, in this case, an initial $55/45$ opinion distribution that is randomly distributed in each community).
+The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
+The community distributed case is compared to the corresponding randomly distributed case (which is, in this case, an initial $55/45$ opinion distribution).
 The formation of echo chambers in both the high and low modularity network are investigated.
-Figure \ref{rec_comm_echo} shows the results for the REC filtering algorithm, while Figure \ref{pr_comm_echo} shows those for the PR filtering algorithm.
+The REC and PR filtering algorithms are compared to each other.
 Error bars are again omitted for an improved visualization.\\
 \newline
-Lets first discuss the results for the REC filtering algorithm.
-The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line).
-The number of neighborhoods where all the nodes carry opinion A thus decreases in favor of neighborhoods where we find both nodes with opinion A and nodes with opinion B.
-Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion A communities, reduces and that the opinions become more mixed.
-The fraction of opinion A remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion A see a decrease in the nodes with opinion A and become more mixed.
-We can also observe a very small increase in the number of echo chambers around the minority opinion B, which indicates the fact that this minority opinion starts to group more together.
+The results for the REC filtering algorithm are depicted in Figure \ref{rec_comm_echo}.
+The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (bump in the blue line).
+The number of neighborhoods where all the nodes carry opinion A decreases in favor of neighborhoods where we find both nodes with opinion A and nodes with opinion B.
+Even though there is only a minor communication between the communities, this limited communication has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion A communities, reduces and that the opinions become more mixed.
+The prevalence of opinion A remains constant over time when using the REC filtering algorithm (Fig.
+\ref{high_mod_rec}).
+The communities with the single opinion A see, however, a decrease in the prevalence of opinion A and become more mixed.
+We can also observe a small increase in the number of echo chambers around the minority opinion B.
+This indicates that the minority opinion starts to group together.
 The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion A and opinion B.
@@
 \newline
-The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion A, which could be observed for the community distributed, high modularity case.
-This is due to the fact that the number of echo chambers around opinion A at $t=0$, originating from the fraction of single opinion A communities, is a lot smaller for the low modularity network than for the high modularity network.
-The reason for this is the much more blended community structure for the low modularity network.
-We can still observe a very small increase in the number of echo chambers around opinion B.
+The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion A (which was observed for the community distributed, high modularity case).
+This is due to the fact that the number of echo chambers around opinion A at $t=0$, originating from the fraction of single opinion A communities, is smaller for the low modularity network than for the high modularity network.
+The reason is the more blended community structure for the low modularity network.
+We still observe a small increase in the number of echo chambers around opinion B.
 This increase in the formation of echo chambers around the minority opinion B seems however hampered compared to the randomly distributed case.
-Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80}, see Figure \ref{sbm}.
+Remember that the increase in the number of echo chambers around the minority opinion for the SBM in combination with the REC filtering algorithm for unbalanced starting conditions that are randomly distributed was observed in Section \ref{netw_struc}, Subsection \ref{20-80} (Fig?
+\ref{sbm}).
 \\
@@
 The results for the PR filtering algorithm are depicted in Figure \ref{pr_comm_echo}.
-They are quite different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Figure \ref{rec_comm_echo}.\\
+They are different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Figure \ref{rec_comm_echo}.\\
 \newline
 Lets first focus on the high modularity network.
-The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (note that these results were obtained for an initial $20/80$ opinion distribution).
+The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (these results were obtained for an initial $20/80$ opinion distribution).
 We can observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion B.
 Remember from Figure \ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion A.
-If we, however, make a fraction equal to $0.1$ of the communities single opinion A communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion A is strongly hampered.
+If we, however, make $10 \%$ of the communities single opinion A communities while giving the remaining ones a $50/50$ opinion distribution, we see that this increase in the formation of echo chambers around the majority opinion A is strongly hampered.
 This hampering is partly an artifact of the way the opinions are distributed.
@@
 If we, however, would look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
-This shows that the observed hampering of the formation of echo chambers around the majority opinion A in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion A.
-Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed cases versus the random distributed case.
+This shows that the observed hampering of the formation of echo chambers around the majority opinion A in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers.
+Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed case versus the random distributed case.
 Beside this hampering of echo chambers around the majority opinion A, the community distributed case also shows an increase of echo chambers around the minority opinion B (which is not visible for the randomly distributed case).
@@
 \newline
-The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network.
+The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases, as was observed for the high modularity network.
 Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion B.
-From Figure \ref{low_mod_pr}, it was already made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of this minority opinion (when using the PR filtering algorithm).
-This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry this minority opinion.
+From Figure \ref{low_mod_pr}, it was made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of the minority opinion (when using the PR filtering algorithm).
+This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry the minority opinion.
 The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case.
@@
   \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
-A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
-\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
+\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
 Each curve is the average of $10^2$ independent simulations.}
@@
 \noindent
-Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, but now a fraction equal to $0.5$ of the communities is only given opinion A, while the others get a $50/50$ distribution.
-The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion A (Figure \ref{echo_01_commOp0_other_50-50}).\\
-\newline
-The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
-A difference with the case where a fraction equal to $0.1$ of the communities are single opinion A communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion B, but instead we observe a decrease in the number of echo chambers around the minority opinion.
+Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, $50 \%$ of the communities is given opinion A, while the others get a $50/50$ distribution.
+The results look more or less the same as for the case where $10 \%$ of the communities is given opinion A (Fig.
+\ref{echo_01_commOp0_other_50-50}).\\
+\newline
+The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
+A difference with the case where $10 \%$ of the communities are single opinion A communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion B.
+Instead we observe a decrease in the number of echo chambers around the minority opinion.
 For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion A and a minority of the nodes carrying the minority opinion B.
@@
 The low modularity, community distributed case seems to wiggle around $1$.
-If we, however, look closely, we can see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller).
-We can observe a small decrease in the formation of echo chambers around the majority opinion A, with a small increase in more mixed opinion neighborhoods as result.
-We can also observe a small decrease in the number of echo chambers around the minority opinion A.
-The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion A communities, while the others get a $50/50$ opinion distribution.\\
+If we look closely, we see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller).
+We observe a small decrease in the formation of echo chambers around the majority opinion A, with a small increase in more mixed opinion neighborhoods as result.
+We also observe a small decrease in the number of echo chambers around the minority opinion B.
+The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of $50 \%$ of the communities being single opinion A communities, while the others get a $50/50$ opinion distribution.\\
 \newline
@@
 The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion A.
-This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (see Figure \ref{high_mod_pr} and Figure \ref{pr_comm}).
+This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (Fig.
+\ref{high_mod_pr} and Fig.
+\ref{pr_comm}).
 The increase in the number of echo chambers around the minority opinion A that was visible in Figure \ref{pr_comm_echo} is no longer present.
@@
   \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
-A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+$50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$.
-\textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm.
+$50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
+\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
 Each curve is the average of $10^2$ independent simulations.}
@@
 
-We made use of simple toy models of real social media platforms to generate results regarding the evolution of opinions and (possible) formation of echo chambers on these platforms.
+We made use of simple toy models of real social media platforms to generate results regarding the evolution of the prevalence of the opinions and formation of echo chambers.
 These toy models consist of several layers, including a network layer and an opinion dynamics layer.
-Simple filtering algorithms were implemented to replicate and simulate the content curation used by these real social media companies.
-Furthermore, a notion of stubbornness was introduced to investigate the interplay between this individual resistance to change, network structure and filtering algorithm.
-For simplicity we only allowed two competing opinions (A and B) in the system.\\
+Simple filtering algorithms were implemented to replicate and simulate the content curation used by real social media companies.
+Furthermore, a notion of stubbornness was introduced to investigate the interplay between individual resistance to change, network structure and filtering algorithm.
+For simplicity we only allowed two competing opinions (A and B).\\
 \newline
@@
 This indicates that it is mainly the high clustering that drives the formation of echo chambers.
-There is however a clear correlation between modularity and formation of echo chambers: a higher modularity results in a larger formation of echo chambers (in case of the PR filtering algorithm).
+There is a correlation between modularity and formation of echo chambers: a higher modularity results in a larger formation of echo chambers (in case of the PR filtering algorithm).
 The enhancement is just not as strong as the enhancement in the formation of echo chambers introduced by a high clustering.
@@
 \newline
-When introducing a notion of stubbornness in the system (with the help of a stubbornness parameter $r$), we found that the interplay between the PR filtering algorithm and this individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
+When introducing a notion of stubbornness (with the help of a stubbornness parameter $r$), we found that the interplay between the PR filtering algorithm and this individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
 The REF and REC filtering algorithms, on the other hand, showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if the stubbornness was further increased.
-This is mainly seen in the high clustering networks, however a tiny bump can also be observed in the high modularity network.
-This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen in both the case of balanced initial conditions and unbalanced initial conditions.\\
+This is mainly seen in the high clustering networks, however a small bump can also be observed in the high modularity network.
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.\\
 % talk about evolution of opinions??
 \newline
-It was also found that how the opinions are initially distributed over the communities in a network with a community structure (SBM) does makes a difference.
-We investigate a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ distribution.
+It was also found that the way the opinions are initially distributed over the communities in a network with a community structure (SBM) does make a difference.
+We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ distribution.
 This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.

### 2021-05-03 draft.tex
made some small changes in ch. 1 and ch. 2

@@
 \\
-Beside that, we will also pay some attention to the possible ways of distributing the opinions and the stubborn actors and how this may effect the opinion dynamics in the system. \\
+Beside that, we will also pay some attention to the possible ways of distributing the opinions and how this may affect the opinion dynamics in the system. \\
 Finally, we will test the model on real-world networks.
-Some question that may arise are: Do the results agree with those obtained from the network models?
+Some question that may arise are: Do the results agree with those obtained from the theoretical network models?
 What are the similarities or differences in the results?
@@
 It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
-One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
+It is important to understand the effects of individual resistance on the formation and evolution of opinions.
+When designing models of opinion dynamics on social networks, this individual stubbornness should not be ignored.\\
 \newline
@@
 Ideally, our theoretical network should exhibit these three properties.
-Some theoretical networks, that are thoroughly studied, do not possess all of these three properties.
+Some theoretical networks, that are thoroughly studied, do not possess all of them.
 However, they might still be used, because of their simplicity and ability to produce analytic results.
@@
 The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
-These models are obviously a simplification of real-world opinion dynamics.
-It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
+These models are a simplification of real-world opinion dynamics.
+It is however shown that they display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
 \\
@@
 The opinion of the agents in the model can be either discrete or continuous.
-This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions.
+This thesis deals with models where each agent can have one of the two opinions, A or B, and thus only deals with the case of discrete opinions.
 This might come across as a huge simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
@@
 An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
-Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
+This interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
 
@@
 
-As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
+This thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
 On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn actors in the network.
@@
 \newline
-In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}) Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
+In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, \cite{Perra2019}) Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
 They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).
 Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}.
-On the other hand, a more heterogeneous contact pattern could hamper this formation of echo chambers \cite{Perra2019}.
+On the other hand, a more heterogeneous contact pattern hampers this formation of echo chambers \cite{Perra2019}.
 If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
@@
 We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
-We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can, for example, observe convergence to one opinion inside the communities.
-We will also pay some attention to networks that have both a high clustering coefficient and a community structure.
+We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient.
+We will also pay some attention to networks that have both a high clustering coefficient and community structure.
 Can we observe echo chambers in those kinds of networks?
@@
 Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness?
-What about the possible interplay between the filtering algorithms and the introduction of stubborn actors?
-Does the distribution of stubborn actors play a role (eg.
-randomly distributed versus clustered together inside communities)?
-\\
+What about the possible interplay between the filtering algorithms and the introduction of stubborn actors?\\
 % partly rewrite this according to results? Convergence to one opinion in communities --> discussed? Clustered stubborn people --> discussed? Include clustered opinions? Include real-world networks?
@@
 
-Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
+Networks or graphs are conceptually simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
 \begin{equation}
@@
 
-Several very useful measurements quantify the notion of length and distance in a network/graph.
+Several useful measurements quantify the notion of length and distance in a network/graph.
 We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
 The length of a path is then defined as the number of edges in the path \cite{Goddard2010}.
-A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}.
+A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (the geodesic path is not necessarily unique) \cite{F.Costa2007}.
 The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}.
 Geodesic distance is often just called distance \cite{Goddard2010}.
-The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\
+The notion of distance enables to define several global network measurements such as average distance, diameter and radius.\\
 \newline
@@
 These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}.
-Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
+Instead, another definition of average distance is used: the global efficiency $E$ \cite{F.Costa2007}
 \begin{equation}
@@
 \end{equation}
-where $n_i$ is the number of edges that actually exist between the nodes in the neighborhood of $i$ and $k_i(k_i-1)/2$ is the maximum number of edges that could exist between them (note that this expression is only valid for an undirected graph, since for a directed graph $e_{ij} \neq e_{ji}$ and we have $k_i(k_i-1)$ possible edges between the neighbors of node i).
+where $n_i$ is the number of edges that actually exist between the nodes in the neighborhood of $i$ and $k_i(k_i-1)/2$ is the maximum number of edges that could exist between them (this expression is only valid for an undirected graph, since for a directed graph $e_{ij} \neq e_{ji}$ and we have $k_i(k_i-1)$ possible edges between the neighbors of node i).
 % should give more mathimatical expression for n_i?
@@
 \end{equation}
-where $N$ denotes, once again, the number of nodes in the network.
-The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Heer2020}\cite{Iyer2013} \cite{Li2017}.
+where $N$ denotes the number of nodes in the network.
+The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage,  of the network \cite{Heer2020}\cite{Iyer2013}\cite{Li2017}.
 \\
@@
 where $L_c$ is the total number of edges in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.\\
+\newline
 To generalize this idea to the whole network, we sum Eq. (\ref{Qc}) over all $n_c$ communities \cite{Albert2016}
@@
 \end{equation}
-The derivation of this formula and thus of Eq.
-(\ref{Qc}) can be found in Appendix \ref{modul}, Section \ref{simplemod}.
-The modularity $Q$ can have values in the interval $\interval{-1}{1}$ and has the following properties \cite{Albert2016}
-\begin{itemize}
-	\item The higher the value of $Q$ is for a certain partition, the better is the corresponding community structure.
-This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure.
-	\item A modularity $Q$ equal to zero corresponds to the situation where all nodes are put in the same community, this is if the whole network is put into a single community.
+The derivation of this formula, and thus of Eq.
+(\ref{Qc}), can be found in Appendix \ref{modul}, Section \ref{simplemod}.
+The modularity $Q$ can have values in the interval $\interval{-1}{1}$ and has the following properties \cite{Albert2016}: (i) The higher the value of $Q$ for a certain partition, the better the corresponding community structure.
+This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure; (ii) A modularity $Q$ equal to zero corresponds to the situation where all nodes are put in the same community, this is, if the whole network is put into a single community.
 Then the two terms in Eq.
-(\ref{mod}) become equal and $Q$ vanishes.
-	\item If every node is put into a separate community, the first term in Eq. (\ref{mod}) vanishes and we are left with $n_c$ negative terms. Hence, this situation corresponds to a negative modularity.
-\end{itemize}
+(\ref{mod}) become equal and $Q$ vanishes; (iii) If every node is put into a separate community, the first term in Eq.
+(\ref{mod}) vanishes and we are left with $n_c$ negative terms.
+Hence, this situation corresponds to a negative modularity.\\
+\newline
 As said before, modularity is often used in community detection algorithms.
@@
 Independently of them, Solomonoff and Rapoport already proposed the model in 1951.
-The model constructs random networks with $N$ nodes in the following way \cite{Albert2002}
-\begin{enumerate}
-	\item Start with $N$ disconnected nodes,
-	\item For each pair of nodes, connect them with a predefined probability $p$.
-\end{enumerate} 
+The model constructs random networks with $N$ nodes in the following way \cite{Albert2002}: (i) Start with $N$ disconnected nodes; (ii) For each pair of nodes, connect them with a predefined probability $p$.\\
+\newline
 If $p=1$, we obtain a network with the maximum number of edges $N(N-1)/2$. This is called the fully connected network.\\
@@
 \end{equation}
-This equation can be simplified to $\left<k\right> = (N-1)p$, and for $N \gg 1$ this approximately becomes $\left<k\right> = Np$.
+This equation can be simplified to $\left<k\right> = (N-1)p$, and for $N \gg 1$ this approximately becomes $\left<k\right> \approx Np$.
 The derivation of this result can be found in Appendix \ref{ER}, Section \ref{degree}.\\
-\newline
+\newpage
+\noindent
 Most real networks are sparse, which means that $\left<k\right> \ll N$. In this case the degree distribution of the random network is well approximated by a Poisson distribution \cite{Albert2014} 
@@
 Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}.
-The construction goes as follows \cite{Watts1998}
-\begin{enumerate}
-	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side).
-	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
-Rewiring is done in such a way that self-loops and edge duplication are avoided.
-\end{enumerate}
+The construction goes as follows \cite{Watts1998}: (i) Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side); (ii) For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
+Rewiring is done in such a way that self-loops and edge duplication are avoided.\\
+\newline
 The probability $\beta$ allows us to tune the graph between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}.
@@
 \newline
-The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
+The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
 (\ref{C}) and Eq.
@@
 \end{equation} 
-where $\left<cc(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
-\begin{itemize}
-	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$.
-	\item The probability that edges are rewired back to each other is negligible.
-\end{itemize}
+where $\left<cc(0)\right>$ is given by Equation (\ref{C0}).
+This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
+\newline
 For rather small values of $\beta$, the clustering coefficient remains high.
@@
 \newline
-The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}
-\begin{itemize}
-	\item The number of nodes $N$, 
-	\item A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities,
-	\item A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.
-\end{itemize}
+The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}: (i) The number of nodes $N$; (ii) A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities; (iii) A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.\\
+\newline
 There are no real restrictions on the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures \cite{Karrer2011}.
@@
 However, a network with a skewed degree distribution violates this assumption of edge independence \cite{Clauset2017}.
-To say it in the words of Karrer and Newman: ``Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...].'' (Karrer and Newman, p.1, reference \cite{Karrer2011}).
+To say it in the words of Karrer and Newman: ``Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...].'' (Karrer and Newman, p.1, \cite{Karrer2011}).
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
@@
 The construction of the model is very similar to that of the regular stochastic block model.
-For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $K$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
+For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $\left<k\right>$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
 The result is a network that can be tuned to have both a community structure and a high clustering coefficient.

### 2021-05-03 draft.tex
made some small changes in ch.3 and ch. 4

@@
 The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
-The current chapter focuses on the second important layer, namely that of the opinion dynamics models.
+The current chapter focuses on the second important layer, that of the opinion dynamics models.
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
@@
 \newline
-Some major attention will also be given to the content curation/filtering algorithms used by social media companies (see Section \ref{filter}).
+Some major attention will also be given to the content curation/filtering algorithms used by social media companies (Section \ref{filter}).
 This is needed because this thesis deals with opinion dynamics on on-line social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
 \newline
-There will also be some attention on the activation mechanism (see Section \ref{actMech}).
+There will also be some attention on the activation mechanism (Section \ref{actMech}).
 This is the mechanism that describes the timings in which content/information is exchanged between users \cite{Perra2019}.
-Such a mechanism may change the network from a static one to a more temporal one, where nodes become active and inactive over time.
+Such a mechanism changes the network from a static one to a more temporal one, where nodes become active and inactive over time.
 \\
@@
 The following updating rule is then imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
-Thus, initially each agent is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
-\newline
-Several studies have been performed on this kind of model, searching for answers on several questions such as: what are the conditions that drive the system to converge to unanimity; and, does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
+Initially each agent is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
+\newline
+Several studies have been performed on this kind of model, searching for answers on several questions such as: What are the conditions that drive the system to converge to unanimity; and, Does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
 \\
@@
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
-This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016} (also note that even when the opinions are equally distributed, there may still be a slight imbalance due to the random assignement of an opinion to each node; instead of a perfect 50/50 distribution, one may end up with a eg.
-49.8/50.2 distribution; in this sense there is still some kind of notion of `minority' and `majority' opinion).
+This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016}.
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
@@
 The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature.
-In this model the fraction of neighbors of user $i$ with opinion A and the fraction of neighbors with opinion B are determined.
-Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
+In this model the fraction of neighbors of user $i$ with opinion A, $p_\text{A}$, and the fraction of neighbors with opinion B, $p_\text{B}$, are determined.
+Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to $p_\text{A}$ (or $p_\text{B}$) \cite{Perra2019}.\\
 \newline
@@
 
-Another simple binary model, maybe even simpler than the majority model, is the voter model.
+Another simple binary model is the voter model.
 The model works as follows: each agent in the network has a binary opinion (A or B); at each time step, an agent is randomly selected along with one of its neighbors and, subsequently, the agent adopts the opinion of this neighbor \cite{Castellano2009}.
@@
 
-In bounded confidence models, each node has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance (note that $\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
+In bounded confidence models, each node has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance ($\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
 This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
@@
 The final state is a situation where several opinion clusters coexist.
-This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (so the difference in opinions between all agents in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
+This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (the difference in opinions between all agents in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
 Only agents inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
@@
 \newline
-The Hegselmann-Krause model is very similar to the Deffuant model.
+The Hegselmann-Krause model is similar to the Deffuant model.
 Again, there are $N$ agents, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
@@
 The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary opinion clusters \cite{Castellano2009}.
-The number of clusters decreases if $\epsilon$ increased and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
+The number of clusters decreases if $\epsilon$ increases and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
 \newline
@@
 \epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to references \cite{Deffuant2006} and \cite{Deffuant2002}.
-
+\newpage
 \section{Algorithmic personalization}\label{filter}
@@
 If social media companies would not use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
-This in turn would make the on-line social experience a lot more unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
+This in turn would make the on-line social experience unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
 The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that humans can process \cite{Bozdag2013}.\\
 \newline
-The backside of such algorithmic filtering is that it fine-tunes the information it shows based on the user.
+The downside of such algorithmic filtering is that it fine-tunes the information it shows based on the user.
 Even two users with the same friends may then be exposed to different content, based on their previous activities and history \cite{Bozdag2013}.
@@
 It is important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
-Perra and Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
-Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}
-\begin{itemize}
-	\item Popularity filtering: promoting content that is popular across the platform,
-	\item Semantic filtering: promoting posts that are similar to previous consumed posts,
-	\item Collaborative filtering: promoting content that people similar or close to us consume.
-\end{itemize} 
+Perra and Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} \cite{Perra2019}.
+Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}: (i) Popularity filtering: promoting content that is popular across the platform; (ii) Semantic filtering: promoting posts that are similar to previous consumed posts; (iii) Collaborative filtering: promoting content that people similar or close to us consume.\\
+\newline
 The following paragraph briefly reviews the filtering algorithms that Perra and Rocha proposed. \\
@@
 At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}
-\begin{itemize}
-	\item The reference filtering algorithm (REF): this filtering algorithm randomly selects $S$ items from the list $L_i(t)$.
-	\item The old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user.
-	\item The recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$.
-	\item The preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
-This filtering algorithm is based on the semantic filtering explained above.
-\end{itemize}
+The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}: (i) The reference filtering algorithm (REF): this filtering algorithm randomly selects $S$ items from the list $L_i(t)$; (ii) The old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user; (iii) The recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$; (iv) The preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
+This filtering algorithm is based on the semantic filtering explained above.\\
+\newline
 Perra and Rocha showed that the REF, OLD and REC filtering algorithms behave similar in many cases and that the PR filtering algorithm often displays different behavior (such as a larger formation of echo chambers compared to the other three filtering algorithms) \cite{Perra2019}.
@@
 
-In the most simple opinion dynamics models, which make use of a static network, users are always active and thus always exchanging opinions/posts with each other.
+Most of the simple opinion dynamics models make use of a static network in which users are always active and always exchanging opinions/posts with each other.
 In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented.
@@
 The underlying network structure does not change (in the sense that each user/node has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
-Hence, in this sense, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
+Hence, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
 This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity, clustering coefficient, etc may behave differently.
@@
 \newline
-In this thesis we follow the same activation mechanism as was proposed in the paper of Perra and Rocha (see reference \cite{Perra2019}).
+In this thesis we follow the same activation mechanism as was proposed in the paper of Perra and Rocha \cite{Perra2019}.
 Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$).
@@
 \newline
-If a user is active, he/she first spreads his current opinion to all his/her neighbors (for example by posting something).
+If a user is active, he/she first spreads his current opinion to all his/her neighbors.
 Then, the active user updates his own opinion according to the opinions on his/her time-line \cite{Perra2019}.
@@
 majority model, probabilistic majority model,...).
-In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.\\
-\newline
-It is important to note that what somebody sees on his/her time-line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors.
-Every user has a ``hidden time-line $L_i(t)$'' which contains the actual time-line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}.
-At the time $t$ that the user becomes active again, the filtering algorithm selects $S$ post from this ``hidden time-line $L_i(t)$'' and shows them on the actual time-line of the user; these $S$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).
+In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.
 
@@
 
-It is well known that in real life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others.
+It is well known that in real life, you can find people that are persistent towards their own opinion and that are not open to opinions of others.
 They are convinced that they carry the truth with them and that other people, with different world-views, are not capable of giving them any new and useful insights.
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme does not need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
-\newline
-Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups a lot bigger.
+However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+\newline
+Knowing the impact of stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups a lot bigger.
 Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore would not reach too many people.
-So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
+So their impact could only be of a minor nature (this is only a statement that could have been given, there are of course plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
 Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
@@
 The model he proposed goes as follows.
-Remember that in the bounded confidence models, the agents have an opinion $x_i$ (which is a real number, often between $\interval{0}{1}$ or between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
+Remember that in the bounded confidence models, the agents have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
 Guillaume Deffuant then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or the extreme opinion $1$ \cite{Deffuant2006}.
@@
 The moderates are given a threshold $\epsilon$, with $\epsilon > \epsilon_e$ \cite{Deffuant2006}.
-This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is again based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
+This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
 \newline
@@
 
-Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models.
+Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is not possible for binary/discrete models.
 There are, however, many other ways to include some kind of stubborn actors in these models.
@@
 In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions.
-However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.
-Whether this resistance is a sign of critical thinking or ignorance/being convinced that you know it best, is up to the reader to decide.
-The models simply introduce some type of resistance and make no distinction between possible underlying reasons for this resistance.\\
+However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
 \newline
@@
 One way of creating stubborn actors in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
-This resistance parameter can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$.
-\begin{itemize}
-	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...).
-	\item If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea.
-	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
+This resistance parameter can have values between 0 and 1.
+If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...).
+If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea.
+If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
 An uniform random number $f$ between 0 and 1 is drawn.
-If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.
-\end{itemize}
+If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.\\
+\newline
 The resistance parameter $r_i$ can be different or the same for all nodes.
@@
 The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
-The more persistent you are, the more friends you need in order to be convinced of another idea.
+The more persistent you are, the more friends you need in order to be convinced of the other idea.
 \\
@@
 One way, which makes use of the majority model, is as follows.
-First the fraction of friends with opinion A and opinion B, $p_A$ and $p_B$ respectively, are determined; then the following rules are used to update the opinion
-\begin{itemize}
-	\item If $p_A > p_B$ and $p_A > T$, opinion A is adopted,
-	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted,
-	\item In all other cases, the current opinion of the node is not changed.
-\end{itemize}
+First the fraction of friends with opinion A and opinion B, $p_\text{A}$ and $p_\text{B}$ respectively, are determined; then the following rules are used to update the opinion: if $p_\text{A} > p_\text{B}$ and $p_\text{A} > T$, opinion A is adopted; if $p_\text{B} > p_\text{A}$ and $p_\text{B} > T$, opinion B is adopted; in all other cases, the current opinion of the node is not changed.\\
+\newline
 If $T = 0$, this model reduces to the pure majority model described in Section \ref{majModel}.
@@
 \newline
-Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study, many more exist or can be invented.
+Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model, many more exist or can be invented.
 
@@
 \newline
-Section \ref{networkmodel} gives an overview of the network models that are used in this thesis.
-In chapter \ref{chap2}, Section \ref{netModel} we already discussed their main properties; here we will mainly discuss why we use these networks.
+Section \ref{networkmodel} gives an overview of the network models that are used.
+In chapter \ref{chap2}, Section \ref{netModel} we discussed their main properties; here we will discuss why we use these networks.
 Which of their properties are of interest in this thesis?
-What are there benefits and possible shortcomings regarding to our goals?\\
-\newline
-In section \ref{realnetwork}, we will discuss the real-world networks used in this thesis.
-The reasons of choosing these specific networks will be explained, together with there possible strengths and weaknesses.\\ % + explain where they come from etc?
+What are their benefits and possible shortcomings regarding to our goals?\\
+\newline
+In section \ref{realnetwork}, we will discuss the used real-world networks.
+The reasons of choosing these specific networks will be explained, together with their possible strengths and weaknesses.\\ % + explain where they come from etc?
 \newline
@@
 
-This thesis uses three main network models, namely
-\begin{itemize}
-	\item The Erd\H{o}s-R\'{e}nyi model (ER),
-	\item The Watts-Strogatz model (WS),
-	\item The stochastic block model, of which two types are implemented
-		\begin{itemize}
-			\item[\tiny$\bullet$] The regular stochastic block model (SBM),
-			\item[\tiny$\bullet$] The stochastic Watts-Strogatz block model (SBM-WS).
-		\end{itemize}
-\end{itemize}
+This thesis uses three main network models: (i) The Erd\H{o}s-R\'{e}nyi model (ER); (ii) The Watts-Strogatz model (WS); (iii) The stochastic block model, of which two types are implemented: the regular stochastic block model (SBM) and the stochastic Watts-Strogatz block model (SBM-WS).\\
+\newline
 Each of these networks serve specific goals which will shortly be discussed in the following paragraphs.\\
@@
 \newline
-The Erd\H{o}s-R\'{e}nyi model is mainly used as a benchmark model.
-In this thesis we focus intensively on two network features/topologies that may impact the formation of echo chambers, namely a high clustering coefficient and a community structure.
+The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
+In this thesis we focus intensively on two network features/topologies that may impact the formation of echo chambers, namely a high clustering coefficient and community structure.
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
@@
 \newline
-It is expected that the random nature of the model and the lack of a high clustering and a community structure will result in the absence of echo chambers.\\
+It is expected that the random nature of the model and the lack of high clustering and community structure will result in the absence of echo chambers.\\
 \newline
@@
 \newline
-The Watts-Strogatz model is used in this thesis as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-\newline
-In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}) Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
-Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and, on the other hand, we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
-We will investigate whether a community structure behaves similar to a high clustering regarding the appearance of echo chambers.\\
+The Watts-Strogatz model is used in as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
+\newline
+In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (reference \cite{Perra2019}) Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
+Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.\\
 \newline
@@
 \newline
-One of the main features of the stochastic block model is the ability to generate networks with a community structure.
+One of the main features of the stochastic block model is the ability to generate networks with community structure.
 We will therefore use this model to investigate the impact of community structure/modularity on the evolution of opinions and the formation of echo chambers.\\
 \newline
-It is expected that a community structure will enhance the formation of echo chambers compared to the random network case.
-The question remains, however, whether the influence of a community structure on the formation of echo chambers is of the same caliber as that of a high clustering.
-It is also of interest to investigate the possible relation between the appearance of echo chambers and the modularity of the network.\\
+It is expected that community structure will enhance the formation of echo chambers compared to the random network case.
+The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering.
+It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
 \newline
@@
 All the above models have a homogeneous degree distribution.
-In this thesis, we did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
+We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
 It might, however, be an interesting extension to investigate such types of models as well.\\
@@
 They are the most simple network models that one can create and each of them have characteristics and topologies that can be easily tuned (clustering for WS, community structure for SBM).
-This makes it really easy to investigate the effect of different network topologies (clustering, community structure) without the need of taking the interplay with other possible complex network features and topologies into account.
+This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account.
 This plus side is, however, also one of the possible downsides.
@@
 But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world.
-Results that are obtained from them should not be thoughtlessly extrapolated to real-world networks.
+Results that are obtained from them should not be thoughtlessly extrapolated to real-world situations.
 

### 2021-05-04 draft.tex
worked on ch.4 real-world networks

@@
 The degree distribution can eg.
-also be a power-law with an exponential cut-off or a log-normal distribution.\\
+also be a power-law with an exponential cut-off, a log-normal distribution, a stretched exponential, etc \cite{Newman2005}.\\
 \newline
@@
 \newline
-The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.\\
-\newline
-The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis.
-But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world.
-Results that are obtained from them should not be thoughtlessly extrapolated to real-world situations.
+The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.
 
@@
 \newline
-One of the decisive properties of the real-world networks that were eventually used, was there size.
+One of the decisive properties of the real-world networks that were eventually used, was their size.
 The main, and actually only, reason for this was the computation time.
@@
 The size of the network models that were used, was around a thousand nodes.
-So for the real-world networks, we also used networks with a size of magnitude of thousand.
-This obviously benefits the computation time; it is, however, unknown how the network size affects the results.
+For the real-world networks, we also used networks with a size of magnitude of thousand to ten-thousand.
+This benefits the computation time; it is, however, unknown how the network size affects the results.
 It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic.
+\\
+\newline
+We used two real-world networks.
+The data was obtained from \textit{The index of Complex networks} (ICON) \cite{ICON}.
+The first real-world network is an on-line friendship network among Finnish users of the social music listening and sharing website Last.fm (data from $2006$).
+Only the largest connected component of this on-line social network is included \cite{ICON}.
+This data set was originally used in reference \cite{Toivonen2009}, which describes the data set.
+This on-line social network has the following properties: number of nodes: $N = 8003$; number of edges: $16824$; modularity: $Q = 0.79$; average clustering coefficient: $\left<cc\right> = 0.31$; average degree: $\left<k\right> = 4.2$.
+The degree distribution is heterogeneous and is best fitted by a log-normal
+\begin{equation}
+	f(k) = \frac{1}{k \sigma \sqrt{2\pi}} \exp(-\frac{(\ln(k) - \mu)^2}{2 \sigma^2})\ ,
+\end{equation}
+with $\mu = 1.0078272$ and $\sigma = 0.95173779$. The degree distribution of the Last.fm social network, together with the log-normal fit, is shown in Figure \ref{deg_distr_lastfm}.\\
+\newline
+The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{ICON}.
+This network was originally used in reference \cite{Boguna2004}, which gives a detailed description of the network data.
+The main properties of this network are: number of nodes: $N = 10680$; number of edges: $24340$; modularity: $Q = 0.85$; average clustering coefficient: $\left<cc\right> = 0.27$; average degree: $\left<k\right> = 4.56$.
+The network has a heterogeneous degree distribution, which is best fitted by two power-laws
+\begin{equation}
+ f(k) = a k^{-\gamma} \ ,
+\end{equation}
+with $a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$.
+The degree distribution for the PGP trust network, together with the power-law fits, is shown in Figure \ref{deg_distr_pgp}.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{Last.fm}
+    \includegraphics[width=8cm, height=7cm]{lastfm_degree_distribution_8x7.png}
+    \label{deg_distr_lastfm}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PGP}
+    \includegraphics[width=8cm, height=7cm]{PGP_degree_distribution_8x7.png}
+    \label{deg_distr_pgp}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real world networks used in this thesis.
+\textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.0078272$, $\sigma = 0.95173779$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$).}
+\label{ev_op_50_50}
+\end{figure}
 
@@
 
-In Chapter \ref{chap3} we already discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
-The goal of this section is to put every thing together and summarize the opinion dynamics model of this thesis.\\
-\newline
-This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}.
-\\
-\newline
-We start with a network with $N$ users.
+In Chapter \ref{chap3} we discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
+The goal of this section is to put everything together and summarize the opinion dynamics model of this thesis.\\
+\newline
+This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} \cite{Perra2019}.
+\\
+\newline
+We start with a network with $N = 10^3$ users.
 Initially, each user is given one of the two opinions A or B.
@@
 This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}.
-If another model instead of the probabilistic majority model is used, eg.
+If another model is used, eg.
 the majority model, this will be explicitly stated.
@@
 It is important to realize that what an active user sees on his/her time-line is only a selection of all the posts that his/her friends have posted.
-Remember from Chapter \ref{chap3}, Section \ref{actMech} that each user has a hidden time-line $L_i(t)$.
+Each user has a hidden time-line $L_i(t)$.
 This hidden time-line contains the time-line of the user at the last time he/she was active ($t_{\text{last}}$) plus all the post that his/her friends posted from $t_{\text{last}}$ until the user becomes active again at time $t$.
@@
 Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
-This might make the model a bit more realistic.
+This might make the model more realistic.
 This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.

### 2021-05-05 draft.tex
latest update

@@
 In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.
-
+\newpage
 % NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where nodes can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, eg.
@@
 \end{equation}
-with $\mu = 1.0078272$ and $\sigma = 0.95173779$. The degree distribution of the Last.fm social network, together with the log-normal fit, is shown in Figure \ref{deg_distr_lastfm}.\\
+with $\mu = 1.0078272$ and $\sigma = 0.95173779$. The degree distribution of the Last.fm social network, together with the log-normal fit, is shown in Fig. \ref{deg_distr_lastfm}.\\
 \newline
@@
 with $a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$.
-The degree distribution for the PGP trust network, together with the power-law fits, is shown in Figure \ref{deg_distr_pgp}.
+The degree distribution for the PGP trust network, together with the power-law fits, is shown in Fig.
+\ref{deg_distr_pgp}.
 
@@
   \captionsetup{format=plain}
-  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real world networks used in this thesis.
+  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real-world networks used in this thesis.
 \textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.0078272$, $\sigma = 0.95173779$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$).}
@@
 This chapter is subdivided into four main sections.
-We first discuss the impact of the network structure on the evolution of the opinions and the formation of echo chambers.
+We first discuss the impact of the network structure on the evolution of the prevalence of the opinions and the formation of echo chambers.
 Then, we introduce some kind of stubbornness in the system.
@@
 In this section we will discuss the results related to the network topologies of the different network models that are used in this thesis.
-We will focus on the effect of a high clustering coefficient and a community structure.
-Two different opinion distributions are investigated, namely balanced starting conditions and unbalanced starting conditions.
-For both opinion distributions we investigate the evolution of the prevalence of the opinions with time and the formation of echo chambers.
-Two filtering algorithms will be compared to each other, namely the REC and the PR filtering algorithm, in order to investigate the possible interplay between those filtering algorithms and the network structures.\\
+We will focus on the effect of a high clustering coefficient and community structure.
+Two different opinion distributions are investigated, balanced starting conditions and unbalanced starting conditions.
+For both starting conditions we investigate the evolution of the prevalence of the opinions with time and the formation of echo chambers.
+Two filtering algorithms will be compared to each other, the REC and the PR filtering algorithms, in order to investigate the possible interplay between those filtering algorithms and the network structures.\\
 %\newline
@@
 \newline
-Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
+Four different network models are implemented, each with a total number of nodes $N = 10^3$ and an average degree $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
@@
 \newline
-We start with the results for the evolution of the opinion fractions in case of an initial $50/50$ opinion distribution.
-Perra and Rocha showed that, for the case of balanced initial opinions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the different filtering algorithms \cite{Perra2019}.
-The results for the network model with community structure (SBM, Figure \ref{sbm_50-50}) and the network model with both community structure and high clustering (SBM-WS, Figure \ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithms.
-A community structure is also not able to break the status quo for both of the filtering algorithms.
+We start with the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution.
+Perra and Rocha showed that, for the case of balanced starting conditions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the filtering algorithms \cite{Perra2019}.
 
@@
   \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ distribution.
-Two different network models with a community structure are represented, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Two different network models with community structure are represented, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model the REC and the PR filtering algorithms are compared.
-Each curve is the average of $10^2$ independent simulations and data points are shown, together with their standard deviation $\sigma$, for every 50 time steps to improve visualization.}
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}
 \label{ev_op_50_50}
@@
 \noindent
+The results for the network model with community structure (SBM, Fig.
+\ref{sbm_50-50}) and the network model with both community structure and high clustering (SBM-WS, Fig.
+\ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithms.
+A community structure is also not able to break the status quo for both of the filtering algorithms.
+The initial prevalence of the opinions is maintained over time.\\
+\newline
 \textbf{Echo chambers}\\
 \newline
-Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
+Fig.
+\ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
 The value shown on the $y$-axis, $F_N(\left<P_{\text{B}}^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
@@
 The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
-Error bars are omitted for the sake of an improved visualization.
 
@@
 \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model with $p = 0.01$; The stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; The stochastic Watts-Strogatz model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; The Watts-Strogatz model with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: ER with $p = 0.01$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; WS with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
 The starting conditions are $P_{\text{A}}(0) = 0.5$ and $P_{\text{B}}(0) = 0.5$.
-Each curve is the average of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.
+Error bars are omitted for an improved visualization.}
 \label{echo_50_50}
@@
 \noindent
-In Figure \ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR filtering algorithms.
+In Fig.
+\ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR filtering algorithms.
 The results for both filtering algorithms are qualitatively the same.
@@
 The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a very limited, sign of echo chambers, for both the REC filtering algorithm and the PR filtering algorithm.
-Networks with a high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
+Networks with high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
 Going to slightly higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the Watts-Strogatz model results in significantly larger echo chambers.
 This suggests that the formation of echo chambers is really sensitive to changes in the clustering coefficient.
-The stochastic block model (SBM) with a high modularity, hence, a clear community structure, does show an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller than compared to networks with a high clustering.
+The stochastic block model (SBM) with a high modularity, a clear community structure, does show an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller compared to networks with high clustering.
 The stochastic block model with a low modularity has no, or very limited, echo chambers, similar as for the Erd\H{o}s-R\'{e}nyi
@@
 This is to be expected, since a stochastic block model with a low modularity, and hence a very blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network.
-The stochastic Watts-Strogatz block model (SBM-WS) has both a community structure and a high average clustering.
+The stochastic Watts-Strogatz block model (SBM-WS) has both community structure and high average clustering.
 The formation of echo chambers when using this model is approximately of the same size as the formation of echo chambers in a Watts-Strogatz model with a similar average clustering.
@@
 \newline
-For an initial $50/50$ opinion distribution none of the network models, nor the filtering algorithms could break this status quo (Fig.
+For balanced starting conditions, none of the network models, nor the filtering algorithms could break the status quo (Fig.
 \ref{ev_op_50_50} and related results in reference \cite{Perra2019}).
@@
 \newline
-We study four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$.
+We study four different network models each with a total number of nodes $N = 10^3$ and an average degree $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
@@
 \newline
-Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion A, which is the minority opinion, over time.
+Fig.
+\ref{ev_op_20_80} shows the results for the evolution of the prevalence of opinion A, which is the minority opinion, over time.
 For each network model two filtering algorithms are compared, the REC filtering algorithm and the PR filtering algorithm.
@@
 The REC filtering algorithm maintains the status quo; the prevalence of opinion A remains stable over time for the four different network models.
-When using the PR filtering algorithm, the results are very different.
+When using the PR filtering algorithm, the results are different.
 The prevalence of the minority opinion, $P_\text{A}(t)$, decreases over time.
@@
 Since the PR filtering algorithm orders the posts according to the users' preferences, nodes with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
-Nodes with the minority opinion A, on the other hand, will obviously see posts with opinion A on their time-line when using the PR filtering algorithm.
+Nodes with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
 However, since there are more nodes with opinion B than nodes with opinion A, the nodes with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
@@
 \newline
-Figure \ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
-This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
-The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Figure \ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Figure \ref{ws_20-80_op}).\\
+Fig.
+\ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
+This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
+The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
+\ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Fig.
+\ref{ws_20-80_op}).\\
 \newline
 This result differs from reference \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
-The results that where obtained in reference \cite{Perra2019} are, however, for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$.
-In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$.
+The results that where obtained in reference \cite{Perra2019} are, however, for networks with a total number of nodes $N = 10^5$ and average degree $\left<k\right> \sim 6$.
+In this thesis results are obtained for $N=10^3$ and average degree $\left<k\right> \sim 10$.
 In both cases the maximum length of the actual time-line is set equal to $S=20$.
 The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
-This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the prevalence of opinion A for an initial $20/80$ opinion distribution and using the PR filtering algorithm; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$.
+This can for example be seen in Fig.
+\ref{ev_op_WS_k=6}, which shows the results for the evolution of the prevalence of opinion A for an initial $20/80$ opinion distribution and using the PR filtering algorithm; Fig.
+\ref{ws1000} displays the results for a WS model with $N=10^3,\ \left<k\right>=6$ and $\beta=0.01$; Fig.
+\ref{ws10000} shows the results for a WS model with $N=10^4,\ \left<k\right>=6$ and $\beta=0.01$.
 In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion A reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with a value of $P_\text{A}(500) = 0.027 \pm 0.023$ in Figure \ref{ws1000} and a value of $P_\text{A}(500) = 0.030 \pm 0.018$ in Figure \ref{ws10000}.
+In this figure we can see that the minority opinion A reaches higher final values of $P_0(500)$ compared to the case depicted in Fig.
+\ref{ws_20-80_op}, with a value of $P_\text{A}(500) = 0.027 \pm 0.023$ in Fig.
+\ref{ws1000} and a value of $P_\text{A}(500) = 0.030 \pm 0.018$ in Fig.
+\ref{ws10000}.
 
@@
   \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model the REC and the PR filtering algorithms are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-The bars represent the standard deviation $\sigma$.}
+Bars represent the standard deviation $\sigma$.}
 \label{ev_op_20_80}
@@
 The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
-The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$ and \textbf{(b)} $N=10^4$.
+The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$.
 In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the PR filtering filtering algorithm is used.
@@
 \newline
-Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
-For each network, the REC and PR filtering algorithm are compared to each other.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.
-Error bars are omitted for a better visualization.\\
+Fig.
+\ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
+For each network, the REC and PR filtering algorithm are compared.
+The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.\\
 \newline
@@
 There are no visible effects of having a clear community structure.
-Figure \ref{sbm} shows the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure).
+Fig.
+\ref{sbm} shows the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure).
 There is (almost) no difference in the results for a high and a low modularity.
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model the REC and the PR filtering algorithms are compared.
 The starting conditions are $P_\text{A}(0) = 0.2$ and $P_\text{B}(0) = 0.8$.
-Each curve is the average of $10^2$ independent simulations.}
+Each curve is the average of $10^2$ independent simulations.
+Error bars are omitted for a better visualization.}
 \label{echo_20_80}
@@
 \noindent
-Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}.
+Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (Fig.
+\ref{er_echo_20-80} and Fig.
+\ref{sbm}).
 When using the REC filtering algorithm, the formation of echo chambers around the majority opinion B is hampered.
@@
 There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion A.
-From figure \ref{ev_op_20_80} it was already concluded that the REC filtering algorithm does not break the prevalence of the opinions.
+From figure \ref{ev_op_20_80} it was concluded that the REC filtering algorithm does not break the prevalence of the opinions.
 The opinion distribution is however slightly shifted from the case where the minority opinion A is more or less randomly distributed over the network to the case where the minority opinion A groups together in opinion clusters.
@@
 \newline
-The PR filtering algorithm shows a completely different tendency.
+The PR filtering algorithm shows a different tendency.
 There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion B.
@@
 This reduced visibility of the minority opinion A in these neighborhoods drives them eventually to adopt opinion B, with an increase in the number of echo chambers around the majority opinion B as a consequence.
-Remember from Figure \ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion.
+Remember from Fig.
+\ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion.
 Hence, the increase in the formation of echo chambers around the majority opinion B is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
@@
 Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
-For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}).
-An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}.
+For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in Fig.
+\ref{ws} and Fig.
+\ref{sbm-ws}).
+An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Fig.
+\ref{ws} and in the case of a SBM-WS network in Fig.
+\ref{sbm-ws}.
 This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity.
-In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero are not actually reaching infinity, but are cut off around values equal to 100/180.
+In Fig.
+\ref{ws} and Fig.
+\ref{sbm-ws}, the spikes around zero are not actually reaching infinity, but are cut off around values equal to 100/180.
 These large spikes, indicating a large increase of echo chambers around the minority opinion, do not necessary mean that we have a huge absolute number of echo chambers around the minority opinion A at $t=500$.
@@
 This enhancing of the formation of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
-Remember from Figure \ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for the high clustering networks.
+Remember from Fig.
+\ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for the high clustering networks.
 Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
@@
 This means that the average fraction of nodes with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
-In case of the PR filtering algorithm this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+In case of the PR filtering algorithm this value is even slightly higher than $1$, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
 
@@
 
-\subsection{Relation between formation of echo chambers and community structure}\label{echoVSmod}
-
-In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
+\subsection{Formation of echo chambers versus community structure}\label{echoVSmod}
+
+In Fig.
+\ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
 The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion A and the normalized fraction of nodes with all neighbors carrying opinion B (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
@@
 The data points represent the value of the echo chamber size in networks with a different modularity.
-The initial opinion distribution was always $50/50$ an was, initially, randomly distributed over the network.
+The initial opinion distribution was $50/50$ and was, initially, randomly distributed over the network.
 \\
@@
 \newline
-Table \ref{tab1} displays the network models that are used in Figure \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
+Table \ref{tab1} displays the network models that are used in Fig. \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
 
@@
 \end{tabular}
-\caption{Network models used in Figure \ref{echo_vs_mod}.}
+\caption{Network models used in Fig. \ref{echo_vs_mod}.}
 \label{tab1}
 \end{table}
-
+\newpage
 \begin{figure}[H]
@@
 The initial opinion distribution is $50/50$.
-The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10.
+The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and average degree around 10.
 The other four data points are the results for SBM's with 10 communities of size 100 ($N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and average degree $\left<k\right> \sim 10$.
 Each data point is the average of $10^2$ independent simulations.
-The bars represent the standard error $SE$.}
+Bars represent the standard error $SE$.}
 \label{echo_vs_mod}
@@
 For most values of the modularity, the REC filtering algorithm produces almost no echo chambers.
-If the modularity becomes high enough we see a small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than two.
+If the modularity becomes high enough we see a small increase in the formation of echo chambers, however even for $Q = 0.908 \pm 0.005$, the size of the echo chambers remains smaller than two.
 This means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
@@
 \newline
-We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness.
-Each of the network models have a total number of nodes $N=10^3$ and an average degree $\left<k\right> \sim 10$.
+We will implement four different network models in order to investigate the possible interplay between network structure and stubbornness.
+Each of the network models have a total number of nodes $N=10^3$ and average degree $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-The interplay between stubbornness and clustering is studies with the Watts-Strogatz model (WS) with mean degree $K=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
+The interplay between stubbornness and clustering is studies with the Watts-Strogatz model (WS) with mean degree $\left<k\right>=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
 The effect of community structure is investigated with the stochastic block model (SBM) with 10 communities of size 100.
@@
 \newline
-For each network model, we compare the REF, REC and PR filtering algorithms to each other to analyze the possible interplay between stubbornness and filtering algorithm.
+For each network model, we compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm.
 
@@
 \newline
-For both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers.
-This is done this for two different starting conditions: a balanced starting conditions and unbalanced starting conditions.
-\\
-\newline
-The opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
+In both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers.
+This is done for two different starting conditions: balanced starting conditions and unbalanced starting conditions.
+\\
+\newline
+The opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
 \newline
@@
 We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.
-Then we will discuss the formation of echo chambers and the evolution of the opinions for an initial $20/80$ opinion distribution.
+Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for an initial $20/80$ opinion distribution.
 
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the PR filtering algorithm and for an initial $50/50$ opinion distribution.
-The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.
+Fig.
+\ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the PR filtering algorithm for balanced starting conditions.
+The $y$-axis represents the formation of echo chambers in the same way as was explained for Fig.
+\ref{echo_vs_mod} in Subsection \ref{echoVSmod}.
 The dotted black line represents an echo chamber size equal to one.
@@
 The stochastic block model with a low modularity behaves similar as the Erd\H{o}s-R\'{e}nyi model (no formation of echo chambers).
-The data points for the SBM with a low modularity are not visible in Figure \ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
+The data points for the SBM with a low modularity are not visible in Fig.
+\ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
 \newline
 The other three networks (WS, SBM-WS and SBM with a high modularity) all show qualitatively the same behavior.
-If an increasing fraction of completely stubborn actors is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn actors.
-This can be explained as follows.
-Having a fraction of completely stubborn actors, reduces the effective number of neighbors that can change their opinion.
-This in turn leads to the hampering of the formation of echo chambers.\\
+If an increasing fraction of completely stubborn actors is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn actors.\\
 \newline
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc \right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc \right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the REF and REC filtering algorithms.
+Fig.
+\ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the REF and REC filtering algorithms.
 The results are obtained for an initial $50/50$ opinion distribution.
-The results for the REF and REC filtering algorithms behave qualitatively the same, but are different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}).
+The results for the REF and REC filtering algorithms behave qualitatively the same, but are different from those obtained with the PR filtering algorithm (Fig.
+\ref{echo_vs_frac_complStubb}).
 Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors, we first observe an increase in the formation of echo chambers, followed by a decrease.
@@
 \newline
-These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn actors enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
-% explain potential reasons (eg echo chambers around stubborn nodes --> keep on pushing their opinion to the neighbors, not the case for PR) + check those + need for futher research.
+These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small) fraction of stubborn actors enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
 If fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
 This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that will not change opinions.
-This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.\\
+This in turn increases the chance of having stubborn neighbors, each with different opinions, who cannot change opinions, which disables the formation of echo chambers.\\
 \newline
@@
 However, the two nodes with opinion A still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion A, and hence reduces the visibility of the posts of the stubborn friend with opinion B.
-The system will now not form an echo chamber, since the stubborn node with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion A to opinion B.
+The system will not form an echo chamber, since the stubborn node with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion A to opinion B.
 The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
@@
 This would explain the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
-If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease again which can also be easily seen with the help of the triangle example.
+If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease which can also be easily seen with the help of the triangle example.
 If the fraction of stubborn actors becomes too high, we might end up with the case where both the node with opinion B and one of the nodes with opinion A are stubborn.
@@
 \newline
-So far the results for the formation of echo chambers in the system for an initial $50/50$ opinion distribution and introducing a fraction of completely stubborn actors.
+So far the results for the formation of echo chambers in the system for balanced starting conditions and introducing a fraction of completely stubborn actors.
 But what about the evolution of the prevalence of the opinions over time?
@@
 \newline
-Lets now study the situation of unbalanced starting conditions, combined with a fraction of stubborn actors.
+Lets now study the situation for unbalanced starting conditions, combined with a fraction of stubborn actors.
 The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and the initial prevalence of opinion B is $P_\text{B}(0) = 0.8$.
@@
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$.
+Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \noindent
-Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion).
+Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion).
 The PR and the REC filtering algorithms are compared.
@@
 Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both.
-In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity.
+In case of a $20/80$ distribution we have a skewed echo chamber distribution, as can be seen in Fig.
+\ref{echo_20_80}, with normalized values going to infinity.
 We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
@@
 \newline
-When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
-This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}.
-For an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}.
-This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was explained for Figure \ref{echo_20_80}.
+When using the PR filtering algorithm (Fig.
+\ref{PR_frac_compl_stubb_20-80}) we observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
+This behavior is similar to the case of balanced starting conditions in combination with the PR filtering algorithm (Fig.
+\ref{echo_vs_frac_complStubb}).
+For unbalanced starting conditions, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for fracRes $=0$), as can be observed in Fig.
+\ref{echo_20_80}.
+This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was explained for Fig.
+\ref{echo_20_80}.
 \\
@@
 It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of unbalanced starting conditions.
-The reason is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}.
+The reason is that this average neglects the skewed echo chamber distributions as depicted in Fig.
+\ref{echo_20_80}.
 When comparing the different network models, these skewed distributions should not be overlooked.
@@
 % explain reason behavior REC!!
-The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, are different than those for the PR filtering algorithm.
+The results for the REC filtering algorithm (Fig.
+\ref{REC_frac_compl_stubb_20-80}) are different than those for the PR filtering algorithm.
 The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease.
-This result is similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}.
+This result is similar to the results for balanced starting conditions in combination with the REC filtering algorithm (Fig.
+\ref{REC_frac_compl_stubb}).
 The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we see a decrease in the formation of echo chambers.
 The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show no significant formation of echo chambers.
-For an increasing fraction of the completely stubborn actors, the echo chambers eventually converge to one (for all the network models), indicating a constant amount of echo chambers over time.\\
-\newline
-We can conclude that the formation of echo chambers versus the fraction of completely stubborn actors, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn actors for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm).
+For an increasing fraction of the completely stubborn actors, the echo chambers eventually converge to one (for all the network models), indicating a constant amount of echo chambers over time (represented by the black dotted line).\\
+\newline
+We can conclude that the formation of echo chambers versus the fraction of completely stubborn actors, for unbalanced starting conditions, qualitatively resembles the results for balanced starting conditions (decrease in formation of echo chambers with an increasing fraction of stubborn actors for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm).
 
@@
   \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 Lets now discuss the evolution of the prevalence of the opinions.
-In Figure \ref{ev_op_20_80}, we saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution.
+In Fig.
+\ref{ev_op_20_80}, we saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of unbalanced starting conditions.
 The PR filtering algorithm, on the other hand, resulted in a decrease of the prevalence of the minority opinion.
 Are these results affected by introducing stubborn actors?
-As for the balanced initial conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
+As for the balanced starting conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
 The status quo is maintained over time.
-Introducing stubborn actors combined with the PR filtering algorithm, on the other hand, does have an impact in case of unbalanced initial conditions.\\
-\newline
-Figure \ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time in case of unbalanced starting conditions and using the PR filtering algorithm.
+Introducing stubborn actors combined with the PR filtering algorithm, on the other hand, does have an impact in case of unbalanced starting conditions.\\
+\newline
+Fig.
+\ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time in case of unbalanced starting conditions and using the PR filtering algorithm.
 We compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS).
@@
 If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
-If we introduce a fraction of stubborn actors in the system, then we will end up with stubborn actors that carry the minority opinion.
+If we introduce a fraction of stubborn actors in the system, we end up with stubborn actors that carry the minority opinion.
 These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
@@
 \newline
-Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod}.
+Fig.
+\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The interpretation of the $y$-axis is again as explained in Subsection \ref{echoVSmod}.
 The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig.
@@
 \noindent
-In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
-Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
+In Fig.
+\ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
+Fig.
+\ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Fig.
+\ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
 \newpage
@@
 \newline
-What about the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution and making all the nodes partly stubborn?
+What about the results for the evolution of the prevalence of the opinions in case of balanced starting conditions and making all the nodes partly stubborn?
 Making all the nodes partly stubborn does not disturb/change the evolution of the opinions.
-The balanced initial conditions remain balanced over time, as was the case for no stubbornness in the system.
+The balanced starting conditions remain balanced over time, as was the case for no stubbornness in the system.
 This result is valid for all the network models (ER, WS, SBM and SBM-WS) and for both the REC and PR filtering algorithm.\\
@@
 \newline
-Figure \ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
+Fig.
+\ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
 The standard error $SE$ is small and not visible in the figure.
-Figure \ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Figure \ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
-The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
+Fig.
+\ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Fig.
+\ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
+The $y$-axis is interpreted in the same way as explained for Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
 \newline
@@
 If $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
-This behavior is slightly different than in case of an initial $50/50$ distribution, combined with the PR filtering algorithm (Fig.
+This behavior is slightly different than in case of balanced starting conditions, combined with the PR filtering algorithm (Fig.
 \ref{echo_vs_all_frac_stubb_PR}).
 In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if $r>0.1$).
-This difference is probably a consequence of the unbalanced initial conditions, which seems to make the system less sensitive to the stubbornness parameter $r$ of all the nodes.\\
-\newline
-As was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of an initial $20/80$ opinion distribution, combined with the PR filtering algorithm is non-zero.
-These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models to each other in case of an initial $20/80$ opinion distribution.
+This difference is probably a consequence of the unbalanced starting conditions, which seems to make the system less sensitive to the stubbornness parameter $r$.\\
+\newline
+As was explained for Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of unbalanced starting conditions, combined with the PR filtering algorithm is non-zero.
+These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models in case of unbalanced starting conditions.
 In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
-Figure \ref{echo_20_80}), as was explained for Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80}.
+Fig.
+\ref{echo_20_80}), as was explained for Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}.
 
@@
 The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
-The increase seems less steep than the increase for an initial $50/50$ opinion distribution, which we can observe in Figure \ref{REC_all_frac_stubb}.
-This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for an initial $20/80$ distribution than for an initial $50/50$ distribution.\\
+The increase seems less steep than the increase for balanced starting conditions, which we can observe in Fig.
+\ref{REC_all_frac_stubb}.
+This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for unbalanced starting conditions than for balanced starting conditions.\\
 \newline
@@
 This increase can even be observed for the low modularity case, albeit less strong.
-The Erd\H{o}s-R\'{e}nyi network wiggles a bit around a formation of echo chambers equal to one.
+The Erd\H{o}s-R\'{e}nyi network wiggles around a formation of echo chambers equal to one.
 \newpage
@@
   \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 The prevalence of opinion A versus time in case of the REC filtering algorithm is not affected by making all the nodes more and more stubborn.
-The prevalence of the opinions is maintained an the status quo is not broken.
+The prevalence of the opinions is maintained and the status quo is not broken.
 The PR filtering algorithm gives us another story.
-The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Figure \ref{ev_op_20_80_all_frac_stubb_PR}.
-If the nodes are not stubborn, the PR filtering algorithm reduces the minority opinion over time.
+The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Fig.
+\ref{ev_op_20_80_all_frac_stubb_PR}.
+If the nodes are not stubborn, the PR filtering algorithm reduces the prevalence of the minority opinion over time.
 Making the nodes more and more stubborn reduces this tendency.
@@
 \ref{ev_op_20_80_stubb_PR}).
-The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the minority opinion compared to making a fraction of the nodes completely stubborn.
-For $r=0.4$, the decrease of the minority opinion is almost not hampered, it is merely slowed down.
-All the nodes need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the minority opinion.
-Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$).
+The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the prevalence of the minority opinion compared to making a fraction of the nodes completely stubborn.
+For $r=0.4$, the decrease of the prevalence of the minority opinion is almost not hampered, it is merely slowed down.
+All the nodes need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the prevalence of the minority opinion.
+Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$, respectively).
 
@@
 \newline
-Figure \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
+Fig.
+\ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
 In this figure, the PR filtering algorithm is used.
 Data points are shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.
-The $y$-axis is interpreted in the same way as explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}.\\
+The $y$-axis is interpreted in the same way as explained for Fig.
+\ref{echo_vs_mod} in Subsection \ref{echoVSmod}.\\
 \newline
@@
 \noindent
-Figure \ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Figure \ref{REF_thres}) and REC (Figure \ref{REC_thres}) filtering algorithms.
+Fig.
+\ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Fig.
+\ref{REF_thres}) and REC (Fig.
+\ref{REC_thres}) filtering algorithms.
 The initial opinion distribution is $50/50$.
@@
 The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
-This can be seen in Figure \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
+This can be seen in Fig.
+\ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
 Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REC/REF filtering algorithms.
@@
 If $T \geqslant 0.5$, it starts to have an impact.
-However, in Figure \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
+However, in Fig.
+\ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
 At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
@@
 \newline
-Figure \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
+Fig.
+\ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
 The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
@@
 This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
-Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Figure \ref{sbm_majT}.
+Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Fig.
+\ref{sbm_majT}.
 For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
@@
 The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
 For each model the REC and the PR filtering algorithms are compared.}
@@
 
-\section{Random versus community distributed}
+\section{Random versus community distributed opinions}
 
 This section is devoted to results regarding the way the opinions are distributed over the network.
-In all the previous results, the two opinions were, initially, always randomly distributed across the network.
+In all the previous results, the two opinions were, initially, randomly distributed across the network.
 It might however be more realistic if people with the same opinion are more or less grouped together initially.
@@
 The total opinion distribution over the entire network is no longer $50/50$, but instead opinion A becomes the majority opinion (hence, opinion B becomes the minority opinion).
-The higher the fraction of single opinion A communities, the bigger the difference between this majority and minority opinion becomes.
-The results will always be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
+The higher the fraction of single opinion A communities, the bigger the difference between this majority and minority opinion.
+The results will be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
 Thus, eg.
@@
 \newline
-We will a network with a high modularity, representing the case of more or less isolated communities, and a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
+We will study a network with a high modularity, representing the case of more or less isolated communities, and a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
 More specifically, the two stochastic block models that are implemented are: (i) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ ($Q = 0.908 \pm 0.005$); (ii) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ ($Q = 0.220 \pm 0.007$).
@@
 \newline
-Figure \ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
+Fig.
+\ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
 A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the other communities get an initial $50/50$ opinion distribution.
-The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in case of a fraction of $0.1$ single opinion A communities).
+The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in this case).
 The figure shows the results for both the high and the low modularity case.
-For each modularity the REC and PR filtering algorithms are compared to each other.\\
-\newline
-In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
-The prevalence of the majority opinion A remains unchanged over time.
-This result is the same for both the high and low modularity case, see Figures \ref{high_mod_rec} and \ref{low_mod_rec}.\\
-\newline
-The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced starting conditions and if the opinions are randomly distributed.
-This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
-Figure \ref{ev_op_20_80}.
-If a fraction equal to $0.1$ of the communities is given one opinion, namely opinion A, while the remaining communities get a $50/50$ opinion distribution, we see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
-This makes sense, since the single opinion communities are more or less isolated from the other communities.
-This has as consequence that the other (mixed opinion) communities live under the assumption that the opinion distribution is $50/50$.
-They do not really feel that opinion A is a majority opinion, since they are isolated from these single opinion A communities.
-In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
-This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present.
-The community distributed case behaves equivalent to the random distributed case.
-In the low modularity network, the community structure is not clearly present, but becomes blended.
-The different communities talk a lot to each other and there is a lot of communication between them.
-As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over the communities.
+For each modularity the REC and PR filtering algorithms are compared.
 
@@
 \end{figure}
+\newpage
 \noindent
-The results found in Figure \ref{commOp0_01_other_50-50} for $10 \%$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
-This can be seen in Figure \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
-The error bars in this figure are omitted for the sake of an improved visualization.
-The community distributed cases are always compared to the corresponding randomly distributed cases.
+In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
+The prevalence of the majority opinion A remains unchanged over time.
+This result is the same for both the high and low modularity case, see Fig.
+\ref{high_mod_rec} and Fig.
+\ref{low_mod_rec}.\\
+\newline
+The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced starting conditions and if the opinions are randomly distributed.
+This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
+Fig.
+\ref{ev_op_20_80}.
+If a fraction equal to $0.1$ of the communities is given one opinion, opinion A, while the remaining communities get a $50/50$ opinion distribution, we see from Fig.
+\ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
+This makes sense, since the single opinion communities are more or less isolated from the mixed opinion communities.
+This has as consequence that the mixed opinion communities live under the assumption that the opinion distribution is $50/50$.
+They do not really feel that opinion A is a majority opinion, since they are isolated from these single opinion A communities.
+In case of the low modularity network, Fig.
+\ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
+This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present.
+The community distributed case behaves equivalent to the random distributed case.
+In the low modularity network, the community structure is not clearly present, but becomes blended.
+The different communities talk a lot to each other and there is a lot of communication between them.
+As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over the communities.
+\\
+\newline
+The results found in Fig.
+\ref{commOp0_01_other_50-50} for $10 \%$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
+This can be seen in Fig.
+\ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
+The community distributed cases are compared to the corresponding randomly distributed cases.
 Results are shown for both a high modularity network and a low modularity network.
-Figure \ref{rec_comm} shows the results for the REC filtering algorithm.
-Figure \ref{pr_comm} shows those for the PR filtering algorithm.\\
+Fig.
+\ref{rec_comm} shows the results for the REC filtering algorithm.
+Fig.
+\ref{pr_comm} shows those for the PR filtering algorithm.\\
 \newline
@@
 In case of the REC filtering algorithm (Fig.
-\ref{rec_comm}) the prevalence of opinion A remains the same after the time evolution for a fixed fraction of single opinion A communities.
+\ref{rec_comm}) the prevalence of opinion A remains the same before and after the time evolution for a fixed fraction of single opinion A communities.
 Giving a fraction of the communities a single opinion, opinion A, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
@@
 \newline
-The PR filtering algorithm (Figure \ref{pr_comm}), on the other hand, shows an increase in the prevalence of opinion A over time for a fixed fraction of single opinion A communities.
+The PR filtering algorithm (Fig.
+\ref{pr_comm}) shows an increase in the prevalence of opinion A over time for a fixed fraction of single opinion A communities.
 Making a fraction of the communities single opinion A communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
 The low modularity network does not exhibit this hampering.
-The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was explained for Figure \ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
+The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was explained for Fig.
+\ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
 
@@
 Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
-\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
-Each curve is the average of $10^2$ independent simulations.}
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Each curve is the average of $10^2$ independent simulations.
+Error bars are omitted for an improved visualization.
+}
 \label{fracOp0_vs_commOp0_other_50-50}
@@
 \newline
-Figure \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
+Fig.
+\ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
 The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
@@
 The formation of echo chambers in both the high and low modularity network are investigated.
-The REC and PR filtering algorithms are compared to each other.
-Error bars are again omitted for an improved visualization.\\
-\newline
-The results for the REC filtering algorithm are depicted in Figure \ref{rec_comm_echo}.
+The REC and PR filtering algorithms are compared.
+\\
+\newline
+The results for the REC filtering algorithm are depicted in Fig.
+\ref{rec_comm_echo}.
 The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (bump in the blue line).
@@
 This increase in the formation of echo chambers around the minority opinion B seems however hampered compared to the randomly distributed case.
-Remember that the increase in the number of echo chambers around the minority opinion for the SBM in combination with the REC filtering algorithm for unbalanced starting conditions that are randomly distributed was observed in Section \ref{netw_struc}, Subsection \ref{20-80} (Fig?
+Remember that the increase in the number of echo chambers around the minority opinion for the SBM in combination with the REC filtering algorithm for unbalanced starting conditions that are randomly distributed was observed in Section \ref{netw_struc}, Subsection \ref{20-80} (Fig.
 \ref{sbm}).
@@
 \newline    % make link to opinion fraction 
-The results for the PR filtering algorithm are depicted in Figure \ref{pr_comm_echo}.
-They are different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Figure \ref{rec_comm_echo}.\\
+The results for the PR filtering algorithm are depicted in Fig.
+\ref{pr_comm_echo}.
+They are different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Fig.
+\ref{rec_comm_echo}.\\
 \newline
 Lets first focus on the high modularity network.
-The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (these results were obtained for an initial $20/80$ opinion distribution).
-We can observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion B.
-Remember from Figure \ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion A.
+The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Fig.
+\ref{sbm} (these results were obtained for an initial $20/80$ opinion distribution).
+We observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion B.
+Remember from Fig.
+\ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion A.
 If we, however, make $10 \%$ of the communities single opinion A communities while giving the remaining ones a $50/50$ opinion distribution, we see that this increase in the formation of echo chambers around the majority opinion A is strongly hampered.
@@
 This shows that the observed hampering of the formation of echo chambers around the majority opinion A in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers.
-Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed case versus the random distributed case.
+Remember from Fig.
+\ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed case versus the random distributed case.
 Beside this hampering of echo chambers around the majority opinion A, the community distributed case also shows an increase of echo chambers around the minority opinion B (which is not visible for the randomly distributed case).
-This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
-\newline
+This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.
+% share data??
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_REC_8x7.png}
+    \label{rec_comm_echo}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_PR_8x7.png}
+    \label{pr_comm_echo}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Each curve is the average of $10^2$ independent simulations.
+Error bars are omitted for an improved visualization.}
+\label{echo_01_commOp0_other_50-50}
+\end{figure}
+\noindent
 The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases, as was observed for the high modularity network.
 Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion B.
-From Figure \ref{low_mod_pr}, it was made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of the minority opinion (when using the PR filtering algorithm).
+From Fig.
+\ref{low_mod_pr}, it was made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of the minority opinion (when using the PR filtering algorithm).
 This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry the minority opinion.
@@
 + make table with data??
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_REC_8x7.png}
-    \label{rec_comm_echo}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_PR_8x7.png}
-    \label{pr_comm_echo}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
-$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
-$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
-\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
-Each curve is the average of $10^2$ independent simulations.}
-\label{echo_01_commOp0_other_50-50}
-\end{figure}
-\noindent
-Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, $50 \%$ of the communities is given opinion A, while the others get a $50/50$ distribution.
+\\
+\newline
+Fig.
+\ref{echo_05_commOp0_other_50-50} shows a similar plot as Fig.
+\ref{echo_01_commOp0_other_50-50}.
+$50 \%$ of the communities is given opinion A, while the others get a $50/50$ distribution.
 The results look more or less the same as for the case where $10 \%$ of the communities is given opinion A (Fig.
@@
 \newline
-The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
+The REC filtering algorithm, Fig.
+\ref{rec_comm05_echo}, shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
 A difference with the case where $10 \%$ of the communities are single opinion A communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion B.
@@
 \newline
-The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Figure \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
+The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Fig.
+\ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
 They show a small increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
 \newline
-Results for the PR filtering algorithm also strongly resemble those depicted in Figure \ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion A communities.
-The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Figure \ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}).
+Results for the PR filtering algorithm also strongly resemble those depicted in Fig.
+\ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion A communities.
+The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Fig.
+\ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}).
 They show a strong increase in the formation of echo chambers around the majority opinion A.\\
 \newline
-The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion A, which was also observed in Figure \ref{pr_comm_echo}.\\
+The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion A, which was also observed in Fig. \ref{pr_comm_echo}.\\
 \newline
 The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion A.
-This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (Fig.
+This was also observed in Fig.
+\ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (Fig.
 \ref{high_mod_pr} and Fig.
 \ref{pr_comm}).
-The increase in the number of echo chambers around the minority opinion A that was visible in Figure \ref{pr_comm_echo} is no longer present.
+The increase in the number of echo chambers around the minority opinion A that was visible in Fig.
+\ref{pr_comm_echo} is no longer present.
 
@@
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
-\textbf{(a)} REC filtering algorithms; \textbf{(b)} PR filtering algorithm.
-Each curve is the average of $10^2$ independent simulations.}
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Each curve is the average of $10^2$ independent simulations.
+Error bars are omitted for an improved visualization.}
 \label{echo_05_commOp0_other_50-50}
@@
 
+\newpage
 \section{Real-world networks}
+
+The results of the two real-world networks: (i) the Last.fm on-line social network; (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
+For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or community structure.
+We will mainly focus on the formation of echo chambers.
+We will only study balanced initial conditions and we will introduce a fraction of completely stubborn actors.
+For both real-world networks, the REC and PR filtering algorithms will be compared.
+
+\subsection{The Last.fm on-line social network}
+
+The formation of echo chambers versus the fraction of completely stubborn actors, fracRes, in the Last.fm on-line social network is studied.
+The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
+The results for the PR and REC filtering algorithm are shown in Fig.
+\ref{echo_vs_fracRes_real}.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_8x7.png}
+    \label{pr_real}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_8x7.png}
+    \label{rec_real}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
+The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+The Last.fm on-line social network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM-WS with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$; (iii) SBM with 53 communities of size 151 with $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$.
+\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Each data point is the average of $50$ independent simulations.
+Bars represent the standard deviation $\sigma$.}
+\label{echo_vs_fracRes_real}
+\end{figure}
+
+\subsection{The PGP trust network}
 
@@
 The main focus was put on the difference between networks with a high clustering coefficient and no community structure (WS) compared to networks with a clear community structure and a low clustering coefficient (SBM).
-It was found that a clear community structure does enhance the formation of echo chambers (mainly when using the PR filtering algorithm).
+It was found that a clear community structure does enhance the formation of echo chambers (mainly when using the PR filtering algorithm) compared to a random network.
 This enhancement is however not as strong as for networks with a high clustering coefficient.
@@
 \newline
-It was also found that the way the opinions are initially distributed over the communities in a network with a community structure (SBM) does make a difference.
-We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ distribution.
+It was found that the way the opinions are initially distributed over the communities in a network with a community structure (SBM) does make a difference.
+We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
 This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.
@@
 \newline
-The REC filtering algorithm did not show any difference in the evolution of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.
+The REC filtering algorithm did not show any difference in the evolution of the prevalence of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.
 For the PR filtering algorithm it was found that the high modularity network hampers the increase in the majority opinion in the community distributed scenario compared to the randomly distributed scenario.
 For the low modularity network, such a discrepancy was not found.\\
+\newline
+Furthermore, it was found that, when using the REC filtering algorithm, the system evolves towards a more mixed opinion distribution in the community distributed scenario.
+This was found for both the high and low modularity network.
+If we make a fraction of the communities single opinion A communities, while giving the remaining communities a $50/50$ opinion distribution, we initially have a large number of echo chambers around opinion A.
+Over time, these opinion A echo chambers decrease; even if there is only a limited amount of communication between the communities, it was found that the single opinion A communities become more and more mixed.
+The total prevalence of opinion A is unchanged when using the REC filtering algorithm, so the initial $50/50$ communities see a small increase in the prevalence of opinion A.
+The randomly distributed scenario does not evolve towards a mixed opinion distribution, instead there is a small increase in the formation of echo chambers around the minority opinion B (for both the high and low modularity network).\\
+\newline
+For the PR filtering algorithm, the randomly distributed scenario resulted in an increase in the formation of echo chambers around the majority opinion A.
+This was observed for both the low and high modularity networks.
+The community distributed scenario for the low modularity network hampers this increase.
+It was found that this is more than just a result of the increased number of initial opinion A echo chambers.
+For the high modularity network, the community distributed scenario showed an even larger hampering of the formation of echo chambers around the majority opinion A.
+This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase in the majority opinion A for the community distributed scenario in case of the high modularity network combined with the PR filtering algorithm.
 % talk about echo chambers??

### 2021-05-06 draft.tex
latest update: worked on discussion

@@
 \textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.0078272$, $\sigma = 0.95173779$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$).}
-\label{ev_op_50_50}
+\label{deg_distr_real_network}
 \end{figure}
-
-% Discuss actual networks we used!! + reference them??
 
@@
 The results for the PR and REC filtering algorithm are shown in Fig.
-\ref{echo_vs_fracRes_real}.
+\ref{echo_vs_fracRes_real}.\\
+\newline
+The PR filtering algorithm shows a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
+This is valid for both the real-world network and the network models.
+The stochastic Watts-Strogatz block model has a similar clustering and community structure as the Last.fm network.
+The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
+This suggests that it is not only the clustering and/or the community structure that drives the formation of echo chambers.
+Other topologies must play a role as well.
+The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous.
+It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.
 
@@
 This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase in the majority opinion A for the community distributed scenario in case of the high modularity network combined with the PR filtering algorithm.
-% talk about echo chambers??
+\\
+% summarize real-world network results
+\newline
+The formation of echo chambers, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature \cite{Cinelli2020}\cite{Nikolov2015}.
+The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.\\
+\newline
+The hampering in the increase of the majority opinion in case of unbalanced starting conditions combined with the introduction of stubborn actors for the PR filtering algorithm agrees partly with related research on continuous opinion dynamics models combined with stubborn actors \cite{Como2013}.
+In this research, Acemo\u{g}lu et al.
+showed that introducing stubborn actors (with different opinions) in an inhomogeneous stochastic gossip process of continuous opinion dynamics leads to a situation in which consensus can no longer be reached \cite{Como2013}.
+This result is also found in the majority opinion dynamics model and in the voter model combined with stubborn actors (equally distributed over the two opinions) \cite{Galam2007}\cite{Mobilia2007}\cite{Yildiz2013}.
+In our case, the possible consensus around the majority opinion, that could have been reached when combining the PR filtering algorithm with unbalanced starting conditions, can no longer be obtained when introducing stubborn actors.
 

### 2021-05-07 draft.tex
latest update

@@
 With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
-An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}.
+An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}\cite{Pawel}.
 This interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
@@
 Another example of introducing some kind of persistence towards the own opinion of the nodes is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
+The model introduced by Friedkin and Johnsen in which each agent adheres to its initial opinion to a certain degree, see Section \ref{classMod}, can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
 
@@
 If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.\\
-\newline
+\newpage
+\noindent
 The resistance parameter $r_i$ can be different or the same for all nodes.
@@
 Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
-This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results.
+This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Pawel}.
 \\
@@
 There is however a need to compare data from (computational) models to experimental, real-world data.
-A need for critical thinking and careful analyzing of results can also not be denied.
+A need for critical thinking and careful analyzing of results can also not be denied \cite{Pawel}.
 % put this section here?? Reference??
@@
 The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
-This suggests that it is not only the clustering and/or the community structure that drives the formation of echo chambers.
+This suggests that it is not only clustering and/or community structure that drives the formation of echo chambers.
 Other topologies must play a role as well.
-The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous.
+The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution.
 It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.
+This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.
+% talk about possible effect of powerlaw community sizes + figure??
+The results for the Watts-Strogatz model are similar to those for the stochastic Watts-Strogatz block model.
+The formation of echo chambers seems to be slightly smaller in the WS than in the SBM-WS network.
+It may be tempting to attribute this to the lack of community structure in the WS model.
+It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers in the WS network than in the SBM-WS network.
+Furthermore, it is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity.
+It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of high clustering.
+% talk about coincidence??
 
@@
 \end{figure}
+\noindent
+The REC filtering algorithm enforces an increase in the formation of echo chambers when a small fraction of stubborn actors is introduced.
+If this fraction is further increased, the formation of echo chambers starts to decrease again.
+The increase is the strongest for the high clustering network models (WS, SBM-WS).
+Even though the Last.fm network has a similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
+As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network.
+% talk about SBM
 
@@
 The main focus was put on the difference between networks with a high clustering coefficient and no community structure (WS) compared to networks with a clear community structure and a low clustering coefficient (SBM).
-It was found that a clear community structure does enhance the formation of echo chambers (mainly when using the PR filtering algorithm) compared to a random network.
+It was found that a clear community structure enhances the formation of echo chambers (mainly when using the PR filtering algorithm) compared to a random network.
 This enhancement is however not as strong as for networks with a high clustering coefficient.
@@
 This is mainly seen in the high clustering networks, however a small bump can also be observed in the high modularity network.
-This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.\\
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.
+It was also found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
 % talk about evolution of opinions??
 \newline
-It was found that the way the opinions are initially distributed over the communities in a network with a community structure (SBM) does make a difference.
+Finally, it was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) does make a difference.
 We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
@@
 \newline
-The formation of echo chambers, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature \cite{Cinelli2020}\cite{Nikolov2015}.
-The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.\\
+The formation of echo chambers in social networks/social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
+The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.
+Banisch and Olbrich showed that in a social feedback model polarization is increased in high modularity networks compared to an Erd\H{o}s-R\'{e}nyi network \cite{Banisch2018}.
+This agrees with our findings that a clear community structure enhances the formation of echo chambers.\\
 \newline
@@
 This result is also found in the majority opinion dynamics model and in the voter model combined with stubborn actors (equally distributed over the two opinions) \cite{Galam2007}\cite{Mobilia2007}\cite{Yildiz2013}.
-In our case, the possible consensus around the majority opinion, that could have been reached when combining the PR filtering algorithm with unbalanced starting conditions, can no longer be obtained when introducing stubborn actors.
+In our case, the possible consensus around the majority opinion, that could have been reached when combining the PR filtering algorithm with unbalanced starting conditions, can no longer be obtained when introducing stubborn actors.\\
+\newline
+Wang et al.
+\cite{Wang2020} proposed an agent based continuous opinion dynamics model in which they investigated the formation of echo chambers.
+They found that stubbornness of individuals protects them from the
+influence of centralized sources of influence; ``A purely stubborn or conformist
+population may both be less susceptible to the emergence of
+two dominant and opposing echo chambers'' (Wang et al., p.10, \cite{Wang2020}).
+This agrees with our result that the PR filtering algorithm, combined with stubborn actors, hinders the formation of echo chambers.
 

### 2021-05-08 draft.tex
latest version

@@
 \selectlanguage{english}
-\newgeometry{top=20mm, bottom=25mm}
 \tableofcontents
 \addcontentsline{toc}{chapter}{Table of Contents}
+\newgeometry{top=20mm, bottom=25mm}
 \listoffigures
@@
 This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.
-% talk about possible effect of powerlaw community sizes + figure??
+Another difference between the Last.fm and the SBM and SBM-WS network is the community structure.
+The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones.
+The stochastic block model and stochastic Watts-Strogatz block model, that we used, only have communities of a single size.
+This does, however, not play a significant role in the formation of echo chambers, as can be seen in Fig.
+\ref{power_vs_single}.
+In this figure, the distribution of echo chambers in the Last.fm network is compared to those in the SBM-WS network with similar clustering and modularity.
+The SBM-WS is modeled in two ways: (i) the communities all have a fixed size; (ii) the communities have the same heterogeneous sizes as in the Last.fm network.
+The PR and REC filtering algorithms are compared.\\ % add parameters of SBM-WS power??
+\newpage
+\noindent
 The results for the Watts-Strogatz model are similar to those for the stochastic Watts-Strogatz block model.
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{PR}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{REC}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_8x7.png}
@@
 As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network.
-% talk about SBM
+The results for the Last.fm network coincide with those of the SBM.
+As for the PR filtering algorithm, this seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the corresponding hampering due to the lack of high clustering.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_powerlaw_vs_regular_REC.png}
+    \label{pr_pow_single}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_powerlaw_vs_regular_PR.png}
+    \label{rec_pow_single}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+The SBM-WS with heterogeneous community sizes is compared to the SBM-WS with a single community size.
+Results are compared to the real-world Last.fm network with similar clustering and modularity.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
+The SBM-WS with heterogeneous community sizes is compared to the SBM-WS with a single community size.
+Results are compared to the real-world Last.fm network with similar clustering and modularity.
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Each curve is the average of $50$ independent simulations.
+Error bars are omitted for an improved visualization.}
+\label{power_vs_single}
+\end{figure}
 
@@
 It was also found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
-% talk about evolution of opinions??
-\newline
-Finally, it was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) does make a difference.
+\newline
+These results give us an insight in the interplay between network structure, filtering algorithms and individual resistance.
+These insights may be crucial to understand phenomena such as polarization in society and to determine the impact of social media on these phenomena.
+They learn us that even though semantic filtering (PR) strongly enhances the formation of echo chambers in the absence of stubborn actors, this enhancing does not stand on its own.
+An interplay between a network with a high clustering and/or community structure is needed to effectively see this increase in echo chambers induced by semantic filtering.
+A random network, combined with semantic filtering, does not enhance the formation of echo chambers.
+Furthermore, if there are no stubborn actors, recent or random filtering (REC/REF) seems to be more protective against the formation of echo chambers than semantic filtering.
+If there are, however, small fractions of stubborn actors in the system, this scenario changes: semantic filtering starts to hamper the formation of echo chambers, whereas recent or random filtering begins to enhance polarization.
+It is often assumed that algorithmic personalization, tailored to ones preferences, is one of the main reasons of the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
+Our results may, however, suggest, that if there is some kind of stubbornness (which is probably the case in the real world), heterogeneous exposure (REF/REC) to different opinions may be more harmful than a homogeneous exposure (PR).
+Bail et al.
+found experimental results that may agree with this finding \cite{Bail2018}.
+This is an interesting result, that should be investigated more thoroughly.\\
+\newline
+Finally, it was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) makes a difference.
 We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
@@
 For the high modularity network, the community distributed scenario showed an even larger hampering of the formation of echo chambers around the majority opinion A.
-This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase in the majority opinion A for the community distributed scenario in case of the high modularity network combined with the PR filtering algorithm.
+This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario in case of the high modularity network combined with the PR filtering algorithm.
 \\
@@
 two dominant and opposing echo chambers'' (Wang et al., p.10, \cite{Wang2020}).
-This agrees with our result that the PR filtering algorithm, combined with stubborn actors, hinders the formation of echo chambers.
+This agrees with our result that the PR filtering algorithm, combined with stubborn actors, hinders the formation of echo chambers.\\
+\newline
+It must be emphasized that there are limitations to our model and methods.
+As already stated before, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses.
+Real-life social behavior and opinion dynamics on social media may be far from simple.
+As a consequence, there may be many dynamics of real-life opinion formation that are not well-captured by our model.
+More empirical data is therefore needed to compare the results, obtained from our model, against.
+Other limitations may be the rather small system sizes that were used in our simulations ($N = 10^3$).
+It is unclear how the system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks.
+Another possible shortcoming is the binary nature of the model.
+There are real-life cases in which people have to choose between two competing opinions.
+These situations, however, do not cover the complete range of possible opinion dynamics situations.
+It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics.
+We also did not differentiate between the agents in our model, except for non-stubborns versus stubborns.
+This combined with the homogeneous degree distribution of our networks models may be another limitation.
+Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status.
+We could adapt our model to mimic this real social media behavior more closely.
+Therefore, a heterogeneous network model should be introduced, combined with an influence parameter.
+This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adapted by the others.
+In the model, the high degree nodes should then be made more influential.
+As a final limitation, we mention the fixed stubbornness of the agents.
+Even though the stubbornness parameter $r$ could be individualized to each node, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
+This idea may be linked to the influence of certain friends/acquaintances: somebody may adapt another viewpoint more easily from certain friends than from others.\\
+\newline
+Some future investigation paths, linked to limitations of our model, were given in the previous paragraph.
+Other future directions may be: (i) investigating the distribution of the stubborn actors (community distributed versus random); (ii) making one opinion more stubborn than the other; (iii) investigating the effect of the activation mechanism: what happens if we switch to the heterogeneous activation mechanism (see Section \ref{actMech})?
+and (iv) making the network temporal in the sense that it becomes possible to add and remove friends.
+Sasahara et al.
+investigated a model in which social influence was combined with social rewiring (unfriending) \cite{Sasahara2020}.
+Their findings were that these processes of influence and unfriending enhanced the formation of echo chambers \cite{Sasahara2020}.
+It might be interesting to incorporate their model features into our model with algorithmic personalization and stubbornness.
 

### 2021-05-10 draft.tex
latest update, worked on lay-out abstract + wrote acknowledgements

@@
 {\cleardoublepage\null \vfill\begin{center}\bfseries \abstractname \end{center}}{\vfill\null}
+\usepackage{abstract}
 
@@
 
+\usepackage{epigraph}
+
+\setlength\epigraphwidth{.8\textwidth}
+\setlength\epigraphrule{0pt}
+
 \addto\captionsenglish{\renewcommand{\contentsname}{Table of Contents}
@@
 \pagestyle{plain}
+
+\newcounter{abstractpage}
+\setcounter{abstractpage}{\value{page}}
+
 \renewcommand{\abstractname}{Acknowledgments}
 \begin{abstract}
+\thispagestyle{plain}
+\setcounter{page}{\value{abstractpage}}
 \addcontentsline{toc}{chapter}{Acknowledgments}
 
+\epigraph{\itshape At times, our own light goes out and is rekindled by a spark from another person.
+Each of us has cause to think with deep gratitude of those who have lighted the flame within us.}{---Albert Schweitzer}
+\vspace*{\fill}
+\noindent
+First and foremost, I would like to thank my supervisor Prof.
+Luis E.C.
+Rocha for his support, feedback and useful insights during the course of this year.
+This thesis would not be of the same quality if it wasn't for his insightful remarks.
+I learned a tremendous deal, not only about social physics and network science, but also about organizing, structuring and reporting relevant results.
+I would not have been able to learn these things without his guidance.\\
+\newline
+I also want to give special thanks to my co-supervisor Prof.
+Jan Ryckebusch.
+Not only for his support during my thesis, but mostly for the enthusiastic and passionate way in which he teaches all his classes.
+His dedication to his students and his willing to learn them valuable insights in the wonderful world of physics are truly admirable.\\
+\newline
+I am especially thankful to Yolan.
+He helped me with every possible problem I encountered during my physics education.
+If it was not for him, I would still be a (very poor) programming beginner.
+His patience and talent for explaining things in an understandable way are incredibly valuable.
+I do not only want to thank him for all his help during my study and my thesis, but also for being the great friend he is.
+This thesis, but also myself, would not have been the same without him.\\
+\newline
+Beside Yolan, I want to give credit to all of my other friends. I would not be the same person without any of them.\\
+\newline
+Finally, I would like to thank my family, and especially my parents.
+They have unbelievable faith in me and support me in every single way.
+I feel great gratitude for how they raised me and for all the opportunities that they have given me, and are still giving me every single day.\\
+\vfill
+\noindent
+Nina Botte.
+
+\setcounter{abstractpage}{\value{page}}
+
 \end{abstract}
+\setcounter{page}{\value{abstractpage}}
+\stepcounter{abstractpage}
 
@@
 \begin{abstract}
+\thispagestyle{plain}
+\setcounter{page}{\value{abstractpage}}
+
 \addcontentsline{toc}{chapter}{Abstract}
+\noindent
 The past decade, the rise of social media has become undeniable.
@@
 This thesis deals with the impact of social media on the formation and evolution of opinions in society.\\
+\newline
 Social media allows for the spread of information at a faster pace and at an unprecedented scale.
@@
 It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
+\newline
 People are, however, more than sheep that thoughtlessly follow their shepherd (contrary to what some books might lead us to believe).
@@
 This aspect of potential stubbornness should not be ignored.\\
-In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.\\
+\newline
+In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
 We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers.
 We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness.
-\\
-Beside that, we will also pay some attention to the possible ways of distributing the opinions and how this may affect the opinion dynamics in the system. \\
+Beside that, we will also pay some attention to the possible ways of distributing the opinions and how this may affect the opinion dynamics in the system.
+\\
+\newline
 Finally, we will test the model on real-world networks.
@@
 What are the similarities or differences in the results?
-What are the possible underlying reason/structures for these similarities/differences?\\
-
+What are the possible underlying reason/structures for these similarities/differences?
 % definetely not finished!! --> add some type of conclusions or not? Change part about distributing stubborn people??
+\setcounter{abstractpage}{\value{page}}
 \end{abstract}
 
+%\setcounter{page}{\value{abstractpage}}
+\stepcounter{abstractpage}
+
 \selectlanguage{dutch}
-\newgeometry{top=0mm, bottom=15mm}
 \begin{abstract}
+\thispagestyle{plain}
+\setcounter{page}{\value{abstractpage}}
+
 \addcontentsline{toc}{chapter}{Samenvatting}
+\noindent
 Sociale media zijn bijna niet meer weg te denken uit ons dagelijks leven.
@@
 \\
+\newline
 Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloedt worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,...
@@
 Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\
+\newline
 Door het toenemende gebruik van sociale media worden deze echter overspoeld door informatie.
@@
 Zo zou het kunnen leiden tot de vorming van  ``echo kamers'' waar gebruikers gevangen raken tussen ideeën die hun eigen visie versterken.
-Dit zou grote gevolgen kunnen hebben op bijvoorbeeld polarisatie in de maatschappij, opkomst van extremisme, enzovoort.
+Dit kan op zijn beurt polarisatie in de maatschappij, opkomst van extremisme, enzovoort versterken.
 Het is daarom uiterst belangrijk om de impact van deze filtering algoritmen grondig in kaart te brengen.\\
+\newline
 Mensen zijn echter ook in staat om kritisch te denken, soms ook over-kritisch of selectief kritisch, maar dat terzijde.
@@
 Deze individuele koppigheid zou men niet mogen verwaarlozen in het ontwerpen van modellen die de dynamica van opinies op sociale media willen onderzoeken.\\
+\newline
 In deze thesis wordt gebruik gemaakt van ``toy models'' die de interactie tussen verschillende netwerk modellen, filtering algoritmen en individuele terughoudenheid om van mening te veranderen onderzoeken in de evolutie van twee opinies.
@@
 \\
+\newline
 Naast het bovenstaande zal er ook kort aandacht gespendeerd worden aan hoe de twee verschillende opinies verdeeld worden over het systeem.
@@
 Er zal nagegaan worden of de resultaten van deze werkelijke netwerken gelijkenissen, dan wel verschillen, vertonen met de data verkregen van de netwerk modellen.
+\setcounter{abstractpage}{\value{page}}
 \end{abstract}
 
-\restoregeometry
+\setcounter{page}{\value{abstractpage}}
+\stepcounter{abstractpage}
+
 \selectlanguage{english}
@@
 The results of the two real-world networks: (i) the Last.fm on-line social network; (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
-For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or community structure.
+For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or modularity.
 We will mainly focus on the formation of echo chambers.
@@
 This is valid for both the real-world network and the network models.
-The stochastic Watts-Strogatz block model has a similar clustering and community structure as the Last.fm network.
+The stochastic Watts-Strogatz block model has similar clustering and community structure as the Last.fm network.
 The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
@@
 It may be tempting to attribute this to the lack of community structure in the WS model.
-It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers in the WS network than in the SBM-WS network.
+It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers.
 Furthermore, it is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity.
@@
 
+Results for the PGP trust network are compared to results for network models with similar average degree, clustering and/or modularity.
+The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.
+
 \chapter{Discussion}
@@
 If there are, however, small fractions of stubborn actors in the system, this scenario changes: semantic filtering starts to hamper the formation of echo chambers, whereas recent or random filtering begins to enhance polarization.
-It is often assumed that algorithmic personalization, tailored to ones preferences, is one of the main reasons of the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
+It is often assumed that algorithmic personalization, tailored to ones preferences (eg.
+semantic filtering), is one of the main reasons of the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
 Our results may, however, suggest, that if there is some kind of stubbornness (which is probably the case in the real world), heterogeneous exposure (REF/REC) to different opinions may be more harmful than a homogeneous exposure (PR).
@@
 \newline
-Finally, it was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) makes a difference.
+It was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) makes a difference.
 We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
@@
 \\
-% summarize real-world network results
-\newline
-The formation of echo chambers in social networks/social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
+\newline
+Finally, results for real-world networks were compared to those obtain from corresponding network models.
+It was found that the formation of echo chambers in the real-world networks is smaller than those in a network model with similar clustering and modularity.
+The reason for this hampered formation of echo chambers was attributed to the heterogeneous degree pattern of the real-world networks.
+This hampering of echo chambers due to a heterogeneous degree distribution was also found by Perra and Rocha \cite{Perra2019} when they compared a network model with a heterogeneous degree distribution to models with homogeneous contact patterns.\\
+\newline
+The formation of echo chambers in social networks/social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature (in the absence of stubborn actors) \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
 The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.
@@
 Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status.
+This may impact the spread of opinions \cite{Fan2020}\cite{Wang2019}.
 We could adapt our model to mimic this real social media behavior more closely.
 Therefore, a heterogeneous network model should be introduced, combined with an influence parameter.
-This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adapted by the others.
+This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adopted by the others.
 In the model, the high degree nodes should then be made more influential.
@@
 Even though the stubbornness parameter $r$ could be individualized to each node, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
-This idea may be linked to the influence of certain friends/acquaintances: somebody may adapt another viewpoint more easily from certain friends than from others.\\
+This idea may be linked to the influence of certain friends/acquaintances: somebody may adopt another viewpoint more easily from certain friends than from others.\\
 \newline

### 2021-05-11 draft.tex
made small changes

@@
 \newline
-Beside Yolan, I want to give credit to all of my other friends. I would not be the same person without any of them.\\
+Beside Yolan, I want to give credit to all of my other friends. They make my life valuable, joyful and are a true enrichment.\\
 \newline
@@
 Deze filtering algoritmen filteren, zoals de naam al impliceert, en individualiseren de informatie die iemand te zien krijgt op zijn/haar tijdslijn.
-Een gebruiker op sociale media krijgt dus niet alle berichten die de vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn/haar sociale media geschiedenis.
+Een gebruiker op sociale media krijgt dus niet alle berichten die zijn/haar vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn/haar sociale media geschiedenis.
 Dit filteren/individualiseren kan, mogelijks, grote gevolgen hebben.
 Zo zou het kunnen leiden tot de vorming van ``echo kamers'' waar gebruikers gevangen raken tussen ideeën die hun eigen visie versterken.
-Dit kan op zijn beurt polarisatie in de maatschappij, opkomst van extremisme, enzovoort versterken.
+Dit kan op zijn beurt polarisatie in de maatschappij, opkomst van extremisme, enzovoort bevorderen.
 Het is daarom uiterst belangrijk om de impact van deze filtering algoritmen grondig in kaart te brengen.\\
@@
 \newline
-One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system.
-Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model, the Watts-Strogatz model, etc.
+Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model.
 These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
@@
 
-As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg.
+The formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg.
 stubbornness), the information someone is exposed to, etc.
@@
 Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}.
-Even if the dynamics/behavior of individuals would be well know, then the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
+Even if the dynamics/behavior of individuals would be well know, the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
 Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}.
 The use of statistical physics may, however, also offer a way out.
-In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}.
+In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}\cite{Gug2015}.
 Instead, the relevant aspects for the global behavior are often higher level features such as symmetries, conservation laws, dimensionality, etc \cite{Castellano2009}.
-This introduces a concept of universality.
+This introduces a concept of universality \cite{Gug2015}.
 With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
@@
 where $N$ is the total number of nodes in the network.
-The degree distribution also allows us to classify networks.
+The degree distribution allows us to classify networks.
 The two most important classes are homogeneous and heterogeneous networks.
@@
 Such a network is called a neutral network \cite{F.Costa2007}.
-If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative.
+If $k_{nn}(k)$ is an increasing function of $k$, nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative.
 If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
@@
 
-Several useful measurements quantify the notion of length and distance in a network/graph.
+Several measurements quantify the notion of length and distance in a network/graph.
 We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
-The length of a path is then defined as the number of edges in the path \cite{Goddard2010}.
+The length of a path is defined as the number of edges in the path \cite{Goddard2010}.
 A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (the geodesic path is not necessarily unique) \cite{F.Costa2007}.
@@
 \begin{equation}\label{avdist}
-	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ .
+	\left<d\right> = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ .
 \end{equation}
@@
 These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}.
-Instead, another definition of average distance is used: the global efficiency $E$ \cite{F.Costa2007}
+Instead, another definition of average distance can be used: the global efficiency $E$ \cite{F.Costa2007}
 \begin{equation}
@@
 \noindent
-Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix in the following way \cite{F.Costa2007}
+Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix \cite{F.Costa2007}
 \begin{align}
@@
 In the case of the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$ this is, however, not the case.
-The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node \cite{Newman2003}.
+The difference between the two definitions is that the average of Eq.
+(\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Eq.
+(\ref{avClus}) gives the same weight to each node \cite{Newman2003}.
 This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.
@@
 \newline
-The modularity for a network with $N$ nodes and $L$ edges and partitioned into $n_c$ communities is calculated as follows.
+The modularity of a network with $N$ nodes and $L$ edges and partitioned into $n_c$ communities is calculated as follows.
 Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
@@
 \end{equation}
-where the factor $\frac{1}{2L}$ is just a normalization factor. Using \cite{Albert2016}
+where the factor $\frac{1}{2L}$ is a normalization factor. Using \cite{Albert2016}
 \begin{equation}
@@
 \newline
-As said before, modularity is often used in community detection algorithms.
+Modularity is often used in community detection algorithms.
 It is thus important to be aware of the limitations of modularity.
@@
 and if there is at least one edge between the two communities; even if $A$ and $B$ are distinct communities \cite{Albert2016}!
-Equation (\ref{reslim}) is called the resolution limit: communities that are smaller than this limit will not be detected by modularity maximization \cite{Albert2016}.
+Eq.
+(\ref{reslim}) is called the resolution limit: communities that are smaller than this limit will not be detected by modularity maximization \cite{Albert2016}.
 
@@
 \begin{equation}\label{degDistRan}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
+	P(k) = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
 \end{equation}
@@
 \begin{split}
-	\left<k\right> &= \sum_{k=0}^{N-1} k\  p_k  \\
+	\left<k\right> &= \sum_{k=0}^{N-1} k\  P(k)  \\
 &= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
@@
 \begin{equation}
-	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!} \ .
+	P(k) = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!} \ .
 \end{equation}
@@
 The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates).
-If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<cc\right> \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}.
+If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<cc\right> \sim \mathcal{O}(1/N)$, which becomes small for large networks \cite{Clauset2011}.
 \\
@@
 A more precise calculation can be found in Appendix \ref{ER}, Section \ref{ERdist}.
-Equation (\ref{diameter}) might, however, be a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
-This is because $d_{\textrm{max}}$ is often dominated by a few extreme paths, where $\left<d\right>$ instead is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
+Eq.
+(\ref{diameter}) might, however, be a better approximation to the average distance $\left<d\right>$ between two randomly chosen nodes \cite{Albert2014}.
+This is because $d_{\textrm{max}}$ is often dominated by a few extreme paths, whereas $\left<d\right>$ is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
 Hence, the average distance of a random network is defined by \cite{Albert2014}
@@
 To summarize: the random Erd\H{o}s-R\'{e}nyi network has a bell shaped degree distribution which approximates a Poisson distribution for $N \gg 1$; the average clustering coefficient becomes small for large networks and the average distance depends logarithmic on the system sizes.
-The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the ``small world phenomenon'' (namely short average distance).\\
+The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the ``small world phenomenon'', a short average distance.\\
 \newline
 Real social networks, however, usually display the ``small world phenomenon'' and thus have both a short average distance and a high clustering coefficient.
-The question then rises: can we construct a model that has both of these properties?
+The question then rises: Can we construct a model that has both of these properties?
 The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
@@
 \newline
-The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq.
+The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (Eq.
 (\ref{C}) and Eq.
@@
 \end{equation} 
-where $\left<cc(0)\right>$ is given by Equation (\ref{C0}).
+where $\left<cc(0)\right>$ is given by Eq.
+(\ref{C0}).
 This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
@@
 
-The stochastic block model (SBM) is another random graph model that is able to produce networks containing a community structure.
+The stochastic block model (SBM) is a random graph model that is able to produce networks containing a community structure.
 The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
@@
 \newline
-The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}: (i) The number of nodes $N$; (ii) A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities; (iii) A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.\\
-\newline
-There are no real restrictions on the edge probabilities, which makes the stochastic block model a very general model that is capable of reproducing many different network structures \cite{Karrer2011}.
+The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}: (i) The number of nodes $N$; (ii) A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities; and (iii) A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.\\
+\newline
+There are no real restrictions on the edge probabilities, which makes the stochastic block model a general model that is capable of reproducing many different network structures \cite{Karrer2011}.
 However, the edge probabilities inside a community are often taken to be bigger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
-In this way, the stochastic block model generates a network with a clear community structure.
+In this way, the stochastic block model generates a network with community structure.
 A stochastic block model that fulfills this condition is called an assortative SBM \cite{Lee2019}.
@@
 \newline
-In this thesis, the stochastic block model is generated in the following way.
-First, the parameters of the model are specified.
-Then, for each community $C_i$ an Erd\H{o}s-R\'{e}nyi network is constructed, where the number of nodes equals the size of the community $C_i$ and the edge probability $p$ is given by $M_{ii}$.
-Finally, edges are added between communities $C_i$ and $C_j$ with probability $p = M_{ij}$.\\
-\newline
 The stochastic block model can be fitted to data \cite{Clauset2017}.
@@
 The consequence of this is that if a network possesses a more skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
-The reason behind this is one of the main assumptions of the stochastic block model, namely the fact that the edge probabilities only depend  on which group the node belongs to \cite{Clauset2017}.
+The reason behind this is one of the main assumptions of the stochastic block model: the edge probabilities only depend on which group the node belongs to \cite{Clauset2017}.
 In other words, nodes that belong to the same group are stochastically equivalent, meaning that they have equivalent connectivity patterns to other nodes \cite{Clauset2017}.
@@
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
-We will not discuss their variation, since that would lead us too far astray, but interested readers can find more detailed information in reference \cite{Karrer2011}.
+We will not discuss their variation, since that would lead us too far astray, but interested readers can find more detailed information in \cite{Karrer2011}.
 \\
@@
 \newline
-As said before, the stochastic block model is a very flexible model and there are many different extensions.
+The stochastic block model is a flexible model and there are many different extensions.
 One extension that is used in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
@@
 They may posses some of the characteristics of real-world networks, but they certainly do not always describe real-world networks appropriately.
-The question then arises: what are the main properties/characteristics of real-world networks, and more specifically of real-world social networks?
+The question then arises: What are the main properties/characteristics of real-world networks, and more specifically of real-world social networks?
 Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
 \newline
-As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties.
+As discussed briefly in the introduction, it is found that real-world networks often posses three characteristic properties.
 These are the two properties of the ``small world phenomenon'', namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
@@
 A short average distance facilitates a fast transmission of information \cite{Newman2003}\cite{Zhang2014}.\\
-\newline
+\newpage
+\noindent
 \textbf{High clustering coefficient}\\
 \newline
-In real-world social networks, one often observes the pattern that two acquaintances/friends of an individual are also acquaintances/friends of each other \cite{RealWorld}.
+In real-world social networks, one often observes the pattern that two friends of an individual are also friends of each other \cite{RealWorld}.
 In graph terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
@@
 Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
-If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges, however simple scale-free networks with an exponent smaller than two cannot exist, since the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}; if $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance \cite{Newman2005}.
+If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges \cite{Newman2005}.
+Simple scale-free networks with an exponent smaller than two, however, cannot exist, since the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}.
+If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance \cite{Newman2005}.
 A diverging variance means that the fluctuations around the average can be arbitrary large, which makes this average degree no longer a characteristic scale for the system \cite{Barabasi2016}.\\
@@
 \newline
-As said earlier, these three main characteristics are valid for on- and off-line social networks.
+These three main characteristics are valid for on- and off-line social networks.
 Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}.
@@
 \newline
-There are, however, some properties that are different between on-line and off-line social networks as well.
+There are, however, some properties that are different between on-line and off-line social networks.
 One of them might be the degree correlations/mixing pattern.
@@
 \label{chap3}
-The study of opinion dynamics in a group of people, as already mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
+The study of opinion dynamics in a group of people, as mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
 The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
-The current chapter focuses on the second important layer, that of the opinion dynamics models.
+The current chapter focuses on the second layer, that of the opinion dynamics models.
 These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
@@
 \newline
-Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.\\
-\newline
-Some major attention will also be given to the content curation/filtering algorithms used by social media companies (Section \ref{filter}).
+Some major attention will be given to the content curation/filtering algorithms used by social media companies (Section \ref{filter}).
 This is needed because this thesis deals with opinion dynamics on on-line social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
 \newline
-There will also be some attention on the activation mechanism (Section \ref{actMech}).
+We will also discuss the activation mechanism (Section \ref{actMech}).
 This is the mechanism that describes the timings in which content/information is exchanged between users \cite{Perra2019}.
-Such a mechanism changes the network from a static one to a more temporal one, where nodes become active and inactive over time.
-\\
-\newline
+Such a mechanism changes the network from a static one to a temporal one, where nodes become active and inactive over time.
+\\
+\newline
+Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.
 There are two major groups of opinion dynamics models, namely binary/discrete models and continuous approaches.
@@
 One of the simplest opinion dynamics models is the binary majority model.
-In this model each agent has one of the two opinions A or B.
+Each agent has one of the two opinions A or B.
 Initially, these opinions can be equally distributed or not.
 If the distribution is not equal, but instead eg.
-80/20, there is a predominant opinion and a minority opinion.
-The following updating rule is then imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
+$80/20$, there is a predominant opinion and a minority opinion.
+The following updating rule is imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
 Initially each agent is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
@@
 Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
-This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini, Chan, O'Donnell, Tamuz and Tan \cite{Benjamini2016}.
+This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini et al.
+\cite{Benjamini2016}.
 It is also found that the rate of convergence increases with increasing edge density and that if the network is sparse enough, consensus will not be reached and the two opinions coexist \cite{Nguyen2020}.
-This can be explained by the fact that if the network is sparse enough, we may have disconnected parts; each of these disconnected groups may then evolve to a different consensus opinion, so that over the entire network different opinions coexist \cite{Nguyen2020}.
+This can be explained by the fact that if the network is sparse enough, we may have disconnected parts; each of these disconnected groups may evolve to a different consensus opinion, so that over the entire network different opinions coexist \cite{Nguyen2020}.
 
@@
 
-The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature.
+The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic nature.
 In this model the fraction of neighbors of user $i$ with opinion A, $p_\text{A}$, and the fraction of neighbors with opinion B, $p_\text{B}$, are determined.
@@
 For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
-Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
+User $i$ no longer simply updates his/her opinion according to the majority opinion carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
 \newline
@@
 Hence, we will not go into further detail.
-Interested readers are referred to references \cite{Castellano2009}, \cite{Fernandez2014} and related references therein.
+Interested readers are referred to \cite{Castellano2009}, \cite{Fernandez2014} and related references therein.
 
@@
 \newline
-In continuous opinion dynamics models, some concepts that are valid for the discrete models no longer apply (eg.
+Some concepts that are valid for the discrete models no longer apply (eg.
 the concept of a majority/minority opinion or the concept of equality of opinions); hence, they require a bit of a different framework \cite{Castellano2009}.
@@
 As mentioned earlier, the threshold $\epsilon$ can be different for each node, so that each node has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
-\epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to references \cite{Deffuant2006} and \cite{Deffuant2002}.
-\newpage
+\epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to \cite{Deffuant2006} and \cite{Deffuant2002}.
+
 \section{Algorithmic personalization}\label{filter}
@@
 In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.
-\newpage
+
 % NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where nodes can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, eg.
@@
 If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.\\
-\newpage
-\noindent
+\newline
 The resistance parameter $r_i$ can be different or the same for all nodes.
@@
 This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.
-A simple model is more tractable and allows for a better understanding of what is going on in the dynamics.
+A simple model is more tractable and allows for a better understanding of what is going on in the dynamics \cite{Jed2019}.
 However, this might be a downside as well.
@@
 Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
-This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Pawel}.
+This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
 \\
@@
 This does not mean that simple models should be thrown away.
-They might still be capable of capturing a lot of the dynamics that are seen in the real world.
-Not because a system appears complex, that it cannot be modeled by something far more simple.
-This has been proven by physics research over and over again.
+They might still be capable of capturing a lot of the dynamics that are seen in the real world \cite{Gug2015}.
+Not because a system appears complex, that it cannot be modeled by something far more simple \cite{Jed2019}.
+This has been proven by physics research over and over again \cite{Gug2015}.
 Beside that, if you do not know the dynamics/underlying processes of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
@@
 \begin{equation}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
+	P(k) = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .
 \end{equation}
@@
 \begin{equation}\label{pk}
-	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ ,
+	P(k) = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ ,
 \end{equation}
@@
 \begin{split}
-	p_k &= \binom{N-1}{k}p^k (1-p)^{N-1-k} = \frac{(N-1)^k}{k!} p^k \mathrm{e}^{-\left<k\right>} \\
+	P(k) &= \binom{N-1}{k}p^k (1-p)^{N-1-k} = \frac{(N-1)^k}{k!} p^k \mathrm{e}^{-\left<k\right>} \\
 &= \frac{(N-1)^k}{k!} \bigg(\frac{\left<k\right>}{N-1} \bigg)^k \mathrm{e}^{-\left<k\right>}\\

### 2021-05-12 draft.tex
made small changes

@@
 They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
-If social media companies would not use any kind of filtering, there would be way to many content to handle and process for the social media user \cite{Bozdag2013}.
+If social media companies would not use any kind of filtering, there would be too much content to handle and process for the social media user \cite{Bozdag2013}.
 This in turn would make the on-line social experience unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
@@
 It is important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
-Perra and Rocha already partly investigated the role of such algorithms on the dynamics of opinions in several networks in one of their previous papers, \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} \cite{Perra2019}.
+Perra and Rocha partly investigated the role of such algorithms on the dynamics of opinions in several networks in \cite{Perra2019}.
 Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}: (i) Popularity filtering: promoting content that is popular across the platform; (ii) Semantic filtering: promoting posts that are similar to previous consumed posts; (iii) Collaborative filtering: promoting content that people similar or close to us consume.\\
@@
 At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}); the following selection methods are used \cite{Perra2019}: (i) The reference filtering algorithm (REF): this filtering algorithm randomly selects $S$ items from the list $L_i(t)$; (ii) The old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user; (iii) The recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$; (iv) The preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and then selects the first $S$ elements.
-This filtering algorithm is based on the semantic filtering explained above.\\
+The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}).
+The following selection methods are used \cite{Perra2019}: (i) the reference filtering algorithm (REF): randomly select $S$ items from the list $L_i(t)$; (ii) the old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user; (iii) the recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$; and (iv) the preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and selects the first $S$ elements.
+This filtering algorithm is based on semantic filtering.\\
 \newline
 Perra and Rocha showed that the REF, OLD and REC filtering algorithms behave similar in many cases and that the PR filtering algorithm often displays different behavior (such as a larger formation of echo chambers compared to the other three filtering algorithms) \cite{Perra2019}.
-In this thesis the REC and REF filtering algorithms are mainly used as a reference/benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
+In this thesis the REC and REF filtering algorithms are used as a reference/benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
 
@@
 
-Most of the simple opinion dynamics models make use of a static network in which users are always active and always exchanging opinions/posts with each other.
+Most of the simple opinion dynamics models make use of a static network in which users are always active and always exchanging opinions.
 In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented.
@@
 \newline
-It might, however, be a more realistic situation if nodes/users become active/inactive over time.
+It might, however, be more realistic if nodes/users become active/inactive over time.
 This would be a better representation of real social media platforms, where users are active for some periods of time and then become inactive again.
-This kind of activation mechanism of the users gives the network a more temporal character.
-The underlying network structure does not change (in the sense that each user/node has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
+This kind of activation mechanism of the users gives the network a temporal character.
+The underlying network structure does not change (each user/node has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
 Hence, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
@@
 In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes.
-Perra and Rocha chose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
+Perra and Rocha chose activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
 They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note that being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (also note that having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+However, there are groups in society, often with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
 \newline
@@
 The model he proposed goes as follows.
-Remember that in the bounded confidence models, the agents have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
+In bounded confidence models, the agents have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
 Guillaume Deffuant then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or the extreme opinion $1$ \cite{Deffuant2006}.
@@
 \newline
-Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in reference \cite{Deffuant2006}. \\
+Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in \cite{Deffuant2006}. \\
 \newline
 Another example of introducing some kind of persistence towards the own opinion of the nodes is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
-The model introduced by Friedkin and Johnsen in which each agent adheres to its initial opinion to a certain degree, see Section \ref{classMod}, can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
+The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each agent adheres to its initial opinion to a certain degree can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
 
@@
 One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
-In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions.
+In binary models, however, there are only two possible opinions and in that sense it is less useful to talk about extreme and moderate opinions.
 However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
@@
 \newline
-There are different possibilities of implementing this kind of model.
+There are different possibilities of implementing this model.
 One way, which makes use of the majority model, is as follows.
@@
 If $T = 1$, the nodes do not change their opinion.
-For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
-\newline
-Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model, many more exist or can be invented.
+For $0 < T < 1$, the model is a majority model, but the nodes are more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
+\newline
+The above examples are only a handful of the possible ways to include some notion of stubbornness in the model, many more exist or can be invented.
 
@@
 \newline
-Finally, Section \ref{opinionmodel} summarizes and discusses the opinion dynamics model that is implemented in this thesis.
-Again we will give some focus on possible advantages and disadvantages of the model.
+Finally, Section \ref{opinionmodel} summarizes and discusses the opinion dynamics model that is implemented. We will focus on possible advantages and disadvantages of the model.
 
@@
 
-This thesis uses three main network models: (i) The Erd\H{o}s-R\'{e}nyi model (ER); (ii) The Watts-Strogatz model (WS); (iii) The stochastic block model, of which two types are implemented: the regular stochastic block model (SBM) and the stochastic Watts-Strogatz block model (SBM-WS).\\
-\newline
-Each of these networks serve specific goals which will shortly be discussed in the following paragraphs.\\
+This thesis uses three network models: (i) the Erd\H{o}s-R\'{e}nyi model (ER); (ii) the Watts-Strogatz model (WS); and (iii) the stochastic block model, of which two types are implemented: the regular stochastic block model (SBM) and the stochastic Watts-Strogatz block model (SBM-WS).\\
+\newline
+Each of these networks serve specific goals which will be shortly discussed in the following paragraphs.\\
 \newline
@@
 The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
-In this thesis we focus intensively on two network features/topologies that may impact the formation of echo chambers, namely a high clustering coefficient and community structure.
+In this thesis we focus intensively on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
@@
 \newline
-The Watts-Strogatz model is used in as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-\newline
-In the paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (reference \cite{Perra2019}) Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
+The Watts-Strogatz model is used as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
+\newline
+Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
 Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
@@
 \newline
-The stochastic Watts-Strogatz block model is used as a model that has both a high clustering and a community structure.
+The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
 We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
@@
 However, most of these companies are not to keen on sharing any form of data.
-This makes it very hard to find accurate and complete social media network data.
+This makes it hard to find accurate and complete social media network data.
 Hence, we chose to test the model on other types of social networks, either on-line or off-line.
@@
 The main, and actually only, reason for this was the computation time.
-If the networks would be to large, it would take weeks, sometimes even months before there would be any kind of results.\\
+If the networks are too big, it would take weeks, sometimes even months before there would be results.\\
 \newline
@@
 Only the largest connected component of this on-line social network is included \cite{ICON}.
-This data set was originally used in reference \cite{Toivonen2009}, which describes the data set.
-This on-line social network has the following properties: number of nodes: $N = 8003$; number of edges: $16824$; modularity: $Q = 0.79$; average clustering coefficient: $\left<cc\right> = 0.31$; average degree: $\left<k\right> = 4.2$.
+This data set was originally used in \cite{Toivonen2009}, which describes the data set.
+This on-line social network has the following properties: $N = 8003$; number of edges: $16824$; $Q = 0.79$; $\left<cc\right> = 0.31$; $\left<k\right> = 4.2$.
 The degree distribution is heterogeneous and is best fitted by a log-normal
@@
 The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{ICON}.
-This network was originally used in reference \cite{Boguna2004}, which gives a detailed description of the network data.
-The main properties of this network are: number of nodes: $N = 10680$; number of edges: $24340$; modularity: $Q = 0.85$; average clustering coefficient: $\left<cc\right> = 0.27$; average degree: $\left<k\right> = 4.56$.
+This network was originally used in \cite{Boguna2004}, which gives a detailed description of the network data.
+The main properties of this network are: $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
 The network has a heterogeneous degree distribution, which is best fitted by two power-laws
@@
 
-In Chapter \ref{chap3} we discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
+In Chapter \ref{chap3} we discussed opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
 The goal of this section is to put everything together and summarize the opinion dynamics model of this thesis.\\
 \newline
-This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} \cite{Perra2019}.
-\\
-\newline
+This thesis follows the opinion dynamics model that was designed by Perra and Rocha in \cite{Perra2019}.
 We start with a network with $N = 10^3$ users.
@@
 Each node is given a stubbornness parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$ the node can update his opinion, if $f \leqslant r_i$ the node will keep his current opinion.\\
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the node can update his opinion, if $f \leqslant r_i$, the node will keep his current opinion.\\
 \newline
@@
 We can however assume that they are not as simple as the ones we used in this thesis.
-Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
+Trying to model these far more complex systems by our simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
 This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
@@
 They might still be capable of capturing a lot of the dynamics that are seen in the real world \cite{Gug2015}.
-Not because a system appears complex, that it cannot be modeled by something far more simple \cite{Jed2019}.
+Not because a system appears complex, that it cannot be modeled by something more simple \cite{Jed2019}.
 This has been proven by physics research over and over again \cite{Gug2015}.
@@
 There is no need to make things too complex if simple works too.
-There is however a need to compare data from (computational) models to experimental, real-world data.
+There is however a need to compare data from (computational) models to empirical, real-world data.
 A need for critical thinking and careful analyzing of results can also not be denied \cite{Pawel}.
@@
 \newline
-Four different network models are implemented, each with a total number of nodes $N = 10^3$ and an average degree $\left<k\right> \sim 10$.
+Four different network models are implemented, each with a total number of nodes $N = 10^3$ and average degree $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
 To investigate the effect of clustering on the dynamics, we use the Watts-Strogatz model (WS).
-Two different values of the rewire probability $\beta$ are used, each with a slightly different value of the clustering coefficient $\left<cc\right>$: $\beta = 0.01$ with $\left<cc\right> = 0.648 \pm 0.003$ and $\beta = 0.06$ with $\left<cc\right> = 0.557 \pm 0.007$.
+Two different values of the rewire probability $\beta$ are used, each with a slightly different value of the clustering coefficient: $\beta = 0.01$ with $\left<cc\right> = 0.648 \pm 0.003$ and $\beta = 0.06$ with $\left<cc\right> = 0.557 \pm 0.007$.
 The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are investigated: a high modularity $Q = 0.908 \pm 0.005$ with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Two different modularities are investigated: a high modularity, $Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
 Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
@@
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and a high modularity ($Q = 0.909 \pm 0.004$).
+This is a network with both high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and high modularity ($Q = 0.909 \pm 0.004$).
 \\
@@
 model.
-This is to be expected, since a stochastic block model with a low modularity, and hence a very blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network.
+This is to be expected, since a stochastic block model with a low modularity, and hence a blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network.
 The stochastic Watts-Strogatz block model (SBM-WS) has both community structure and high average clustering.
@@
 For balanced starting conditions, none of the network models, nor the filtering algorithms could break the status quo (Fig.
-\ref{ev_op_50_50} and related results in reference \cite{Perra2019}).
+\ref{ev_op_50_50} and related results in \cite{Perra2019}).
 Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a more clustered opinion distribution and thus lead to polarization and formation of echo chambers.
@@
 \newline
-We study four different network models each with a total number of nodes $N = 10^3$ and an average degree $\left<k\right> \sim 10$.
+We study four different network models each with a total number of nodes $N = 10^3$ and average degree $\left<k\right> \sim 10$.
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
@@
 The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are investigated: a high modularity $Q = 0.908 \pm 0.005$ with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Two different modularities are investigated: a high modularity, $Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
 Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
@@
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both a high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and a high modularity ($Q = 0.909 \pm 0.004$).\\
+This is a network with both high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and high modularity ($Q = 0.909 \pm 0.004$).\\
 \newline
@@
 \newline
-This result differs from reference \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
-The results that where obtained in reference \cite{Perra2019} are, however, for networks with a total number of nodes $N = 10^5$ and average degree $\left<k\right> \sim 6$.
+This result differs from results in \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
+The results that where obtained in \cite{Perra2019} are, however, for networks with a total number of nodes $N = 10^5$ and average degree $\left<k\right> \sim 6$.
 In this thesis results are obtained for $N=10^3$ and average degree $\left<k\right> \sim 10$.
@@
 Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for a better visualization.}
+Error bars are omitted for an improved visualization.}
 \label{echo_20_80}
@@
 When using the REC filtering algorithm, the formation of echo chambers around the majority opinion B is hampered.
-This is probably a consequence of the absence of clustering and the presence of short average path lengths in these models, since $80\%$ of the nodes carry this opinion.
+This is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
 There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion A.
-From figure \ref{ev_op_20_80} it was concluded that the REC filtering algorithm does not break the prevalence of the opinions.
+From Fig.
+\ref{ev_op_20_80} it was concluded that the REC filtering algorithm does not break the prevalence of the opinions.
 The opinion distribution is however slightly shifted from the case where the minority opinion A is more or less randomly distributed over the network to the case where the minority opinion A groups together in opinion clusters.
@@
 This means that the average fraction of nodes with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
-In case of the PR filtering algorithm this value is even slightly higher than $1$, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+In case of the PR filtering algorithm this value is slightly higher than $1$, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
 
@@
 The data points represent the value of the echo chamber size in networks with a different modularity.
-The initial opinion distribution was $50/50$ and was, initially, randomly distributed over the network.
+The initial opinion distribution was $50/50$ and was randomly distributed over the network.
 \\
@@
 The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-The interplay between stubbornness and clustering is studies with the Watts-Strogatz model (WS) with mean degree $\left<k\right>=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
+The interplay between stubbornness and clustering is studied with the Watts-Strogatz model (WS) with mean degree $\left<k\right>=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
 The effect of community structure is investigated with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are implemented: a high modularity, $Q = 0.908 \pm 0.005$, with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity, $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+Two different modularities are implemented: a high modularity, $Q = 0.908 \pm 0.005$, with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
 Finally, the interplay between stubbornness and both clustering and community structure is studied with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
@@
 The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
-What happens if we use the REF/REC filtering algorithm?
+What happens if we use the REF/REC filtering algorithms?
 Again we have one node with opinion B and two nodes with opinion A in the triangle.
@@
 This would explain the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
-If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithm, starts to decrease which can also be easily seen with the help of the triangle example.
+If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithms, starts to decrease which can also be easily seen with the help of the triangle example.
 If the fraction of stubborn actors becomes too high, we might end up with the case where both the node with opinion B and one of the nodes with opinion A are stubborn.
@@
 The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases.
-For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slowly decrease between $r\sim 0.7$ and $r\sim 0.9$.
-Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers.
-The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model.
-The SBM-WS model, on the other hand, displays a continuous increase until $r \sim 0.9$.
-In both case we have a sharp drop from $r \sim 0.9$ to $r=1$.
+For the REF filtering algorithm, this increases reaches a maximum around $r = 0.7$ and then starts to slowly decrease between $r = 0.7$ and $r = 0.9$.
+Between $r = 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers.
+The REC filtering algorithm also seems to have a maximum around $r = 0.7$ for the WS model.
+The SBM-WS model, on the other hand, displays a continuous increase until $r = 0.9$.
+In both case we have a sharp drop from $r = 0.9$ to $r=1$.
 If $r=1$, the nodes are completely stubborn and thus keep their initial opinion.
@@
 This is represented by the black dotted line.
-It is remarkable that for high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is large and even larger than if the nodes are not stubborn.
+It is remarkable that for high levels of the stubbornness of the nodes $r$, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the nodes are not stubborn.
 We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithms in case of networks with a high clustering coefficient.
@@
 \ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of unbalanced starting conditions, combined with the PR filtering algorithm is non-zero.
-These figures are not really suited to (quantitatively) compare the formation of echo chambers in the different network models in case of unbalanced starting conditions.
+These figures are not suited to (quantitatively) compare the formation of echo chambers in the different network models in case of unbalanced starting conditions.
 In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
@@
 \newline
-It becomes clear that the PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
-It can also be observed that the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
+The PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
+Furthermore, the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
 This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
@@
 We will study a network with a high modularity, representing the case of more or less isolated communities, and a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
-More specifically, the two stochastic block models that are implemented are: (i) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ ($Q = 0.908 \pm 0.005$); (ii) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ ($Q = 0.220 \pm 0.007$).
-Both networks have a total number of nodes equal to $N = 10^3$ and average degree $\left<k\right> = 10.79 \pm 0.14$ and $\left<k\right> = 10.17 \pm 0.14$, respectively.\\
+More specifically, the two stochastic block models that are implemented are: (i) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ ($Q = 0.908 \pm 0.005$); and (ii) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ ($Q = 0.220 \pm 0.007$).
+Both networks have a total number of nodes $N = 10^3$ and average degree $\left<k\right> = 10.79 \pm 0.14$ and $\left<k\right> = 10.17 \pm 0.14$, respectively.\\
 \newline
@@
 \ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
-A fraction equal to $0.1$ of the communities is given a single opinion, namely opinion A, while the other communities get an initial $50/50$ opinion distribution.
+A fraction equal to $0.1$ of the communities is given a single opinion, opinion A, while the other communities get an initial $50/50$ opinion distribution.
 The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in this case).
@@
   \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($55/45$ in this case) is randomly distributed over each community.
+Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.
 The stochastic block model with 10 communities of size 100 is implemented with two different modularities: $Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$, $\left<k\right> = 10.79 \pm 0.14$ and $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$, $\left<k\right> = 10.17 \pm 0.14$.
@@
 \newline
-The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced starting conditions and if the opinions are randomly distributed.
+The PR filtering algorithm leads to an increase in the prevalence of the majority opinion in case of unbalanced starting conditions and if the opinions are randomly distributed.
 This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
@@
 \ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
-This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present.
+This is also visible in the results since the hampering in the increase of the majority opinion is no longer present.
 The community distributed case behaves equivalent to the random distributed case.
@@
 The community distributed cases are compared to the corresponding randomly distributed cases.
-Results are shown for both a high modularity network and a low modularity network.
+Results are shown for both a high modularity and a low modularity network.
 Fig.
@@
 The communities with the single opinion A see, however, a decrease in the prevalence of opinion A and become more mixed.
-We can also observe a small increase in the number of echo chambers around the minority opinion B.
+We also observe a small increase in the number of echo chambers around the minority opinion B.
 This indicates that the minority opinion starts to group together.
@@
 
-The results of the two real-world networks: (i) the Last.fm on-line social network; (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
+The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
 For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or modularity.

### 2021-05-13 draft.tex
made some small changes, also in appendix

@@
 Finally, we will run the opinion dynamics model of this thesis on some real-world networks.
-We will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
+We will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or modularity.
 
@@
 The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
-The results for the PR and REC filtering algorithm are shown in Fig.
+The results for the PR and REC filtering algorithms are shown in Fig.
 \ref{echo_vs_fracRes_real}.\\
@@
 This is valid for both the real-world network and the network models.
-The stochastic Watts-Strogatz block model has similar clustering and community structure as the Last.fm network.
+The stochastic Watts-Strogatz block model has similar clustering and modularity as the Last.fm network.
 The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
-This suggests that it is not only clustering and/or community structure that drives the formation of echo chambers.
+This suggests that it is not only clustering and/or modularity that drives the formation of echo chambers.
 Other topologies must play a role as well.
@@
 The REC filtering algorithm enforces an increase in the formation of echo chambers when a small fraction of stubborn actors is introduced.
-If this fraction is further increased, the formation of echo chambers starts to decrease again.
+If this fraction is further increased, the formation of echo chambers starts to decrease.
 The increase is the strongest for the high clustering network models (WS, SBM-WS).
@@
 Furthermore, a notion of stubbornness was introduced to investigate the interplay between individual resistance to change, network structure and filtering algorithm.
-For simplicity we only allowed two competing opinions (A and B).\\
+For simplicity we only allowed two competing opinions (A and B).
+Furthermore, we studied two different settings: balanced starting conditions and unbalanced starting conditions.\\
 \newline
@@
 
-% put here: be carefull with extrapolating data/results to real world cases, need for more real data to compare with (eg data from role of facebook in elections in US), this lack of comparison between real data and network model data/analytical work = one of the downfalls/backsides of social physics??
-% relevance to society of knowing the impact of social media on opinion dynamics 
-% important to know the limits of the model, may be intuitive, but intuition maybe misleading (learned in physics) --> why we need experimental data to compare with!
-
 \chapter{Conclusions and outlook}
@@
 \end{equation}
-where $c_i$ represents the community to which node $i$ belongs and where the Dirac-delta function $\delta_{c_i, c_j}$ expresses the fact that only pairs of nodes that belong to the same community are taken into account.
+where $c_i$ represents the community to which node $i$ belongs and the Dirac-delta function $\delta_{c_i, c_j}$ expresses the fact that only pairs of nodes that belong to the same community are taken into account.
 We can then rewrite the first term as a sum over communities \cite{Albert2016}
 \begin{equation}\label{1}
-	\frac{1}{2L}\sum_{i, j = 1}^N A_{ij}\delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{2L}\sum_{i, j \in C_c} A_{ij} = \sum_{c=1}^{n_c}\frac{L_c}{L} \ ,
+	\frac{1}{2L}\sum_{i, j = 1}^N A_{ij}\delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{2L}\sum_{(i, j) \in C_c} A_{ij} = \sum_{c=1}^{n_c}\frac{L_c}{L} \ ,
 \end{equation}
@@
 \begin{equation}\label{2}
-	\frac{1}{2L}\sum_{i, j = 1}^N \frac{k_ik_j}{2L} \delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{(2L)^2}\sum_{i, j \in C_c} k_ik_j = \sum_{c=1}^{n_c}\frac{k_c^2}{4L^2} \ ,
+	\frac{1}{2L}\sum_{i, j = 1}^N \frac{k_ik_j}{2L} \delta_{c_i, c_j} = \sum_{c=1}^{n_c}\frac{1}{(2L)^2}\sum_{(i, j) \in C_c} k_ik_j = \sum_{c=1}^{n_c}\frac{k_c^2}{4L^2} \ ,
 \end{equation}
@@
 \begin{equation}\label{Q_AB}
-	\Delta Q_{AB} = \bigg[\frac{L_{AB}}{L} - \bigg(\frac{k_{AB}}{2L} \bigg)^2 \bigg] - \bigg[\frac{L_A}{L} - \bigg(\frac{k_A}{2L} \bigg)^2 + \frac{L_B}{2L} - \bigg(\frac{k_B}{2L} \bigg)^2 \bigg] \ ,
+	\Delta Q_{AB} = \bigg[\frac{L_{AB}}{L} - \bigg(\frac{k_{AB}}{2L} \bigg)^2 \bigg] - \bigg[\frac{L_A}{L} - \bigg(\frac{k_A}{2L} \bigg)^2 + \frac{L_B}{L} - \bigg(\frac{k_B}{2L} \bigg)^2 \bigg] \ ,
 \end{equation}
 where $L_{AB} = L_A + L_B + l_{AB}$ is the total number of edges in the merged community; $L_A$ and $L_B$ are the number of edges in communities $A$ and $B$ respectively and $l_{AB}$ is the number of direct edges between communities $A$ and $B$ \cite{Albert2016}.
-$k_{AB} = k_A + k_B$ is the total degree of nodes in the merged community \cite{Albert2016}.
+$k_{AB} = k_A + k_B$ is the total degree of nodes in the merged community; $k_A$ and $k_B$ are the total degree of the nodes in community $A$ and $B$ respectively \cite{Albert2016}.
 Inserting the formulas for $L_{AB}$ and $k_{AB}$ in Eq.
@@
 \end{equation}
-where the right hand side is equal to Equation (\ref{avDegr}) with $n = N-1$ and we thus obtain $\left<k\right> = (N-1)p$.
+where the right hand side is equal to Eq. (\ref{avDegr}) with $n = N-1$ and we obtain $\left<k\right> = (N-1)p$.
 
@@
 \begin{equation}
-	\ln[(1-p)^{N-1-k}] \approx (N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right> \ ,
+	\ln[(1-p)^{N-1-k}] \approx -(N-1-k) \frac{\left<k\right>}{N-1} = - \left<k\right>\bigg(1 - \frac{\left<k\right>}{N-1} \bigg) \approx - \left<k\right> \ ,
 \end{equation}
@@
 It is then clear that the final, backwards step can span at most $K/2$ nodes \cite{Pele2015}; otherwise we would end up at a different node from which we started, which would definitely not form a triangle.\\
+\newline
 Hence, the number of triangles for a given node is given by choosing the two forward target nodes from the $K/2$ possibilities \cite{Pele2015}
@@
 (\ref{triangle}) and Eq.
-(\ref{triple}) with the number of nodes $N$.
+(\ref{triple}) by the number of nodes $N$.
 We then obtain the final result for the clustering coefficient in the regular ring lattice \cite{Pele2015}
 \begin{equation}
-	\left<cc\right> = \frac{3\cdot N \frac{1}{4} K \bigg(\frac{K}{2} - 1 \bigg)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)} \ .
+	\left<cc\right> = \frac{3\cdot N \frac{1}{4} K \big(\frac{K}{2} - 1 \big)}{N \frac{1}{2} K (K-1)} = \frac{3(K-2)}{4(K-1)} \ .
 \end{equation}
@@
 \begin{equation}
-	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ ,
+	\left<d\right> = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ ,
 \end{equation}
@@
 Assume we have a regular ring lattice with $N$ nodes and each node has $K$ neighbors.
-Now we can calculate the distance between node $i$ and every other node in the ring lattice
-\begin{itemize}
-	\item There are $K$ nodes adjacent to node $i$, so there are $K$ nodes with a distance $d_{ij} = 1$
-	\item There are $K$ nodes that are neighbors of the neighbors of node $i$, so there are $K$ nodes with a distance $d_{ij} = 2$
-	\item ...
-	\item There are $K$ nodes that have a distance $d_{ij} = \frac{N}{K}$ between them and node $i$; its easy to see that these are the nodes that are the furthest away from node $i$ 
-\end{itemize}  
+We can then calculate the distance between node $i$ and every other node in the ring lattice as follows: (i) There are $K$ nodes adjacent to node $i$, so there are $K$ nodes with a distance $d_{ij} = 1$; (ii) There are $K$ nodes that are neighbors of the neighbors of node $i$, so there are $K$ nodes with a distance $d_{ij} = 2$; (iii) ...; and (iv) There are $K$ nodes that have a distance $d_{ij} = \frac{N}{K}$ between them and node $i$; its easy to see that these are the nodes that are the furthest away from node $i$.\\
+\newline
 So the average distance for node $i$ becomes \cite{chap9}
 \begin{equation}
-	l_i = \frac{1}{N-1} K \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ ,
+	d_i = \frac{1}{N-1} K \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ ,
 \end{equation}
@@
 \begin{equation}
-	l = \frac{1}{N} \sum_i l_i = \frac{1}{N} N l_i \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ .
+	\left<d\right> = \frac{1}{N} \sum_i d_i = \frac{1}{N} N d_i \approx \frac{K}{N} \bigg(1 + 2 + 3 + ... + \frac{N}{K} \bigg) \ .
 \end{equation}
@@
 \end{equation}
-where the last approximation again holds for $x \gg 1$; we finally obtain \cite{chap9}
-\begin{equation}
-	l \approx \frac{K}{N} \frac{\frac{N^2}{K^2}}{2} = \frac{N}{2K} \ .
+where the last approximation holds for $x \gg 1$; we finally obtain \cite{chap9}
+\begin{equation}
+	\left<d\right> \approx \frac{K}{N} \frac{\frac{N^2}{K^2}}{2} = \frac{N}{2K} \ .
 \end{equation}

### 2021-05-16 draft.tex
wrote conclusion

@@
 
-\chapter{Conclusions and outlook}
+\chapter{Conclusion}
+
+Simple toy models were implemented to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
+Our goal was to get a better understanding of the impact of social media and algorithmic personalization on phenomena such as polarization and formation of echo chambers (opinion clusters) in society.
+\\
+\newline
+We found that topological correlations such as high clustering strongly enhance the formation of opinion clusters.
+Beside clustering, community structure also favors the formation of echo chambers.
+Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity.
+The heterogeneous contact pattern of these real-world networks is mostly likely the cause of this hampering.
+If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system.
+This is an intuitive result: if you shield people from opposing views and strengthen them in their own beliefs, it feels natural that the system will evolve towards a more polarized state.
+It is then expected that a more heterogeneous exposure to information reduces this polarization.
+We found that this is indeed the case if people are not stubborn.
+If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
+This is an important result, since it may change our beliefs/assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
+More research on this topic is definitely needed in the near future.
+Finally, we found that distributing the opinions in communities instead of randomly distributing them may be hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
+\newline
+Our results strongly suggest that the effects of algorithmic personalization do not stand on their own.
+Instead, the interplay with certain network topologies may enhance or hamper these effects.
+Beside that, individual resistance strongly affects the impact of these filtering algorithms.
+It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization.
+It is the interplay between many different aspects of the system (such as topological correlations, distribution of the opinions, individual resistance, content curration) that may potentially enhance or hamper the formation of opinion clusters.
 % also discuss need to investigate effect of changing N, S and p_act

### 2021-05-18 draft.tex
added results for PGP trust network

@@
 \newline
-Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloedt worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,...
+Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,...
 Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen.
@@
 \addcontentsline{toc}{chapter}{Table of Contents}
-\newgeometry{top=20mm, bottom=25mm}
+%\newgeometry{top=20mm, bottom=25mm}
 \listoffigures
@@
 \addcontentsline{toc}{chapter}{List of Tables}
-\restoregeometry
+%\restoregeometry
 
@@
 The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
-For each real-world network, the results will be compared to results obtained from network models with the same average degree, clustering and/or modularity.
+For each real-world network, the results will be compared to results obtained from network models with similar average degree, clustering and/or modularity.
 We will mainly focus on the formation of echo chambers.
@@
 It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.
-This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.
+This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.\\
+\newline 
 Another difference between the Last.fm and the SBM and SBM-WS network is the community structure.
@@
 The results for the Last.fm network coincide with those of the SBM.
-As for the PR filtering algorithm, this seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the corresponding hampering due to the lack of high clustering.
+This seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to the lack of clustering.
 
@@
 Results for the PGP trust network are compared to results for network models with similar average degree, clustering and/or modularity.
-The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.
+The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
+\newline
+In Fig.
+\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of completely stubborn actors (fracRes) is studied for the real-world PGP trust network and for the corresponding network models.
+This is done for both the PR and REC filtering algorithms.
+\\
+\newline
+The PR filtering algorithm (Fig.
+\ref{pr_real_PGP}) shows a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors for both the PGP trust network and the corresponding network models.
+The real-world PGP network has a significantly smaller formation of echo chambers compared to the network model with similar clustering and modularity (SBM-WS).
+This is most likely due to the heterogeneous degree pattern of the PGP network.
+\\
+The REC filtering algorithm (Fig.
+\ref{rec_real_PGP}) shows an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease if the fraction of stubborn actors is further increased.
+This is visible for both the PGP trust network and the network models.
+Again, the PGP network has a smaller formation of echo chambers than the network model with similar clustering and modularity (SBM-WS), due to its heterogeneous degree distribution.
+The data points for the SBM and the PGP trust network coincide, which may suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
+% compare powerlaw community distribution??
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{PR}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_PGP_8x7.png}
+    \label{pr_real_PGP}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{REC}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_PGP_8x7.png}
+    \label{rec_real_PGP}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
+The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+The PGP trust network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM-WS with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$; (iii) SBM with 60 communities of size 178 with $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$.
+\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
+Each data point is the average of $50$ independent simulations.
+Bars represent the standard deviation $\sigma$.}
+\label{echo_vs_fracRes_real_PGP}
+\end{figure}
 
@@
 It must be emphasized that there are limitations to our model and methods.
-As already stated before, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses.
+First, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses.
 Real-life social behavior and opinion dynamics on social media may be far from simple.
@@
 Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity.
-The heterogeneous contact pattern of these real-world networks is mostly likely the cause of this hampering.
+The heterogeneous contact pattern of these real-world networks is most likely the cause of this hampering.
 If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system.
@@
 More research on this topic is definitely needed in the near future.
-Finally, we found that distributing the opinions in communities instead of randomly distributing them may be hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
+Finally, we found that distributing the opinions in communities instead of randomly distributing them may hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
 \newline

### 2021-05-19 draft.tex
incorporated feedback ch. 1 and ch.2

@@
 This filtering and individualizing of information may have some potentially dangerous downsides in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
-It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
+It is important to know the effects of such algorithmic personalization on the formation of phenomena such as polarization, spread of extremism and formation of echo chambers.\\
 \newline
@@
 \newline
-In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
-We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers.
-We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness.
-Beside that, we will also pay some attention to the possible ways of distributing the opinions and how this may affect the opinion dynamics in the system.
-\\
-\newline
-Finally, we will test the model on real-world networks.
-Some question that may arise are: Do the results agree with those obtained from the theoretical network models?
-What are the similarities or differences in the results?
-What are the possible underlying reason/structures for these similarities/differences?
-% definetely not finished!! --> add some type of conclusions or not? Change part about distributing stubborn people??
+In this thesis toy models are designed to study the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
+We will investigate whether network properties such as community structure and clustering may enhance or hamper the formation of echo chambers.
+We will use filtering algorithms to determine their possible effects on the formation of echo chambers and we will study the impact of individual stubbornness.
+We will also study different ways of distributing the opinions and how this may affect the opinion dynamics in the system.
+\\
+\newline
+Finally, we will test the model on real-world networks. We will investigate whether the results agree with those obtained from the theoretical network models.
+
 \setcounter{abstractpage}{\value{page}}
@@
 \newline
-Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding,...
+Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding.
 Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen.
@@
 Een gebruiker op sociale media krijgt dus niet alle berichten die zijn/haar vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn/haar sociale media geschiedenis.
-Dit filteren/individualiseren kan, mogelijks, grote gevolgen hebben.
+Dit filteren en individualiseren kan, mogelijks, grote gevolgen hebben.
 Zo zou het kunnen leiden tot de vorming van ``echo kamers'' waar gebruikers gevangen raken tussen ideeën die hun eigen visie versterken.
@@
 Mensen zijn echter ook in staat om kritisch te denken, soms ook over-kritisch of selectief kritisch, maar dat terzijde.
-Dit kritisch denken kan ertoe leiden dat mensen niet zomaar hun mening veranderen omdat hun vrienden en/of omgeving dat suggereren.
+Dit kritisch denken kan ertoe leiden dat mensen niet zomaar hun mening veranderen omdat hun vrienden of omgeving dat suggereren.
 Ze zijn meer overtuigd van hun eigen gelijk of menen dat hun eigen visie een correcter standpunt inneemt.
@@
 In deze thesis wordt gebruik gemaakt van ``toy models'' die de interactie tussen verschillende netwerk modellen, filtering algoritmen en individuele terughoudenheid om van mening te veranderen onderzoeken in de evolutie van twee opinies.
-Er worden verschillende netwerk modellen geïntroduceerd die de impact van onder andere clustering en/of groepsstructuur op de evolutie van opinies en de vorming van echo kamers moeten onderzoeken.
+Er worden verschillende netwerk modellen geïntroduceerd die de impact van onder andere clustering en groepsstructuur op de evolutie van opinies en de vorming van echo kamers moeten onderzoeken.
 Daarnaast wordt er gebruik gemaakt van enkel simpele filtering algoritmen om hun impact op de dynamica van de opinies te onderzoeken.
@@
 Naast het bovenstaande zal er ook kort aandacht gespendeerd worden aan hoe de twee verschillende opinies verdeeld worden over het systeem.
-Is er bijvoorbeeld een verschil tussen de twee meningen willekeurig over het systeem te verspreiden versus deze in een meer groepsstructuur te verdelen?
+We zullen de opinies in groepsstructuren verdelen en dit scenario vergelijken met het scenario waarin de opinies willekeurig verdeeld worden.
 Om de thesis te besluiten zal het opinie-dynamica model, dat in de thesis gebruikt werd, getest worden op netwerken verkregen van data uit de werkelijke wereld (bv.
@@
 
-% ADD MORE REFERENCES (previous research, contextualization etc)
+% ADD MORE REFERENCES (previous research, contextualization)
 
@@
 Nowadays most of us have, beside a real-world life, a virtual life on social media.
-This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale.
+This increased engagement on social media may have an impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale.
 \\
@@
 People are, however, still bound by temporal and cognitive constraints.
-In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}.
+In order to assure a more pleasant and convenient experience on their platforms, social media companies use algorithms to order and filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}.
 It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
@@
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\
-\newline
+On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate models to reproduce the opinion dynamics and opinion formation processes found in real life.\\
+\newline
+Section \ref{statPhys} introduces concepts such as statistical and social physics.
 Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks.
-The second layer, the opinion dynamics layer, is introduced in Section \ref{Opinion}.
-Section \ref{statPhys} introduces concepts such as statistical and social physics.
+The second layer, the opinion dynamics, is introduced in Section \ref{Opinion}.
 Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are summarized and discussed.
 
+\section{Statistical and social physics}\label{statPhys}
+
+An important tool in the study of opinion dynamics in large groups of people is the use of statistical physics.
+This branch of physics offers a framework that relates microscopic properties of atoms and molecules to macroscopic observed behavior and has a wide range of applications inside and outside physics.
+The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
+The application of the theory of statistical physics on social phenomena is referred to as social physics \cite{Castellano2009}\cite{Galam2008}\cite{Galam1982}\cite{Stauffer2012}.
+The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies.
+These `macroscopic' groups display regularities, transitions from disorder to order, the emergence of consensus, universality \cite{Buchanan2007}.
+The statistical physics approach of these social systems tries to explain the regularities at large scale as collective effects of the interactions among individuals \cite{Sirbu2016}.\\
+\newline
+The statistical approach of social phenomena has one difficulty compared to the description of gases and other physical phenomena.
+In physics, the microscopic constituents are atoms or molecules, whose dynamics and interactions with each other are often well known and can be put in relatively simple laws with only a few parameters \cite{Castellano2009}.
+Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}.
+Even if the dynamics of individuals would be well know, the interactions between them would still be complicated and it would be hard to describe them with simple laws and only a few parameters \cite{Castellano2009}.
+Hence, extracting useful and qualitative information out of social models is a challenging task \cite{Castellano2009}.
+The use of statistical physics may, however, offer a way out.
+In statistical physics it is found that different systems behave qualitatively the same and are governed by the same laws near the critical point \cite{Hu2018}\cite{Kadanoff2010}.
+This lead to the concept of universality: many different microscopic problems have the same macroscopic manifestation \cite{Gug2015}\cite{Kadanoff2015}.
+This idea may then be extended to social systems \cite{Hu2018}: we should not take the unique character of every individual into account, instead these individual characteristics can be incorporated in a sort of averaged way \cite{Gug2015}.
+One can then proceed to build models of social behavior where only the most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
+An important factor herein is the comparison with empirical data to investigate whether the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}\cite{Pawel}.
+This interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in physical research and science in general.
+\newpage
+\noindent
 \section{Complex networks}\label{complNet}
@@
 Complex networks have become more and more popular for analyzing complex, dynamical systems.
-They are used in fields such as physics, economics, social sciences, biology, etc \cite{Costa2008}.
-These different fields are very diverge, but they have at least one common ground: they often deal with a large number of variables/components that interact with each other.
+They are used in fields such as physics, economics, social sciences and biology \cite{Costa2008}.
+These different fields are diverse, but they have at least one common ground: they often deal with a large number of components that interact with each other.
 In other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of the individual components alone \cite{Mata2020}.
-Complex systems often display phenomena such as non-linearity, emergence, spontaneous order,...
+Complex systems often display phenomena such as non-linearity, emergence, spontaneous order.
 \\
@@
 One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
-A network, also often called a graph (both terms will be used in this thesis), is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
-\newline
-Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems.
+A network is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis nodes will mostly be called actors; edges will be called edges) \cite{Mata2020}.\\
+\newline
+Representing a complex system as a network makes the system appear simpler and tractable, while it still includes the non-linearity of these systems.
 One can find the language to describe networks in mathematical graph theory.
-However, complex systems in real-life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
+However, complex systems in real-life often deal with a large number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
 One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\
@@
 \newline
-Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc.
-These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks.
+Some other important concepts in network theory are the degree distribution, clustering, community structure and connectivity.
+These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate on the case of social networks.
 It is found that many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale-free degree distribution represents one of the three general properties of social networks.
-The other two are short distances and high clustering.
-These two are also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
+The other two are short distances and high clustering, also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 \newline
-Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model.
+The theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model.
 These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
 
-% talk about difference between simple networks and complex network??
-
-% talk about using real-world networks to compare against network models??
-
 \section{Opinion dynamics}\label{Opinion}
 
-The formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg.
-stubbornness), the information someone is exposed to, etc.
+The formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (e.g.
+stubbornness) and the information someone is exposed to.
 Opinion dynamics models should try to capture these complex processes in a simplified way.
@@
 \newline
-The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
+The basic idea of all opinion dynamics models is that the nodes or actors in a social network have a variable that represents their opinion and that is updated according to some predefined rules.
 These models are a simplification of real-world opinion dynamics.
-It is however shown that they display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}.
-\\
-\newline
-The opinion of the agents in the model can be either discrete or continuous.
-This thesis deals with models where each agent can have one of the two opinions, A or B, and thus only deals with the case of discrete opinions.
-This might come across as a huge simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg.
-republicans or democrats, cat or dog, renting or buying a house,...).
-\\
-\newline
-The rules that determine how an agent updates his or her opinion can include many different aspects.
+It is however shown that they display aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization and formation of echo chambers (clusters of people with the same opinion) \cite{Sirbu2016}.
+\\
+\newline
+The opinion of the actors in the model can be either discrete or continuous.
+This thesis deals with models where each actor can have one of the two opinions, A or B, and thus only deals with the case of discrete opinions.
+This might come across as a simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (e.g.
+republicans or democrats, renting or buying a house).
+\\
+\newline
+The rules that determine how an actor updates his or her opinion can include many different aspects.
 They can, for example, be as simple as a majority rule (i.e.
-the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating.
-A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}.
-\\
-\newline
-One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others.
-Furthermore, the agents can have a parameter that determines whether they are active or not.
-The introduction of stubborn actors will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
-\newline
-Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms.
-The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
+the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a probabilistic way of updating.
+An overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}.
+\\
+\newline
+One can also include a resistance parameter that determines the hesitation of an actor to change to a new opinion or a parameter that determines the influence of an actor on others.
+Furthermore, the actors can have a parameter that determines whether they are active or not.
+The introduction of stubborn actors will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of actors is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
+\newline
+Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define different filtering algorithms.
+The filtering algorithms that social media companies use, are corporate secrets, but they typically have three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
 Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
@@
 
-\section{Statistical and social physics}\label{statPhys}
-
-Another important tool in the study of opinion dynamics/formations in large groups of people is the use of statistical physics.
-This branch of physics offers a framework that relates microscopic properties of atoms, molecules,...
-to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.
-The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
-The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics.
-The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies.
-These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc \cite{Buchanan2007}.
-The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.\\
-\newline
-The statistical approach of social phenomena has one big difficulty compared to the description of gases and other physical phenomena.
-In regular physics, the microscopic constituents are `simple' atoms or molecules, whose dynamics and interactions with each other are often well known and can be put in relatively simple laws with only a few parameters \cite{Castellano2009}.
-Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}.
-Even if the dynamics/behavior of individuals would be well know, the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}.
-Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}.
-The use of statistical physics may, however, also offer a way out.
-In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}\cite{Gug2015}.
-Instead, the relevant aspects for the global behavior are often higher level features such as symmetries, conservation laws, dimensionality, etc \cite{Castellano2009}.
-This introduces a concept of universality \cite{Gug2015}.
-With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
-An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}\cite{Pawel}.
-This interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.
-
-% talk somewhere about rise of social physics, the need to use statistical physics (because of macroscopic phenomena as scale invariance, emergence, polarization,...) --> partly done?
-
 \section{This thesis: hypotheses and goals}\label{goal}
 
-This thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
-On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn actors in the network.
-Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\
-\newline
-In one of their previous papers (\textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, \cite{Perra2019}) Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions.
-They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).
-Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}.
+This thesis will study the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
+On one hand, we will determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we will study the effect of stubborn actors in the network.
+Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms will be included.\\
+\newline
+Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions \cite{Perra2019}.
+They also considered the effect of nudging (meaning that one opinion is pushed to all actors in the network).
+Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering) could result in the formation of echo chambers.
 On the other hand, a more heterogeneous contact pattern hampers this formation of echo chambers \cite{Perra2019}.
-If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
+If the two opinions were, initially, not equally distributed, one of the filtering algorithms (semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}.
 In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.
-They, however, did not investigate the effect of stubborn actors in the network, nor did they investigate networks with a community structure.
+They, however, did not study the effect of stubborn actors in the network, nor did they investigate networks with community structure.
 This thesis will build on their model and incorporate these new features.
-We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
-We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient.
-We will also pay some attention to networks that have both a high clustering coefficient and community structure.
-Can we observe echo chambers in those kinds of networks?
-If yes, can we determine which property might be the main driver for the formation of echo chambers: high clustering or community structure?
-\\
-\newline
-On the other hand, we want to determine the impact of stubborn actors on the formation and evolution of opinions in the network.
-Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness?
-What about the possible interplay between the filtering algorithms and the introduction of stubborn actors?\\
-% partly rewrite this according to results? Convergence to one opinion in communities --> discussed? Clustered stubborn people --> discussed? Include clustered opinions? Include real-world networks?
+We will compare networks that have community structure and relatively low clustering to networks without community structure, but with high clustering (a detailed description of concepts like clustering and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
+We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with high clustering.
+We will also study networks that have both high clustering and community structure to identify the conditions and driving forces leading to the formation of echo chambers.\\
+\newline
+We will also study the impact of stubborn actors on the formation and evolution of opinions in the network.
+We aim to determine whether there is an interplay between network structure, filtering algorithm and stubbornness.
 
@@
 
-Networks or graphs are conceptually simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
+Networks are conceptually simple and flexible objects that are made of a set of nodes $V$ and a set of edges $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
 \begin{equation}
@@
 \end{equation}
-A graph with no self-loops is a simple graph.
+A network with no self-loops is a simple network.
 The nodes $x$ and $y$ of an edge $\{x, y\}$ are called the endpoints of the edge.
 It is possible that nodes are not joined by any edge, such nodes are called isolated or disconnected.
-If two or more edges have the same pair of endpoints, the graph is called a multi-graph.
-This thesis is not concerned with multi-graphs, nor are self-loops allowed.
-Graphs in which the edges have an orientation are called directed graphs \cite{Costa2018}.
-If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
-\newline
-A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.
+If two or more edges have the same pair of endpoints, the network is called a multi-network.
+This thesis is not concerned with multi-networks, nor are self-loops allowed.
+Networks in which the edges have an orientation are called directed networks \cite{Costa2018}.
+If the edges have weights $w_{ij}$, one speaks of a weighted network.\\
+\newline
+A network can be represented in different ways. One of the most common ways is by use of the adjacency matrix.
 
@@
 
-The adjacency matrix \textbf{A} is a $N \times N$ matrix (for a graph with $N$ nodes) where the elements indicate whether two nodes are connected by an edge or not \cite{Mata2020}
+The adjacency matrix \textbf{A} is a $N \times N$ matrix (for a network with $N$ nodes) where the elements indicate whether two nodes are connected by an edge \cite{Mata2020}
 \begin{equation}
@@
 \end{equation}
-For a simple, undirected graph, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$).
-Directed graphs can have asymmetric adjacency matrices.
-If we are dealing with weighted graphs, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
+For a simple, undirected network, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$).
+Directed networks can have asymmetric adjacency matrices.
+In weighted networks, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
 \begin{equation}
@@
 
-The adjacency matrix contains a lot of information such as the degree of a node.
-The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$.
-The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
+The adjacency matrix contains information such as the degree of a node.
+The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ and represents the number of nearest neighbors of node $i$.
+The degree can be obtained from the adjacency matrix \cite{Mata2020}
 \begin{equation}
@@
 \end{equation}
-In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$.
+In the case of a directed network one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$.
 These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
@@
 \end{align}
-The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$.
-For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
+The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$. For weighted networks the degree is generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
 \begin{equation}
@@
 \end{equation}
-From the degree of the nodes in the network one can construct the degree distribution.
-The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$ \cite{Newman2003}.
-The average degree can be obtained from the degree distribution \cite{Mata2020}
+The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$.
+The degree distribution gives the probability that a randomly selected node has a degree $k$ \cite{Newman2003}.
+The average degree can be obtained from the degree distribution
 \begin{equation}
@@
 \end{equation}
-where $N$ is the total number of nodes in the network.
+where $N$ is the total number of nodes in the network \cite{Mata2020}.
 The degree distribution allows us to classify networks.
 The two most important classes are homogeneous and heterogeneous networks.
-Homogeneous networks have a bell curved degree distribution, eg.
+Homogeneous networks have a bell curved degree distribution, e.g.
 a Poisson distribution.
@@
 Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$.
-They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree is not a characteristic scale for the network \cite{Barabasi2016}\cite{Mata2020}.
-
-\subsection{Connectivity}
-
-In physics literature, the degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
-This is different from the mathematical/graph definition of connectivity, where connectivity refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subgraphs.\\
-\newline
-Two nodes $u$ and $v$ of a graph are connected if there exist a path between them. If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
+They are also referred to as scale-free networks, since they do not posses a characteristic length scale, i.e.
+the average degree is not a characteristic scale for the network \cite{Barabasi2016}\cite{Mata2020}.
+
+\subsection{Connectivity}\label{connect}
+
+The degree of a node is often called the connectivity \cite{F.Costa2007}\cite{Mendes2002}.
+It also refers to the minimum number of nodes or edges that need to be removed to separate the remaining nodes in isolated subnetworks.\\
+\newline
+Two nodes $u$ and $v$ of a network are connected if there is a path between them. If the two nodes are connected by a path that contains only one edge, the two nodes are called adjacent.\\
 \newline
 The average degree $\left<k\right>$ is a global measurement of the connectivity of the network \cite{Costa2008}.
-If the average connectivity $\left<k\right>$ is too small (this is, if there are too few edges), there will be many isolated nodes and only a few clusters with a small number of nodes.
+If the average connectivity $\left<k\right>$ is too small (that is, if there are too few edges), there will be many isolated nodes and only a few clusters with a small number of nodes.
 If more and more edges are added to the network, the small clusters will grow and will tend to connect to each other to form larger clusters.
@@
 
-Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks.
-More precisely, the notion of assortativity and disassortativity has to do with degree correlations.
-\\
+Another concept that is closely related to the degree of a node is assortative and disassortative networks. The notion of assortativity and disassortativity has to do with degree correlations. \\
 \newline
 The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}.
-The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
-\begin{equation}
-	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)} \ .
-\end{equation}
+The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$
+\begin{equation}
+	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)} \ ,
+\end{equation}
+where a neighbor or nearest neighbor of a node $i$ is an adjacent node of node $i$, the meaning of adjacent nodes is given in Section \ref{connect} \cite{F.Costa2007}.
 The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
@@
 \newline
-The terms assortativity and disassortativity are also often used without taking degree correlations into account.
-Instead, one may look at a particular node property to classify networks as either assortative or disassortative \cite{Thed2014}.
+Assortativity and disassortativity are most often applied to the degree of the nodes in the network, which is referred to as degree assortativity \cite{Noldus2015}.
+Assortativity may, however, also be applied to other characteristics of the nodes \cite{Noldus2015}\cite{Thed2014}.
 In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.
@@
 
-Several measurements quantify the notion of length and distance in a network/graph.
-We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
+Several measurements quantify the notion of length and distance in a network.
+A path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
 The length of a path is defined as the number of edges in the path \cite{Goddard2010}.
@@
 \newline
-The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
+The average distance is the mean value of the distance $d_{ij}$ \cite{F.Costa2007}
 \begin{equation}\label{avdist}
@@
 The clustering coefficient is another important measure of network topology.
-It is a measure for the number of triangles in a network and determines the connectivity in the neighborhood of a node $i$: if a node $i$ has a high clustering coefficient, its neighbors are likely to be directly connected to each other \cite{Li2017}.
+It measures the number of triangles in a network and determines the connectivity in the neighborhood of a node $i$: if a node $i$ has high clustering, its neighbors are connected to each other \cite{Li2017}.
 \\
@@
 The clustering coefficient is thus a way to measure the degree to which nodes in a network tend to cluster \cite{Li2017}.
-The local clustering coefficient of a node $i$ is defined as \cite{Mata2020}
+The local clustering coefficient of a node $i$ is defined as
 \begin{equation}\label{clus}
-	cc_i = \frac{2n_i}{k_i(k_i-1)} \ ,
-\end{equation}
-where $n_i$ is the number of edges that actually exist between the nodes in the neighborhood of $i$ and $k_i(k_i-1)/2$ is the maximum number of edges that could exist between them (this expression is only valid for an undirected graph, since for a directed graph $e_{ij} \neq e_{ji}$ and we have $k_i(k_i-1)$ possible edges between the neighbors of node i).
+	cc_i = \frac{2e_i}{k_i(k_i-1)} \ ,
+\end{equation}
+where $e_i$ is the number of edges that actually exist between the nodes in the neighborhood of $i$ and $k_i(k_i-1)/2$ is the maximum number of edges that could exist between them \cite{Mata2020} (this expression is only valid for an undirected network, since for a directed network $e_{ij} \neq e_{ji}$ (where $e_{ij}$ denotes an edge from node $i$ to node $j$) and we have $k_i(k_i-1)$ possible edges between the neighbors of node i).
 % should give more mathimatical expression for n_i?
-The average clustering coefficient is then given as the average of the local clustering coefficients \cite{Newman2003}
+The average clustering coefficient is 
 \begin{equation}\label{avClus}
@@
 \end{equation}
-where $N$ denotes the number of nodes in the network.
-The clustering coefficient is, beside a measure for the connectivity in the network, also linked to the robustness, or resilience against random damage, of the network \cite{Heer2020}\cite{Iyer2013}\cite{Li2017}.
+where $N$ denotes the number of nodes in the network \cite{Newman2003}.
+The clustering coefficient is linked to the robustness, or resilience against random damage, of the network \cite{Heer2020}\cite{Iyer2013}\cite{Li2017}.
 \\
@@
 \\
-\newpage
-\noindent
+\newline
 Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix \cite{F.Costa2007}
@@
 \end{align}
+\newpage
+\noindent
 where $A_{ij}$ are the elements of the adjacency matrix. It is possible to define the transitivity of one node as \cite{F.Costa2007}
@@
 \end{align}
-It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
-It is then obvious that the local clustering coefficient $cc_i$ (Eq.
+$N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$.
+The local clustering coefficient $cc_i$ (Eq.
 (\ref{clus})) of a node and the transitivity $T_i$ (Eq.
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
-In the case of the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$ this is, however, not the case.
+This however not the case for the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$.
 The difference between the two definitions is that the average of Eq.
@@
 
-Network communities are another feature that (complex) networks may posses.
+Network communities are another clustering structure of networks.
 A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}.
 More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}.
-In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}.
+The connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}.
 The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}.
@@
 Let us consider three possible definitions: maximum cliques, strong communities and weak communities \cite{Albert2016}.\\
+\newline
+\textbf{Maximum cliques}\\
+\newline
+A clique is a complete subnetwork, where a complete subnetwork is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}.
+A community based on this definition would then be the largest clique in the network.
+This definition of community might, however, be too restrictive and large cliques do not appear often in networks \cite{Albert2016}.\\
 \newpage
 \noindent
-\textbf{Maximum cliques}\\
-\newline
-A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}.
-A community based on this definition would then be the largest clique in the network.
-This definition of community might, however, be too restrictive and, beside that, large cliques do not appear very frequent in networks \cite{Albert2016}.\\
-\newline
 \textbf{Strong communities}\\
@@
 \end{equation}
-where the factor $\frac{1}{2L}$ is a normalization factor. Using \cite{Albert2016}
-\begin{equation}
-	p_{ij} = \frac{k_ik_j}{2L} \ ,
-\end{equation}
-we can derive a simpler form for $Q_c$ \cite{Albert2016}
+where the factor $\frac{1}{2L}$ is a normalization factor.
+If we randomly wire the network, using the degree preserving null model, we find the probability that nodes with degree $k_i$ and degree $k_j$ are connected to each other \cite{Bara2016}
+\begin{equation}\label{pij}
+	p_{ij} = \frac{k_ik_j}{2L} \ .
+\end{equation}
+This equation implies that hubs, due to their large number of edges, are more likely to connect to other hubs than to small degree nodes \cite{Bara2016}.
+Furthermore, Eq.
+(\ref{pij}) serves as a benchmark to define degree correlations as assortativity and disassortativity.
+A network displays degree correlations if the number of edges between high and low degree nodes systematically deviates from what is expected by chance (given by Eq.
+(\ref{pij})) \cite{Bara2016}.
+\\
+\newline
+Using Eq. (\ref{pij}), we can derive a simpler form for $Q_c$
 \begin{equation}\label{Qc}
@@
 \end{equation}
-where $L_c$ is the total number of edges in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.\\
-\newline
-To generalize this idea to the whole network, we sum Eq. (\ref{Qc}) over all $n_c$ communities \cite{Albert2016}
+where $L_c$ is the total number of edges in the community $C_c$ and $k_c$ is the total degree of the nodes in this community \cite{Albert2016}.
+To generalize this idea to the whole network, we sum Eq.
+(\ref{Qc}) over all $n_c$ communities \cite{Albert2016}
 \begin{equation}\label{mod}
@@
 Modularity is often used in community detection algorithms.
-It is thus important to be aware of the limitations of modularity.
-One of the biggest shortcomings of modularity maximization as community detection is its resolution limit.
+One of the shortcomings of modularity maximization as community detection is its resolution limit.
 Modularity maximization forces small communities into larger ones \cite{Albert2016}.
@@
 \end{equation}
-and if there is at least one edge between the two communities; even if $A$ and $B$ are distinct communities \cite{Albert2016}!
+and if there is at least one edge between the two communities; even if $A$ and $B$ are distinct communities \cite{Albert2016}.
 Eq.
@@
 These theoretical models usually have a simpler representation and well known properties that can be derived analytically.
-They are widely studied and used extensively.
-The study of complex networks relies heavily on the knowledge and understanding of these models.
+They are widely studied and used.
+The study of complex networks relies on the knowledge and understanding of these models.
 In this section we will review some of the most important ones.
@@
 \end{equation}
-The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. \\
+The derivation can be found in Appendix \ref{ER}, Section \ref{poiss}. The Erd\H{o}s-R\'{e}nyi network thus has a homogeneous, bell shaped, degree distribution.\\
 \newline
@@
 \end{equation}
-This can be easily understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$.
+This can be understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$.
 The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates).
@@
 \end{equation}
-Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network.
-Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the ``small world phenomenon''; the other property is a high clustering coefficient \cite{Easley2010}.
-\\
-\newline
-To summarize: the random Erd\H{o}s-R\'{e}nyi network has a bell shaped degree distribution which approximates a Poisson distribution for $N \gg 1$; the average clustering coefficient becomes small for large networks and the average distance depends logarithmic on the system sizes.
-The Erd\H{o}s-R\'{e}nyi network thus possesses one of the two properties of the ``small world phenomenon'', a short average distance.\\
-\newline
-Real social networks, however, usually display the ``small world phenomenon'' and thus have both a short average distance and a high clustering coefficient.
+Since for large $N$, $\ln{N} \ll N$, we find that the average distance is smaller than the size of the network.
+Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of the ``small world phenomenon''; the other property is high clustering \cite{Easley2010}.
+
+\subsection{The Watts-Strogatz model}
+
+Erd\H{o}s-R\'{e}nyi networks have small average distances and thus posses one of the two defining properties of the ``small world phenomenon''.
+Real social networks, however, usually display the ``small world phenomenon'' and thus have both short average distances and high clustering.
 The question then rises: Can we construct a model that has both of these properties?
-The answer is yes and this is exactly what Duncan Watts and Steve Strogatz did.
-
-\subsection{The Watts-Strogatz model}
-
-In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances.
-Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}.
-The construction goes as follows \cite{Watts1998}: (i) Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side); (ii) For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
+The answer is yes and this is what Duncan Watts and Steve Strogatz did in 1998.
+They constructed a network model that generates random networks with small world properties by interpolating between a regular ring lattice and a random network \cite{Watts1998}.\\
+\newline
+The construction of their model goes as follows \cite{Watts1998}: (i) Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side); (ii) For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$.
 Rewiring is done in such a way that self-loops and edge duplication are avoided.\\
 \newline
-The probability $\beta$ allows us to tune the graph between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}.
+The probability $\beta$ allows us to tune the network between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}.
 Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$.
-The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
+The first inequality assures sparse networks, whereas the second inequality guarantees that a random network will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
 \newline
@@
 \end{align}
-Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random graph (which is small), even for small $\beta$ \cite{Newman2000}.
+Watts and Strogatz showed by numerical simulations that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random network (which is small), even for small $\beta$ \cite{Newman2000}.
 The clustering coefficient is approximately given by \cite{Barrat1999}
@@
 \newline
-For rather small values of $\beta$, the clustering coefficient remains high.
+For small values of $\beta$, the clustering coefficient remains high.
 The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
@@
 
-The stochastic block model (SBM) is a random graph model that is able to produce networks containing a community structure.
+The stochastic block model (SBM) is a random network model that is able to produce networks containing community structure.
 The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
 The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected with each other than with the rest of the network \cite{F.Costa2007}.
-Hence, these networks posses a modular/community structure and this is exactly the structure that is easily modeled by the stochastic block model.
+Hence, these networks posses a modular structure and this is exactly the structure that is easily modeled by the stochastic block model.
 The stochastic block model is also often used as a benchmark for community detection algorithms \cite{Abbe2018}.\\
 \newline
-The stochastic block model in its most simple form is defined by the following parameters \cite{Clauset2017}: (i) The number of nodes $N$; (ii) A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities; and (iii) A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.\\
-\newline
-There are no real restrictions on the edge probabilities, which makes the stochastic block model a general model that is capable of reproducing many different network structures \cite{Karrer2011}.
-However, the edge probabilities inside a community are often taken to be bigger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
+The stochastic block model in its simplest form is defined by the following parameters \cite{Clauset2017}: (i) The number of nodes $N$; (ii) A partition of the node set $\{1, 2, ..., N\}$ into $k$ disjoint subsets $C_1, ..., C_k$; these groups are the communities; and (iii) A symmetric $k\times k$ matrix $M$; the diagonal elements $M_{ii}$, $i = 1, ..., k$, represent the edge probabilities inside the community $C_i$ and the off-diagonal elements $M_{ij}$, $i, j = 1, ..., k$ with $i \neq j$, give the edge probabilities between nodes in community $C_i$ and nodes in community $C_j$.\\
+\newline
+There are no restrictions on the edge probabilities, which makes the stochastic block model a general model capable of reproducing many different network structures \cite{Karrer2011}.
+However, the edge probabilities inside a community are often larger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
 In this way, the stochastic block model generates a network with community structure.
@@
 Again, we are left with the parameters $N$ and $p$ of the Erd\H{o}s-R\'{e}nyi model.
-In both cases, the mathematical properties of the Erd\H{o}s-R\'{e}nyi model are obtained: a Poisson degree distribution, an average distance that is logarithmic with the system size $N$, vanishing clustering coefficient for large $N$, etc \cite{Clauset2017}.
-The structure within the groups or communities of the stochastic block model is that of a random graph, while the structure between the communities is that of a random bipartite graph \cite{Clauset2017}.\\
+In both cases, the mathematical properties of the Erd\H{o}s-R\'{e}nyi model are obtained: a Poisson degree distribution, an average distance that is logarithmic with the system size $N$ and vanishing clustering coefficient for large $N$ \cite{Clauset2017}.
+The structure within the groups or communities of the stochastic block model is a random network, while the structure between the communities is a random bipartite network \cite{Clauset2017}.\\
 \newline
@@
 The stochastic block model is a flexible model and there are many different extensions.
-One extension that is used in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
+One extension that is introduced in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
 In this way, we obtain a version of the stochastic block model that has both a community structure and a high clustering coefficient.\\
 \newline
-The construction of the model is very similar to that of the regular stochastic block model.
+The construction of the model is similar to that of the regular stochastic block model.
 For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $\left<k\right>$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
 The result is a network that can be tuned to have both a community structure and a high clustering coefficient.
-This is different from the regular stochastic block model that can only account for a community structure and lacks a high clustering.
+This is different from the regular stochastic block model that can only account for a community structure and lacks high clustering.
 % talk about community sizes from power law??
@@
 \section{Real-world social networks}
-The models discussed in the sections above, describe real-world networks in an idealized, simplified way.
+
+Many natural and man-made systems can be described with the help of complex networks.
+Complex networks are suited to represent almost any discrete system \cite{Costa2008}.
+Social networks are applications of complex networks to social systems.
+In social networks the nodes usually represent an individual and edges can e.g.
+represent friendship ties between these individuals; such a network is called a friendship network.
+Other examples of social networks are sexual networks, in which the edges denote sexual contacts between the individuals and email networks, in which directed edges represent an email send from individual $i$ to individual $j$ \cite{Costa2008}.
+There are also plenty of other real-world network such as collaboration networks, where the nodes are usually scientists and edges indicate professional ties between them; citation networks, where nodes represent scientific papers and the directed edges report citations and protein-protein interaction networks, in which nodes stand for proteins and edges indicate physical interactions between them \cite{Costa2008}.
+These examples of real-world networks are only a handful.
+A comprehensive review can be found in \cite{Costa2008}.\\
+\newline
+The models discussed in the previous sections are idealized and simplified.
 They may posses some of the characteristics of real-world networks, but they certainly do not always describe real-world networks appropriately.
-The question then arises: What are the main properties/characteristics of real-world networks, and more specifically of real-world social networks?
+The question then arises: What are the main characteristics of real-world networks, and more specifically of real-world social networks?
 Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
@@
 In real-world social networks, one often observes the pattern that two friends of an individual are also friends of each other \cite{RealWorld}.
-In graph terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
+In network terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
 This property is not only found in social networks, but also in a wide variety of different real-world networks \cite{RealWorld}.\\
@@
 The term scale-free refers to the fact that there is no characteristic scale in the network.
-Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs).
+Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a high degree (these nodes are referred to as hubs).
 If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges \cite{Newman2005}.
@@
 And even if the degree distribution is heterogeneous or fat-tailed, it does not always need to follow a power-law.
-The degree distribution can eg.
-also be a power-law with an exponential cut-off, a log-normal distribution, a stretched exponential, etc \cite{Newman2005}.\\
+The degree distribution can e.g.
+also be a power-law with an exponential cut-off, a log-normal distribution or a stretched exponential \cite{Newman2005}.\\
 \newline
@@
 Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}.
-They also often posses a modular/community structure \cite{Ferrara2012}\cite{McGlohon2011}.
+They also often posses a modular structure \cite{Ferrara2012}\cite{McGlohon2011}.
 \\
@@
 There are, however, some properties that are different between on-line and off-line social networks.
-One of them might be the degree correlations/mixing pattern.
-It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern (this distinguishes them from other real-world networks, eg.
+One of them might be the degree correlations or mixing pattern.
+It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern (this distinguishes them from other real-world networks, e.g.
 biological networks \cite{F.Costa2007}, which often have a disassortative pattern) \cite{Hu2009}\cite{Zhang2014}.
@@
 The `ordinary people' do not have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
-In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance/influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
+In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance and influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
 This leads to a disassortative mixing pattern.
@@
 The study of opinion dynamics in a group of people, as mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
-The network layer, which describes the underlying structure and correlations between the agents/people, is intensively described in the previous chapter, Chapter \ref{chap2}.
+The network layer, which describes the underlying structure and correlations between the actors, is intensively described in the previous chapter, Chapter \ref{chap2}.
 The current chapter focuses on the second layer, that of the opinion dynamics models.
@@
 \newline
-Some major attention will be given to the content curation/filtering algorithms used by social media companies (Section \ref{filter}).
+We will study models for the filtering algorithms used by social media companies (Section \ref{filter}).
 This is needed because this thesis deals with opinion dynamics on on-line social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
@@
 We will also discuss the activation mechanism (Section \ref{actMech}).
-This is the mechanism that describes the timings in which content/information is exchanged between users \cite{Perra2019}.
-Such a mechanism changes the network from a static one to a temporal one, where nodes become active and inactive over time.
+This is the mechanism that describes the timings in which content or information is exchanged between users \cite{Perra2019}.
+Such a mechanism changes the network from a static one to a temporal one, where actors become active and inactive over time.
 \\
@@
 Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.
-There are two major groups of opinion dynamics models, namely binary/discrete models and continuous approaches.
+There are two major groups of opinion dynamics models, namely binary or discrete models and continuous approaches.
 These will be discussed in Section \ref{Bin} and Section \ref{Con}, respectively.
@@
 
-Binary models are models where each agent can have one of the two opinions A or B.
+Binary models are models where each actor can have one of the two opinions A or B.
 A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
@@
 One of the simplest opinion dynamics models is the binary majority model.
-Each agent has one of the two opinions A or B.
+Each actor has one of the two opinions A or B.
 Initially, these opinions can be equally distributed or not.
-If the distribution is not equal, but instead eg.
+If the distribution is not equal, but instead e.g.
 $80/20$, there is a predominant opinion and a minority opinion.
-The following updating rule is imposed on the network: each agent/individual adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
-Initially each agent is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
+The following updating rule is imposed on the network: each actor adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
+Initially each actor is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
 \newline
@@
 \newline
-Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the agents adopt the same opinion \cite{Nguyen2020}.
+Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the actors adopt the same opinion \cite{Nguyen2020}.
 This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini et al.
@@
 Another simple binary model is the voter model.
-The model works as follows: each agent in the network has a binary opinion (A or B); at each time step, an agent is randomly selected along with one of its neighbors and, subsequently, the agent adopts the opinion of this neighbor \cite{Castellano2009}.
-The agent thus mimics its neighbor \cite{Castellano2009}.
+The model works as follows: each actor in the network has a binary opinion (A or B); at each time step, an actor is randomly selected along with one of its neighbors and, subsequently, the actor adopts the opinion of this neighbor \cite{Castellano2009}.
+The actor thus mimics its neighbor \cite{Castellano2009}.
 \\
@@
 
-In continuous opinion dynamics models, the opinion variable of the agents in the network is no longer discrete, in contrast to binary or discrete models, but takes real values (often in the interval $\interval{0}{1}$ or $\interval{-1}{1}$).\\
-\newline
-Some concepts that are valid for the discrete models no longer apply (eg.
-the concept of a majority/minority opinion or the concept of equality of opinions); hence, they require a bit of a different framework \cite{Castellano2009}.
-\\
-\newline
-The basis of most continuous models is a set of $N$ nodes or agents, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
-In contrast to discrete/binary models, all agents usually start with different opinions \cite{Castellano2009}.
-The possible outcomes are more complex than in the discrete models and often have opinion clusters emerging in the final states (where opinion clusters refer to clusters of nodes with the same opinion) \cite{Castellano2009}.
+In continuous opinion dynamics models, the opinion variable of the actors in the network is no longer discrete, in contrast to binary or discrete models, but takes real values (often in the interval $\interval{0}{1}$ or $\interval{-1}{1}$).\\
+\newline
+Some concepts that are valid for the discrete models no longer apply (e.g.
+the concept of a majority and minority opinion or the concept of equality of opinions); hence, they require a bit of a different framework \cite{Castellano2009}.
+\\
+\newline
+The basis of most continuous models is a set of $N$ nodes or actors, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
+In contrast to discrete or binary models, all actors usually start with different opinions \cite{Castellano2009}.
+The possible outcomes are more complex than in the discrete models and often have opinion clusters emerging in the final states (where opinion clusters refer to clusters of actors with the same opinion) \cite{Castellano2009}.
 The number of opinion clusters can be one (consensus), two (polarization) or more (fragmentation) \cite{Castellano2009}.
@@
 
-In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance/influence of the neighbors) \cite{Krause2002}.
+In the classical model, there are $N$ nodes or actors, each with an opinion $x_i$ and each actor $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance or influence of the neighbors) \cite{Krause2002}.
 The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
@@
 \end{equation}
-where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
+where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each actor at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
 \newline
 Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}.
-The model they proposed assumes that an agent $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}.
+The model they proposed assumes that an actor $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}.
 This variation thus has the following updating rule (in matrix notation) \cite{Krause2002}
@@
 This model can be used for situations in which so called ``hardening of positions'' occur \cite{Krause2002}.
-This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}.
+This means that, over time, the actors put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}.
 
@@
 
-In bounded confidence models, each node has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance ($\epsilon$ can be the same for all nodes or can be different for each node) \cite{Castellano2009}\cite{Deffuant2006}.
-This threshold determines which opinions of neighbors are close enough to the agent's opinion in order to influence that agent; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
-In other words, an agent with opinion $x_i$ is only affected/influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
+In bounded confidence models, each actor has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance ($\epsilon$ can be the same for all actors or can be different for each actor) \cite{Castellano2009}\cite{Deffuant2006}.
+This threshold determines which opinions of neighbors are close enough to the actor's opinion in order to influence that actor; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
+In other words, an actor with opinion $x_i$ is only influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
 \\
@@
 \newline
-Assume we have a population of $N$ agents, each with an opinion $x_i$ that takes real values in the interval $\interval{0}{1}$ \cite{Castellano2009}.
-Furthermore, the threshold $\epsilon$ is the same for all the agents.
-At each time step, an agent $i$ is randomly selected together with one of its neighbors $j$ (which is also randomly selected) \cite{Castellano2009}.
-If the opinions of the two nodes $i$ and $j$ are too far apart nothing happens; this is the case if $\abs{x_i(t) - x_j(t)} \geqslant \epsilon$.
+Assume we have a population of $N$ actors, each with an opinion $x_i$ that takes real values in the interval $\interval{0}{1}$ \cite{Castellano2009}.
+Furthermore, the threshold $\epsilon$ is the same for all the actors.
+At each time step, an actor $i$ is randomly selected together with one of its neighbors $j$ (which is also randomly selected) \cite{Castellano2009}.
+If the opinions of the two actors $i$ and $j$ are too far apart nothing happens; this is the case if $\abs{x_i(t) - x_j(t)} \geqslant \epsilon$.
 If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, then the opinions are updated to \cite{Castellano2009}\cite{Deffuant2000}
@@
 \newline
-The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two agents move closer to each other \cite{Castellano2009}.
+The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two actors move closer to each other \cite{Castellano2009}.
 The final state is a situation where several opinion clusters coexist.
-This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (the difference in opinions between all agents in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
-Only agents inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
+This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (the difference in opinions between all actors in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
+Only actors inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
 Usually, the number and sizes of the clusters depends on the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
@@
 The Hegselmann-Krause model is similar to the Deffuant model.
-Again, there are $N$ agents, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
-Each agent interacts with those neighbors with an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
+Again, there are $N$ actors, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
+Each actor interacts with those neighbors with an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
 The difference with the Deffuant model lies in the updating rule \cite{Castellano2009}.
@@
 There are, of course, many other variations of the bounded confidence model.
-As mentioned earlier, the threshold $\epsilon$ can be different for each node, so that each node has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
+As mentioned earlier, the threshold $\epsilon$ can be different for each actor, so that each actor has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
 \epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to \cite{Deffuant2006} and \cite{Deffuant2002}.
@@
 \newline
-At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the opinions/posts that he/she received from his/her friends \cite{Perra2019}.
+At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the posts that he/she received from his/her friends \cite{Perra2019}.
 The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}).
@@
 Perra and Rocha showed that the REF, OLD and REC filtering algorithms behave similar in many cases and that the PR filtering algorithm often displays different behavior (such as a larger formation of echo chambers compared to the other three filtering algorithms) \cite{Perra2019}.
-In this thesis the REC and REF filtering algorithms are used as a reference/benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
+In this thesis the REC and REF filtering algorithms are used as a benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
 
@@
 \newline
-It might, however, be more realistic if nodes/users become active/inactive over time.
-This would be a better representation of real social media platforms, where users are active for some periods of time and then become inactive again.
+It might, however, be more realistic if actors or users become active/inactive over time.
+This would be a better representation of social media platforms, where users are active for some periods of time and then become inactive again.
 This kind of activation mechanism of the users gives the network a temporal character.
-The underlying network structure does not change (each user/node has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
-Hence, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent.
-This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity, clustering coefficient, etc may behave differently.
+The underlying network structure does not change (each user or actor has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
+Hence, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information and the network structure behaves like if this user or actor is absent.
+This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity and clustering may behave differently.
 \\
@@
 In this thesis we follow the same activation mechanism as was proposed in the paper of Perra and Rocha \cite{Perra2019}.
-Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$).
-In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes.
+Every time step users are made active with some probability $p_i$ (thus, every time step, each actor becomes active with a probability $p_i$).
+In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the actors.
 Perra and Rocha chose activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
@@
 Then, the active user updates his own opinion according to the opinions on his/her time-line \cite{Perra2019}.
-This updating goes according to the implemented opinion dynamics model (eg.
-majority model, probabilistic majority model,...).
-In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.
-
-% NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where nodes can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, eg.
+This updating goes according to the implemented opinion dynamics model (e.g.
+majority model, probabilistic majority model).
+In every time step the updating of all active actors is done simultaneously \cite{Perra2019}.
+
+% NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where actors can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, e.g.
 if a friend has an opinion far from your, you might be more likely to defriend that one and, vice versa, you might be more likely to become new friends with somebody who has an opinion close to yours)
@@
 Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation/link between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
-\newline
-Knowing the impact of stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups a lot bigger.
+However, there are groups in society, often with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
+\newline
+Knowing the impact of stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups bigger.
 Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore would not reach too many people.
-So their impact could only be of a minor nature (this is only a statement that could have been given, there are of course plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
+So their impact could only be of a minor nature (this is only a statement that could have been given, there are of course plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
 Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
-Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge, and largely unknown, impacts.
-\\
-\newline
-What is the impact of having stubborn/extreme people in your society?
+Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have (largely unknown) impacts.
+\\
+\newline
+What is the impact of having stubborn and extreme people in your society?
 Does it matter if the stubborn people are grouped together or if they are distributed evenly throughout your system?
-Are there ways to reduce the possible impacts of stubbornness and/or extremism?
+Are there ways to reduce the possible impacts of stubbornness and extremism?
 Can these stubborn groups lead to polarization or even overtake the opinions in the system?
-Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism/polarization?
-To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.
+Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism and polarization?
+To answer these questions, several models have been proposed to include some kind of stubbornness in the system under study.
 
@@
 
-The bounded confidence models, discussed in Section \ref{bounded}, can be easily extended to included some notion of stubbornness.
+The bounded confidence models, discussed in Section \ref{bounded}, can be extended to included some notion of stubbornness.
 Guillaume Deffuant proposed such an extension where stubbornness is correlated to extremism \cite{Deffuant2006}.
 The model he proposed goes as follows.
-In bounded confidence models, the agents have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence agent $i$ (agents with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing agent $i$).
+In bounded confidence models, the actors have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence actor $i$ (actors with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing actor $i$).
 Guillaume Deffuant then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or the extreme opinion $1$ \cite{Deffuant2006}.
 These extremists are given a threshold equal to $\epsilon_e$ \cite{Deffuant2006}.
-The other agents are considered to be moderate and can have opinions between $-1$ and $1$.
+The other actors are considered to be moderate and can have opinions between $-1$ and $1$.
 The moderates are given a threshold $\epsilon$, with $\epsilon > \epsilon_e$ \cite{Deffuant2006}.
@@
 \newline
-Another example of introducing some kind of persistence towards the own opinion of the nodes is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
-The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each agent adheres to its initial opinion to a certain degree can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
+Another example of introducing some kind of persistence towards the own opinion of the actors is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
+The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each actor adheres to its initial opinion to a certain degree can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
 
@@
 
-Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is not possible for binary/discrete models.
+Playing around with the threshold $\epsilon_i$ of the actors to include a notion of stubbornness, such as one can do for the bounded confidence models, is not possible for binary or discrete models.
 There are, however, many other ways to include some kind of stubborn actors in these models.
@@
 \newline
-One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
+One remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
 In binary models, however, there are only two possible opinions and in that sense it is less useful to talk about extreme and moderate opinions.
@@
 \newline
-\textbf{Including a stubbornness parameter}\\
-\newline
-One way of creating stubborn actors in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$.
+\textbf{Including a resistance parameter}\\
+\newline
+One way of creating stubborn actors in the binary model is by giving each actor $i$ a resistance parameter $r_i$.
 This resistance parameter can have values between 0 and 1.
-If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...).
-If $r_i = 1$, the node is completely stubborn and there is no chance of convincing him/here of the other idea.
-If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
+If $r_i = 0$, the actor is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority).
+If $r_i = 1$, the actor is completely stubborn and there is no chance of convincing him/here of the other idea.
+If $0 < r_i < 1$, the actor has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
 An uniform random number $f$ between 0 and 1 is drawn.
-If $f > r_i$, the node will update his/her opinion; if $f \leqslant r_i$, the node is not convinced and keeps his/her own opinion.\\
-\newline
-The resistance parameter $r_i$ can be different or the same for all nodes.
-One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
+If $f > r_i$, the actor will update his/her opinion; if $f \leqslant r_i$, the actor is not convinced and keeps his/her own opinion.\\
+\newline
+The resistance parameter $r_i$ can be different or the same for all actors.
+One can also play around with the fraction of actors that is stubborn, for example, there can be a fraction of actors that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
 Many other possibilities can be found according to how the stubborn actors are distributed and according to how stubborn they are made.
-This makes the model very flexible.\\
+This makes the model flexible.\\
 \newline
@@
 In the threshold model one introduces a notion of stubbornness by introducing a threshold parameter $T$.
-This parameter can be either the same for all the nodes ($T$) or can be made individual ($T_i$).
-The main idea is that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
+This parameter can be either the same for all the actors ($T$) or can be made individual ($T_i$).
+The main idea is that actors can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
 The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
@@
 One way, which makes use of the majority model, is as follows.
-First the fraction of friends with opinion A and opinion B, $p_\text{A}$ and $p_\text{B}$ respectively, are determined; then the following rules are used to update the opinion: if $p_\text{A} > p_\text{B}$ and $p_\text{A} > T$, opinion A is adopted; if $p_\text{B} > p_\text{A}$ and $p_\text{B} > T$, opinion B is adopted; in all other cases, the current opinion of the node is not changed.\\
+First the fraction of friends with opinion A and opinion B, $p_\text{A}$ and $p_\text{B}$ respectively, are determined; then the following rules are used to update the opinion: if $p_\text{A} > p_\text{B}$ and $p_\text{A} > T$, opinion A is adopted; if $p_\text{B} > p_\text{A}$ and $p_\text{B} > T$, opinion B is adopted; in all other cases, the current opinion of the actor is not changed.\\
 \newline
 If $T = 0$, this model reduces to the pure majority model described in Section \ref{majModel}.
-If $T = 1$, the nodes do not change their opinion.
-For $0 < T < 1$, the model is a majority model, but the nodes are more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
+If $T = 1$, the actors do not change their opinion.
+For $0 < T < 1$, the model is a majority model, but the actors are more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
 \newline
@@
 
-The previous two chapters, Chapter \ref{chap2} and \ref{chap3}, gave a broad introduction/overview of networks and opinion dynamics models respectively.
+The previous two chapters, Chapter \ref{chap2} and \ref{chap3}, gave a broad overview of networks and opinion dynamics models respectively.
 In this chapter we will summarize the specific networks and opinion dynamics models that are used in this thesis.
@@
 In section \ref{realnetwork}, we will discuss the used real-world networks.
-The reasons of choosing these specific networks will be explained, together with their possible strengths and weaknesses.\\ % + explain where they come from etc?
+The reasons of choosing these specific networks will be explained, together with their possible strengths and weaknesses.\\ % + explain where they come from?
 \newline
@@
 One of the main features of the stochastic block model is the ability to generate networks with community structure.
-We will therefore use this model to investigate the impact of community structure/modularity on the evolution of opinions and the formation of echo chambers.\\
+We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
 \newline
@@
 \newline
-The above models obviously have their added value and benefits.
+The above models have their added value and benefits.
 One of their most important plus points is their simplicity.
-They are the most simple network models that one can create and each of them have characteristics and topologies that can be easily tuned (clustering for WS, community structure for SBM).
+They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM).
 This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account.
@@
 \section{Real-world networks}\label{realnetwork}
-% talk about reasons why you chose the networks you chose: eg. size (computation time),...
+% talk about reasons why you chose the networks you chose: e.g. size (computation time),...
 
@@
 Hence, we chose to test the model on other types of social networks, either on-line or off-line.
-Some possibilities are any form of friendship networks, e-mail networks, collaboration networks, etc.
+Some possibilities are any form of friendship networks, e-mail networks or collaboration networks.
 \\
@@
 This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}.
-If another model is used, eg.
+If another model is used, e.g.
 the majority model, this will be explicitly stated.
@@
 \begin{algorithm}
-\caption{Updating the opinions of the active nodes at time $t$}
+\caption{Updating the opinions of the active actors at time $t$}
 \label{alg:update}
@@
 	\State \(S \leftarrow 20\) \Comment{Maximum length of time-line $T$}
-	\State \(r\) \Comment{Stubbornness parameter of each node, takes values between $0$ and $1$}
+	\State \(r\) \Comment{Stubbornness parameter of each actor, takes values between $0$ and $1$}
 	\ForAll{active user}
@@
 Lets briefly summarize this concept of stubbornness.
-Each node is given a stubbornness parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the node can update his opinion, if $f \leqslant r_i$, the node will keep his current opinion.\\
+Each actor is given a stubbornness parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update his opinion, if $f \leqslant r_i$, the actor will keep his current opinion.\\
 \newline
@@
 \newline
-In Algorithm \ref{alg:update}, the pseudo-code for updating the opinions of the active nodes in case of using a stubbornness parameter $r$ at time step $t$ is displayed. \\
+In Algorithm \ref{alg:update}, the pseudo-code for updating the opinions of the active actors in case of using a stubbornness parameter $r$ at time step $t$ is displayed. \\
 \newline
@@
 Beside that, the cases where people have to choose between only two opinions are rather limited.
-Filtering algorithms, used by real social media companies, are unfortunately not publicly shared.
+Filtering algorithms, used by social media companies, are unfortunately not publicly shared.
 We can however assume that they are not as simple as the ones we used in this thesis.
-Trying to model these far more complex systems by our simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model.
+Trying to model these far more complex systems by our simple model, might lead to the fact that part of the dynamics might be missed or not well captured by our model.
 This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
@@
 This does not mean that simple models should be thrown away.
-They might still be capable of capturing a lot of the dynamics that are seen in the real world \cite{Gug2015}.
+They might still be capable of capturing the dynamics that are seen in the real world \cite{Gug2015}.
 Not because a system appears complex, that it cannot be modeled by something more simple \cite{Jed2019}.
 This has been proven by physics research over and over again \cite{Gug2015}.
-Beside that, if you do not know the dynamics/underlying processes of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
+Beside that, if you do not know the dynamics of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
 There is no need to make things too complex if simple works too.
@@
 \label{50-50}
-In this configuration $50 \%$ of the nodes start with opinion A ($P_{\text{A}}(0) = 0.5$) and $50 \%$ of the nodes start with opinion B ($P_{\text{B}}(0) = 0.5$).
+In this configuration $50 \%$ of the actors start with opinion A ($P_{\text{A}}(0) = 0.5$) and $50 \%$ of the actors start with opinion B ($P_{\text{B}}(0) = 0.5$).
 The opinions are randomly distributed over the network.
@@
 Fig.
-\ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
+\ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
 The value shown on the $y$-axis, $F_N(\left<P_{\text{B}}^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion B (these are thus neighborhoods in which most users adopt opinion A).
+The region on the $x$-axis close to zero describes neighborhoods in which none, or few users, adopt opinion B (these are thus neighborhoods in which most users adopt opinion A).
 Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion B.
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The results for both filtering algorithms are qualitatively the same.
-The main difference is that the formation of echo chambers, using the PR filtering algorithm, is about ten times as big compared to the formation of echo chambers using the REC filtering algorithm.
-\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a very limited, sign of echo chambers, for both the REC filtering algorithm and the PR filtering algorithm.
+The main difference is that the formation of echo chambers, using the PR filtering algorithm, is about ten times as large compared to the formation of echo chambers using the REC filtering algorithm.
+\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a limited, sign of echo chambers, for both the REC filtering algorithm and the PR filtering algorithm.
 Networks with high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
@@
 The stochastic block model (SBM) with a high modularity, a clear community structure, does show an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller compared to networks with high clustering.
-The stochastic block model with a low modularity has no, or very limited, echo chambers, similar as for the Erd\H{o}s-R\'{e}nyi
+The stochastic block model with a low modularity has no, or limited, formation of echo chambers, similar as for the Erd\H{o}s-R\'{e}nyi
 model.
@@
 This decrease is a consequence of the PR filtering algorithm.
-Since the PR filtering algorithm orders the posts according to the users' preferences, nodes with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
-Nodes with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
-However, since there are more nodes with opinion B than nodes with opinion A, the nodes with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
-When using the PR filtering algorithm, there is an imbalance in the sense that nodes with the minority opinion A have a chance of switching to the majority opinion B, whereas nodes with the majority opinion B have (almost) no chance of switching to the minority opinion A.
+Since the PR filtering algorithm orders the posts according to the users' preferences, actors with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
+Actors with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
+However, since there are more actors with opinion B than actors with opinion A, the actors with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
+When using the PR filtering algorithm, there is an imbalance in the sense that actors with the minority opinion A have a chance of switching to the majority opinion B, whereas actors with the majority opinion B have (almost) no chance of switching to the minority opinion A.
 This imbalance leads to the decrease of the minority opinion A.\\
@@
 Fig.
-\ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
+\ref{echo_20_80} shows the distribution of the fraction of friends of actor $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
 For each network, the REC and PR filtering algorithm are compared.
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 This is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
-There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion A.
+There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of actors with the majority of their neighbors carrying the minority opinion A.
 From Fig.
@@
 The PR filtering algorithm shows a different tendency.
-There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion B.
-This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion A in the neighborhood of nodes with the majority opinion B.
+There is an increase in the fraction of actors with (almost) all their neighbors carrying the majority opinion B.
+This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion A in the neighborhood of actors with the majority opinion B.
 This reduced visibility of the minority opinion A in these neighborhoods drives them eventually to adopt opinion B, with an increase in the number of echo chambers around the majority opinion B as a consequence.
@@
 Hence, the increase in the formation of echo chambers around the majority opinion B is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
-This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the nodes carrying the minority opinion A becomes equal to zero.
+This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the actors carrying the minority opinion A becomes equal to zero.
 \\
@@
 \ref{sbm-ws}).
-An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Fig.
+An important note must be made here: the absolute value of neighborhoods with all the actors carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Fig.
 \ref{ws} and in the case of a SBM-WS network in Fig.
@@
 \ref{sbm-ws}, the spikes around zero are not actually reaching infinity, but are cut off around values equal to 100/180.
-These large spikes, indicating a large increase of echo chambers around the minority opinion, do not necessary mean that we have a huge absolute number of echo chambers around the minority opinion A at $t=500$.
+These large spikes, indicating a large increase of echo chambers around the minority opinion, do not necessary mean that we have a large absolute number of echo chambers around the minority opinion A at $t=500$.
 They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion A at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion A at $t=500$).
@@
 \ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for the high clustering networks.
-Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters.
+Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining actors with this minority opinion to group together and form opinion clusters.
 The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
-This means that the average fraction of nodes with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
+This means that the average fraction of actors with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
 In case of the PR filtering algorithm this value is slightly higher than $1$, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
@@
 \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
-The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion A and the normalized fraction of nodes with all neighbors carrying opinion B (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
+The $y$-axis represents the average of the normalized fraction of actors with all neighbors carrying opinion A and the normalized fraction of actors with all neighbors carrying opinion B (with normalized we mean that the computed value for the fraction of actors with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
 This is a measure for the formation of echo chambers (opinion clusters) in the system.
@@
 Each data point is the average of $10^2$ independent simulations.
-The dispersion/variation, which is measured by the standard deviation, in the set of the $10^2$ obtained values is quite large.
+The dispersion, which is measured by the standard deviation, in the set of the $10^2$ obtained values is quite large.
 However, if we run the $10^2$ simulations over and over again, each time calculating the average, the obtained value of the average is always more or less the same.
@@
 	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 
-In this section, we study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of nodes completely stubborn, (ii) all nodes partly stubborn.\\
+In this section, we study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.\\
 \newline
@@
 \newline
-We make a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$).
-We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.
-In all cases, the stubborn nodes are randomly distributed over the network.\\
+We make a fraction of the actors (fracRes) completely stubborn ($r_i = 1$).
+We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$.
+In all cases, the stubborn actors are randomly distributed over the network.\\
 \newline
@@
 	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 This is especially clear for the high clustering networks (WS and SBM-WS).
-The SBM with a high modularity displays a tiny bump.
+The SBM with a high modularity displays a small bump.
 The ER and the SBM with a low modularity (data points for this model lie behind the ER data points) do not seem to have a significant bump.\\
@@
 If fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
-This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that will not change opinions.
+This is probably due to the fact that too many actors become stubborn, so there are more and more actors that will not change opinions.
 This in turn increases the chance of having stubborn neighbors, each with different opinions, who cannot change opinions, which disables the formation of echo chambers.\\
 \newline
-This discrepancy between the results for the PR filtering algorithm and those for the REF/REC filtering algorithms in combination with a small fraction of stubborn actors can be potentially explained with the help of the following simple example.
-Assume we have three nodes that form a triangle.
-Two nodes initially carry opinion A and one node carries opinion B.
-Lets first discuss the PR filtering algorithm and assume that none of the nodes are stubborn.
-In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the nodes adopt the same opinion).
-This can be easily seen: the one node with opinion B is likely to switch to opinion A, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
-The two nodes with opinion A will have more chance of keeping that opinion than switching to opinion B in case of the PR filtering algorithm.
+This discrepancy between the results for the PR filtering algorithm and those for the REF and REC filtering algorithms in combination with a small fraction of stubborn actors can be potentially explained with the help of the following simple example.
+Assume we have three actors that form a triangle.
+Two actors initially carry opinion A and one actor carries opinion B.
+Lets first discuss the PR filtering algorithm and assume that none of the actors are stubborn.
+In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion).
+This can be seen as follows: the one actor with opinion B is likely to switch to opinion A, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
+The two actors with opinion A will have more chance of keeping that opinion than switching to opinion B in case of the PR filtering algorithm.
 This is because the PR filtering algorithm will promote the posts of the friend with the same opinion A over the posts of the friend with opinion B.
-The end result will be that there is a high chance that all the three nodes will adopt opinion A and hence, form an echo chamber.
-Now assume that we make the node with opinion B stubborn.
-This node is no longer capable of switching to opinion A.
-However, the two nodes with opinion A still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion A, and hence reduces the visibility of the posts of the stubborn friend with opinion B.
-The system will not form an echo chamber, since the stubborn node with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the nodes with opinion A to opinion B.
-The conclusion is that, in case of the PR filtering algorithm, making a fraction of the nodes stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
-What happens if we use the REF/REC filtering algorithms?
-Again we have one node with opinion B and two nodes with opinion A in the triangle.
-Assume first that none of the nodes are stubborn.
+The end result will be that there is a high chance that all the three actors will adopt opinion A and hence, form an echo chamber.
+Now assume that we make the actor with opinion B stubborn.
+This actor is no longer capable of switching to opinion A.
+However, the two actors with opinion A still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion A, and hence reduces the visibility of the posts of the stubborn friend with opinion B.
+The system will not form an echo chamber, since the stubborn actor with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the actors with opinion A to opinion B.
+The conclusion is that, in case of the PR filtering algorithm, making a fraction of the actors stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
+What happens if we use the REF or REC filtering algorithms?
+Again we have one actor with opinion B and two actors with opinion A in the triangle.
+Assume first that none of the actors are stubborn.
 For the PR filtering algorithm, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber.
-For the REF/REC filtering algorithms this chance is however smaller, which can be seen as follows.
-The node with opinion B still has a high chance of switching to opinion A, because all his friends (which are the two other nodes in the triangle) carry this opinion.
-However, the two nodes with opinion A will now not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
+For the REF and REC filtering algorithms this chance is however smaller, which can be seen as follows.
+The actor with opinion B still has a high chance of switching to opinion A, because all his friends (which are the two other actors in the triangle) carry this opinion.
+However, the two actors with opinion A will now not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
 Instead, they have an equal chance of keeping their opinion as to change to the other opinion.
-This is because both the REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the nodes with opinion A.
-Both opinion A and opinion B have an equal chance of being promoted on the time-line of the nodes with opinion A.
-This reduced tendency of the nodes with opinion A to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
-This would explain why, in case of no stubborn actors, the REC/REF filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
-Now, lets again assume that we make the node with opinion B stubborn.
+This is because both the REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the actors with opinion A.
+Both opinion A and opinion B have an equal chance of being promoted on the time-line of the actors with opinion A.
+This reduced tendency of the actors with opinion A to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
+This would explain why, in case of no stubborn actors, the REF and REC filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
+Now, lets again assume that we make the actor with opinion B stubborn.
 In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle.
-For the REC/REF filtering algorithms, things are different.
-The stubborn node with opinion B will not change opinion.
-The two nodes with opinion A, on the other hand, will now have a higher chance of switching to opinion B (which was not the case when using the PR filtering algorithm).
+For the REF and REC filtering algorithms, things are different.
+The stubborn actor with opinion B will not change opinion.
+The two actors with opinion A, on the other hand, will now have a higher chance of switching to opinion B (which was not the case when using the PR filtering algorithm).
 This can be seen as follows.
-The REF/REC filtering algorithms do not discriminate between the two different opinions of the two friends of the nodes with opinion A.
-If the node with opinion B is, however, stubborn it will keep on pushing that opinion on the time-lines of the two nodes with opinion A.
-This will eventually lead to an increase of posts with opinion B, compared to posts with opinion A, on the time-lines of the two nodes with opinion A.
-The result will then be that the nodes with opinion A will have a higher chance of switching to opinion B, which enhances the formation of echo chambers in the triangle.
+The REF and REC filtering algorithms do not discriminate between the two different opinions of the two friends of the actors with opinion A.
+If the actor with opinion B is, however, stubborn it will keep on pushing that opinion on the time-lines of the two actors with opinion A.
+This will eventually lead to an increase of posts with opinion B, compared to posts with opinion A, on the time-lines of the two actors with opinion A.
+The result will then be that the actors with opinion A will have a higher chance of switching to opinion B, which enhances the formation of echo chambers in the triangle.
 This would explain the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
-If too many nodes become stubborn, this increased ability of the system to form echo chambers, in case of using the REF/REC filtering algorithms, starts to decrease which can also be easily seen with the help of the triangle example.
-If the fraction of stubborn actors becomes too high, we might end up with the case where both the node with opinion B and one of the nodes with opinion A are stubborn.
-This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three nodes in the triangle carry the same opinion.\\
+If too many actors become stubborn, this increased ability of the system to form echo chambers, in case of using the REF or REC filtering algorithms, starts to decrease which can also be seen with the help of the triangle example.
+If the fraction of stubborn actors becomes too high, we might end up with the case where both the actor with opinion B and one of the actors with opinion A are stubborn.
+This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three actors in the triangle carry the same opinion.\\
 % make figure 
@@
 This result is valid for all the four implemented network models (ER, WS, SBM and SBM-WS).
-Network topologies such as a high clustering coefficient and/or a community structure do not change this result.
+Network topologies such as high clustering or community structure do not change this result.
 \\% Include figure??
@@
 The interpretation of the $y$-axis is slightly different than before.
-In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both.
@@
 This result is valid for all the different network topologies.
-If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion.
+If there are no stubborn actors, the PR filtering algorithm forces actors with the minority opinion to switch to the majority opinion.
 If we introduce a fraction of stubborn actors in the system, we end up with stubborn actors that carry the minority opinion.
@@
 \newline
-In the second configuration, all the nodes are partly stubborn.
+In the second configuration, all the actors are partly stubborn.
 They still have some chance of changing their opinion, they are however less eager to do so.
-All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
+All the actors are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the actors are completely willing to change opinions) to $r=1$ (all the actors are completely stubborn).\\
 \newline
@@
 Fig.
-\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The interpretation of the $y$-axis is again as explained in Subsection \ref{echoVSmod}.
-The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig.
+The results are similar to the case where a fraction of the actors is completely stubborn combined with the PR filtering algorithm (Fig.
 \ref{echo_vs_frac_complStubb}).\\
@@
 The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers.
-This is not changed by making the nodes more and more stubborn.
+This is not changed by making the actors more and more stubborn.
 The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\
@@
 The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
-If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
+If the actors become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
 This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
-If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
-If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
+If $r$ increases and the actors become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
+If $r=1$, the actors retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
 This situation is represented by the black dotted line.
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+	\caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 In Fig.
-\ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution.
+\ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the actors, $r$, is depicted for an initial $50/50$ opinion distribution.
 Fig.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 \ref{echo_vs_all_frac_stubb_PR}).
-The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases.
+The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the actors $r$ increases.
 For the REF filtering algorithm, this increases reaches a maximum around $r = 0.7$ and then starts to slowly decrease between $r = 0.7$ and $r = 0.9$.
@@
 In both case we have a sharp drop from $r = 0.9$ to $r=1$.
-If $r=1$, the nodes are completely stubborn and thus keep their initial opinion.
+If $r=1$, the actors are completely stubborn and thus keep their initial opinion.
 Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
 This is represented by the black dotted line.
-It is remarkable that for high levels of the stubbornness of the nodes $r$, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the nodes are not stubborn.
-We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithms in case of networks with a high clustering coefficient.
-This is the opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes).
+It is remarkable that for high levels of the stubbornness of the actors $r$, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the actors are not stubborn.
+We can conclude that making the actors partly stubborn enhances the formation of echo chambers when using the REF and REC filtering algorithms in case of networks with a high clustering coefficient.
+This is the opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the actors).
 The reason for this discrepancy is probably of a similar nature as for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Completely stubborn actors'' paragraph.
@@
 \newline
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points are hidden behind those of the ER network) do not have any dependency on the stubbornness level of the nodes.
-This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithms that drives this increase in the formation of echo chambers with increasing $r$.
+The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points are hidden behind those of the ER network) do not have any dependency on the stubbornness level of the actors.
+This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF or REC filtering algorithms that drives this increase in the formation of echo chambers with increasing $r$.
 % talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
@@
 \newline
-What about the results for the evolution of the prevalence of the opinions in case of balanced starting conditions and making all the nodes partly stubborn?
-Making all the nodes partly stubborn does not disturb/change the evolution of the opinions.
+What about the results for the evolution of the prevalence of the opinions in case of balanced starting conditions and making all the actors partly stubborn?
+Making all the actors partly stubborn does not disturb the evolution of the opinions.
 The balanced starting conditions remain balanced over time, as was the case for no stubbornness in the system.
@@
 Fig.
-\ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the nodes for an initial $20/80$ distribution.
+\ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the actors for an initial $20/80$ distribution.
 The standard error $SE$ is small and not visible in the figure.
@@
 These figures are not suited to (quantitatively) compare the formation of echo chambers in the different network models in case of unbalanced starting conditions.
-In order to perform such a (quantitative) comparison, one should look at the complete distribution (see eg.
+In order to perform such a (quantitative) comparison, one should look at the complete distribution (see e.g.
 Fig.
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $20/80$ opinion distribution.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the nodes ($r = 0$, $r = 0.4$ and $r = 0.8$).
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the actors ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the actors ($r = 0$, $r = 0.4$ and $r = 0.8$).
 Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
@@
 \noindent
-We will now discuss the results for the evolution of the prevalence of the opinions in case of unbalanced starting conditions and making all the nodes partly stubborn.
+We will now discuss the results for the evolution of the prevalence of the opinions in case of unbalanced starting conditions and making all the actors partly stubborn.
 We investigate four different network models.
@@
 \newline
-The prevalence of opinion A versus time in case of the REC filtering algorithm is not affected by making all the nodes more and more stubborn.
+The prevalence of opinion A versus time in case of the REC filtering algorithm is not affected by making all the actors more and more stubborn.
 The prevalence of the opinions is maintained and the status quo is not broken.
 The PR filtering algorithm gives us another story.
-The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the nodes partly stubborn are given in Fig.
+The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the actors partly stubborn are given in Fig.
 \ref{ev_op_20_80_all_frac_stubb_PR}.
-If the nodes are not stubborn, the PR filtering algorithm reduces the prevalence of the minority opinion over time.
-Making the nodes more and more stubborn reduces this tendency.
+If the actors are not stubborn, the PR filtering algorithm reduces the prevalence of the minority opinion over time.
+Making the actors more and more stubborn reduces this tendency.
 This result is more or less similar to the corresponding case in which a fraction of completely stubborn actors was introduced in the system (Fig.
 \ref{ev_op_20_80_stubb_PR}).
-The only difference is that making all the nodes partly stubborn seems to be less effective in reducing the decrease of the prevalence of the minority opinion compared to making a fraction of the nodes completely stubborn.
+The only difference is that making all the actors partly stubborn seems to be less effective in reducing the decrease of the prevalence of the minority opinion compared to making a fraction of the actors completely stubborn.
 For $r=0.4$, the decrease of the prevalence of the minority opinion is almost not hampered, it is merely slowed down.
-All the nodes need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the prevalence of the minority opinion.
+All the actors need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the prevalence of the minority opinion.
 Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$, respectively).
@@
 We do not make use of a probabilistic majority model, but a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
-The rest of the model (filtering algorithms, activation mechanism,...) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
+The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
 The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
-The threshold parameter $T$ is equal for all the nodes and varies between $T=0$ and $T=1$.\\
+The threshold parameter $T$ is equal for all the actors and varies between $T=0$ and $T=1$.\\
 \newline
@@
 This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
-This is because the majority opinion of the neighbors of a node will always be larger than the threshold if $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
+This is because the majority opinion of the neighbors of an actor will always be larger than the threshold if $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
 If $T \geqslant 0.5$, it starts to have an effect.
-The nodes become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
+The actors become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
 \newline
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.
-The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the threshold parameter of all the actors, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.
+The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the actors, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
 The majority threshold model is implemented.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the nodes, $T$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the threshold parameter of all the actors, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.
+The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the actors, $T$, for an initial $50/50$ opinion distribution.
 The majority threshold model is implemented.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
 This is not due to any network features, but a direct consequence of the usage of the majority model.
-As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes kind of a huge, single echo chamber).
+As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes a large, single echo chamber).
 The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
@@
 \ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REC/REF filtering algorithms.
+Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REF or REC filtering algorithms.
 This is another indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.
@@
 \ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
-At this point, none of the nodes can be convinced of another opinion and hence all the nodes keep their initial opinion.
+At this point, none of the actors can be convinced of another opinion and hence all the actors keep their initial opinion.
 The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.
@@
 \ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
-The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of nodes with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of actors with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
 The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
@@
 The results will be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
-Thus, eg.
+Thus, e.g.
 if we have 10 communities of size 100 and we make $10 \%$ of those communities single opinion A communities while the others get a $50/50$ distribution, the total initial opinion distribution is $55/45$.
@@
 The PR filtering algorithm leads to an increase in the prevalence of the majority opinion in case of unbalanced starting conditions and if the opinions are randomly distributed.
-This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg.
+This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see e.g.
 Fig.
@@
 In the low modularity network, the community structure is not clearly present, but becomes blended.
-The different communities talk a lot to each other and there is a lot of communication between them.
+The different communities talk to each other and there is a lot of communication between them.
 As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over the communities.
@@
 Fig.
-\ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
+\ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
 The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
@@
 The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (bump in the blue line).
-The number of neighborhoods where all the nodes carry opinion A decreases in favor of neighborhoods where we find both nodes with opinion A and nodes with opinion B.
+The number of neighborhoods where all the actors carry opinion A decreases in favor of neighborhoods where we find both actors with opinion A and actors with opinion B.
 Even though there is only a minor communication between the communities, this limited communication has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion A communities, reduces and that the opinions become more mixed.
@@
 \ref{sbm} (these results were obtained for an initial $20/80$ opinion distribution).
-We observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion B.
+We observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of actors with the minority opinion B.
 Remember from Fig.
@@
 Beside this hampering of echo chambers around the majority opinion A, the community distributed case also shows an increase of echo chambers around the minority opinion B (which is not visible for the randomly distributed case).
-This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.
+This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion actors (around $30 \%$) is able to survive and they group together in opinion clusters.
 % share data??
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 $10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases, as was observed for the high modularity network.
-Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion B.
+Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the actors carry the minority opinion B.
 From Fig.
 \ref{low_mod_pr}, it was made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of the minority opinion (when using the PR filtering algorithm).
-This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry the minority opinion.
+This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the actors carry the minority opinion.
 The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case.
@@
 Instead we observe a decrease in the number of echo chambers around the minority opinion.
-For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion A and a minority of the nodes carrying the minority opinion B.
+For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the actors carrying the majority opinion A and a minority of the actors carrying the minority opinion B.
 \\
@@
 The low modularity, community distributed case seems to wiggle around $1$.
-If we look closely, we see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller).
+If we look closely, we see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are smaller).
 We observe a small decrease in the formation of echo chambers around the majority opinion A, with a small increase in more mixed opinion neighborhoods as result.
@@
 \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
-They show a small increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
+They show a small increase in the formation of echo chambers around the minority opinion, indicating the clustering of the actors that carry this opinion.\\
 \newline
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 $50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
 The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
-This suggests that it is not only clustering and/or modularity that drives the formation of echo chambers.
+This suggests that it is not only clustering or modularity that drives the formation of echo chambers.
 Other topologies must play a role as well.
@@
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The SBM-WS with heterogeneous community sizes is compared to the SBM-WS with a single community size.
 Results are compared to the real-world Last.fm network with similar clustering and modularity.
-Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion B at $t = 500$.
+Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 
-We made use of simple toy models of real social media platforms to generate results regarding the evolution of the prevalence of the opinions and formation of echo chambers.
+We made use of simple toy models of social media platforms to generate results regarding the evolution of the prevalence of the opinions and formation of echo chambers.
 These toy models consist of several layers, including a network layer and an opinion dynamics layer.
-Simple filtering algorithms were implemented to replicate and simulate the content curation used by real social media companies.
+Simple filtering algorithms were implemented to replicate and simulate the content curation used by social media companies.
 Furthermore, a notion of stubbornness was introduced to investigate the interplay between individual resistance to change, network structure and filtering algorithm.
@@
 This is mainly seen in the high clustering networks, however a small bump can also be observed in the high modularity network.
-This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF/REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF and REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.
 It was also found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
@@
 They learn us that even though semantic filtering (PR) strongly enhances the formation of echo chambers in the absence of stubborn actors, this enhancing does not stand on its own.
-An interplay between a network with a high clustering and/or community structure is needed to effectively see this increase in echo chambers induced by semantic filtering.
+An interplay between a network with a high clustering or community structure is needed to effectively see this increase in echo chambers induced by semantic filtering.
 A random network, combined with semantic filtering, does not enhance the formation of echo chambers.
-Furthermore, if there are no stubborn actors, recent or random filtering (REC/REF) seems to be more protective against the formation of echo chambers than semantic filtering.
+Furthermore, if there are no stubborn actors, recent or random filtering (REC or REF) seems to be more protective against the formation of echo chambers than semantic filtering.
 If there are, however, small fractions of stubborn actors in the system, this scenario changes: semantic filtering starts to hamper the formation of echo chambers, whereas recent or random filtering begins to enhance polarization.
-It is often assumed that algorithmic personalization, tailored to ones preferences (eg.
+It is often assumed that algorithmic personalization, tailored to ones preferences (e.g.
 semantic filtering), is one of the main reasons of the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
-Our results may, however, suggest, that if there is some kind of stubbornness (which is probably the case in the real world), heterogeneous exposure (REF/REC) to different opinions may be more harmful than a homogeneous exposure (PR).
+Our results may, however, suggest, that if there is some kind of stubbornness (which is probably the case in the real world), heterogeneous exposure (REF or REC) to different opinions may be more harmful than a homogeneous exposure (PR).
 Bail et al.
@@
 \newline
-The formation of echo chambers in social networks/social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature (in the absence of stubborn actors) \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
+The formation of echo chambers in social networks and social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature (in the absence of stubborn actors) \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
 The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.
@@
 It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics.
-We also did not differentiate between the agents in our model, except for non-stubborns versus stubborns.
+We also did not differentiate between the actors in our model, except for non-stubborns versus stubborns.
 This combined with the homogeneous degree distribution of our networks models may be another limitation.
@@
 This may impact the spread of opinions \cite{Fan2020}\cite{Wang2019}.
-We could adapt our model to mimic this real social media behavior more closely.
+We could adapt our model to mimic this social media behavior more closely.
 Therefore, a heterogeneous network model should be introduced, combined with an influence parameter.
-This parameter should be used to determine the influence of an agent on others: the opinion of a more influential agent is easier adopted by the others.
+This parameter should be used to determine the influence of an actor on others: the opinion of a more influential actor is easier adopted by the others.
 In the model, the high degree nodes should then be made more influential.
-As a final limitation, we mention the fixed stubbornness of the agents.
-Even though the stubbornness parameter $r$ could be individualized to each node, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
-This idea may be linked to the influence of certain friends/acquaintances: somebody may adopt another viewpoint more easily from certain friends than from others.\\
+As a final limitation, we mention the fixed stubbornness of the actors.
+Even though the stubbornness parameter $r$ could be individualized to each actor, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
+This idea may be linked to the influence of certain friends: somebody may adopt another viewpoint easier from certain friends than from others.\\
 \newline
@@
 If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
-This is an important result, since it may change our beliefs/assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
+This is an important result, since it may change our assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
 More research on this topic is definitely needed in the near future.
@@
 \end{equation}
-which characterizes a random graph. In the case $k \ll N$, we can simplify the binomial in the following way \cite{Albert2014}
+which characterizes a random network. In the case $k \ll N$, we can simplify the binomial in the following way \cite{Albert2014}
 \begin{equation}\label{bin}
@@
 \end{equation}
-Since every node in the regular ring lattice is identical, we can easily find the total number of triangles and connected triples in the network by multiplying Eq.
+Since every node in the regular ring lattice is identical, we can find the total number of triangles and connected triples in the network by multiplying Eq.
 (\ref{triangle}) and Eq.
@@
 \end{equation}
-where the last approximation holds for $N \gg 1$. Since every node in the regular ring lattice is identical we can easily find the average distance of the network \cite{chap9}
+where the last approximation holds for $N \gg 1$. Since every node in the regular ring lattice is identical we can find the average distance of the network \cite{chap9}
 \begin{equation}

### 2021-05-20 draft.tex
rewrote chapter 3

@@
 In order to assure a more pleasant and convenient experience on their platforms, social media companies use algorithms to order and filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}.
-It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
+It is of importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
 \newline
@@
 The use of statistical physics may, however, offer a way out.
-In statistical physics it is found that different systems behave qualitatively the same and are governed by the same laws near the critical point \cite{Hu2018}\cite{Kadanoff2010}.
+In statistical physics different systems behave qualitatively the same and are governed by the same laws near the critical point \cite{Hu2018}\cite{Kadanoff2010}.
 This lead to the concept of universality: many different microscopic problems have the same macroscopic manifestation \cite{Gug2015}\cite{Kadanoff2015}.
@@
 These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate on the case of social networks.
-It is found that many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
+Many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale-free degree distribution represents one of the three general properties of social networks.
@@
 The opinion of the actors in the model can be either discrete or continuous.
-This thesis deals with models where each actor can have one of the two opinions, A or B, and thus only deals with the case of discrete opinions.
+This thesis deals with models where each actor can have one of two opinions, A or B, and thus only deals with the case of discrete opinions.
 This might come across as a simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (e.g.
@@
 \newline
-Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix \cite{F.Costa2007}
+Let us denote the number of triangles as $N_{\Delta}$ and the number of connected triples as $N_3$. These two numbers can be obtained from the adjacency matrix
 \begin{align}
@@
 \noindent
-where $A_{ij}$ are the elements of the adjacency matrix. It is possible to define the transitivity of one node as \cite{F.Costa2007}
+where $A_{ij}$ are the elements of the adjacency matrix \cite{F.Costa2007}. It is possible to define the transitivity of one node as
 \begin{equation}\label{trans}
@@
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
-This however not the case for the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$.
+This is not the case for the average clustering coefficient $\left<cc\right>$ and the global transitivity $T$.
 The difference between the two definitions is that the average of Eq.
@@
 This last quantity is given by $p_{ij}$, which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
-Hence, the modularity of a community $Q_c$ becomes \cite{Albert2016}
+Hence, the modularity of a community $Q_c$ becomes
 \begin{equation}
@@
 \end{equation}
-where the factor $\frac{1}{2L}$ is a normalization factor.
+where the factor $\frac{1}{2L}$ is a normalization factor \cite{Albert2016}.
 If we randomly wire the network, using the degree preserving null model, we find the probability that nodes with degree $k_i$ and degree $k_j$ are connected to each other \cite{Bara2016}
@@
 Modularity maximization forces small communities into larger ones \cite{Albert2016}.
-If two communities $A$ and $B$ are merged, the modularity changes by \cite{Albert2016}
+If two communities $A$ and $B$ are merged, the modularity changes by
 \begin{equation}
@@
 \end{equation}
-where $l_{AB}$ is the number of direct edges that connect nodes in community $A$ with nodes in community $B$; $k_A$ and $k_B$ are the total degree of nodes in communities $A$ and $B$ respectively.
+where $l_{AB}$ is the number of direct edges that connect nodes in community $A$ with nodes in community $B$; $k_A$ and $k_B$ are the total degree of nodes in communities $A$ and $B$ respectively \cite{Albert2016}.
 The derivation of this formula is found in Appendix \ref{modul}, Section \ref{modchange}.
@@
 Watts and Strogatz showed by numerical simulations that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random network (which is small), even for small $\beta$ \cite{Newman2000}.
-The clustering coefficient is approximately given by \cite{Barrat1999}
+The clustering coefficient is approximately given by
 \begin{equation}
@@
 where $\left<cc(0)\right>$ is given by Eq.
-(\ref{C0}).
+(\ref{C0}) \cite{Barrat1999}.
 This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
@@
 \newline
-An extension that we are going to explore a bit deeper is the stochastic Watts-Strogatz block model (SBM-WS).\\
+An extension that we are going to explore deeper is the stochastic Watts-Strogatz block model (SBM-WS).\\
 \newline
@@
 The models discussed in the previous sections are idealized and simplified.
-They may posses some of the characteristics of real-world networks, but they certainly do not always describe real-world networks appropriately.
-The question then arises: What are the main characteristics of real-world networks, and more specifically of real-world social networks?
-Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
-\newline
-As discussed briefly in the introduction, it is found that real-world networks often posses three characteristic properties.
+They may posses some of the characteristics of real-world networks, but they do not always describe real-world networks appropriately.
+Here, we will list some of the main  characteristics of real-world networks, and more specifically of real-world social networks.
+An important question that can be asked is whether there is a significant difference between off-line social networks and on-line social networks.\\
+\newline
+As discussed briefly in the introduction, real-world networks often posses three characteristic properties.
 These are the two properties of the ``small world phenomenon'', namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
 These properties are valid for both off-line and on-line social networks \cite{Zhang2014}.
-Lets discuss them in a bit more detail.\\
+Lets discuss them in more detail.\\
 \newline
@@
 He asked randomly selected people in Boston and Omaha to forward a letter to a distant target person.
-The letter could, however, only be sent to acquaintances, thought to be closer to the target person \cite{RealWorld}.
-The remarkable result of this experiment was that the average number of steps between the sender and the target person was only around six.
-This phenomenon is now often referred to as ``six degrees of separation'' \cite{RealWorld}.\\
-\newline
-This phenomenon is also observed in many other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
+The letter could only be sent to acquaintances, thought to be closer to the target person \cite{RealWorld}.
+The result of this experiment was that the average number of steps between the sender and the target person was around six.
+This phenomenon is often referred to as ``six degrees of separation'' \cite{RealWorld}.\\
+\newline
+This phenomenon is also observed in other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
 Such networks usually have an average distance that depends logarithmically on the system size instead of linearly \cite{Newman2003}.
 A short average distance facilitates a fast transmission of information \cite{Newman2003}\cite{Zhang2014}.\\
-\newpage
-\noindent
+\newline
 \textbf{High clustering coefficient}\\
@@
 In real-world social networks, one often observes the pattern that two friends of an individual are also friends of each other \cite{RealWorld}.
-In network terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
+In graph terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}.
 This property is not only found in social networks, but also in a wide variety of different real-world networks \cite{RealWorld}.\\
@@
 \newline
-Many real-world networks are found to have a power-law degree distribution \cite{RealWorld}
+Many real-world networks are found to have a power-law degree distribution
 \begin{equation}
@@
 \end{equation}
-where the exponent $\gamma$ often lies between 2 and 3.
+where the exponent $\gamma$ often lies between 2 and 3 \cite{RealWorld}.
 Networks with a power-law degree distribution are also called heterogeneous or scale-free networks.
-The term scale-free refers to the fact that there is no characteristic scale in the network.
+The term scale-free refers to the fact that there is no characteristic scale in the network: the system is invariant over different scales.
 Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a high degree (these nodes are referred to as hubs).
 If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges \cite{Newman2005}.
-Simple scale-free networks with an exponent smaller than two, however, cannot exist, since the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}.
-If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance \cite{Newman2005}.
-A diverging variance means that the fluctuations around the average can be arbitrary large, which makes this average degree no longer a characteristic scale for the system \cite{Barabasi2016}.\\
-\newline
-It should be noted that not all real-world networks have a heterogeneous degree distribution.
-And even if the degree distribution is heterogeneous or fat-tailed, it does not always need to follow a power-law.
+Simple scale-free networks with an exponent smaller than two cannot exist, because the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}.
+If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance in the limit of infinite system size ($N \rightarrow \infty$) \cite{Newman2005}.
+Real networks always have finite sizes, so there is always a finite average degree and variance.
+This variance would, however, be meaningless \cite{Newman2005}.
+A diverging variance means that the fluctuations around the average can be arbitrary large \cite{Barabasi2016}.\\
+\newline
+Not all real-world networks have a heterogeneous degree distribution.
+Even if the degree distribution is heterogeneous or fat-tailed, it does not always follow a power-law.
 The degree distribution can e.g.
@@
 These three main characteristics are valid for on- and off-line social networks.
-Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}.
+Another property that is found in most on- and off-line networks is the appearance of a giant component, that is, most nodes belong to the same connected component \cite{Latapy}.
 They also often posses a modular structure \cite{Ferrara2012}\cite{McGlohon2011}.
@@
 There are, however, some properties that are different between on-line and off-line social networks.
-One of them might be the degree correlations or mixing pattern.
-It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern (this distinguishes them from other real-world networks, e.g.
-biological networks \cite{F.Costa2007}, which often have a disassortative pattern) \cite{Hu2009}\cite{Zhang2014}.
-For a long time, it was thought that this is also valid for on-line social networks.
-However, recent research has shown that a lot of on-line social networks display disassortative mixture \cite{Hu2009}\cite{Zhang2014}.
-This discrepancy between real-life social networks and on-line social networks may be explained as follows.
+One of them is the degree correlations or mixing pattern.
+It is often said that most real-life social networks have an assortative mixing pattern (this would distinguish them from most other real-world networks, e.g.
+biological networks \cite{F.Costa2007}, which often have a disassortative pattern) \cite{Fisher2017}\cite{Hu2009}\cite{Zhang2014}.
+This assumption of the assortative nature of social networks has however been questioned \cite{Fisher2017}\cite{Whitney2010}.
+For a long time, it was thought that this assortative mixing is also valid for on-line social networks.
+However, recent research has shown that a number of on-line social networks display disassortative mixture \cite{Holme2004}\cite{Hu2009}\cite{Zhang2014}.
+This discrepancy between real-life and on-line social networks may be explained as follows.
 In real life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}.
 The `ordinary people' do not have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure.
-In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance and influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
+In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his or her importance and influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
 This leads to a disassortative mixing pattern.
-% also talk about size of groups that people can maintain?
+
+\section{Temporal networks}
+
+Real-life systems are often not static but evolve in time.
+In order to represent these systems appropriately, we need temporal networks.
+In temporal networks, the edges (or nodes) are not continuously active \cite{Holme2012}.
+Instead, edges (or nodes) become active and inactive over time and time becomes an explicit element of the network representation \cite{Holme2012}.
+An example of such temporal networks in real life are communication networks such as the phone network where edges represent instantaneous contacts \cite{Holme2012}.
+Other examples of applications of temporal networks are the application to the dynamics of functional brain connectivity \cite{Thompson2017} and the application to infrastructure networks \cite{Salama2020}.\\
+\newline
+Temporal networks require different methods for analyzing topological and temporal structures \cite{Holme2012}\cite{Holme2021}.
+The time ordering can e.g.
+influence the transitivity of edges: in a static network, if A is connected to B and B is connected to C, then A is indirectly connected to C; in temporal networks this does not need to be so, if A is connected to B at some time $t$ and B is connected to C at some time $t'$ later, then A is not indirectly connected to C \cite{Holme2012}.\\
+\newline
+Not all temporal systems should be modeled by a temporal network.
+Projecting a temporal system on a static network looses some information, but the analysis of temporal networks is more complex and involved.
+A trade-off should be made to see whether the more complex structure of temporal networks outweighs the loss of information when modeling the system as a static network \cite{Holme2012}.
+Another situation in which it is not necessary to model the system as a temporal network is when the dynamical system on the network is too rapid compared to the dynamics of the edges \cite{Holme2012}.\\
+\newline
+We will not go into too much detail. A review of temporal networks and related measurements is found in \cite{Holme2012} and \cite{Holme2021}.
 
@@
 \label{chap3}
-The study of opinion dynamics in a group of people, as mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer.
-The network layer, which describes the underlying structure and correlations between the actors, is intensively described in the previous chapter, Chapter \ref{chap2}.
-The current chapter focuses on the second layer, that of the opinion dynamics models.
-These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way.
+The study of opinion dynamics in a group of people has two layers: a network layer and an opinion dynamics layer.
+The network layer, which describes the underlying structure and correlations between the actors, is described in the previous chapter, Chapter \ref{chap2}.
+This chapter focuses on the second layer, that of the opinion dynamics models.
+These models describe how opinions are formed and how they evolve in a group of people that interact socially.
 Hence, these opinion dynamics are the dynamics that run on the underlying network structure.\\
 \newline
-We will study models for the filtering algorithms used by social media companies (Section \ref{filter}).
-This is needed because this thesis deals with opinion dynamics on on-line social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.\\
-\newline
-We will also discuss the activation mechanism (Section \ref{actMech}).
-This is the mechanism that describes the timings in which content or information is exchanged between users \cite{Perra2019}.
-Such a mechanism changes the network from a static one to a temporal one, where actors become active and inactive over time.
-\\
-\newline
-Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis.
-There are two major groups of opinion dynamics models, namely binary or discrete models and continuous approaches.
+Several opinion dynamics models have been proposed; here we will review some of the most important ones.
+There are two groups of opinion dynamics models, namely binary or discrete models and continuous approaches.
 These will be discussed in Section \ref{Bin} and Section \ref{Con}, respectively.
-This thesis only deals with binary opinion dynamics models, so these will be thoroughly described; the continuous models will be quickly reviewed in order to give a more complete overview.\\
-\newline
-Finally, in Section \ref{stubb}, the introduction of stubborn actors is discussed.
+This thesis only deals with binary opinion dynamics models; the continuous models will be reviewed in order to give a more complete overview.\\
+\newline
+In section \ref{actMech}, we will pay attention to the activation mechanism.
+This is the mechanism that describes the timings in which information is exchanged between actors \cite{Perra2019}.
+Such mechanism changes the network from a static one to a temporal one, where actors become active and inactive over time.
+Then, in Section \ref{stubb}, the introduction of stubborn actors will be discussed.
+Section \ref{filter} will study filtering algorithms used by social media companies.
+This is needed because this thesis deals with opinion dynamics on on-line social networks, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.
+Finally, in Section \ref{modelThesis} the opinion dynamics model and the filtering algorithms of this thesis will be summarized.
 
@@
 
-Binary models are models where each actor can have one of the two opinions A or B.
-A binary model could reflect real-life cases where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
-\newline
-In this section, some binary models will be discussed; note, however, that the models described here are far from all the models that currently exist.
+Binary models are models where each actor can have one of two opinions A or B.
+A binary model could reflect real-life where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
+\newline
+In this section, the most important binary models will be discussed.
 
@@
 One of the simplest opinion dynamics models is the binary majority model.
-Each actor has one of the two opinions A or B.
-Initially, these opinions can be equally distributed or not.
-If the distribution is not equal, but instead e.g.
-$80/20$, there is a predominant opinion and a minority opinion.
-The following updating rule is imposed on the network: each actor adopts the majority opinion of his/her nearest neighbors \cite{Nguyen2020}.
-Initially each actor is assigned an opinion; then the opinions are evolved in time: each individual iteratively updates his/her opinion according to the majority opinion of his/her direct neighbors \cite{Nguyen2020}.\\
-\newline
-Several studies have been performed on this kind of model, searching for answers on several questions such as: What are the conditions that drive the system to converge to unanimity; and, Does an opinion, that is initially a majority opinion, remain a majority opinion in the final state \cite{Nguyen2020}?
-\\
-\newline
-Most of these studies conclude that, at the end of the opinion evolution, a complete consensus state is obtained where all the actors adopt the same opinion \cite{Nguyen2020}.
-This is even the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini et al.
+Each actor has one of two opinions, A or B.
+Initially, these opinions can be equally distributed or not; the starting conditions can be balanced or unbalanced.
+Balanced starting conditions indicate the initial situation where $50 \%$ of the actors is given opinion A and $50 \%$ of the actors is given opinion B.
+The opinions are usually randomly assigned; this is however not necessary and opinions can e.g.
+also be assigned according to the community a certain actor belongs to.
+If the starting condition are unbalanced, we have a majority and a minority opinion.
+E.g.
+$80 \%$ of the actors can be assigned opinion A and $20 \%$ of the actors is then given opinion B.
+Hence, opinion A is the predominant or majority opinion and opinion B is the minority opinion.
+The following updating rule is imposed on the network: each actor adopts the majority opinion of its nearest neighbors \cite{Nguyen2020}.
+Initially each actor is assigned an opinion; then the opinions are evolved in time: each actor iteratively updates its opinion according to the majority opinion of its direct neighbors \cite{Nguyen2020}.\\
+\newline
+Several studies have been performed on this kind of model with the aim to find the conditions that drive the system to converge to unanimity or to see whether an initial majority opinion remains a majority opinion in the final state \cite{Nguyen2020}.\\
+\newline
+Most of these studies conclude that, in the stationary state, a consensus is obtained where all the actors adopt the same opinion \cite{Nguyen2020}.
+This is also the case when the opinions are equally and randomly distributed in an Erd\H{o}s-R\'{e}nyi network (with the side note that the degree needs to be high enough), as was shown by Benjamini et al.
 \cite{Benjamini2016}.
@@
 
-The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic nature.
-In this model the fraction of neighbors of user $i$ with opinion A, $p_\text{A}$, and the fraction of neighbors with opinion B, $p_\text{B}$, are determined.
-Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to $p_\text{A}$ (or $p_\text{B}$) \cite{Perra2019}.\\
-\newline
-For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
-User $i$ no longer simply updates his/her opinion according to the majority opinion carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
-\newline
-This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}.
+The probabilistic majority model is similar to the majority model, but has, as the name suggests, a probabilistic nature.
+In this model the fraction of neighbors of actor $i$ with opinion A, $p_\text{A}$, and the fraction of neighbors with opinion B, $p_\text{B}$, are determined.
+Then, the opinion of actor $i$ is updated: it takes on opinion A (or B) with a probability equal to $p_\text{A}$ (or $p_\text{B}$) \cite{Perra2019}.\\
+\newline
+For example, if $70\%$ of the neighbors of actor $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), actor $i$ will update its opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
+Actor $i$ no longer updates its opinion according to the majority opinion carried by its neighbors, but the majority opinion still has a higher probability to be adopted by this actor.\\
+\newline
+This model is stochastic, which is an important component of opinion dynamics \cite{Castellano2009}\cite{Perra2019}.
+Opinion dynamics processes are not purely deterministic; individual behavior, social influence and influence of external factors such as mass media and political campaigns may vary in an unpredictable way \cite{Castellano2009}.
 
@@
 \newline
-The pressure of the majority in only felt in an averaged way, in the sense that a neighbor that carries the majority opinion has more chance to be randomly selected \cite{Castellano2009}.\\
-\newline
-Lots of studies have been performed on this model.
-In this thesis, however, we will not make use of the voter model.
+The pressure of the majority in only felt in an averaged way, in the sense that a neighbor that carries the majority opinion has a higher chance to be randomly selected \cite{Castellano2009}.\\
+\newline
+This model has been extensively studied.
+We will not use the voter model.
 Hence, we will not go into further detail.
@@
 Some concepts that are valid for the discrete models no longer apply (e.g.
-the concept of a majority and minority opinion or the concept of equality of opinions); hence, they require a bit of a different framework \cite{Castellano2009}.
-\\
-\newline
-The basis of most continuous models is a set of $N$ nodes or actors, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
+the concept of a majority and minority opinion or equality of opinions); hence, they require a different framework \cite{Castellano2009}.
+\\
+\newline
+The basis of most continuous models is a set of $N$ actors, each having an opinion $x_i$ represented by a real number that takes values in some interval \cite{Castellano2009}.
 In contrast to discrete or binary models, all actors usually start with different opinions \cite{Castellano2009}.
@@
 The number of opinion clusters can be one (consensus), two (polarization) or more (fragmentation) \cite{Castellano2009}.
-\\
-\newline
-Several models have been proposed, we will only discuss a fraction of them.
 
@@
 
-In the classical model, there are $N$ nodes or actors, each with an opinion $x_i$ and each actor $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance or influence of the neighbors) \cite{Krause2002}.
-The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
+In the classical model, there are $N$ actors, each with an opinion $x_i$ and each actor $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance or influence of the neighbors) \cite{Krause2002}.
+The most simple model has fixed weights $w$ and the updating rule is
 \begin{equation}\label{refMod}
@@
 \end{equation}
-where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each actor at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
+where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each actor at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$ \cite{Krause2002}.\\
 \newline
@@
 The model they proposed assumes that an actor $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}.
-This variation thus has the following updating rule (in matrix notation) \cite{Krause2002}
+This variation thus has the following updating rule (in matrix notation)
 \begin{equation}
@@
 \end{equation}
-where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix. If all the $g_i$ are equal to zero, we obtain the model of Eq. (\ref{refMod}) \cite{Krause2002}.\\
-\newline
-A possible time variant extension is \cite{Krause2002}
+where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix  \cite{Krause2002}.
+If all the $g_i$ are equal to zero, we obtain the model of Eq.
+(\ref{refMod}) \cite{Krause2002}.\\
+\newline
+A possible time variant extension is
 \begin{equation}
@@
 \end{equation}
-where now the weights are time dependent.
+where now the weights are time dependent \cite{Krause2002}.
 This model can be used for situations in which so called ``hardening of positions'' occur \cite{Krause2002}.
@@
 \end{equation}
-The parameter $\mu$ is called the convergence parameter; its value lies in the interval $\interval{0}{1/2}$ \cite{Castellano2009}. \\
+The parameter $\mu$ is called the convergence parameter whose value lies in the interval $\interval{0}{0.5}$ \cite{Castellano2009}. \\
 \newline
 The model is based on a compromise strategy: after some discussion and a constructive debate, the opinions of the two actors move closer to each other \cite{Castellano2009}.
-The final state is a situation where several opinion clusters coexist.
-This is because the opinions move closer and closer to each other in several clusters; once each cluster is sufficiently far apart from the others (the difference in opinions between all actors in the two clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
-Only actors inside the same cluster keep on interacting with each other, which makes the opinions inside a cluster converge to a single value \cite{Castellano2009}.
-Usually, the number and sizes of the clusters depends on the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
+The stationary state is a situation where several opinion clusters coexist.
+This is because the opinions move closer and closer to each other in several of such opinion clusters; once each opinion cluster is sufficiently far apart from the others (the difference in opinions between all actors in the two opinion clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
+Only actors inside the same opinion cluster keep on interacting with each other, which makes the opinions inside an opinion cluster converge to a single value \cite{Castellano2009}.
+Usually, the number and sizes of the opinion clusters depends on the threshold $\epsilon$, whereas $\mu$ determines the convergence rate \cite{Castellano2009}.\\
 \newline
@@
 The Hegselmann-Krause model is similar to the Deffuant model.
-Again, there are $N$ actors, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
+There are $N$ actors, each with an opinion $x_i$ and a constant threshold $\epsilon$ \cite{Castellano2009}.
 Each actor interacts with those neighbors with an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
@@
 \newline
-The dynamics of this model is similar to the Deffuant model and the final state is also characterized by a number of stationary opinion clusters \cite{Castellano2009}.
+The dynamics of this model is similar to the Deffuant model and the stationary state is also characterized by a number of opinion clusters \cite{Castellano2009}.
 The number of clusters decreases if $\epsilon$ increases and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
 \newline
-There are, of course, many other variations of the bounded confidence model.
+There are variations of the bounded confidence model.
 As mentioned earlier, the threshold $\epsilon$ can be different for each actor, so that each actor has an opinion $x_i$ and a threshold $\epsilon_i$ or the confidence interval can be made asymmetric $\interval{x_i - \epsilon_l}{x_i + \epsilon_r}$ with $\epsilon_l \neq
-\epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model. Interested readers are referred to \cite{Deffuant2006} and \cite{Deffuant2002}.
-
-\section{Algorithmic personalization}\label{filter}
-
-On-line social platforms use algorithms to order and filter the posts that appear on the time-line of an individual \cite{Perra2019}.
-They do this in order to make the experience on their platforms more pleasant, interesting and convenient \cite{Perra2019}.
-If social media companies would not use any kind of filtering, there would be too much content to handle and process for the social media user \cite{Bozdag2013}.
-This in turn would make the on-line social experience unpleasant and inconvenient; the user might feel overwhelmed by the number of information he/she is exposed to \cite{Bozdag2013}.
-The reason behind this is that people are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that humans can process \cite{Bozdag2013}.\\
-\newline
-The downside of such algorithmic filtering is that it fine-tunes the information it shows based on the user.
-Even two users with the same friends may then be exposed to different content, based on their previous activities and history \cite{Bozdag2013}.
-This may lead to the formation of echo chambers in which users get `trapped' in content that confirms their own mindset \cite{Bozdag2013}.
-\\
-\newline
-It is important to know and be aware of the possible impacts of algorithmic personalization on the formation and evolution of opinions.
-Perra and Rocha partly investigated the role of such algorithms on the dynamics of opinions in several networks in \cite{Perra2019}.
-Since filtering algorithms are often not transparently shared by social media companies, they proposed several personalization algorithms based on three simple ways of filtering \cite{Perra2019}: (i) Popularity filtering: promoting content that is popular across the platform; (ii) Semantic filtering: promoting posts that are similar to previous consumed posts; (iii) Collaborative filtering: promoting content that people similar or close to us consume.\\
-\newline
-The following paragraph briefly reviews the filtering algorithms that Perra and Rocha proposed. \\
-\newline
-At every time $t$ each platform user has a hidden list $L_i(t)$ that contains all the posts that he/she received from his/her friends \cite{Perra2019}.
-The filtering algorithm then selects $S$ items from this list and shows them on the user's time-line (this is the fundamental idea behind algorithmic personalization: only showing a fraction of the possible posts \cite{Perra2019}).
-The following selection methods are used \cite{Perra2019}: (i) the reference filtering algorithm (REF): randomly select $S$ items from the list $L_i(t)$; (ii) the old filtering algorithm (OLD): the $S$ oldest post in $L_i(t)$ are displayed on the time-line of the user; (iii) the recent filtering algorithm (REC): shows the $S$ most recent items from $L_i(t)$; and (iv) the preference filtering algorithm (PR): sorts the hidden list $L_i(t)$ according to the current opinion of the user and selects the first $S$ elements.
-This filtering algorithm is based on semantic filtering.\\
-\newline
-Perra and Rocha showed that the REF, OLD and REC filtering algorithms behave similar in many cases and that the PR filtering algorithm often displays different behavior (such as a larger formation of echo chambers compared to the other three filtering algorithms) \cite{Perra2019}.
-In this thesis the REC and REF filtering algorithms are used as a benchmark algorithm and the PR filtering algorithm is used as the main filtering algorithm.
+\epsilon_r$ \cite{Krause2002}. Another extension is the relative agreement interaction model \cite{Deffuant2006}\cite{Deffuant2002}.
 
@@
 
-Most of the simple opinion dynamics models make use of a static network in which users are always active and always exchanging opinions.
-In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented.
-Another possibility is updating all users at the same time during each time step.
-\\
-\newline
-It might, however, be more realistic if actors or users become active/inactive over time.
+Most of the standard opinion dynamics models use a static network in which actors are always active and exchanging opinions.
+In these models the opinion dynamics might go as follows: each time step a random actor is selected after which its opinion is updated according to the opinion dynamics model that is implemented.
+Another possibility is updating all actors at the same time during each time step.
+\\
+\newline
+It might be more realistic if actors become active/inactive over time.
 This would be a better representation of social media platforms, where users are active for some periods of time and then become inactive again.
-This kind of activation mechanism of the users gives the network a temporal character.
-The underlying network structure does not change (each user or actor has a fixed number of neighbors that does not change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which are not.
-Hence, this does affect the network structure in some way, because if a user, for example, does not become active for a long period of time, he/she does not participate in the exchange of information and the network structure behaves like if this user or actor is absent.
+This kind of activation mechanism of the actors gives the network a temporal character.
+The underlying network structure does not change (each actor has a fixed number of neighbors that does not change over time; no new friends are made or, conversely, no old friends are being defriended).
+However, by making actors active and inactive the exchange of information between actors becomes time dependent on which actors are active or not.
+Hence, this does affect the network structure in some way, because if an actor, for example, does not become active for a long period of time, it does not participate in the exchange of information and the network structure behaves like if this actor is absent.
 This may affect the dynamics, since it may seem as if the network has a different average degree and some properties like connectivity and clustering may behave differently.
@@
 \newline
-In this thesis we follow the same activation mechanism as was proposed in the paper of Perra and Rocha \cite{Perra2019}.
-Every time step users are made active with some probability $p_i$ (thus, every time step, each actor becomes active with a probability $p_i$).
-In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the actors.
-Perra and Rocha chose activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
-They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
-However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}.
-\\
-\newline
-For simplicity, this thesis implements the constant activation probability variant.
-It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
-\\
-\newline
-If a user is active, he/she first spreads his current opinion to all his/her neighbors.
-Then, the active user updates his own opinion according to the opinions on his/her time-line \cite{Perra2019}.
-This updating goes according to the implemented opinion dynamics model (e.g.
-majority model, probabilistic majority model).
-In every time step the updating of all active actors is done simultaneously \cite{Perra2019}.
-
-% NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where actors can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, e.g.
-if a friend has an opinion far from your, you might be more likely to defriend that one and, vice versa, you might be more likely to become new friends with somebody who has an opinion close to yours)
+In this thesis, actors are made active with some probability $p_i$ every time step (thus, every time step, each actor becomes active with a probability $p_i$).
+The activation probability is set equal to $p = 0.1$, which is constant for all the actors.
+This constant activation probability is chosen for simplicity.
 
@@
 
-It is well known that in real life, you can find people that are persistent towards their own opinion and that are not open to opinions of others.
+In real life, some people are persistent towards their own opinion and are not open to the opinions of others.
 They are convinced that they carry the truth with them and that other people, with different world-views, are not capable of giving them any new and useful insights.
 They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs.
-Of course this is put a bit simplistic and there are many shades of gray between this black and white statement.
-However, there are groups in society, often with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often widely accepted that there is some kind of correlation between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be seen as a form of critical thinking and as a form of taking things with a grain of salt).\\
-\newline
-Knowing the impact of stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of more extreme and persistent groups bigger.
-Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore would not reach too many people.
-So their impact could only be of a minor nature (this is only a statement that could have been given, there are of course plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, all examples of groups that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
-Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
-Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have (largely unknown) impacts.
-\\
-\newline
-What is the impact of having stubborn and extreme people in your society?
-Does it matter if the stubborn people are grouped together or if they are distributed evenly throughout your system?
-Are there ways to reduce the possible impacts of stubbornness and extremism?
-Can these stubborn groups lead to polarization or even overtake the opinions in the system?
-Or is introducing some kind of stubbornness, on the other hand, a protective factor towards extremism and polarization?
-To answer these questions, several models have been proposed to include some kind of stubbornness in the system under study.
+This is a bit simplistic and there are shades of gray between this black and white statement.
+However, there are groups in society, with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often the case that there is some correlation between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be a form of critical thinking and a form of taking things with a grain of salt).\\
+\newline
+Knowing the impact of stubborn groups on the opinions of others is of importance, especially with the rise of social media.
+Social media facilitates the spread of information and ideas.
+This is, unfortunately, also true for extreme and persistent groups.
+They become capable of reaching and influencing larger groups of people.
+Before the era of social media, one could maybe argue that these extreme and stubborn groups were often isolated and therefore would not reach too many people.
+So their impact could only be minor (this is only a statement that could have been given, there are plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, communism exploited by, among other, Stalin and Mao in the 20th century, all examples of groups or regimes that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
+Today, this argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no (time and spatial) constraints.
+Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have unknown impact.
+\\
+\newline
+Several models that include some form of stubbornness have been proposed to study the impact of stubborn and extreme groups on society.
+These models aim to study whether the distribution of the stubborns (clustered versus random) matters; whether stubborn groups can lead to polarization or can overtake the opinions in the system or, contrary, whether stubbornness is protective against extremism and polarization.
+We will review some of the most important models.
 
@@
 
-The bounded confidence models, discussed in Section \ref{bounded}, can be extended to included some notion of stubbornness.
+The bounded confidence models, discussed in Section \ref{bounded}, can be extended to included stubbornness.
 Guillaume Deffuant proposed such an extension where stubbornness is correlated to extremism \cite{Deffuant2006}.
-The model he proposed goes as follows.
-In bounded confidence models, the actors have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence actor $i$ (actors with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing actor $i$).
-Guillaume Deffuant then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or the extreme opinion $1$ \cite{Deffuant2006}.
+In his model, the actors have an opinion $x_i$ (which is a real number between $\interval{-1}{1}$) and a threshold $\epsilon_i$, which determines the range of opinions that can influence actor $i$ (actors with opinions between $\interval{x_i - \epsilon_i}{x_i + \epsilon_i}$ are capable of influencing actor $i$).
+He then introduced a fraction of extremists, $f_e$, which can have either the extreme opinion $-1$ or $+1$ \cite{Deffuant2006}.
 These extremists are given a threshold equal to $\epsilon_e$ \cite{Deffuant2006}.
@@
 \newline
-Since this thesis is not concerned with bounded confidence models, we will not discuss the results of this model here. Interested readers can find them in \cite{Deffuant2006}. \\
-\newline
 Another example of introducing some kind of persistence towards the own opinion of the actors is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
-The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each actor adheres to its initial opinion to a certain degree can also be seen as introducing a notion of stubbornness \cite{Noorazar2020}.
+The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each actor adheres to its initial opinion to a certain degree can also be seen as introducing a form of stubbornness \cite{Noorazar2020}.
 
@@
 Playing around with the threshold $\epsilon_i$ of the actors to include a notion of stubbornness, such as one can do for the bounded confidence models, is not possible for binary or discrete models.
-There are, however, many other ways to include some kind of stubborn actors in these models.
+There are, however, other ways to include stubborn actors in these models.
 In the following, two different possibilities will be discussed.
@@
 \newline
-One remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
+In bounded confidence models, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
 In binary models, however, there are only two possible opinions and in that sense it is less useful to talk about extreme and moderate opinions.
@@
 \newline
-\textbf{Including a resistance parameter}\\
+\textbf{Resistance parameter}\\
 \newline
@@
 This resistance parameter can have values between 0 and 1.
-If $r_i = 0$, the actor is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority).
-If $r_i = 1$, the actor is completely stubborn and there is no chance of convincing him/here of the other idea.
-If $0 < r_i < 1$, the actor has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so.
+If $r_i = 0$, the actor is not stubborn and can be easily convinced by the other opinion (according to the opinion dynamics model that is implemented: majority, probabilistic majority).
+If $r_i = 1$, the actor is completely stubborn and there is no chance of convincing it of the other idea.
+If $0 < r_i < 1$, the actor has some resistance but can still be convinced to switch to the other opinion, even though it will be less likely to do so.
 An uniform random number $f$ between 0 and 1 is drawn.
-If $f > r_i$, the actor will update his/her opinion; if $f \leqslant r_i$, the actor is not convinced and keeps his/her own opinion.\\
+If $f > r_i$, the actor will update its opinion; if $f \leqslant r_i$, the actor is not convinced and keeps its own opinion.\\
 \newline
 The resistance parameter $r_i$ can be different or the same for all actors.
-One can also play around with the fraction of actors that is stubborn, for example, there can be a fraction of actors that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$).
-Many other possibilities can be found according to how the stubborn actors are distributed and according to how stubborn they are made.
+One can also control the fraction of stubborn actors, for example, there can be a fraction of actors that is completely stubborn ($r = 1$), while the rest is not stubborn ($r = 0$).
+Other possibilities can be found according to how the stubborn actors are distributed and how stubborn they are made.
 This makes the model flexible.\\
@@
 \newline
-In the threshold model one introduces a notion of stubbornness by introducing a threshold parameter $T$.
-This parameter can be either the same for all the actors ($T$) or can be made individual ($T_i$).
-The main idea is that actors can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$).
-The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
+In the threshold model one introduces stubbornness by introducing a threshold parameter $T$.
+This parameter can be either the same for all the actors ($T$) or individualized ($T_i$).
+The main idea is that actors can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is larger than $T$ ($p > T$).
+The idea behind the threshold model is that people are hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
 The more persistent you are, the more friends you need in order to be convinced of the other idea.
@@
 If $T = 1$, the actors do not change their opinion.
-For $0 < T < 1$, the model is a majority model, but the actors are more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
-\newline
-The above examples are only a handful of the possible ways to include some notion of stubbornness in the model, many more exist or can be invented.
-
-% NOTE: maybe make somewhere in thesis link to real-world examples; for example the election system in USA: formation of echo chambers, different states voting for a different party
+For $0 < T < 1$, the model is a majority model, but the actors are more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.
+
+\section{Algorithmic personalization}\label{filter}
+
+People are cognitive and time constrained \cite{Perra2019}; hence, there is only a limited amount of information that humans can process \cite{Bozdag2013}.
+This leads to the fact that on-line social platforms use algorithms to order and filter the posts that appear on the time-line of an individual in order to make the on-line social experience more convenient \cite{Perra2019}.
+If social media companies would not use any kind of filtering, there would be too much content to handle for the social media user \cite{Bozdag2013}.
+This in turn would make the on-line social experience unpleasant and inconvenient; the user might feel overwhelmed by the number of information he or she is exposed to \cite{Bozdag2013}.
+\\
+\newline
+The downside of algorithmic filtering is that it fine-tunes the information it shows based on the user.
+Even two users with the same friends may then be exposed to different content, based on their previous activities \cite{Bozdag2013}.
+This may lead to the formation of echo chambers in which users get `trapped' in content that confirms their own mindset \cite{Bozdag2013}.
+\\
+\newline
+It is important to know and be aware of the possible impact of algorithmic personalization on the formation and evolution of opinions.
+Perra and Rocha partly investigated the role of such algorithms on the dynamics of opinions in several networks \cite{Perra2019}.
+Filtering algorithms are often not transparently shared by social media companies or social platforms.
+Personalization algorithms are based on three filtering strategies \cite{Perra2019}: (i) Popularity filtering: promoting content that is popular across the platform; (ii) Semantic filtering: promoting posts that are similar to previous consumed posts; (iii) Collaborative filtering: promoting content that people similar or close to us consume.\\
+\newline
+The following section will discuss the opinion dynamics model and filtering algorithms adopted in this thesis.
+
+\section{This thesis}\label{modelThesis}
+
+\subsection{Opinion dynamics model}
+
+% introduce dynamics of social platform under study and link to social networks from previous chapter!!
+
+This thesis follows the opinion dynamics model that was designed by Perra and Rocha \cite{Perra2019}.
+We start with a network with $N = 10^3$ actors.
+Initially, each actor is given one of two opinions, A or B.
+Each time step a random fraction equal to $p= 0.1$ of these actors becomes active, where $p$ is the constant activation probability.
+Every time step, all active actors update their opinion synchronously.
+This updating is done as follows.
+First, the active actor broadcasts its current opinion to all its neighbors.
+Then, it updates its own opinion according to the opinions on its time-line $R_i(t)$.
+This time-line contains only a selection of all the opinions received from its friends, which is the fundamental idea behind algorithmic personalization.
+Each user has a hidden time-line $L_i(t)$ which contains the time-line of the actor at the last time it was active ($R_i(t_{\text{last}})$) plus all the post that its friends posted during $\interval[right open fence =), open right]{t_\text{last}}{t}$.
+If an actor becomes active, the filtering algorithm select $S = 20$ items from this hidden time-line $L_i(t)$.
+These $20$ items then form the actual time-line $R_i(t)$ of the actor.
+Updating is done according to the probabilistic majority model: an actor who sees $20\%$ of the posts promoting opinion A and thus $80\%$ promoting opinion B will adopt opinion A with a chance $p_\text{A} = 0.2$ and opinion B with a chance $p_\text{B} = 0.8$.
+If another model is used, e.g.
+the majority model, this will be explicitly stated.
+% Add pictorial representation of model??
+
+\subsection{Filtering algorithms}
+
+Perra and Rocha proposed different ways of selecting the $S$ items from the hidden time-line $L_i(t)$ \cite{Perra2019}.
+We will implement three of their algorithms: (i) REF filtering: $S$ random items from the hidden time-line $L_i(t)$ are selected and show on the actual time-line $R_i(t)$; (ii) REC filtering: the $S$ most recent posts in the hidden time-line $L_i(t)$ are displayed on the actual time-line $R_i(t)$ and (iii) PR filtering: this way of filtering is based on semantic filtering explained in Section \ref{filter}; the post in the hidden time-line $L_i(t)$ are ordered according to the current opinion of the actor (if the actor has opinion A, the hidden time-line will be ordered as AAAA...BBB), then the $S$ first items are selected and shown on the time-line $R_i(t)$.\\
+\newline
+The base-line algorithms are the REF and REC filtering algorithms. Filtering tailored to the user's preferences is accomplished with the PR filtering algorithm.
+
+\subsection{Stubbornness}
+
+If we want to introduce stubbornness in the model we will mainly use the resistance parameter, as explained in Section \ref{stubbBin}.
+Lets briefly summarize this concept of stubbornness.
+Each actor is given a resistance parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update its opinion, if $f \leqslant r_i$, the actor will keep its current opinion.\\
+\newline
+Some results for the majority threshold model, also explained in Section \ref{stubbBin}, will be shared as well.
+This is, however, not the main stubbornness model that we will focus on.
+If this model is used, this will be explicitly stated.
+\\
+\newline
+Some pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
+The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
+
+\begin{algorithm}
+\caption{Updating the opinions of the active actors at time $t$}
+\label{alg:update}
+\begin{algorithmic}[1]
+	\State \(L_i \leftarrow \text{hidden time-line}\)
+	\State \(R_i \leftarrow \text{actual time-line}\)
+	\State \(S \leftarrow 20\) \Comment{Maximum length of time-line $R$}
+	\State \(r\) \Comment{Stubbornness parameter of each actor, takes values between $0$ and $1$}
+	\ForAll{active actor $i$}
+		\State \(\text{op} \leftarrow \text{current opinion}\)
+		\ForAll{neighbor $k$} \Comment{Add opinion of active actor $i$ to hidden time-line of each neighbor $k$}
+			\State \(L_k.\textsc{add}\text{(op)}\)
+		\EndFor
+		\State \(f \leftarrow \text{rand(0, 1)}\) \Comment{Draw random number between $0$ and $1$}
+		\If{$f > r$}
+			\State \(\text{op} \leftarrow \Call{Update opinion}{L_i}\) \Comment{Update opinion of active actor $i$}
+		\Else
+			\State \(\text{op} \leftarrow \text{current opinion}\) \Comment{Do not update opinion of active actor $i$}
+		\EndIf
+	\EndFor
+
+	\Function{Update opinion}{$L_i$}
+		\State \(\textsc{Order}(L_i)\) \Comment{Order hidden time-line according to REF, REC or PR filtering}
+		\If{$\text{len}(L_i) > S$}
+			\State \(R_i \leftarrow L_i.\textsc{resize}(S)\) \Comment{Time-line $R_i$ = $S$ first elements of ordered hidden time-line}
+		\EndIf
+		\State \(\text{op} \leftarrow \textsc{probMaj}(R_i)\) \Comment{Update opinion according to probabilistic majority model}
+	\State \Return op \Comment{Return updated opinion}
+	\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
 \chapter{Materials and methods in this thesis}\label{chap4}
@@
 The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
-In this thesis we focus intensively on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
+In this thesis we focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
@@
 In other words, network models may lack many features of real-world networks.
-It is thus always a bit tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
+It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
 \newline
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{Last.fm}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{lastfm_degree_distribution_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PGP}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{PGP_degree_distribution_8x7.png}
@@
 \end{figure}
-
-\section{Opinion dynamics model}\label{opinionmodel}
-
-In Chapter \ref{chap3} we discussed opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness.
-The goal of this section is to put everything together and summarize the opinion dynamics model of this thesis.\\
-\newline
-This thesis follows the opinion dynamics model that was designed by Perra and Rocha in \cite{Perra2019}.
-We start with a network with $N = 10^3$ users.
-Initially, each user is given one of the two opinions A or B.
-Each time step a random fraction equal to $p= 0.1$ of these users becomes active, where $p$ is the activation probability.
-As discussed in Chapter \ref{chap3}, Section \ref{actMech}, this thesis makes use of a constant activation probability.
-Every time step, all active users update their opinion synchronously.
-This updating is done as follows.
-First, the active user sends his current opinion to all his neighbors (active or not).
-Then, he updates his own opinion according to the opinions on his time-line.
-This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}.
-If another model is used, e.g.
-the majority model, this will be explicitly stated.
-If not, one can assume that the probabilistic majority model is implemented.
-It is important to realize that what an active user sees on his/her time-line is only a selection of all the posts that his/her friends have posted.
-Each user has a hidden time-line $L_i(t)$.
-This hidden time-line contains the time-line of the user at the last time he/she was active ($t_{\text{last}}$) plus all the post that his/her friends posted from $t_{\text{last}}$ until the user becomes active again at time $t$.
-If a user becomes active, the filtering algorithm (REF, REC or PR), discussed in Chapter \ref{chap3}, Section \ref{filter}, selects $S$ items from this hidden time-line and shows them on the actual time-line of the user.
-We will usually set $S = 20$, which quantifies the maximum length of the actual time-line.\\
-
-\begin{algorithm}
-\caption{Updating the opinions of the active actors at time $t$}
-\label{alg:update}
-\begin{algorithmic}[1]
-	\State \(L \leftarrow \text{hidden time-line}\)
-	\State \(T \leftarrow \text{actual time-line}\)
-	\State \(S \leftarrow 20\) \Comment{Maximum length of time-line $T$}
-	\State \(r\) \Comment{Stubbornness parameter of each actor, takes values between $0$ and $1$}
-	\ForAll{active user}
-		\State \(\text{op} \leftarrow \text{current opinion}\)
-		\ForAll{neighbor} \Comment{Add opinion of active user to hidden time-line of each neighbor}
-			\State \(L.\textsc{add}\text{(op)}\)
-		\EndFor
-		\State \(f \leftarrow \text{rand(0, 1)}\) \Comment{Draw random number between $0$ and $1$}
-		\If{$f > r$}
-			\State \(\text{op} \leftarrow \Call{Update opinion}{L}\) \Comment{Update opinion of active user}
-		\Else
-			\State \(\text{op} \leftarrow \text{current opinion}\) \Comment{Do not update opinion of active user}
-		\EndIf
-	\EndFor
-
-	\Function{Update opinion}{$L$}
-		\State \(\textsc{Order}(L)\) \Comment{Order hidden time-line according to REF, REC or PR filtering}
-		\If{$\text{len}(L) > S$}
-			\State \(T \leftarrow L.\textsc{resize}(S)\) \Comment{Time-line $T$ = $S$ first elements of ordered hidden time-line}
-		\EndIf
-		\State \(\text{op} \leftarrow \textsc{probMaj}(T)\) \Comment{Update opinion according to probabilistic majority model}
-	\State \Return op \Comment{Return updated opinion}
-	\EndFunction
-\end{algorithmic}
-\end{algorithm}
-\noindent
-If we want to introduce some kind of stubbornness in the model we will mainly use the stubbornness parameter, as was explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
-Lets briefly summarize this concept of stubbornness.
-Each actor is given a stubbornness parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update his opinion, if $f \leqslant r_i$, the actor will keep his current opinion.\\
-\newline
-Some results for the majority threshold model, also explained in Chapter \ref{chap3}, Section \ref{stubbBin}, will be shared as well.
-This is, however, not the main stubbornness model that we will focus on.
-If this model is used, this will be explicitly stated.
-Otherwise one can assume that we are making use of the stubbornness parameter.\\
-\newline
-In Algorithm \ref{alg:update}, the pseudo-code for updating the opinions of the active actors in case of using a stubbornness parameter $r$ at time step $t$ is displayed. \\
-\newline
-This model is a simple one, since it only allows for two different opinions in the system.
-The designed filtering algorithms are simple and intuitive as well.
-Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
-This might make the model more realistic.
-This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.
-A simple model is more tractable and allows for a better understanding of what is going on in the dynamics \cite{Jed2019}.
-However, this might be a downside as well.
-Real-world opinion dynamics are, probably, far from simple.
-People do not update their opinion according to a simple probabilistic opinion dynamics model.
-Beside that, the cases where people have to choose between only two opinions are rather limited.
-Filtering algorithms, used by social media companies, are unfortunately not publicly shared.
-We can however assume that they are not as simple as the ones we used in this thesis.
-Trying to model these far more complex systems by our simple model, might lead to the fact that part of the dynamics might be missed or not well captured by our model.
-This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
-\\
-\newline
-This does not mean that simple models should be thrown away.
-They might still be capable of capturing the dynamics that are seen in the real world \cite{Gug2015}.
-Not because a system appears complex, that it cannot be modeled by something more simple \cite{Jed2019}.
-This has been proven by physics research over and over again \cite{Gug2015}.
-Beside that, if you do not know the dynamics of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
-There is no need to make things too complex if simple works too.
-There is however a need to compare data from (computational) models to empirical, real-world data.
-A need for critical thinking and careful analyzing of results can also not be denied \cite{Pawel}.
-% put this section here?? Reference??
 
@@
 The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} is used.
-We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
+We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each actor has a maximum length equal to $S=20$.\\
 \newline
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10x100_10-001-0001_50-50_PR_REC.png}
@@
 The value shown on the $y$-axis, $F_N(\left<P_{\text{B}}^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The region on the $x$-axis close to zero describes neighborhoods in which none, or few users, adopt opinion B (these are thus neighborhoods in which most users adopt opinion A).
-Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion B.
+The region on the $x$-axis close to zero describes neighborhoods in which none, or few actors, adopt opinion B (these are thus neighborhoods in which most actors adopt opinion A).
+Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the actors adopt opinion B.
 The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}.
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_PR_8x7.png}
@@
 The initial prevalence of opinion B is $P_\text{B}(0) = 0.8$, hence this is the majority opinion.
-The opinions are, initially, randomly distributed over the system.\\ %Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
+The opinions are, initially, randomly distributed over the system.\\ %Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each actor has a maximum length $S = 20$.\\
 \newline
@@
 This decrease is a consequence of the PR filtering algorithm.
-Since the PR filtering algorithm orders the posts according to the users' preferences, actors with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
+Since the PR filtering algorithm orders the posts according to the actors' preferences, actors with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
 Actors with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
@@
   \begin{subfigure}[t]{0.49\textwidth}
-    \subcaption{ER}
+    \subcaption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
@@
   \begin{subfigure}[t]{0.49\textwidth}
-    \subcaption{WS}
+    \subcaption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
@@
   \begin{subfigure}[t]{0.49\textwidth}    
-    \subcaption{SBM}
+    \subcaption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
@@
   \begin{subfigure}[t]{0.49\textwidth}
-    \subcaption{SBM-WS}
+    \subcaption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{$N = 10^3$}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{$N = 10^4$}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
@@
   \begin{subfigure}[t]{0.49\textwidth}
-    \caption{REF}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
@@
   \begin{subfigure}[t]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
@@
 In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion).
-This can be seen as follows: the one actor with opinion B is likely to switch to opinion A, since all his friends carry this opinion and hence, this opinion will be strongly pushed on his time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
+This can be seen as follows: the one actor with opinion B is likely to switch to opinion A, since all its friends carry this opinion and hence, this opinion will be strongly pushed on its time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
 The two actors with opinion A will have more chance of keeping that opinion than switching to opinion B in case of the PR filtering algorithm.
@@
 For the REF and REC filtering algorithms this chance is however smaller, which can be seen as follows.
-The actor with opinion B still has a high chance of switching to opinion A, because all his friends (which are the two other actors in the triangle) carry this opinion.
+The actor with opinion B still has a high chance of switching to opinion A, because all its friends (which are the two other actors in the triangle) carry this opinion.
 However, the two actors with opinion A will now not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REF}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REF_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_all_frac_stubb_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_REC_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
@@
 This model is explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
-We do not make use of a probabilistic majority model, but a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
+We do not use a probabilistic majority model, but a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
 The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
-The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the users is $S=20$.
+The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the actors is $S=20$.
 The threshold parameter $T$ is equal for all the actors and varies between $T=0$ and $T=1$.\\
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REF}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REF_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_threshold_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{ER}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_majT=0_one_sim_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{WS}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_majT=0_one_sim_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_majT=0_one_sim_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{SBM-WS}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_majT=0_one_sim_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{High modularity, REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_high_mod_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{High modularity, PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_high_mod_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-   \caption{Low modularity, REC}
+   \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_REC_low_mod_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{Low modularity, PR}
+    \caption{}
     \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0=01_other=50-50_PR_low_mod_8x7.png}
@@
 The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was explained for Fig.
-\ref{commOp0_01_other_50-50} and has to do with whether or not the single opinion A communities can influence the $50/50$ communities.
+\ref{commOp0_01_other_50-50} and has to do with whether the single opinion A communities can influence the $50/50$ communities.
 
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_commOp0_other=50-50_PR_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm01_distr_PR_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_powerlaw_vs_regular_REC.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_powerlaw_vs_regular_PR.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{PR}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_PR_real_network_PGP_8x7.png}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \caption{REC}
+    \caption{}
   	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_fracRes_REC_real_network_PGP_8x7.png}
@@
 
+ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS
+
+''' Perra and Rocha chose activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
+They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
+However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}.
+\\
+\newline
+ It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics. '''
+
++ ADD THIS (about opinion dynamics model)
+
+''' This model is a simple one, since it only allows for two different opinions in the system.
+The designed filtering algorithms are simple and intuitive as well.
+Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
+This might make the model more realistic.
+This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.
+A simple model is more tractable and allows for a better understanding of what is going on in the dynamics \cite{Jed2019}.
+However, this might be a downside as well.
+Real-world opinion dynamics are, probably, far from simple.
+People do not update their opinion according to a simple probabilistic opinion dynamics model.
+Beside that, the cases where people have to choose between only two opinions are rather limited.
+Filtering algorithms, used by social media companies, are unfortunately not publicly shared.
+We can however assume that they are not as simple as the ones we used in this thesis.
+Trying to model these far more complex systems by our simple model, might lead to the fact that part of the dynamics might be missed or not well captured by our model.
+This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
+\\
+\newline
+This does not mean that simple models should be thrown away.
+They might still be capable of capturing the dynamics that are seen in the real world \cite{Gug2015}.
+Not because a system appears complex, that it cannot be modeled by something more simple \cite{Jed2019}.
+This has been proven by physics research over and over again \cite{Gug2015}.
+Beside that, if you do not know the dynamics of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
+There is no need to make things too complex if simple works too.
+There is however a need to compare data from (computational) models to empirical, real-world data.
+A need for critical thinking and careful analyzing of results can also not be denied \cite{Pawel}.
+'''
+
 \chapter{Conclusion}
@@
 \end{equation}
-In order to simplify this equation we can make use of the following formula \cite{Hopcroft2006}
+In order to simplify this equation we use the following formula \cite{Hopcroft2006}
 \begin{equation} 

### 2021-05-21 draft.tex
Update draft.tex
changed section 3.6

@@
 
+We want to model opinion dynamics on social media platforms, with focus on Facebook.
+Each time we access Facebook we experience algorithmic personalization.
+The time-line or News Feed of each user on Facebook is the complex combination of posts from, among other, Facebook friends, companies and mass media.
+In 2013, Facebook reported that on average 1500 potential stories of friends and pages we follow or companies that are sponsored could be shown on the News Feed of each user.
+% add ref Facebook
+No user could ever process this amount of information, hence, Facebook selects around 300 of them, which then appear on your personalized News Feed.
+Their selection methods are based on (i) how often we interact with certain friends, pages or prominent figures; (ii) the amount of likes, shares and comments a post receives from your friends and others; (iii) how much you have interacted with certain types of posts before and (iv) whether you and other people on Facebook are hiding or reporting a certain post.
+\\
+\newline
+If we translate Facebook to a social network, we end up with a temporal network.
+Nodes represent the Facebook users, which become active and inactive over time.
+Edges represent friendship ties on Facebook, that is, they link an user to his Facebook friends and the pages he or she follows.
+Each node then has an attribute that needs to mimic the Facebook News Feed and which is susceptible to the filtering algorithm.
+
 \subsection{Opinion dynamics model}
-
-% introduce dynamics of social platform under study and link to social networks from previous chapter!!
 
@@
 Each time step a random fraction equal to $p= 0.1$ of these actors becomes active, where $p$ is the constant activation probability.
-Every time step, all active actors update their opinion synchronously.
+Then, all active actors update their opinion synchronously.
 This updating is done as follows.
@@
 This time-line contains only a selection of all the opinions received from its friends, which is the fundamental idea behind algorithmic personalization.
-Each user has a hidden time-line $L_i(t)$ which contains the time-line of the actor at the last time it was active ($R_i(t_{\text{last}})$) plus all the post that its friends posted during $\interval[right open fence =), open right]{t_\text{last}}{t}$.
-If an actor becomes active, the filtering algorithm select $S = 20$ items from this hidden time-line $L_i(t)$.
-These $20$ items then form the actual time-line $R_i(t)$ of the actor.
+Each actor has a hidden time-line $L_i(t)$ which contains the time-line of the actor at the last time it was active ($R_i(t_{\text{last}})$) plus all the post that its friends posted during $\interval[right open fence =), open right]{t_\text{last}}{t}$.
+If an actor becomes active, the filtering algorithm select $S = 20$ items from this hidden time-line $L_i(t)$ and shows them on the time-line $R_i(t)$ of the actor.
 Updating is done according to the probabilistic majority model: an actor who sees $20\%$ of the posts promoting opinion A and thus $80\%$ promoting opinion B will adopt opinion A with a chance $p_\text{A} = 0.2$ and opinion B with a chance $p_\text{B} = 0.8$.
@@
 Perra and Rocha proposed different ways of selecting the $S$ items from the hidden time-line $L_i(t)$ \cite{Perra2019}.
-We will implement three of their algorithms: (i) REF filtering: $S$ random items from the hidden time-line $L_i(t)$ are selected and show on the actual time-line $R_i(t)$; (ii) REC filtering: the $S$ most recent posts in the hidden time-line $L_i(t)$ are displayed on the actual time-line $R_i(t)$ and (iii) PR filtering: this way of filtering is based on semantic filtering explained in Section \ref{filter}; the post in the hidden time-line $L_i(t)$ are ordered according to the current opinion of the actor (if the actor has opinion A, the hidden time-line will be ordered as AAAA...BBB), then the $S$ first items are selected and shown on the time-line $R_i(t)$.\\
+We will implement three of their algorithms: (i) REF filtering: $S$ random items from the hidden time-line $L_i(t)$ are selected and shown on the actual time-line $R_i(t)$; (ii) REC filtering: the $S$ most recent posts in the hidden time-line $L_i(t)$ are displayed on the actual time-line $R_i(t)$ and (iii) PR filtering: this way of filtering is based on semantic filtering explained in Section \ref{filter}; the posts on the hidden time-line $L_i(t)$ are ordered according to the current opinion of the actor (if the actor has opinion A, the hidden time-line will be ordered as AAAA...BBB), then the $S$ first items are selected and shown on the time-line $R_i(t)$.\\
 \newline
@@
 Some results for the majority threshold model, also explained in Section \ref{stubbBin}, will be shared as well.
-This is, however, not the main stubbornness model that we will focus on.
+This is not the main stubbornness model that we will focus on.
 If this model is used, this will be explicitly stated.

### 2021-05-21 draft.tex
Update draft.tex
made changes in Ch.2 temporal networks+ Ch.3

@@
 In temporal networks, the edges (or nodes) are not continuously active \cite{Holme2012}.
-Instead, edges (or nodes) become active and inactive over time and time becomes an explicit element of the network representation \cite{Holme2012}.
-An example of such temporal networks in real life are communication networks such as the phone network where edges represent instantaneous contacts \cite{Holme2012}.
+Instead, edges (or nodes) become active and inactive over time, which becomes an explicit element of the network representation \cite{Holme2012}.
+An example of temporal networks in real life are communication networks such as the phone network where edges represent instantaneous contacts \cite{Holme2012}.
 Other examples of applications of temporal networks are the application to the dynamics of functional brain connectivity \cite{Thompson2017} and the application to infrastructure networks \cite{Salama2020}.\\
@@
 These models aim to study whether the distribution of the stubborns (clustered versus random) matters; whether stubborn groups can lead to polarization or can overtake the opinions in the system or, contrary, whether stubbornness is protective against extremism and polarization.
-We will review some of the most important models.
+We will review the most important models.
 
@@
 Each actor is given a resistance parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update its opinion, if $f \leqslant r_i$, the actor will keep its current opinion.\\
-\newline
-Some results for the majority threshold model, also explained in Section \ref{stubbBin}, will be shared as well.
-This is not the main stubbornness model that we will focus on.
-If this model is used, this will be explicitly stated.
-\\
-\newline
-Some pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
-The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update its opinion, if $f \leqslant r_i$, the actor will keep its current opinion.
 
@@
 \end{algorithm}
+\noindent
+Some pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
+The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
 

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.2, erdos-renyi model

@@
 Since for large $N$, $\ln{N} \ll N$, we find that the average distance is smaller than the size of the network.
-Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of the ``small world phenomenon''; the other property is high clustering \cite{Easley2010}.
+Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of the ``small world phenomenon''; the other property is high clustering \cite{Easley2010}.\\
+\newline
+\textbf{Erd\H{o}s-R\'{e}nyi model in this thesis}\\
+\newline
+The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
+We focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
+It is therefore necessary to have a model that lacks both of these features to compare the results with.
+The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the appropriate comparison model.
+It is also used in the regular stochastic block model to construct the constituent communities.\\
+\newline
+It is expected that the random nature of the model and the lack of high clustering and community structure will result in the absence of echo chambers.
 
@@
 
-This thesis uses three network models: (i) the Erd\H{o}s-R\'{e}nyi model (ER); (ii) the Watts-Strogatz model (WS); and (iii) the stochastic block model, of which two types are implemented: the regular stochastic block model (SBM) and the stochastic Watts-Strogatz block model (SBM-WS).\\
-\newline
-Each of these networks serve specific goals which will be shortly discussed in the following paragraphs.\\
-\newline
-\textbf{The Erd\H{o}s-R\'{e}nyi model}\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
-In this thesis we focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
-It is therefore necessary to have a model that lacks both of these features to compare the results with.
-The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the perfect comparison model.
-\\
-\newline
-It is expected that the random nature of the model and the lack of high clustering and community structure will result in the absence of echo chambers.\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model is also used in the regular stochastic block model to construct the constituent communities.\\
-\newline
 \textbf{The Watts-Strogatz model}\\

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.2 Watts-Strogatz model

@@
 Real social networks, however, usually display the ``small world phenomenon'' and thus have both short average distances and high clustering.
-The question then rises: Can we construct a model that has both of these properties?
+The question then rises whether we can construct a model that has both of these properties.
 The answer is yes and this is what Duncan Watts and Steve Strogatz did in 1998.
@@
 The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
-The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.
+The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.\\
+\newline
+\textbf{The Watts-Strogatz model in this thesis}\\
+\newline
+The Watts-Strogatz model is used to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
+\newline
+Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
+Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.
+The Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.
 
@@
 
-\textbf{The Watts-Strogatz model}\\
-\newline
-The Watts-Strogatz model is used as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
-\newline
-Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
-Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
-We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.\\
-\newline
-Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
-\newline
 \textbf{The stochastic block model}\\
@@
 The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
-We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
-\newline
-All the above models have a homogeneous degree distribution.
-We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
-It might, however, be an interesting extension to investigate such types of models as well.\\
-\newline
-The above models have their added value and benefits.
-One of their most important plus points is their simplicity.
-They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM).
-This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account.
-This plus side is, however, also one of the possible downsides.
-Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
-In other words, network models may lack many features of real-world networks.
-It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
-\newline
-The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.
+We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).
 
@@
 
-ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS
+ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS (about activation probabilities)
 
@@
 
++ ADD THIS (limitations of network models)
+
+'''All the above models have a homogeneous degree distribution.
+We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
+It might, however, be an interesting extension to investigate such types of models as well.\\
+\newline
+The above models have their added value and benefits.
+One of their most important plus points is their simplicity.
+They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM).
+This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account.
+This plus side is, however, also one of the possible downsides.
+Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
+In other words, network models may lack many features of real-world networks.
+It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
+\newline
+The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.'''
+
 \chapter{Conclusion}

### 2021-05-21 draft.tex
Update draft.tex
small change in Ch. 2 Watts-Strogatz

@@
 \newline
-Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}.
-Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient.
+Perra and Rocha showed that high clustering enhances the formation of echo chambers \cite{Perra2019}.
+Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses community structure without having high clustering.
 We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.2 SBM

@@
 The result is a network that can be tuned to have both a community structure and a high clustering coefficient.
-This is different from the regular stochastic block model that can only account for a community structure and lacks high clustering.
-% talk about community sizes from power law??
+This is different from the regular stochastic block model that can only account for a community structure and lacks high clustering.\\
+\newline
+\textbf{The stochastic block model in this thesis}\\
+\newline
+The stochastic block model is used to generate networks with community structure.
+We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
+\newline
+It is expected that community structure will enhance the formation of echo chambers compared to the random network case.
+The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering.
+It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
+\newline
+The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
+We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).
 
@@
 \chapter{Materials and methods in this thesis}\label{chap4}
-
-The previous two chapters, Chapter \ref{chap2} and \ref{chap3}, gave a broad overview of networks and opinion dynamics models respectively.
-In this chapter we will summarize the specific networks and opinion dynamics models that are used in this thesis.
-\\
-\newline
-Section \ref{networkmodel} gives an overview of the network models that are used.
-In chapter \ref{chap2}, Section \ref{netModel} we discussed their main properties; here we will discuss why we use these networks.
-Which of their properties are of interest in this thesis?
-What are their benefits and possible shortcomings regarding to our goals?\\
-\newline
-In section \ref{realnetwork}, we will discuss the used real-world networks.
-The reasons of choosing these specific networks will be explained, together with their possible strengths and weaknesses.\\ % + explain where they come from?
-\newline
-Finally, Section \ref{opinionmodel} summarizes and discusses the opinion dynamics model that is implemented. We will focus on possible advantages and disadvantages of the model.
-
-\section{Network models}\label{networkmodel}
-
-\textbf{The stochastic block model}\\
-\newline
-One of the main features of the stochastic block model is the ability to generate networks with community structure.
-We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
-\newline
-It is expected that community structure will enhance the formation of echo chambers compared to the random network case.
-The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering.
-It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
-\newline
-The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
-We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).
 
@@
 
-ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS (about activation probabilities)
+ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS (about activation probabilities) + PAY ATTENTION TO FEEDBACK (see file)
 

### 2021-05-21 draft.tex
Update draft.tex
Changed Ch.2 Real-world social networks + removed chapter 4

@@
 Other examples of social networks are sexual networks, in which the edges denote sexual contacts between the individuals and email networks, in which directed edges represent an email send from individual $i$ to individual $j$ \cite{Costa2008}.
-There are also plenty of other real-world network such as collaboration networks, where the nodes are usually scientists and edges indicate professional ties between them; citation networks, where nodes represent scientific papers and the directed edges report citations and protein-protein interaction networks, in which nodes stand for proteins and edges indicate physical interactions between them \cite{Costa2008}.
-These examples of real-world networks are only a handful.
-A comprehensive review can be found in \cite{Costa2008}.\\
+There are also plenty of other real-world network such as collaboration networks, where the nodes are usually scientists and edges indicate professional ties between them; citation networks, where nodes represent scientific papers and the directed edges report citations and protein-protein interaction networks, in which nodes stand for proteins and edges indicate physical interactions between them \cite{Costa2008}.\\
 \newline
@@
 If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance in the limit of infinite system size ($N \rightarrow \infty$) \cite{Newman2005}.
-Real networks always have finite sizes, so there is always a finite average degree and variance.
-This variance would, however, be meaningless \cite{Newman2005}.
+Real networks always have finite sizes, so they have a finite average degree and variance.
+This variance is, however, meaningless \cite{Newman2005}.
 A diverging variance means that the fluctuations around the average can be arbitrary large \cite{Barabasi2016}.\\
@@
 In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his or her importance and influence by the number of fans \cite{Hu2009}\cite{Zhang2014}.
-This leads to a disassortative mixing pattern.
+This leads to a disassortative mixing pattern.\\
+\newline
+\textbf{Real-world social networks in this thesis}\\
+\newline
+Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies.
+However, most of these companies are not to keen on sharing any form of data.
+This makes it hard to find accurate and complete social media network data.
+Hence, we chose to test the model on other types of social networks, either on-line or off-line, e.g.
+friendship networks, e-mail networks or collaboration networks.
+\\
+\newline
+We used two real-world networks.
+The first real-world network is an on-line friendship network among Finnish users of the social music listening and sharing website Last.fm (data from $2006$).
+Only the largest connected component of this on-line social network is included \cite{ICON}.
+This data set was originally used in \cite{Toivonen2009}.
+This on-line social network has size $N = 8003$; number of edges: $16824$; $Q = 0.79$; $\left<cc\right> = 0.31$; $\left<k\right> = 4.2$.
+The degree distribution is heterogeneous and is best fitted by a log-normal
+\begin{equation}
+	f(k) = \frac{1}{k \sigma \sqrt{2\pi}} \exp(-\frac{(\ln(k) - \mu)^2}{2 \sigma^2})\ ,
+\end{equation}
+with $\mu = 1.008$ and $\sigma = 0.95$; see Fig. \ref{deg_distr_lastfm}. Nodes indicate Finnish users of the Last.fm website, edges represent friendship ties between them.\\
+\newline
+The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
+This network has $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
+The network has a heterogeneous degree distribution, which is best fitted by two power-laws
+\begin{equation}
+ f(k) = a k^{-\gamma} \ ,
+\end{equation}
+with $a = 0.39$, $\gamma = 1.54$ for $k < 20$ and $a = 5.59$, $\gamma = 2.49$ for $k \geqslant 20$; see Fig.
+\ref{deg_distr_pgp}.
+In this network nodes represent peers that have signed public keys of another based on trust, edges indicate links between peers that have mutually signed their keys \cite{Boguna2004}.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+    \includegraphics[width=8cm, height=7cm]{lastfm_degree_distribution_8x7.png}
+    \label{deg_distr_lastfm}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+    \includegraphics[width=8cm, height=7cm]{PGP_degree_distribution_8x7.png}
+    \label{deg_distr_pgp}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real-world networks used in this thesis.
+\textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.008$, $\sigma = 0.95$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39$, $\gamma = 1.54$ for $k < 20$ and $a = 5.59$, $\gamma = 2.49$ for $k \geqslant 20$).
+Curves are fitted to the data with \textsc{scipy.optimize.curve$\_$fit}.}
+\label{deg_distr_real_network}
+\end{figure}
+\noindent
+One of the decisive properties of the real-world networks that were used, was their size.
+The main, and actually only, reason for this was the computation time.
+If the networks are too big, it would take weeks, sometimes even months before there would be results.
+The size of the network models that were implemented, was around a thousand nodes.
+For the real-world networks, we also used networks with a size of magnitude of thousand to ten-thousand.
+This benefits the computation time.
 
@@
 The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
-
-\chapter{Materials and methods in this thesis}\label{chap4}
-
-\section{Real-world networks}\label{realnetwork}
-% talk about reasons why you chose the networks you chose: e.g. size (computation time),...
-
-Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies.
-However, most of these companies are not to keen on sharing any form of data.
-This makes it hard to find accurate and complete social media network data.
-Hence, we chose to test the model on other types of social networks, either on-line or off-line.
-Some possibilities are any form of friendship networks, e-mail networks or collaboration networks.
-\\
-\newline
-One of the decisive properties of the real-world networks that were eventually used, was their size.
-The main, and actually only, reason for this was the computation time.
-If the networks are too big, it would take weeks, sometimes even months before there would be results.\\
-\newline
-The size of the network models that were used, was around a thousand nodes.
-For the real-world networks, we also used networks with a size of magnitude of thousand to ten-thousand.
-This benefits the computation time; it is, however, unknown how the network size affects the results.
-It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic.
-\\
-\newline
-We used two real-world networks.
-The data was obtained from \textit{The index of Complex networks} (ICON) \cite{ICON}.
-The first real-world network is an on-line friendship network among Finnish users of the social music listening and sharing website Last.fm (data from $2006$).
-Only the largest connected component of this on-line social network is included \cite{ICON}.
-This data set was originally used in \cite{Toivonen2009}, which describes the data set.
-This on-line social network has the following properties: $N = 8003$; number of edges: $16824$; $Q = 0.79$; $\left<cc\right> = 0.31$; $\left<k\right> = 4.2$.
-The degree distribution is heterogeneous and is best fitted by a log-normal
-\begin{equation}
-	f(k) = \frac{1}{k \sigma \sqrt{2\pi}} \exp(-\frac{(\ln(k) - \mu)^2}{2 \sigma^2})\ ,
-\end{equation}
-with $\mu = 1.0078272$ and $\sigma = 0.95173779$. The degree distribution of the Last.fm social network, together with the log-normal fit, is shown in Fig. \ref{deg_distr_lastfm}.\\
-\newline
-The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{ICON}.
-This network was originally used in \cite{Boguna2004}, which gives a detailed description of the network data.
-The main properties of this network are: $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
-The network has a heterogeneous degree distribution, which is best fitted by two power-laws
-\begin{equation}
- f(k) = a k^{-\gamma} \ ,
-\end{equation}
-with $a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$.
-The degree distribution for the PGP trust network, together with the power-law fits, is shown in Fig.
-\ref{deg_distr_pgp}.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-    \includegraphics[width=8cm, height=7cm]{lastfm_degree_distribution_8x7.png}
-    \label{deg_distr_lastfm}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-    \includegraphics[width=8cm, height=7cm]{PGP_degree_distribution_8x7.png}
-    \label{deg_distr_pgp}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real-world networks used in this thesis.
-\textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.0078272$, $\sigma = 0.95173779$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$).}
-\label{deg_distr_real_network}
-\end{figure}
 
@@
 
++ ADD THIS (limitations on size) --> needed? or already included?
+
+'''it is, however, unknown how the network size affects the results.
+It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic'''
+
 \chapter{Conclusion}

### 2021-05-21 draft.tex
Update draft.tex
made changes in samenvatting

@@
 Goed en kwaad gaan echter vaak hand in hand en men kan niet ontkennen dat het gebruik van sociale media ook veel potentiële, soms ongekende, nadelen kent.
-Eén van die, tot nog toe, ongekende kanten van sociale media is de impact die het heeft op de dynamica van opinies in de maatschappij.
-\\
-\newline
-Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving, hun opvoeding.
-Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke en/of tijdsgrenzen.
+Eén van die, tot nog toe, ongekende kanten van sociale media is de impact op de dynamica van opinies in de maatschappij.
+\\
+\newline
+Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving en hun opvoeding.
+Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke- en tijdsgrenzen.
 Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\
 \newline
-Door het toenemende gebruik van sociale media worden deze echter overspoeld door informatie.
+Door het toenemende gebruik worden sociale media echter overspoeld door informatie.
 Om hun gebruikers niet het gevoel te geven dat ze verdrinken in wat ze te zien krijgen op hun sociale media platform, gebruiken deze sociale media platformen filtering algoritmen.
-Deze filtering algoritmen filteren, zoals de naam al impliceert, en individualiseren de informatie die iemand te zien krijgt op zijn/haar tijdslijn.
-Een gebruiker op sociale media krijgt dus niet alle berichten die zijn/haar vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn/haar sociale media geschiedenis.
+Deze filtering algoritmen filteren, zoals de naam al impliceert, en individualiseren de informatie die iemand te zien krijgt op zijn of haar tijdslijn.
+Een gebruiker op sociale media krijgt dus niet alle berichten die zijn of haar vrienden posten te zien, maar enkel een selectie die onder andere gemaakt wordt op basis van de gebruiker zijn of haar sociale media geschiedenis.
 Dit filteren en individualiseren kan, mogelijks, grote gevolgen hebben.

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.5 network structure

@@
 We first discuss the impact of the network structure on the evolution of the prevalence of the opinions and the formation of echo chambers.
-Then, we introduce some kind of stubbornness in the system.
+Then, we introduce stubbornness in the system.
 The impact of how the opinions are initially distributed in a network with community structure (SBM) will be discussed afterwards.
-Finally, we will run the opinion dynamics model of this thesis on some real-world networks.
-We will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or modularity.
+Finally, we will study the opinion dynamics model in real-world networks.
+We will discuss the results and compare them to results from network models with a similar average degree, clustering or modularity.\\
+\newline
+Four different network models are studied: (i) Erd\H{o}s-R\'{e}nyi model (ER); (ii) Watts-Strogatz model (WS); (iii) stochastic block model (SBM) and (iv) stochastic Watts-Strogatz block model (SBM-WS).
+Each model has $N = 1000$ nodes and average degree $\left<k\right> \sim 10$.
+The ER model has an edge probability $p = 0.01$.
+For the WS model, we use two rewiring probabilities: $\beta = 0.01$ ($\left<cc\right> = 0.648 \pm 0.003$) and $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+The SBM has 10 communities of size 100 nodes each.
+We study one configuration with high modularity ($Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$) and one configuration with low modularity ($Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$).
+The stochastic Watts-Strogatz block model has 10 communities of size 100 nodes each.
+Each community has Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
+Edges are added between communities with a probability $p_{\text{add}} = 0.001$ ($\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$).\\
+\newline
+The opinion dynamics model discussed in Section \ref{modelThesis} is used.
+We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each actor has a maximum length equal to $S=20$.
+If another model is used or if parameters (either those of the network models or the opinion dynamics model) are changed, this will be explicitly stated in the relevant sections.
+Opinions and stubborns (if present) are randomly distributed over the network, unless specified otherwise.
 
@@
 
-In this section we will discuss the results related to the network topologies of the different network models that are used in this thesis.
-We will focus on the effect of a high clustering coefficient and community structure.
-Two different opinion distributions are investigated, balanced starting conditions and unbalanced starting conditions.
+Real-world social networks often have high clustering and modular structure.
+We want to study the impact of these topologies on the dynamics of opinions.
+In this section, we will discuss the results related to different network topologies with focus on the effect of clustering and community structure.
+Two different opinion distributions are investigated, balanced and unbalanced starting conditions.
 For both starting conditions we investigate the evolution of the prevalence of the opinions with time and the formation of echo chambers.
-Two filtering algorithms will be compared to each other, the REC and the PR filtering algorithms, in order to investigate the possible interplay between those filtering algorithms and the network structures.\\
-%\newline
-%Before sharing the results, a quick disclaimer: Some of the results that are shared in this section, were already obtained by prof.
-Nicola Perra and prof.
-Luis E.C.
-Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} (see reference \cite{Perra2019}).
-These results include the effect of a high clustering coefficient (Watts-Strogatz model) and a random network model.
-Similar results (albeit with a different network size and average degree) are included in this section to give a complete overview and to be able to compare a community structure to a high clustering coefficient.
+Two filtering algorithms will be compared, the REC and the PR filtering algorithms, in order to investigate the possible interplay between filtering algorithms and network structure.
 
@@
 \label{50-50}
-In this configuration $50 \%$ of the actors start with opinion A ($P_{\text{A}}(0) = 0.5$) and $50 \%$ of the actors start with opinion B ($P_{\text{B}}(0) = 0.5$).
-The opinions are randomly distributed over the network.
-The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} is used.
-We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each actor has a maximum length equal to $S=20$.\\
-\newline
-Four different network models are implemented, each with a total number of nodes $N = 10^3$ and average degree $\left<k\right> \sim 10$.
-The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
-To investigate the effect of clustering on the dynamics, we use the Watts-Strogatz model (WS).
-Two different values of the rewire probability $\beta$ are used, each with a slightly different value of the clustering coefficient: $\beta = 0.01$ with $\left<cc\right> = 0.648 \pm 0.003$ and $\beta = 0.06$ with $\left<cc\right> = 0.557 \pm 0.007$.
-The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are investigated: a high modularity, $Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
-Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
-Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and high modularity ($Q = 0.909 \pm 0.004$).
-\\
+
+In this configuration $50 \%$ of the actors start with opinion A ($P_{\text{A}}(0) = 0.5$) and $50 \%$ of the actors start with opinion B ($P_{\text{B}}(0) = 0.5$).\\
 \newline
@@
 We start with the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution.
-Perra and Rocha showed that, for the case of balanced starting conditions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the filtering algorithms \cite{Perra2019}.
+For both the REC and the PR filtering algorithms, the community structure is not able to break the status quo for both of the filtering algorithms.
+The initial prevalence of the opinions is maintained over time (Fig.
+\ref{ev_op_50_50}).
+This is also the case for networks with high clustering \cite{Perra2019}.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for an initial $50/50$ distribution.
-Two different network models with community structure are represented, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(b)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+  \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for balanced starting conditions.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for balanced starting conditions.
+\textbf{(a)} SBM; \textbf{(b)} SBM-WS.
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
 For each model the REC and the PR filtering algorithms are compared.
+The dotted black line represents the initial prevalence of opinion A.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 \end{figure}
+\newpage
 \noindent
-The results for the network model with community structure (SBM, Fig.
-\ref{sbm_50-50}) and the network model with both community structure and high clustering (SBM-WS, Fig.
-\ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithms.
-A community structure is also not able to break the status quo for both of the filtering algorithms.
-The initial prevalence of the opinions is maintained over time.\\
-\newline
 \textbf{Echo chambers}\\
@@
 \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$.
-The value shown on the $y$-axis, $F_N(\left<P_{\text{B}}^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
 The region on the $x$-axis close to zero describes neighborhoods in which none, or few actors, adopt opinion B (these are thus neighborhoods in which most actors adopt opinion A).
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_{\text{B}}^{\text{nn}}\right>$, for balanced starting conditions.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$, for balanced starting conditions.
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Four network models are used: (i) ER; (ii) SBM; (iii) SBM-WS and (iv) WS.
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+For the WS model, we use two different values for the rewire probability: $\beta = 0.01$ ($\left<cc\right> = 0.648 \pm 0.003$) and $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+The dotted black line represents a constant amount of echo chambers over time.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: ER with $p = 0.01$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; WS with $\left<k\right> = 10$, two slightly different values of the clustering coefficient are used ($\left<cc\right> = 0.557 \pm 0.007$ and $\left<cc\right> = 0.648 \pm 0.003$).
-The starting conditions are $P_{\text{A}}(0) = 0.5$ and $P_{\text{B}}(0) = 0.5$.
 Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for an improved visualization.}
+Error bars are omitted for improved visualization.}
 \label{echo_50_50}
@@
 The results for both filtering algorithms are qualitatively the same.
-The main difference is that the formation of echo chambers, using the PR filtering algorithm, is about ten times as large compared to the formation of echo chambers using the REC filtering algorithm.
-\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model (ER) showns no, or a limited, sign of echo chambers, for both the REC filtering algorithm and the PR filtering algorithm.
-Networks with high clustering (WS and SBM-WS), on the other hand, show a clear sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and one around one).
-Going to slightly higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the Watts-Strogatz model results in significantly larger echo chambers.
-This suggests that the formation of echo chambers is really sensitive to changes in the clustering coefficient.
-The stochastic block model (SBM) with a high modularity, a clear community structure, does show an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller compared to networks with high clustering.
-The stochastic block model with a low modularity has no, or limited, formation of echo chambers, similar as for the Erd\H{o}s-R\'{e}nyi
-model.
-This is to be expected, since a stochastic block model with a low modularity, and hence a blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network.
-The stochastic Watts-Strogatz block model (SBM-WS) has both community structure and high average clustering.
-The formation of echo chambers when using this model is approximately of the same size as the formation of echo chambers in a Watts-Strogatz model with a similar average clustering.
-This suggests that it is mainly the high clustering that drives the system to form echo chambers, while the community structure plays a less important role.
-However, as said before, the stochastic block model with a high modularity also has a (smaller) formation of echo chambers, so having a clear and strong community structure definitely plays some role in phenomena such as polarization and the formation of opinion clusters.
+The difference is that the formation of echo chambers, using the PR filtering algorithm, is about ten times as large compared to the formation of echo chambers using the REC filtering algorithm.
+\\
+\newline
+The ER model shows no, or a limited, sign of echo chambers, for both the REC and PR filtering algorithms.
+Networks with high clustering (WS and SBM-WS), on the other hand, show a sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and another around one).
+Going to slightly higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the WS model results in significantly larger echo chambers.
+This suggests that the formation of echo chambers is sensitive to changes in the clustering coefficient.
+The SBM with high modularity, community structure, shows an appearance of polarization (mainly when using the PR filtering algorithm), however, the formation of echo chambers is significantly smaller compared to networks with high clustering.
+The stochastic block model with low modularity has no, or limited, formation of echo chambers, similar as for the ER model.
+This is expected, since a SBM with low modularity, and hence, almost non-existing community structure, resembles an ER network.
+The SBM-WS has both community structure and high average clustering.
+The formation of echo chambers when using this model is approximately of the same size as the formation of echo chambers in a WS model with similar average clustering.
+This suggests that it is the high clustering that drives the dynamics to form echo chambers, while the community structure plays a less important role.
 \\
@@
 \ref{ev_op_50_50} and related results in \cite{Perra2019}).
-Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a more clustered opinion distribution and thus lead to polarization and formation of echo chambers.
+Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a clustered opinion distribution and thus lead to polarization and formation of echo chambers.
 
@@
 We will now investigate the effect of unbalanced starting conditions, that is one opinion is initially more prevalent than the other.
-The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$, opinion A thus becomes the minority opinion.
-The initial prevalence of opinion B is $P_\text{B}(0) = 0.8$, hence this is the majority opinion.
-The opinions are, initially, randomly distributed over the system.\\ %Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each actor has a maximum length $S = 20$.\\
-\newline
-We study four different network models each with a total number of nodes $N = 10^3$ and average degree $\left<k\right> \sim 10$.
-The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with an edge probability $p = 0.01$.
-To investigate the effect of clustering on the dynamics, we use the Watts-Strogatz model (WS).
-The rewire probability is $\beta = 0.06$, which gives a clustering coefficient equal to $\left<cc\right> = 0.557 \pm 0.007$.
-The impact of community structure on the dynamics is studied with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are investigated: a high modularity, $Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$; and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
-Finally, the impact of both clustering and community structure is investigated with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
-Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-This is a network with both high clustering ($\left<cc\right> = 0.553 \pm 0.004$) and high modularity ($Q = 0.909 \pm 0.004$).\\
+The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$, the minority opinion, and of opinion B is $P_\text{B}(0) = 0.8$, the majority opinion.\\
 \newline
@@
 \ref{ev_op_20_80} shows the results for the evolution of the prevalence of opinion A, which is the minority opinion, over time.
-For each network model two filtering algorithms are compared, the REC filtering algorithm and the PR filtering algorithm.
-\\
-\newline
-The REC filtering algorithm maintains the status quo; the prevalence of opinion A remains stable over time for the four different network models.
-When using the PR filtering algorithm, the results are different.
+For each network model the REC and PR filtering algorithms are compared.
+\\
+\newline
+The REC filtering algorithm maintains the status quo; the prevalence of opinion A remains stable over time for the four network models.
+When using the PR filtering algorithm, results are different.
 The prevalence of the minority opinion, $P_\text{A}(t)$, decreases over time.
@@
 Actors with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
-However, since there are more actors with opinion B than actors with opinion A, the actors with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
+However, because there are more actors with opinion B than actors with opinion A, the actors with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
 When using the PR filtering algorithm, there is an imbalance in the sense that actors with the minority opinion A have a chance of switching to the majority opinion B, whereas actors with the majority opinion B have (almost) no chance of switching to the minority opinion A.
-This imbalance leads to the decrease of the minority opinion A.\\
-\newline
-Fig.
-\ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
-This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
-The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
-\ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Fig.
-\ref{ws_20-80_op}).\\
-\newline
-This result differs from results in \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
-The results that where obtained in \cite{Perra2019} are, however, for networks with a total number of nodes $N = 10^5$ and average degree $\left<k\right> \sim 6$.
-In this thesis results are obtained for $N=10^3$ and average degree $\left<k\right> \sim 10$.
-In both cases the maximum length of the actual time-line is set equal to $S=20$.
-The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
-This can for example be seen in Fig.
-\ref{ev_op_WS_k=6}, which shows the results for the evolution of the prevalence of opinion A for an initial $20/80$ opinion distribution and using the PR filtering algorithm; Fig.
-\ref{ws1000} displays the results for a WS model with $N=10^3,\ \left<k\right>=6$ and $\beta=0.01$; Fig.
-\ref{ws10000} shows the results for a WS model with $N=10^4,\ \left<k\right>=6$ and $\beta=0.01$.
-In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the maximum length of the actual time-line is $S=20$.
-In this figure we can see that the minority opinion A reaches higher final values of $P_0(500)$ compared to the case depicted in Fig.
-\ref{ws_20-80_op}, with a value of $P_\text{A}(500) = 0.027 \pm 0.023$ in Fig.
-\ref{ws1000} and a value of $P_\text{A}(500) = 0.030 \pm 0.018$ in Fig.
-\ref{ws10000}.
+This imbalance leads to the decrease of the minority opinion A.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
 For each model the REC and the PR filtering algorithms are compared.
+The dotted black line represents the initial prevalence of opinion A.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 \end{figure}
+\noindent
+Fig.
+\ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
+This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
+The high clustering in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
+\ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Fig.
+\ref{ws_20-80_op}).\\
+\newline
+This result differs from results in \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
+The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$.
+The results that where obtained in \cite{Perra2019} are for networks with size $N = 10^5$ nodes and average degree $\left<k\right> \sim 6$.
+In this thesis results are obtained for $N=10^3$ and average degree $\left<k\right> \sim 10$.
+In both cases the maximum length of the actual time-line is $S=20$.
+Fig.
+\ref{ev_op_WS_k=6} shows the results for the evolution of the prevalence of opinion A for an initial $20/80$ opinion distribution and using the PR filtering algorithm; Fig.
+\ref{ws1000} displays results for a WS model with $N=10^3$ and $\left<k\right> = 6$; Fig.
+\ref{ws10000} shows results for a WS model with $N=10^4$ and $\left<k\right> = 6$.
+We see that the minority opinion A reaches higher final values of $P_0(500)$ compared to the case depicted in Fig.
+\ref{ws_20-80_op}, with a value of $P_\text{A}(500) = 0.027 \pm 0.023$ in Fig.
+\ref{ws1000} and a value of $P_\text{A}(500) = 0.030 \pm 0.018$ in Fig.
+\ref{ws10000}.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$.
-The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditionsn for a WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$.
+The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
 The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$.
-In both cases the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$ and the PR filtering filtering algorithm is used.
+In both networks the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$.
+The PR filtering filtering algorithm is used.
+The dotted black represents the initial prevalence of opinion A.
+The dotted grey line represents a vanished prevalence of opinion A.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 Fig.
-\ref{echo_20_80} shows the distribution of the fraction of friends of actor $i$ with the same opinion B at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS).
-For each network, the REC and PR filtering algorithm are compared.
-The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the ``Echo chambers'' paragraph.\\
-\newline
-The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same.
-Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz block model agree with each other as well.
-This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
-\newline
-There are no visible effects of having a clear community structure.
-Fig.
-\ref{sbm} shows the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure).
-There is (almost) no difference in the results for a high and a low modularity.
-\newpage
+\ref{echo_20_80} shows the distribution of the fraction of friends of actor $i$ with the same opinion B at $t=500$.
+The REC and PR filtering algorithms are compared for the ER, WS, SBM and SBM-WS.\\
+\newline
+The results for the ER model and the SBM (both the high and low modularity configuration) look qualitatively the same.
+Similarly, the results for the WS model and the SBM-WS agree with each other as well.
+This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is the clustering that drives these differences.
+
 \begin{figure}[H]
@@
   \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
+  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for unbalanced starting conditions.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$ for unbalanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+For each model the REC and the PR filtering algorithms are compared.
+The dotted black line represents a constant amount of echo chambers over time.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-Four different network models are used, each with $N = 10^3$ and average degree $\left<k\right> \sim 10$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model the REC and the PR filtering algorithms are compared.
-The starting conditions are $P_\text{A}(0) = 0.2$ and $P_\text{B}(0) = 0.8$.
 Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for an improved visualization.}
+Error bars are omitted for improved visualization.}
 \label{echo_20_80}
@@
 \noindent
-Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (Fig.
+There are no visible effects of having a community structure.
+Fig.
+\ref{sbm} shows the results for a SBM with high modularity (community structure) and low modularity (negligible community structure).
+There is (almost) no difference in the results for high and low modularity.\\
+\newline
+Lets first discuss the results for the ER model and the SBM (Fig.
 \ref{er_echo_20-80} and Fig.
@@
 \newline
-Networks with a high clustering coefficient (WS and SBM-WS) behave completely different.
-For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in Fig.
+Networks with high clustering (WS and SBM-WS) behave completely different.
+For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in Fig.
 \ref{ws} and Fig.
 \ref{sbm-ws}).
-An important note must be made here: the absolute value of neighborhoods with all the actors carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Fig.
+The absolute value of neighborhoods with all the actors carrying the minority opinion A at $t=0$ is equal to zero in the case of a WS network in Fig.
 \ref{ws} and in the case of a SBM-WS network in Fig.
@@
 \newline
-This enhancing of the formation of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm.
 Remember from Fig.
@@
 
-% talk about small spike around 1 for PR method??
-% explain black dotted line??
-
 \subsection{Formation of echo chambers versus community structure}\label{echoVSmod}
@@
 \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
-The $y$-axis represents the average of the normalized fraction of actors with all neighbors carrying opinion A and the normalized fraction of actors with all neighbors carrying opinion B (with normalized we mean that the computed value for the fraction of actors with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
-This is a measure for the formation of echo chambers (opinion clusters) in the system.
 The data points represent the value of the echo chamber size in networks with a different modularity.
-The initial opinion distribution was $50/50$ and was randomly distributed over the network.
-\\
-\newline
-The data points are shown together with the standard error $SE = \sigma /\sqrt{N}$.
-The standard error $SE$ is chosen here instead of the standard deviation $\sigma$.
-This is because the standard deviation is quite large.
-Each data point is the average of $10^2$ independent simulations.
-The dispersion, which is measured by the standard deviation, in the set of the $10^2$ obtained values is quite large.
-However, if we run the $10^2$ simulations over and over again, each time calculating the average, the obtained value of the average is always more or less the same.
-This suggests that the value of the calculated average lies close to the real average.
-Since we are mostly interested in the correctness of the value of the average, which is quantified by the standard error, we chose to depict the results together with this standard error.\\
-\newline
-Table \ref{tab1} displays the network models that are used in Fig. \ref{echo_vs_mod} (all with $N = 10^3$ and average degree $\left<k\right> \sim 10$).
+The initial opinion distribution is $50/50$.
+\\
+\newline
+Table \ref{tab1} displays the network models used in Fig. \ref{echo_vs_mod} (all with $N = 10^3$ and $\left<k\right> \sim 10$).
 
@@
 \end{table}
-\newpage
+\noindent
+For most values of the modularity, the REC filtering algorithm produces almost no echo chambers.
+If the modularity becomes high enough we see a small increase in the formation of echo chambers, however even for $Q = 0.908 \pm 0.005$, the size of the echo chambers remains smaller than two.
+This means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.
+For the PR filtering algorithm, the increase in the size of the echo chambers with increasing modularity is more significant.
+From a modularity equal to $Q = 0.476 \pm 0.009$, we see a sharp increase in the formation of echo chambers.
+For $Q < 0.476$, the formation of echo chambers remains small, even when using the PR filtering algorithm.
+These results suggest that community structure enhances the formation of echo chambers, mostly when using the PR filtering algorithm.
+A high modularity is however required to see the effect of community structure on the formation of opinion clusters.
+
 \begin{figure}[H]
@@
 	\captionsetup{format=plain}
-	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-The initial opinion distribution is $50/50$.
-The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and average degree around 10.
-The other four data points are the results for SBM's with 10 communities of size 100 ($N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and average degree $\left<k\right> \sim 10$.
+	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{The relation between the formation of echo chambers and the community structure, measured by the modularity $Q$, for both the PR and REC filtering algorithms.
+The data points with $Q = 0$ are the result for an ER network; the other eight data points are results for SBM's, each with different modularity (modularity is changed by changing the edge probabilities).
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B  (normalized means that the value for the fraction of actors with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
 Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard error $SE$.}
+Bars represent the standard error $SE$.
+The data points are shown together with the standard error $SE = \sigma /\sqrt{N}$, because we are mostly interested in the correctness of the value of the average, which is quantified by the standard error.}
 \label{echo_vs_mod}
 \end{figure}
-\noindent
-The PR and REC filtering algorithms are compared.
-For most values of the modularity, the REC filtering algorithm produces almost no echo chambers.
-If the modularity becomes high enough we see a small increase in the formation of echo chambers, however even for $Q = 0.908 \pm 0.005$, the size of the echo chambers remains smaller than two.
-This means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\
-\newline
-For the PR filtering algorithm, the increase in the size of the echo chambers with increasing modularity is more significant.
-From a modularity equal to $Q = 0.476 \pm 0.009$, we see a sharp increase in the formation of echo chambers.
-For $Q < 0.476$, the formation of echo chambers remains small, even when using the PR filtering algorithm.\\
-\newline
-These results suggest that the community structure enhances the formation of echo chambers, mostly when using the PR filtering algorithm.
-A strong community structure, high modularity, is however required to see the effect of community structure on the formation of opinion clusters.
+
 \newpage

### 2021-05-21 draft.tex
Update draft.tex
changed clustering and/or modularity to clustering or modularity

@@
 The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
-For each real-world network, the results will be compared to results obtained from network models with similar average degree, clustering and/or modularity.
+For each real-world network, the results will be compared to results obtained from network models with similar average degree, clustering or modularity.
 We will mainly focus on the formation of echo chambers.
@@
 
-Results for the PGP trust network are compared to results for network models with similar average degree, clustering and/or modularity.
+Results for the PGP trust network are compared to results for network models with similar average degree, clustering or modularity.
 The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\

### 2021-05-21 draft.tex
Update draft.tex
changed beginning of results stubborn actors

@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditionsn for a WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
+A WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$ is implemented.
 The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
@@
 \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
 Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
@@
 
+Stubbornness is an important feature of real-life opinion dynamics.
+Understanding the impact of individual resistance and the interplay with filtering algorithms and network topologies is crucial.
 In this section, we will study the effect of stubborn actors on the opinion dynamics.
-We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}.
-The results of the first model, including a stubbornness parameter, are given in Section \ref{stubbpar}.
+We focus on two different stubbornness models, discussed in Section \ref{stubbBin}.
+The results of the first model, including a resistance parameter, are given in Section \ref{stubbpar}.
 The results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
 \newline
-We will implement four different network models in order to investigate the possible interplay between network structure and stubbornness.
-Each of the network models have a total number of nodes $N=10^3$ and average degree $\left<k\right> \sim 10$.
-The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
-The interplay between stubbornness and clustering is studied with the Watts-Strogatz model (WS) with mean degree $\left<k\right>=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$.
-The effect of community structure is investigated with the stochastic block model (SBM) with 10 communities of size 100.
-Two different modularities are implemented: a high modularity, $Q = 0.908 \pm 0.005$, with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
-Finally, the interplay between stubbornness and both clustering and community structure is studied with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100.
-Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> = 10$ and rewire probability $\beta = 0.01$.
-Edges are added between communities with a probability $p_{\text{add}} = 0.001$.
-The clustering coefficient of this network is $\left<cc\right> = 0.553 \pm 0.004$ and the modularity is $Q = 0.909 \pm 0.004$.\\
-\newline
-For each network model, we compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm.
-
-\subsection{Including a stubbornness parameter}
+We will compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm.
+The ER, WS, SBM and SBM-WS are implemented to study the interplay between network structure and stubbornness.
+
+\subsection{Resistance parameter}
 \label{stubbpar}
 
-In this section, we study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.\\
-\newline
+We study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.
 In both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers.
-This is done for two different starting conditions: balanced starting conditions and unbalanced starting conditions.
-\\
-\newline
-The opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
+This is done for balanced and unbalanced starting conditions.
+\\
 \newline
@@
 \newline
-We make a fraction of the actors (fracRes) completely stubborn ($r_i = 1$).
-We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$.
-In all cases, the stubborn actors are randomly distributed over the network.\\
-\newline
-We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared.
-Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for an initial $20/80$ opinion distribution.
+We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$. \\
+\newline
+We will first discuss the formation of echo chambers for balanced starting conditions, after which the results for the opinion evolution will be shared.
+Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for unbalanced starting conditions.
 
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and balanced starting conditions.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for balanced starting conditions and using the PR filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.

### 2021-05-22 draft.tex
Update draft.tex
changed results: stubborn actors

@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for balanced starting conditions.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) over time for balanced starting conditions.
+  \caption[Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) for balanced starting conditions.]{Evolution of the prevalence of opinion A ($P_{\text{A}}(t)$) for balanced starting conditions.
 \textbf{(a)} SBM; \textbf{(b)} SBM-WS.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
 \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
 A WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$ is implemented.
-The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions.
+The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
 The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$.
@@
 
-We study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.
-In both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers.
+We study the effect of a resistance parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.
+For both configurations, we study the evolution of the prevalence of the opinions and the formation of echo chambers.
 This is done for balanced and unbalanced starting conditions.
-\\
-\newline
-\textbf{Completely stubborn actors}\\
-\newline
-We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$. \\
-\newline
-We will first discuss the formation of echo chambers for balanced starting conditions, after which the results for the opinion evolution will be shared.
-Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for unbalanced starting conditions.
+
+\subsubsection{Completely stubborn actors}\label{complStubbAct}
+
+We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, $f_s=0$, to all the actors are stubborn, $f_s=1$. \\
+\newline
+First, we discuss the formation of echo chambers for balanced starting conditions, after which the results for the opinion evolution are shared.
+Then, the formation of echo chambers and the evolution of the prevalence of the opinions for unbalanced starting conditions are discussed.\\
+\newline
+\textbf{Balanced starting conditions}\\
+\newline
+We start with a configuration in which $50 \%$ of the actors is given opinion A and $50 \%$ is given opinion B.
+Fig.
+\ref{echo_vs_frac_complStubb} shows the formation of echo chambers versus the fraction of stubborn actors, $f_s$.
+The PR filtering algorithm is used.
 
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and balanced starting conditions.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for balanced starting conditions and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus fraction of stubborn actors for the PR filtering algorithm and balanced starting conditions.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
 Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
@@
 Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
+The PR filtering algorithm is used.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
@@
 \noindent
-Fig.
-\ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the PR filtering algorithm for balanced starting conditions.
-The $y$-axis represents the formation of echo chambers in the same way as was explained for Fig.
-\ref{echo_vs_mod} in Subsection \ref{echoVSmod}.
-The dotted black line represents an echo chamber size equal to one.
-This means that the number of echo chambers remains constant over time.\\
-\newline
-The Erd\H{o}s-R\'{e}nyi model (ER) shows no appearance of echo chambers.
-Introducing a fraction of completely stubborn actors does not change this behavior.
-The stochastic block model with a low modularity behaves similar as the Erd\H{o}s-R\'{e}nyi model (no formation of echo chambers).
-The data points for the SBM with a low modularity are not visible in Fig.
-\ref{echo_vs_frac_complStubb}, since they lie behind the ER data points.\\
-\newline
-The other three networks (WS, SBM-WS and SBM with a high modularity) all show qualitatively the same behavior.
-If an increasing fraction of completely stubborn actors is introduced, the formation of echo chambers decreases with this increasing fraction of stubborn actors.\\
-\newline
-The WS and SBM-WS networks are the high clustering networks.
+The ER model shows no appearance of echo chambers.
+Introducing a fraction of stubborn actors does not change this behavior.
+The SBM with low modularity behaves similar as the ER model.
+The data points for the SBM with low modularity lie behind the ER data points and are not visible in Fig.
+\ref{echo_vs_frac_complStubb}.
+The other networks (WS, SBM-WS and SBM with high modularity) all show qualitatively the same behavior.
+If an increasing fraction of stubborn actors is introduced, the formation of echo chambers decreases.\\
+\newline
+The WS and SBM-WS networks have high clustering.
 For a small fraction of stubborn actors they show the largest echo chambers.
-The stochastic block model with a high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for fracRes $\leqslant 0.1$.
-This leads us to make the same conclusion as before: a community structure enhances the formation of echo chambers compared to a random network, however not as significant as networks with a high clustering coefficient.
+The SBM with high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for $f_s \leqslant 0.1$.
+These results confirm that community structure enhances the formation of echo chambers compared to random networks, however not as significant as networks with high clustering.
+
 \begin{figure}[H]
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of stubborn actors for the REF and REC filtering algorithms and balanced starting conditions.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc \right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \noindent
-Fig.
-\ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn actors, fracRes, when using the REF and REC filtering algorithms.
-The results are obtained for an initial $50/50$ opinion distribution.
-The results for the REF and REC filtering algorithms behave qualitatively the same, but are different from those obtained with the PR filtering algorithm (Fig.
+Results for the REF and REC filtering algorithms (Fig.
+\ref{echo_vs_frac_complStubb_REF-REC}) behave qualitatively the same, but are different from those obtained with the PR filtering algorithm (Fig.
 \ref{echo_vs_frac_complStubb}).
-Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors, we first observe an increase in the formation of echo chambers, followed by a decrease.
-This is especially clear for the high clustering networks (WS and SBM-WS).
-The SBM with a high modularity displays a small bump.
-The ER and the SBM with a low modularity (data points for this model lie behind the ER data points) do not seem to have a significant bump.\\
-\newline
-These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small) fraction of stubborn actors enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure).
-If fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease.
-This is probably due to the fact that too many actors become stubborn, so there are more and more actors that will not change opinions.
-This in turn increases the chance of having stubborn neighbors, each with different opinions, who cannot change opinions, which disables the formation of echo chambers.\\
-\newline
-This discrepancy between the results for the PR filtering algorithm and those for the REF and REC filtering algorithms in combination with a small fraction of stubborn actors can be potentially explained with the help of the following simple example.
+Instead of a decrease in the formation of echo chambers with increasing fraction of stubborn actors, we first observe an increase, followed by a decrease.
+This is especially visible for the high clustering networks (WS and SBM-WS).
+The SBM with high modularity displays a small bump.
+The ER and the SBM with low modularity (data points for this model lie behind the ER data points) do not seem to have a significant dependency on the fraction of stubborn actors.\\
+\newline
+These results suggest that the interplay between the REF or REC filtering algorithms and the presence of a (small) fraction of stubborn actors enhances the formation of echo chambers (mainly for networks with high clustering or community structure).
+If $f_s \geqslant 0.25$, the formation of echo chambers starts to decrease.
+This is due to the fact that too many actors become stubborn.
+This increases the chance of having stubborn neighbors, each with different opinions, who cannot change opinions.
+Hence, the formation of echo chambers is disabled.\\
+\newline
+The discrepancy between the results for the PR filtering algorithm and those for the REF and REC filtering algorithms in combination with a small fraction of stubborn actors may be explained with the help of the following example.
 Assume we have three actors that form a triangle.
@@
 Lets first discuss the PR filtering algorithm and assume that none of the actors are stubborn.
-In case of the PR filtering algorithm this triangle has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion).
-This can be seen as follows: the one actor with opinion B is likely to switch to opinion A, since all its friends carry this opinion and hence, this opinion will be strongly pushed on its time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter here, since there are no neighbors with opinion B).
-The two actors with opinion A will have more chance of keeping that opinion than switching to opinion B in case of the PR filtering algorithm.
-This is because the PR filtering algorithm will promote the posts of the friend with the same opinion A over the posts of the friend with opinion B.
-The end result will be that there is a high chance that all the three actors will adopt opinion A and hence, form an echo chamber.
+This triangle then has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion), which can be seen as follows: the actor with opinion B is likely to switch to opinion A, since all its friends carry this opinion and hence, this opinion will be recommended on its time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter, since there are no neighbors with opinion B).
+The two actors with opinion A will have more chance of keeping that opinion than to switch to opinion B.
+This is because the PR filtering algorithm will promote posts of the friend with the same opinion A over posts of the friend with opinion B.
+There is thus a high chance that all three actors will adopt opinion A and form an echo chamber.
 Now assume that we make the actor with opinion B stubborn.
 This actor is no longer capable of switching to opinion A.
-However, the two actors with opinion A still have more chance of keeping this opinion in case of the PR filtering algorithm, because this filtering algorithm keeps on promoting the posts of the neighbor with the same opinion A, and hence reduces the visibility of the posts of the stubborn friend with opinion B.
-The system will not form an echo chamber, since the stubborn actor with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the actors with opinion A to opinion B.
+The two actors with opinion A still have more chance of keeping this opinion: the PR filtering algorithm keeps on promoting posts of the neighbor with the same opinion A, and hence reduces the visibility of posts of the stubborn friend with opinion B.
+The system will be less likely to form an echo chamber: the stubborn actor with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the actors with opinion A to opinion B.
 The conclusion is that, in case of the PR filtering algorithm, making a fraction of the actors stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
-What happens if we use the REF or REC filtering algorithms?
+Lets now study what happens if we use the REF or REC filtering algorithms.
 Again we have one actor with opinion B and two actors with opinion A in the triangle.
 Assume first that none of the actors are stubborn.
-For the PR filtering algorithm, we argued that in this case, the triangle has a high chance of evolving towards an echo chamber.
-For the REF and REC filtering algorithms this chance is however smaller, which can be seen as follows.
-The actor with opinion B still has a high chance of switching to opinion A, because all its friends (which are the two other actors in the triangle) carry this opinion.
-However, the two actors with opinion A will now not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
+For the PR filtering algorithm, we argued that the triangle has a high chance of evolving towards an echo chamber.
+For the REF and REC filtering algorithms this chance is smaller.
+This can be understood as follows.
+The actor with opinion B has high chance of switching to opinion A, because all its friends carry this opinion.
+However, the two actors with opinion A will not have a higher chance of keeping this opinion than to switch to opinion B, as was the case for the PR filtering algorithm.
 Instead, they have an equal chance of keeping their opinion as to change to the other opinion.
-This is because both the REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the actors with opinion A.
+This is because the REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the actors with opinion A.
 Both opinion A and opinion B have an equal chance of being promoted on the time-line of the actors with opinion A.
-This reduced tendency of the actors with opinion A to not change their opinion (compared to using the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
-This would explain why, in case of no stubborn actors, the REF and REC filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
+This reduced tendency of the actors with opinion A to not change their opinion (compared to the PR filtering algorithm) hampers the ability of the triangle to form an echo chamber.
+This explains why, in case of no stubborn actors, the REF and REC filtering algorithms show a lower formation of echo chambers compared to the PR filtering algorithm.
 Now, lets again assume that we make the actor with opinion B stubborn.
-In case of the PR filtering algorithm this reduced the formation of echo chambers in the triangle.
+In case of the PR filtering algorithm, this reduced the formation of echo chambers in the triangle.
 For the REF and REC filtering algorithms, things are different.
 The stubborn actor with opinion B will not change opinion.
-The two actors with opinion A, on the other hand, will now have a higher chance of switching to opinion B (which was not the case when using the PR filtering algorithm).
+The two actors with opinion A, on the other hand, will have a higher chance of switching to opinion B (which was not the case for the PR filtering algorithm).
 This can be seen as follows.
-The REF and REC filtering algorithms do not discriminate between the two different opinions of the two friends of the actors with opinion A.
-If the actor with opinion B is, however, stubborn it will keep on pushing that opinion on the time-lines of the two actors with opinion A.
-This will eventually lead to an increase of posts with opinion B, compared to posts with opinion A, on the time-lines of the two actors with opinion A.
-The result will then be that the actors with opinion A will have a higher chance of switching to opinion B, which enhances the formation of echo chambers in the triangle.
-This would explain the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
-If too many actors become stubborn, this increased ability of the system to form echo chambers, in case of using the REF or REC filtering algorithms, starts to decrease which can also be seen with the help of the triangle example.
-If the fraction of stubborn actors becomes too high, we might end up with the case where both the actor with opinion B and one of the actors with opinion A are stubborn.
-This strongly reduces the ability to form echo chambers in the triangle since we can no longer have a final situation in which all the three actors in the triangle carry the same opinion.\\
+The REF and REC filtering algorithms do not discriminate between the two different opinions of the friends of the actors with opinion A.
+If the actor with opinion B is, however, stubborn it will keep on pushing that opinion on the time-lines of the actors with opinion A.
+This will lead to an increase of posts with opinion B, compared to posts with opinion A.
+The actors with opinion A will have a higher chance of switching to opinion B, which enhances the formation of echo chambers in the triangle.
+This explains the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
+If too many actors become stubborn, this increased ability of the system to form echo chambers starts to decrease which can also be seen with the help of the triangle example.
+If the fraction of stubborn actors becomes too high, we might end up with both the actor with opinion B and one of the actors with opinion A that are stubborn.
+This reduces the ability to form echo chambers since we can no longer have a final situation in which all the three actors in the triangle carry the same opinion.\\
 % make figure 
 \newline
-So far the results for the formation of echo chambers in the system for balanced starting conditions and introducing a fraction of completely stubborn actors.
-But what about the evolution of the prevalence of the opinions over time?
-Are these affected by introducing stubborn actors in the system?
-The short answer is no.
-Introducing stubborn actors in the system does not affect the evolution of the prevalence of the opinions over time for neither the REC, nor the PR filtering algorithm.
+It is found that stubborn actors influence the formation of echo chambers.
+Combined with the PR filtering algorithm, they lead to a decrease in the formation of echo chambers.
+Combined with the REF or REC filtering algorithms they induce an increase, followed by a decrease if the fraction of stubborns becomes too high.
+We now want to study whether the stubborns also influence the evolution of the prevalence of the opinions.
+The answer is no.
+Introducing stubborn actors in the system does not affect the prevalence of the opinions over time for neither the REC, nor the PR filtering algorithm.
 The balanced opinion distribution is maintained over time.
-This result is valid for all the four implemented network models (ER, WS, SBM and SBM-WS).
 Network topologies such as high clustering or community structure do not change this result.
-\\% Include figure??
-\newline
-Lets now study the situation for unbalanced starting conditions, combined with a fraction of stubborn actors.
-The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and the initial prevalence of opinion B is $P_\text{B}(0) = 0.8$.
+\\
+\newline
+\textbf{Unbalanced starting conditions}\\
+\newline
+Lets now study the situation for unbalanced starting conditions. The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and of opinion B is $P_\text{B}(0) = 0.8$.
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus fraction of stubborn actors for the PR and REC filtering algorithms and unbalanced starting conditions.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for unbalanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard error $SE$.}
+Bars represent the standard error $SE$; these are small and not visible.}
 \label{echo_vs_frac_complStubb_PR-REC_20-80}
@@
 \noindent
-Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion).
-The PR and the REC filtering algorithms are compared.
-The standard errors $SE$ are small and not visible in the figure.
-The dotted black line represent the situation in which the number of echo chambers remains constant over time.
-The interpretation of the $y$-axis is slightly different than before.
-In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both.
-In case of a $20/80$ distribution we have a skewed echo chamber distribution, as can be seen in Fig.
+In Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers versus the fraction of stubborn actors, $f_s$ is shown for the PR and the REC filtering algorithms.
+The interpretation of the $y$-axis is different than for configurations with balanced starting conditions.
+For balanced starting conditions, the echo chamber size was interpreted as the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
+We first normalized the echo chambers around opinion A and those around opinion B and then took the average of both.
+For unbalanced starting conditions, we have skewed echo chamber distributions, as can be seen in Fig.
 \ref{echo_20_80}, with normalized values going to infinity.
-We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities.
-Hence, we opted for a different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B), then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before).
+We still want to have an average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers, but we want to avoid problems with those infinities.
+Hence, we opted for a different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (the sum of the number of echo chambers around opinion A and those around opinion B), then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done for balanced starting conditions).
 By taking the average, we neglect the skewed echo chamber distribution.
-This is, however, not a problem.
-We want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers.
-\\
+This is not a problem.
+We want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of the echo chambers.
+It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as an average in case of unbalanced starting conditions.
+When comparing the different network models, the skewed distributions should not be overlooked.
+The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other.\\
 \newline
 When using the PR filtering algorithm (Fig.
-\ref{PR_frac_compl_stubb_20-80}) we observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors.
-This behavior is similar to the case of balanced starting conditions in combination with the PR filtering algorithm (Fig.
+\ref{PR_frac_compl_stubb_20-80}) we observe a decrease in the formation of echo chambers with increasing fraction of stubborn actors.
+This behavior is similar to the situation with balanced starting conditions (Fig.
 \ref{echo_vs_frac_complStubb}).
-For unbalanced starting conditions, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for fracRes $=0$), as can be observed in Fig.
-\ref{echo_20_80}.
-This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was explained for Fig.
-\ref{echo_20_80}.
-\\
-\newline
-The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher.
-It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of unbalanced starting conditions.
-The reason is that this average neglects the skewed echo chamber distributions as depicted in Fig.
-\ref{echo_20_80}.
-When comparing the different network models, these skewed distributions should not be overlooked.
-The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other.
-\\
-\newline
-% explain reason behavior REC!!
+The formation of echo chambers in the ER network and the SBM with low modularity is non-zero (for $f_s=0$).
+This agrees with results found in Fig.
+\ref{echo_20_80} and is a consequence of the fact that unbalanced starting conditions, combined with the PR filtering algorithm, lead to an increase in the prevalence of the majority opinion.
+\\
+\newline
 The results for the REC filtering algorithm (Fig.
 \ref{REC_frac_compl_stubb_20-80}) are different than those for the PR filtering algorithm.
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease.
-This result is similar to the results for balanced starting conditions in combination with the REC filtering algorithm (Fig.
+The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease.
+This result is similar to results for balanced starting conditions (Fig.
 \ref{REC_frac_compl_stubb}).
-The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we see a decrease in the formation of echo chambers.
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show no significant formation of echo chambers.
-For an increasing fraction of the completely stubborn actors, the echo chambers eventually converge to one (for all the network models), indicating a constant amount of echo chambers over time (represented by the black dotted line).\\
-\newline
-We can conclude that the formation of echo chambers versus the fraction of completely stubborn actors, for unbalanced starting conditions, qualitatively resembles the results for balanced starting conditions (decrease in formation of echo chambers with an increasing fraction of stubborn actors for the PR filtering algorithm versus an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm).
+The SBM with high modularity does not show an increase in the formation of echo chambers, but remains flat for $f_s \leqslant 0.2$; if the fraction of stubborn actors increases further, the formation of echo chambers decreases and converges to one, indicating a constant amount of echo chambers over time.
+The ER network and the SBM with low modularity show no significant formation of echo chambers.\\
+\newline
+We can conclude that the formation of echo chambers versus the fraction of stubborn actors, for unbalanced starting conditions, qualitatively resembles the results for balanced starting conditions: decrease in formation of echo chambers with increasing fraction of stubborn actors for the PR filtering algorithm versus increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm.
+There is, however, a quantitative difference: the formation of echo chambers is lower when using unbalanced starting conditions compared to using balanced starting conditions.
+The reason for this is probably the higher initial number of echo chambers when using unbalanced starting conditions.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+The PR filtering algorithm is used and three fractions of stubborn actors ($f_s = 0$, $f_s = 0.4$ and $f_s = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents the initial prevalence of opinion A.
+The PR filtering algorithm is used and three fractions of stubborn actors ($f_s = 0$, $f_s = 0.4$ and $f_s = 0.8$) are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-Bars represent the standard deviation $\sigma$.}%more explanation?
+Bars represent the standard deviation $\sigma$.}
 \label{ev_op_20_80_stubb_PR}
@@
 \noindent
-Lets now discuss the evolution of the prevalence of the opinions.
+For balanced starting conditions we found that introducing stubborn actors did not affect the prevalence of the opinions over time.
+Lets now study whether this is also the case for unbalanced starting conditions.
 In Fig.
-\ref{ev_op_20_80}, we saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of unbalanced starting conditions.
-The PR filtering algorithm, on the other hand, resulted in a decrease of the prevalence of the minority opinion.
-Are these results affected by introducing stubborn actors?
-As for the balanced starting conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
+\ref{ev_op_20_80}, we saw that a configuration with no stubborn actors maintained the prevalence of the initial opinions if the REC filtering algorithm is used.
+The PR filtering algorithm resulted in a decrease in the prevalence of the minority opinion.
+Introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
 The status quo is maintained over time.
-Introducing stubborn actors combined with the PR filtering algorithm, on the other hand, does have an impact in case of unbalanced starting conditions.\\
+Introducing stubborn actors combined with the PR filtering algorithm does have an impact.\\
 \newline
 Fig.
-\ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time in case of unbalanced starting conditions and using the PR filtering algorithm.
-We compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS).
-For each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$).
-\\
-\newline
-If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of unbalanced starting conditions.
-Introducing a fraction of stubborn actors changes this scenario.
-For an increasing fraction of stubborn actors, the decrease of the minority opinion becomes less and less strong.
-This result is valid for all the different network topologies.
-If there are no stubborn actors, the PR filtering algorithm forces actors with the minority opinion to switch to the majority opinion.
-If we introduce a fraction of stubborn actors in the system, we end up with stubborn actors that carry the minority opinion.
-These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them.
-The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.
-\\
-\newline
-\textbf{All actors partly stubborn}\\
-\newline
-In the second configuration, all the actors are partly stubborn.
-They still have some chance of changing their opinion, they are however less eager to do so.
-All the actors are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the actors are completely willing to change opinions) to $r=1$ (all the actors are completely stubborn).\\
-\newline
-We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate the results for unbalanced starting conditions.\\
-\newline
+\ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time when using the PR filtering algorithm.
+We compare four network models (ER, WS, SBM (high and low modularity) and SBM-WS).
+For each model we compare the results for three different fractions of stubborn actors ($f_s=0$, $f_s=0.4$ and $f_s=0.8$).
+\\
+\newline
+If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion.
+For an increasing fraction of stubborn actors, this decrease becomes less and less strong.
+This result is valid for the different network topologies.
+For no stubborn actors, the PR filtering algorithm forces actors with the minority opinion to switch to the majority opinion.
+If we introduce a fraction of stubborn actors, we will have stubborn actors that carry the minority opinion.
+These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them and hamper the decrease of the minority opinion.
+
+\subsubsection{All actors partly stubborn}
+
+In the second configuration, we assume that all actors are given the same resistance parameter $r$ ($0 \leqslant r \leqslant 1$), i.e.
+they have a chance to change their opinion.
+We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate results for unbalanced starting conditions.\\
+\newline
+\textbf{Balanced starting conditions}\\
+\newline
+We start with $50 \%$ of the actors with opinion A and $50 \%$ with opinion B.
 Fig.
-\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The interpretation of the $y$-axis is again as explained in Subsection \ref{echoVSmod}.
-The results are similar to the case where a fraction of the actors is completely stubborn combined with the PR filtering algorithm (Fig.
+\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the resistance parameter of the actors, $r$ when using the PR filtering algorithm.
+Results resemble those of the configuration in which a fraction of the actors is completely stubborn (Section \ref{complStubbAct}, Fig.
 \ref{echo_vs_frac_complStubb}).\\
 \newline
-The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers.
+The ER model has no appearance of echo chambers.
 This is not changed by making the actors more and more stubborn.
-The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\
-\newline
-The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$.
-If the actors become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases.
+The same result is obtained for the SBM with low modularity (the data points lie behind those of the ER network).
+The SBM with high modularity has a (small) formation of echo chambers for $r=0$.
+If $r$ increases, the formation of echo chambers decreases.
 This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
@@
 If $r=1$, the actors retain their initial opinion and the number of echo chambers before and after the time evolution remains the same.
-This situation is represented by the black dotted line.
 
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
+	\caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
+The PR filtering algorithm is used.]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represent a constant amount of echo chambers over time.
+The PR filtering algorithm is used.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \end{figure}
-\noindent
-In Fig.
-\ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the actors, $r$, is depicted for an initial $50/50$ opinion distribution.
-Fig.
-\ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Fig.
-\ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.
-\newpage
+
 \begin{figure}[H]
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $50/50$ opinion distribution.
+  \caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
+The REF and REC filtering algorithms are used]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represent a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \noindent
-We can make similar observations as we did in the ``Completely stubborn actors'' paragraph.
-The results for the REF and REC filtering algorithms are qualitatively similar to each other and differ from the results obtained with the PR filtering algorithm (Fig.
+Results for the REF and REC filtering algorithms (Fig.
+\ref{echo_vs_all_frac_stubb_REF-REC}) are qualitatively similar to each other and differ from results obtained with the PR filtering algorithm (Fig.
 \ref{echo_vs_all_frac_stubb_PR}).
-The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the actors $r$ increases.
-For the REF filtering algorithm, this increases reaches a maximum around $r = 0.7$ and then starts to slowly decrease between $r = 0.7$ and $r = 0.9$.
+The formation of echo chambers in networks with high clustering (WS and SBM-WS) increases if the resistance of the actors increases.
+For the REF filtering algorithm, this increase reaches a maximum around $r = 0.7$ and slowly decreases between $r = 0.7$ and $r = 0.9$.
 Between $r = 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers.
-The REC filtering algorithm also seems to have a maximum around $r = 0.7$ for the WS model.
-The SBM-WS model, on the other hand, displays a continuous increase until $r = 0.9$.
-In both case we have a sharp drop from $r = 0.9$ to $r=1$.
-If $r=1$, the actors are completely stubborn and thus keep their initial opinion.
-Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
-This is represented by the black dotted line.
-It is remarkable that for high levels of the stubbornness of the actors $r$, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the actors are not stubborn.
-We can conclude that making the actors partly stubborn enhances the formation of echo chambers when using the REF and REC filtering algorithms in case of networks with a high clustering coefficient.
-This is the opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the actors).
-The reason for this discrepancy is probably of a similar nature as for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Completely stubborn actors'' paragraph.
-If we zoom in on the stochastic block model with a high modularity, we see a slight increase in the formation of echo chambers.
-The increase is however not as strong as for the high clustering networks.\\
-\newline
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points are hidden behind those of the ER network) do not have any dependency on the stubbornness level of the actors.
-This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF or REC filtering algorithms that drives this increase in the formation of echo chambers with increasing $r$.
-% talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
-\\
-\newline
-What about the results for the evolution of the prevalence of the opinions in case of balanced starting conditions and making all the actors partly stubborn?
-Making all the actors partly stubborn does not disturb the evolution of the opinions.
-The balanced starting conditions remain balanced over time, as was the case for no stubbornness in the system.
-This result is valid for all the network models (ER, WS, SBM and SBM-WS) and for both the REC and PR filtering algorithm.\\
-\newline
-Lets now study the results for unbalanced starting conditions ($20/80$).
-Opinion A is the minority opinion, which makes opinion B the majority opinion.
+The REC filtering algorithm seems to have a maximum around $r = 0.7$ for the WS model.
+The SBM-WS model displays a continuous increase until $r = 0.9$.
+We observe a sharp drop from $r = 0.9$ to $r=1$.
+If $r=1$, the actors are completely stubborn and hence, retain their initial opinion: the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
+For high levels of the resistance of the actors, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the actors are not stubborn.
+If we zoom in on the SBM with high modularity, we see an increase in the formation of echo chambers.
+The increase is however not as strong as for the networks with high clustering.\\
+\newline
+The ER network and the SBM with low modularity (data points are hidden behind those of the ER network) do not show a dependency on the resistance level of the actors.
+This suggests that it is the interplay between high clustering and, to a lesser extent, community structure (high modularity), and the REF or REC filtering algorithms that drives the increase in the formation of echo chambers with increasing $r$.
+This is the opposite of the situation where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with increasing $r$).
+The reason for this discrepancy is probably of a similar nature as for the discrepancy in case of introducing a fraction of completely stubborn actors, Section \ref{complStubbAct}.\\
+\newline
+We conclude that giving all actors the same resistance parameter behaves qualitatively similar to the scenario in which a fraction of completely stubborn actors is introduced (Section \ref{complStubbAct}): decrease in the formation of echo chambers when using PR filtering; increase in polarization when using REF or REC filtering, followed by a decrease if the resistance becomes high enough.
+The question then rises whether this similarity between the two configurations is also found for the evolution of the prevalence of the opinions.
+The conclusion is yes: making all actors partly stubborn or introducing a fraction of completely stubborn actors are both not able to break the initial $50/50$ opinion distribution.\\
+\newline
+\textbf{Unbalanced starting conditions}\\
+\newline
+The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and of opinion B is $P_\text{B}(0) = 0.8$.
 First, we discuss the formation of echo chambers.
@@
 \newline
-Fig.
-\ref{echo_vs_all_frac_stubb_PR-REC_20-80} shows the results for the formation of echo chambers versus the stubbornness parameter $r$ of all the actors for an initial $20/80$ distribution.
-The standard error $SE$ is small and not visible in the figure.
-Fig.
-\ref{PR_all_frac_stubb_20-80} gives the results for the PR filtering algorithm and Fig.
-\ref{REC_all_frac_stubb_20-80} gives the results for the REC filtering algorithm.
-The $y$-axis is interpreted in the same way as explained for Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80}.\\
-\newline
 In case of the PR filtering algorithm (Fig.
-\ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for all the network models if $r \leqslant 0.4$.
-If $r > 0.4$, the formation of echo chambers starts to decrease for all the network models.
-This behavior is slightly different than in case of balanced starting conditions, combined with the PR filtering algorithm (Fig.
+\ref{PR_all_frac_stubb_20-80}), the formation of echo chambers remains more or less flat for $r \leqslant 0.4$.
+If $r > 0.4$, the formation of echo chambers starts to decrease.
+This behavior is different than in the configuration with balanced starting conditions (Fig.
 \ref{echo_vs_all_frac_stubb_PR}).
-In that case, we could already observe a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with a high modularity which showed a decrease if $r>0.1$).
-This difference is probably a consequence of the unbalanced starting conditions, which seems to make the system less sensitive to the stubbornness parameter $r$.\\
-\newline
-As was explained for Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with a low modularity in case of unbalanced starting conditions, combined with the PR filtering algorithm is non-zero.
-These figures are not suited to (quantitatively) compare the formation of echo chambers in the different network models in case of unbalanced starting conditions.
-In order to perform such a (quantitative) comparison, one should look at the complete distribution (see e.g.
-Fig.
-\ref{echo_20_80}), as was explained for Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80}.
+In that case, we observed a decrease in the formation of echo chambers for $r > 0$ (except for the SBM with high modularity which showed a decrease if $r>0.1$).\\
+\newline
+As explained for Fig.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with low modularity for unbalanced starting conditions, combined with the PR filtering algorithm, is non-zero.
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the actors, $r$, for an initial $20/80$ opinion distribution.
+  \caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for unbalanced starting conditions.
+The PR and REC filtering algorithms are used.]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for unbalanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represent a constant amount of echo chambers over time.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard error $SE$.}
+Bars represent the standard error $SE$; these are small and not visible.}
 \label{echo_vs_all_frac_stubb_PR-REC_20-80}
@@
 \noindent
-The results for the REC filtering algorithm (Fig.
-\ref{REC_all_frac_stubb_20-80}) resemble those for balanced starting conditions (Fig.
+Results for the REC filtering algorithm (Fig.
+\ref{REC_all_frac_stubb_20-80}) qualitatively resemble those for balanced starting conditions (Fig.
 \ref{REC_all_frac_stubb}).
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
-The increase seems less steep than the increase for balanced starting conditions, which we can observe in Fig.
-\ref{REC_all_frac_stubb}.
-This again seems to suggest that the system is less sensitive to the stubbornness parameter $r$ for unbalanced starting conditions than for balanced starting conditions.\\
-\newline
-The stochastic block model with a high modularity shows a slight increase in the formation of echo chambers up to $r = 0.8$.
-This increase can even be observed for the low modularity case, albeit less strong.
-The Erd\H{o}s-R\'{e}nyi network wiggles around a formation of echo chambers equal to one.
-\newpage
+Networks with high clustering (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
+The SBM with high modularity shows a small increase in the formation of echo chambers up to $r = 0.8$.
+This increase can even be observed for the low modularity configuration, albeit less strong.
+The ER network wiggles around a formation of echo chambers equal to one.\\
+\newline
+As for the configuration in which a fraction of completely stubborn actors is introduced, the formation of echo chambers is smaller for unbalanced starting conditions than for balanced starting conditions.
+This quantitative difference is most likely a consequence of the higher initial number of echo chambers in case of unbalanced starting conditions.
+
 \begin{figure}[H]
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the actors ($r = 0$, $r = 0.4$ and $r = 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different values of the stubbornness parameter $r$ of all the actors ($r = 0$, $r = 0.4$ and $r = 0.8$).
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represent the initial prevalence of opinion A.
+The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-Bars represent the standard deviation $\sigma$.}%more explanation?
+Bars represent the standard deviation $\sigma$.}
 \label{ev_op_20_80_all_frac_stubb_PR}
@@
 \noindent
-We will now discuss the results for the evolution of the prevalence of the opinions in case of unbalanced starting conditions and making all the actors partly stubborn.
-We investigate four different network models.
-For each model the results for three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$) are compared.
-Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
-\newline
-The prevalence of opinion A versus time in case of the REC filtering algorithm is not affected by making all the actors more and more stubborn.
-The prevalence of the opinions is maintained and the status quo is not broken.
-The PR filtering algorithm gives us another story.
-The results for the prevalence of opinion A over time for an initial $20/80$ opinion distribution in case of the PR filtering algorithm and making all the actors partly stubborn are given in Fig.
+Lets now study the effect of the resistance parameter of the actors on the evolution of the prevalence of the opinions in case of unbalanced starting conditions.
+Three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$) are compared.
+When using the REC filtering algorithm, the prevalence of opinion A versus time is not affected by making the actors more and more stubborn.
+Results for the prevalence of opinion A over time when using the PR filtering algorithm are given in Fig.
 \ref{ev_op_20_80_all_frac_stubb_PR}.
-If the actors are not stubborn, the PR filtering algorithm reduces the prevalence of the minority opinion over time.
-Making the actors more and more stubborn reduces this tendency.
-This result is more or less similar to the corresponding case in which a fraction of completely stubborn actors was introduced in the system (Fig.
+If the actors are not stubborn, the PR filtering algorithm reduces the prevalence of the minority opinion.
+Making the actors more and more stubborn hampers this tendency.
+This qualitatively resembles the situation in which a fraction of completely stubborn actors is introduced (Section \ref{complStubbAct}, Fig.
 \ref{ev_op_20_80_stubb_PR}).
-The only difference is that making all the actors partly stubborn seems to be less effective in reducing the decrease of the prevalence of the minority opinion compared to making a fraction of the actors completely stubborn.
+The quantitative difference is that giving all actors the same resistance parameter seems to be less effective in reducing the decrease of the minority opinion compared to making a fraction of the actors completely stubborn.
 For $r=0.4$, the decrease of the prevalence of the minority opinion is almost not hampered, it is merely slowed down.
-All the actors need to be really stubborn ($r=0.8$) to see a hampering in the decrease of the prevalence of the minority opinion.
-Even then, the final prevalence of the minority opinion remains lower than for the case where a fraction of completely stubborn actors equal to fracRes $=0.4$ was introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$, respectively).
+The actors need to have high resistance levels ($r=0.8$) to see a hampering in the decrease.
+Even then, the final prevalence of the minority opinion remains smaller compared to the situation in which a fraction of completely stubborn actors $f_s=0.4$ is introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$, respectively).
 
@@
 
-In this subsection, the results of the majority threshold model are shared and discussed.
-This model is explained in Chapter \ref{chap3}, Section \ref{stubbBin}.
-We do not use a probabilistic majority model, but a classic majority model, as explainded in Chapter \ref{chap3}, Section \ref{majModel}, is implemented.
-The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Chapter \ref{chap4}, Section \ref{opinionmodel}.
-The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line of the actors is $S=20$.
-The threshold parameter $T$ is equal for all the actors and varies between $T=0$ and $T=1$.\\
-\newline
-For this model we only investigate the case of balanced starting conditions.
-We focus on the formation of echo chambers versus the threshold parameter $T$ and we compare three different filtering algorithms: the PR, REF and REC filtering algorithms.
+The results for the majority threshold model are shared and discussed.
+This model is explained in Section \ref{stubbBin}.
+We do not use a probabilistic majority model, but a classic majority model, as explainded in Section \ref{majModel}, is implemented.
+The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Section \ref{modelThesis}.
+The threshold parameter $T$ is equal for all the actors and varies between $T=0$ and $T=1$.
+We only investigate balanced starting conditions ($50/50$), with the focus on the formation of echo chambers.
+The PR, REF and REC filtering algorithms are compared.
 \\
@@
 Fig.
-\ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ for an initial $50/50$ distribution.
-In this figure, the PR filtering algorithm is used.
-Data points are shown together with their standard deviation $\sigma$ to emphasize on the spread in the data.
-The $y$-axis is interpreted in the same way as explained for Fig.
-\ref{echo_vs_mod} in Subsection \ref{echoVSmod}.\\
-\newline
-The high clustering models (WS and SBM-WS) show a large formation of echo chambers, which remain more or less flat for $T \leqslant 0.5$.
-This makes sense, since for $T \leqslant 0.5$, the threshold has no effect.
-This is because the majority opinion of the neighbors of an actor will always be larger than the threshold if $T \leqslant 0.5$ (the sum of the two opinion fractions is always equal to one, so the fraction of neighbors with the majority opinion will always take values between $0.5$ and $1$).
+\ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ when using the PR filtering algorithm.
+The high clustering models (WS and SBM-WS) show a significant formation of echo chambers, which remains more or less flat for $T \leqslant 0.5$.
+This is expected, because if $T \leqslant 0.5$, the threshold has no effect: the majority opinion of the neighbors of an actor will be larger than the threshold (the sum of the two opinion fractions equals one, so the fraction of neighbors with the majority opinion will take values between $0.5$ and $1$).
 If $T \geqslant 0.5$, it starts to have an effect.
-The actors become more hesitant to change their opinion and the formation of echo chambers starts to decrease.\\
-\newline
-Similar observations can be made for the stochastic block model with a high modularity.
-The only difference is of a quantitative nature (the formation of echo chambers is not as large as for the high clustering networks).\\
-\newline
-The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points lie behind those of the Erd\H{o}s-R\'{e}nyi network) do not enhance the formation of echo chambers in any way.
+The actors become more hesitant to change their opinion and need a higher fraction of neighbors to convince them.
+Hence, the formation of echo chambers starts to decrease.
+Similar observations can be made for the SBM with high modularity.
+The difference is of a quantitative nature: the formation of echo chambers is not as large as for the networks with high clustering.
+The ER network and the SBM with low modularity (data points lie behind those of the ER network) do not enhance the formation of echo chambers.
 Including a threshold $T$ does not change this tendency.
-% make coupling to ev of op??
-% explain opinion dynamics or refer to relevant sections!!!
+
 \begin{figure}[H]
@@
 	\captionsetup{format=plain}
-	\caption[Formation of echo chambers versus the threshold parameter of all the actors, $T$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.
-The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the actors, $T$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm.
-The majority threshold model is implemented.
+	\caption[Formation of echo chambers versus the threshold parameter $T$, for balanced starting conditions.
+The PR filtering algorithm is used and the majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter $T$, for balanced starting conditions.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
+The PR filtering algorithm is used.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard deviation $\sigma$.}
+Bars represent the standard deviation $\sigma$ to emphasize on the spread in the data.}
 \label{echo_vs_threshold_PR}
@@
 Fig.
-\ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF (Fig.
-\ref{REF_thres}) and REC (Fig.
-\ref{REC_thres}) filtering algorithms.
-The initial opinion distribution is $50/50$.
-The results for the REF and REC filtering algorithms resemble each other.
-Data points are again shown together with their standard deviation $\sigma$ to emphasize the large spread in the data (the standard deviations are of the order $\sigma \sim 150$).
+\ref{echo_vs_threshold_REF-REC} shows the formation of echo chambers versus the threshold $T$ for the REF and REC filtering algorithms.
+Results for the REF and REC filtering algorithms resemble each other.
+The standard deviations are of the order $\sigma \sim 150$.
 These large errors are a result of the majority model.
-If we run the majority model once, the model forces the system to converge to a single opinion.
-To which of the two opinions the model converges is a matter of random fluctuations.
-On average, the opinion distribution will remain $50/50$.
-However, each run the system converges to one of the two opinions, with the result that the number of echo chambers around this opinion strongly increases.
-After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (and thus resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in almost no echo chambers around opinion B).
+If we simulate the majority model once, the model forces the system to converge to a single opinion.
+To which of the two opinions the model converges is matter of random fluctuations.
+In the average of different independent simulations, the opinion distribution will remain $50/50$.
+However, each independent simulation the system converges to one of the two opinions, with the result that the number of echo chambers around that opinion strongly increases.
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in no echo chambers around opinion B).
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus the threshold parameter of all the actors, $T$, for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter of all the actors, $T$, for an initial $50/50$ opinion distribution.
-The majority threshold model is implemented.
+  \caption[Formation of echo chambers versus the threshold parameter $T$ for balanced starting conditions.
+The REF and REC filtering algorithms are used and the majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter $T$ for balanced starting conditions.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
-Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$.
 Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard deviation $\sigma$.}
+Bars represent the standard deviation $\sigma$ to emphasize on the spread in the data.}
 \label{echo_vs_threshold_REF-REC}
@@
 \noindent
-Another consequence of this converging to one of the two opinions every single simulation, is the fact that the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a strong appearance of echo chambers for $T \leqslant 0.5$.
-This is not due to any network features, but a direct consequence of the usage of the majority model.
-As explained above, each simulation the system converges to one of the two opinions with as result a strong increase in the echo chambers around that opinion (as a matter of fact the whole system becomes a large, single echo chamber).
-The PR filtering algorithm seems to hamper this convergence to one opinion each run and hence, this artificial creation of echo chambers.
+The ER network and the SBM with low modularity show a significant appearance of echo chambers for $T \leqslant 0.5$.
+This is not due to any network topologies, but a consequence of the convergence to one of the two opinions each independent simulation.
+This convergence results in a strong increase in the echo chambers around that opinion (the whole system becomes a large, single echo chamber).
+The PR filtering algorithm seems to hamper this convergence and hence, this artificial creation of echo chambers.
 This can be seen in Fig.
-\ref{echo_vs_threshold_PR} where the ER and the SBM with a low modularity do not show any creation of opinion clusters.
-Also note that the standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REF or REC filtering algorithms.
-This is another indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.
+\ref{echo_vs_threshold_PR} where the ER and the SBM with low modularity do not show a formation of opinion clusters.
+The standard deviations when using the PR filtering algorithm are significantly smaller (of the order $\sigma \sim 50$) than those when using the REF or REC filtering algorithms; another indication that the PR filtering algorithm hampers the convergence to a single opinion each simulation.
 If $T \geqslant 0.5$, it starts to have an impact.
-However, in Fig.
-\ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers for all the network models until the echo chambers size reaches values equal to one for $T=1$.
-At this point, none of the actors can be convinced of another opinion and hence all the actors keep their initial opinion.
-The number of echo chambers does not change over time, which results in an echo chamber size equal to one in the figure.
-A constant amount of echo chambers over time is represented by the black dotted line.\\
+In Fig.
+\ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers until the echo chambers size reaches values equal to one for $T=1$.
+At this point, none of the actors can be convinced of another opinion and hence all actors retain their initial opinion.
+The number of echo chambers does not change over time, which results in an echo chamber size equal to one.
+\\
 \newline
 Fig.
-\ref{ev_op_majT_one_sim} illustrates the particular behavior of the majority model more clearly.
-The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of actors with opinion A; the reason for this is that it makes a comparison of the results easier and more clear; it must however be emphasized that every simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
+\ref{ev_op_majT_one_sim} illustrates this particular behavior of the majority model.
+The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of actors with opinion A, since this makes a comparison of the results easier; it must be emphasized that every independent simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
 The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
@@
 The PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
-Furthermore, the high clustering networks (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
-This may indicate that a high clustering is a protective factor against the convergence to one opinion each independent simulation in case of the majority model.
-Having a clear community structure (high modularity) in case of the stochastic block model seems to be another hindering of this convergence, as can be seen in Fig.
+Furthermore, networks with high clustering (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
+This indicates that high clustering is a protective factor against the convergence to one opinion each independent simulation.
+Having community structure (high modularity) in case of the SBM is another hindering of this convergence, as can be seen in Fig.
 \ref{sbm_majT}.
-For the REC filtering algorithm, the high modularity stochastic block model converges to lower values than the low modularity stochastic block model.
-For the PR filtering algorithm, this is not really the case, but this is probably due to the fact that the PR filtering algorithm already strongly hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
+For the REC filtering algorithm, the SBM with high modularity converges to lower values than the SBM with low modularity.
+For the PR filtering algorithm, this is not the case, but this is probably due to the fact that the PR filtering algorithm already hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ opinion distribution.
-Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $50/50$ distribution.
-The majority threshold model is implemented with $T=0$ and results are shown for a single simulation.
-Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} ER, $p=0.01$; \textbf{(b)} WS with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$.
-For each model the REC and the PR filtering algorithms are compared.}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for balanced starting conditions.
+Results for one simulation of the majority threshold model with $T=0$.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for balanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents the initial prevalence of opinion A.
+For each model the REC and the PR filtering algorithms are compared.
+Results are shown for a single simulation.}
 \label{ev_op_majT_one_sim}
@@
 
-The formation of echo chambers versus the fraction of completely stubborn actors, fracRes, in the Last.fm on-line social network is studied.
+The formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, in the Last.fm on-line social network is studied.
 The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
@@
   \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, for an initial $50/50$ opinion distribution.
 The Last.fm on-line social network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM-WS with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$; (iii) SBM with 53 communities of size 151 with $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$.
@@
 In Fig.
-\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of completely stubborn actors (fracRes) is studied for the real-world PGP trust network and for the corresponding network models.
+\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of completely stubborn actors is studied for the real-world PGP trust network and for the corresponding network models.
 This is done for both the PR and REC filtering algorithms.
@@
   \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution.
+The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, for an initial $50/50$ opinion distribution.
 The PGP trust network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM-WS with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$; (iii) SBM with 60 communities of size 178 with $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$.

### 2021-05-24 draft.tex
Update draft.tex
added something in preambule to show numbering of subsubsection
latest update

@@
 
+\setcounter{tocdepth}{3}
+\setcounter{secnumdepth}{3}
+
 \begin{document}
@@
 \newline
-\textbf{The Watts-Strogatz model in this thesis}\\
+\textbf{Watts-Strogatz model in this thesis}\\
 \newline
@@
 \newline
-\textbf{The stochastic block model in this thesis}\\
+\textbf{Stochastic block model in this thesis}\\
 \newline
@@
 This kind of activation mechanism of the actors gives the network a temporal character.
-The underlying network structure does not change (each actor has a fixed number of neighbors that does not change over time; no new friends are made or, conversely, no old friends are being defriended).
+The underlying network structure does not change (each actor has a fixed number of neighbors that does not change over time; no new friends are made or, conversely, no old friends are being unfriended).
 However, by making actors active and inactive the exchange of information between actors becomes time dependent on which actors are active or not.
@@
 
-\begin{algorithm}
+\begin{algorithm}[H]
 \caption{Updating the opinions of the active actors at time $t$}
@@
 \noindent
-Some pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
+Pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
 The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
@@
     \caption{}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC_8x7.png}
     \label{sbm_50-50}
@@
 \end{figure}
-\newpage
 \noindent
@@
     \caption{}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC.png}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_50-50_REC_8x7.png}
       \end{subfigure}
@@
 Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-For the WS model, we use two different values for the rewire probability: $\beta = 0.01$ ($\left<cc\right> = 0.648 \pm 0.003$) and $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+For the WS model, we use two different values for the rewire probability: $\beta = 0.01$ ($\left<cc\right> = 0.648 \pm 0.003$; WS2) and $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$; WS1).
 The dotted black line represents a constant amount of echo chambers over time.
@@
 When using the PR filtering algorithm, there is an imbalance in the sense that actors with the minority opinion A have a chance of switching to the majority opinion B, whereas actors with the majority opinion B have (almost) no chance of switching to the minority opinion A.
-This imbalance leads to the decrease of the minority opinion A.
+This imbalance leads to the decrease of the minority opinion A.\\
+\newline
+Fig.
+\ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
+This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
+The high clustering in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
+\ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Fig.
+\ref{ws_20-80_op}).
 
@@
 \noindent
-Fig.
-\ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
-This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
-The high clustering in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
-\ref{sbm-ws_20-80_op}) and an average value of $P_\text{A}(500) = 0.009 \pm 0.015$ for the WS network (Fig.
-\ref{ws_20-80_op}).\\
-\newline
 This result differs from results in \cite{Perra2019}, where the minority opinion was able to survive over time for all the network models, even for the PR filtering algorithm, and where a minority opinion prevalence at $t=500$ between $P_\text{A}(500) = 0.06$ and $P_\text{A}(500) = 0.09$ was reported \cite{Perra2019}.
@@
 Similarly, the results for the WS model and the SBM-WS agree with each other as well.
-This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is the clustering that drives these differences.
+This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is the clustering that drives these differences.\\
+\newline
+There are no visible effects of having a community structure.
+Fig.
+\ref{sbm} shows the results for a SBM with high modularity (community structure) and low modularity (negligible community structure).
+There is (almost) no difference in the results for high and low modularity.
 
@@
 \noindent
-There are no visible effects of having a community structure.
-Fig.
-\ref{sbm} shows the results for a SBM with high modularity (community structure) and low modularity (negligible community structure).
-There is (almost) no difference in the results for high and low modularity.\\
-\newline
 Lets first discuss the results for the ER model and the SBM (Fig.
@@
 A high modularity is however required to see the effect of community structure on the formation of opinion clusters.
-
+\newpage
 \begin{figure}[H]
@@
 \end{figure}
-
-\newpage
+\noindent
 \section{Stubborn actors}
@@
 \newline
-Lets now study the situation for unbalanced starting conditions. The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and of opinion B is $P_\text{B}(0) = 0.8$.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
-    \label{PR_frac_compl_stubb_20-80}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
-    \label{REC_frac_compl_stubb_20-80}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of stubborn actors for the PR and REC filtering algorithms and unbalanced starting conditions.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for unbalanced starting conditions.
-\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
-Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
-The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represents a constant amount of echo chambers over time.
-The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
-Each data point is the average of $10^2$ independent simulations.
-Bars represent the standard error $SE$; these are small and not visible.}
-\label{echo_vs_frac_complStubb_PR-REC_20-80}
-\end{figure}
-\noindent
+Lets now study the situation for unbalanced starting conditions. The initial prevalence of opinion A is $P_\text{A}(0) = 0.2$ and of opinion B is $P_\text{B}(0) = 0.8$.\\
+\newline
 In Fig.
@@
 When comparing the different network models, the skewed distributions should not be overlooked.
-The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other.\\
-\newline
+The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_PR_20-80_8x7.png}
+    \label{PR_frac_compl_stubb_20-80}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_frac_compl_stubb_REC_20-80_8x7.png}
+    \label{REC_frac_compl_stubb_20-80}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of stubborn actors for the PR and REC filtering algorithms and unbalanced starting conditions.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for unbalanced starting conditions.
+\textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represents a constant amount of echo chambers over time.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+Each data point is the average of $10^2$ independent simulations.
+Bars represent the standard error $SE$; these are small and not visible.}
+\label{echo_vs_frac_complStubb_PR-REC_20-80}
+\end{figure}
+\noindent
 When using the PR filtering algorithm (Fig.
@@
 \newline
-We can conclude that the formation of echo chambers versus the fraction of stubborn actors, for unbalanced starting conditions, qualitatively resembles the results for balanced starting conditions: decrease in formation of echo chambers with increasing fraction of stubborn actors for the PR filtering algorithm versus increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm.
+We can conclude that the formation of echo chambers versus the fraction of stubborn actors for unbalanced starting conditions qualitatively resembles the results for balanced starting conditions: decrease in formation of echo chambers with increasing fraction of stubborn actors for the PR filtering algorithm versus increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease for the REC filtering algorithm.
 There is, however, a quantitative difference: the formation of echo chambers is lower when using unbalanced starting conditions compared to using balanced starting conditions.
-The reason for this is probably the higher initial number of echo chambers when using unbalanced starting conditions.
+The reason for this is probably the higher initial number of echo chambers when using unbalanced starting conditions.\\
+\newline
+For balanced starting conditions we found that introducing stubborn actors did not affect the prevalence of the opinions over time.
+Lets now study whether this is also the case for unbalanced starting conditions.
+In Fig.
+\ref{ev_op_20_80}, we saw that a configuration with no stubborn actors maintained the prevalence of the initial opinions if the REC filtering algorithm is used.
+The PR filtering algorithm resulted in a decrease in the prevalence of the minority opinion.
+Introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
+The status quo is maintained over time.
+Introducing stubborn actors combined with the PR filtering algorithm does have an impact.\\
+\newline
+Fig.
+\ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time when using the PR filtering algorithm.
+We compare four network models (ER, WS, SBM (high and low modularity) and SBM-WS).
+For each model we compare the results for three different fractions of stubborn actors ($f_s=0$, $f_s=0.4$ and $f_s=0.8$).\\
+\newline
+If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion.
+For an increasing fraction of stubborn actors, this decrease becomes less and less strong.
+This result is valid for the different network topologies.
+For no stubborn actors, the PR filtering algorithm forces actors with the minority opinion to switch to the majority opinion.
+If we introduce a fraction of stubborn actors, we will have stubborn actors that carry the minority opinion.
+These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them and hamper the decrease of the minority opinion.
 
@@
 \end{figure}
-\noindent
-For balanced starting conditions we found that introducing stubborn actors did not affect the prevalence of the opinions over time.
-Lets now study whether this is also the case for unbalanced starting conditions.
-In Fig.
-\ref{ev_op_20_80}, we saw that a configuration with no stubborn actors maintained the prevalence of the initial opinions if the REC filtering algorithm is used.
-The PR filtering algorithm resulted in a decrease in the prevalence of the minority opinion.
-Introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm.
-The status quo is maintained over time.
-Introducing stubborn actors combined with the PR filtering algorithm does have an impact.\\
-\newline
-Fig.
-\ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time when using the PR filtering algorithm.
-We compare four network models (ER, WS, SBM (high and low modularity) and SBM-WS).
-For each model we compare the results for three different fractions of stubborn actors ($f_s=0$, $f_s=0.4$ and $f_s=0.8$).
-\\
-\newline
-If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion.
-For an increasing fraction of stubborn actors, this decrease becomes less and less strong.
-This result is valid for the different network topologies.
-For no stubborn actors, the PR filtering algorithm forces actors with the minority opinion to switch to the majority opinion.
-If we introduce a fraction of stubborn actors, we will have stubborn actors that carry the minority opinion.
-These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them and hamper the decrease of the minority opinion.
 
@@
 As explained for Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with low modularity for unbalanced starting conditions, combined with the PR filtering algorithm, is non-zero.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers in the ER network and the SBM with low modularity for unbalanced starting conditions, combined with the PR filtering algorithm, is non-zero.\\
+\newline
+Results for the REC filtering algorithm (Fig.
+\ref{REC_all_frac_stubb_20-80}) qualitatively resemble those for balanced starting conditions (Fig.
+\ref{REC_all_frac_stubb}).
+Networks with high clustering (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
+The SBM with high modularity shows a small increase in the formation of echo chambers up to $r = 0.8$.
+This increase can even be observed for the low modularity configuration, albeit less strong.
+The ER network wiggles around a formation of echo chambers equal to one.\\
+\newline
+As for the configuration in which a fraction of completely stubborn actors is introduced, the formation of echo chambers is smaller for unbalanced starting conditions than for balanced starting conditions.
+This quantitative difference is most likely a consequence of the higher initial number of echo chambers in case of unbalanced starting conditions.
 
@@
 \label{echo_vs_all_frac_stubb_PR-REC_20-80}
-\end{figure}
-\noindent
-Results for the REC filtering algorithm (Fig.
-\ref{REC_all_frac_stubb_20-80}) qualitatively resemble those for balanced starting conditions (Fig.
-\ref{REC_all_frac_stubb}).
-Networks with high clustering (WS and SBM-WS) show an increase in the formation of echo chambers up to $r=0.8$, after which a sharp drop to one, for $r=1$, is observed.
-The SBM with high modularity shows a small increase in the formation of echo chambers up to $r = 0.8$.
-This increase can even be observed for the low modularity configuration, albeit less strong.
-The ER network wiggles around a formation of echo chambers equal to one.\\
-\newline
-As for the configuration in which a fraction of completely stubborn actors is introduced, the formation of echo chambers is smaller for unbalanced starting conditions than for balanced starting conditions.
-This quantitative difference is most likely a consequence of the higher initial number of echo chambers in case of unbalanced starting conditions.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
-The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
-\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
-The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represent the initial prevalence of opinion A.
-The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
-Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
-Bars represent the standard deviation $\sigma$.}
-\label{ev_op_20_80_all_frac_stubb_PR}
 \end{figure}
@@
 Even then, the final prevalence of the minority opinion remains smaller compared to the situation in which a fraction of completely stubborn actors $f_s=0.4$ is introduced ($P_\text{A}(500) = 0.064 \pm 0.017$ versus $P_\text{A}(500) = 0.099 \pm 0.009$, respectively).
+\newpage
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_all_frac_stubb_PR_20-80_8x7.png}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
+The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+The black dotted line represent the initial prevalence of opinion A.
+The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
+Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
+Bars represent the standard deviation $\sigma$.}
+\label{ev_op_20_80_all_frac_stubb_PR}
+\end{figure}
 
@@
 The number of echo chambers does not change over time, which results in an echo chamber size equal to one.
-\\
-\newline
-Fig.
-\ref{ev_op_majT_one_sim} illustrates this particular behavior of the majority model.
-The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of actors with opinion A, since this makes a comparison of the results easier; it must be emphasized that every independent simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
-The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
-For each model the REC and PR filtering algorithms are compared.
-The threshold parameter is set equal to $T=0$, so there is no effect of stubbornness.\\
-\newline
-The PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
-Furthermore, networks with high clustering (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
-This indicates that high clustering is a protective factor against the convergence to one opinion each independent simulation.
-Having community structure (high modularity) in case of the SBM is another hindering of this convergence, as can be seen in Fig.
-\ref{sbm_majT}.
-For the REC filtering algorithm, the SBM with high modularity converges to lower values than the SBM with low modularity.
-For the PR filtering algorithm, this is not the case, but this is probably due to the fact that the PR filtering algorithm already hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
 
@@
 \end{figure}
+\noindent
+Fig.
+\ref{ev_op_majT_one_sim} illustrates this particular behavior of the majority model.
+The figure shows the evolution of the prevalence of opinion A for a single simulation (we always chose results that show an increase in the fraction of actors with opinion A, since this makes a comparison of the results easier; it must be emphasized that every independent simulation there is an equal probability of converging to opinion A as to opinion B, the opinion to which the system eventually converges is a consequence of random fluctuations).
+The figure shows the results for each of the four network models: ER, WS, SBM (high and low modularity) and SBM-WS.
+For each model the REC and PR filtering algorithms are compared.
+The threshold parameter is set equal to $T=0$, so there is no effect of stubbornness.\\
+\newline
+The PR filtering algorithm hampers the convergence to a single opinion each independent simulation.
+Furthermore, networks with high clustering (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
+This indicates that high clustering is a protective factor against the convergence to one opinion each independent simulation.
+Having community structure (high modularity) in case of the SBM is another hindering of this convergence, as can be seen in Fig.
+\ref{sbm_majT}.
+For the REC filtering algorithm, the SBM with high modularity converges to lower values than the SBM with low modularity.
+For the PR filtering algorithm, this is not the case, but this is probably due to the fact that the PR filtering algorithm already hampers the convergence to a single opinion; having a high or low modularity does not give an additional effect on top of this PR filtering effect.
 
@@
 
-This section is devoted to results regarding the way the opinions are distributed over the network.
-In all the previous results, the two opinions were, initially, randomly distributed across the network.
-It might however be more realistic if people with the same opinion are more or less grouped together initially.
-We did introduce a network model with a community structure (the SBM) to represent the group structures found in real-life systems.
-But we have not yet distributed the opinions according to those groups.
-\\
-\newline
-We study the stochastic block model (SBM) and we give a fraction of the communities a single opinion, namely opinion A.
+Previously, the two opinions were, initially, randomly distributed across the network.
+It might be more realistic if people with the same opinion are grouped together initially.
+We did introduce a network model with a community structure (SBM) to represent group structures found in real-life.
+But we have not yet distributed the opinions according to these groups.
+\\
+\newline
+We study the SBM and give a fraction of the communities a single opinion, opinion A.
 The remaining communities are given a $50/50$ opinion distribution.
-This situation may, in an idealized way, represent the situation in which you have groups of people following the same idea together with groups where the two ideas coexist.
-The total opinion distribution over the entire network is no longer $50/50$, but instead opinion A becomes the majority opinion (hence, opinion B becomes the minority opinion).
-The higher the fraction of single opinion A communities, the bigger the difference between this majority and minority opinion.
-The results will be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community.
-Thus, e.g.
+This situation may represent real-life with groups of people following the same idea together with groups where the two ideas coexist.
+The total opinion distribution over the entire network is no longer $50/50$; opinion A becomes the majority opinion.
+The higher the fraction of single opinion A communities, the bigger the difference between the majority and minority opinion.
+Results will be compared to the configuration in which the total opinion distribution is randomly distributed over each community.
+E.g.
 if we have 10 communities of size 100 and we make $10 \%$ of those communities single opinion A communities while the others get a $50/50$ distribution, the total initial opinion distribution is $55/45$.
-This situation is then compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\
-\newline
-We will study a network with a high modularity, representing the case of more or less isolated communities, and a network with a low modularity, representing a more blended community structure in which there is a lot of communication between the different communities.
-More specifically, the two stochastic block models that are implemented are: (i) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ ($Q = 0.908 \pm 0.005$); and (ii) SBM with 10 communities of size 100 with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ ($Q = 0.220 \pm 0.007$).
-Both networks have a total number of nodes $N = 10^3$ and average degree $\left<k\right> = 10.79 \pm 0.14$ and $\left<k\right> = 10.17 \pm 0.14$, respectively.\\
-\newline
-The opinion dynamics model is described in Chapter \ref{chap4}, Section \ref{opinionmodel}.
-The activation probability is set equal to $p_{\text{act}} = 0.1$ and the maximum length of the actual time-line is $S=20$.
-\\
-\newline
-In subsection \ref{groupVSrandomOP} we will discuss the results for the evolution of the prevalence of the opinions. Subsection \ref{groupVSrandomECHO} will discuss the formation of echo chambers.
+This configuration is then compared to the configuration in which a $55/45$ distribution is randomly distributed over each community.\\
+\newline
+Two configurations of the SBM are implemented: a high modularity configuration ($Q = 0.908 \pm 0.005$) with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ and a low modularity configuration ($Q = 0.220 \pm 0.007$) with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$.
+For both configurations the SBM has 10 communities of size 100 nodes each.
+The high modularity configuration has $\left<k\right> = 10.79 \pm 0.14$; the low modularity configuration has $\left<k\right> = 10.17 \pm 0.14$.\\
+\newline
+In section \ref{groupVSrandomOP} the results for the evolution of the prevalence of the opinions are discussed. Section \ref{groupVSrandomECHO} discusses the formation of echo chambers.
 
@@
 
-First, we discuss how the evolution of the opinions are impacted by distributing the opinions in communities instead of randomly distributing them.\\
-\newline
-Fig.
-\ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
-A fraction equal to $0.1$ of the communities is given a single opinion, opinion A, while the other communities get an initial $50/50$ opinion distribution.
-The results are compared to the random case in which each community initially gets the corresponding total opinion distribution ($55/45$ in this case).
-The figure shows the results for both the high and the low modularity case.
-For each modularity the REC and PR filtering algorithms are compared.
+First, we discuss how the evolution of the prevalence of the opinions are impacted by distributing the opinions in communities instead of randomly distributing them.
 
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
-Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for making a fraction equal to $0.1$ of the communities single opinion A communities, while the others get an initial $50/50$ opinion distribution.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$).
+$10 \%$ of the communities is given a single opinion, opinion A, while the others get an initial $50/50$ opinion distribution.
+Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$).
+First row: high modularity configuration \textbf{(a)} REC; \textbf{(b)} PR.
+second row: low modularity configuration \textbf{(c)} REC; \textbf{(d)} PR.
+$10 \%$ of the communities is given a single opinion, opinion A.
 Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.
-The stochastic block model with 10 communities of size 100 is implemented with two different modularities: $Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$, $\left<k\right> = 10.79 \pm 0.14$ and $Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$, $\left<k\right> = 10.17 \pm 0.14$.
-$N = 10^3$ for both modularities.
-\textbf{(a)} High modularity, REC; \textbf{(b)} High modularity, PR; \textbf{(c)} Low modularity, REC; \textbf{(d)} Low modularity, PR.
+The dotted black line represents the initial prevalence of opinion A.
 Each curve is the average of $10^2$ independent simulations and data points are shown every 50 time steps to improve visualization.
@@
 \end{figure}
-\newpage
 \noindent
-In case of the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
+Fig.
+\ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
+$10 \%$ of the communities is given a single opinion, opinion A.
+Results are compared to the case in which the total opinion distribution ($55/45$) is randomly distributed.
+The high and low modularity configuration are depicted.
+For each modularity the REC and PR filtering algorithms are compared.\\
+\newline
+For the REC filtering algorithm there is no difference between distributing the opinions randomly versus distributing them in communities.
 The prevalence of the majority opinion A remains unchanged over time.
-This result is the same for both the high and low modularity case, see Fig.
+This result is valid for both the high and low modularity configuration, Fig.
 \ref{high_mod_rec} and Fig.
@@
 \newline
-The PR filtering algorithm leads to an increase in the prevalence of the majority opinion in case of unbalanced starting conditions and if the opinions are randomly distributed.
-This is a result that was found in Section \ref{netw_struc}, Subsection \ref{20-80}, see e.g.
-Fig.
-\ref{ev_op_20_80}.
-If a fraction equal to $0.1$ of the communities is given one opinion, opinion A, while the remaining communities get a $50/50$ opinion distribution, we see from Fig.
-\ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case.
-This makes sense, since the single opinion communities are more or less isolated from the mixed opinion communities.
-This has as consequence that the mixed opinion communities live under the assumption that the opinion distribution is $50/50$.
-They do not really feel that opinion A is a majority opinion, since they are isolated from these single opinion A communities.
-In case of the low modularity network, Fig.
-\ref{low_mod_pr}, this isolation argument of the different communities no longer holds.
-This is also visible in the results since the hampering in the increase of the majority opinion is no longer present.
+If the opinions are randomly distributed and if the starting conditions are unbalanced, the PR filtering algorithm leads to an increase in the prevalence of the majority opinion.
+This confirms results found in Section \ref{20-80} (Fig.
+\ref{ev_op_20_80}).
+In the community distributed scenario, we see from Fig.
+\ref{high_mod_pr} that the high modularity configuration hampers the increase in the majority opinion compared to the randomly distributed case.
+This is expected, since the single opinion communities are more or less isolated from the mixed opinion communities.
+The consequence is that the mixed opinion communities live under the assumption that the opinion distribution is $50/50$.
+They do not feel the majority of opinion A, since they are isolated from these single opinion A communities.
+For the low modularity configuration, Fig.
+\ref{low_mod_pr}, this isolation argument no longer holds.
+This is visible in the results: the hampering in the increase of the majority opinion is no longer present.
 The community distributed case behaves equivalent to the random distributed case.
-In the low modularity network, the community structure is not clearly present, but becomes blended.
-The different communities talk to each other and there is a lot of communication between them.
-As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion A in the same way as if the opinions would be randomly distributed over the communities.
-\\
-\newline
-The results found in Fig.
-\ref{commOp0_01_other_50-50} for $10 \%$ of single opinion A communities remain valid if we increase the fraction of single opinion A communities.
+The community structure is not clearly present; there is a lot of communication between the communities.
+Hence, the mixed opinion communities feel the presence of the single opinion communities and feel the majority of opinion A in a similar way as if the opinions would be randomly distributed over the communities.\\
+\newline
+If we increase the fraction of single opinion A communities, these observations remain valid.
 This can be seen in Fig.
-\ref{fracOp0_vs_commOp0_other_50-50}, which depicts the average prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
-The community distributed cases are compared to the corresponding randomly distributed cases.
-Results are shown for both a high modularity and a low modularity network.
-Fig.
-\ref{rec_comm} shows the results for the REC filtering algorithm.
-Fig.
-\ref{pr_comm} shows those for the PR filtering algorithm.\\
-\newline
-The higher the fraction of single opinion A communities, the higher the initial average prevalence of opinion A.
-This is seen in both figures by the increase of the solid black line, which represents the initial prevalence of opinion A.
-In case of the REC filtering algorithm (Fig.
+\ref{fracOp0_vs_commOp0_other_50-50}, which depicts the prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
+The community distributed cases are compared to the randomly distributed cases.
+Results are shown for both a high and a low modularity configuration.\\
+\newline
+The higher the fraction of single opinion A communities, the higher the initial prevalence of opinion A.
+This is seen for both the REC and PR filtering algorithms by the increase of the solid black line.
+For the REC filtering algorithm (Fig.
 \ref{rec_comm}) the prevalence of opinion A remains the same before and after the time evolution for a fixed fraction of single opinion A communities.
-Giving a fraction of the communities a single opinion, opinion A, and the others an initial $50/50$ opinion distribution does not make a difference compared to randomly distributing the same total initial opinion distribution.
-This result is valid for both the high and low modularity case.\\
+Giving a fraction of the communities a single opinion, opinion A, does not make a difference compared to randomly distributing the same total initial opinion distribution.
+This result is valid for both the high and low modularity configuration.\\
 \newline
@@
 \ref{pr_comm}) shows an increase in the prevalence of opinion A over time for a fixed fraction of single opinion A communities.
-Making a fraction of the communities single opinion A communities, while the other communities get a $50/50$ opinion distribution, hampers this increase compared to the corresponding randomly distributed case, if we are dealing with a high modularity network with a clear and almost isolated community structure.
-The low modularity network does not exhibit this hampering.
-The reason for this hampering in case of the high modularity network and the absence of it for the low modularity network was explained for Fig.
+For the high modularity configuration, making a fraction of the communities single opinion A communities hampers this increase compared to the corresponding randomly distributed case.
+The low modularity configuration does not exhibit this hampering.
+The reason for the hampering in case of a high modularity and the absence of it for a low modularity configuration was explained for Fig.
 \ref{commOp0_01_other_50-50} and has to do with whether the single opinion A communities can influence the $50/50$ communities.
@@
   \caption[Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
-Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity network.
+Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity configuration.
 Results for the REC and PR filtering algorithms.]{Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
-Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
 \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Community distributed cases are compared to the corresponding randomly distributed cases for both a high and a low modularity configuration.
+The solid black line represents the initial prevalence of opinion A.
+The dotted lines represent the final prevalence of opinion A.
 Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for an improved visualization.
-}
+Error bars are omitted for improved visualization.}
 \label{fracOp0_vs_commOp0_other_50-50}
@@
 
-We will investigate the effect of community distributing the opinions versus randomly distributing them on the formation of echo chambers. \\
+We give a fraction of the communities a single opinion, opinion A, while the remaining ones get an initial $50/50$ opinion distribution.
+For a high modularity configuration, we found that this community distributed scenario hampers the increase in the majority opinion, opinion A, when using the PR filtering algorithm.
+We will now study the effect of this community distributed scenario on the formation of echo chambers.
+\\
 \newline
 Fig.
-\ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$ for the case where a fraction equal to $0.1$ of the communities is made a single opinion A community while the other communities get an initial $50/50$ opinion distribution.
-The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, represents the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$.
-The community distributed case is compared to the corresponding randomly distributed case (which is, in this case, an initial $55/45$ opinion distribution).
-The formation of echo chambers in both the high and low modularity network are investigated.
+\ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$.
+In the first row, $10 \%$ of the communities is given opinion A; in de second row, $50 \%$ of the communities are single opinion A communities .
+The community distributed case is compared to the random distributed case: $55/45$ in the 1st row; $75/25$ in the 2nd row.
+Both the high and low modularity configuration are studied.
 The REC and PR filtering algorithms are compared.
-\\
-\newline
-The results for the REC filtering algorithm are depicted in Fig.
-\ref{rec_comm_echo}.
-The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion A (leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (bump in the blue line).
-The number of neighborhoods where all the actors carry opinion A decreases in favor of neighborhoods where we find both actors with opinion A and actors with opinion B.
-Even though there is only a minor communication between the communities, this limited communication has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion A communities, reduces and that the opinions become more mixed.
-The prevalence of opinion A remains constant over time when using the REC filtering algorithm (Fig.
-\ref{high_mod_rec}).
-The communities with the single opinion A see, however, a decrease in the prevalence of opinion A and become more mixed.
-We also observe a small increase in the number of echo chambers around the minority opinion B.
-This indicates that the minority opinion starts to group together.
-The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion A and opinion B.
-Again, the opinion fractions remain constant over time, but are shifted from a random distribution to a more clustered situation.
-The increase remains however smaller than two, which indicates that there are less than twice as much echo chambers at $t=500$ than at $t=0$.\\
-\newline
-The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion A (which was observed for the community distributed, high modularity case).
-This is due to the fact that the number of echo chambers around opinion A at $t=0$, originating from the fraction of single opinion A communities, is smaller for the low modularity network than for the high modularity network.
-The reason is the more blended community structure for the low modularity network.
-We still observe a small increase in the number of echo chambers around opinion B.
-This increase in the formation of echo chambers around the minority opinion B seems however hampered compared to the randomly distributed case.
-Remember that the increase in the number of echo chambers around the minority opinion for the SBM in combination with the REC filtering algorithm for unbalanced starting conditions that are randomly distributed was observed in Section \ref{netw_struc}, Subsection \ref{20-80} (Fig.
-\ref{sbm}).
-\\
-\newline    % make link to opinion fraction 
-The results for the PR filtering algorithm are depicted in Fig.
-\ref{pr_comm_echo}.
-They are different from those obtained with the REC filtering algorithm, which we just discussed and which are depicted in Fig.
-\ref{rec_comm_echo}.\\
-\newline
-Lets first focus on the high modularity network.
-The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Fig.
-\ref{sbm} (these results were obtained for an initial $20/80$ opinion distribution).
-We observe a strong increase in the formation of echo chambers around the majority opinion A, with a corresponding decrease in all the neighborhoods that have a fraction of actors with the minority opinion B.
-Remember from Fig.
-\ref{high_mod_pr} that the PR filtering algorithm leads to an increase in the majority opinion A.
-If we, however, make $10 \%$ of the communities single opinion A communities while giving the remaining ones a $50/50$ opinion distribution, we see that this increase in the formation of echo chambers around the majority opinion A is strongly hampered.
-This hampering is partly an artifact of the way the opinions are distributed.
-If we give entire communities only opinion A while others get a $50/50$ distribution, we have a large number of echo chambers around opinion A at $t=0$ due to these single opinion communities.
-Moreover, the number of opinion A echo chambers at $t=0$ is higher than in the corresponding randomly distributed case.
-If we, however, would look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
-This shows that the observed hampering of the formation of echo chambers around the majority opinion A in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers.
-Remember from Fig.
-\ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion A (and thus the decrease of the minority opinion B); this hampering is probably one of the reasons for the lower absolute number of echo chambers around the majority opinion at $t=500$ for the community distributed case versus the random distributed case.
-Beside this hampering of echo chambers around the majority opinion A, the community distributed case also shows an increase of echo chambers around the minority opinion B (which is not visible for the randomly distributed case).
-This may also be a consequence of the hampering in the decrease of the minority opinion for the community distributed case; hence, a significant fraction of minority opinion actors (around $30 \%$) is able to survive and they group together in opinion clusters.
-% share data??
-
+\newpage
 \begin{figure}[H]
@@
   \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
+    \label{rec_comm05_echo}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \caption{}
+  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
+    \label{pr_comm05_echo}
+  \end{subfigure}
   \captionsetup{format=plain}
   \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
-$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity and a low modularity network.
+$10 \%$ and $50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
+The community distributed case is compared to the corresponding randomly distributed case ($55/45$ or $75/25$) for both a high modularity and a low modularity configuration.
 Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
+First row: $10 \%$ single opinion A communities; \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Second row: $50 \%$ single opinion A communities: \textbf{(c)} REC filtering algorithm; \textbf{(d)} PR filtering algorithm.
+The community distributed case is compared to the randomly distributed case for both a high and a low modularity configuration.
+The black dotted line represents a constant amount of echo chambers over time.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-$10 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($55/45$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
-\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
 Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for an improved visualization.}
+Error bars are omitted for improved visualization.}
 \label{echo_01_commOp0_other_50-50}
@@
 \noindent
-The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases, as was observed for the high modularity network.
-Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion A and a corresponding decrease in all the neighborhoods where a fraction of the actors carry the minority opinion B.
-From Fig.
-\ref{low_mod_pr}, it was made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of the minority opinion (when using the PR filtering algorithm).
-This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the actors carry the minority opinion.
-The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case.
-However, as for the high modularity case, if we would look at the absolute number of echo chambers around the majority opinion at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case.
-% make conclusion about this ?
-+ make table with data??
-\\
-\newline
-Fig.
-\ref{echo_05_commOp0_other_50-50} shows a similar plot as Fig.
-\ref{echo_01_commOp0_other_50-50}.
-$50 \%$ of the communities is given opinion A, while the others get a $50/50$ distribution.
-The results look more or less the same as for the case where $10 \%$ of the communities is given opinion A (Fig.
-\ref{echo_01_commOp0_other_50-50}).\\
-\newline
-The REC filtering algorithm, Fig.
-\ref{rec_comm05_echo}, shows a decrease in the number of echo chambers around the majority opinion A with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case.
-A difference with the case where $10 \%$ of the communities are single opinion A communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion B.
-Instead we observe a decrease in the number of echo chambers around the minority opinion.
-For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the actors carrying the majority opinion A and a minority of the actors carrying the minority opinion B.
-\\
-\newline
-The low modularity, community distributed case seems to wiggle around $1$.
-If we look closely, we see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are smaller).
-We observe a small decrease in the formation of echo chambers around the majority opinion A, with a small increase in more mixed opinion neighborhoods as result.
-We also observe a small decrease in the number of echo chambers around the minority opinion B.
-The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of $50 \%$ of the communities being single opinion A communities, while the others get a $50/50$ opinion distribution.\\
-\newline
-The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Fig.
-\ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}.
-They show a small increase in the formation of echo chambers around the minority opinion, indicating the clustering of the actors that carry this opinion.\\
-\newline
-Results for the PR filtering algorithm also strongly resemble those depicted in Fig.
-\ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion A communities.
-The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Fig.
-\ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}).
-They show a strong increase in the formation of echo chambers around the majority opinion A.\\
-\newline
-The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion A, which was also observed in Fig. \ref{pr_comm_echo}.\\
-\newline
-The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion A.
-This was also observed in Fig.
-\ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion A, hampers the increase in this majority opinion compared to the randomly distributed case (Fig.
-\ref{high_mod_pr} and Fig.
+Results for $10 \%$ and $50 \%$ single opinion A communities qualitatively resemble each other.
+Lets first discuss results for the REC filtering algorithm.
+The main observation is that the community distributed scenario for the high modularity configuration shows an increase in more mixed neighborhoods, with a corresponding decrease in single opinion neighborhoods.
+This is different from the random distributed scenario where an increase in single opinion neighborhoods (mainly around the minority opinion) is observed.
+While distributing the opinions randomly increases polarization, the community distributed scenario favors mixed opinion neighborhoods and hence, hampers polarization.
+The community distributed scenario has initially a large number of echo chambers around opinion A, due to the single opinion A communities.
+Even though there is only a minor communication between the communities, this still leads to a situation in which these single opinion A communities become more mixed, which reduces polarization.
+If we zoom in on the low modularity configuration, we see a similar result, albeit less significant.\\
+\newline
+If the opinions are randomly distributed, the PR filtering algorithm enhances the formation of echo chambers around the majority opinion for both the high and low modularity configuration.
+This confirms results found in Section \ref{20-80}.
+The community distributed scenario hampers this increase.
+This is partly a consequence of the larger initial number of echo chambers around the majority opinion, due to the initial single opinion A communities.
+If we would, however, look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed scenario than for the random distributed scenario.
+For the high modularity configuration, the hampering is even stronger than for the low modularity configuration.
+This is a consequence of the fact that for the high modularity configuration the increase in the majority opinion is hampered in the community distributed scenario (Fig.
 \ref{pr_comm}).
-The increase in the number of echo chambers around the minority opinion A that was visible in Fig.
-\ref{pr_comm_echo} is no longer present.
-
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
-    \label{rec_comm05_echo}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \caption{}
-  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
-    \label{pr_comm05_echo}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
-$50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network.
-Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
-$50 \%$ of the communities is given a single opinion, opinion A, while the remaining communities get an initial $50/50$ opinion distribution.
-The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($Q = 0.908 \pm 0.005$ with $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ and $\left<k\right> = 10.79 \pm 0.14$) and a low modularity ($Q = 0.220 \pm 0.007$ with $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ and $\left<k\right> = 10.17 \pm 0.14$), both with $N=10^3$.
-\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Each curve is the average of $10^2$ independent simulations.
-Error bars are omitted for an improved visualization.}
-\label{echo_05_commOp0_other_50-50}
-\end{figure}
-
-\newpage
+Another consequence of this hampering is that for $10 \%$ of single opinion A communities an increase in the formation of echo chambers around the minority opinion B is observed.
+A large fraction of the minority opinion is able to survive and they cluster together in opinion clusters.
+For $50 \%$ of single opinion A communities, this can no longer be observed: too few actors with the minority opinion B are able to survive.
+We, however, observe a hampering in the disappearance of neighborhoods with a fraction of actors that carry the minority opinion.
+
 \section{Real-world networks}
 
-The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Chapter \ref{chap4}, Section \ref{realnetwork}, will be shared and discussed.
-For each real-world network, the results will be compared to results obtained from network models with similar average degree, clustering or modularity.
-We will mainly focus on the formation of echo chambers.
-We will only study balanced initial conditions and we will introduce a fraction of completely stubborn actors.
-For both real-world networks, the REC and PR filtering algorithms will be compared.
+Testing the opinion dynamics model on network models, gives us useful insights in the interplay between network topologies and model features such as filtering algorithm and stubbornness.
+It is, however, interesting to test the model on real-world networks to get a better understanding of real-world opinion dynamics.\\
+\newline
+The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Section \ref{realnetwork}, are shared and discussed.
+For each real-world network, the results are compared to results obtained from network models with similar average degree, clustering or modularity.
+We focus on the formation of echo chambers.
+We study balanced starting conditions and we introduce a fraction of completely stubborn actors.
+For both real-world networks, the REC and PR filtering algorithms are compared.
 
@@
 
-The formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, in the Last.fm on-line social network is studied.
-The results are compared to corresponding results from network models, all with $N = 8003$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) the stochastic block model with 53 communities of size 151, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) the stochastic Watts-Strogatz block model with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
+The formation of echo chambers versus the fraction of stubborn actors, $f_s$, in the Last.fm on-line social network is studied.
+Results are compared to results from network models, all with $N = 8003$: (i) WS model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM with 53 communities of size 151 nodes each, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) SBM-WS with 53 communities of size 151 nodes each, each community has the following WS parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
 The results for the PR and REC filtering algorithms are shown in Fig.
@@
 This is valid for both the real-world network and the network models.
-The stochastic Watts-Strogatz block model has similar clustering and modularity as the Last.fm network.
+The SBM-WS has similar clustering and modularity as the Last.fm network.
 The Last.fm network has, however, a smaller formation of echo chambers than the SBM-WS.
@@
 Other topologies must play a role as well.
-The biggest difference between the real-world Last.fm network and the stochastic Watts-Strogatz block model lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution.
-It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.
-This agrees with the result that Perra and Rocha \cite{Perra2019} found: a heterogeneous contact pattern hampers the formation of echo chambers.\\
-\newline 
-Another difference between the Last.fm and the SBM and SBM-WS network is the community structure.
-The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones.
-The stochastic block model and stochastic Watts-Strogatz block model, that we used, only have communities of a single size.
-This does, however, not play a significant role in the formation of echo chambers, as can be seen in Fig.
-\ref{power_vs_single}.
-In this figure, the distribution of echo chambers in the Last.fm network is compared to those in the SBM-WS network with similar clustering and modularity.
-The SBM-WS is modeled in two ways: (i) the communities all have a fixed size; (ii) the communities have the same heterogeneous sizes as in the Last.fm network.
-The PR and REC filtering algorithms are compared.\\ % add parameters of SBM-WS power??
-\newpage
-\noindent
-The results for the Watts-Strogatz model are similar to those for the stochastic Watts-Strogatz block model.
-The formation of echo chambers seems to be slightly smaller in the WS than in the SBM-WS network.
-It may be tempting to attribute this to the lack of community structure in the WS model.
-It must however be noted that the implemented WS model has a slightly smaller clustering coefficient ($\left<cc\right> = 0.317 \pm 0.004$) than the SBM-WS ($\left<cc\right> = 0.344 \pm 0.002$), which may also explain the somewhat smaller formation of echo chambers.
-Furthermore, it is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity.
+The most significant difference between the real-world Last.fm network and the SBM-WS lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution.
+It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.\\ % add parameters of SBM-WS power??
+\newline
+It is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity.
 It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of high clustering.
-% talk about coincidence??
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, for an initial $50/50$ opinion distribution.
-The Last.fm on-line social network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM-WS with 53 communities of size 151, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$; (iii) SBM with 53 communities of size 151 with $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$.
+  \caption[Formation of echo chambers versus fraction of stubborn actors for balanced starting conditions.
+The PR and REC filtering algorithms are compared.
+The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The Last.fm on-line social network is compared to corresponding network models: (i) WS; (ii) SBM-WS and (iii) SBM.
+The dotted black line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Each data point is the average of $50$ independent simulations.
@@
 If this fraction is further increased, the formation of echo chambers starts to decrease.
-The increase is the strongest for the high clustering network models (WS, SBM-WS).
+The increase is the strongest for the network models with high clustering (WS, SBM-WS).
 Even though the Last.fm network has a similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
@@
 The results for the Last.fm network coincide with those of the SBM.
-This seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to the lack of clustering.
+This seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
+\\
+\newline
+Another difference between the Last.fm and the SBM and SBM-WS network is the community structure.
+The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones.
+The SBM and SBM-WS only have communities of a single, fixed size.
+This does not play a significant role in the formation of echo chambers, as can be seen in Fig.
+\ref{power_vs_single}.
+In this figure, the distribution of echo chambers in the Last.fm network is compared to those in the SBM-WS network with similar clustering and modularity.
+The SBM-WS is modeled in two ways: (i) the communities all have a fixed size; (ii) the communities have the same heterogeneous sizes as in the Last.fm network.
+The PR and REC filtering algorithms are compared.
 
@@
 Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
+\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
 The SBM-WS with heterogeneous community sizes is compared to the SBM-WS with a single community size.
 Results are compared to the real-world Last.fm network with similar clustering and modularity.
-\textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+The dotted black line represents a constant amount of echo chambers over time.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
 Each curve is the average of $50$ independent simulations.
-Error bars are omitted for an improved visualization.}
+Error bars are omitted for improved visualization.}
 \label{power_vs_single}
@@
 Results for the PGP trust network are compared to results for network models with similar average degree, clustering or modularity.
-The used network models are, all with $N = 10680$: (i) the Watts-Strogatz model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) the stochastic block model with 60 communities of size 178, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) the stochastic Watts-Strogatz block model with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
+These models are, all with $N = 10680$: (i) WS model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM with 60 communities of size 178 nodes each, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) SBM-WS with 60 communities of size 178 nodes each, each community has the following WS parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
 \newline
 In Fig.
-\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of completely stubborn actors is studied for the real-world PGP trust network and for the corresponding network models.
+\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of stubborn actors is studied for the real-world PGP trust network and for the network models.
 This is done for both the PR and REC filtering algorithms.
@@
 The PR filtering algorithm (Fig.
-\ref{pr_real_PGP}) shows a decrease in the formation of echo chambers with an increasing fraction of completely stubborn actors for both the PGP trust network and the corresponding network models.
+\ref{pr_real_PGP}) shows a decrease in the formation of echo chambers with an increasing fraction of stubborn actors for both the PGP trust network and the network models.
 The real-world PGP network has a significantly smaller formation of echo chambers compared to the network model with similar clustering and modularity (SBM-WS).
 This is most likely due to the heterogeneous degree pattern of the PGP network.
-\\
-The REC filtering algorithm (Fig.
-\ref{rec_real_PGP}) shows an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease if the fraction of stubborn actors is further increased.
-This is visible for both the PGP trust network and the network models.
-Again, the PGP network has a smaller formation of echo chambers than the network model with similar clustering and modularity (SBM-WS), due to its heterogeneous degree distribution.
-The data points for the SBM and the PGP trust network coincide, which may suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
-% compare powerlaw community distribution??
 
@@
   \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $50/50$ opinion distribution.
-The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of completely stubborn actors, $f_s$, for an initial $50/50$ opinion distribution.
-The PGP trust network is compared to corresponding network models: (i) WS with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM-WS with 60 communities of size 178, each community has the following Watts-Strogatz parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$; (iii) SBM with 60 communities of size 178 with $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$.
+  \caption[Formation of echo chambers versus fraction of stubborn actors for balanced starting conditions.
+The PR and REC filtering algorithms are compared.
+The PGP trust network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
+The PGP trust network is compared to network models: (i) WS; (ii) SBM-WS and (iii) SBM.
+The dotted black line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure).
 Each data point is the average of $50$ independent simulations.
@@
 \end{figure}
+\noindent
+The REC filtering algorithm (Fig.
+\ref{rec_real_PGP}) shows an increase in the formation of echo chambers with an increasing fraction of stubborn actors, followed by a decrease if the fraction of stubborn actors is further increased.
+This is visible for both the PGP trust network and the network models.
+The PGP network has a smaller formation of echo chambers than the network model with similar clustering and modularity (SBM-WS), due to its heterogeneous degree distribution.
+The data points for the SBM and the PGP trust network coincide, which may suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
 
@@
 Simple filtering algorithms were implemented to replicate and simulate the content curation used by social media companies.
-Furthermore, a notion of stubbornness was introduced to investigate the interplay between individual resistance to change, network structure and filtering algorithm.
+Furthermore, stubbornness was introduced to study the interplay between individual resistance to change, network structure and filtering algorithm.
 For simplicity we only allowed two competing opinions (A and B).
-Furthermore, we studied two different settings: balanced starting conditions and unbalanced starting conditions.\\
-\newline
-We investigated the effect of the network structure on the dynamics of the opinions in the system.
-The main focus was put on the difference between networks with a high clustering coefficient and no community structure (WS) compared to networks with a clear community structure and a low clustering coefficient (SBM).
-It was found that a clear community structure enhances the formation of echo chambers (mainly when using the PR filtering algorithm) compared to a random network.
-This enhancement is however not as strong as for networks with a high clustering coefficient.
-Networks that have both a high clustering coefficient and a high modularity (clear community structure) (SBM-WS) behave similar as the high clustering networks (WS).
-This indicates that it is mainly the high clustering that drives the formation of echo chambers.
-There is a correlation between modularity and formation of echo chambers: a higher modularity results in a larger formation of echo chambers (in case of the PR filtering algorithm).
-The enhancement is just not as strong as the enhancement in the formation of echo chambers introduced by a high clustering.
-\\
-% also talk about evolution of opinions??
-\newline
-When introducing a notion of stubbornness (with the help of a stubbornness parameter $r$), we found that the interplay between the PR filtering algorithm and this individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
-The REF and REC filtering algorithms, on the other hand, showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if the stubbornness was further increased.
-This is mainly seen in the high clustering networks, however a small bump can also be observed in the high modularity network.
-This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF and REC filtering algorithms and stubbornness is seen for both balanced starting conditions and unbalanced starting conditions.
-It was also found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
-\newline
-These results give us an insight in the interplay between network structure, filtering algorithms and individual resistance.
+We studied two starting conditions: balanced and unbalanced starting conditions.\\
+\newline
+First, we studied the effect of network structure on the dynamics of the opinions.
+Focus was put on the difference between networks with high clustering and no community structure (WS) compared to networks with community structure and low clustering (SBM).
+It was found that community structure enhances the formation of echo chambers (mainly when using the PR filtering algorithm) compared to random networks.
+There is a correlation between modularity and formation of echo chambers: higher modularity results in larger formation of echo chambers (in case of the PR filtering algorithm).
+The enhancement is, however, not as strong as the enhancement in the formation of echo chambers introduced by high clustering.
+Networks that have both high clustering and high modularity (SBM-WS) behave similar as networks with high clusteing (WS).
+This indicates that it is mainly high clustering that drives the formation of echo chambers.\\
+\newline
+When introducing stubbornness, we found that the interplay between the PR filtering algorithm and individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
+The REF and REC filtering algorithms showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if the stubbornness was further increased.
+This is mainly seen in networks with high clustering, however a small bump can also be observed in networks with high modularity.
+This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF and REC filtering algorithms and stubbornness is seen for both balanced and unbalanced starting conditions.
+It was found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
+\newline
+These results give us insight in the interplay between network structure, filtering algorithms and individual resistance.
 These insights may be crucial to understand phenomena such as polarization in society and to determine the impact of social media on these phenomena.
-They learn us that even though semantic filtering (PR) strongly enhances the formation of echo chambers in the absence of stubborn actors, this enhancing does not stand on its own.
-An interplay between a network with a high clustering or community structure is needed to effectively see this increase in echo chambers induced by semantic filtering.
-A random network, combined with semantic filtering, does not enhance the formation of echo chambers.
-Furthermore, if there are no stubborn actors, recent or random filtering (REC or REF) seems to be more protective against the formation of echo chambers than semantic filtering.
-If there are, however, small fractions of stubborn actors in the system, this scenario changes: semantic filtering starts to hamper the formation of echo chambers, whereas recent or random filtering begins to enhance polarization.
+They learn us that even though semantic filtering (PR) enhances the formation of echo chambers in the absence of stubborn actors, this enhancement does not stand on its own.
+An interplay between networks with high clustering or community structure is needed to effectively see the increase in echo chambers induced by semantic filtering.
+Random networks, combined with semantic filtering, do not enhance the formation of echo chambers.
+If there are no stubborn actors, recent or random filtering (REC or REF) seem to be more protective against the formation of echo chambers than semantic filtering.
+If there is a small fraction of stubborn actors, this scenario changes: semantic filtering starts to hamper the formation of echo chambers, whereas recent or random filtering begins to enhance polarization.
 It is often assumed that algorithmic personalization, tailored to ones preferences (e.g.
-semantic filtering), is one of the main reasons of the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
-Our results may, however, suggest, that if there is some kind of stubbornness (which is probably the case in the real world), heterogeneous exposure (REF or REC) to different opinions may be more harmful than a homogeneous exposure (PR).
+semantic filtering), is one of the reasons for the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
+Our results may, however, suggest, that if there is stubbornness (which is probably the case in real life), heterogeneous exposure (REF or REC) to different opinions may be more harmful than homogeneous exposure (PR).
 Bail et al.
 found experimental results that may agree with this finding \cite{Bail2018}.
-This is an interesting result, that should be investigated more thoroughly.\\
-\newline
-It was found that the initial distribution of the opinions over the communities in a network with community structure (SBM) makes a difference.
+This is an interesting result, that should be studied more thoroughly.\\
+\newline
+It was found that the initial distribution of the opinions over the communities in networks with community structure (SBM) makes a difference.
 We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
-This scenario was then compared to the case in which the total opinion distribution was randomly distributed over the system.
-This was done for both a high modularity and a low modularity network and for both the REC and PR filtering algorithms.
-\\
-\newline
-The REC filtering algorithm did not show any difference in the evolution of the prevalence of the opinions between this (initial) community distributed and (initial) randomly distributed scenario.
-For the PR filtering algorithm it was found that the high modularity network hampers the increase in the majority opinion in the community distributed scenario compared to the randomly distributed scenario.
+This scenario was then compared to the scenario in which the total opinion distribution was randomly distributed.
+This was done for both a high modularity and a low modularity configuration and for both the REC and PR filtering algorithms.
+\\
+\newline
+The REC filtering algorithm did not show a difference in the evolution of the prevalence of the opinions between the (initial) community distributed and (initial) random distributed scenario.
+For the PR filtering algorithm, it was found that the high modularity configuration hampers the increase in the majority opinion in the community distributed scenario compared to the random distributed scenario.
 For the low modularity network, such a discrepancy was not found.\\
 \newline
-Furthermore, it was found that, when using the REC filtering algorithm, the system evolves towards a more mixed opinion distribution in the community distributed scenario.
-This was found for both the high and low modularity network.
-If we make a fraction of the communities single opinion A communities, while giving the remaining communities a $50/50$ opinion distribution, we initially have a large number of echo chambers around opinion A.
-Over time, these opinion A echo chambers decrease; even if there is only a limited amount of communication between the communities, it was found that the single opinion A communities become more and more mixed.
-The total prevalence of opinion A is unchanged when using the REC filtering algorithm, so the initial $50/50$ communities see a small increase in the prevalence of opinion A.
-The randomly distributed scenario does not evolve towards a mixed opinion distribution, instead there is a small increase in the formation of echo chambers around the minority opinion B (for both the high and low modularity network).\\
-\newline
-For the PR filtering algorithm, the randomly distributed scenario resulted in an increase in the formation of echo chambers around the majority opinion A.
-This was observed for both the low and high modularity networks.
-The community distributed scenario for the low modularity network hampers this increase.
+Furthermore, it was found that, when using the REC filtering algorithm, the system evolves towards a configuration with more mixed opinion neighborhoods in the community distributed scenario.
+This reduces polarization and was found for both the high and low modularity configuration.
+The random distributed scenario does not evolve towards mixed opinion neighborhoods, instead a small increase in the formation of echo chambers around the minority opinion B is observed.
+Hence, polarization increases in this scenario.\\
+\newline
+For the PR filtering algorithm, the random distributed scenario resulted in an increase in the formation of echo chambers around the majority opinion A (for both the low and high modularity configuration).
+The community distributed scenario for the low modularity configuration hampers this increase.
 It was found that this is more than just a result of the increased number of initial opinion A echo chambers.
 For the high modularity network, the community distributed scenario showed an even larger hampering of the formation of echo chambers around the majority opinion A.
-This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario in case of the high modularity network combined with the PR filtering algorithm.
-\\
-\newline
-Finally, results for real-world networks were compared to those obtain from corresponding network models.
-It was found that the formation of echo chambers in the real-world networks is smaller than those in a network model with similar clustering and modularity.
+This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario compared to the random distributed scenario.
+\\
+\newline
+Finally, results for real-world networks were compared to those from network models.
+It was found that the formation of echo chambers in the real-world networks is smaller than those in network models with similar clustering and modularity.
 The reason for this hampered formation of echo chambers was attributed to the heterogeneous degree pattern of the real-world networks.
-This hampering of echo chambers due to a heterogeneous degree distribution was also found by Perra and Rocha \cite{Perra2019} when they compared a network model with a heterogeneous degree distribution to models with homogeneous contact patterns.\\
+This hampering of echo chambers due to a heterogeneous degree distribution was also found by Perra and Rocha \cite{Perra2019} when they compared network models with heterogeneous degree distribution to models with homogeneous contact patterns.\\
 \newline
@@
 The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.
-Banisch and Olbrich showed that in a social feedback model polarization is increased in high modularity networks compared to an Erd\H{o}s-R\'{e}nyi network \cite{Banisch2018}.
-This agrees with our findings that a clear community structure enhances the formation of echo chambers.\\
+Banisch and Olbrich showed that in a social feedback model polarization is increased in high modularity networks compared to Erd\H{o}s-R\'{e}nyi networks \cite{Banisch2018}.
+This agrees with our findings that community structure enhances the formation of echo chambers.\\
 \newline
@@
 It must be emphasized that there are limitations to our model and methods.
-First, the simpleness of the model, which on one hand can be seen as a strength, may also be one of the weaknesses.
 Real-life social behavior and opinion dynamics on social media may be far from simple.
-As a consequence, there may be many dynamics of real-life opinion formation that are not well-captured by our model.
-More empirical data is therefore needed to compare the results, obtained from our model, against.
-Other limitations may be the rather small system sizes that were used in our simulations ($N = 10^3$).
-It is unclear how the system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks.
+As a consequence, there may be other dynamics of real-life opinion formation that are not well-captured by our model.
+More empirical data is therefore needed to compare the results, obtained from our model, against \cite{Castellano2009}\cite{Pawel}.
+Other limitations may be the small system sizes that were used in our simulations ($N = 10^3$).
+It is unclear how system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks.
+Limitations due to the activation mechanism, may be the fact that we used a constant activation probability.
+Analyses of real social networks show that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
+It might be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
 Another possible shortcoming is the binary nature of the model.
 There are real-life cases in which people have to choose between two competing opinions.
-These situations, however, do not cover the complete range of possible opinion dynamics situations.
+These situations, however, do not cover the complete range of opinion dynamics situations.
 It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics.
 We also did not differentiate between the actors in our model, except for non-stubborns versus stubborns.
-This combined with the homogeneous degree distribution of our networks models may be another limitation.
+This combined with the homogeneous degree distribution of our network models may be another limitation.
 Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status.
-This may impact the spread of opinions \cite{Fan2020}\cite{Wang2019}.
-We could adapt our model to mimic this social media behavior more closely.
+This may impact the dynamics of opinions \cite{Fan2020}\cite{Wang2019}.
+We could adapt our model to mimic this social media behavior.
 Therefore, a heterogeneous network model should be introduced, combined with an influence parameter.
-This parameter should be used to determine the influence of an actor on others: the opinion of a more influential actor is easier adopted by the others.
+This parameter is used to determine the influence of an actor on others: the opinion of a more influential actor is easier adopted by the others.
 In the model, the high degree nodes should then be made more influential.
-As a final limitation, we mention the fixed stubbornness of the actors.
+The fixed stubbornness of the actors might be another shortcoming.
 Even though the stubbornness parameter $r$ could be individualized to each actor, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
-This idea may be linked to the influence of certain friends: somebody may adopt another viewpoint easier from certain friends than from others.\\
-\newline
-Some future investigation paths, linked to limitations of our model, were given in the previous paragraph.
-Other future directions may be: (i) investigating the distribution of the stubborn actors (community distributed versus random); (ii) making one opinion more stubborn than the other; (iii) investigating the effect of the activation mechanism: what happens if we switch to the heterogeneous activation mechanism (see Section \ref{actMech})?
-and (iv) making the network temporal in the sense that it becomes possible to add and remove friends.
+This idea may be linked to the influence of certain friends: somebody may adopt another viewpoint easier from certain friends than from others.
+As a final limitation, we mention the homogeneous degree distribution of our network models.
+Real-world social networks often have heterogeneous contact patterns.
+It is therefore recommended to test the opinion dynamics model on heterogeneous network models, in order to study the impact of a heterogeneous degree distribution.\\
+\newline
+Future directions may be: (i) investigating the distribution of the stubborn actors (community distributed versus random); (ii) making one opinion more stubborn than the other; (iii) investigating the effect of the activation mechanism: switching to the heterogeneous activation mechanism and (iv) making the network temporal in the sense that it becomes possible to add and remove friends.
 Sasahara et al.
 investigated a model in which social influence was combined with social rewiring (unfriending) \cite{Sasahara2020}.
-Their findings were that these processes of influence and unfriending enhanced the formation of echo chambers \cite{Sasahara2020}.
+Their findings were that processes of influence and unfriending enhanced the formation of echo chambers \cite{Sasahara2020}.
 It might be interesting to incorporate their model features into our model with algorithmic personalization and stubbornness.
 
-ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS (about activation probabilities) + PAY ATTENTION TO FEEDBACK (see file)
-
-''' Perra and Rocha chose activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}.
-They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
-However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}.
-\\
-\newline
- It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics. '''
-
-+ ADD THIS (about opinion dynamics model)
-
-''' This model is a simple one, since it only allows for two different opinions in the system.
-The designed filtering algorithms are simple and intuitive as well.
-Some randomness is introduced by using the probabilistic majority model instead of the classic majority model.
-This might make the model more realistic.
-This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models.
-A simple model is more tractable and allows for a better understanding of what is going on in the dynamics \cite{Jed2019}.
-However, this might be a downside as well.
-Real-world opinion dynamics are, probably, far from simple.
-People do not update their opinion according to a simple probabilistic opinion dynamics model.
-Beside that, the cases where people have to choose between only two opinions are rather limited.
-Filtering algorithms, used by social media companies, are unfortunately not publicly shared.
-We can however assume that they are not as simple as the ones we used in this thesis.
-Trying to model these far more complex systems by our simple model, might lead to the fact that part of the dynamics might be missed or not well captured by our model.
-This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results \cite{Castellano2009}\cite{Pawel}.
-\\
-\newline
-This does not mean that simple models should be thrown away.
-They might still be capable of capturing the dynamics that are seen in the real world \cite{Gug2015}.
-Not because a system appears complex, that it cannot be modeled by something more simple \cite{Jed2019}.
-This has been proven by physics research over and over again \cite{Gug2015}.
-Beside that, if you do not know the dynamics of complex phenomena yet, the best thing you can do is try to model it with the most simple model that comes to mind.
-There is no need to make things too complex if simple works too.
-There is however a need to compare data from (computational) models to empirical, real-world data.
-A need for critical thinking and careful analyzing of results can also not be denied \cite{Pawel}.
-'''
-
-+ ADD THIS (limitations of network models)
-
-'''All the above models have a homogeneous degree distribution.
-We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model.
-It might, however, be an interesting extension to investigate such types of models as well.\\
-\newline
-The above models have their added value and benefits.
-One of their most important plus points is their simplicity.
-They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM).
-This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account.
-This plus side is, however, also one of the possible downsides.
-Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role.
-In other words, network models may lack many features of real-world networks.
-It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
-\newline
-The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.'''
-
-+ ADD THIS (limitations on size) --> needed? or already included?
-
-'''it is, however, unknown how the network size affects the results.
-It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic'''
-
 \chapter{Conclusion}
 
-Simple toy models were implemented to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions.
-Our goal was to get a better understanding of the impact of social media and algorithmic personalization on phenomena such as polarization and formation of echo chambers (opinion clusters) in society.
-\\
-\newline
-We found that topological correlations such as high clustering strongly enhance the formation of opinion clusters.
+Simple toy models were implemented to study the interplay between network structure, filtering algorithms and individual resistance to change competing opinions.
+Our goal was to get a better understanding of the impact of algorithmic personalization on polarization and formation of echo chambers (opinion clusters) in society.
+\\
+\newline
+We found that topological correlations such as high clustering enhance the formation of opinion clusters.
 Beside clustering, community structure also favors the formation of echo chambers.
@@
 The heterogeneous contact pattern of these real-world networks is most likely the cause of this hampering.
-If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system.
+If people are not stubborn, we found that semantic filtering and hence, a homogeneous exposure to information increases polarization in the system compared to random or recent filtering.
 This is an intuitive result: if you shield people from opposing views and strengthen them in their own beliefs, it feels natural that the system will evolve towards a more polarized state.
-It is then expected that a more heterogeneous exposure to information reduces this polarization.
-We found that this is indeed the case if people are not stubborn.
-If, on the contrary, people have some level of stubbornness, we found the surprising result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
-This is an important result, since it may change our assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
-More research on this topic is definitely needed in the near future.
+Heterogeneous exposure to information (random or recent filtering) reduces this polarization if people are not stubborn.
+If, on the contrary, people have some level of stubbornness, we found the result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
+This is important, since it may change our assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
+More research on this topic is needed.
 Finally, we found that distributing the opinions in communities instead of randomly distributing them may hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
 \newline
-Our results strongly suggest that the effects of algorithmic personalization do not stand on their own.
+Our results suggest that the effects of algorithmic personalization do not stand on their own.
 Instead, the interplay with certain network topologies may enhance or hamper these effects.
-Beside that, individual resistance strongly affects the impact of these filtering algorithms.
+Beside that, individual resistance affects the impact of these filtering algorithms.
 It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization.
-It is the interplay between many different aspects of the system (such as topological correlations, distribution of the opinions, individual resistance, content curration) that may potentially enhance or hamper the formation of opinion clusters.
-% also discuss need to investigate effect of changing N, S and p_act
+It is the interplay between topological correlations, distribution of the opinions, individual resistance and content curration that potentially enhance or hamper the formation of opinion clusters.
+
 \newpage

### 2021-05-25 draft.tex
latest update

@@
 This filtering and individualizing of information may have some potentially dangerous downsides in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
-It is important to know the effects of such algorithmic personalization on the formation of phenomena such as polarization, spread of extremism and formation of echo chambers.\\
+It is important to know the effect of algorithmic personalization on the formation of phenomena such as polarization, spread of extremism and formation of echo chambers.\\
 \newline
 People are, however, more than sheep that thoughtlessly follow their shepherd (contrary to what some books might lead us to believe).
-They are capable of critical thinking and may be hesitant to change their current opinion.
+They are capable of critical thinking and may be hesitant to change their opinion.
 This aspect of potential stubbornness should not be ignored.\\
@@
 \newline
-Finally, we will test the model on real-world networks. We will investigate whether the results agree with those obtained from the theoretical network models.
+Finally, we will test the model on real-world networks. We will investigate whether the results agree with results from the theoretical network models.
 
@@
 \newline
-It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion.
+Some people change their opinions more easily than others, that is, once in a while you encounter somebody that is stubborn and persistent towards its own opinion.
 It is important to understand the effects of individual resistance on the formation and evolution of opinions.
@@
 In order to assure a more pleasant and convenient experience on their platforms, social media companies use algorithms to order and filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}.
-It is of importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\
+It is of importance to know the effect of these filtering algorithms on the evolution of opinions in the population.\\
 \newline
@@
 Section \ref{statPhys} introduces concepts such as statistical and social physics.
-Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks.
+Section \ref{complNet} gives an introduction to the first layer, that of complex networks.
 The second layer, the opinion dynamics, is introduced in Section \ref{Opinion}.
@@
 The use of statistical physics may, however, offer a way out.
-In statistical physics different systems behave qualitatively the same and are governed by the same laws near the critical point \cite{Hu2018}\cite{Kadanoff2010}.
+In statistical physics it is found that different systems behave qualitatively the same and are governed by the same laws near the critical point \cite{Hu2018}\cite{Kadanoff2010}.
 This lead to the concept of universality: many different microscopic problems have the same macroscopic manifestation \cite{Gug2015}\cite{Kadanoff2015}.
@@
 This interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in physical research and science in general.
-\newpage
-\noindent
+
 \section{Complex networks}\label{complNet}
@@
 In other words, in all these fields one encounters complex systems, systems where it is not possible to predict collective behavior based on the properties of the individual components alone \cite{Mata2020}.
-Complex systems often display phenomena such as non-linearity, emergence, spontaneous order.
+Complex systems often display phenomena such as non-linearity, emergence and spontaneous order.
 \\
@@
 One can find the language to describe networks in mathematical graph theory.
-However, complex systems in real-life often deal with a large number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
+However, complex systems in real life often deal with a large number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}.
 One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\
@@
 Some other important concepts in network theory are the degree distribution, clustering, community structure and connectivity.
-These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate on the case of social networks.
+These concepts will be explained in depth in Section \ref{netDef}, but let us already anticipate on the case of social networks.
 Many real-life social systems have a power-law degree distribution \cite{Muchnik2013}.
 This heterogeneous or scale-free degree distribution represents one of the three general properties of social networks.
-The other two are short distances and high clustering, also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}.
+The other two are short distances and high clustering, also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
@@
 The theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model.
-These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}.
+These will be explained thoroughly in Section \ref{netModel}.
 
@@
 \newline
-The rules that determine how an actor updates his or her opinion can include many different aspects.
+The rules that determine how an actor updates his or her opinion can include different aspects.
 They can, for example, be as simple as a majority rule (i.e.
 the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a probabilistic way of updating.
-An overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}.
+An overview of possible opinion dynamics models is given in Section \ref{Bin} and Section \ref{Con}.
 \\
@@
 Furthermore, the actors can have a parameter that determines whether they are active or not.
-The introduction of stubborn actors will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of actors is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
+The introduction of stubborn actors will be discussed in Section \ref{stubb}; the activation mechanism of actors is discussed in Section \ref{actMech}.\\
 \newline
@@
 Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
-In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in this thesis will be given.
-
+In Section \ref{filter} and Section \ref{modelThesis} a detailed description of the filtering algorithms used in this thesis will be given.
+\newpage
 \section{This thesis: hypotheses and goals}\label{goal}
 
-This thesis will study the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions.
+This thesis will study the interplay between social network structure and individual resistance to change in the evolution of two competing opinions.
 On one hand, we will determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we will study the effect of stubborn actors in the network.
@@
 They also considered the effect of nudging (meaning that one opinion is pushed to all actors in the network).
-Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering) could result in the formation of echo chambers.
+Their main findings were that algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering) could result in the formation of echo chambers.
 On the other hand, a more heterogeneous contact pattern hampers this formation of echo chambers \cite{Perra2019}.
@@
 This thesis will build on their model and incorporate these new features.
-We will compare networks that have community structure and relatively low clustering to networks without community structure, but with high clustering (a detailed description of concepts like clustering and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}).
+We will compare networks that have community structure and relatively low clustering to networks without community structure, but with high clustering (a detailed description of concepts like clustering and community structure will be given in Section \ref{netDef}).
 We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with high clustering.
@@
 \newline
-We will also study the impact of stubborn actors on the formation and evolution of opinions in the network.
+The impact of stubborn actors on the formation and evolution of opinions in the network will be studied.
 We aim to determine whether there is an interplay between network structure, filtering algorithm and stubbornness.
@@
 \end{equation}
-where a neighbor or nearest neighbor of a node $i$ is an adjacent node of node $i$, the meaning of adjacent nodes is given in Section \ref{connect} \cite{F.Costa2007}.
+where a neighbor of a node $i$ is an adjacent node of node $i$, the meaning of adjacent nodes is given in Section \ref{connect} \cite{F.Costa2007}.
 The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
@@
 \end{equation}
-The derivation can be found in Appendix \ref{ER}, Section \ref{degdis}. From this we can calculate the average degree of the network. The expected mean degree is given by \cite{Hopcroft2006}
+The derivation can be found in Appendix \ref{ER}, Section \ref{degdis}. From this we can calculate the average degree of the network \cite{Hopcroft2006}
 \begin{equation}\label{avDeg}
@@
 The Erd\H{o}s-R\'{e}nyi model is used as a benchmark model.
-We focus on two network topologies that may impact the formation of echo chambers: a high clustering coefficient and community structure.
+We focus on two network topologies that may impact the formation of echo chambers: high clustering and community structure.
 It is therefore necessary to have a model that lacks both of these features to compare the results with.
@@
 For small values of $\beta$, the clustering coefficient remains high.
-The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
+The model of Watts and Strogatz is thus able to generate random networks with high clustering and short average distances.
 The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.\\
@@
 The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected with each other than with the rest of the network \cite{F.Costa2007}.
-Hence, these networks posses a modular structure and this is exactly the structure that is easily modeled by the stochastic block model.
+Hence, these networks posses a modular structure and this is a structure that is easily modeled by the stochastic block model.
 The stochastic block model is also often used as a benchmark for community detection algorithms \cite{Abbe2018}.\\
@@
 \newline
-There are no restrictions on the edge probabilities, which makes the stochastic block model a general model capable of reproducing many different network structures \cite{Karrer2011}.
+There are no restrictions on the edge probabilities, which makes the stochastic block model a general model capable of reproducing different network structures \cite{Karrer2011}.
 However, the edge probabilities inside a community are often larger than the probabilities between communities ($M_{ii} > M_{ij}$, $i \neq j$).
@@
 One extension that is introduced in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
-In this way, we obtain a version of the stochastic block model that has both a community structure and a high clustering coefficient.\\
+In this way, we obtain a version of the stochastic block model that has both community structure and high clustering.\\
 \newline
@@
 For a SBM-WS with $n_c$ communities of specified sizes, we first construct $n_c$ Watts-Strogatz networks (for each constituent Watts-Strogatz network the rewire probability $\beta$ and the mean degree $\left<k\right>$ are specified), then edges are (with some predefined probability) randomly added between the different Watts-Strogatz communities.
-The result is a network that can be tuned to have both a community structure and a high clustering coefficient.
-This is different from the regular stochastic block model that can only account for a community structure and lacks high clustering.\\
+The result is a network that can be tuned to have both community structure and high clustering.
+This is different from the regular stochastic block model that can only account for community structure and lacks high clustering.\\
 \newline
@@
 \newline
-It is expected that community structure will enhance the formation of echo chambers compared to the random network case.
+It is expected that community structure will enhance the formation of echo chambers compared to random networks.
 The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering.
@@
 Other examples of social networks are sexual networks, in which the edges denote sexual contacts between the individuals and email networks, in which directed edges represent an email send from individual $i$ to individual $j$ \cite{Costa2008}.
-There are also plenty of other real-world network such as collaboration networks, where the nodes are usually scientists and edges indicate professional ties between them; citation networks, where nodes represent scientific papers and the directed edges report citations and protein-protein interaction networks, in which nodes stand for proteins and edges indicate physical interactions between them \cite{Costa2008}.\\
+Other real-world networks are collaboration networks, where the nodes are usually scientists and edges indicate professional ties between them; citation networks, where nodes represent scientific papers and the directed edges report citations and protein-protein interaction networks, in which nodes stand for proteins and edges indicate physical interactions between them \cite{Costa2008}.\\
 \newline
@@
 Here, we will list some of the main characteristics of real-world networks, and more specifically of real-world social networks.
-An important question that can be asked is whether there is a significant difference between off-line social networks and on-line social networks.\\
+A relevant question is whether there is a significant difference between off-line social networks and on-line social networks.\\
 \newline
 As discussed briefly in the introduction, real-world networks often posses three characteristic properties.
-These are the two properties of the ``small world phenomenon'', namely a short average distance and a high clustering coefficient, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
+These are the two properties of the ``small world phenomenon'', namely short average distances and high clustering, and a heterogeneous or power-law degree distribution \cite{RealWorld}.
 These properties are valid for both off-line and on-line social networks \cite{Zhang2014}.
@@
 \newline
-\textbf{High clustering coefficient}\\
+\textbf{High clustering}\\
 \newline
@@
 \newline
-Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies.
+Ideally, we would test the opinion dynamics model that is used throughout this thesis on network data directly obtained from social network companies.
 However, most of these companies are not to keen on sharing any form of data.
@@
 \newline
-We used two real-world networks.
+We will use two real-world networks.
 The first real-world network is an on-line friendship network among Finnish users of the social music listening and sharing website Last.fm (data from $2006$).
@@
 \end{equation}
-with $\mu = 1.008$ and $\sigma = 0.95$; see Fig. \ref{deg_distr_lastfm}. Nodes indicate Finnish users of the Last.fm website, edges represent friendship ties between them.\\
-\newline
-The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
+with $\mu = 1.008$ and $\sigma = 0.95$; Fig. \ref{deg_distr_lastfm}. Nodes indicate Finnish users of the Last.fm website, edges represent friendship ties between them.\\
+\newline
+The second real-world network that is used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
 This network has $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
@@
 \end{equation}
-with $a = 0.39$, $\gamma = 1.54$ for $k < 20$ and $a = 5.59$, $\gamma = 2.49$ for $k \geqslant 20$; see Fig.
+with $a = 0.39$, $\gamma = 1.54$ for $k < 20$ and $a = 5.59$, $\gamma = 2.49$ for $k \geqslant 20$; Fig.
 \ref{deg_distr_pgp}.
@@
 \noindent
-One of the decisive properties of the real-world networks that were used, was their size.
-The main, and actually only, reason for this was the computation time.
-If the networks are too big, it would take weeks, sometimes even months before there would be results.
-The size of the network models that were implemented, was around a thousand nodes.
-For the real-world networks, we also used networks with a size of magnitude of thousand to ten-thousand.
+One of the decisive properties of the real-world networks that are used, is their size.
+The main, and actually only, reason for this is the computation time.
+If the networks are too big, it takes weeks, sometimes even months before there would be results.
+The size of the network models that are implemented, is around a thousand nodes.
+For the real-world networks, we also use networks with a size of magnitude of thousand to ten-thousand.
 This benefits the computation time.
@@
 \newline
-The following section will discuss the opinion dynamics model and filtering algorithms adopted in this thesis.
+The next section will discuss the opinion dynamics model and filtering algorithms adopted in this thesis.
 
@@
 
-We want to model opinion dynamics on social media platforms, with focus on Facebook.
+We model opinion dynamics on social media platforms, with focus on Facebook.
 Each time we access Facebook we experience algorithmic personalization.
 The time-line or News Feed of each user on Facebook is the complex combination of posts from, among other, Facebook friends, companies and mass media.
-In 2013, Facebook reported that on average 1500 potential stories of friends and pages we follow or companies that are sponsored could be shown on the News Feed of each user.
-% add ref Facebook
+In 2013, Facebook reported that on average 1500 potential stories of friends and pages we follow or companies that are sponsored could be shown on the News Feed of each user \cite{facebook}.
 No user could ever process this amount of information, hence, Facebook selects around 300 of them, which then appear on your personalized News Feed.
-Their selection methods are based on (i) how often we interact with certain friends, pages or prominent figures; (ii) the amount of likes, shares and comments a post receives from your friends and others; (iii) how much you have interacted with certain types of posts before and (iv) whether you and other people on Facebook are hiding or reporting a certain post.
+Their selection methods are based on (i) how often we interact with certain friends, pages or prominent figures; (ii) the amount of likes, shares and comments a post receives from your friends and others; (iii) how much you have interacted with certain types of posts before and (iv) whether you and other people on Facebook are hiding or reporting a certain post \cite{facebook}.
 \\
@@
 If we want to introduce stubbornness in the model we will mainly use the resistance parameter, as explained in Section \ref{stubbBin}.
-Lets briefly summarize this concept of stubbornness.
+Lets summarize this concept of stubbornness.
 Each actor is given a resistance parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor can update its opinion, if $f \leqslant r_i$, the actor will keep its current opinion.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor updates its opinion, if $f \leqslant r_i$, the actor keeps its current opinion.
 
@@
 This model is explained in Section \ref{stubbBin}.
-We do not use a probabilistic majority model, but a classic majority model, as explainded in Section \ref{majModel}, is implemented.
+We do not use a probabilistic majority model, but a classic majority model, explainded in Section \ref{majModel}, is implemented.
 The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Section \ref{modelThesis}.

### 2021-05-26 draft.tex
latest update

@@
 
-\section{Real-world social networks}
+\section{Real-world social networks}\label{realWorld}
 
@@
 The study of opinion dynamics in a group of people has two layers: a network layer and an opinion dynamics layer.
-The network layer, which describes the underlying structure and correlations between the actors, is described in the previous chapter, Chapter \ref{chap2}.
+The network layer, which describes the underlying structure and correlations between individuals, is described in the previous chapter, Chapter \ref{chap2}.
 This chapter focuses on the second layer, that of the opinion dynamics models.
@@
 This is the mechanism that describes the timings in which information is exchanged between actors \cite{Perra2019}.
-Such mechanism changes the network from a static one to a temporal one, where actors become active and inactive over time.
+Such mechanism changes the network from a static to a temporal one, where actors become active and inactive over time.
 Then, in Section \ref{stubb}, the introduction of stubborn actors will be discussed.
@@
 Each actor has one of two opinions, A or B.
-Initially, these opinions can be equally distributed or not; the starting conditions can be balanced or unbalanced.
+Initially, these opinions can be equally distributed or not, i.e.
+the starting conditions can be balanced or unbalanced.
 Balanced starting conditions indicate the initial situation where $50 \%$ of the actors is given opinion A and $50 \%$ of the actors is given opinion B.
 The opinions are usually randomly assigned; this is however not necessary and opinions can e.g.
-also be assigned according to the community a certain actor belongs to.
+also be assigned according to which community a certain actor belongs to.
 If the starting condition are unbalanced, we have a majority and a minority opinion.
@@
 \newline
-Several studies have been performed on this kind of model with the aim to find the conditions that drive the system to converge to unanimity or to see whether an initial majority opinion remains a majority opinion in the final state \cite{Nguyen2020}.\\
+Several studies have been performed on this model with the aim to find the conditions that drive the system to converge to unanimity or to see whether an initial majority opinion remains a majority opinion in the final state \cite{Nguyen2020}.\\
 \newline
@@
 \newline
-For example, if $70\%$ of the neighbors of actor $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), actor $i$ will update its opinion to opinion A with a probability equal to $0.7$ and to opinion B with a probability equal to $0.3$ \cite{Perra2019}.
+For example, if $80\%$ of the neighbors of actor $i$ have an opinion A (and thus $20 \%$ of the neighbors have opinion B), actor $i$ will update its opinion to opinion A with a probability equal to $0.8$ and to opinion B with a probability equal to $0.2$ \cite{Perra2019}.
 Actor $i$ no longer updates its opinion according to the majority opinion carried by its neighbors, but the majority opinion still has a higher probability to be adopted by this actor.\\
@@
 \newline
-Some concepts that are valid for the discrete models no longer apply (e.g.
+Some concepts that are valid for discrete models no longer apply (e.g.
 the concept of a majority and minority opinion or equality of opinions); hence, they require a different framework \cite{Castellano2009}.
@@
 The stationary state is a situation where several opinion clusters coexist.
-This is because the opinions move closer and closer to each other in several of such opinion clusters; once each opinion cluster is sufficiently far apart from the others (the difference in opinions between all actors in the two opinion clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
+The opinions move closer and closer to each other in several of such opinion clusters; once each opinion cluster is sufficiently far apart from the others (the difference in opinions between all actors in the two opinion clusters is larger than $\epsilon$), they no longer interact with each other \cite{Castellano2009}.
 Only actors inside the same opinion cluster keep on interacting with each other, which makes the opinions inside an opinion cluster converge to a single value \cite{Castellano2009}.
@@
 \newline
-The dynamics of this model is similar to the Deffuant model and the stationary state is also characterized by a number of opinion clusters \cite{Castellano2009}.
+The dynamics is similar to the Deffuant model and the stationary state is also characterized by a number of opinion clusters \cite{Castellano2009}.
 The number of clusters decreases if $\epsilon$ increases and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
@@
 This would be a better representation of social media platforms, where users are active for some periods of time and then become inactive again.
-This kind of activation mechanism of the actors gives the network a temporal character.
+This activation mechanism of the actors gives the network a temporal character.
 The underlying network structure does not change (each actor has a fixed number of neighbors that does not change over time; no new friends are made or, conversely, no old friends are being unfriended).
@@
 This is a bit simplistic and there are shades of gray between this black and white statement.
-However, there are groups in society, with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often the case that there is some correlation between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to any other opinion does not necessarily need to be a bad thing, it can be a form of critical thinking and a form of taking things with a grain of salt).\\
+However, there are groups in society, with more extreme opinions, that are know to be inflexible towards changing that (extreme) opinion (being stubborn and being extreme does not need to go hand in hand, you can be stubborn without being extreme and vice versa; however, it is often the case that there is some correlation between people with extreme opinions and being more persistent towards these opinions compared to people with more moderate beliefs; this idea of correlating extreme people with a stubborn character is often used in models \cite{Deffuant2006}\cite{Deffuant2002}\cite{Turner}) (having some resistance to change your opinion to another opinion does not necessarily need to be a bad thing, it can be a form of critical thinking and taking things with a grain of salt).\\
 \newline
@@
 Social media facilitates the spread of information and ideas.
-This is, unfortunately, also true for extreme and persistent groups.
-They become capable of reaching and influencing larger groups of people.
+This is, unfortunately, also true for persistent groups with extreme ideas: they become capable of reaching and influencing larger groups of people.
 Before the era of social media, one could maybe argue that these extreme and stubborn groups were often isolated and therefore would not reach too many people.
@@
 \newline
-Several models that include some form of stubbornness have been proposed to study the impact of stubborn and extreme groups on society.
+Several models that include stubbornness have been proposed to study the impact of stubborn and extreme groups on society.
 These models aim to study whether the distribution of the stubborns (clustered versus random) matters; whether stubborn groups can lead to polarization or can overtake the opinions in the system or, contrary, whether stubbornness is protective against extremism and polarization.
@@
 
-The bounded confidence models, discussed in Section \ref{bounded}, can be extended to included stubbornness.
+The bounded confidence models, discussed in Section \ref{bounded}, can be extended to include stubbornness.
 Guillaume Deffuant proposed such an extension where stubbornness is correlated to extremism \cite{Deffuant2006}.
@@
 \newline
-Another example of introducing some kind of persistence towards the own opinion of the actors is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
-The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each actor adheres to its initial opinion to a certain degree can also be seen as introducing a form of stubbornness \cite{Noorazar2020}.
+Another example of introducing persistence towards the own opinion of the actors is the ``hardening of positions'' that was briefly discussed at the end of Section \ref{classMod}.
+The model introduced by Friedkin and Johnsen (Section \ref{classMod}) in which each actor adheres to its initial opinion to a certain degree can also be seen as introducing stubbornness \cite{Noorazar2020}.
 
@@
 
-Playing around with the threshold $\epsilon_i$ of the actors to include a notion of stubbornness, such as one can do for the bounded confidence models, is not possible for binary or discrete models.
+Playing around with the threshold $\epsilon_i$ of the actors to include stubbornness, such as one can do for the bounded confidence models, is not possible for binary or discrete models.
 There are, however, other ways to include stubborn actors in these models.
@@
 In bounded confidence models, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions.
-In binary models, however, there are only two possible opinions and in that sense it is less useful to talk about extreme and moderate opinions.
+In binary models there are only two possible opinions and in that sense it is less useful to talk about extreme and moderate opinions.
 However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
@@
 This resistance parameter can have values between 0 and 1.
-If $r_i = 0$, the actor is not stubborn and can be easily convinced by the other opinion (according to the opinion dynamics model that is implemented: majority, probabilistic majority).
+If $r_i = 0$, the actor is not stubborn and can be easily convinced by the other opinion.
 If $r_i = 1$, the actor is completely stubborn and there is no chance of convincing it of the other idea.
@@
 No user could ever process this amount of information, hence, Facebook selects around 300 of them, which then appear on your personalized News Feed.
-Their selection methods are based on (i) how often we interact with certain friends, pages or prominent figures; (ii) the amount of likes, shares and comments a post receives from your friends and others; (iii) how much you have interacted with certain types of posts before and (iv) whether you and other people on Facebook are hiding or reporting a certain post \cite{facebook}.
+Their selection methods are based on (i) how often we interact with certain friends, pages or prominent figures; (ii) the amount of likes, shares and comments a post receives from our friends and others; (iii) how much we have interacted with certain types of posts before and (iv) whether we and other people on Facebook are hiding or reporting a certain post \cite{facebook}.
 \\
@@
 Nodes represent the Facebook users, which become active and inactive over time.
-Edges represent friendship ties on Facebook, that is, they link an user to his Facebook friends and the pages he or she follows.
+Undirected edges represent friendship ties on Facebook, that is, they link an user to his Facebook friends and the pages he or she follows.
 Each node then has an attribute that needs to mimic the Facebook News Feed and which is susceptible to the filtering algorithm.
@@
 \noindent
-Pseudo-code is given in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
-The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed.
+The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
 
@@
 This chapter is subdivided into four main sections.
-We first discuss the impact of the network structure on the evolution of the prevalence of the opinions and the formation of echo chambers.
+We first discuss the impact of network structure on the evolution of the prevalence of the opinions and the formation of echo chambers.
 Then, we introduce stubbornness in the system.
@@
 Four different network models are studied: (i) Erd\H{o}s-R\'{e}nyi model (ER); (ii) Watts-Strogatz model (WS); (iii) stochastic block model (SBM) and (iv) stochastic Watts-Strogatz block model (SBM-WS).
-Each model has $N = 1000$ nodes and average degree $\left<k\right> \sim 10$.
+Each model has $N = 10^3$ nodes and average degree $\left<k\right> \sim 10$.
 The ER model has an edge probability $p = 0.01$.
@@
 We study one configuration with high modularity ($Q = 0.908 \pm 0.005$, with an edge probability inside the communities $p_{\text{cl}} = 0.1$ and an edge probability between communities $p_{\text{add}} = 0.001$) and one configuration with low modularity ($Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$).
-The stochastic Watts-Strogatz block model has 10 communities of size 100 nodes each.
-Each community has Watts-Strogatz parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
+The SBM-WS has 10 communities of size 100 nodes each.
+Each community has WS parameters: mean degree $\left<k\right> =10$ and rewire probability $\beta = 0.01$.
 Edges are added between communities with a probability $p_{\text{add}} = 0.001$ ($\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$).\\
@@
 Two different opinion distributions are investigated, balanced and unbalanced starting conditions.
-For both starting conditions we investigate the evolution of the prevalence of the opinions with time and the formation of echo chambers.
-Two filtering algorithms will be compared, the REC and the PR filtering algorithms, in order to investigate the possible interplay between filtering algorithms and network structure.
+For both starting conditions we investigate the evolution of the prevalence of the opinions and the formation of echo chambers.
+Two filtering algorithms will be compared, the REC and the PR filtering algorithms, in order to investigate the interplay between filtering algorithms and network structure.
 
@@
 \newline
-We start with the results for the evolution of the prevalence of the opinions in case of an initial $50/50$ opinion distribution.
-For both the REC and the PR filtering algorithms, the community structure is not able to break the status quo for both of the filtering algorithms.
+We start with the results for the evolution of the prevalence of the opinions.
+For both the REC and the PR filtering algorithms, the community structure is not able to break the status quo.
 The initial prevalence of the opinions is maintained over time (Fig.
@@
 \textbf{(a)} SBM; \textbf{(b)} SBM-WS.
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 For each model the REC and the PR filtering algorithms are compared.
@@
 Four network models are used: (i) ER; (ii) SBM; (iii) SBM-WS and (iv) WS.
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 For the WS model, we use two different values for the rewire probability: $\beta = 0.01$ ($\left<cc\right> = 0.648 \pm 0.003$; WS2) and $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$; WS1).
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
 The dotted black line represents a constant amount of echo chambers over time.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
 Each curve is the average of $10^2$ independent simulations.
@@
 \newline
-Fig.
-\ref{ev_op_20_80} shows the results for the evolution of the prevalence of opinion A, which is the minority opinion, over time.
-For each network model the REC and PR filtering algorithms are compared.
-\\
+Fig. \ref{ev_op_20_80} shows the results for the evolution of the prevalence of opinion A, which is the minority opinion. For each network model the REC and PR filtering algorithms are compared. \\
 \newline
@@
 This decrease is a consequence of the PR filtering algorithm.
-Since the PR filtering algorithm orders the posts according to the actors' preferences, actors with the majority opinion B will mainly see posts with the same opinion B on their time-line, which limits the chance of changing their opinion to the minority opinion.
-Actors with the minority opinion A, on the other hand, will see posts with opinion A on their time-line when using the PR filtering algorithm.
-However, because there are more actors with opinion B than actors with opinion A, the actors with the minority opinion A will also see a significant number of posts with the majority opinion B on their time-line.
+Since the PR filtering algorithm orders the posts according to the actors' preferences, actors with the majority opinion B will mainly see posts with the same opinion B on their time-line.
+This limits the chance of changing their opinion to the minority opinion.
+Actors with the minority opinion A will see posts with opinion A on their time-line when using the PR filtering algorithm.
+However, because there are more actors with opinion B than actors with opinion A, the actors with the minority opinion A will also see a number of posts with the majority opinion B on their time-line.
 When using the PR filtering algorithm, there is an imbalance in the sense that actors with the minority opinion A have a chance of switching to the majority opinion B, whereas actors with the majority opinion B have (almost) no chance of switching to the minority opinion A.
@@
 \ref{ev_op_20_80} shows that the minority opinion A eventually vanishes.
-This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with a low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with a high modularity.
+This is the case for the low clustering networks, with $P_\text{A}(500) = 0.001 \pm 0.003$ for the ER network, $P_\text{A}(500) = 0.002 \pm 0.003$ for the SBM with low modularity and $P_\text{A}(500) = 0.003 \pm 0.006$ for the SBM with high modularity.
 The high clustering in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_\text{A}(500) = 0.007 \pm 0.014$ for the SBM-WS network (Fig.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$ configuration).
 For each model the REC and the PR filtering algorithms are compared.
@@
 The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
-The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$.
+Results for WS with $\left<k\right> = 6$ and $\beta = 0.01$ are depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$.
 In both networks the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$.
@@
 \newline
-There are no visible effects of having a community structure.
+There are no visible effects of having community structure.
 Fig.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 For each model the REC and the PR filtering algorithms are compared.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
 The dotted black line represents a constant amount of echo chambers over time.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
 Each curve is the average of $10^2$ independent simulations.
@@
 This is probably a consequence of the absence of clustering and the presence of short average path lengths in these models.
-There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of actors with the majority of their neighbors carrying the minority opinion A.
+There is even a little spike around $\left<P_\text{B}^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of actors with the majority of their neighbors carrying the minority opinion A.
 From Fig.
-\ref{ev_op_20_80} it was concluded that the REC filtering algorithm does not break the prevalence of the opinions.
-The opinion distribution is however slightly shifted from the case where the minority opinion A is more or less randomly distributed over the network to the case where the minority opinion A groups together in opinion clusters.
+\ref{ev_op_20_80} it was concluded that the REC filtering algorithm does not break the initial prevalence of the opinions.
+The opinion distribution is however slightly shifted from the case where the minority opinion A is randomly distributed over the network to the case where the minority opinion A starts to group together in opinion clusters.
 \\
@@
 Remember from Fig.
-\ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion.
+\ref{ev_op_20_80}, that the PR filtering algorithm reduces the initial prevalence of the minority opinion.
 Hence, the increase in the formation of echo chambers around the majority opinion B is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution.
@@
 \newline
-Networks with high clustering (WS and SBM-WS) behave completely different.
-For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in Fig.
+Networks with high clustering (WS and SBM-WS) behave completely different than those with absence of clustering.
+For both the REC and PR filtering algorithms the high clustering enhances the formation of echo chambers around the minority opinion A (visible in the spike around $\left<P_\text{B}^{\text{nn}}\right> = 0$ in Fig.
 \ref{ws} and Fig.
@@
 Remember from Fig.
-\ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for the high clustering networks.
-Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining actors with this minority opinion to group together and form opinion clusters.
-The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm.
-This means that the average fraction of actors with all their neighbors carrying the majority opinion B remains the same before and after the time evolution.
-In case of the PR filtering algorithm this value is slightly higher than $1$, indicating a small increase in the number of majority opinion echo chambers after the time evolution.
+\ref{ev_op_20_80} that in case of the PR filtering algorithm the minority opinion A decreases to less than $1 \%$ of the entire population for networks with high clustering.
+Even though the prevalence of the minority opinion decreases, the high clustering forces the remaining actors with this minority opinion to group together and form opinion clusters.
 
@@
 
+We found that community structure enhances the formation of echo chambers compared to random networks when using the PR filtering algorithm.
+We will now study this relation between formation of echo chambers and community structure more deeply.\\
+\newline
 In Fig.
-\ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted.
-The data points represent the value of the echo chamber size in networks with a different modularity.
+\ref{echo_vs_mod}, the size of the echo chambers versus modularity is depicted.
+Data points represent the value of the echo chamber size in networks with different modularity.
 The initial opinion distribution is $50/50$.
-\\
-\newline
-Table \ref{tab1} displays the network models used in Fig. \ref{echo_vs_mod} (all with $N = 10^3$ and $\left<k\right> \sim 10$).
+Table \ref{tab1} displays the network models used in Fig.
+\ref{echo_vs_mod} (all with $N = 10^3$ and $\left<k\right> \sim 10$).
+The SBM's have 10 communities of size 100 nodes each.
 
@@
 \hline
-$0$ & Erd\H{o}s-R\'{e}nyi model; $p=0.01$ \\
-$0.220 \pm 0.007$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$ \\
-$0.476 \pm 0.009$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.05$ and $p_{\text{add}} = 0.005$ \\
-$0.624 \pm 0.008$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.07$ and $p_{\text{add}} = 0.004$ \\
-$0.908 \pm 0.005$ & SBM with 10 communities of size 100; $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$ \\
+$0$ & ER model; $p=0.01$ \\
+$0.220 \pm 0.007$ & SBM; $p_{\text{cl}} = 0.03$, $p_{\text{add}} = 0.008$ \\
+$0.476 \pm 0.009$ & SBM; $p_{\text{cl}} = 0.05$, $p_{\text{add}} = 0.005$ \\
+$0.624 \pm 0.008$ & SBM; $p_{\text{cl}} = 0.07$, $p_{\text{add}} = 0.004$ \\
+$0.908 \pm 0.005$ & SBM; $p_{\text{cl}} = 0.1$, $p_{\text{add}} = 0.001$ \\
 \end{tabular}
@@
 These results suggest that community structure enhances the formation of echo chambers, mostly when using the PR filtering algorithm.
-A high modularity is however required to see the effect of community structure on the formation of opinion clusters.
+High modularity is however required to see the effect of community structure on the formation of opinion clusters.
 \newpage
@@
 	\captionsetup{format=plain}
-	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{The relation between the formation of echo chambers and the community structure, measured by the modularity $Q$, for both the PR and REC filtering algorithms.
-The data points with $Q = 0$ are the result for an ER network; the other eight data points are results for SBM's, each with different modularity (modularity is changed by changing the edge probabilities).
+	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{Relation between the formation of echo chambers and the community structure, measured by the modularity $Q$, for both the PR and REC filtering algorithms.
+Data points with $Q = 0$ are the result for an ER network; the other eight data points are results for SBM's, each with different modularity, which is changed by changing the edge probabilities.
+The dotted black line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B  (normalized means that the value for the fraction of actors with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
@@
 Bars represent the standard error $SE$.
-The data points are shown together with the standard error $SE = \sigma /\sqrt{N}$, because we are mostly interested in the correctness of the value of the average, which is quantified by the standard error.}
+The data points are shown together with the standard error $SE = \sigma /\sqrt{n}$ ($n$ is the number of independent simulations), because we are most interested in the correctness of the average.}
 \label{echo_vs_mod}
@@
 \newline
-We will compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm.
+We will compare the REF, REC and PR filtering algorithms to analyze the interplay between stubbornness and filtering algorithm.
 The ER, WS, SBM and SBM-WS are implemented to study the interplay between network structure and stubbornness.
-
+\newpage
 \subsection{Resistance parameter}
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 The black dotted line represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Each data point is the average of $10^2$ independent simulations.
@@
 The ER model shows no appearance of echo chambers.
-Introducing a fraction of stubborn actors does not change this behavior.
+Introducing a fraction of stubborn actors does not change this.
 The SBM with low modularity behaves similar as the ER model.
@@
 For a small fraction of stubborn actors they show the largest echo chambers.
-The SBM with high modularity has smaller echo chambers than the high clustering networks, but still has some formation of echo chambers for $f_s \leqslant 0.1$.
+The SBM with high modularity has smaller echo chambers than networks with high clustering, but still has some formation of echo chambers for $f_s \leqslant 0.1$.
 These results confirm that community structure enhances the formation of echo chambers compared to random networks, however not as significant as networks with high clustering.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represents a constant amount of echo chambers over time.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
+The dotted black line represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 Instead of a decrease in the formation of echo chambers with increasing fraction of stubborn actors, we first observe an increase, followed by a decrease.
-This is especially visible for the high clustering networks (WS and SBM-WS).
+This is especially visible for networks with high clustering (WS and SBM-WS).
 The SBM with high modularity displays a small bump.
@@
 This increases the chance of having stubborn neighbors, each with different opinions, who cannot change opinions.
-Hence, the formation of echo chambers is disabled.\\
-\newline
-The discrepancy between the results for the PR filtering algorithm and those for the REF and REC filtering algorithms in combination with a small fraction of stubborn actors may be explained with the help of the following example.
+Hence, the formation of echo chambers is hindered.\\
+\newline
+The discrepancy between results for the PR filtering algorithm and those for the REF and REC filtering algorithms in combination with a small fraction of stubborn actors may be explained with the help of the following example.
 Assume we have three actors that form a triangle.
@@
 Lets first discuss the PR filtering algorithm and assume that none of the actors are stubborn.
-This triangle then has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion), which can be seen as follows: the actor with opinion B is likely to switch to opinion A, since all its friends carry this opinion and hence, this opinion will be recommended on its time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter, since there are no neighbors with opinion B).
+The triangle then has a high chance of forming an echo chamber (which is obtained if all the actors adopt the same opinion), which can be seen as follows: the actor with opinion B is likely to switch to opinion A, since all its friends carry this opinion and hence, this opinion will be recommended on its time-line (the fact that the PR filtering algorithm wants to promote posts with opinion B does not matter, since there are no neighbors with opinion B).
 The two actors with opinion A will have more chance of keeping that opinion than to switch to opinion B.
@@
 This explains the increase in the formation of echo chambers for a small fraction of stubborn actors when using the REF or REC filtering algorithms.
-If too many actors become stubborn, this increased ability of the system to form echo chambers starts to decrease which can also be seen with the help of the triangle example.
-If the fraction of stubborn actors becomes too high, we might end up with both the actor with opinion B and one of the actors with opinion A that are stubborn.
+If too many actors become stubborn, this increased ability of the system to form echo chambers starts to decrease.
+This is because, we might end up with both the actor with opinion B and one of the actors with opinion A that are stubborn.
 This reduces the ability to form echo chambers since we can no longer have a final situation in which all the three actors in the triangle carry the same opinion.\\
@@
 The answer is no.
-Introducing stubborn actors in the system does not affect the prevalence of the opinions over time for neither the REC, nor the PR filtering algorithm.
+Introducing stubborn actors in the system does not affect the initial prevalence of the opinions over time for neither the REC, nor the PR filtering algorithm.
 The balanced opinion distribution is maintained over time.
@@
 In Fig.
-\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers versus the fraction of stubborn actors, $f_s$ is shown for the PR and the REC filtering algorithms.
+\ref{echo_vs_frac_complStubb_PR-REC_20-80}, the formation of echo chambers versus the fraction of stubborn actors, $f_s$, is shown for the PR and the REC filtering algorithms.
 The interpretation of the $y$-axis is different than for configurations with balanced starting conditions.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 The black dotted line represents a constant amount of echo chambers over time.
-The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
 Each data point is the average of $10^2$ independent simulations.
@@
 \ref{echo_vs_frac_complStubb}).
-The formation of echo chambers in the ER network and the SBM with low modularity is non-zero (for $f_s=0$).
+The formation of echo chambers in the ER network and the SBM with low modularity is non-zero.
 This agrees with results found in Fig.
@@
 \ref{REC_frac_compl_stubb_20-80}) are different than those for the PR filtering algorithm.
-The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease.
+Networks with high clustering (WS and SBM-WS) show an increase in the formation of echo chambers with increasing fraction of stubborn actors, followed by a decrease.
 This result is similar to results for balanced starting conditions (Fig.
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
-The PR filtering algorithm is used and three fractions of stubborn actors ($f_s = 0$, $f_s = 0.4$ and $f_s = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions with stubborn actors.
+The PR filtering algorithm is used and three fractions of stubborn actors ($f_s = 0$, $f_s = 0.4$ and $f_s = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions with stubborn actors.
 \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represents the initial prevalence of opinion A.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The dotted black line represents the initial prevalence of opinion A.
 The PR filtering algorithm is used and three fractions of stubborn actors ($f_s = 0$, $f_s = 0.4$ and $f_s = 0.8$) are compared.
@@
 Fig.
-\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the resistance parameter of the actors, $r$ when using the PR filtering algorithm.
-Results resemble those of the configuration in which a fraction of the actors is completely stubborn (Section \ref{complStubbAct}, Fig.
+\ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the resistance parameter of the actors, $r$, when using the PR filtering algorithm.
+Results resemble those for the configuration in which a fraction of the actors is completely stubborn (Section \ref{complStubbAct}, Fig.
 \ref{echo_vs_frac_complStubb}).\\
@@
 If $r$ increases, the formation of echo chambers decreases.
-This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
+This is especially visible for networks with high clustering (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$.
 If $r$ increases and the actors become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represent a constant amount of echo chambers over time.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
+The dotted black line represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Each data point is the average of $10^2$ independent simulations.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represent a constant amount of echo chambers over time.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
+The dotted black line represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 If $r=1$, the actors are completely stubborn and hence, retain their initial opinion: the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one.
-For high levels of the resistance of the actors, up to $r = 0.9$, the formation of echo chambers is large and even larger than if the actors are not stubborn.
+For high resistance levels of the actors, up to $r = 0.9$, the formation of echo chambers is larger than if the actors are not stubborn.
 If we zoom in on the SBM with high modularity, we see an increase in the formation of echo chambers.
-The increase is however not as strong as for the networks with high clustering.\\
+The increase is however not as strong as for networks with high clustering.\\
 \newline
@@
 This suggests that it is the interplay between high clustering and, to a lesser extent, community structure (high modularity), and the REF or REC filtering algorithms that drives the increase in the formation of echo chambers with increasing $r$.
-This is the opposite of the situation where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with increasing $r$).
+This is the opposite from the situation where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with increasing $r$).
 The reason for this discrepancy is probably of a similar nature as for the discrepancy in case of introducing a fraction of completely stubborn actors, Section \ref{complStubbAct}.\\
@@
   \captionsetup{format=plain}
-  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
-The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions.
+  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions when the actors have a resistance parameter $r$.
+The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) for unbalanced starting conditions when the actors have a resistance parameter $r$.
 \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS.
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represent the initial prevalence of opinion A.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The dotted black line represents the initial prevalence of opinion A.
 The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
@@
 
-The results for the majority threshold model are shared and discussed.
+Results for the majority threshold model are shared and discussed.
 This model is explained in Section \ref{stubbBin}.
@@
 \ref{echo_vs_threshold_PR} shows the formation of echo chambers versus the threshold parameter $T$ when using the PR filtering algorithm.
-The high clustering models (WS and SBM-WS) show a significant formation of echo chambers, which remains more or less flat for $T \leqslant 0.5$.
-This is expected, because if $T \leqslant 0.5$, the threshold has no effect: the majority opinion of the neighbors of an actor will be larger than the threshold (the sum of the two opinion fractions equals one, so the fraction of neighbors with the majority opinion will take values between $0.5$ and $1$).
+Networks with high clustering (WS and SBM-WS) show a significant formation of echo chambers, which remains more or less flat for $T \leqslant 0.5$.
+This is expected, because if $T \leqslant 0.5$, the threshold has no effect: the majority opinion of the neighbors of an actor will be larger than the threshold (the fraction of neighbors with the majority opinion always take values between $0.5$ and $1$).
 If $T \geqslant 0.5$, it starts to have an effect.
@@
 The ER network and the SBM with low modularity (data points lie behind those of the ER network) do not enhance the formation of echo chambers.
-Including a threshold $T$ does not change this tendency.
+The threshold $T$ does not change this.
 
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represents a constant amount of echo chambers over time.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
+The dotted black line represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Each data point is the average of $10^2$ independent simulations.
@@
 In the average of different independent simulations, the opinion distribution will remain $50/50$.
-However, each independent simulation the system converges to one of the two opinions, with the result that the number of echo chambers around that opinion strongly increases.
-After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (resulting in large echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in no echo chambers around opinion B).
+However, each independent simulation the system converges to one of the two opinions, with the result that the number of echo chambers around that opinion increases significantly.
+After the $10^2$ independent simulations, the spread in the formation of echo chambers will be large, since there will have been simulations that converge to opinion B (resulting in large numbers of echo chambers around opinion B) and there will have been simulations that converge to opinion A (resulting in no echo chambers around opinion B).
 
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represents the initial prevalence of opinion A.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+The dotted black line represents the initial prevalence of opinion A.
 For each model the REC and the PR filtering algorithms are compared.
@@
 Furthermore, networks with high clustering (WS and SBM-WS) hinder the convergence to one opinion, even for the REC filtering algorithm.
-This indicates that high clustering is a protective factor against the convergence to one opinion each independent simulation.
 Having community structure (high modularity) in case of the SBM is another hindering of this convergence, as can be seen in Fig.
@@
 It might be more realistic if people with the same opinion are grouped together initially.
-We did introduce a network model with a community structure (SBM) to represent group structures found in real-life.
-But we have not yet distributed the opinions according to these groups.
-\\
-\newline
-We study the SBM and give a fraction of the communities a single opinion, opinion A.
+We did introduce a network model with community structure (SBM) to represent group structures found in real-life.
+But we have not yet distributed the opinions according to these communities.
+\\
+\newline
+We study the SBM and give a fraction of the communities a single opinion, opinion A ($f_\text{A}$).
 The remaining communities are given a $50/50$ opinion distribution.
 This situation may represent real-life with groups of people following the same idea together with groups where the two ideas coexist.
-The total opinion distribution over the entire network is no longer $50/50$; opinion A becomes the majority opinion.
+The total opinion distribution is not balanced, but opinion A becomes the majority opinion.
 The higher the fraction of single opinion A communities, the bigger the difference between the majority and minority opinion.
@@
 In section \ref{groupVSrandomOP} the results for the evolution of the prevalence of the opinions are discussed. Section \ref{groupVSrandomECHO} discusses the formation of echo chambers.
-
+\newpage
 \subsection{Evolution of the prevalence of the opinions}\label{groupVSrandomOP}
@@
 First row: high modularity configuration \textbf{(a)} REC; \textbf{(b)} PR.
-second row: low modularity configuration \textbf{(c)} REC; \textbf{(d)} PR.
-$10 \%$ of the communities is given a single opinion, opinion A.
+Second row: low modularity configuration \textbf{(c)} REC; \textbf{(d)} PR.
+$f_\text{A} = 0.1$.
 Results are compared to the case where the total initial opinion distribution ($55/45$) is randomly distributed over each community.
@@
 \end{figure}
+\newpage
 \noindent
@@
 \ref{commOp0_01_other_50-50} shows the results for the prevalence of opinion A versus time.
-$10 \%$ of the communities is given a single opinion, opinion A.
+$10 \%$ of the communities is given a single opinion, opinion A ($f_\text{A} = 0.1$).
 Results are compared to the case in which the total opinion distribution ($55/45$) is randomly distributed.
@@
 The consequence is that the mixed opinion communities live under the assumption that the opinion distribution is $50/50$.
-They do not feel the majority of opinion A, since they are isolated from these single opinion A communities.
+They do not feel the majority of opinion A.
 For the low modularity configuration, Fig.
@@
 This is visible in the results: the hampering in the increase of the majority opinion is no longer present.
-The community distributed case behaves equivalent to the random distributed case.
+The community distributed case behaves equivalent to the randomly distributed case.
 The community structure is not clearly present; there is a lot of communication between the communities.
-Hence, the mixed opinion communities feel the presence of the single opinion communities and feel the majority of opinion A in a similar way as if the opinions would be randomly distributed over the communities.\\
+Hence, the mixed opinion communities feel the presence of the single opinion communities and feel the majority of opinion A in a similar way as if the opinions are randomly distributed over the communities.\\
 \newline
@@
 This can be seen in Fig.
-\ref{fracOp0_vs_commOp0_other_50-50}, which depicts the prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities.
-The community distributed cases are compared to the randomly distributed cases.
+\ref{fracOp0_vs_commOp0_other_50-50}, which depicts the prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities, $f_\text{A}$.
+Community distributed cases are compared to randomly distributed cases.
 Results are shown for both a high and a low modularity configuration.\\
@@
 For the REC filtering algorithm (Fig.
-\ref{rec_comm}) the prevalence of opinion A remains the same before and after the time evolution for a fixed fraction of single opinion A communities.
+\ref{rec_comm}) the prevalence of opinion A remains the same before and after the time evolution for fixed $f_\text{A}$.
 Giving a fraction of the communities a single opinion, opinion A, does not make a difference compared to randomly distributing the same total initial opinion distribution.
@@
 The PR filtering algorithm (Fig.
-\ref{pr_comm}) shows an increase in the prevalence of opinion A over time for a fixed fraction of single opinion A communities.
+\ref{pr_comm}) shows an increase in the prevalence of opinion A over time for fixed $f_\text{A}$.
 For the high modularity configuration, making a fraction of the communities single opinion A communities hampers this increase compared to the corresponding randomly distributed case.
@@
 The reason for the hampering in case of a high modularity and the absence of it for a low modularity configuration was explained for Fig.
-\ref{commOp0_01_other_50-50} and has to do with whether the single opinion A communities can influence the $50/50$ communities.
+\ref{commOp0_01_other_50-50} and has to do with whether the single opinion A communities can influence the mixed communities.
 
@@
   \caption[Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
-Community distributed cases are compared to the corresponding randomly distributed cases for both a high modularity and a low modularity configuration.
-Results for the REC and PR filtering algorithms.]{Average prevalence of opinion A before and after the time evolution versus the fraction of single opinion A communities.
+Community distributed cases are compared to randomly distributed cases for both a high and low modularity configuration.
+Results for the REC and PR filtering algorithms.]{Average prevalence of opinion A, $\left<P_\text{A}\right>$, before and after the time evolution versus the fraction of single opinion A communities, $f_\text{A}$.
 \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Community distributed cases are compared to the corresponding randomly distributed cases for both a high and a low modularity configuration.
+Community distributed cases are compared to randomly distributed cases for both a high and low modularity configuration.
 The solid black line represents the initial prevalence of opinion A.
@@
 \ref{echo_01_commOp0_other_50-50} shows the distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$.
-In the first row, $10 \%$ of the communities is given opinion A; in de second row, $50 \%$ of the communities are single opinion A communities .
-The community distributed case is compared to the random distributed case: $55/45$ in the 1st row; $75/25$ in the 2nd row.
+In the first row, $10 \%$ of the communities is given opinion A ($f_\text{A} = 0.1$); in the second row, $50 \%$ of the communities are single opinion A communities ($f_\text{A} = 0.5$).
+Community distributed case are compared to the randomly distributed cases: $55/45$ in the 1st row; $75/25$ in the 2nd row.
 Both the high and low modularity configuration are studied.
@@
 \newpage
+
 \begin{figure}[H]
@@
 Results for the REC and PR filtering algorithms.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$.
-First row: $10 \%$ single opinion A communities; \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-Second row: $50 \%$ single opinion A communities: \textbf{(c)} REC filtering algorithm; \textbf{(d)} PR filtering algorithm.
+First row: $f_\text{A} = 0.1$: \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
+Second row: $f_\text{A} = 0.5$: \textbf{(c)} REC filtering algorithm; \textbf{(d)} PR filtering algorithm.
 The community distributed case is compared to the randomly distributed case for both a high and a low modularity configuration.
@@
 \noindent
-Results for $10 \%$ and $50 \%$ single opinion A communities qualitatively resemble each other.
+Results for $f_\text{A} = 0.1$ and $f_\text{A} = 0.5$ qualitatively resemble each other.
 Lets first discuss results for the REC filtering algorithm.
 The main observation is that the community distributed scenario for the high modularity configuration shows an increase in more mixed neighborhoods, with a corresponding decrease in single opinion neighborhoods.
-This is different from the random distributed scenario where an increase in single opinion neighborhoods (mainly around the minority opinion) is observed.
+This is different from the randomly distributed scenario where an increase in single opinion neighborhoods (mainly around the minority opinion) is observed.
 While distributing the opinions randomly increases polarization, the community distributed scenario favors mixed opinion neighborhoods and hence, hampers polarization.
@@
 This is partly a consequence of the larger initial number of echo chambers around the majority opinion, due to the initial single opinion A communities.
-If we would, however, look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed scenario than for the random distributed scenario.
+If we would, however, look at the absolute number of echo chambers around opinion A at $t=500$, we would see that this number is lower for the community distributed scenario than for the randomly distributed scenario.
 For the high modularity configuration, the hampering is even stronger than for the low modularity configuration.
@@
 \ref{pr_comm}).
-Another consequence of this hampering is that for $10 \%$ of single opinion A communities an increase in the formation of echo chambers around the minority opinion B is observed.
+Another consequence of this hampering is that for $f_\text{A} = 0.1$, an increase in the formation of echo chambers around the minority opinion B is observed.
 A large fraction of the minority opinion is able to survive and they cluster together in opinion clusters.
-For $50 \%$ of single opinion A communities, this can no longer be observed: too few actors with the minority opinion B are able to survive.
+For $f_\text{A} = 0.5$, this can no longer be observed: too few actors with the minority opinion B are able to survive.
 We, however, observe a hampering in the disappearance of neighborhoods with a fraction of actors that carry the minority opinion.
@@
 \newline
-The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Section \ref{realnetwork}, are shared and discussed.
-For each real-world network, the results are compared to results obtained from network models with similar average degree, clustering or modularity.
+The results of the two real-world networks: (i) the Last.fm on-line social network; and (ii) the PGP trust network, discussed in Section \ref{realWorld}, are shared and discussed.
+For each real-world network, the results are compared to results from network models with similar average degree, clustering or modularity.
 We focus on the formation of echo chambers.
@@
 The PR and REC filtering algorithms are compared.
-The Last.fm on-line social network is compared to corresponding network models.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
+The Last.fm on-line social network is compared to network models.]{Formation of echo chambers versus the fraction of stubborn actors, $f_s$, for balanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
-The Last.fm on-line social network is compared to corresponding network models: (i) WS; (ii) SBM-WS and (iii) SBM.
+The Last.fm on-line social network is compared to network models: (i) WS; (ii) SBM-WS and (iii) SBM.
 The dotted black line represents a constant amount of echo chambers over time.
@@
 \textbf{(a)} REC filtering algorithm; \textbf{(b)} PR filtering algorithm.
-The SBM-WS with heterogeneous community sizes is compared to the SBM-WS with a single community size.
+The SBM-WS with heterogeneous community sizes (SBM-WS1) is compared to the SBM-WS with a single community size (SBM-WS2).
 Results are compared to the real-world Last.fm network with similar clustering and modularity.

### 2021-05-27 draft.tex
latest update

@@
 We want to study the impact of these topologies on the dynamics of opinions.
-In this section, we will discuss the results related to different network topologies with focus on the effect of clustering and community structure.
+In this section, we will discuss results related to different network topologies with focus on the effect of clustering and community structure.
 Two different opinion distributions are investigated, balanced and unbalanced starting conditions.
@@
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The black dotted line represents a constant amount of echo chambers over time.
+The dotted black line represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
@@
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
-The black dotted line represents a constant amount of echo chambers over time.
+The dotted black line represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The black dotted line represent a constant amount of echo chambers over time.
+The dotted black line represents a constant amount of echo chambers over time.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
@@
 The community distributed case is compared to the randomly distributed case for both a high and a low modularity configuration.
-The black dotted line represents a constant amount of echo chambers over time.
+The dotted black line represents a constant amount of echo chambers over time.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
@@
 The formation of echo chambers versus the fraction of stubborn actors, $f_s$, in the Last.fm on-line social network is studied.
-Results are compared to results from network models, all with $N = 8003$: (i) WS model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM with 53 communities of size 151 nodes each, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) SBM-WS with 53 communities of size 151 nodes each, each community has the following WS parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
+Results are compared to results from network models, all with a size of $N = 8003$ nodes: (i) WS model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM with 53 communities of size 151 nodes each, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) SBM-WS with 53 communities of size 151 nodes each, each community has WS parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
 The results for the PR and REC filtering algorithms are shown in Fig.
@@
 Other topologies must play a role as well.
-The most significant difference between the real-world Last.fm network and the SBM-WS lies in the degree distribution: the Last.fm network has a heterogeneous degree distribution instead of a homogeneous degree distribution.
-It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to a network model with similar clustering and modularity, but with a homogeneous degree distribution.\\ % add parameters of SBM-WS power??
-\newline
-It is interesting to notice that the Last.fm on-line social network coincides with the SBM with a similar modularity.
-It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of high clustering.
+The most significant difference between the real-world Last.fm network and the SBM-WS lies in the degree distribution: the Last.fm network has a heterogeneous instead of a homogeneous degree distribution.
+It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to network models with similar clustering and modularity, but with a homogeneous degree distribution.\\ % add parameters of SBM-WS power??
+\newline
+It is interesting to notice that the Last.fm on-line social network coincides with the SBM with similar modularity.
+It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of clustering.
 
@@
 The Last.fm on-line social network is compared to network models: (i) WS; (ii) SBM-WS and (iii) SBM.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 The dotted black line represents a constant amount of echo chambers over time.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Each data point is the average of $50$ independent simulations.
@@
 If this fraction is further increased, the formation of echo chambers starts to decrease.
-The increase is the strongest for the network models with high clustering (WS, SBM-WS).
-Even though the Last.fm network has a similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
+The increase is the strongest for network models with high clustering (WS, SBM-WS).
+Even though the Last.fm network has similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
 As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network.
@@
 Results are compared to the real-world Last.fm network with similar clustering and modularity.
+The distribution is normalized by dividing for the same quantity computed at $t=0$.
 The dotted black line represents a constant amount of echo chambers over time.
-The distribution is normalized by dividing for the same quantity computed at $t=0$.
 Each curve is the average of $50$ independent simulations.
@@
 Results for the PGP trust network are compared to results for network models with similar average degree, clustering or modularity.
-These models are, all with $N = 10680$: (i) WS model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM with 60 communities of size 178 nodes each, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) SBM-WS with 60 communities of size 178 nodes each, each community has the following WS parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
+These models are, all with a size of $N = 10680$ nodes: (i) WS model with $\left<k\right> = 4$, $\beta = 0.2$ and $\left<cc\right> = 0.266 \pm 0.003$; (ii) SBM with 60 communities of size 178 nodes each, $p_\text{cl} = 0.02$, $p_\text{add} = 0.00005$, $\left<k\right> = 4.06 \pm 0.02$ and $Q = 0.868 \pm 0.002$; (iii) SBM-WS with 60 communities of size 178 nodes each, each community has WS parameters: $\left<k\right> = 4$, $\beta = 0.12$; edges are added between communities with probability $p_\text{add} = 0.00007$, $\left<k\right> = 4.74 \pm 0.01$, $\left<cc\right> = 0.263 \pm 0.003$ and $Q = 0.842 \pm 0.002$.\\
 \newline
 In Fig.
-\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of stubborn actors is studied for the real-world PGP trust network and for the network models.
+\ref{echo_vs_fracRes_real_PGP} the formation of echo chambers versus the fraction of stubborn actors, $f_s$, is studied for the real-world PGP trust network and for the network models.
 This is done for both the PR and REC filtering algorithms.
@@
 The PGP trust network is compared to network models: (i) WS; (ii) SBM-WS and (iii) SBM.
+The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 The dotted black line represents a constant amount of echo chambers over time.
-The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
 Each data point is the average of $50$ independent simulations.
@@
 These toy models consist of several layers, including a network layer and an opinion dynamics layer.
-Simple filtering algorithms were implemented to replicate and simulate the content curation used by social media companies.
+Filtering algorithms were implemented to replicate and simulate the content curation used by social media companies.
 Furthermore, stubbornness was introduced to study the interplay between individual resistance to change, network structure and filtering algorithm.
@@
 When introducing stubbornness, we found that the interplay between the PR filtering algorithm and individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
-The REF and REC filtering algorithms showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if the stubbornness was further increased.
+The REF and REC filtering algorithms showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if stubbornness was further increased.
 This is mainly seen in networks with high clustering, however a small bump can also be observed in networks with high modularity.
 This interplay between the PR filtering algorithm and individual resistance versus the interplay between the REF and REC filtering algorithms and stubbornness is seen for both balanced and unbalanced starting conditions.
-It was found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
+Furthermore, it was found that including stubborn actors hampers the increase in the majority opinion (and thus prevents the formation of a final consensus state) in case of unbalanced starting conditions combined with the PR filtering algorithm.\\
 \newline
@@
 found experimental results that may agree with this finding \cite{Bail2018}.
-This is an interesting result, that should be studied more thoroughly.\\
+This is a result that should be studied more thoroughly.\\
 \newline
@@
 This scenario was then compared to the scenario in which the total opinion distribution was randomly distributed.
-This was done for both a high modularity and a low modularity configuration and for both the REC and PR filtering algorithms.
-\\
-\newline
-The REC filtering algorithm did not show a difference in the evolution of the prevalence of the opinions between the (initial) community distributed and (initial) random distributed scenario.
+This was done for both a high and low modularity configuration and for both the REC and PR filtering algorithms.
+\\
+\newline
+The REC filtering algorithm did not show a difference in the evolution of the prevalence of the opinions between the (initial) community distributed and (initial) randomly distributed scenario.
 For the PR filtering algorithm, it was found that the high modularity configuration hampers the increase in the majority opinion in the community distributed scenario compared to the random distributed scenario.
-For the low modularity network, such a discrepancy was not found.\\
-\newline
+For the low modularity network, such a discrepancy was not found.
 Furthermore, it was found that, when using the REC filtering algorithm, the system evolves towards a configuration with more mixed opinion neighborhoods in the community distributed scenario.
 This reduces polarization and was found for both the high and low modularity configuration.
-The random distributed scenario does not evolve towards mixed opinion neighborhoods, instead a small increase in the formation of echo chambers around the minority opinion B is observed.
+The randomly distributed scenario does not evolve towards mixed opinion neighborhoods, instead a small increase in the formation of echo chambers around the minority opinion B is observed.
 Hence, polarization increases in this scenario.\\
@@
 For the high modularity network, the community distributed scenario showed an even larger hampering of the formation of echo chambers around the majority opinion A.
-This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario compared to the random distributed scenario.
+This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario compared to the randomly distributed scenario.
 \\
@@
 \newline
-The formation of echo chambers in social networks and social media, and the enhancement of them when using algorithmic personalization such as semantic filtering, agrees with results found in literature (in the absence of stubborn actors) \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
+The formation of echo chambers in social networks and social media, and the enhancement of them when using algorithmic personalization such as semantic filtering (in the absence of stubborn actors), agrees with results found in literature \cite{Baumann2019}\cite{Cinelli2020}\cite{Nikolov2015}\cite{Sirbu2018}.
 The enhancement of the formation of echo chambers in clustered communities is also found in literature on complex contagious processes \cite{Becker2017}.
@@
 It is unclear how system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks.
-Limitations due to the activation mechanism, may be the fact that we used a constant activation probability.
+Limitations due to the activation mechanism, may be that we used a constant activation probability.
 Analyses of real social networks show that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
-It might be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.
+It might be an interesting extension to switch to the heterogeneous activation probabilities and investigate possible impacts on the dynamics.
 Another possible shortcoming is the binary nature of the model.
@@
 These situations, however, do not cover the complete range of opinion dynamics situations.
-It would be interesting to run the model on situations where more than two opinions are allowed or to switch to continuous opinion dynamics.
+It would be interesting to run the model on configurations where more than two opinions are allowed or to switch to continuous opinion dynamics.
 We also did not differentiate between the actors in our model, except for non-stubborns versus stubborns.
@@
 Real-world social networks often have heterogeneous contact patterns.
-It is therefore recommended to test the opinion dynamics model on heterogeneous network models, in order to study the impact of a heterogeneous degree distribution.\\
+It is therefore recommended to test the opinion dynamics model on heterogeneous network models.\\
 \newline
@@
 Heterogeneous exposure to information (random or recent filtering) reduces this polarization if people are not stubborn.
-If, on the contrary, people have some level of stubbornness, we found the result that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
+If, on the contrary, people have some level of stubbornness, we found that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
 This is important, since it may change our assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
@@
 It is the interplay between topological correlations, distribution of the opinions, individual resistance and content curration that potentially enhance or hamper the formation of opinion clusters.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3cm]{drawing.png}
+    \label{pr_real_PGP}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3cm]{drawing1.png}
+    \label{pr_real_PGP}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3cm]{drawing2.png}
+    \label{pr_real_PGP}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3cm]{drawing3.png}
+    \label{pr_real_PGP}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Formation of echo chambers versus fraction of stubborn actors for balanced starting conditions.
+The PR and REC filtering algorithms are compared.
+The PGP trust network is compared to corresponding network models.]{test.}
+\label{echo_vs_fracRes_real_PGP}
+\end{figure}
 

### 2021-05-28 draft.tex
latest update

@@
 The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
-We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).
+We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
+\newline
+Fig. \ref{netRepr} visualizes the three main network models we used: (i) ER; (ii) WS and (iii) SBM.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.32\textwidth}
+    \caption{}
+    \includegraphics[width=5cm, height=5cm]{ER.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\textwidth}
+    \caption{}
+    \includegraphics[width=5cm, height=5cm]{WS.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\textwidth}
+    \caption{}
+    \includegraphics[width=5cm, height=5cm]{SBM.png}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Visualization of the three main network models used in this thesis; (i) ER; (ii) WS and (iii) SBM.]{Visualization of the three main network models used in this thesis.
+\textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM.
+The large number of triangles, indicating high clustering, are clearly visible for the WS model.
+The communities in the SBM are indicated with different colors.
+The definition of weak communities defines the community structure.}
+\label{netRepr}
+\end{figure}
 
@@
 Here, we will list some of the main characteristics of real-world networks, and more specifically of real-world social networks.
-A relevant question is whether there is a significant difference between off-line social networks and on-line social networks.\\
-\newline
+A relevant question is whether there is a significant difference between off-line social networks and on-line social networks.
+\newpage
+\noindent
 As discussed briefly in the introduction, real-world networks often posses three characteristic properties.
@@
 
-We model opinion dynamics on social media platforms, with focus on Facebook.
+We model opinion dynamics on social networks to replicate the opinion dynamics on social media platforms, with focus on Facebook.
 Each time we access Facebook we experience algorithmic personalization.
@@
 Undirected edges represent friendship ties on Facebook, that is, they link an user to his Facebook friends and the pages he or she follows.
-Each node then has an attribute that needs to mimic the Facebook News Feed and which is susceptible to the filtering algorithm.
+Each node has an attribute that needs to mimic the Facebook News Feed and which is susceptible to the filtering algorithm.
 
@@
 Updating is done according to the probabilistic majority model: an actor who sees $20\%$ of the posts promoting opinion A and thus $80\%$ promoting opinion B will adopt opinion A with a chance $p_\text{A} = 0.2$ and opinion B with a chance $p_\text{B} = 0.8$.
-If another model is used, e.g.
-the majority model, this will be explicitly stated.
-% Add pictorial representation of model??
 
@@
 \newline
-The base-line algorithms are the REF and REC filtering algorithms. Filtering tailored to the user's preferences is accomplished with the PR filtering algorithm.
+The base-line algorithms are the REF and REC filtering algorithms. Filtering tailored to the user's preferences is accomplished with the PR filtering algorithm.\\
+\newline
+Fig. \ref{op_dyn_schem} shows a schematic representation of the opinion dynamics model. The activation mechanism and filtering algorithms are represented.
+
+\begin{figure}[H]
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3.5cm]{drawing.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3.5cm]{drawing1.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3.5cm]{drawing2.png}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \caption{}
+  	\includegraphics[width=4cm, height=3.5cm]{drawing3.png}
+  \end{subfigure}
+  \captionsetup{format=plain}
+  \caption[Schematic representation of the opinion dynamics model.]{Schematic representation of the opinion dynamics model.
+Focus lies on the central, colored actor.
+The grey bar represents its hidden time-line $L_i(t)$.
+The central actor is connected to six other actors.
+We assume that these are only connected to this central actor, so their opinion will not be changed until the central actor becomes active.
+\textbf{(a)} Initial situation: every actor is randomly assigned an opinion; the hidden time-line of the central actor is empty; \textbf{(b)} and \textbf{(c)} Actors become active (indicated in red) and push their opinion on the hidden time-line of the central actor; \textbf{(d)} The central actor becomes active: $S = 4$ items of its hidden time-line are selected and shown on its actual time-line.
+(i) REF: the hidden time-line is randomly shuffled; (ii) REC: the hidden time-line is ordered from most recent to oldest; (iii) PR: the hidden time-line is ordered according to the current opinion of the central actor.
+For each filtering algorithm the probability that the central actor changes its opinion is indicated.
+Only the first four posts are taken into account, the rest is deleted.}
+\label{op_dyn_schem}
+\end{figure}
 
@@
 
-If we want to introduce stubbornness in the model we will mainly use the resistance parameter, as explained in Section \ref{stubbBin}.
+If we introduce stubbornness in the model we mainly use the resistance parameter, as explained in Section \ref{stubbBin}.
 Lets summarize this concept of stubbornness.
@@
 For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor updates its opinion, if $f \leqslant r_i$, the actor keeps its current opinion.
+\\
+\newline
+The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed in Algorithm \ref{alg:update}.
 
 \begin{algorithm}[H]
-\caption{Updating the opinions of the active actors at time $t$}
+\caption{Updating the opinions of the active actors at time $t$.}
 \label{alg:update}
@@
 \end{algorithm}
-\noindent
-The pseudo-code for updating the opinions of the active actors in case of using a resistance parameter $r$ at time step $t$ is displayed in Algorithm \ref{alg:update} to make the model more comprehensive and understandable.
 
@@
 
-\begin{figure}[H]
-  \begin{subfigure}[b]{0.24\textwidth}
-    \caption{}
-  	\includegraphics[width=4cm, height=3cm]{drawing.png}
-    \label{pr_real_PGP}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \caption{}
-  	\includegraphics[width=4cm, height=3cm]{drawing1.png}
-    \label{pr_real_PGP}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \caption{}
-  	\includegraphics[width=4cm, height=3cm]{drawing2.png}
-    \label{pr_real_PGP}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \caption{}
-  	\includegraphics[width=4cm, height=3cm]{drawing3.png}
-    \label{pr_real_PGP}
-  \end{subfigure}
-  \captionsetup{format=plain}
-  \caption[Formation of echo chambers versus fraction of stubborn actors for balanced starting conditions.
-The PR and REC filtering algorithms are compared.
-The PGP trust network is compared to corresponding network models.]{test.}
-\label{echo_vs_fracRes_real_PGP}
-\end{figure}
-
 \newpage

### 2021-05-29 draft.tex
latest update

@@
 This filtering and individualizing of information may have some potentially dangerous downsides in the sense that people might get trapped in ``echo chambers'' that confirm their own beliefs.
-It is important to know the effect of algorithmic personalization on the formation of phenomena such as polarization, spread of extremism and formation of echo chambers.\\
+It is important to know the effect of algorithmic personalization on phenomena such as polarization, spread of extremism and formation of echo chambers.\\
 \newline
@@
 We will investigate whether network properties such as community structure and clustering may enhance or hamper the formation of echo chambers.
-We will use filtering algorithms to determine their possible effects on the formation of echo chambers and we will study the impact of individual stubbornness.
-We will also study different ways of distributing the opinions and how this may affect the opinion dynamics in the system.
-\\
-\newline
-Finally, we will test the model on real-world networks. We will investigate whether the results agree with results from the theoretical network models.
+We will use filtering algorithms to determine their possible effect on the formation of echo chambers and we will study the impact of individual stubbornness.
+We will also study different ways of distributing the opinions and how this may affect the opinion dynamics.
+\\
+\newline
+Finally, we will test the model on real-world networks. We will investigate whether the results agree with results from theoretical network models.
 
@@
 \noindent
-Sociale media zijn bijna niet meer weg te denken uit ons dagelijks leven.
+Sociale media zijn niet meer weg te denken uit ons dagelijks leven.
 De laatste tientallen jaren is het gebruik van sociale media, zowel op persoonlijk vlak als op professioneel vlak, in snel tempo toegenomen.
@@
 Het gebruik van sociale media heeft ongetwijfeld veel voordelen.
-Goed en kwaad gaan echter vaak hand in hand en men kan niet ontkennen dat het gebruik van sociale media ook veel potentiële, soms ongekende, nadelen kent.
+Goed en kwaad gaan echter vaak hand in hand en men kan niet ontkennen dat het gebruik van sociale media ook veel potentiële, vaak ongekende, nadelen kent.
 Eén van die, tot nog toe, ongekende kanten van sociale media is de impact op de dynamica van opinies in de maatschappij.
@@
 \newline
-Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar sterk beïnvloed worden door onder andere de mensen rond hen, hun omgeving en hun opvoeding.
+Het is gekend dat mensen niet enkel hun mening uit zichzelf vormen, maar beïnvloed worden door onder andere de mensen rond hen, hun omgeving en hun opvoeding.
 Sociale media laten onmiddellijke interacties toe tussen mensen van overal op de wereld, zonder daarbij belemmerd te worden door fysieke- en tijdsgrenzen.
-Dit kan een potentieel grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\
+Dit kan een grote impact hebben op hoe meningen zich vormen en hoe ze evolueren.\\
 \newline
@@
 Dit kan op zijn beurt polarisatie in de maatschappij, opkomst van extremisme, enzovoort bevorderen.
-Het is daarom uiterst belangrijk om de impact van deze filtering algoritmen grondig in kaart te brengen.\\
+Het is daarom belangrijk om de impact van deze filtering algoritmen grondig in kaart te brengen.\\
 \newline
@@
 Er worden verschillende netwerk modellen geïntroduceerd die de impact van onder andere clustering en groepsstructuur op de evolutie van opinies en de vorming van echo kamers moeten onderzoeken.
-Daarnaast wordt er gebruik gemaakt van enkel simpele filtering algoritmen om hun impact op de dynamica van de opinies te onderzoeken.
-Tot slot wordt er een notie van koppigheid in het systeem geïntroduceerd om mogelijke effecten hiervan in kaart te kunnen brengen.
+Daarnaast wordt er gebruik gemaakt van simpele filtering algoritmen om hun impact op de dynamica van de opinies te onderzoeken.
+Verder wordt er een notie van koppigheid geïntroduceerd om mogelijke effecten hiervan in kaart te kunnen brengen.
 \\
@@
 We zullen de opinies in groepsstructuren verdelen en dit scenario vergelijken met het scenario waarin de opinies willekeurig verdeeld worden.
-Om de thesis te besluiten zal het opinie-dynamica model, dat in de thesis gebruikt werd, getest worden op netwerken verkregen van data uit de werkelijke wereld (bv.
+Om de thesis te besluiten zal het opinie-dynamica model, dat in de thesis gebruikt wordt, getest worden op netwerken verkregen van data uit de werkelijke wereld (bv.
 van email data of telefoonnetwerken).
-Er zal nagegaan worden of de resultaten van deze werkelijke netwerken gelijkenissen, dan wel verschillen, vertonen met de data verkregen van de netwerk modellen.
+Er zal nagegaan worden of de resultaten van deze werkelijke netwerken gelijkenissen, dan wel verschillen, vertonen met de data verkregen van netwerk modellen.
 \setcounter{abstractpage}{\value{page}}
@@
 Some people change their opinions more easily than others, that is, once in a while you encounter somebody that is stubborn and persistent towards its own opinion.
-It is important to understand the effects of individual resistance on the formation and evolution of opinions.
+It is important to understand the effect of individual resistance on the formation and evolution of opinions.
 When designing models of opinion dynamics on social networks, this individual stubbornness should not be ignored.\\
@@
 Opinion dynamics models have two important layers.
-On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate models to reproduce the opinion dynamics and opinion formation processes found in real life.\\
-\newline
+On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate models to reproduce the opinion dynamics and opinion formation processes found in real life.
+\newpage
+\noindent
 Section \ref{statPhys} introduces concepts such as statistical and social physics.
@@
 This lead to the concept of universality: many different microscopic problems have the same macroscopic manifestation \cite{Gug2015}\cite{Kadanoff2015}.
-This idea may then be extended to social systems \cite{Hu2018}: we should not take the unique character of every individual into account, instead these individual characteristics can be incorporated in a sort of averaged way \cite{Gug2015}.
+This idea may be extended to social systems \cite{Hu2018}: we should not take the unique character of every individual into account, instead these individual characteristics can be incorporated in a sort of averaged way \cite{Gug2015}.
 One can then proceed to build models of social behavior where only the most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}.
@@
 \newline
-One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks.
+One of the tools to deal with these complex systems and to give us more insight in the underlying structures are complex networks.
 A network is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis nodes will mostly be called actors; edges will be called edges) \cite{Mata2020}.\\
@@
 Temporal networks are networks where the edges are not always active, but instead become active for some periods of time \cite{Holme2012}.
-In temporal networks time becomes an explicit element in the network representation \cite{Holme2012}.
+Hence, time becomes an explicit element in the network representation \cite{Holme2012}.
 Static networks have been widely studied and are often convenient for their analytic tractability, whereas temporal networks are, in some cases, more realistic \cite{Mata2020}.\\
@@
 This heterogeneous or scale-free degree distribution represents one of the three general properties of social networks.
-The other two are short distances and high clustering, also referred to as ``small world phenomenon'' (note that in the literature the term ``small world'' sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both high clustering and short distances) \cite{Muchnik2013}.
+The other two are short distances and high clustering, also referred to as ``small world phenomenon'' (in the literature the term ``small world'' sometimes only refers to the property of short distances; in this thesis, it is used to indicate both high clustering and short distances) \cite{Muchnik2013}.
 Ideally, our theoretical network should exhibit these three properties.
 Some theoretical networks, that are thoroughly studied, do not possess all of them.
-However, they might still be used, because of their simplicity and ability to produce analytic results.
+They might, however, still be used, because of their simplicity and ability to produce analytic results.
 When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems.
@@
 \newline
-The theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model.
-These will be explained thoroughly in Section \ref{netModel}.
+The theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model and the Watts-Strogatz model. These will be explained in Section \ref{netModel}.
 
@@
 An overview of possible opinion dynamics models is given in Section \ref{Bin} and Section \ref{Con}.
-\\
-\newline
 One can also include a resistance parameter that determines the hesitation of an actor to change to a new opinion or a parameter that determines the influence of an actor on others.
@@
 \newline
-Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define different filtering algorithms.
+Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define filtering algorithms.
 The filtering algorithms that social media companies use, are corporate secrets, but they typically have three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
@@
 This thesis will study the interplay between social network structure and individual resistance to change in the evolution of two competing opinions.
-On one hand, we will determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we will study the effect of stubborn actors in the network.
+On one hand, we will determine the influence of network structures on the formation and evolution of opinions, whereas, on the other hand, we will study the effect of stubborn actors in the network.
 Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms will be included.\\
 \newline
-Perra and Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions \cite{Perra2019}.
-They also considered the effect of nudging (meaning that one opinion is pushed to all actors in the network).
+Perra and Rocha investigated both the effect of filtering algorithms and network topologies on the formation and evolution of two competing opinions \cite{Perra2019}.
+They also considered the impact of nudging (meaning that one opinion is pushed to all actors in the network).
 Their main findings were that algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering) could result in the formation of echo chambers.
@@
 \end{equation}
-In the case of a directed network one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$.
+For directed networks one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$.
 These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
@@
 \end{equation}
-The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$.
+The degree distribution $P(k)$ represents the fraction of nodes with degree $k$.
 The degree distribution gives the probability that a randomly selected node has a degree $k$ \cite{Newman2003}.
@@
 The local clustering coefficient $cc_i$ (Eq.
-(\ref{clus})) of a node and the transitivity $T_i$ (Eq.
+(\ref{clus})) and the transitivity $T_i$ (Eq.
 (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}.
@@
 (\ref{avClus}) gives the same weight to each node \cite{Newman2003}.
-This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.
+This may lead to slightly different values since nodes with a higher degree may be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.
 
@@
 Network communities are another clustering structure of networks.
-A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}.
+A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to other nodes in the network \cite{Saha2015}.
 More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}.
@@
 A community based on this definition would then be the largest clique in the network.
-This definition of community might, however, be too restrictive and large cliques do not appear often in networks \cite{Albert2016}.\\
+This definition of community might be too restrictive and large cliques do not appear often in networks \cite{Albert2016}.\\
 \newpage
@@
 Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
-The modularity of a community $Q_c$ is then calculated by measuring the difference between the network's real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
+The modularity of a community, $Q_c$, is calculated by measuring the difference between the network's real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
 This last quantity is given by $p_{ij}$, which is obtained by randomizing the original network while keeping the expected degree of each node unchanged \cite{Albert2016}.
@@
 The modularity $Q$ can have values in the interval $\interval{-1}{1}$ and has the following properties \cite{Albert2016}: (i) The higher the value of $Q$ for a certain partition, the better the corresponding community structure.
-This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure; (ii) A modularity $Q$ equal to zero corresponds to the situation where all nodes are put in the same community, this is, if the whole network is put into a single community.
+This idea is used in community detection algorithms, where modularity is maximized in order to find the optimal community structure; (ii) A modularity $Q$ equal to zero corresponds to the situation where all nodes are put in the same community, that is, if the whole network is put into a single community.
 Then the two terms in Eq.
@@
 This is because $d_{\textrm{max}}$ is often dominated by a few extreme paths, whereas $\left<d\right>$ is averaged over all node pairs and thus suppresses these fluctuations \cite{Albert2014}.
-Hence, the average distance of a random network is defined by \cite{Albert2014}
+Hence, the average distance of a random network is defined as \cite{Albert2014}
 \begin{equation}\label{d}
@@
 The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}.
-The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$.
-\\
-\newline
-The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (Eq.
+The regular ring lattice thus has high clustering and an average path length that grows linearly with the system size $N$.
+\\
+\newline
+The random network ($\beta = 1$), on the other hand, has a small clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (Eq.
 (\ref{C}) and Eq.
@@
 \end{align}
-Watts and Strogatz showed by numerical simulations that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random network (which is small), even for small $\beta$ \cite{Newman2000}.
+Watts and Strogatz showed by numerical simulations that the average distance for networks with $0 < \beta < 1$ is comparable with the average distance of random networks (which is small), even for small $\beta$ \cite{Newman2000}.
 The clustering coefficient is approximately given by
@@
 (\ref{C0}) \cite{Barrat1999}.
-This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
+This can be seen as follows: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; and (ii) The probability that edges are rewired back to each other is negligible.\\
 \newline
@@
 The stochastic block model originates in social sciences and was developed to describe group structures in friendship networks \cite{Funke2019}.
-The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely interconnected with each other than with the rest of the network \cite{F.Costa2007}.
+The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely connected with each other than with the rest of the network \cite{F.Costa2007}.
 Hence, these networks posses a modular structure and this is a structure that is easily modeled by the stochastic block model.
@@
 Since the SBM is made of random networks inside each group and random bipartite networks between each pair of groups, the degree distribution of the full network is always a mixture of Poisson distributions \cite{Clauset2017}.
-The consequence of this is that if a network possesses a more skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
+The consequence of this is that if a network possesses a skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
 The reason behind this is one of the main assumptions of the stochastic block model: the edge probabilities only depend on which group the node belongs to \cite{Clauset2017}.
 In other words, nodes that belong to the same group are stochastically equivalent, meaning that they have equivalent connectivity patterns to other nodes \cite{Clauset2017}.
-However, a network with a skewed degree distribution violates this assumption of edge independence \cite{Clauset2017}.
+A network with a skewed degree distribution violates this assumption of edge independence \cite{Clauset2017}.
 To say it in the words of Karrer and Newman: ``Just as the fitting of a straight line to intrinsically curved data is likely to miss important features of the data, so a fit of the simple stochastic block model to the structure of a complex network is likely to miss much [...].'' (Karrer and Newman, p.1, \cite{Karrer2011}).
 Karrer and Newman proposed a variation of the SBM, the degree corrected SBM, that deals with this limitation.
-We will not discuss their variation, since that would lead us too far astray, but interested readers can find more detailed information in \cite{Karrer2011}.
+We will not discuss their variation, since that would lead us too far astray; interested readers can find more information in \cite{Karrer2011}.
 \\
@@
 \newline
-\textbf{The stochastic Watts-Strogatz block model (SBM-WS)}\\
-\newline
-The stochastic block model is a flexible model and there are many different extensions.
+\textbf{The stochastic Watts-Strogatz block model}\\
+\newline
+The stochastic block model is a flexible model and there are different extensions.
 One extension that is introduced in this thesis is a stochastic block model that uses the Watts-Strogatz model instead of the Erd\H{o}s-R\'{e}nyi model to construct the constituent communities.
@@
 The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering.
-It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
-\newline
+It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.
+\newpage
+\noindent
 The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure.
@@
 \newline
-Fig. \ref{netRepr} visualizes the three main network models we used: (i) ER; (ii) WS and (iii) SBM.
+Fig. \ref{netRepr} visualizes the three network models used in this thesis: (i) ER; (ii) WS and (iii) SBM.
 
@@
   \captionsetup{format=plain}
-  \caption[Visualization of the three main network models used in this thesis; (i) ER; (ii) WS and (iii) SBM.]{Visualization of the three main network models used in this thesis.
+  \caption[Visualization of the three network models used in this thesis; (i) ER; (ii) WS and (iii) SBM.]{Visualization of the three network models used in this thesis.
 \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM.
-The large number of triangles, indicating high clustering, are clearly visible for the WS model.
+The large number of triangles, indicating high clustering, are visible for the WS model.
 The communities in the SBM are indicated with different colors.
@@
 Here, we will list some of the main characteristics of real-world networks, and more specifically of real-world social networks.
-A relevant question is whether there is a significant difference between off-line social networks and on-line social networks.
+A relevant question is whether there is a difference between off-line and on-line social networks.
 \newpage
@@
 \newline
-This phenomenon is also observed in other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
+This is also observed in other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}.
 Such networks usually have an average distance that depends logarithmically on the system size instead of linearly \cite{Newman2003}.
@@
 The term scale-free refers to the fact that there is no characteristic scale in the network: the system is invariant over different scales.
-Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system), the heterogeneous network has many nodes with a low degree and a few nodes with a high degree (these nodes are referred to as hubs).
+Compared to homogeneous networks, where each node has a degree close to the average degree $\left<k\right>$ (hence, this average degree serves as a characteristic scale of the system),  heterogeneous networks have many nodes with a low degree and a few nodes with a high degree (these nodes are referred to as hubs).
 If the exponent $\gamma$ is smaller or equal to two ($\gamma \leqslant 2$), the average degree diverges \cite{Newman2005}.
 Simple scale-free networks with an exponent smaller than two cannot exist, because the largest hubs grow faster than the network size $N$ \cite{Barabasi2016}.
-If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance in the limit of infinite system size ($N \rightarrow \infty$) \cite{Newman2005}.
+If $2 < \gamma < 3$ (which is the case for most real-world networks), there is a finite average degree, but a diverging variance in the limit of infinite system size ($N \rightarrow \infty$), indicating that the fluctuations around the average degree can be arbitrary large \cite{Newman2005}.
 Real networks always have finite sizes, so they have a finite average degree and variance.
-This variance is, however, meaningless \cite{Newman2005}.
-A diverging variance means that the fluctuations around the average can be arbitrary large \cite{Barabasi2016}.\\
+This variance is, however, meaningless \cite{Newman2005}.\\
 \newline
@@
 \newline
-Ideally, we would test the opinion dynamics model that is used throughout this thesis on network data directly obtained from social network companies.
+Ideally, we would test the opinion dynamics model of this thesis on network data directly obtained from social network companies.
 However, most of these companies are not to keen on sharing any form of data.
@@
 \newline
-The second real-world network that is used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
+The second real-world network is the giant component of the on-line social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
 This network has $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
@@
 These will be discussed in Section \ref{Bin} and Section \ref{Con}, respectively.
-This thesis only deals with binary opinion dynamics models; the continuous models will be reviewed in order to give a more complete overview.\\
+This thesis deals with binary opinion dynamics models; the continuous models will be reviewed in order to give a more complete overview.\\
 \newline
@@
 Binary models are models where each actor can have one of two opinions A or B.
-A binary model could reflect real-life where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
+A binary model reflects real-life where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
 \newline
@@
 E.g.
-$80 \%$ of the actors can be assigned opinion A and $20 \%$ of the actors is then given opinion B.
+$80 \%$ of the actors can be assigned opinion A and $20 \%$ opinion B.
 Hence, opinion A is the predominant or majority opinion and opinion B is the minority opinion.
-The following updating rule is imposed on the network: each actor adopts the majority opinion of its nearest neighbors \cite{Nguyen2020}.
+The following updating rule is then imposed: each actor adopts the majority opinion of its nearest neighbors \cite{Nguyen2020}.
 Initially each actor is assigned an opinion; then the opinions are evolved in time: each actor iteratively updates its opinion according to the majority opinion of its direct neighbors \cite{Nguyen2020}.\\
@@
 \newline
-Some concepts that are valid for discrete models no longer apply (e.g.
-the concept of a majority and minority opinion or equality of opinions); hence, they require a different framework \cite{Castellano2009}.
-\\
+Some concepts that are valid for discrete models no longer apply (e.g. the concept of a majority and minority opinion); hence, they require a different framework \cite{Castellano2009}. \\
 \newline
@@
 
-In the classical model, there are $N$ actors, each with an opinion $x_i$ and each actor $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance or influence of the neighbors) \cite{Krause2002}.
-The most simple model has fixed weights $w$ and the updating rule is
+In the classical model, there are $N$ actors, each with an opinion $x_i$.
+Each actor $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance or influence of the neighbors) \cite{Krause2002}.
+The simplest model has fixed weights $w$ and the updating rule is
 \begin{equation}\label{refMod}
@@
 In bounded confidence models, each actor has, beside an opinion $x_i$, a threshold $\epsilon_i$, also called an uncertainty or tolerance ($\epsilon$ can be the same for all actors or can be different for each actor) \cite{Castellano2009}\cite{Deffuant2006}.
-This threshold determines which opinions of neighbors are close enough to the actor's opinion in order to influence that actor; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
+This threshold determines which opinions of neighbors are close enough to the actor's opinion to influence that actor; opinions that are beyond this threshold are ignored \cite{Deffuant2006}.
 In other words, an actor with opinion $x_i$ is only influenced by those neighbors which have an opinion in the interval $\interval{x_i-\epsilon}{x_i+\epsilon}$ \cite{Castellano2009}.
@@
 Furthermore, the threshold $\epsilon$ is the same for all the actors.
-At each time step, an actor $i$ is randomly selected together with one of its neighbors $j$ (which is also randomly selected) \cite{Castellano2009}.
+Each time step, an actor $i$ is randomly selected together with one of its neighbors $j$ (also randomly selected) \cite{Castellano2009}.
 If the opinions of the two actors $i$ and $j$ are too far apart nothing happens; this is the case if $\abs{x_i(t) - x_j(t)} \geqslant \epsilon$.
-If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, then the opinions are updated to \cite{Castellano2009}\cite{Deffuant2000}
+If instead $\abs{x_i(t) - x_j(t)} < \epsilon$, the opinions are updated to \cite{Castellano2009}\cite{Deffuant2000}
 \begin{equation}
@@
 The dynamics is similar to the Deffuant model and the stationary state is also characterized by a number of opinion clusters \cite{Castellano2009}.
-The number of clusters decreases if $\epsilon$ increases and above a critical value $\epsilon_c$ of the threshold, there can only be one cluster \cite{Castellano2009}.\\
+The number of opinion clusters decreases if $\epsilon$ increases and above a critical value $\epsilon_c$ of the threshold, there can only be one opinion cluster \cite{Castellano2009}.\\
 \newline
@@
 This is, unfortunately, also true for persistent groups with extreme ideas: they become capable of reaching and influencing larger groups of people.
-Before the era of social media, one could maybe argue that these extreme and stubborn groups were often isolated and therefore would not reach too many people.
+Before the era of social media, one could argue that these extreme and stubborn groups were often isolated and therefore would not reach too many people.
 So their impact could only be minor (this is only a statement that could have been given, there are plenty of examples throughout human history that may contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, communism exploited by, among other, Stalin and Mao in the 20th century, all examples of groups or regimes that can be considered extreme and that have had a (relatively) wide reach and a significant impact on society).
@@
 The moderates are given a threshold $\epsilon$, with $\epsilon > \epsilon_e$ \cite{Deffuant2006}.
-This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates, which is based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
+This last condition implicitly denotes the resistance of the extremes towards other opinions in contrast to the more flexible nature of the moderates.
+This is based on the idea that people with extreme opinions will be more convinced of that opinion and will be less likely to change it.\\
 \newline
@@
 This parameter can be either the same for all the actors ($T$) or individualized ($T_i$).
-The main idea is that actors can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is larger than $T$ ($p > T$).
+Actors can then only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is larger than $T$ ($p > T$).
 The idea behind the threshold model is that people are hesitant to change their opinion, but if they have enough friends with the other opinion, they might get convinced of that opinion as well.
@@
 One way, which makes use of the majority model, is as follows.
-First the fraction of friends with opinion A and opinion B, $p_\text{A}$ and $p_\text{B}$ respectively, are determined; then the following rules are used to update the opinion: if $p_\text{A} > p_\text{B}$ and $p_\text{A} > T$, opinion A is adopted; if $p_\text{B} > p_\text{A}$ and $p_\text{B} > T$, opinion B is adopted; in all other cases, the current opinion of the actor is not changed.\\
+First the fraction of friends with opinion A and opinion B, $p_\text{A}$ and $p_\text{B}$ respectively, are determined; then the following rules are used to update the opinion: if $p_\text{A} > p_\text{B}$ and $p_\text{A} > T$, opinion A is adopted; if $p_\text{B} > p_\text{A}$ and $p_\text{B} > T$, opinion B is adopted; in all other cases, the opinion of the actor is not changed.\\
 \newline
@@
 \newline
-It is important to know and be aware of the possible impact of algorithmic personalization on the formation and evolution of opinions.
+It is important to know and be aware of the impact of algorithmic personalization on the formation and evolution of opinions.
 Perra and Rocha partly investigated the role of such algorithms on the dynamics of opinions in several networks \cite{Perra2019}.
 Filtering algorithms are often not transparently shared by social media companies or social platforms.
-Personalization algorithms are based on three filtering strategies \cite{Perra2019}: (i) Popularity filtering: promoting content that is popular across the platform; (ii) Semantic filtering: promoting posts that are similar to previous consumed posts; (iii) Collaborative filtering: promoting content that people similar or close to us consume.\\
+Personalization algorithms are based on three filtering strategies \cite{Perra2019}: (i) popularity filtering: promoting content that is popular across the platform; (ii) semantic filtering: promoting posts that are similar to previous consumed posts; and (iii) collaborative filtering: promoting content that people similar or close to us consume.\\
 \newline
@@
 Nodes represent the Facebook users, which become active and inactive over time.
-Undirected edges represent friendship ties on Facebook, that is, they link an user to his Facebook friends and the pages he or she follows.
+Undirected edges represent friendship ties on Facebook, that is, they link an user to his or her Facebook friends and the pages he or she follows.
 Each node has an attribute that needs to mimic the Facebook News Feed and which is susceptible to the filtering algorithm.
@@
 
-Perra and Rocha proposed different ways of selecting the $S$ items from the hidden time-line $L_i(t)$ \cite{Perra2019}.
+Perra and Rocha proposed different ways of selecting the $S = 20$ items from the hidden time-line $L_i(t)$ \cite{Perra2019}.
 We will implement three of their algorithms: (i) REF filtering: $S$ random items from the hidden time-line $L_i(t)$ are selected and shown on the actual time-line $R_i(t)$; (ii) REC filtering: the $S$ most recent posts in the hidden time-line $L_i(t)$ are displayed on the actual time-line $R_i(t)$ and (iii) PR filtering: this way of filtering is based on semantic filtering explained in Section \ref{filter}; the posts on the hidden time-line $L_i(t)$ are ordered according to the current opinion of the actor (if the actor has opinion A, the hidden time-line will be ordered as AAAA...BBB), then the $S$ first items are selected and shown on the time-line $R_i(t)$.\\
@@
 We assume that these are only connected to this central actor, so their opinion will not be changed until the central actor becomes active.
-\textbf{(a)} Initial situation: every actor is randomly assigned an opinion; the hidden time-line of the central actor is empty; \textbf{(b)} and \textbf{(c)} Actors become active (indicated in red) and push their opinion on the hidden time-line of the central actor; \textbf{(d)} The central actor becomes active: $S = 4$ items of its hidden time-line are selected and shown on its actual time-line.
+\textbf{(a)} Initial situation: every actor is randomly assigned an opinion; the hidden time-line of the central actor is empty; \textbf{(b)} and \textbf{(c)} Actors become active (indicated in red) and push their opinion on the hidden time-line of the central actor; \textbf{(d)} The central actor becomes active: $S = 4$ items of its hidden time-line are selected and shown on its actual time-line (indicated in red).
 (i) REF: the hidden time-line is randomly shuffled; (ii) REC: the hidden time-line is ordered from most recent to oldest; (iii) PR: the hidden time-line is ordered according to the current opinion of the central actor.
@@
 Each actor is given a resistance parameter $r_i$, with $0 \leqslant r_i \leqslant 1$; $r_i = 0$ means not stubborn and $r_i = 1$ means completely stubborn.
-For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor updates its opinion, if $f \leqslant r_i$, the actor keeps its current opinion.
+For all intermediate values of $r_i$, a random number $f$ between zero and one is drawn; if $f > r_i$, the actor updates its opinion, if $f \leqslant r_i$, the actor keeps its opinion.
 \\
@@
 		\State \(\textsc{Order}(L_i)\) \Comment{Order hidden time-line according to REF, REC or PR filtering}
-		\If{$\text{len}(L_i) > S$}
 		\State \(R_i \leftarrow L_i.\textsc{resize}(S)\) \Comment{Time-line $R_i$ = $S$ first elements of ordered hidden time-line}
-		\EndIf
 		\State \(\text{op} \leftarrow \textsc{probMaj}(R_i)\) \Comment{Update opinion according to probabilistic majority model}

### 2021-05-30 draft.tex
latest update

@@
 We first discuss the impact of network structure on the evolution of the prevalence of the opinions and the formation of echo chambers.
-Then, we introduce stubbornness in the system.
-The impact of how the opinions are initially distributed in a network with community structure (SBM) will be discussed afterwards.
-Finally, we will study the opinion dynamics model in real-world networks.
-We will discuss the results and compare them to results from network models with a similar average degree, clustering or modularity.\\
+Then, we introduce stubbornness.
+The impact of how the opinions are initially distributed in a network with community structure (SBM) are discussed afterwards.
+Finally, we study the opinion dynamics model in real-world networks.
+We discuss the results and compare them to results from network models with similar average degree, clustering or modularity.\\
 \newline
@@
 For both starting conditions we investigate the evolution of the prevalence of the opinions and the formation of echo chambers.
-Two filtering algorithms will be compared, the REC and the PR filtering algorithms, in order to investigate the interplay between filtering algorithms and network structure.
+Two filtering algorithms will be compared, the REC and the PR filtering algorithms, to investigate the interplay between filtering algorithms and network structure.
 
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each curve is the average of $10^2$ independent simulations.
@@
 \noindent
-In Fig.
-\ref{echo_50_50} two different filtering algorithms are compared, the REC and the PR filtering algorithms.
+Two filtering algorithms are compared, the REC and the PR filtering algorithms.
 The results for both filtering algorithms are qualitatively the same.
@@
 \newline
-The ER model shows no, or a limited, sign of echo chambers, for both the REC and PR filtering algorithms.
+The ER model shows no, or a limited, sign of the formation of echo chambers, for both the REC and PR filtering algorithms.
 Networks with high clustering (WS and SBM-WS), on the other hand, show a sign of polarization and formation of echo chambers (visible in the two spikes, one around zero and another around one).
-Going to slightly higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the WS model results in significantly larger echo chambers.
+Going to higher values of the clustering coefficient ($\left<cc\right> = 0.557 \pm 0.007$ to $\left<cc\right> = 0.648 \pm 0.003$) in the WS model results in significantly larger echo chambers.
 This suggests that the formation of echo chambers is sensitive to changes in the clustering coefficient.
@@
 \ref{ev_op_50_50} and related results in \cite{Perra2019}).
-Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can potentially shift the initial randomly distributed opinions to a clustered opinion distribution and thus lead to polarization and formation of echo chambers.
+Despite this preserving of the balanced opinion distribution, the interplay of filtering algorithms and network structure can shift the initial randomly distributed opinions to a clustered opinion distribution and thus lead to polarization and formation of echo chambers.
 
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$ configuration).
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 For each model the REC and the PR filtering algorithms are compared.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 For each model the REC and the PR filtering algorithms are compared.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each curve is the average of $10^2$ independent simulations.
@@
 \noindent
-Lets first discuss the results for the ER model and the SBM (Fig.
+Lets first discuss results for the ER model and the SBM (Fig.
 \ref{er_echo_20-80} and Fig.
@@
 The PR filtering algorithm shows a different tendency.
-There is an increase in the fraction of actors with (almost) all their neighbors carrying the majority opinion B.
+There is an increase in the fraction of actors with all their neighbors carrying the majority opinion B.
 This is a consequence from the fact that the PR filtering algorithm reduces the visibility of the minority opinion A in the neighborhood of actors with the majority opinion B.
@@
 \ref{ws} and Fig.
-\ref{sbm-ws}, the spikes around zero are not actually reaching infinity, but are cut off around values equal to 100/180.
-These large spikes, indicating a large increase of echo chambers around the minority opinion, do not necessary mean that we have a large absolute number of echo chambers around the minority opinion A at $t=500$.
-They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion A at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion A at $t=500$).
+\ref{sbm-ws}, the spikes around zero are not reaching infinity, but are cut off around values equal to 100 or 180.
+These large spikes, indicating a large increase in echo chambers around the minority opinion, do not necessarily mean that we have a large absolute number of echo chambers around the minority opinion A at $t=500$.
+They merely indicate that, compared to the beginning at $t=0$, the increase is large (if there are zero echo chambers around the minority opinion A at $t=0$, the increase is infinite even if there is only one echo chamber around the minority opinion A at $t=500$).
 If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion B and that only a minor fraction carries the minority opinion A.\\
@@
 	\captionsetup{format=plain}
-	\caption[Relation between the formation of echo chambers and community structure (modularity $Q$).]{Relation between the formation of echo chambers and the community structure, measured by the modularity $Q$, for both the PR and REC filtering algorithms.
+	\caption[Formation of echo chambers versus community structure (modularity $Q$).]{Formation of echo chambers versus community structure, measured by the modularity $Q$, for both the PR and REC filtering algorithms.
 Data points with $Q = 0$ are the result for an ER network; the other eight data points are results for SBM's, each with different modularity, which is changed by changing the edge probabilities.
-The dotted black line represents a constant amount of echo chambers over time.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B  (normalized means that the value for the fraction of actors with all neighbors having opinion A or all neighbors having opinion B at $t=500$ is divided by the corresponding value at $t=0$).
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
 Bars represent the standard error $SE$.
-The data points are shown together with the standard error $SE = \sigma /\sqrt{n}$ ($n$ is the number of independent simulations), because we are most interested in the correctness of the average.}
+Data points are shown together with the standard error $SE = \sigma /\sqrt{n}$ ($n$ is the number of independent simulations), because we are most interested in the correctness of the average.}
 \label{echo_vs_mod}
@@
 We focus on two different stubbornness models, discussed in Section \ref{stubbBin}.
-The results of the first model, including a resistance parameter, are given in Section \ref{stubbpar}.
-The results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
+Results of the first model, including a resistance parameter, are given in Section \ref{stubbpar}.
+Results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
 \newline
@@
 
-We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, $f_s=0$, to all the actors are stubborn, $f_s=1$. \\
+We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, $f_s=0$, to all actors are stubborn, $f_s=1$. \\
 \newline
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
@@
 The SBM with low modularity behaves similar as the ER model.
-The data points for the SBM with low modularity lie behind the ER data points and are not visible in Fig.
+The data points for this model lie behind the ER data points and are not visible in Fig.
 \ref{echo_vs_frac_complStubb}.
@@
 The WS and SBM-WS networks have high clustering.
-For a small fraction of stubborn actors they show the largest echo chambers.
-The SBM with high modularity has smaller echo chambers than networks with high clustering, but still has some formation of echo chambers for $f_s \leqslant 0.1$.
+For a small fraction of stubborn actors they have the largest formation of echo chambers.
+The SBM with high modularity has a smaller formation of echo chambers than networks with high clustering, but still has some formation of echo chambers for $f_s \leqslant 0.1$.
 These results confirm that community structure enhances the formation of echo chambers compared to random networks, however not as significant as networks with high clustering.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 The SBM with high modularity displays a small bump.
-The ER and the SBM with low modularity (data points for this model lie behind the ER data points) do not seem to have a significant dependency on the fraction of stubborn actors.\\
+The ER and the SBM with low modularity (data points lie behind the ER data points) do not seem to have a significant dependency on the fraction of stubborn actors.\\
 \newline
@@
 The system will be less likely to form an echo chamber: the stubborn actor with opinion B cannot switch to opinion A and the PR filtering algorithm reduces the chance of changing the opinion of the actors with opinion A to opinion B.
-The conclusion is that, in case of the PR filtering algorithm, making a fraction of the actors stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
+In case of the PR filtering algorithm, making a fraction of the actors stubborn reduces the ability of the system to form echo chambers compared to the case where there are no stubborn actors.
 Lets now study what happens if we use the REF or REC filtering algorithms.
@@
 This is not a problem.
-We want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of the echo chambers.
+We want to know how the total number of echo chambers is impacted by the presence of stubborn actors and are less interested in the distribution of the echo chambers.
 It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as an average in case of unbalanced starting conditions.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 For balanced starting conditions we found that introducing stubborn actors did not affect the prevalence of the opinions over time.
-Lets now study whether this is also the case for unbalanced starting conditions.
+Lets now study whether this is also true for unbalanced starting conditions.
 In Fig.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The dotted black line represents the initial prevalence of opinion A.
@@
 This is not changed by making the actors more and more stubborn.
-The same result is obtained for the SBM with low modularity (the data points lie behind those of the ER network).
+The same result is obtained for the SBM with low modularity (data points lie behind those of the ER network).
 The SBM with high modularity has a (small) formation of echo chambers for $r=0$.
@@
 	\caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
-The PR filtering algorithm is used.]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
+The PR filtering algorithm is used.]{Formation of echo chambers versus the resistance parameter $r$ for balanced starting conditions.
 Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
@@
   \caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
-The REF and REC filtering algorithms are used]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for balanced starting conditions.
+The REF and REC filtering algorithms are used]{Formation of echo chambers versus the resistance parameter $r$ for balanced starting conditions.
 \textbf{(a)} REF filtering algorithm; \textbf{(b)} REC filtering algorithm.
@@
 For the REF filtering algorithm, this increase reaches a maximum around $r = 0.7$ and slowly decreases between $r = 0.7$ and $r = 0.9$.
-Between $r = 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers.
+Between $r = 0.9$ and $r=1$, there is a steep drop in the formation of echo chambers.
 The REC filtering algorithm seems to have a maximum around $r = 0.7$ for the WS model.
@@
 First, we discuss the formation of echo chambers.
-Then, we investigate the results for the evolution of the prevalence of the opinions.\\
+Then, we investigate results for the evolution of the prevalence of the opinions.\\
 \newline
@@
   \caption[Formation of echo chambers versus the resistance parameter of the actors, $r$, for unbalanced starting conditions.
-The PR and REC filtering algorithms are used.]{Formation of echo chambers versus the resistance parameter of the actors, $r$, for unbalanced starting conditions.
+The PR and REC filtering algorithms are used.]{Formation of echo chambers versus the resistance parameter $r$ for unbalanced starting conditions.
 \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$).
-The dotted black line represents a constant amount of echo chambers over time.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each data point is the average of $10^2$ independent simulations.
@@
 Lets now study the effect of the resistance parameter of the actors on the evolution of the prevalence of the opinions in case of unbalanced starting conditions.
-Three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$) are compared.
+Three different values of the stubbornness parameter ($r=0$, $r=0.4$ and $r=0.8$) are compared.
 When using the REC filtering algorithm, the prevalence of opinion A versus time is not affected by making the actors more and more stubborn.
@@
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The dotted black line represents the initial prevalence of opinion A.
-The PR filtering algorithm is used and three values of the resistance parameter $r$ ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
+The PR filtering algorithm is used and three values of the resistance parameter ($r = 0$, $r = 0.4$ and $r = 0.8$) are compared.
 Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.
@@
 This model is explained in Section \ref{stubbBin}.
-We do not use a probabilistic majority model, but a classic majority model, explainded in Section \ref{majModel}, is implemented.
-The rest of the model (filtering algorithms, activation mechanism) is unchanged and still follows the rules as stated in Section \ref{modelThesis}.
-The threshold parameter $T$ is equal for all the actors and varies between $T=0$ and $T=1$.
-We only investigate balanced starting conditions ($50/50$), with the focus on the formation of echo chambers.
-The PR, REF and REC filtering algorithms are compared.
-\\
+We do not use a probabilistic majority model, but a classic majority model, explainded in Section \ref{majModel}.
+The rest of the model (filtering algorithms, activation mechanism) is unchanged and follows the rules as stated in Section \ref{modelThesis}.
+The threshold parameter $T$ is equal for all actors and varies between $T=0$ and $T=1$.
+We investigate balanced starting conditions, with focus on the formation of echo chambers.\\
 \newline
@@
 Networks with high clustering (WS and SBM-WS) show a significant formation of echo chambers, which remains more or less flat for $T \leqslant 0.5$.
-This is expected, because if $T \leqslant 0.5$, the threshold has no effect: the majority opinion of the neighbors of an actor will be larger than the threshold (the fraction of neighbors with the majority opinion always take values between $0.5$ and $1$).
+This is expected, because if $T \leqslant 0.5$, the threshold has no effect: the majority opinion of the neighbors of an actor is larger than the threshold (the fraction of neighbors with the majority opinion always take values between $0.5$ and $1$).
 If $T \geqslant 0.5$, it starts to have an effect.
@@
 	\caption[Formation of echo chambers versus the threshold parameter $T$, for balanced starting conditions.
-The PR filtering algorithm is used and the majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter $T$, for balanced starting conditions.
+The PR filtering algorithm is used and the majority threshold model is implemented.]{Formation of echo chambers versus the threshold parameter $T$ for balanced starting conditions.
 Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER.
 The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$).
-Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$)) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
+Two configurations are implemented for the SBM: a high modularity ($Q = 0.908 \pm 0.005$) and a low modularity ($Q = 0.220 \pm 0.007$) configuration.
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 The PR filtering algorithm is used.
@@
 This is not due to any network topologies, but a consequence of the convergence to one of the two opinions each independent simulation.
-This convergence results in a strong increase in the echo chambers around that opinion (the whole system becomes a large, single echo chamber).
+This convergence results in a strong increase in the echo chambers around that opinion (the whole system becomes a large echo chamber).
 The PR filtering algorithm seems to hamper this convergence and hence, this artificial creation of echo chambers.
@@
 In Fig.
-\ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of the echo chambers until the echo chambers size reaches values equal to one for $T=1$.
+\ref{echo_vs_threshold_REF-REC} the effect of the threshold does not become visible until $T > 0.8$, after which we see a decrease in the formation of echo chambers until the echo chambers size reaches values equal to one for $T=1$.
 At this point, none of the actors can be convinced of another opinion and hence all actors retain their initial opinion.
-The number of echo chambers does not change over time, which results in an echo chamber size equal to one.
 
@@
 Results are compared to the case in which the total opinion distribution ($55/45$) is randomly distributed.
-The high and low modularity configuration are depicted.
-For each modularity the REC and PR filtering algorithms are compared.\\
+The first row shows results for the high modularity configuration; the second row shows those for the low modularity configuration.
+The REC and PR filtering algorithms are compared.\\
 \newline
@@
 \ref{fracOp0_vs_commOp0_other_50-50}, which depicts the prevalence of opinion A before (solid black line) and after (dotted lines) the time evolution versus the fraction of single opinion A communities, $f_\text{A}$.
-Community distributed cases are compared to randomly distributed cases.
-Results are shown for both a high and a low modularity configuration.\\
+Community distributed cases are compared to randomly distributed cases.\\
 \newline
@@
 The community distributed case is compared to the randomly distributed case for both a high and a low modularity configuration.
-The dotted black line represents a constant amount of echo chambers over time.
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each curve is the average of $10^2$ independent simulations.

### 2021-05-30 draft.tex
latest update

@@
 
-Testing the opinion dynamics model on network models, gives us useful insights in the interplay between network topologies and model features such as filtering algorithm and stubbornness.
-It is, however, interesting to test the model on real-world networks to get a better understanding of real-world opinion dynamics.\\
+Studying the opinion dynamics model on network models, gives us insights in the interplay between network topologies and model features such as filtering algorithm and stubbornness.
+It is, however, interesting to study the model on real-world networks to get a better understanding of real-world opinion dynamics.\\
 \newline
@@
 Results are compared to results from network models, all with a size of $N = 8003$ nodes: (i) WS model with $\left<k\right> = 4$, $\beta = 0.15$ and $\left<cc\right> = 0.317 \pm 0.004$; (ii) SBM with 53 communities of size 151 nodes each, $p_\text{cl} = 0.023$, $p_\text{add} = 0.0001$, $\left<k\right> = 4.24 \pm 0.03$ and $Q = 0.811 \pm 0.004$; (iii) SBM-WS with 53 communities of size 151 nodes each, each community has WS parameters: $\left<k\right> = 4$, $\beta = 0.025$; edges are added between communities with probability $p_\text{add} = 0.0001$, $\left<k\right> = 4.77 \pm 0.01$, $\left<cc\right> = 0.344 \pm 0.002$ and $Q = 0.833 \pm 0.003$.
-The results for the PR and REC filtering algorithms are shown in Fig.
+Results for the PR and REC filtering algorithms are shown in Fig.
 \ref{echo_vs_fracRes_real}.\\
@@
 Other topologies must play a role as well.
-The most significant difference between the real-world Last.fm network and the SBM-WS lies in the degree distribution: the Last.fm network has a heterogeneous instead of a homogeneous degree distribution.
+A difference between the real-world Last.fm network and the SBM-WS lies in the degree distribution: the Last.fm network has a heterogeneous instead of a homogeneous degree distribution.
 It is probably this heterogeneous contact pattern that hampers the formation of echo chambers in the real-world network compared to network models with similar clustering and modularity, but with a homogeneous degree distribution.\\ % add parameters of SBM-WS power??
@@
 It is interesting to notice that the Last.fm on-line social network coincides with the SBM with similar modularity.
-It seems that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of the same size as the hampering due to a lack of clustering.
+This may suggest that the hampering of the formation of echo chambers due to a heterogeneous contact pattern is of a similar size as the hampering due to a lack of clustering.
 
@@
 The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each data point is the average of $50$ independent simulations.
@@
 The increase is the strongest for network models with high clustering (WS, SBM-WS).
-Even though the Last.fm network has similar clustering as these high clustering models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
+Even though the Last.fm network has similar clustering as these models, the formation of echo chambers and the corresponding increase in that formation are significantly smaller.
 As for the PR filtering algorithm, this is probably a consequence of the heterogeneous degree pattern of the Last.fm network.
-The results for the Last.fm network coincide with those of the SBM.
+Results for the Last.fm network coincide with those for the SBM.
 This seems to suggest that the hampering in the formation of echo chambers due to a heterogeneous degree distribution is of a similar size as the hampering due to a lack of clustering.
@@
 \newline
-Another difference between the Last.fm and the SBM and SBM-WS network is the community structure.
+Another difference, beside the degree distribution, between the Last.fm and the SBM and SBM-WS network is the community structure.
 The Last.fm network has a community structure where the community sizes follow a heterogeneous distribution: plenty of small communities and a few large ones.
-The SBM and SBM-WS only have communities of a single, fixed size.
-This does not play a significant role in the formation of echo chambers, as can be seen in Fig.
+The SBM and SBM-WS have communities of a single, fixed size.
+This difference in community structure does not play a noticeable role in the formation of echo chambers, as can be seen in Fig.
 \ref{power_vs_single}.
@@
 The distribution is normalized by dividing for the same quantity computed at $t=0$.
-The dotted black line represents a constant amount of echo chambers over time.
+The dotted black line ($y=1$) represents a constant amount of echo chambers over time.
 Each curve is the average of $50$ independent simulations.
@@
 \ref{pr_real_PGP}) shows a decrease in the formation of echo chambers with an increasing fraction of stubborn actors for both the PGP trust network and the network models.
-The real-world PGP network has a significantly smaller formation of echo chambers compared to the network model with similar clustering and modularity (SBM-WS).
+The real-world PGP network has a smaller formation of echo chambers compared to network models with similar clustering and modularity (SBM-WS).
 This is most likely due to the heterogeneous degree pattern of the PGP network.
@@
 We made use of simple toy models of social media platforms to generate results regarding the evolution of the prevalence of the opinions and formation of echo chambers.
-These toy models consist of several layers, including a network layer and an opinion dynamics layer.
+These toy models consist of two layers, a network layer and an opinion dynamics layer.
 Filtering algorithms were implemented to replicate and simulate the content curation used by social media companies.
@@
 \newline
-When introducing stubbornness, we found that the interplay between the PR filtering algorithm and individual resistance resulted in a decrease in the formation of echo chambers for the both the high clustering (WS, SBM-WS) and the high modularity networks (SBM, SBM-WS).
+When introducing stubbornness, we found that the interplay between the PR filtering algorithm and individual resistance resulted in a decrease in the formation of echo chambers for both the high clustering (WS, SBM-WS) and high modularity networks (SBM).
 The REF and REC filtering algorithms showed an increase in the formation of echo chambers with increasing stubbornness, followed by a decrease if stubbornness was further increased.
@@
 semantic filtering), is one of the reasons for the appearance of echo chambers on social media \cite{Ge2020}\cite{Mohseni2018}\cite{Stark2020}.
-Our results may, however, suggest, that if there is stubbornness (which is probably the case in real life), heterogeneous exposure (REF or REC) to different opinions may be more harmful than homogeneous exposure (PR).
+Our results may, however, suggest, that if there is stubbornness, heterogeneous exposure (REF or REC) to different opinions may be more harmful than homogeneous exposure (PR).
 Bail et al.
@@
 It was found that the initial distribution of the opinions over the communities in networks with community structure (SBM) makes a difference.
-We investigated a scenario in which a fraction of the communities was given a single opinion, opinion A, while the remaining communities were given an initial $50/50$ opinion distribution.
+We investigated a scenario in which a fraction of the communities was given a single opinion, while the remaining communities were given an initial $50/50$ opinion distribution.
 This scenario was then compared to the scenario in which the total opinion distribution was randomly distributed.
@@
 \newline
-The REC filtering algorithm did not show a difference in the evolution of the prevalence of the opinions between the (initial) community distributed and (initial) randomly distributed scenario.
-For the PR filtering algorithm, it was found that the high modularity configuration hampers the increase in the majority opinion in the community distributed scenario compared to the random distributed scenario.
+The REC filtering algorithm did not show a difference in the evolution of the prevalence of the opinions between the community distributed and randomly distributed scenario.
+For the PR filtering algorithm, it was found that the high modularity configuration hampers the increase in the majority opinion in the community distributed scenario compared to the randomly distributed scenario.
 For the low modularity network, such a discrepancy was not found.
@@
 This reduces polarization and was found for both the high and low modularity configuration.
-The randomly distributed scenario does not evolve towards mixed opinion neighborhoods, instead a small increase in the formation of echo chambers around the minority opinion B is observed.
+The randomly distributed scenario does not evolve towards mixed opinion neighborhoods, instead a small increase in the formation of echo chambers around the minority opinion is observed.
 Hence, polarization increases in this scenario.\\
 \newline
-For the PR filtering algorithm, the random distributed scenario resulted in an increase in the formation of echo chambers around the majority opinion A (for both the low and high modularity configuration).
+For the PR filtering algorithm, the randomly distributed scenario resulted in an increase in the formation of echo chambers around the majority opinion (for both the low and high modularity configuration).
 The community distributed scenario for the low modularity configuration hampers this increase.
-It was found that this is more than just a result of the increased number of initial opinion A echo chambers.
-For the high modularity network, the community distributed scenario showed an even larger hampering of the formation of echo chambers around the majority opinion A.
-This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion A than for the low modularity network; (ii) the hampering in the increase of the majority opinion A for the community distributed scenario compared to the randomly distributed scenario.
+It was found that this is more than a result of the increased number of initial opinion A echo chambers due to the single opinion communities.
+For the high modularity network, the community distributed scenario showed an even larger hampering in the formation of echo chambers around the majority opinion.
+This is probably the result of: (i) an even higher initial number of echo chambers around the majority opinion than for the low modularity network; (ii) the hampering in the increase in the majority opinion that occurs for the high modularity, community distributed scenario.
 \\
@@
 Finally, results for real-world networks were compared to those from network models.
-It was found that the formation of echo chambers in the real-world networks is smaller than those in network models with similar clustering and modularity.
+It was found that the formation of echo chambers in real-world networks is smaller than those in network models with similar clustering.
 The reason for this hampered formation of echo chambers was attributed to the heterogeneous degree pattern of the real-world networks.
-This hampering of echo chambers due to a heterogeneous degree distribution was also found by Perra and Rocha \cite{Perra2019} when they compared network models with heterogeneous degree distribution to models with homogeneous contact patterns.\\
+This hampering of echo chambers due to a heterogeneous degree distribution was also found by Perra and Rocha \cite{Perra2019} when they compared network models with heterogeneous degree distributions to models with homogeneous contact patterns.\\
 \newline
@@
 \newline
-The hampering in the increase of the majority opinion in case of unbalanced starting conditions combined with the introduction of stubborn actors for the PR filtering algorithm agrees partly with related research on continuous opinion dynamics models combined with stubborn actors \cite{Como2013}.
+The hampering in the increase of the majority opinion in case of unbalanced starting conditions combined with the introduction of stubborn actors for the PR filtering algorithm agrees with related research on continuous opinion dynamics models combined with stubborn actors \cite{Como2013}.
 In this research, Acemo\u{g}lu et al.
@@
 Even though the stubbornness parameter $r$ could be individualized to each actor, we did not include the possibility to be more stubborn towards opinions coming from one friend than from another.
-This idea may be linked to the influence of certain friends: somebody may adopt another viewpoint easier from certain friends than from others.
-As a final limitation, we mention the homogeneous degree distribution of our network models.
-Real-world social networks often have heterogeneous contact patterns.
-It is therefore recommended to test the opinion dynamics model on heterogeneous network models.\\
-\newline
-Future directions may be: (i) investigating the distribution of the stubborn actors (community distributed versus random); (ii) making one opinion more stubborn than the other; (iii) investigating the effect of the activation mechanism: switching to the heterogeneous activation mechanism and (iv) making the network temporal in the sense that it becomes possible to add and remove friends.
+This idea may be linked to the influence of certain friends: somebody may adopt another viewpoint easier from certain friends than from others.\\
+\newline
+Future directions may be: (i) investigating the distribution of the stubborn actors (community distributed versus random); (ii) making one opinion more stubborn than the other; (iii) investigating the effect of the activation mechanism: switching to the heterogeneous activation mechanism; (iv) studying the opinion dynamics model on heterogeneous network models; (v) introducing nudging; (vi) switching to continuous opinion dynamics models; (vii) introducing an influence parameter and (viii) making the network temporal in the sense that it becomes possible to add and remove friends.
 Sasahara et al.
@@
 
-Simple toy models were implemented to study the interplay between network structure, filtering algorithms and individual resistance to change competing opinions.
-Our goal was to get a better understanding of the impact of algorithmic personalization on polarization and formation of echo chambers (opinion clusters) in society.
+Simple toy models were implemented to study the interplay between network structure, filtering algorithms and individual resistance.
+Our goal was to get a better understanding of the impact of algorithmic personalization and stubbornness on polarization and formation of echo chambers in society.
 \\
@@
 Beside clustering, community structure also favors the formation of echo chambers.
-Real-world networks have a lower formation of echo chambers than network models with similar clustering and modularity.
+Real-world networks have a lower formation of echo chambers than network models with similar clustering.
 The heterogeneous contact pattern of these real-world networks is most likely the cause of this hampering.
@@
 If, on the contrary, people have some level of stubbornness, we found that homogeneous exposure to information may be more protective against opinion clusters than heterogeneous exposure.
-This is important, since it may change our assumptions that algorithmic personalization is at the root of the increase in polarization and extremism often seen on social media.
+This is important, since it may change our assumptions that algorithmic personalization, tailored to one's preferences, is at the root of the increase in polarization and extremism often seen on social media.
 More research on this topic is needed.
-Finally, we found that distributing the opinions in communities instead of randomly distributing them may hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.\\
-\newline
-Our results suggest that the effects of algorithmic personalization do not stand on their own.
+Finally, we found that distributing the opinions in communities instead of randomly distributing them may hinder the convergence towards the majority opinion in case of semantic filtering, if the modularity is high enough.
+Furthermore, this community distributed scenario seems to reduce polarization compared to the randomly distributed scenario.
+This might feel counterintuitive: it might be expected that if you initially distributed the opinions in a more polarized configuration, this would enhance polarization.
+Our results suggest that this does not necessarily need to be so.\\
+\newline
+Our findings suggest that the effects of algorithmic personalization do not stand on their own.
 Instead, the interplay with certain network topologies may enhance or hamper these effects.

### 2021-05-31 draft.tex
final version?

@@
 \newline
-One important feature of opinion dynamics on social media is the use of algorithmic personalization.
+One important aspect of opinion dynamics on social media is the use of algorithmic personalization.
 Social media allow for simultaneous or asynchronous interactions between people without geographical constraints and thus allow for the spread of information at a faster pace and at an unprecedented scale \cite{Perra2019}.
@@
 The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}.
-The application of the theory of statistical physics on social phenomena is referred to as social physics \cite{Castellano2009}\cite{Galam2008}\cite{Galam1982}\cite{Stauffer2012}.
+The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics \cite{Castellano2009}\cite{Galam2008}\cite{Galam1982}\cite{Stauffer2012}.
 The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies.
@@
 
-The formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (e.g.
+The formation of an individual's opinion is the result of the interplay of factors such as self-reflection, peer pressure, the individual's personality (e.g.
 stubbornness) and the information someone is exposed to.
@@
 Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define filtering algorithms.
-The filtering algorithms that social media companies use, are corporate secrets, but they typically have three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
+The filtering algorithms that social media companies use, are corporate secrets, but typically have three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}.
 Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}.
@@
 
-Another concept that is closely related to the degree of a node is assortative and disassortative networks. The notion of assortativity and disassortativity has to do with degree correlations. \\
+Another concept that is closely related to the degree of a node is assortative and disassortative networks. Assortativity and disassortativity have to do with degree correlations. \\
 \newline
@@
 
-Several measurements quantify the notion of length and distance in a network.
+Several measurements quantify length and distance in a network.
 A path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}.
@@
 \newline
-The average distance is the mean value of the distance $d_{ij}$ \cite{F.Costa2007}
+The average distance is the mean value of the geodesic $d_{ij}$ \cite{F.Costa2007}
 \begin{equation}\label{avdist}
@@
 \end{equation}
-This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}.
-This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes.
+This definition diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}.
+This can be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes.
 These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}.
@@
 The modularity of a network with $N$ nodes and $L$ edges and partitioned into $n_c$ communities is calculated as follows.
-Each community has $N_c$ nodes and $L_c$ edges ($c = 1,..., n_c$) \cite{Albert2016}.
+Each community has $N_c$ nodes and $L_c$ edges ($c = 1, ..., n_c$) \cite{Albert2016}.
 The modularity of a community, $Q_c$, is calculated by measuring the difference between the network's real wiring diagram (which is given by the adjacency matrix $A_{ij}$) and the expected number of edges between $i$ and $j$ if the network is randomly wired (where $i$ and $j$ are two nodes that lie in the same community $C_c$) \cite{Albert2016}.
@@
 We focus on two network topologies that may impact the formation of echo chambers: high clustering and community structure.
-It is therefore necessary to have a model that lacks both of these features to compare the results with.
-The Erd\H{o}s-R\'{e}nyi model is one of the most simple random network models that does not display both of these features and is hence the appropriate comparison model.
+It is therefore necessary to have a model that lacks both of these topologies to compare the results with.
+The Erd\H{o}s-R\'{e}nyi model is one of the simplest random network models that does not display both of these topologies and is hence the appropriate comparison model.
 It is also used in the regular stochastic block model to construct the constituent communities.\\
@@
 The idea behind the development of this model is that often in real networks, there are groups of nodes that are more densely connected with each other than with the rest of the network \cite{F.Costa2007}.
-Hence, these networks posses a modular structure and this is a structure that is easily modeled by the stochastic block model.
+Hence, these networks posses a modular structure and this is a structure that can be modeled by the stochastic block model.
 The stochastic block model is also often used as a benchmark for community detection algorithms \cite{Abbe2018}.\\
@@
 Since the SBM is made of random networks inside each group and random bipartite networks between each pair of groups, the degree distribution of the full network is always a mixture of Poisson distributions \cite{Clauset2017}.
-The consequence of this is that if a network possesses a skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
+The consequence is that if a network possesses a skewed degree distribution, then fitting the SBM to the data will most likely give poor results \cite{Karrer2011}.
 The reason behind this is one of the main assumptions of the stochastic block model: the edge probabilities only depend on which group the node belongs to \cite{Clauset2017}.
@@
 This data set was originally used in \cite{Toivonen2009}.
-This on-line social network has size $N = 8003$; number of edges: $16824$; $Q = 0.79$; $\left<cc\right> = 0.31$; $\left<k\right> = 4.2$.
+This on-line social network has size $N = 8003$ nodes; number of edges: $16824$; $Q = 0.79$; $\left<cc\right> = 0.31$; $\left<k\right> = 4.2$.
 The degree distribution is heterogeneous and is best fitted by a log-normal
@@
 The second real-world network is the giant component of the on-line social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{Boguna2004}\cite{ICON}.
-This network has $N = 10680$; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
+This network has $N = 10680$ nodes; number of edges: $24340$; $Q = 0.85$; $\left<cc\right> = 0.27$; $\left<k\right> = 4.56$.
 The network has a heterogeneous degree distribution, which is best fitted by two power-laws
@@
 One of the decisive properties of the real-world networks that are used, is their size.
-The main, and actually only, reason for this is the computation time.
-If the networks are too big, it takes weeks, sometimes even months before there would be results.
+The main, and actually only, reason is the computation time.
+If the networks are too large, it takes weeks, sometimes even months before there would be results.
 The size of the network models that are implemented, is around a thousand nodes.
@@
 This benefits the computation time.
-
+\newpage
 \section{Temporal networks}
@@
 Binary models are models where each actor can have one of two opinions A or B.
-A binary model reflects real-life where there exist two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
+A binary model reflects real-life where there are two competing opinions, such as a vote in an election system with two competing candidates or choosing between Windows and Linux \cite{Nguyen2020}.\\
 \newline
@@
 \newline
-In this thesis, actors are made active with some probability $p_i$ every time step (thus, every time step, each actor becomes active with a probability $p_i$).
-The activation probability is set equal to $p = 0.1$, which is constant for all the actors.
-This constant activation probability is chosen for simplicity.
+In this thesis, each actor becomes active with a probability $p_i$ every time step. The activation probability is $p = 0.1$, which is constant for all the actors and is chosen for simplicity.
 
@@
 In real life, some people are persistent towards their own opinion and are not open to the opinions of others.
-They are convinced that they carry the truth with them and that other people, with different world-views, are not capable of giving them any new and useful insights.
+They are convinced that they carry the truth with them and that people, with different world-views, are not capable of giving them any new and useful insights.
 They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs.
@@
 Social media facilitates the spread of information and ideas.
-This is, unfortunately, also true for persistent groups with extreme ideas: they become capable of reaching and influencing larger groups of people.
+This is also true for persistent groups with extreme ideas: they become capable of reaching and influencing larger groups of people.
 Before the era of social media, one could argue that these extreme and stubborn groups were often isolated and therefore would not reach too many people.
@@
 Table \ref{tab1} displays the network models used in Fig.
-\ref{echo_vs_mod} (all with $N = 10^3$ and $\left<k\right> \sim 10$).
+\ref{echo_vs_mod} (all with $N = 10^3$ nodes and $\left<k\right> \sim 10$).
 The SBM's have 10 communities of size 100 nodes each.
@@
 
-Stubbornness is an important feature of real-life opinion dynamics.
+Normally, people do not change their opinion thoughtlessly, hence stubbornness is an important aspect of real-life opinion dynamics.
 Understanding the impact of individual resistance and the interplay with filtering algorithms and network topologies is crucial.
@@
 There is, however, a quantitative difference: the formation of echo chambers is lower when using unbalanced starting conditions compared to using balanced starting conditions.
-The reason for this is probably the higher initial number of echo chambers when using unbalanced starting conditions.\\
+The reason is probably the higher initial number of echo chambers when using unbalanced starting conditions.\\
 \newline
@@
 The total opinion distribution is not balanced, but opinion A becomes the majority opinion.
-The higher the fraction of single opinion A communities, the bigger the difference between the majority and minority opinion.
+The higher the fraction of single opinion A communities, the larger the difference between the majority and minority opinion.
 Results will be compared to the configuration in which the total opinion distribution is randomly distributed over each community.
@@
 \newline
-The hampering in the increase of the majority opinion in case of unbalanced starting conditions combined with the introduction of stubborn actors for the PR filtering algorithm agrees with related research on continuous opinion dynamics models combined with stubborn actors \cite{Como2013}.
+The hampering in the increase of the majority opinion, and hence the inability to reach consensus around this majority opinion, in case of unbalanced starting conditions combined with the introduction of stubborn actors for the PR filtering algorithm agrees with related research on continuous opinion dynamics models combined with stubborn actors \cite{Como2013}.
 In this research, Acemo\u{g}lu et al.
 showed that introducing stubborn actors (with different opinions) in an inhomogeneous stochastic gossip process of continuous opinion dynamics leads to a situation in which consensus can no longer be reached \cite{Como2013}.
-This result is also found in the majority opinion dynamics model and in the voter model combined with stubborn actors (equally distributed over the two opinions) \cite{Galam2007}\cite{Mobilia2007}\cite{Yildiz2013}.
-In our case, the possible consensus around the majority opinion, that could have been reached when combining the PR filtering algorithm with unbalanced starting conditions, can no longer be obtained when introducing stubborn actors.\\
+This result is also found in the majority opinion dynamics model and in the voter model combined with stubborn actors (equally distributed over the two opinions) \cite{Galam2007}\cite{Mobilia2007}\cite{Yildiz2013}.\\
 \newline
@@
 More empirical data is therefore needed to compare the results, obtained from our model, against \cite{Castellano2009}\cite{Pawel}.
-Other limitations may be the small system sizes that were used in our simulations ($N = 10^3$).
+Other limitations may be the small system sizes in our simulations ($N = 10^3$ nodes).
 It is unclear how system size affects the results and it is recommended that the model is tested on both larger network models and real-world networks.
-Limitations due to the activation mechanism, may be that we used a constant activation probability.
+Limitations due to the activation mechanism, may be the constant activation probability.
 Analyses of real social networks show that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}.
 It might be an interesting extension to switch to the heterogeneous activation probabilities and investigate possible impacts on the dynamics.
-Another possible shortcoming is the binary nature of the model.
-There are real-life cases in which people have to choose between two competing opinions.
-These situations, however, do not cover the complete range of opinion dynamics situations.
-It would be interesting to run the model on configurations where more than two opinions are allowed or to switch to continuous opinion dynamics.
-We also did not differentiate between the actors in our model, except for non-stubborns versus stubborns.
-This combined with the homogeneous degree distribution of our network models may be another limitation.
+Another shortcoming may be the binary nature of the model.
+There are real-life situations in which people have to choose between two competing opinions.
+These, however, do not cover all the possibilities.
+It would be interesting to study the model on configurations where more than two opinions are allowed or to switch to continuous opinion dynamics.
+Furthermore, we did not differentiate between the actors in our model, except for non-stubborns versus stubborns.
+This combined with the homogeneous degree distribution of our network models can be another limitation.
 Real-life social media platforms display a heterogeneous contact pattern, where the `hubs' are often more influential than other users, not only by their larger reach, but also by their status.
@@
 Furthermore, this community distributed scenario seems to reduce polarization compared to the randomly distributed scenario.
-This might feel counterintuitive: it might be expected that if you initially distributed the opinions in a more polarized configuration, this would enhance polarization.
+This might feel counterintuitive: it might be expected that if the opinions are initially distributed in a more polarized configuration, this would enhance polarization.
 Our results suggest that this does not necessarily need to be so.\\
@@
 It would therefore be incorrect to only point to algorithmic personalization as the cause of polarization.
-It is the interplay between topological correlations, distribution of the opinions, individual resistance and content curration that potentially enhance or hamper the formation of opinion clusters.
+It is the interplay between topological correlations, distribution of the opinions, individual resistance and content curation that potentially enhance or hamper the formation of opinion clusters.
 
